%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{array}
\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

This material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

{\LARGE{}EPIC MONSTERS (G-W)}

\vspace{12pt}
\subsubsection*{{\LARGE{}GENIUS LOCI }}

\subsubsection*{\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{179pt}|}
\hline
\section*{ } & \section*{Colossal Ooze }\tabularnewline
\hline
H\textbf{it Dice:} & 70d10+1,400 (1,785 hp) \tabularnewline
\hline
I\textbf{nitiative:}- & 2 (Dex) \tabularnewline
\hline
S\textbf{peed:} & 5 ft., burrow 5 ft. \tabularnewline
\hline
A\textbf{C:} & 0 (-2 Dex, -8 size) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +52/+88\tabularnewline
\hline
A\textbf{ttack:} &  Slam +64 (4d10+20) melee\tabularnewline
\hline
F\textbf{ull Attack:} & 10 slam +64 (4d10+20) melee\tabularnewline
\hline
S\textbf{pace/Reach:} & Varies (see text) \tabularnewline
\hline
S\textbf{pecial Attacks:} & Enslave, improved grab, constrict 4d10+30 \tabularnewline
\hline
S\textbf{pecial Qualities:} & Blindsight 200 ft., fast healing 50, ooze traits 
\tabularnewline
\hline
S\textbf{aves:} & Fort +43, Ref +21, Will +30 \tabularnewline
\hline
A\textbf{bilities:} & Str 50, Dex 6, Con 50, Int -, Wis 24, Cha 26 \tabularnewline
\hline
S\textbf{kills:} & Jump +5\tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary \tabularnewline
\hline
C\textbf{hallenge Rating:} & 30 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Usually any evil \tabularnewline
\hline
A\textbf{dvancement:} & None \tabularnewline
\hline
\end{tabular}

A genius loci is difficult to spot, since it looks like any other section of landscape. 
A genius loci itself has no intelligence, but rather ``emulates'' the sentience 
of any creature it currently enslaves (see below), if any.  Each genius loci has 
only a single enslaved creature at one time, though other creatures may also inhabit 
the surface of the genius loci. Any creature entering a genius loci who makes a 
successful Survival check (DC 31) realizes that something is amiss, though it is 
difficult to tell exactly what. Genius loci do not speak directly, though they 
may do so through their currently enslaved thrall. 

\section*{COMBAT }

A genius loci can directly attack any creature standing on its surface with animated 
parts of itself, delivering slam attacks. It also directs its enslaved thrall against 
those it attacks. If its slave is slain, it selects another of its foes as its 
next, newest slave. Once a genius loci animates a portion of itself to attack the 
boundaries of its form become clear, and the creature becomes open to attack (Survival 
checks to determine the extent of the genius loci are no longer necessary). The 
animated part of the genius loci has a reach of 15 feet. 

\textbf{Enslave (Su):} A genius loci can use the epic spell \textit{enslave }once 
per round as a free action until it successfully enslaves a foe (Will DC 53 to 
resist). The DC is Charisma-based. Once it has a slave, it cannot enslave another 
until its current slave dies, is \textit{imprisone}d, or is otherwise incapaticiated. 
A slave becomes free if its genius loci master is slain. A genius loci forces its 
slave to remain with it, though it may send its slave off to temporarily visit 
nearby locations in order to draw in unsuspecting victims. In fact, the ``purposes'' 
of the genius loci are to some degree influenced by the current slave. The genius 
loci has no Intelligence score of its own; thus, its emulation of its slave's intelligence 
sometimes transfers attitudes and goals. However, a slave can only go so far in 
influencing its mobile master, and cannot knowingly attempt to free itself from 
the genius loci that keeps it close. 

\textbf{Improved Grab (Ex):} If the genius loci hits a Gargantuan or smaller creature 
with a slam attack, it deals normal Damage and attempts to start a grapple as a 
free action without provoking an attack of opportunity. The genius loci has the 
option to conduct the grapple normally, or simply use the part of its body it used 
in the improved grab to hold the opponent. Each successful grapple check it makes 
during successive rounds automatically deals slam attack Damage, in addition to 
constriction Damage. 

\textbf{Constrict (Ex):} A genius loci crushes its opponent, dealing 4d10+30 points 
of bludgeoning Damage, after making a successful grapple check. 

\textbf{Ooze Traits:} An ooze is an amorphous or mutable creature. Oozes are immune 
to poison, \textit{sleep, }paralysis, stunning, and polymorphing. They have no 
clear front or back and are therefore not subject to critical hits or flanking. 
Oozes are blind but have the blindsight special quality. They have no Intelligence 
scores and are therefore immune to all mind-affecting effects (charms, compulsions, 
phantasms, patterns, and morale effects). 

\vspace{12pt}
{\LARGE{}GIBBERING ORB }

\begin{tabular}{|>{\raggedright}p{66pt}|>{\raggedright}p{259pt}|}
\hline
  & Huge Aberration\tabularnewline
\hline
H\textbf{it Dice:} & 27d8+216 (337 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +20 (+12 Dex, +8 Superior Initiative) \tabularnewline
\hline
S\textbf{peed:} & 5 ft., fly 20 ft. (good) \tabularnewline
\hline
A\textbf{C:} & 48 (+12 Dex, -2 size, +16 natural armor, +12 insight) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +20/+39\tabularnewline
\hline
A\textbf{ttack:} &  24 eye rays +31 ranged touch, 12 bites +28 (2d8+5/19-20) melee\tabularnewline
\hline
F\textbf{ull Attack:} & 24 eye rays +31 ranged touch, 12 bites +28 (2d8+5/19-20) 
melee\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./10 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Gibbering, improved grab, swallow whole, eye rays, 
spell-like abilities \tabularnewline
\hline
S\textbf{pecial Qualities:} & All-around vision, flight, amorphous, SR 37, DR 10/epic 
\tabularnewline
\hline
S\textbf{aves:} & Fort +21, Ref +21, Will +24 \tabularnewline
\hline
A\textbf{bilities:} & Str 32, Dex 35, Con 27, Int 40, Wis 24, Cha 22 \tabularnewline
\hline
S\textbf{kills:} & Concentration +38, Escape Artist +42, Jump -4, Knowledge (arcana) 
+37, Hide +34, Listen +37, Move Silently +34, Search +37, Speak Language (any five), 
Spot +37\tabularnewline
\hline
F\textbf{eats:} & Flyby Attack, Improved Initiative, Improved Critical (bite), 
Iron Will, Multiattack, Weapon Focus (eye ray), Weapon Focus (bite) \tabularnewline
\hline
E\textbf{pic Feats:} & Epic Fortitude, Polyglot, Superior Initiative \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any land and underground\tabularnewline
\hline
O\textbf{rganization:} & Solitary \tabularnewline
\hline
C\textbf{hallenge Rating:} & 27 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic evil \tabularnewline
\hline
A\textbf{dvancement:} & 27-54 HD (Huge); 55-108 HD (Gargantuan) \tabularnewline
\hline
\end{tabular}

A gibbering orb can speak all languages.

A gibbering orb's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

\section*{COMBAT }

The gibbering orb has the ability to bite its foes by extending a pseudopod with 
one of its mouths protruding from the end. The orb can extend four of these at 
any one foe, or a total of twelve in any given round. At the same time, the legions 
of eyes have the ability to cast a host of spells at a rapid rate. 

\textbf{Gibbering (Su):} The cacophony of speech emanating from the scores of mouths 
that make up the orb forces all within 60 feet of the creature to succeed at a 
Will save each round (DC 40) or suffer the effects of an \textit{insanity }spell. 

\textbf{Improved Grab (Ex):} To use this ability, the gibbering orb must hit with 
a bite attack. If it gets a hold, the orb automatically deals bite Damage and can 
attempt to swallow whole. 

\textbf{Swallow Whole (Ex):} A gibbering orb can attempt to swallow a Large or 
smaller opponent by making a successful grapple check. Once inside, the opponent 
takes 4d8 points of constriction Damage plus 3d10 points of acid Damage per round. 
A swallowed creature can cut its way out by using claws or a light slashing weapon 
to deal 20 points of Damage in a single blow (AC 16, DR 10/epic still applies). 
Once the creature exits, the fluidlike body structure of the orb closes the wound; 
other swallowed opponents must cut their own way out. When a creature dies by being 
swallowed whole (or when a creature killed by the gibbering orb in some other fashion 
is eaten by it), the gibbering orb absorbs the creature's known spells, prepared 
spells, and spell-like abilities. The orb can use any two of those abilities per 
round as a free action. Each originates from an eye that is not producing an eye 
ray (see below) that round. Stolen spells and spell-like abilities are lost after 
24 hours. 

\textbf{Eye Rays (Su):} Two dozen of the eyes can each produce a magical ray each 
round, with each eye emulating a spell from among the list of spells below as if 
cast by a 27th-level caster. The save DCs, where applicable, are 16 + spell level. 
A gibbering orb has no directional limitations on where it can point its eye rays, 
because the eyes orbiting around its body drift and float wherever needed. However, 
a gibbering orb can never aim more than five rays at any single target, due to 
limitations of aiming. All rays have a range of 150 feet. Each of these effects 
functions as a ray, regardless of the normal parameters of the spell it resembles. 
That is, each is usable against a single target and requires a ranged touch attack. 
The eye rays are: \textit{cone of cold, disintegrate, dominate monster, energy 
drain, feeblemind, finger of death, flesh to stone, greater dispelling, harm, hold 
monster, horrid wilting, implosion, inflict critical wounds, lightning bolt, magic 
missile, mage's disjunction, irresistible dance, polymorph any object, power word 
blind, power word kill, power word stun, prismatic spray, slay living, }and \textit{temporal 
stasis. }

\textbf{Spell-Like Abilities: }At will---Two stolen spells or spell-like abilities 
per round. Caster level 27th, save DC 16 + spell level. 

\textbf{All-Around Vision (Ex): }The gibbering orb's many eyes give it a +4 racial 
bonus on Spot and Search checks. 

\textbf{Flight (Ex):} The gibbering orb's body is naturally buoyant. This buoyancy 
allows it to fly as the spell, as a free action, at a speed of 20 feet. This buoyancy 
also grants it a permanent \textit{feather fall }effect with personal range. 

\textbf{Amorphous (Ex):} A gibbering orb is not subject to critical hits. It has 
no clear front or back, so it cannot be flanked. 

\vspace{12pt}
{\LARGE{}GLOOM }

\begin{tabular}{|>{\raggedright}p{63pt}|>{\raggedright}p{263pt}|}
\hline
  & Medium-Size Monstrous Humanoid \tabularnewline
\hline
H\textbf{it Dice:} & 25d8+225 (337 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +26 (+18 Dex, +8 Superior Initiative) \tabularnewline
\hline
S\textbf{peed:} & 30 ft. \tabularnewline
\hline
A\textbf{C:} & 40 (+18 Dex, +12 insight) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +25/+36\tabularnewline
\hline
A\textbf{ttack:} & +\textit{10 keen dagger of human dread }+54 (1d4+21/17-20) melee\tabularnewline
\hline
F\textbf{ull Attack:} & +\textit{10 keen dagger of human dread }+54/+49/+44/+39 
(1d4+21/17-20) melee\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Fear gaze, sneak attack +13d6 \tabularnewline
\hline
S\textbf{pecial Qualities:} & Blindsight 60 ft., opportunist, quiescence, spell-like 
abilities, SR 35, DR 10/epic \tabularnewline
\hline
S\textbf{aves:} & Fort +17, Ref +32, Will +21 \tabularnewline
\hline
A\textbf{bilities:} & Str 32, Dex 46, Con 29, Int 26, Wis 25, Cha 30 \tabularnewline
\hline
S\textbf{kills:} & Bluff +38, Climb +39, Hide +46, Intimidate +38, Jump +39, Listen 
+35, Move Silently +66, Spot +35, Survival +35, Tumble +46\tabularnewline
\hline
F\textbf{eats:} & Dodge, Improved Initiative, Improved Critical (dagger), Mobility, 
Spring Attack, Weapon Finesse, Weapon Focus (dagger) \tabularnewline
\hline
E\textbf{pic Feats:} & Sneak Attack of Opportunity, Superior Initiative \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any urban \tabularnewline
\hline
O\textbf{rganization:} & Solitary \tabularnewline
\hline
C\textbf{hallenge Rating:} & 25 \tabularnewline
\hline
T\textbf{reasure:} & Nonstandard (just its dagger) \tabularnewline
\hline
A\textbf{lignment:} & Usually lawful evil \tabularnewline
\hline
A\textbf{dvancement:} & 25+ HD (Medium-size) \tabularnewline
\hline
\end{tabular}

In the hand of any other creature but a gloom, the dagger is a \textit{+5 keen 
dagge}r. 

A gloom's natural weapons are treated as epic for the purpose of overcoming damage 
reduction.

COMBAT 

\textbf{Fear Gaze (Su):} Viewing the gloom's face inspires terror. Creatures that 
meet the gloom's gaze must succeed at a Will save (DC 32) or suffer the effects 
of a \textit{fear }spell as cast by a 20th-level caster.   The DC is Charisma-based.

\textbf{Opportunist (Ex):} This ability functions like the rogue class feature 
of the same name. 

\textbf{Sneak Attack (Ex):} A gloom is able to sneak attack as a 25th-level rogue, 
dealing 13d6 extra Damage. 

\textbf{Quiescence (Su):} A gloom is constantly silent, as per the \textit{silence 
}spell, although the area does not extend beyond the creature itself. This quality 
gives it a +20 racial bonus on Move Silently checks (it is still possible for a 
gloom to create noise, such as by bumping into something that scrapes on the floor 
or by sounding a chime). The gloom can lower this effect at will. 

\textbf{Spell-Like Abilities: }At will---\textit{shadow walk; }3/day--- \textit{true 
strike. }Caster level 20th; save DC 20 + spell level.  The DC is Charisma-based.

\vspace{12pt}
\section*{{\LARGE{}GOLEM}}

\section*{\begin{tabular}{|>{\raggedright}p{51pt}|>{\raggedright}p{131pt}|>{\raggedright}p{131pt}|}
\hline
} & M\textbf{ithral Golem } & A\textbf{damantine Golem }\tabularnewline
\hline
  & Huge Construct  & Huge Construct \tabularnewline
\hline
H\textbf{it Dice:}  & 36d10+40 (238 hp)  & 54d10+40 (337 hp) \tabularnewline
\hline
I\textbf{nitiative:}  & +4 (Dex) - & 1 (Dex) \tabularnewline
\hline
S\textbf{peed:}  & 40 ft.  & 20 ft. (can't run) \tabularnewline
\hline
A\textbf{C:}  & 42 (-2 size, +4 Dex, +26 natural, +4 haste)  & 37 (-2 size, -1 
Dex, +30 natural) \tabularnewline
\hline
B\textbf{ase Attack/}\linebreak{}
\textbf{Grapple:} & +27/+49 & +40/+68\tabularnewline
\hline
A\textbf{ttack:} & Slam +39 (4d10+14) melee & Slam +58 (6d10+20) melee\tabularnewline
\hline
F\textbf{ull Attack:} & 2 slams +39 (4d10+14) melee & 2 slams +58 (6d10+20) melee\tabularnewline
\hline
S\textbf{pace/Reach:}  & 15 ft./15 ft.  & 15 ft./15 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} --- &   & Trample \tabularnewline
\hline
S\textbf{pecial Qualities:}  & Construct traits, magic immunity, DR 15/epic and 
adamantine, alacrity & Construct traits, magic immunity, DR 20/epic and adamantine\tabularnewline
\hline
S\textbf{aves:}  & Fort +12, Ref +16, Will +12  & Fort +18, Ref +17, Will +18 \tabularnewline
\hline
S\textbf{kills:} & Jump +18 & Jump +14\tabularnewline
\hline
A\textbf{bilities:}  & Str 39, Dex 19, Con -, Int -, Wis 11, Cha 1  & Str 51, Dex 
9, Con -, Int -, Wis 11, Cha 1 \tabularnewline
\hline
C\textbf{limate/Terrain:}  & Any land  & Any land \tabularnewline
\hline
O\textbf{rganization:}  & Solitary  & Solitary \tabularnewline
\hline
C\textbf{hallenge Rating:}  & 21  & 25 \tabularnewline
\hline
T\textbf{reasure: } & None  & None \tabularnewline
\hline
A\textbf{lignment:}  & Always neutral  & Always neutral \tabularnewline
\hline
A\textbf{dvancement:}  & 37-54 HD (Huge); 55-72 HD (Gargantuan)  & 55-82 HD (Huge); 
83-108 HD (Gargantuan) \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\section*{MITHRAL GOLEM }

A mithral golem cannot speak or make any vocal noise, nor does it have any distinguishable 
odor. Unlike most golems, it can run just as well as a normal creature. 

A mithral golem's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

Combat

\textbf{Alacrity (Su):} Once per round, the mithral golem may take an extra standard 
action (either before or after its other actions in the round). 

\textbf{Magic Immunity (Ex):} A mithral golem is immune to all magical and supernatural 
effects, except as follows. A \textit{slow }spell negates its alacrity for 1d4 
rounds, while a \textit{haste }spell restores 1d6 hit points per level of the caster 
(maximum 10d6) or restores its alacrity, if previously negated by a \textit{slow 
}spell. Multiple \textit{slow }spells simply extend the duration of the effect. 

\section*{Construction }

A mithral golem's body is sculpted from 8 1/2 tons of pure iron and is then polymorphed 
into mithral (using \textit{polymorph any objec}t). A mithral golem costs 250,000 
gold pieces to create, which includes 5,000 gp for the body. Assembling the body 
requires a successful Craft (armorsmithing) or Craft (weaponsmithing) check (DC 
30). The creator must be 25th level, have the Craft Epic Magic Arms and Armor feat, 
and be able to cast arcane spells. Completing the ritual drains 5,000 XP from the 
creator and requires \textit{geas/quest, haste, }and \textit{wis}h. 

\vspace{12pt}
\section*{ADAMANTINE GOLEM }

An adamantine golem cannot speak or make any vocal noise, nor does it have any 
distinguishable odor. 

An adamantine golem's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

\section*{Combat }

\textbf{Trample (Ex):} As a standard action during its turn each round, the adamantine 
golem can literally run over an opponent at least one size category smaller than 
itself. The trample deals 8d10+30 points of bludgeoning Damage. Trampled opponents 
can either attempt attacks of opportunity at a -4 penalty or Reflex saves (DC 57) 
for half Damage. The DC is Strength-based.

\textbf{Magic Immunity (Ex):} An adamantine golem is immune to all magical and 
supernatural effects.

\section*{Construction }

An adamantine golem's body is sculpted from 45,000 pounds of pure iron and is then 
polymorphed into adamantine (using \textit{polymorph any objec}t). The golem costs 
500,000 gold pieces to create, which includes 25,000 gp for the body. Assembling 
the body requires a successful Craft (armorsmithing) or Craft (weaponsmithing) 
check (DC 40). The creator must be 30th level, have the Craft Epic Magic Arms and 
Armor feat, and be able to cast arcane spells. Completing the ritual drains 10,000 
XP from the creator and requires \textit{crushing hand, geas/quest, }and \textit{wis}h. 

\vspace{12pt}
{\LARGE{}HA-NAGA }

\begin{tabular}{|>{\raggedright}p{63pt}|>{\raggedright}p{262pt}|}
\hline
  & Colossal Aberration \tabularnewline
\hline
H\textbf{it Dice:} & 20d8+220 (310 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +14 (Dex) \tabularnewline
\hline
S\textbf{peed:} & 60 ft., fly 120 ft. (perfect) \tabularnewline
\hline
A\textbf{C:} & 40 (-8 size, +14 Dex, +24 natural armor) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +15/+39\tabularnewline
\hline
A\textbf{ttack:} & Coil whip +21 (4d6+8) melee\tabularnewline
\hline
F\textbf{ull Attack:} & Coil whip +21 (4d6+8) melee, sting +19 (2d8+4 plus poison) 
melee, bite +13 (4d8+4) melee\tabularnewline
\hline
S\textbf{pace/Reach:} & 30 ft./ 20 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Charming gaze, poison, improved grab, constrict 4d6+12 
\tabularnewline
\hline
S\textbf{pecial Qualities:} & Flight, SR 30, DR 5/epic \tabularnewline
\hline
S\textbf{aves:} & Fort +17, Ref +22, Will +22 \tabularnewline
\hline
A\textbf{bilities:} & Str 27, Dex 38, Con 32, Int 35, Wis 31, Cha 36 \tabularnewline
\hline
S\textbf{kills:} & Appraise +35, Concentration +34, Bluff +24, Diplomacy +38, Escape 
Artist +37, Hide +21, Jump +20, Knowledge (arcana, history, religion) +35, Listen 
+33, Move Silently +37, Search +35, Sense Motive +24, Spellcraft +37, Spot +33\tabularnewline
\hline
F\textbf{eats:} & Dodge, Flyby Attack, Lightning Reflexes, Mobility, Multiattack, 
Quicken Spell, Weapon Finesse \tabularnewline
\hline
C\textbf{limate/Terrain:} & Temperate and warm land or underground \tabularnewline
\hline
O\textbf{rganization:} & Solitary or nest (2-4) \tabularnewline
\hline
C\textbf{hallenge Rating:} & 22 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic evil \tabularnewline
\hline
A\textbf{dvancement:} & 21-40 HD (Colossal) \tabularnewline
\hline
\end{tabular}

A ha-naga's natural weapons are treated as epic for the purpose of overcoming damage 
reduction.

COMBAT 

\textbf{Flight (Su):} As per the \textit{fly }spell, 120 ft. (perfect). This ability 
gives the ha-naga a +6 circumstance bonus on Move Silently checks. 

\textbf{Charming Gaze (Su):} As \textit{mass charm monster, }90 ft., Will save 
(DC 33). The DC is Charisma-based.

\textbf{Chameleon Ability (Ex):} Ha-nagas can blend in with their surroundings, 
giving them a +8 circumstance bonus on Hide checks. 

\textbf{Improved Grab (Ex):} To use this ability, the ha-naga must hit with its 
coil whip attack. If it succeeds, it can constrict. 

\textbf{Constrict (Ex):} A ha-naga deals 4d6+12 points Damage with a successful 
grapple attack against Huge or smaller opponents. 

\textbf{Poison (Ex):} Sting, Fort save (DC 31); initial and secondary Damage 2d8 
temporary Con.  The DC is Constitution-based.

\textbf{Spells:} Ha-nagas can cast spells as 21st-level sorcerers, and can also 
cast cleric spells and spells from the domains of Chaos and Evil as arcane spells 
(save DC 33 + spell level). The DC is Charisma-based.

\vspace{12pt}
{\LARGE{}HAGUNEMNON (PROTEAN) }

\begin{tabular}{|>{\raggedright}p{64pt}|>{\raggedright}p{261pt}|}
\hline
  & Large Aberration (Shapechanger) \tabularnewline
\hline
H\textbf{it Dice:} & 44d8+616 (814 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +17 (+13 Dex, +4 Improved Initiative) \tabularnewline
\hline
S\textbf{peed:} & 50 ft. \tabularnewline
\hline
A\textbf{C:} & 50 (-1 size, +13 Dex, +28 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +33/+58\tabularnewline
\hline
A\textbf{ttack:} & Slam +54 (2d6+21/19-20) melee\tabularnewline
\hline
F\textbf{ull Attack:} & 5 slams +54 (2d6+21/19-20) melee\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Psionics, destabilize form \tabularnewline
\hline
S\textbf{pecial Qualities:} & Alter shape, DR 10/epic, darkvision 120 ft., immunities, 
regeneration 50, spell resistance 39 \tabularnewline
\hline
S\textbf{aves:} & Fort +40, Ref +39, Will +32 \tabularnewline
\hline
A\textbf{bilities:} & Str 53, Dex 37, Con 39, Int 20, Wis 23, Cha 34 \tabularnewline
\hline
S\textbf{kills:} & Bluff +59, Diplomacy +53, Disguise +59, Hide +35, Intimidate 
+47, Jump +29, Listen +55, Move Silently +35, Sense Motive +29, Spot +55\tabularnewline
\hline
F\textbf{eats:} & Alertness, Cleave, Combat Reflexes, Endurance, Great Cleave, 
Great Fortitude, Improved Initiative, Iron Will, Lightning Reflexes, Multiattack, 
Power Attack, Weapon Focus (slam) \tabularnewline
\hline
E\textbf{pic Feats:} & Blinding Speed, Devastating Critical (slam), Overwhelming 
Critical (slam) \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary \tabularnewline
\hline
C\textbf{hallenge Rating:} & 29 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Always chaotic neutral \tabularnewline
\hline
A\textbf{dvancement:} & 45-59 HD (Large); 60-74 HD (Huge); 75-89 HD (Huge) \tabularnewline
\hline
\end{tabular}

A hagunemnon's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

COMBAT 

\textbf{Psionics (Sp):} At will ---\textit{detect thoughts, dimension door, ethereal 
jaunt, knock, nondetection, plane shift, suggestio}n. Caster level 20th, save DC 
22. 

\textbf{Destabilize Form (Su):} If a protean hits with two or more slam attacks 
in the same round, it partly destabilizes its foe's physical form, dealing an additional 
4d6+31 points of Damage as well as 2 points of temporary Constitution Damage. If 
the victim is drained to 0 Constitution, its dead body is only so much clear fluid 
that drains away unless somehow preserved by friends of the victim. 

\textbf{Alter Shape (Ex):} A protean can assume the shape of any combination of 
physical nondeific creatures at the same time as a free action. In fact, a protean's 
form constantly boils, and it requires a move-equivalent action each round for 
a protean to maintain a certain shape (even if that shape is a combination of several 
shapes). Whatever its present form, the protean retains all its own special qualities. 
Plus, it gains the advantage of up to four extraordinary abilities from the forms 
it mimics (but not spell-like or supernatural powers). The assumed form can be 
no smaller than a flea and no larger than 200 feet in its largest dimension (make 
sure to take into account rules for reach and size modifiers to AC and melee attacks). 
Incorporeal traits can also be assumed, which counts as a single extraordinary 
ability. If a hagunemnon assumes a partial form that confers an extraordinary ability 
already possessed by the creature, only the better of the two abilities is retained. 
No matter its form, the protean can never make more than five attacks using a full-round 
action. However, it may substitute a melee attack form for one of its slam attacks, 
using its own base attack bonus and Strength modifier to Damage, but dealing base 
Damage appropriate to the attack type. 

\vspace{12pt}
{\LARGE{}HOARY HUNTER}

\begin{tabular}{|>{\raggedright}p{57pt}|>{\raggedright}p{128pt}|>{\raggedright}p{128pt}|}
\hline
  & H\textbf{oary Hunter } & H\textbf{oary Steed }\tabularnewline
\hline
  & Medium-Size Fey (Cold)  & Large Magical Beast (Cold) \tabularnewline
\hline
H\textbf{it Dice:}  & 46d6+598 (759 hp)  & 12d10+36 (102 hp) \tabularnewline
\hline
I\textbf{nitiative:}  & +19 (+11 Dex, +8 Superior Initiative)  & +8 (+4 Dex, +4 
Improved Initiative) \tabularnewline
\hline
S\textbf{peed:}  & 30 ft.  & 40 ft., fly 90 ft. (good) \tabularnewline
\hline
A\textbf{C:}  & 46 (+11 Dex, +15 insight, +10 natural)  & 23 (-1 size, +4 Dex, 
+10 natural) \tabularnewline
\hline
B\textbf{ase Attack/}\linebreak{}
\textbf{Grapple:} & +23/+37 & +12/+22\tabularnewline
\hline
A\textbf{ttack:} & +\textit{6 keen longsword of binding }+46 (1d8+27/17-20 plus 
1d6 plus \textit{binding}) melee & Hoof +17 (3d6+6) melee\tabularnewline
\hline
F\textbf{ull Attack:} & +\textit{6 keen longsword of binding }+46/+41/+36/+31 (1d8+27/17-20 
plus 1d6 plus \textit{binding}) melee & 2 hooves +17 (3d6+6) melee, bite +15 (5d8+3) 
melee \tabularnewline
\hline
S\textbf{pace/Reach:}  & 5 ft./5 ft.  & 10 ft./5 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:}  & Spell-like abilities  & \tabularnewline
\hline
S\textbf{pecial Qualities:}  & Cold immunity, fire vulnerability, SR 36, DR 10/epic 
and cold iron  & Cold immunity, fire vulnerability, air walk, astral projection 
and etherealness, magic circle against good, misty breath, SR 20, DR 5/magic and 
cold iron, immunities \tabularnewline
\hline
S\textbf{aves:}  & Fort +28, Ref +36, Will +31  & Fort +11, Ref +12, Will +5 \tabularnewline
\hline
A\textbf{bilities:}  & Str 38, Dex 33, Con 36, Int 21, Wis 23, Cha 26  & Str 22, 
Dex 18, Con 17, Int 6, Wis 13, Cha 14 \tabularnewline
\hline
S\textbf{kills:}  & Diplomacy +14, Hide +60, Intimidate +57, Knowledge (geography, 
nature) +54, Listen +55, Move Silently +60, Ride +60, Search +54, Sense Motive 
+55, Spot +55, Survival +55 & Jump +10, Listen +8, Spot +13\tabularnewline
\hline
F\textbf{eats: } & Blind-Fight, Cleave, Great Cleave, Improved Critical (longsword), 
Improved Initiative, Mounted Combat, Power Attack, Ride-by Attack, Spirited Charge, 
Track, Trample, Weapon Focus (longsword)  & Alertness, Endurance, Flyby Attack, 
Improved Initiative, Multiattack\tabularnewline
\hline
E\textbf{pic Feats: } & Dire Charge, Epic Weapon Focus (longsword), Overwhelming 
Critical (longsword), Superior Initiative  & \tabularnewline
\hline
C\textbf{limate/Terrain:}  & Any cold  & Any cold \tabularnewline
\hline
O\textbf{rganization:}  & Solitary or company (2-5)  & Solitary (with rider) \tabularnewline
\hline
C\textbf{hallenge Rating:}  & 25  & 9 \tabularnewline
\hline
T\textbf{reasure: } & None  & None \tabularnewline
\hline
A\textbf{lignment:}  & Always neutral evil  & Always neutral evil \tabularnewline
\hline
A\textbf{dvancement:}  & 47+ HD (Medium)  & 13-24 HD (Large); 25-36 HD (Huge) \tabularnewline
\hline
\end{tabular}

A hoary hunter's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

COMBAT 

A critical hit (or the blow that would kill or render a foe unconscious) from the 
hoary hunter's blade does not wound the hunter's prey. Instead, the victim is bodily 
transported as per a \textit{soul binding }spell heightened to 16th level (DC 30) 
into a diamond at the end of the sword's hilt. 

\textbf{Spell-Like Abilities: }At will---\textit{discern location, true strike, 
fog cloud, hold monster, plane shift; }3/day---\textit{dimensional anchor, dominate 
monster, greater dispel magic; }1/day---mage's \textit{disjunction, contingent 
recall and resurrection*. }Caster level 23rd; save DC 18 + spell level. \textit{*Contingent 
recall and resurrection }is a variation on an epic spell. When slain, the hoary 
hunter is actually transported, with its mount, back to the fey realm, where it 
is resurrected. The only way to truly slay the hoary hunter is to defeat it in 
its home realm after this epic spell has been exhausted for the day. 

\vspace{12pt}
\section*{HOARY STEED }

A hoary steed's natural weapons are treated as magic for the purpose of overcoming 
damage reduction.

Combat 

\textbf{Air Walk (Su):} A hoary steed can ride through the air as well as on the 
ground, as though affected by a permanent \textit{air walk }spell. 

\textbf{Astral Projection and Etherealness (Su):} These abilities function just 
like the spells of the same names as cast by a 20th-level caster. 

\textbf{Magic Circle Against Good (Su):} This ability continuously duplicates the 
effects of the spell. A hoary steed cannot suppress this ability. 

\textbf{Misty Breath (Su):} The breath of a hoary steed functions like a constant 
\textit{obscuring mist }spell. 

\textbf{Immunities (Ex):} A hoary steed is immune to all \textit{charm }and \textit{hold 
}spells or abilities. 

\vspace{12pt}
{\LARGE{}HUNEFER }

\begin{tabular}{|>{\raggedright}p{63pt}|>{\raggedright}p{262pt}|}
\hline
  & Medium-Size Undead \tabularnewline
\hline
H\textbf{it Dice:} & 50d12+3 (603 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +16 (+12 Dex, +4 Improved Initiative)\tabularnewline
\hline
S\textbf{peed:} & 100 ft. \tabularnewline
\hline
A\textbf{C:} & 52 (+12 Dex, +20 natural, +10 insight) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +25/+43\tabularnewline
\hline
A\textbf{ttack:} & Slam +44 (3d6+18/19-20 (+1d6 on critical hit) plus hunefer rot)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 slams +44 (3d6+18/19-20 (+1d6 on critical hit) plus hunefer 
rot)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Despair, hunefer rot, spell-like abilities \tabularnewline
\hline
S\textbf{pecial Qualities:} & Blindsight 300 ft., DR 10/-, fast healing 30, SR 
37, undead traits, fire vulnerability \tabularnewline
\hline
S\textbf{aves:} & Fort +18, Ref +30, Will +41 \tabularnewline
\hline
A\textbf{bilities:} & Str 47, Dex 35, Con -, Int 18, Wis 38, Cha 36 \tabularnewline
\hline
S\textbf{kills:} & Concentration +66, Jump +46, Knowledge (religion, the planes) 
+57, Listen +67, Search +57, Sense Motive +67, Spellcraft +57, Spot +67\tabularnewline
\hline
F\textbf{eats:} & Cleave, Combat Reflexes, Dodge, Expertise, Great Cleave, Great 
Fortitude, Improved Critical (slam), Improved Disarm, Improved Initiative, Lightning 
Reflexes, Mobility, Power Attack, Toughness, Weapon Focus (slam) \tabularnewline
\hline
E\textbf{pic Feats:} & Devastating Critical (slam), Overwhelming Critical (slam), 
Spell Stowaway \textit{(greater teleport}) \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or court (5-9) \tabularnewline
\hline
C\textbf{hallenge Rating:} & 25 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil \tabularnewline
\hline
A\textbf{dvancement:} & 51-56 HD (Medium-size); 57-60 HD (Large); 61-66 HD (Huge); 
67-70 HD (Gargantuan); 71+ HD (Colossal) \tabularnewline
\hline
\end{tabular}

Hunefers speak all spoken languages, though it takes one 1d4 rounds to switch between 
languages. 

COMBAT 

\textbf{Despair (Su):} The mere sight of a hunefer requires viewers to succeed 
at a Will saving throw (DC 48) or be paralyzed with fear for 1d4 rounds. Regardless 
of the outcome of the save, a subject is not again vulnerable to that hunefer's 
despair ability for 24 hours. The DC is Charisma-based.

\textbf{Hunefer Rot (Su):} Supernatural disease---slam, Fort save (DC 35), incubation 
period instantaneous; Damage 1d6 temporary Con. Unlike normal diseases, hunefer 
rot requires a victim to make a successful saving throw every round or take another 
1d6 points of temporary Constitution Damage. The rot continues until the victim 
reaches Constitution 0 (and dies) or receives a \textit{remove disease }spell or 
similar magic. An afflicted creature that dies shrivels away into sand unless both 
\textit{remove disease }and \textit{raise dead }(or better) are cast on the remains 
within 2 rounds. If the remains are not so treated, on the third round the dust 
swirls and forms an 18 HD mummy with the dead foe's equipment under the hunefer's 
command. (The \textit{mummy dust }epic spell has statistics for an 18 HD mummy.) 

\textbf{Spell-Like Abilities: }At will---\textit{chain lightning, displacement, 
haste, knock, greater dispel magic, passwall, scrying, greater teleport, true seeing, 
unholy aura, weird; }1/day---\textit{ruin }(epic spell). Caster level 27th; save 
DC 23 + spell level. The DC is Charisma-based.

\textbf{Undead Traits:} Immune to poison, \textit{sleep, }paralysis, stunning, 
disease, death, effects, necromantic effects, mind-affecting effects, and any effect 
requiring a Fortitude save unless it also works on objects. Not subject to critical 
hits, subdual dam-age, ability Damage, ability drain, or energy drain. Negative 
energy heals. Not at risk of death from massive Damage, but destroyed at 0 hit 
points or less. Darkvision 60 ft. Cannot be raised; resurrection works only if 
creature is willing. 

\vspace{12pt}
{\LARGE{}LAVAWIGHT }

\begin{tabular}{|>{\raggedright}p{64pt}|>{\raggedright}p{261pt}|}
\hline
  & Medium-Size Undead (Fire) \tabularnewline
\hline
H\textbf{it Dice:} & 32d12 (208 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +12 (+8 Dex, +4 Improved Initiative) \tabularnewline
\hline
S\textbf{peed:} & 60 ft. \tabularnewline
\hline
A\textbf{C:} & 46 (+8 Dex, +28 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +16/+37\tabularnewline
\hline
A\textbf{ttack:} & Claw +40 (3d8+21/19-20 (+1d6 on critical hit) plus blazefire) 
melee\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +40 (3d8+21/19-20 (+1d6 on critical hit) plus blazefire) 
melee, 1 skull butt +32 (2d6+10 plus blazefire) melee\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Rend 6d8+31, blazefire, spell-like abilities \tabularnewline
\hline
S\textbf{pecial Qualities:} & Undead traits, fire subtype, heat aura, SR 34, DR 
10/- , fiery hardness \tabularnewline
\hline
S\textbf{kills:} & Jump +33, Knowledge (religion) +33, Listen +41, Search +35, 
Spot +41\tabularnewline
\hline
F\textbf{eats:} & Cleave, Great Cleave, Great Fortitude, Improved Critical (claw), 
Improved Initiative, Power Attack, Weapon Focus (claw) \tabularnewline
\hline
E\textbf{pic Feats:} & Blinding Speed (x2), Epic Weapon Focus (claw), Overwhelming 
Critical (claw) \tabularnewline
\hline
S\textbf{aves:} & Fort +12, Ref +18, Will +24 \tabularnewline
\hline
A\textbf{bilities:} & Str 52, Dex 27, Con -, Int 10, Wis 22, Cha 28 \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary or pair \tabularnewline
\hline
C\textbf{hallenge Rating:} & 23 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Chaotic evil \tabularnewline
\hline
A\textbf{dvancement:} & 33-42 HD (Medium-size); 43-55 HD (Huge)\tabularnewline
\hline
\end{tabular}

Lavawights speak Common and Infernal. 

COMBAT 

\textbf{Rend (Ex): }If the lavawight hits with both of its claws, it latches onto 
the opponent's body and tears the flesh. This attack automatically deals an additional 
6d8+31 points of Damage. 

\textbf{Blazefire (Su): }Living creatures taking Damage from a lavawight's attack 
find themselves ignited with blaze-fire; they must succeed at a Fortitude save 
(DC 35) or permanently lose 4 hit points. The DC is Charisma-based. The opponent 
must continue to save every round for the next 6 rounds (7 rounds total) to avoid 
being permanently drained of 4 more hit points each round. The lavawight heals 
the same amount of Damage whenever a creature is drained of hit points, gaining 
any excess as temporary hit points. Tehse temporary hit points last a maximum of 
1 hour. If an opponent is slain by blazefire, only blackened ash remains of the 
victim. Hit points lost to the blazefire never heal naturally and cannot be magically 
restored---they are gone for good. 

\textbf{Spell-Like Abilities: }At will--- \textit{dimension door, flame arrow, 
spell immunity (sleet storm}), \textit{fireball, wall of fire. }Caster level 22nd; 
save DC 19 + spell level. The DC is Charisma-based.

\textbf{Undead Traits:} Immune to poison, \textit{sleep, }paralysis, stunning, 
disease, death, effects, necromantic effects, mind-affecting effects, and any effect 
requiring a Fortitude save unless it also works on objects. Not subject to critical 
hits, subdual Damage, ability Damage, ability drain, or energy drain. Negative 
energy heals. Not at risk of death from massive Damage, but destroyed at 0 hit 
points or less. Darkvision 60 ft. Cannot be raised; resurrection works only if 
creature is willing. 

\textbf{Heat Aura (Su):} A 10-foot-radius spread heat aura surrounds a lavawight. 
All creatures of the fire subtype in the field (including the lavawight) are treated 
as if having turn resistance +6 (if undead) and fast healing 10. Creatures subject 
to fire Damage take 2d10 points of fire Damage each round they remain within the 
heat aura. 

\textbf{Fire Subtype:} Fire immunity, cold vulnerability.. 

\textbf{Fiery Hardness (Ex):} The lavawight's molten skin provides it with Damage 
reduction of 10/-. 

\vspace{12pt}
{\LARGE{}LEGENDARY ANIMAL}

\begin{tabular}{|>{\raggedright}p{60pt}|>{\raggedright}p{126pt}|>{\raggedright}p{126pt}|}
\hline
  & L\textbf{egendary Bear } & L\textbf{egendary Tiger }\tabularnewline
\hline
  & Large Animal  & Large Animal \tabularnewline
\hline
H\textbf{it Dice:}  & 20d8+140 (230 hp)  & 26d8+188 (305 hp) \tabularnewline
\hline
I\textbf{nitiative:}  & +6 (+2 Dex, +4 Improved Initiative)  & +8 (+4 Dex, +4 Improved 
Initiative) \tabularnewline
\hline
S\textbf{peed:}  & 50 ft.  & 50 ft. \tabularnewline
\hline
A\textbf{C:}  & 21 (-1 size, +2 Dex, +10 natural)  & 23 (-1 size, +4 Dex, +10 natural) 
\tabularnewline
\hline
B\textbf{ase Attack/}\linebreak{}
\textbf{Grapple:} & +15/+32 & +19/+34\tabularnewline
\hline
A\textbf{ttack:} & Claw +27 (2d6+13) melee & Claw +29 (2d6+11) melee\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +27 (2d6+13) melee, bite +22 (4d6+6) melee & 2 
claws +29 (2d6+11) melee, bite +24 (2d8+5) melee\tabularnewline
\hline
S\textbf{pace/Reach:}  & 10 ft./5 ft.  & 10 ft./5 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:}  & Improved grab  & Pounce, improved grab, rake 2d6+5 
\tabularnewline
\hline
S\textbf{pecial Qualities:}  & Scent  & Scent \tabularnewline
\hline
S\textbf{aves:}  & Fort +19, Ref +14, Will +11  & Fort +22, Ref +19, Will +12 \tabularnewline
\hline
A\textbf{bilities:}  & Str 36, Dex 14, Con 24, Int 2, Wis 16, Cha 12  & Str 32, 
Dex 18, Con 24, Int 2, Wis 14, Cha 10 \tabularnewline
\hline
S\textbf{kills:}  & Climb +15, Jump +21, Listen +14, Spot +14, Swim +20 & Climb 
+13, Hide +15, Jump +19, Listen +12, Move Silently +15, Spot +12, Swim +16\tabularnewline
\hline
F\textbf{eats: } & Alertness, Athletic, Endurance, Improved Initiative, Iron Will, 
Run, Track & Alertness, Athletic, Endurance, Improved Initiative, Iron Will, Run, 
Toughness(x2), Track\tabularnewline
\hline
C\textbf{limate/Terrain:}  & Any forest, hill, mountains, plains, and underground 
 & Any forest, hill, mountains, plains, or underground \tabularnewline
\hline
O\textbf{rganization:}  & Solitary or pair  & Solitary or pair \tabularnewline
\hline
C\textbf{hallenge Rating:}  & 9  & 10 \tabularnewline
\hline
T\textbf{reasure: } & None  & None \tabularnewline
\hline
A\textbf{lignment:}  & Always neutral  & Always neutral \tabularnewline
\hline
A\textbf{dvancement:}  & 21-40 HD (Large)  & 27-48 HD (Large) \tabularnewline
\hline
\end{tabular}

LEGENDARY BEAR 

COMBAT 

\textbf{Improved Grab (Ex): }To use this ability, a legendary bear must hit with 
a claw attack. 

LEGENDARY TIGER 

COMBAT 

\textbf{Pounce (Ex):} If a legendary tiger leaps upon a foe during the first round 
of combat, it can make a full attack even if it has already taken a move action. 

\textbf{Improved Grab (Ex): }To use this ability, the legendary tiger must hit 
with a bite attack. If it gets a hold, it can rake. 

\textbf{Rake (Ex):} A legendary tiger can make two attacks (+30 melee) against 
a held creature with its hind legs for 2d6+5 points of Damage each. If the legendary 
tiger pounces on an opponent, it can also rake. 

\textbf{Skills:} Legendary tigers receive a +8 racial bonus on Hide and Move Silently 
checks. *In areas of tall grasses or heavy undergrowth, the Hide bonus improves 
to +16. 

\vspace{12pt}
{\LARGE{}LESHAY }

\begin{tabular}{|>{\raggedright}p{66pt}|>{\raggedright}p{259pt}|}
\hline
  & Medium-Size Fey \tabularnewline
\hline
H\textbf{it Dice:} & 50d6+650 (825 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +21 (+17 Dex, +4 Improved Initiative) \tabularnewline
\hline
S\textbf{peed:} & 40 ft. \tabularnewline
\hline
A\textbf{C:} & 52 (+17 Dex, +20 insight, +5 natural armor) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +25/+30\tabularnewline
\hline
A\textbf{ttack:} & +10 \textit{keen brilliant energy bastard sword} +53 (1d10+15/17-20) 
melee\tabularnewline
\hline
F\textbf{ull Attack:} & 2 \textit{+10 keen brilliant energy bastard swords }+53/+48/+43/+38 
(1d10+15/17-20) melee\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Gaze, spell-like abilities, leShay weapons \tabularnewline
\hline
S\textbf{pecial Qualities:} & Superior two-weapon fighting, DR 15/epic and cold 
iron, elf traits, immune to poison and disease, low-light vision, SR 42, fast healing 
10 \tabularnewline
\hline
S\textbf{aves:} & Fort +29, Ref +44, Will +35 \tabularnewline
\hline
A\textbf{bilities:} & Str 21, Dex 45, Con 37, Int 33, Wis 23, Cha 47 \tabularnewline
\hline
S\textbf{kills:} & Bluff +71, Concentration +66, Diplomacy +77, Disguise +71, Escape 
Artist +70, Gather Information +24, Hide +70, Intimidate +24, Jump +5, Knowledge 
(local, nature) +59, Listen +59, Move Silently +70, Search +59, Speak Language 
(any five), Spot +59\tabularnewline
\hline
F\textbf{eats:} & Blind-Fight, Cleave, Dodge, Expertise, Improved Critical (leShay 
weapon), Improved Initiative, Iron Will, Mobility, Power Attack, Quick Draw, Spring 
Attack, Weapon Finesse, Weapon Focus (leShay weapon), Whirlwind Attack \tabularnewline
\hline
E\textbf{pic Feats:} & Blinding Speed, Polyglot, Spell Stowaway \textit{(heal}) 
\tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary \tabularnewline
\hline
C\textbf{hallenge Rating:} & 28 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Any \tabularnewline
\hline
A\textbf{dvancement:} & None \tabularnewline
\hline
\end{tabular}

A leshay's natural weapons are treated as epic for the purpose of overcoming damage 
reduction.

\section*{COMBAT }

\textbf{Spell-Like Abilities: }At will---\textit{alter self, detect thoughts, displacement, 
freedom, greater dispelling, heal, greater invisibility, knock, speak with plants, 
spell turning, greater teleport, water breathing. }Caster level 28th; save DC 28 
+ spell level. The DCs are Charisma-based.

\textbf{LeShay Weapons (Su):} Each leShay carries two personal melee weapons that 
he or she manifests from his or her own life essence. The weapon can change form 
as he or she shifts from class to class, typically appearing as a bastard sword 
(one in each hand). When not in combat, the weapons simply do not exist, but leShay 
can ``draw'' their weapons from nowhere as a free action. LeShay weapons are considered 
light weapons for a leShay, regardless of their size. 

\textbf{Superior Two-Weapon Fighting (Ex):} A leShay fights with a leShay weapon 
in each hand. Because these weapons are extensions of its own body, the leShay 
does not take an attack or Damage penalty for attacking with two weapons. 

\textbf{Gaze (Su):} All creatures within 30 feet who meet the gaze of a leShay 
become affected as if by \textit{charm monste}r. A Will save (DC 53) negates the 
effect. Each opponent within range of a gaze attack must attempt a saving throw 
each round at the beginning of his or her turn in the initiative order. A leShay 
can also actively gaze as an attack action by choosing a target within range, who 
must then attempt a saving throw. LeShay are immune to their own gaze. The DC is 
Charisma-based.

\textbf{Elf Traits:} Immune to magic \textit{sleep }spells and effects; +2 racial 
bonus on Will saves against enchantment spells or effects; low-light vision (can 
see twice as far as a human in low-light conditions); entitled to a Search check 
when within 5 feet of a secret or concealed door as though actively looking for 
it; Martial Weapon Proficiency (composite longbow, composite shortbow, longbow, 
longsword, and rapier) as bonus feats; +2 racial bonus on Listen, Spot, and Search 
checks (already figured into the statistics given above). 

\vspace{12pt}
{\LARGE{}LIVING VAULT }

\begin{tabular}{|>{\raggedright}p{68pt}|>{\raggedright}p{257pt}|}
\hline
  & Colossal Construct \tabularnewline
\hline
H\textbf{it Dice:} & 96d10+80 (608 hp)\tabularnewline
\hline
I\textbf{nitiative:}- & 2 (Dex) \tabularnewline
\hline
S\textbf{peed:} & 10 ft. (can't run) \tabularnewline
\hline
A\textbf{C:} & 60 (-8 size, -2 Dex, +60 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +72/+123\tabularnewline
\hline
A\textbf{ttack:} & Slam +99 (10d6+35) melee\tabularnewline
\hline
F\textbf{ull Attack:} & 2 slams +99 (10d6+35) melee\tabularnewline
\hline
S\textbf{pace/Reach:} & 30 ft./20 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & I\textit{mprisonment}\tabularnewline
\hline
S\textbf{pecial Qualities:} & Recognition, \textit{safekeeping, }construct traits, 
magic immunity, DR 15/epic and adamantine \tabularnewline
\hline
S\textbf{aves:} & Fort +32, Ref +30, Will +33 \tabularnewline
\hline
S\textbf{kills:} & Jump +0\tabularnewline
\hline
A\textbf{bilities:} & Str 80, Dex 7, Con -, Int -, Wis 12, Cha 1 \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary \tabularnewline
\hline
C\textbf{hallenge Rating:} & 33 \tabularnewline
\hline
T\textbf{reasure:} & Standard (see text) \tabularnewline
\hline
A\textbf{lignment:} & Neutral \tabularnewline
\hline
A\textbf{dvancement:} & None \tabularnewline
\hline
\end{tabular}

To access a vault, its creator must call the vault from where it is sequestered 
through the use of a \textit{sending }spell. The vault has no obvious opening; 
however, items can be stored and retrieved by naming them (the living vault provides 
an inventory to its creator on request). The item to be stored appears to sink 
into the living vault's form as if made of mud. Likewise, items retrieved pop out 
of the living vault's form. Storing and retrieving items in this fashion takes 
one round per transaction. Creatures other than the vault's master who wish to 
access the inventory must destroy the vault to do so. If a vault is destroyed, 
all the objects and creatures stored within immediately emerge. 

A living vault's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

COMBAT 

\textit{\textbf{Imprisonment }}\textbf{(Sp):} As a free action twice per day, a 
living vault can project a 60-foot cone. Creatures caught in the cone must make 
a Will save (DC 58) or be affected as by the \textit{imprisonment }spell, except 
creatures are stored within the vault's interior. The living vault can hold two 
Gargantuan creatures, eight Huge creatures, thirty-two Large creatures, one hundred 
twenty-eight Medium-size creatures, and so on. This assumes that the rest of the 
vault is also relatively empty, so a large hoard of treasure reduces the number 
of creatures a living vault can hold at any one time. If \textit{freedom }is later 
cast upon the very spot where the victim of \textit{imprisonment }disappeared, 
the victim is freed, though it doesn't appear where it disappeared, but in a space 
adjacent to wherever the living vault is currently physically located. 

\textbf{Recognition (Ex):} Once made, a living vault responds only to its owner; 
no other being can access the vault's interior. The vault is not fooled by disguised 
creatures, and can distinguish between its true master and false duplicates, even 
penetrating disguises used in conjunction with \textit{shapechange, alter self, 
disguise self, polymorph, seeming, }and \textit{simulacrum }and similar effects. 
The vault does, however, respond to the product of a \textit{clone }spell. 

\textit{\textbf{Safekeeping }}\textbf{(Sp):} At will, a living vault can use a 
power that combines the effects of \textit{sequester }and \textit{plane shift }to 
render the vault invisible to any form of sight or divination, and at the same 
time transfer it to a random portion of the Astral Plane. When in safekeeping mode, 
the living vault responds to a \textit{sending }by its master, returning to the 
location from which it left and becoming tangible once again. Creatures and objects 
that are stored within the vault are treated as if under the effects of \textit{sequester, 
}whether the vault itself is or not. Often, creatures are brought into the vault 
via \textit{imprisonment, }but sometimes unconscious or otherwise subdued creatures 
are stored as an object, in which case they receive no saving throw to resist the 
effect. 

\textbf{Construct Traits:} Immune to mind-affecting effects (charms, compulsions, 
phantasms, patterns, and morale effects), and to poison, \textit{sleep, }paralysis, 
stunning, disease, death effects, necromantic effects, and any effect that requires 
a Fortitude save unless it also works on objects. Cannot heal Damage (though regeneration 
and fast healing still apply, if present). Not subject to critical hits, subdual 
Damage, ability Damage, ability drain, or energy drain. Not at risk of death from 
massive Damage, but destroyed when reduced to 0 hit points or less; cannot be raised 
or resurrected. Darkvision 60 ft. 

\textbf{Magic Immunity (Ex):} A living vault is immune to all magical and supernatural 
effects, except as follows. A \textit{freedom }spell cast on it causes it to defer 
the use of its own \textit{imprisonment }power for 10 rounds. 

\vspace{12pt}
{\LARGE{}MERCANE }

\begin{tabular}{|>{\raggedright}p{66pt}|>{\raggedright}p{259pt}|}
\hline
  & Large Outsider\tabularnewline
\hline
H\textbf{it Dice:} & 7d8+21 (52 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +2 (Dex) \tabularnewline
\hline
S\textbf{peed:} & 30 ft. \tabularnewline
\hline
A\textbf{C:} & 15 (-1 size, +2 Dex, +4 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +7/+13\tabularnewline
\hline
A\textbf{ttack:} & Masterwork falchion +9(2d4+3) melee\tabularnewline
\hline
F\textbf{ull Attack:} & Masterwork falchion +9/+4 (2d4+3) melee\textbf{ }\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft. \tabularnewline
\hline
S\textbf{pecial Qualities:} & SR 25, spell-like abilities, telepathy\tabularnewline
\hline
S\textbf{aves:} & Fort +8, Ref +7, Will +8 \tabularnewline
\hline
A\textbf{bilities:} & Str 15, Dex 15, Con 16, Int 20, Wis 17, Cha 15 \tabularnewline
\hline
S\textbf{kills:} & Appraise +19, Bluff +12, Diplomacy +12, Gather Information +12, 
Intimidate +12, Knowledge (arcana) +15, Knowledge (the planes) +15, Listen +13, 
Profession (bookkeeper) +13, Sense Motive +13, Sleight of Hand +14, Spellcraft 
+15, Spot +12 \tabularnewline
\hline
F\textbf{eats:} & Expertise, Improved Disarm, Improved Feint\tabularnewline
\hline
C\textbf{limate/Terrain:} & Any land and underground \tabularnewline
\hline
O\textbf{rganization:} & Company (1-4 mercanes and 3-18 5th-level fighter bodyguards) 
or progression (1 mercane, 3 14th-level fighters, 1 21st-level fighter/2nd-level 
Union Sentinel) \tabularnewline
\hline
C\textbf{hallenge Rating:} & 5 \tabularnewline
\hline
T\textbf{reasure:} & Double standard \tabularnewline
\hline
A\textbf{lignment:} & Always lawful neutral \tabularnewline
\hline
A\textbf{dvancement:} & By character class \tabularnewline
\hline
L\textbf{evel Adjustment:} & +7\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
Mercanes speak Abyssal, Celestial, Draconic, Infernal, and at least two other languages. 

\section*{COMBAT }

\textbf{Spell-Like Abilities: }3/day---\textit{dimension door, invisibility; }1/day---\textit{secret 
chest, plane shift. }Caster level 14th; save DC 12 + spell level. 

\textbf{Telepathy (Su):} Mercanes can communicate telepathically with any creature 
within 100 feet that has a language. 

\textbf{Skills:} Mercanes receive a +4 racial bonus on Appraise checks. 

\vspace{12pt}
\section*{MERCANE CHARACTERS }

A mercane's favored class is wizard. Player character mercanes have a level adjustment 
of +7. 

\vspace{12pt}
{\LARGE{}MU SPORE }

\begin{tabular}{|>{\raggedright}p{66pt}|>{\raggedright}p{259pt}|}
\hline
  & Colossal Plant \tabularnewline
\hline
H\textbf{it Dice:} & 35d8+315 (472 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +5 (-3 Dex, +8 Superior Initiative) \tabularnewline
\hline
S\textbf{peed:} & 40 ft. \tabularnewline
\hline
A\textbf{C:} & 27 (-3 Dex, -8 size, +28 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +26/+55\tabularnewline
\hline
A\textbf{ttack:} & Tendril +31 (2d6+13/19-20) melee\tabularnewline
\hline
F\textbf{ull Attack:} & 4 tendrils +31 (2d6+13/19-20) melee, 1 bite +26 (4d6+6/19-20) 
melee\tabularnewline
\hline
S\textbf{pace/Reach:} & 30 ft./30 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Spore cough, improved grab, swallow whole \tabularnewline
\hline
S\textbf{pecial Qualities:} & Acid resistance 20, blindsight 210 ft., DR 10/epic, 
fast healing 10, plant traits, sticky\tabularnewline
\hline
S\textbf{aves:} & Fort +28, Ref +8, Will +20 \tabularnewline
\hline
A\textbf{bilities:} & Str 36, Dex 5, Con 29, Int 18, Wis 28, Cha 28 \tabularnewline
\hline
S\textbf{kills:} & Bluff +47, Diplomacy +13, Gather Information +13, Hide +35, 
Intimidate +13, Jump +13, Knowledge (local, nature) +42, Listen +47, Spot +47\tabularnewline
\hline
F\textbf{eats:} & Cleave, Great Cleave, Improved Critical (bite), Improved Critical 
(tendril), Improved initiative, Lightning Reflexes, Power Attack\tabularnewline
\hline
E\textbf{pic Feats:} & Epic Reflexes, Epic Toughness (x3), Superior Initiative\tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or polyp (3-7) \tabularnewline
\hline
C\textbf{hallenge Rating:} & 21 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic neutral \tabularnewline
\hline
A\textbf{dvancement:} & 36+ HD (Colossal) \tabularnewline
\hline
\end{tabular}

Mu spores speak Common and Terran. 

A mu spore's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

COMBAT

\textbf{Spore Cough (Su):} Once every 1d4 rounds a mu spore can release a cloud 
of burrowing spores. This spore cough is treated as a 100-foot-cone breath weapon. 
The burrowing spores deal 20d8 points of Damage to all creatures and structures 
in the area, or half Damage to any creatures that make a Reflex save (DC 36). The 
DC is Charisma-based.

\textbf{Improved Grab (Ex):} If a mu spore hits with a tendril, it deals normal 
Damage and attempts to start a grapple as a free action without provoking an attack 
of opportunity. Improved grab works only against Gargantuan and smaller opponents. 
The mu spore has the option to conduct the grapple normally, transfer the victim 
to its mouth as a free action, or simply use its tendril to hold the opponent. 
Each successful grapple check it makes during successive rounds automatically deals 
the Damage given for the tendril. 

\textbf{Swallow Whole (Ex):} If the mu spore begins its turn with an opponent held 
in its mouth, it can attempt a new grapple check (as though attempting to pin the 
opponent). If it succeeds, it swallows its prey and the opponent takes bite Damage. 
A swallowed creature takes the twice the Damage given for the mu spore's bite every 
round. A swallowed creature is considered grappled, while the creature that did 
the swallowing is not. A swallowed creature can try to cut its way free with any 
light piercing or slashing weapon, if it can deal 33 points of Damage in this fashion, 
or it can just try to escape the grapple. If the swallowed creature chooses the 
latter course, success puts it back in the mu spore's mouth, where it may be bitten 
or swallowed again. 

\textbf{Sticky (Ex): }Sticky tendrils cover a mu spore, many more than the four 
main tendrils it uses for combat. Any creature that touches or successfully attacks 
a mu spore with a melee weapon is subject to a free improved grab check by the 
mu spore, as if the mu spore had success-fully attacked the creature with a tendril. 
A creature wielding a melee weapon must immediately relinquish its grip on its 
weapon, or proceed with the grapple check. If the subject relinquishes its grip 
on its weapon, the weapon is swallowed the following round. Creatures who attack 
the mu spore in melee without a weapon they can release (such as those with natural 
attacks only) do not have this option. 

\textbf{Plant Traits:} Immune to poison, \textit{sleep, }paralysis, stunning, and 
polymorphing; not subject to critical hits or mind-affecting effects. 

\vspace{12pt}
{\LARGE{}NEH-THALGGU (BRAIN COLLECTOR) }

\begin{tabular}{|>{\raggedright}p{66pt}|>{\raggedright}p{259pt}|}
\hline
  & Huge Aberration (Extraplanar, Incorporeal) \tabularnewline
\hline
H\textbf{it Dice:} & 32d8+192 (336 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +12 (+4 Dex, +8 Superior Initiative) \tabularnewline
\hline
S\textbf{peed:} & 60 ft. or via \textit{dimension door }\tabularnewline
\hline
A\textbf{C:} & 35 (-2 size, +4 Dex, +3 deflection, +20 insight) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +24/-\tabularnewline
\hline
A\textbf{ttack:} & Bite +26 (4d10 + 12 plus poison) melee\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +26 (4d10 + 12 plus poison ) melee, 10 head-tentacles 
+21 (2d10 plus ability drain ) melee touch attack\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./10 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Extract brains (ranged attack), poison, spells \tabularnewline
\hline
S\textbf{pecial Qualities:} & D\textit{imensional travel, }DR 10/epic, incorporeal 
traits, amorphous physiology, manifest maw, SR 30, darkvision 60 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +16, Ref +14, Will +23 \tabularnewline
\hline
A\textbf{bilities:} & Str -, Dex 19, Con 22, Int 20, Wis 20, Cha 17 \tabularnewline
\hline
S\textbf{kills:} & Concentration +41, Hide +39, Knowledge (arcana, the planes) 
+40, Move Silently +39, Search +40, Spellcraft +40\tabularnewline
\hline
F\textbf{eats:} & Blind-Fight, Combat Reflexes, Dodge, Eschew Materials, Expertise, 
Improved Initiative, Maximize Spell, Mobility, Power Attack, Spring Attack \tabularnewline
\hline
E\textbf{pic Feats:} & Superior Initiative \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 26 \tabularnewline
\hline
T\textbf{reasure:} & Triple standard \tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic neutral, neutral evil, or chaotic evil \tabularnewline
\hline
A\textbf{dvancement:} & 33-66 HD (Gargantuan); 67-112 HD (Colossal) \tabularnewline
\hline
\end{tabular}

A creature whose brain has been harvested by a brain collector cannot be raised, 
reincarnated, or resurrected while the brain is in the creature, because the collector 
preserves and draws upon the soul and basic personality of the creature for as 
long as it retains the brain. Neh-thalggus' own language is a silent sign language 
``spoken'' with their writhing head-tentacles. They can also communicate telepathically 
with any creature that has a language within 100 feet. 

A neh-thalggu's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

COMBAT 

\textbf{Poison (Ex):} A neh-thalggu's bite injects a debilitating poison (DC 32) 
that Damages the victim's Constitution. He or she loses half his or her current 
Constitution immediately on a failed save and must save again 1 minute later or 
lose half of his or her remaining Constitution. If the character survives, he or 
she heals the ability Damage at the normal rate (1 point per day). The save DC 
is Constitution-based.

\textbf{Head-Tentacles (Ex): }The long, whiplike tentacles that frame either side 
of a brain collector's face can deliver a dangerous touch attack that, if successful, 
causes the target to dehydrate and wither (2d10 points of desiccation Damage). 
In addition, each successful tentacle attack drains 1 point of Strength, 1 point 
of Dexterity, and 1 point of Constitution. This is permanent ability drain and 
may only be reversed by spells such as \textit{restoration }and \textit{greater 
restoratio}n. It can bring all ten tentacles to bear on a single target facing 
it or divide its attacks against up to ten targets that it threatens. 

\textit{\textbf{Extract Brains }}\textbf{(Sp):} Once every 1d4 rounds, as a full-round 
action, a brain collector can extract the brain from a target creature in line 
of sight. This attack is psionic in nature and can be blocked by a \textit{dimensional 
anchor }currently in force on the target. The target gets a Will save (DC 31) to 
resist the extraction. The save DC is Intelligence-based. If the save fails, his 
or her brain is drawn out intact through the skull by extradimensional means and 
sucked up by the brain collector, lodging in an unused storage sac above and behind 
its eyes. If the save succeeds, he or she takes 9d6 points of Damage and is stunned 
for 1d4+1 rounds. A brain collector prefers to absorb brains of high-level arcane 
spellcasters but is fully capable of extracting those of other foes as a highly 
effective attack. When encountered, assume a neh-thalggu has a full retinue of 
thirteen stolen brains. Each brain less than the full thirteen bestows one negative 
level on the neh-thalggu (though these never convert to actual level loss), which 
is a powerful incentive for the creature to always keep its brain-sacs filled. 
A neh-thalggu is free to draw on all the Knowledge skills of each brain it currently 
stores, using the base ranks in a skill possessed by each brain, and adjusted by 
the neh-thalggu's own skill modifiers for the Knowledge skill in question (or its 
Intelligence modifier, for a skill it has no ranks in). 

\textit{\textbf{Dimensional Travel }}\textbf{(Sp):} A brain collector's preferred 
method of locomotion is via \textit{dimension door, }which it can do as a quickened 
action, once per round. It can also use \textit{teleport without error }or \textit{plane 
shift }at will as a move-equivalent action.

\textbf{Spells (Sp):} A fully grown brain collector can cast arcane spells as a 
13th-level sorcerer (one level of spell-casting ability per brain).

\textbf{Incorporeal:} A neh-thalggu is not wholly in our reality but always remains 
partially extradimensional. Thus it can be harmed only by other incorporeal creatures, 
+1 or better weapons, magic, or psionics, with a 50\% chance to ignore any Damage 
from a corporeal source. It can pass through solid objects at will, and its own 
attacks pass through armor (except for its bite attack, which is treated as if 
a corporeal attack). It always moves silently unless it chooses otherwise. 

\textbf{Amorphous Physiology (Ex):} A brain collector does not have fixed organs. 
As such, it is immune to critical hits, death from massive Damage, sneak attacks, 
and coup de grace. 

\textbf{Manifest Maw:} Though it is an incorporeal creature, a neh-thalggu can 
manifest its mouth in corporeal form as a standard action. While so manifested, 
the mouth can deliver bite attacks against corporeal creatures or pick up objects. 
The mouth bites as if it had Str 35. 

\textbf{Feats: }A neh-thalggu gains Power Attack as a bonus feat, even though it 
has no Strength score. 

\vspace{12pt}
{\LARGE{}PARAGON CREATURE}

\section*{CREATING A PARAGON CREATURE }

``Paragon'' is a template that can be added to any creature (referred to hereafter 
as the base creature). 

The base creature's type remains unchanged. The paragon creature uses all the base 
creature's statistics and special abilities except as noted here. 

\textbf{HD:} A paragon creature always has maximum hit points. Paragon creatures 
also gain an additional 12 hit points per HD. 

\textbf{Speed:} A paragon creature's speed triples, for all movement types. 

\textbf{AC:} Paragon creatures gain a +12 insight bonus to AC and a +12 luck bonus 
to AC. They also gain a +5 natural armor bonus (if the creature already has natural 
armor, use whichever is better). 

\textbf{Attacks:} A paragon creature makes all its attacks with a +25 luck bonus 
on the attack roll. 

\textbf{Damage:} A paragon creature gains a +20 luck bonus on Damage rolls for 
all melee and thrown ranged attacks. 

\textbf{Special Attacks:} A paragon creature's special attacks, if any, all gain 
a +13 insight bonus, if applicable. The +13 insight bonus may only be applied to 
a given special ability once. 

\textbf{Spell-Like Abilities (Sp) or Psionics (Sp):} If the base creature has spell-like 
abilities, it gains +15 to its caster level to use those abilities. A paragon creature 
also gains the ability to use \textit{greater dispel magic, haste, }and \textit{see 
invisibility }three times per day, even if it did not have spell-like abilities 
before, at 15th caster level. 

\textbf{Special Qualities:} A paragon creature retains all the special qualities 
of the base creature and also gains the following.  

\parindent=3pt
Fire and cold resistance 10. If the creature already possesses such resistance, 
use whichever is better.  

Damage reduction 10/epic. If the creature already possesses Damage reduction, use 
whichever is better.  

Spell resistance equal to the paragon creature's CR +25. If the creature already 
possesses spell resistance, use whichever is higher.  

Fast healing 20. If the creature already possesses fast healing, use whichever 
is better.  

\parindent=7pt
A paragon creature's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

\parindent=0pt
\textbf{Saves:} The paragon creature gains a +10 insight bonus on all its saving 
throws. 

\textbf{Abilities:} All ability scores are 15 points higher than those of the base 
creature. 

\textbf{Skills:} The paragon creature gains a +10 competence bonus on all its skill 
checks. 

\textbf{Feats: }Same as the base creature, plus one bonus feat. 

\textbf{Climate/Terrain:} Any land and underground. 

\textbf{Organization:} Same as the base creature. 

\textbf{Challenge Rating:} As base creature +15.

\textbf{Treasure: }Standard for a creature of the adjusted CR. 

\textbf{Alignment:} Same as the base creature. 

\textbf{Advancement:} Same as the base creature.

\vspace{12pt}
{\LARGE{}PRISMASAURUS }

\begin{tabular}{|>{\raggedright}p{63pt}|>{\raggedright}p{263pt}|}
\hline
  & Huge Magical Beast \tabularnewline
\hline
H\textbf{it Dice:} & 60d10+620 (950 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +15 (+7 Dex, +8 Superior Initiative) \tabularnewline
\hline
S\textbf{peed:} & 30 ft. \tabularnewline
\hline
A\textbf{C:} & 55 (-2 size, +7 Dex, +40 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +60/+79\tabularnewline
\hline
A\textbf{ttack:} & Bite +72 (8d10+11/ 19-20 (+1d6 on critical hit)) melee\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +72 melee (8d10+11/ 19-20 (+1d6 on critical hit)), 
tail +67 (10d6+5) melee\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./10 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Prismatic emanations \tabularnewline
\hline
S\textbf{pecial Qualities:} & Immunities, prismatic blur, SR 38, DR 10/- \tabularnewline
\hline
S\textbf{aves:} & Fort +41, Ref +39, Will +24 \tabularnewline
\hline
A\textbf{bilities:} & Str 32, Dex 25, Con 29, Int 4, Wis 19, Cha 10 \tabularnewline
\hline
S\textbf{kills:} & Listen +37, Spot +38\tabularnewline
\hline
F\textbf{eats:} & Alertness, Cleave, Combat Reflexes, Dodge, Endurance, Great Cleave, 
Improved Critical (bite), Improved Initiative, Power Attack, Weapon Focus (bite), 
Weapon Focus (tail) \tabularnewline
\hline
E\textbf{pic Feats:} & Devastating Critical (bite), Epic Endurance, Epic Toughness 
(x4), Epic Weapon Focus (bite), Epic Weapon Focus (tail), Overwhelming Critical 
(bite), Superior Initiative \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any sunny land \tabularnewline
\hline
O\textbf{rganization:} & Solitary or pair \tabularnewline
\hline
C\textbf{hallenge Rating:} & 28 \tabularnewline
\hline
T\textbf{reasure:} & None \tabularnewline
\hline
A\textbf{lignment:} & Always neutral \tabularnewline
\hline
A\textbf{dvancement:} & 61-90 HD (Huge); 91-150 HD (Gargantuan) \tabularnewline
\hline
\end{tabular}

COMBAT 

\textbf{Prismatic Emanations (Su):} The reflected and refracted light from the 
prismasaurus's crystalline ridge functions very similarly to the \textit{prismatic 
spray }spell. Any creature of less than 8 HD that is within 30 feet of the creature 
is automatically blinded for 2d4 rounds. In addition, any creature within 20 feet 
of the prismasaurus is randomly struck by one or more rays of light, as determined 
by the table that accompanies the \textit{prismatic spray }spell\textit{. }This 
effect is as the \textit{prismatic spray }spell in a 30-foot-radius spread, but 
is otherwise as cast by a 20th-level caster. The save DC for the \textit{prismatic 
spray} effects is 49.

\textbf{Immunities (Su): }It is immune to poison, gases, petrification, and mind-affecting 
attacks. 

\textbf{Prismatic Blur (Ex):} The reflected and refracted light that constantly 
surrounds the creature blurs the creature's outline and makes it difficult to properly 
target. All melee and ranged attacks against the creature have a 50\% miss chance. 

\vspace{12pt}
{\LARGE{}PSEUDONATURAL TROLL }

\begin{tabular}{|>{\raggedright}p{65pt}|>{\raggedright}p{260pt}|}
\hline
  & Large Outsider (Extraplanar)\tabularnewline
\hline
H\textbf{it Dice:} & 6d8+66 (114 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +7 (Dex) \tabularnewline
\hline
S\textbf{peed:} & 60 ft. \tabularnewline
\hline
A\textbf{C:} & 51 (-1 size, +7 Dex, +35 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +4/+40\tabularnewline
\hline
A\textbf{ttack:} & Tentacle rake +35 (2d8+17) melee\tabularnewline
\hline
F\textbf{ull Attack:} & 5 tentacle rakes +35 (2d8+17) melee\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Constant insight, improved grab, rend 4d8+25, rotting 
constriction \tabularnewline
\hline
S\textbf{pecial Qualities:} & DR 5/epic, SR 30, acid and electricity resistance 
20, regeneration 5, scent, spell-like abilities, darkvision 90 ft. \tabularnewline
\hline
S\textbf{aves:} & Fort +16, Ref +9, Will +8 \tabularnewline
\hline
A\textbf{bilities:} & Str 45, Dex 24, Con 33, Int 6, Wis 19, Cha 6 \tabularnewline
\hline
S\textbf{kills:} & Climb +26, Hide +16, Jump +38, Listen +15, Move Silently +16, 
Spot +15\tabularnewline
\hline
F\textbf{eats:} & Alertness, Iron Will, Track \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary or gang (2-4) \tabularnewline
\hline
C\textbf{hallenge Rating:} & 21 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil \tabularnewline
\hline
A\textbf{dvancement:} & By character class \tabularnewline
\hline
\end{tabular}

Pseudotrolls speak Giant. 

A pseudotroll's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

COMBAT 

\textbf{Constant Insight (Su):} A pseudotroll makes all its attacks (including 
grapples) with a +15 insight bonus (calculated into the above statistics). The 
creature is not affected by the miss chance that applies to attacks against a concealed 
target. 

\textbf{Improved Grab (Ex): }If a pseudonatural troll hits an opponent smaller 
than it with a tentacle, it deals normal Damage and attempts to start a grapple 
as a free action without provoking an attack of opportunity. 

\textbf{Rend (Ex): }If a pseudotroll hits with two or more tentacle rake attacks, 
it latches onto the opponent's body and tears the flesh. This attack automatically 
deals an additional 4d8+25 points of Damage, and can be used concurrently with 
its improved grab ability. 

\textbf{Rotting Constriction (Ex): }Once a pseudotroll has hold of an opponent, 
each successful grapple check it makes during subsequent rounds automatically drains 
2d4 points of Constitution. At the same time, the pseudotroll regains 10 lost hit 
points. 

\textbf{Regeneration (Ex):} Fire and acid deal normal Damage to a pseudotroll. 
If a pseudotroll loses a tentacle or body part, the lost portion regrows in 1 minute. 
The creature can reattach the severed member instantly by holding it to the stump. 

\textbf{Spell-Like Abilities: }At will---\textit{blur, dimension door, shield, 
unhallow. }Caster level 20th; save DC 8 + spell level. The DC is Charisma-based.

\textbf{Alternate Form (Su):} At will, a pseudotroll can take the form of a grotesque, 
tentacled mass (or another appropriately gruesome form), but all its abilities 
remain unchanged despite the alien appearance. Changing shape is a standard action. 
Other creatures receive a -1 morale penalty on their attack rolls against a pseudotroll 
in this alternate form. 

\vspace{12pt}
\section*{{\LARGE{}CREATING A PSEUDONATURAL CREATURE }}

``Pseudonatural'' is a template that can be added to any corporeal creature (referred 
to hereafter as the base creature). 

\textbf{Size and Type: }The creature's type changes to outsider (extraplanar). 
Do not recalculate base attack bonus, saves, or skill points. Size is unchanged.

\textbf{HD: }A pseudonatural creature always has maximum hit points. 

\textbf{Speed:} A pseudonatural creature's speed doubles, for all movement types. 

\textbf{AC:} Pseudonatural creatures gain a +35 natural armor bonus. If the creature 
already has an armor bonus, use whichever is better. 

\textbf{Attack/Full Attack:} A pseudonatural creature substitutes tentacle rake 
attacks for its melee attacks while in pseudonatural form. These tentacle attacks 
always use the creature's best attack bonus.  It also gains additional tentacle 
rake attacks (also at its best attack bonus) as noted on the table below.

\textbf{Damage:} Tentacle rakes from a pseudonatural creature deal Damage equal 
to 2d8 + Str modifier. 

\textbf{Special Attacks:} A pseudonatural creature retains all the special attacks 
of the base creature and also gains the following. 

\textit{\textbf{Constant Insight (Su):}}\textit{ }The creature makes all its attacks 
with a +15 insight bonus. The creature is not affected by the miss chance that 
applies to attacks against a concealed target. 

\textit{\textbf{Improved Grab (Ex):}}\textit{ }If the creature hits an opponent 
smaller than it with a tentacle, it deals normal Damage and attempts to start a 
grapple as a free action without provoking an attack of opportunity. 

\textit{\textbf{Rotting Constriction (Ex):}}\textit{ }Once the creature has hold 
of an opponent, each successful grapple check it makes during subsequent rounds 
permanently drains 2d4 points of Constitution. At the same time, the creature regains 
10 lost hit points. 

\textbf{Spell-Like Abilities: }At will---\textit{blur, dimension door, shield, 
unhallow. }Caster level 20th.  The DCs are Charisma-based.

\textbf{Special Qualities:} A pseudonatural creature retains all the special qualities 
of the base creature and also gains the following.  

\parindent=3pt
Electricity and acid resistance of a varying amount related to its Hit Dice (see 
the table below). If the creature already possesses such resistance, use whichever 
is better.  

Damage reduction of a varying amount related to its Hit Dice (see the table below). 
If the creature already possesses Damage reduction, use whichever is better.  

\parindent=7pt
Spell resistance equal to the creature's HD x5. If the creature already possesses 
spell resistance, use whichever is higher. 

\parindent=3pt
One extra tentacle rake attack (the creature can use another of its many tentacles 
to attack at no penalty while in pseudonatural form) for each 4 HD the creature 
has.  

A pseudonatural creature's natural weapons are treated as epic for the purpose 
of overcoming damage reduction.

\vspace{12pt}
\parindent=0pt
\begin{tabular}{|>{\raggedright}p{51pt}|>{\raggedright}p{98pt}|>{\raggedright}p{61pt}|>{\raggedright}p{91pt}|}
\hline
\subsection*{H\textbf{it Dice}} & E\textbf{lectricity and Acid Resistance} & D\textbf{amage 
Reduction} & E\textbf{xtra Tentacle Rake Attacks}\tabularnewline
\hline
1-3 & 15 & 5/epic & 1\tabularnewline
\hline
4-7 & 20 & 5/epic & 2\tabularnewline
\hline
8-11 & 25 & 10/epic & 3\tabularnewline
\hline
12-15 & 30 & 10/epic & 4\tabularnewline
\hline
16-19 & 35 & 15/epic & 5\tabularnewline
\hline
21-24 & 40 & 15/epic & 6\tabularnewline
\hline
each 4 more HD & +5 & 15/epic & +1\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
If the creature already has one or more of these special qualities, use the better 
value. 

\textit{\textbf{Alternate Form (Su):}}\textit{ }At will, a pseudonatural creature 
can take the form of a grotesque, tentacled mass (or another appropriately gruesome 
form), but all its abilities remain unchanged despite the alien appearance. Changing 
shape is a standard action. Other creatures receive a -1 morale penalty on their 
attack rolls against pseudonatural creatures in this alternate form. 

\textbf{Saves:} Same as the base creature. 

\textbf{Abilities:} Increase from the base creature as follows: Str +22, Con +10, 
Dex +10, Wisdom +10, Intelligence at least 3. 

\textbf{Skills:} Same as the base creature. 

\textbf{Feats: }Same as the base creature. 

\textbf{Climate/Terrain:} Any land and underground.

\textbf{Organization:} Same as the base creature. 

\textbf{Challenge Rating:} Up to 6 HD, as base creature +16; 7 HD to 15 HD, as 
base creature +13; 16+ HD, as base creature +10 

\textbf{Treasure: }Same as the base creature. 

\textbf{Alignment:} Same as base creature. 

\textbf{Advancement:} Same as the base creature. 

\vspace{12pt}
{\LARGE{}RUIN SWARM}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{228pt}|}
\hline
  & Tiny Vermin (Swarm)\tabularnewline
\hline
H\textbf{it Dice:} & 50d8+500 (725 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +16\tabularnewline
\hline
S\textbf{peed:} & 20 ft., climb 20 ft., fly 90 ft. (perfect)\tabularnewline
\hline
A\textbf{rmor Class:} & 30 (+2 size, +16 Dex, +2 natural), touch 28, flatfooted 
14\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +37/---\tabularnewline
\hline
A\textbf{ttack:} & Swarm (5d6)\tabularnewline
\hline
F\textbf{ull Attack:} & Swarm (5d6)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./0 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Blinding, distraction (DC 45)\tabularnewline
\hline
S\textbf{pecial Qualities:} & Blindsense, fast healing 15, swarm traits, vermin 
traits\tabularnewline
\hline
S\textbf{aves:} & Fort +37, Ref +34, Will +22\tabularnewline
\hline
A\textbf{bilities:} & Str 3, Dex 42, Con 30, Int---, Wis 23, Cha 32\tabularnewline
\hline
S\textbf{kills:}--- & \tabularnewline
\hline
F\textbf{eats:}--- & \tabularnewline
\hline
E\textbf{nvironment:} & Any\tabularnewline
\hline
O\textbf{rganization:} & Solitary, cloud (2-4 swarms), or plague (5-8 swarms)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 23\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & None\tabularnewline
\hline
\end{tabular}

Combat

\textbf{Blinding (Ex): }Any living creature that begins its turn with a ruin swarm 
in its space must make a Fortitude save (DC 51) or be blinded for one round. The 
save DC is Dexterity-based.

\textbf{Blindsense (Ex): }The ruin swarm notices and locates creatures within 200 
ft. Opponents still have 100\% concealment against the swarm (but swarm attacks 
ignore concealment).

\textbf{Distraction (Ex): }Any living creature vulnerable to the swarm's damage 
that begins its turn with a swarm in its square is nauseated for 1 round; a DC 
45 Fortitude save negates the effect. Even with a successful save, spellcasting 
or concentrating on spells within the area of a swarm requires a Concentration 
check (DC 20 + spell level). Using skills requiring patience and concentration 
requires a Concentration check (DC 20). The save DC is Constitution-based.

\textbf{Swarm Traits: }A swarm has no clear front or back and no discernable anatomy, 
so it is not subject to critical hits or flanking. A swarm made up of Tiny creatures 
takes half damage from

slashing and piercing weapons.

Reducing a swarm to 0 hit points or fewer causes the swarm to break up, though 
damage taken until that point does not degrade its ability to attack or resist 
attack. Swarms are never staggered or reduced to a dying state by damage. Also, 
they cannot be tripped, grappled, or bull rushed, and they cannot grapple another.

A swarm is immune to any spell or effect that targets a specific number of creatures 
(including single-target spells such as \textit{disintegrate}), with the exception 
of mind-affecting effects if the swarm has an intelligence score and a hive mind. 
A swarm takes a -10 penalty on saving throws against spells or effects that affect 
an area, such as many evocation spells or grenadelike weapons. If the area effect 
attack does not allow a saving throw, the swarm takes double damage instead.

A swarm rendered unconscious by means of subdual damage becomes disorganized and 
dispersed, and does not re-form until its hit points exceed its subdual damage.

\textbf{Vermin Traits: }A ruin swarm is immune to all mind-affecting effects (charms, 
compulsions, phantasms, patterns, and morale effects). It also has darkvision (60-foot 
range).

\vspace{12pt}
{\LARGE{}SHADOW OF THE VOID }

\begin{tabular}{|>{\raggedright}p{66pt}|>{\raggedright}p{259pt}|}
\hline
  & Large Undead (Cold, Incorporeal)\tabularnewline
\hline
H\textbf{it Dice:} & 35d12 (227 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +17 (+9 Dex, +8 Superior Initiative) \tabularnewline
\hline
S\textbf{peed:} & Fly 60 ft. (perfect) \tabularnewline
\hline
A\textbf{C:} & 48 (-1 size, +9 Dex, +10 deflection, +20 insight) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +17/-\tabularnewline
\hline
A\textbf{ttack:} & Incorporeal touch +25 (2d6 plus blightfire) melee\tabularnewline
\hline
F\textbf{ull Attack:} & Incorporeal touch +25 (2d6 plus blightfire) melee\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Blightfire, create spawn, spell-like abilities\tabularnewline
\hline
S\textbf{pecial Qualities:} & Undead traits, incorporeal traits, turn resistance 
+6, cold subtype, cold aura, SR 36, DR 10/epic \tabularnewline
\hline
S\textbf{aves:} & Fort +13, Ref +20, Will +26 \tabularnewline
\hline
A\textbf{bilities:} & Str -, Dex 29, Con -, Int 17, Wis 25, Cha 31 \tabularnewline
\hline
S\textbf{kills:} & Diplomacy +48, Hide +47, Intimidate +48, Listen +45, Search 
+41, Sense Motive +45, Spot +45\tabularnewline
\hline
F\textbf{eats:} & Alertness, Blind-Fight, Combat Reflexes, Dodge, Expertise, Great 
Fortitude, Improved Initiative, Mobility, Spring Attack, Whirlwind Attack \tabularnewline
\hline
E\textbf{pic Feats:} & Blinding Speed, Improved Combat Reflexes, Spellcasting Harrier, 
Superior Initiative \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary or gang (2-5) \tabularnewline
\hline
C\textbf{hallenge Rating:} & 26 \tabularnewline
\hline
T\textbf{reasure:} & None \tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil \tabularnewline
\hline
A\textbf{dvancement:} & 36-42 HD (Large); 43-70 HD (Huge) \tabularnewline
\hline
\end{tabular}

A shadow of the void's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

Combat 

\textbf{Blightfire (Su):} Living creatures taking Damage from a shadow of the void's 
incorporeal touch attack find themselves ignited with blightfire; cold, black flames 
cascade over their bodies, and they must succeed at a Fort save (DC 37) or permanently 
lose 6 points of Constitution. The opponent must continue to save every round for 
the next 6 rounds (7 rounds total) to avoid being permanently drained of 6 more 
Con points. The creature heals 30 points of Damage whenever it drains 6 points 
of Constitution, gaining any excess as temporary hit points. These temporary hit 
points last a maximum of 1 hour. If the opponent is slain by blight-fire, only 
icy fragments of the victim remain, until they reform as a winterwight.  The save 
DC is Charisma-dased.

\textbf{Create Spawn (Su):} Any humanoid slain by a shadow of the void becomes 
a winterwight in 1d4 rounds. Winterwights are under the command of the shadow of 
the void that created them and remain enslaved until its death. They do not possess 
any of the abilities they had in life, though they do possess general knowledge 
of the slain creature. This knowledge is used for evil purposes, if possible. 

\textbf{Spell-Like Abilities: }3/day---\textit{deeper darkness, freezing sphere. 
}Caster level 24th; save DC 20 + spell level. The save DC is Charisma-based.

\textbf{Incorporeal Traits:} Can be harmed only by other incorporeal creatures, 
magic weapons, or magic, with a 50\% chance to ignore any Damage from a corporeal 
source. Can pass through solid objects at will, and own attacks pass through armor. 
Always moves silently. 

\textbf{Undead Traits:} Immune to poison, \textit{sleep, }paralysis, stunning, 
disease, death, effects, necromantic effects, mind-affecting effects, and any effect 
requiring a Fortitude save unless it also works on objects. Not subject to critical 
hits, subdual Damage, ability Damage, ability drain, or energy drain. Negative 
energy heals. Not at risk of death from massive Damage, but destroyed at 0 hit 
points or less. Darkvision 60 ft. Cannot be raised; resurrection works only if 
creature is willing. 

\textbf{Cold Subtype: }Immunity to cold. Vulnerability to fire, (takes half again 
as much (+50\%) damage as normal from fire, regardless of whether a saving throw 
is allowed, or if the save is a success or failure). 

\textbf{Cold Aura (Su):} A 10-foot-radius spread cold aura surrounds a shadow of 
the void. All creatures of the cold sub-type in the area (including the shadow 
of the void) are treated as if having turn resistance +6 (if undead) and fast healing 
10. Creatures subject to cold Damage take 2d10 points of cold Damage each round 
they remain within the cold aura. 

\vspace{12pt}
{\LARGE{}SHAPE OF FIRE }

\begin{tabular}{|>{\raggedright}p{66pt}|>{\raggedright}p{259pt}|}
\hline
  & Large Undead (Fire, Incorporeal) \tabularnewline
\hline
H\textbf{it Dice:} & 35d12 (227 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +17 (+9 Dex, +8 Superior Initiative) \tabularnewline
\hline
S\textbf{peed:} & Fly 60 ft. (perfect) \tabularnewline
\hline
A\textbf{C:} & 48 (-1 size, +9 Dex, +10 deflection, +20 insight) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +17/-\tabularnewline
\hline
A\textbf{ttack:} & Incorporeal touch +25 (2d6 plus blazefire) melee\tabularnewline
\hline
F\textbf{ull Attack:} & Incorporeal touch +25 (2d6 plus blazefire) melee\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Blazefire, create spawn, spell-like abilities \tabularnewline
\hline
S\textbf{pecial Qualities:} & Undead traits, incorporeal traits, turn resistance 
+6, fire subtype, heat aura, SR 36, DR 10/epic \tabularnewline
\hline
S\textbf{aves:} & Fort +13, Ref +20, Will +26 \tabularnewline
\hline
A\textbf{bilities:} & Str -, Dex 29, Con -, Int 17, Wis 25, Cha 31 \tabularnewline
\hline
S\textbf{kills:} & Diplomacy +48, Hide +47, Intimidate +48, Listen +45, Search 
+41, Sense Motive +45, Spot +45\tabularnewline
\hline
F\textbf{eats:} & Alertness, Blind-Fight, Combat Reflexes, Dodge, Expertise, Great 
Fortitude, Improved Initiative, Mobility, Spring Attack, Whirlwind Attack \tabularnewline
\hline
E\textbf{pic Feats:} & Blinding Speed, Improved Combat Reflexes, Spellcasting Harrier, 
Superior Initiative \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary or gang (2-5) \tabularnewline
\hline
C\textbf{hallenge Rating:} & 26 \tabularnewline
\hline
T\textbf{reasure:} & None \tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil \tabularnewline
\hline
A\textbf{dvancement:} & 36-42 HD (Large); 43-70 HD (Huge) \tabularnewline
\hline
\end{tabular}

A shape of fire's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

COMBAT 

\textbf{Blazefire (Su): }Living creatures taking Damage from a shape of fire's 
incorporeal touch attack find them-selves ignited with blazefire; white-hot, lambent 
flames cascade over their bodies, and they must succeed at a Fort save (DC 37) 
or permanently lose 10 hit points. The opponent must continue to save every round 
for the next 6 rounds (7 rounds total) to avoid being permanently drained of 10 
more hit points each round. The shape of fire heals the same amount of Damage whenever 
a creature is drained, gaining any excess hit points as temporary hit points. These 
temporary hit points last a maximum of 1 hour. If the opponent is slain by blazefire, 
only blackened ash remains of the victim. Hit points lost to the blazefire never 
heal naturally and cannot be magically restored---they are gone for good. The save 
DC is Charisma-based.

\textbf{Create Spawn (Su):} Any humanoid slain by a shape of fire becomes a lavawight 
in 1d4 rounds. Lavawights are under the command of the shape of fire that created 
them and remain enslaved until its death. They do not possess any of the abilities 
they had in life. 

\textbf{Spell-Like Abilities: }3/day---\textit{fire storm, incendiary cloud. }Caster 
level 24th; save DC 20 + spell level. The save DC is Charisma-based.

\textbf{Incorporeal Traits:} Can be harmed only by other incorporeal creatures, 
magic weapons, or magic, with a 50\% chance to ignore any Damage from a corporeal 
source. Can pass through solid objects at will, and own attacks pass through armor. 
Always moves silently. 

\textbf{Undead Traits:} Immune to poison, \textit{sleep, }paralysis, stunning, 
disease, death, effects, necromantic effects, mind-affecting effects, and any effect 
requiring a Fortitude save unless it also works on objects. Not subject to critical 
hits, subdual Damage, ability Damage, ability drain, or energy drain. Negative 
energy heals. Not at risk of death from massive Damage, but destroyed at 0 hit 
points or less. Darkvision 60 ft. Cannot be raised; resurrection works only if 
creature is willing. 

\textbf{Fire Subtype:} Immunity to fire. Vulnerability to cold (takes half again 
as much (+50\%) damage as normal from cold, regardless of whether a saving throw 
is allowed, or if the save is a success or failure).

\textbf{Heat Aura (Su):} A 10-foot-radius spread heat aura surrounds a shape of 
fire. All creatures of the fire subtype in the area (including the shape of fire) 
are treated as if having turn resistance +6 (if undead) and fast healing 10. Creatures 
subject to fire Damage take 2d10 points of fire Damage each round they remain within 
the heat aura. 

\vspace{12pt}
{\LARGE{}SIRRUSH }

\begin{tabular}{|>{\raggedright}p{64pt}|>{\raggedright}p{262pt}|}
\hline
  & Large Magical Beast \tabularnewline
\hline
H\textbf{it Dice:} & 40d10+680 (900 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +23 (+15 Dex, +8 Superior Initiative) \tabularnewline
\hline
S\textbf{peed:} & 90 ft. \tabularnewline
\hline
A\textbf{C:} & 44 (+15 Dex, -1 size, +20 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +40/+60\tabularnewline
\hline
A\textbf{ttack:} & Claw +56 (2d6+16/19-20 (+1d6 on critical hit)) melee\tabularnewline
\hline
F\textbf{ull Attack:} & 4 claws +56 (2d6+16/19-20 (+1d6 on critical hit)) melee, 
bite +54 (4d6+8) melee\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Pounce, stunning roar \tabularnewline
\hline
S\textbf{pecial Qualities:} & Blindsight 300 ft., DR 10/epic, darkvision 60 ft., 
fast healing 20, low-light vision, SR 39, scent; acid, cold, electricity, fire, 
and sonic resistance 10 \tabularnewline
\hline
S\textbf{aves:} & Fort +39, Ref +37, Will +27 \tabularnewline
\hline
A\textbf{bilities:} & Str 42, Dex 40, Con 44, Int 21, Wis 38, Cha 28 \tabularnewline
\hline
S\textbf{kills:} & Climb +59, Hide +58, Jump +83, Listen +57, Move Silently +58, 
Spot +57, Survival +57\tabularnewline
\hline
F\textbf{eats:} & Cleave, Great Cleave, Improved Critical (claw), Improved Initiative, 
Multiattack, Power Attack, Weapon Focus (bite), Weapon Focus (claw) \tabularnewline
\hline
E\textbf{pic Feats:} & Blinding Speed (x4), Overwhelming Critical (claw), Superior 
Initiative \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or pack (4-9 sirrushes and 1-2 three-headed 
sirrushes) \tabularnewline
\hline
C\textbf{hallenge Rating:} & 24 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic neutral \tabularnewline
\hline
A\textbf{dvancement:} & 41-50 HD (Large), 51-60 HD (Huge), 61+ HD (Gargantuan) 
\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
{\LARGE{}THREE-HEADED SIRRUSH }

\begin{tabular}{|>{\raggedright}p{66pt}|>{\raggedright}p{259pt}|}
\hline
  & Large Magical Beast \tabularnewline
\hline
H\textbf{it Dice:} & 45d10+855 (1,102 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +25 (+17 Dex, +8 Superior Initiative) \tabularnewline
\hline
S\textbf{peed:} & 120 ft. \tabularnewline
\hline
A\textbf{C:} & 50 (+17 Dex, -1 size, +24 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +45/+67\tabularnewline
\hline
A\textbf{ttack:} & Claw +63 (3d6+18/19-20 (+1d6 on critical hit)) melee\tabularnewline
\hline
F\textbf{ull Attack:} & 4 claws +63 melee, 3 bites +61 melee\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Pounce, stunning roar \tabularnewline
\hline
S\textbf{pecial Qualities:} & Blindsight 350 ft., DR 10/epic, darkvision 60 ft., 
fast healing 25, low-light vision, SR 42, scent; acid, cold, electricity, fire, 
and sonic resistance 15 \tabularnewline
\hline
S\textbf{aves:} & Fort +45, Ref +43, Will +33 \tabularnewline
\hline
A\textbf{bilities:} & Str 47, Dex 45, Con 49, Int 26, Wis 43, Cha 33 \tabularnewline
\hline
S\textbf{kills:} & Balance +65, Climb +66, Hide +65, Intimidate +59, Jump +102, 
Listen +64, Move Silently +65, Search +56, Spot +64, Survival +64\tabularnewline
\hline
F\textbf{eats:} & Cleave, Great Cleave, Improved Critical (claw), Improved Initiative, 
Multiattack, Power Attack, Weapon Focus (bite), Weapon Focus (claw) \tabularnewline
\hline
E\textbf{pic Feats:} & Blinding Speed (x6), Overwhelming Critical (claw), Superior 
Initiative \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or pack (1-2 three-headed sirrushes and 
4-9 sirrushes). \tabularnewline
\hline
C\textbf{hallenge Rating:} & 28 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic neutral \tabularnewline
\hline
A\textbf{dvancement:} & 46-55 HD (Large); 56-65 HD (Huge); 66+ HD (Gargantuan) 
\tabularnewline
\hline
\end{tabular}

Sirrushes speak Sylvan and Draconic. 

A sirrush's natural weapons are treated as epic for the purpose of overcoming damage 
reduction.

COMBAT 

\textbf{Stunning Roar (Ex):} Every 1d4 rounds, a sirrush can loose a sonic attack 
of such volume that it stuns all creatures in a 60-foot spread for 1d4 rounds if 
they fail a Fortitude saving throw (DC 47, or DC 51 for a three-headed sirrush). 
The DC is 

\textbf{Pounce (Ex):} If a sirrush charges or leaps upon a foe during its first 
round of combat, it can make a full attack even if it has already taken a move 
action. 

\textbf{Headloss Resistance (Ex):} A sirrush's thick bony plate protects its neck 
like a shield and provides resistance from effects that could normally behead it, 
such as a vorpal weapon. When a sirrush or three-headed sir-rush would otherwise 
lose its head, it instead makes a Fortitude saving throw (DC 10 + 1/2 HD of attacker 
+ attacker's Str modifier). On a failed save, the sirrush's head leaves its shoulders 
normally, but on a successful save, the sirrush is unaffected by the beheading. 
A sirrush dies when beheaded; a three-headed sirrush goes on fighting until all 
its heads are lost (the only penalty it incurs is one or two fewer bite attacks 
when it makes a full attack). 

\vspace{12pt}
{\LARGE{}TAYELLAH }

\begin{tabular}{|>{\raggedright}p{66pt}|>{\raggedright}p{259pt}|}
\hline
  & Gargantuan Magical Beast \tabularnewline
\hline
H\textbf{it Dice:} & 34d10+448 (635 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +27 (+19 Dex, +8 Superior Initiative) \tabularnewline
\hline
S\textbf{peed:} & 100 ft. \tabularnewline
\hline
A\textbf{C:} & 44 (+19 Dex, -4 size, +15 insight, +4 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +34/+57\tabularnewline
\hline
A\textbf{ttack:} & Claw +52 (2d8+11) melee\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +52 (2d8+11) melee, 3 bites +47 (2d6+5) melee, 
sting +39 (3d6+5 plus poison) melee\textbf{ }\tabularnewline
\hline
S\textbf{pace/Reach:} & 20 ft./15 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Pounce, improved grab, rake 4d6+15 \tabularnewline
\hline
S\textbf{pecial Qualities:} & SR 34, DR 10/epic \tabularnewline
\hline
S\textbf{aves:} & Fort +31, Ref +38, Will +22 \tabularnewline
\hline
A\textbf{bilities:} & Str 32, Dex 48, Con 34, Int 14, Wis 32, Cha 19 \tabularnewline
\hline
S\textbf{kills:} & Hide +44, Jump +39, Listen +48, Move Silently +56, Spot +48\tabularnewline
\hline
F\textbf{eats:} & Combat Reflexes, Improved Initiative, Iron Will, Track, Weapon 
Focus (bite), Weapon Focus (claw), Weapon Finesse\tabularnewline
\hline
E\textbf{pic Feats:} & Epic Toughness (x2), Epic Weapon Focus (bite), Epic Weapon 
Focus (claw), Superior Initiative,\tabularnewline
\hline
C\textbf{limate/Terrain:} & Any temperate or cold land \tabularnewline
\hline
O\textbf{rganization:} & Solitary or pair \tabularnewline
\hline
C\textbf{hallenge Rating:} & 24 \tabularnewline
\hline
T\textbf{reasure:} & None \tabularnewline
\hline
A\textbf{lignment:} & Always neutral \tabularnewline
\hline
A\textbf{dvancement:} & 35-68 HD (Gargantuan); 69-102 HD (Colossal) \tabularnewline
\hline
\end{tabular}

A tayellah's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

COMBAT 

\textbf{Poison (Ex):} Sting, Fort save (DC 39); initial and secondary Damage 2d10 
temporary Con. The save DC is Constitution-based.

\textbf{Pounce (Ex):} If a tayellah leaps upon a foe during the first round of 
combat, it can make a full attack even if it has already taken a move action. 

\textbf{Improved Grab (Ex):} To use this ability, the tayellah must hit with two 
of its claw attacks. If it gets a hold, it can rake. 

\textbf{Rake (Ex):} A tayellah that gets a hold can make two additional rake attacks 
(+44 melee) with legs it normally only uses for walking, dealing 4d6+5 points of 
Damage each. If the tayellah pounces on an opponent, it can also rake. 

\vspace{12pt}
{\LARGE{}THORCIASID }

\begin{tabular}{|>{\raggedright}p{64pt}|>{\raggedright}p{261pt}|}
\hline
  & Medium-Size Aberration \tabularnewline
\hline
H\textbf{it Dice:} & 29d8+348 (478 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +22 (+14 Dex, +8 Superior Initiative) \tabularnewline
\hline
S\textbf{peed:} & 70 ft. \tabularnewline
\hline
A\textbf{C:} & 42 (+14 Dex, +18 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +21/+27\tabularnewline
\hline
A\textbf{ttack:} & Forelimb +36 (0 plus ability drain) melee touch\tabularnewline
\hline
F\textbf{ull Attack:} & 2 forelimbs +36 (0 plus ability drain) melee touch, 2 antennae 
+30 (antenna 0 plus energy drain) melee touch\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Ability drain, energy drain, cocoon \tabularnewline
\hline
S\textbf{pecial Qualities:} & DR 10/epic, darkvision 240 ft., scent, SR 34, fire 
resistance 30 \tabularnewline
\hline
S\textbf{aves:} & Fort +21, Ref +25, Will +25 \tabularnewline
\hline
A\textbf{bilities:} & Str 22, Dex 38, Con 34, Int 25, Wis 29, Cha 31 \tabularnewline
\hline
S\textbf{kills:} & Bluff +42, Concentration +44, Diplomacy +38, Escape Artist +46, 
Hide +46, Intimidate +38, Jump +22, Listen +41, Move Silently +46, Spot +41, Survival 
+25\tabularnewline
\hline
F\textbf{eats:} & Dodge, Expertise, Improved Disarm, Improved Initiative, Mobility, 
Spring Attack, Weapon Finesse, Weapon Focus (forelimb) \tabularnewline
\hline
E\textbf{pic Feats:} & Blinding Speed, Superior Initiative \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or infestation (4-7) \tabularnewline
\hline
C\textbf{hallenge Rating:} & 22 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Usually neutral \tabularnewline
\hline
A\textbf{dvancement:} & 30-36 HD (Medium-size); 37-42 HD (Large) \tabularnewline
\hline
\end{tabular}

Thorciasids speak Common, Elven, Dwarven, and Giant. 

A thorciasid's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

COMBAT 

\textbf{Cocoon (Ex):} A thorciasid can eject a stream of noxious goo that hardens 
on contact, potentially fixing a subject in place and hindering its movement and 
actions. Using a standard action, the thorciasid makes a ranged touch attack against 
any foe it can see up to 60 feet away. If it succeeds at the ranged touch attack, 
the subject makes a grapple check against the cocoon. This is accomplished as if 
the thorciasid itself were making the grapple check with a competence bonus of 
+10, for a total check modifier of +37. On a failed check for the subject, the 
goo hardens into a partial cocoon, and the victim is treated as if grappled (even 
though the thorciasid is free to do as it desires). The subject can attempt to 
break the grapple of the cocoon (or use Escape Artist) each round, making an opposed 
grapple check against the cocoon (check modifier +37 for the cocoon). A thorciasid 
can reinforce a partial cocoon by spending a standard action secreting additional 
goo. Each action so spent provides an additional +5 to the cocoon's grapple check 
modifier. Visually, the victim becomes more and more covered and is finally com-pletely 
encased in a cocoon. Even fully encased victims can still breathe, and a thorciasid 
can drain a victim's life force at its leisure. 

\textbf{Ability Drain (Su):} On a successful melee touch attack with a forelimb, 
the thorciasid permanently drains 1d4+1 points of Strength, 1d4+1 points of Dexterity, 
and 1 point of Constitution from the victim. The thorciasid regains 20 lost hit 
points with each successful ability drain. 

\textbf{Energy Drain (Su): }On a successful melee touch attack with an antenna, 
the thorciasid gives the victim one negative level. Each negative level bestowed 
upon a victim gives the thorciasid +1 bonus to its Constitution that lasts for 
24 hours. The negative levels similarly last for 24 hours, at which time the victim 
must make a Fortitude saving throw (DC 34) to avoid level loss. The save DC is 
Charisma-based.

\vspace{12pt}
{\LARGE{}TITAN,ELDER }

\begin{tabular}{|>{\raggedright}p{66pt}|>{\raggedright}p{259pt}|}
\hline
  & Colossal Outsider (Extraplanar)\tabularnewline
\hline
H\textbf{it Dice:} & 70d8+700 (1,015 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +0 \tabularnewline
\hline
S\textbf{peed:} & 150 ft. \tabularnewline
\hline
A\textbf{C:} & 58 (-8 size, +32 natural, +24 insight) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +70/+104\tabularnewline
\hline
A\textbf{ttack:} & C\textit{olossal +5 warhammer }+87 (4d8+30/19-20 (+2d6 on critical 
hit)) melee; or \textit{Colossal +5 javelin }+70 (2d10+22/19-20) ranged\tabularnewline
\hline
F\textbf{ull Attack:} & C\textit{olossal +5 warhammer }+87/+82/+77/+72 (4d8+30/19-20 
(+2d6 on critical hit)) melee; or \textit{Colossal +5 javelin }+70/+65/+60/+55 
(2d10+22/19-20) ranged\tabularnewline
\hline
S\textbf{pace/Reach:} & 30 ft./30 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Spell-like abilities, spells \tabularnewline
\hline
S\textbf{pecial Qualities:} & DR 20/epic, SR 40 \tabularnewline
\hline
S\textbf{aves:} & Fort +47, Ref +37, Will +50 \tabularnewline
\hline
A\textbf{bilities:} & Str 45, Dex 10, Con 31, Int 33, Wis 37, Cha 26 \tabularnewline
\hline
S\textbf{kills:} & Concentration +83, Craft (any five) +45, Decipher Script +84, 
Diplomacy +45, Gather Information +87, Intimidate +56, Jump +90, Knowledge (all) 
+66, Listen +86, Sense Motive +86, Speak Language (any five), Spellcraft +90, Spot 
+86\tabularnewline
\hline
F\textbf{eats:} & Blind-Fight, Cleave, Expertise, Great Cleave, Improved Critical 
(warhammer), Improved Critical (javelin), Improved Sunder, Power Attack, Silent 
Spell, Still Spell, Weapon Focus (warhammer), Weapon Focus (javelin) \tabularnewline
\hline
E\textbf{pic Feats:} & Automatic Silent Spell (0- through 9th level spells), Automatic 
Still Spell (0- through 9th level spells), Devastating Critical (warhammer), Epic 
Spellcasting, Epic Weapon Focus (warhammer), Epic Weapon Focus (javelin), Overwhelming 
Critical (warhammer), Polyglot \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary \tabularnewline
\hline
C\textbf{hallenge Rating:} & 30 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Always neutral \tabularnewline
\hline
A\textbf{dvancement:} & 71-140 HD (Colossal) \tabularnewline
\hline
\end{tabular}

An elder titan's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

\section*{COMBAT }

\textbf{Spell-Like Abilities: }At will---\textit{alter self, analyze dweomer, commune 
with nature, cure critical wounds, eyebite, freedom of movement, fly, fog cloud, 
produce flame, protection from energy, read magic, remove curse, sanctuary, sending, 
speak with animals, soften earth and stone, speak with plants, summon monster VI, 
greater teleport, wind wall; }3/day---\textit{antilife shell, astral projection, 
contact other plane, greater dispel magic, greater scrying, invisibility purge, 
plane shift. }Caster level 29th; save DC 18 + spell level. The save DCs are Charisma-based.

\textbf{Spells:} An elder titan can use arcane spells as a 29th-level wizard or 
divine spells as a 29th-level cleric, from the cleric list and from the Knowledge 
and Magic domains. He will also possess the knowledge of at least three epic spells 
and the means by which to develop them, though he may not have done so himself. 

\vspace{12pt}
{\LARGE{}TREANT,ELDER }

\begin{tabular}{|>{\raggedright}p{65pt}|>{\raggedright}p{260pt}|}
\hline
  & Colossal Plant\tabularnewline
\hline
H\textbf{it Dice:} & 50d8+880 (1,105 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +3 (-1 Dex, +4 Improved Initiative) \tabularnewline
\hline
S\textbf{peed:} & 60 ft. \tabularnewline
\hline
A\textbf{C:} & 41 (-1 Dex, -8 size, +40 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +37/+72\tabularnewline
\hline
A\textbf{ttack:} & Slam +51 (10d6+19/19-20(+1d6 on critical hit)) melee\tabularnewline
\hline
F\textbf{ull Attack:} & 2 slams +51 (10d6+19/19-20 (+1d6 on a critical hit)) melee\tabularnewline
\hline
A\textbf{ttacks:} & 2 slams +49 melee \tabularnewline
\hline
D\textbf{amage:} & Slam 10d6+19/19-20 \tabularnewline
\hline
S\textbf{pace/Reach:} & 30 ft./30 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Animate trees, trample, triple Damage against objects, 
spell-like abilities \tabularnewline
\hline
S\textbf{pecial Qualities:} & Plant traits, SR 29, DR 10/- \tabularnewline
\hline
S\textbf{aves:} & Fort +45, Ref +17, Will +33 \tabularnewline
\hline
A\textbf{bilities:} & Str 48, Dex 8, Con 42, Int 19, Wis 33, Cha 35 \tabularnewline
\hline
S\textbf{kills:} & Diplomacy +61, Hide -12, Intimidate +62, Jump +31, Knowledge 
(any one) +57, Listen +63, Sense Motive +56, Spot +43, Survival +51\tabularnewline
\hline
F\textbf{eats:} & Alertness, Endurance, Great Fortitude, Improved Critical (slam), 
Improved Initiative, Iron Will, Lightning Reflexes, Power Attack, Weapon Focus 
(slam) \tabularnewline
\hline
E\textbf{pic Feats:} & Epic Endurance, Epic Reflexes, Epic Toughness (x4), Epic 
Weapon Focus (slam), Epic Will \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any forest \tabularnewline
\hline
O\textbf{rganization:} & Solitary or with grove of treants \tabularnewline
\hline
C\textbf{hallenge Rating:} & 25 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Always neutral good \tabularnewline
\hline
A\textbf{dvancement:} & 51-150 HD (Colossal) \tabularnewline
\hline
\end{tabular}

Elder treants speak the language of treants, plus Common and Sylvan. 

COMBAT 

\textit{\textbf{Animate Trees }}\textbf{(Sp):} An elder treant can animate trees 
within 500 feet at will, controlling up to six trees at a time. It takes a full 
round for a normal tree to uproot itself. Thereafter it moves and fights as a treant 
in all physical respects. Animated trees lose their ability to move if the elder 
treant who animated them is incapacitated or moves out of range. 

\textbf{Trample (Ex): }An elder treant can trample Gargantuan or smaller creatures 
for 8d20+40 points of Damage. Opponents who do not make attacks of opportunity 
against the elder treant can attempt a Reflex save (DC 54) to halve the Damage. 
The save DC is Strength-based.

\textbf{Triple Damage against Objects (Ex): }An elder treant that makes a full 
attack against an object or structure deals triple Damage. 

\textbf{Plant Traits:} Immune to poison, \textit{sleep, }paralysis, stunning, and 
polymorphing; not subject to critical hits or mind-affecting effects. 

\textbf{Skills: }Elder treants receive skills as though they were fey. *They have 
a +24 racial bonus on Hide checks made in forested areas. 

\textbf{Spell-Like Abilities: }At will---\textit{animal messenger, command plants, 
speak with plants, stone tell; }3/day---mass cure light wounds\textit{ }(plants 
only), \textit{plant growth, sunburst, wall of thorns; }1/day---\textit{shambler. 
}Caster level 23rd; save DC 22 + spell level. The save DCs are Charisma-based.

\vspace{12pt}
{\LARGE{}UMBRAL BLOT (BLACKBALL) }

\begin{tabular}{|>{\raggedright}p{65pt}|>{\raggedright}p{261pt}|}
\hline
  & Medium-Size Construct \tabularnewline
\hline
H\textbf{it Dice:} & 57d10 +120 (433 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +18 (+10 Dex, +8 Superior Initiative) \tabularnewline
\hline
S\textbf{peed:} & Fly 90 ft. (perfect) \tabularnewline
\hline
A\textbf{C:} & 40 (+10 Dex, +20 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +42/+42\tabularnewline
\hline
A\textbf{ttack:} & Disintegrating touch +42 (5d6 plus disintegrating touch (Fort 
DC 38)) melee touch\tabularnewline
\hline
F\textbf{ull Attack:} & Disintegrating touch +42 (5d6 plus disintegrating touch 
(Fort DC 38)) melee touch\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Disintegrating touch, vortex \tabularnewline
\hline
S\textbf{pecial Qualities:} & Blindsight 200 ft., construct traits, fast healing 
10, planar travel, SR 44; acid, cold, electricity, fire, and sonic resistance 30 
\tabularnewline
\hline
S\textbf{aves:} & Fort +19, Ref +29, Will +29 \tabularnewline
\hline
A\textbf{bilities:} & Str 10, Dex 30, Con -, Int 14, Wis 30, Cha 30 \tabularnewline
\hline
S\textbf{kills:} & Hide +62, Listen +52, Move Silently +62, Sense Motive +43, Spot 
+52, Survival +43\tabularnewline
\hline
F\textbf{eats:} & Alertness, Combat Reflexes, Improved Initiative, Skill Focus 
(Sense Motive), Skill Focus (Survival), Stealthy, Track\tabularnewline
\hline
E\textbf{pic Feats:} & Epic Prowess (x3), Epic Skill Focus (Hide), Epic Skill Focus 
(Listen), Epic Skill Focus (Sense Motive), Epic Skill Focus (Spot), Epic Toughness 
(x5), Superior Initiative\tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary \tabularnewline
\hline
C\textbf{hallenge Rating:} & 32 \tabularnewline
\hline
T\textbf{reasure:} & None \tabularnewline
\hline
A\textbf{lignment:} & Always neutral \tabularnewline
\hline
A\textbf{dvancement:} & 58-63 HD (Medium-size); 64-79 HD (Large); 80-95 HD (Huge); 
96-171 HD (Gargantuan) \tabularnewline
\hline
\end{tabular}

COMBAT 

\textbf{Disintegrating Touch (Ex):} Any material object that comes into contact 
with a blackball is immediately disintegrated unless it succeeds at a Fortitude 
save (DC 38). A character or object that has been disintegrated by an umbral blot 
disappears completely, leaving behind not even dust to mark its passing. Those 
who make a successful saving throw still take 5d6 points of Damage from the disintegrating 
touch. Likewise, weapons or objects that save take a like amount of Damage. (Remember, 
tended or held objects save with the same bonus as their owners.) 

\textbf{Vortex (Ex):} Normally a blackball insulates itself somehow from the air 
around it (otherwise it would perpetually be at the center of a howling wind-storm). 
If it chooses, instead of allowing the air to bend around it, the blackball can 
suspend this insulation, causing a sudden rush of wind to pour toward the blackball 
from all directions. This vortex sucks all the air from a 30-foot-by-30-foot-by-30-foot 
room in a single round, creating a sudden influx of air in its direction. All flying 
or floating creatures within 30 feet of the umbral blot who fail a Reflex save 
(DC 38) are swept along with the wind into contact with the blackball. Nonflying 
creatures within 30 feet who fail a Reflex save (DC 19) are pulled into contact 
with the umbral blot. Contact with an umbral blot could lead to disintegration, 
as noted above. 

\textbf{Spell Immunities (Ex):} In addition to the spells that an umbral blot is 
immune to because of its construct traits, it is immune to \textit{disintegration 
}spells and variations thereof. 

\textbf{Planar Travel (Ex):} A blackball can fold space at will, allowing it to 
use \textit{ethereal jaunt, dimension door, greater teleport, }or \textit{plane 
shift }at will as a standard action. 

\textbf{Construct Traits:} Immune to mind-affecting effects (charms, compulsions, 
phantasms, patterns, and morale effects), and to poison, \textit{sleep, }paralysis, 
stunning, disease, death effects, necromantic effects, and any effect that requires 
a Fortitude save unless it also works on objects. Cannot heal Damage (though regeneration 
and fast healing still apply, if present). Not subject to critical hits, subdual 
Damage, ability Damage, ability drain, or energy drain. Not at risk of death from 
massive Damage, but destroyed when reduced to 0 hit points or less; cannot be raised 
or resurrected. Darkvision 60 ft. 

\vspace{12pt}
{\LARGE{}UVUUDAUM }

\begin{tabular}{|>{\raggedright}p{66pt}|>{\raggedright}p{259pt}|}
\hline
  & Large Outsider (Evil, Extraplanar) \tabularnewline
\hline
H\textbf{it Dice:} & 38d8+646 (817 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +14 (Dex) \tabularnewline
\hline
S\textbf{peed:} & 80 ft., climb 40 ft. \tabularnewline
\hline
A\textbf{C:} & 52 (+14 Dex, -1 size, +29 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +38/+56\tabularnewline
\hline
A\textbf{ttack:} & Head spike +51 (10d6+21/19-20 plus Wisdom drain) melee\tabularnewline
\hline
F\textbf{ull Attack:} & Head spike +51 (10d6+21/19-20 plus Wisdom drain) melee\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Confusion aura, spell-like abilities, Wisdom drain 
\tabularnewline
\hline
S\textbf{pecial Qualities:} & Blindsight 500 ft., DR 10/epic and good, electricity 
resistance 30, fast healing 20, regeneration 5, SR 39 \tabularnewline
\hline
S\textbf{aves:} & Fort +40, Ref +37, Will +41 \tabularnewline
\hline
A\textbf{bilities:} & Str 39, Dex 38, Con 44, Int 32, Wis 38, Cha 46 \tabularnewline
\hline
S\textbf{kills:} & Balance +18, Climb +32, Concentration +58, Craft (alchemy) +52, 
Diplomacy +22, Escape Artist +55, Gather Information +22, Hide +51, Intimidate 
+59, Jump +69, Knowledge (arcana, history, local, religion, the planes) +52, Listen 
+55, Move Silently +55, Search +52, Sense Motive +55, Spellcraft +56, Spot +55, 
Swim +55, Tumble +59\tabularnewline
\hline
F\textbf{eats:} & Cleave, Combat Casting, Combat Reflexes, Great Cleave, Improved 
Critical (head spike), Lightning Reflexes, Power Attack \tabularnewline
\hline
E\textbf{pic Feats:} & Epic Reflexes, Improved Combat Reflexes, Spell Stowaway 
\textit{(time stop}), Tenacious Magic \textit{(haste}), Tenacious Magic \textit{(displacement}), 
Tenacious Magic \textit{(stoneskin}) \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or madness (3-6) \tabularnewline
\hline
C\textbf{hallenge Rating:} & 27 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Usually neutral evil \tabularnewline
\hline
A\textbf{dvancement:} & As character class \tabularnewline
\hline
\end{tabular}

Uvuudaums can speak telepathically to any creature within 500 feet that has a language. 

A uvuudaum's natural weapons are treated as epic and evil for the purpose of overcoming 
damage reduction.

COMBAT 

\textbf{Confusion Aura (Su):} This ability operates continuously, as the \textit{confusion 
}spell but with a 30-foot radius. Potential victims can make a Will save (DC 47) 
to resist confusion. Those affected by the aura can make a new save every 4 rounds 
to throw off the effect. Once it has successfully resisted, a subject remains immune 
to that uvuudaum's confusion aura for 24 hours. The save DC is Charisma-based.

\textbf{Wisdom Drain (Su):} This effect permanently reduces a living opponent's 
Wisdom score by 2d4 points when the creature hits with its head spike, or twice 
that amount on a critical hit. The creature regains 5 lost hit points (or 10 on 
a critical hit ) whenever it drains Wisdom, gaining any excess hit points as temporary 
hit points. Thesse temporary hit points last a maximum of 1 hour. An opponent can 
avoid the Wisdom drain with a Fortitude save (DC 47). The save DC is Charisma-based.

\textbf{Spell-Like Abilities: }At will---\textit{confusion, dimension door, greater 
dispel magic, displacement, haste, invisibility, lightning bolt, magic missile, 
mage armor, polymorph, shield, stoneskin, true strike, web; }3/day---\textit{chain 
lightning, cone of cold, disintegrate, dominate person, plane shift, prismatic 
spray, scrying, greater teleport, wall of force; }1/day---\textit{contingent resurrection, 
nailed to the sky, time duplicate }(epic spells), \textit{time stop. }Caster level 
27th; save DC 28 + spell level. The save DCs are Charisma-based.

\textbf{Regeneration (Ex):} Uvuudaums take normal damage from holy weapons, cold, 
and fire. 

\vspace{12pt}
{\LARGE{}VERMIURGE }

\begin{tabular}{|>{\raggedright}p{66pt}|>{\raggedright}p{259pt}|}
\hline
  & Large Aberration \tabularnewline
\hline
H\textbf{it Dice:} & 42d8+546 (735 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +11 (+7 Dex, +4 Improved Initiative) \tabularnewline
\hline
S\textbf{peed:} & 50 ft., fly 90 ft. (perfect) \tabularnewline
\hline
A\textbf{C:} & 40 (+3 Dex, -1 size, +24 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +31/+47\tabularnewline
\hline
A\textbf{ttack:} & Sting +43 (1d6+12/19-20 plus poison) melee\tabularnewline
\hline
F\textbf{ull Attack:} & Sting +43 (1d6+12/19-20 plus poison) melee, 4 pincers +41 
(2d8+6/19-20) melee, bite +40 (3d6+6 plus poison) melee\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Aura of doom, concealing aura, frightful presence, 
poison, spell-like abilities \tabularnewline
\hline
S\textbf{pecial Qualities:} & DR 10/epic, darkvision 60 ft., fast healing 10, immune 
to all mind-affecting effects, low-light vision, SR 34; scent \tabularnewline
\hline
S\textbf{aves:} & Fort +29, Ref +21, Will +38 \tabularnewline
\hline
A\textbf{bilities:} & Str 34, Dex 25, Con 36, Int 14, Wis 40, Cha 44 \tabularnewline
\hline
S\textbf{kills:} & Heal +24, Hide +33, Jump +45, Knowledge (nature) +34, Listen 
+47, Move Silently +33, Spot +47\tabularnewline
\hline
F\textbf{eats:} & Alertness, Cleave, Combat Reflexes, Dodge, Great Cleave, Great 
Fortitude, Improved Bull Rush, Improved Critical (pincers), Improved Critical (sting), 
Improved Initiative, Multiattack, Power Attack, Weapon Focus (pincers), Weapon 
Focus (sting) \tabularnewline
\hline
E\textbf{pic Feats:} & Blinding Speed\tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary \tabularnewline
\hline
C\textbf{hallenge Rating:} & 24 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Usually lawful neutral\tabularnewline
\hline
A\textbf{dvancement:} & 43-52 HD (Large); 53-70 HD (Huge) \tabularnewline
\hline
\end{tabular}

Vermiurges speak Terran. 

A vermiurge's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

COMBAT 

\textbf{Aura of Doom (Su):} A vermiurge is constantly affected as if by a \textit{crown 
of vermin }spell (epic spell) cast by a 42nd-level caster. As per the spell, a 
vermiurge may completely suppress the aura of insects as desired. A vermiurge's 
aura of doom constantly replenishes itself at a rate of 200 vermin per round (unlimited 
duration). When vermin die off either by being slain or by dealing Damage to another 
creature, the aura does not permanently collapse, even if all the vermin are slain 
or if all 1,000 vermin deal their Damage in 1 round. As soon as any vermin are 
replenished, they participate normally in the \textit{crown of vermin }effect. 

\textbf{Concealing Aura (Ex):} Its aura of doom constantly provides a vermiurge 
with nine-tenths concealment, so all attacks made against it have a 40\% miss chance. 

\textbf{Frightful Presence (Su):} This ability takes effect automatically when 
the vermiurge first engages its aura of doom. It affects only opponents with fewer 
Hit Dice or levels than the creature has. The affected creature must make a successful 
Will save (DC 48) or become shaken. Success indicates that the target is immune 
to that creature's frightful presence for one day. The save DC is Charisma-based.

\textbf{Poison (Ex):} Sting or bite, Fort save (DC 44); initial Damage 1d6 temporary 
Con, secondary Damage 2d6 temporary Con. The save DC is Constitution-based.

\vspace{12pt}
{\LARGE{}WINTERWIGHT }

\begin{tabular}{|>{\raggedright}p{66pt}|>{\raggedright}p{259pt}|}
\hline
  & Medium-Size Undead (Cold) \tabularnewline
\hline
H\textbf{it Dice:} & 32d12 (208 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +12 (+8 Dex, +4 Improved Initiative) \tabularnewline
\hline
S\textbf{peed:} & 60 ft. \tabularnewline
\hline
A\textbf{C:} & 46 (+8 Dex, +28 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +16/+37\tabularnewline
\hline
A\textbf{ttack:} & +Claw +40 (3d8+21/19-20 (+1d6 on critical hit) plus blight-fire) 
melee\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +40 (3d8+21/19-20 (+1d6 on critical hit) plus blight-fire) 
melee, skull butt +32 (2d6+10 plus blightfire) melee\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Rend 6d8+31, blightfire, spell-like abilities \tabularnewline
\hline
S\textbf{pecial Qualities:} & Undead traits, cold subtype, cold aura, SR 34, DR 
10/-\tabularnewline
\hline
S\textbf{kills:} & Jump +21, Intimidate +44, Knowledge (religion) +35, Listen +43, 
Spot +43\tabularnewline
\hline
F\textbf{eats:} & Cleave, Improved Critical (claw), Improved Initiative, Power 
Attack, Weapon Focus (claw) \tabularnewline
\hline
E\textbf{pic Feats:} & Overwhelming Critical (claw), Blinding Speed, Epic Fortitude, 
Epic Weapon Focus (claw) \tabularnewline
\hline
S\textbf{aves:} & Fort +14, Ref +18, Will +24 \tabularnewline
\hline
A\textbf{bilities:} & Str 52, Dex 27, Con ---, Int 10, Wis 22, Cha 28 \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary or pair \tabularnewline
\hline
C\textbf{hallenge Rating:} & 23 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil \tabularnewline
\hline
A\textbf{dvancement:} & 33-42 HD (Medium-size); 43-55 HD (Huge) \tabularnewline
\hline
\end{tabular}

Winterwights speak Common and Infernal. 

A winterwight's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

COMBAT 

\textbf{Rend (Ex):} If the winterwight hits with both of its claws, it latches 
onto the opponent's body and tears the flesh. This attack automatically deals an 
additional 6d8+31 points of Damage. 

\textbf{Blightfire (Su):} When a winterwight deals Damage to a living opponent, 
a night-black flame begins to burn on the opponent's body. If the opponent fails 
a Fortitude save (DC 35), it takes 4 points of permanent Constitution drain. The 
opponent must continue to save every round for the next 4 rounds (5 rounds total) 
to avoid being permanently drained of an additional 4 points of Constitution per 
round. The creature regains 10 lost hit points whenever it drains 4 points of Constitution, 
gaining any excess hit points as temporary hit points. These temporary hit points 
last a maximum of 1 hour. If the opponent is slain by blightfire, only icy fragments 
of the victim remain. The save DC is Charisma-based.

\textbf{Spell-Like Abilities: }At will---\textit{cone of cold, control weather, 
dimension door, spell immunity (fireball}), \textit{sleet storm }(a winterwight 
is immune to the effects of a \textit{sleet storm }spell), \textit{wall of ic}e. 
Caster level 23rd; save DC 19 + spell level. The save DCs are Charisma-based.

\textbf{Undead Traits:} Immune to poison, \textit{sleep, }paralysis, stunning, 
disease, death, effects, necromantic effects, mind-affecting effects, and any effect 
requiring a Fortitude save unless it also works on objects. Not subject to critical 
hits, subdual Damage, ability Damage, ability drain, or energy drain. Negative 
energy heals. Not at risk of death from massive Damage, but destroyed at 0 hit 
points or less. Darkvision 60 ft. Cannot be raised; resurrection works only if 
creature is willing. 

\textbf{Cold Subtype:} Immunity to cold. Vulnerability to fire (takes half again 
as much (+50\%) damage as normal from fire, regardless of whether a saving throw 
is allowed, or if the save is a success or failure). 

\textbf{Cold Aura (Su):} A 10-foot-radius spread cold aura surrounds a winterwight. 
All creatures of the cold sub-type in the field (including the winterwight) are 
treated as if having turn resistance +6 (if undead) and fast healing 10. Creatures 
subject to cold Damage take 2d10 points of cold Damage each round they remain within 
the cold aura. 

\vspace{12pt}
{\LARGE{}WORM THAT WALKS }

\begin{tabular}{|>{\raggedright}p{50pt}|>{\raggedright}p{275pt}|}
\hline
  & Medium-Size Aberration \tabularnewline
\hline
H\textbf{it Dice:} & HD 23d8 + 49 (152 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +8 (+4 Dex, +4 Improved Initiative) \tabularnewline
\hline
S\textbf{peed:} & 30 ft. \tabularnewline
\hline
A\textbf{C:} & 47 (+4 Dex, +8 \textit{bracers}, +3 \textit{ring}, +2 \textit{amulet}, 
+20 insight) \tabularnewline
\hline
A\textbf{ttacks:} & +\textit{4 staff }+15/+10 melee \tabularnewline
\hline
D\textbf{amage:} & +\textit{4 staff }1d6+4\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Spell-like abilities, engulf, frightful presence \tabularnewline
\hline
S\textbf{pecial Qualities:} & Blindsight 300 ft., SR 36, discorporate, immunities\tabularnewline
\hline
S\textbf{aves:} & Fort +11, Ref +14, Will +17 \tabularnewline
\hline
A\textbf{bilities:} & Str 10, Dex 14 (18 with \textit{glove}s), Con 13 (15 with 
\textit{Ioun ston}e), Int 20 (26 with \textit{headban}d), Wis 12, Cha 8\tabularnewline
\hline
S\textbf{kills:} & Concentration +22, Craft (alchemy) +28, Listen +24, Knowledge 
(arcana, local, nature, religion) +28, Listen +21, Move Silently +24, Spellcraft 
+28\tabularnewline
\hline
F\textbf{eats:} & Brew Potion, Combat Casting, Craft Staff, Craft Wondrous Item, 
Heighten Spell, Improved Initiative, Lightning Reflexes, Maximize Spell, Quicken 
Spell, Scribe Scroll, Spell Penetration, Toughness \tabularnewline
\hline
E\textbf{pic Feats:} & Improved Spell Capacity (10th), Improved Spell Capacity 
(11th) \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary \tabularnewline
\hline
C\textbf{hallenge Rating:} & 26 \tabularnewline
\hline
T\textbf{reasure:} & As character \tabularnewline
\hline
A\textbf{lignment:} & Any evil \tabularnewline
\hline
A\textbf{dvancement:} & As character \tabularnewline
\hline
\end{tabular}

A worm that walks rarely speaks, although it retains knowledge of all languages 
it knew in life, as well as any it has learned since.

COMBAT 

\textbf{Spells:} A worm that walks can cast any spells it could cast as a character. 
This sample worm was a 23rd-level wizard (spells per day: 4/12/6/6/6/5/5/5/5/4/1/1). 
The last spell slots are 10th-level and 11th-level slots, respectively, and are 
available for 0-9th level meta-magic spells. 

\textbf{Spell-Like Abilities: }1/day---\textit{animal growth }(vermin), \textit{animal 
messenger }(vermin), \textit{animal shapes }(vermin), \textit{animal trance }(vermin), 
\textit{colossal vermin }(as \textit{giant vermi}n, but it can increase the creature's 
size from Large to Gargantuan and from Huge to Colossal), \textit{creeping doom, 
giant vermin, summon swarm }(vermin), \textit{summon vermin }(as \textit{summon 
nature's ally, }except it summons 10 HD of vermin per level), and \textit{vermin 
plague }(as \textit{insect plagu}e). Caster level 20th; save DC 9 + spell level. 
The save DCs are Charisma-based.

\textbf{Engulf (Ex):} A worm that walks can choose to engulf an opponent who is 
no more than one size category larger than itself. The worm attempts to embrace 
its victim, and with a successful melee touch attack, the victim is immediately 
swallowed up and surrounded by a mass of vermin, taking 100 points of damage. A 
victim who spends a full-round action can break free of the embrace and move up 
to half its speed away from the worm if desired, but can do nothing else. Otherwise, 
each round a victim remains embraced, it takes another 100 points of Damage. Constructs 
are immune to this attack. 

\textbf{Spells:} The worm that walks can cast any spells it could cast when a living 
character. It also gains several related abilities.

\textbf{Frightful Presence (Su):} When a worm that walks engulfs a victim, witnesses 
must make a Will save (DC 10 + 1/2 the worm's HD + the worm's Cha modifier). Those 
who make the save are shaken. Those who fail by 5 or fewer points are frightened. 
Those who fail by 6 to10 points are panicked. Those who fail by 11 or more points 
are cowering. All these conditions last for 1d4 rounds. Those who have seen a worm 
use this attack before gain a +5 bonus on their saving throws. The victim of the 
worm's engulf attack has a -5 penalty on his or her saving throw. 

\textbf{Discorporate (Ex):} If gravely threatened, a worm can discorporate as a 
free action, simply falling into a pile of individual vermin that slither quickly 
away. So long as any of the component vermin survive, they can breed and create 
a new body to house the wizard's intelligence and personality. Discorporating is 
a dangerous tactic, because once separated, the vermin are treated no differently 
than other vermin. Also, the worm stands a good chance of losing all its equipment. 
However, discorporation almost assures that at least one maggot (if not dozens) 
will manage to crawl away, and so provide for the worm's continued existence. 

\textbf{Immunities (Ex): }A worm that walks has no discernable anatomy, so it is 
not subject to critical hits or flanking.

\textbf{Possessions: }Scrolls of \textit{fireball }(10th-level caster), greater 
\textit{teleport, acid fog, incendiary cloud, shapechange, wail of the banshee, 
clenched fist, horrid wilting, maze, summon monster I}X; \textit{potions of cure 
serious wounds, alter self, haste, fly. }Other magic gear: \textit{amulet of natural 
armor +2, bracers of armor +8, ring of protection +3, +4 staff}, \textit{rod of 
absorption, gloves of Dexterity +4, headband of intellect +6, pink Ioun stone, 
wings of flying.}

\vspace{12pt}
\section*{{\LARGE{}CREATING A WORM THAT WALKS}}

``Worm that walks'' is a template that can be added to any evil wizard or sorcerer. 
It uses all the original character's statistics, special abilities, and equipment, 
except as noted here. 

\textbf{Type: }The character's type changes to aberration (it is a creature composed 
of hundreds of discrete crawling worms). 

\textbf{Hit Dice:} Increase to d8. 

\textbf{AC:} The mass of worms that make up this creature, each looking out for 
danger, in sum provide a +20 insight bonus to AC. 

\textbf{Special Attack:} A worm that walks retains all the character's special 
attacks. It also gains one special attack, engulf. 

\textbf{Engulf (Ex)}: A worm that walks can choose to engulf an opponent who is 
no more than one size category larger than itself. The worm attempts to embrace 
its victim, and with a successful melee touch attack, the victim is immediately 
swallowed up and surrounded by a mass of vermin, taking 100 points of Damage as 
the biting vermin nibble away. A victim who spends a full-round action can break 
free of the embrace and move up to half its speed away from the worm if desired, 
but can do nothing else. Otherwise, each round a victim remains embraced, it takes 
another 100 points of Damage. Constructs are immune to this attack. 

\textbf{Spells:} A worm that walks can cast any spells it could cast as a living 
character. 

\textbf{Spell-Like Abilities: }1/day--- \textit{animal growth }(vermin), \textit{animal 
messenger }(vermin), \textit{animal shapes }(vermin), \textit{animal trance }(vermin), 
\textit{colossal vermin }(as \textit{giant vermi}n, but it can increase the creature's 
size from Large to Gargantuan and from Huge to Colossal), \textit{creeping doom, 
giant vermin, summon swarm }(vermin), \textit{summon vermin }(as \textit{summon 
nature's ally, }except it summons 10 HD of vermin per level), and \textit{vermin 
plague }(as \textit{insect plagu}e). Caster level 20th. 

\textbf{Special Qualities:} A worm that walks retains any special abilities it 
had in life and gains those mentioned below. 

\textit{\textbf{Blindsight (Ex):}}\textit{ }Worms that walk have blindsight 300 
ft. 

\textit{\textbf{Spell Resistance (Ex):}}\textit{ }A worm that walks has spell resistance 
equal to its Challenge Rating +10. 

\textit{\textbf{Frightful Presence (Su):}}\textit{ }When a worm that walks engulfs 
a victim, witnesses must make a Will save (DC 10 + 1/2 the worm's HD + the worm's 
Cha modifier). Those who make the save are shaken. Those who fail by 5 or fewer 
points are frightened. Those who fail by 6 to 10 points are panicked. Those who 
fail by 11 or more points are cowering. All these conditions last for 1d4 rounds. 
Those who have seen a worm use this attack before gain a +5 bonus on their saving 
throws. The victim of the worm's engulf attack has a -5 penalty on his or her saving 
throw. 

\textit{\textbf{Discorporate (Ex):}}\textit{ }If gravely threatened, a worm can 
discorporate as a free action, simply falling into a pile of individual vermin 
that slither quickly away. So long as any of the component vermin survive, they 
can breed and create a new body to house the wizard's intelligence and personality. 
Discorporating is a dangerous tactic, because once separated, the vermin are treated 
no differently than other vermin. Also, the worm stands a good chance of losing 
all its equipment. However, discorporation almost assures that at least one maggot 
(if not dozens) will manage to crawl away, and so provide for the worm's continued 
existence. 

\textbf{Immunities (Ex): }A worm that walks has no discernable anatomy, so it is 
not subject to critical hits or flanking.

\textbf{Saves:} Same as the character. 

\textbf{Abilities:} Same as the character. 

\textbf{Skills:} Same as the character, except that a worm that walks receives 
a +20 racial bonus on Hide, Intuit Direction, Listen, and Move Silently checks. 

\textbf{Feats: }Same as the character. 

\textbf{Epic Feats: }Same as the character. 

\textbf{Climate/Terrain:} Any. 

\textbf{Organization:} Solitary, occasionally with minions or master. 

\textbf{Challenge Rating:} Same as the character + 3. 

\textbf{Treasure: }Same as the character. 

\textbf{Alignment:} Any evil. 

\textbf{Advancement:} By character class. 

\vspace{12pt}
\section*{WORM THAT WALKS CHARACTERS }

By making eldritch preparations on a burial plot, a powerful spellcaster can improve 
the likelihood of a worm that walks emerging from the ground. The body is buried 
in an elaborate ritual that, the spellcaster hopes, will attract a critical mass 
of worms or maggots. The spellcaster performing the ritual must spend 10,000 gp 
in rare reagents for the ceremony. Furthermore, the ritual drains 2,000 XP from 
the spell-caster and requires the following spells: \textit{limited wish, polymorph 
any object, summon swarm }(heightened to 7th level),\textit{ }and \textit{sympathy. 
}Even if the ritual is performed correctly, there is only a chance that the deceased 
spellcaster will arise as a worm that walks. For each prepared but uncast arcane 
spell the deceased had at the moment of death (or unused spell slots if a spontaneous 
caster), there's a 1\% chance that a worm that walks will slither from the grave 
soil in 1d4 days. Some worms that walk arise spontaneously from ordinary burial 
plots, but such an event is exceedingly rare. And an epic spell can create worms 
that walk every time without fail. 

\vspace{12pt}
\section*{GATHERING OF MAGGOTS }

Conjuration (Healing) 

\textbf{Spellcraft DC: }49 

\textbf{Components:} V, S , D F 

\textbf{Casting Time:} 7 days 

\textbf{Target: }Dead creature touched 

\textbf{Duration:} Permanent 

\textbf{Saving Throw: }Fortitude negates (harmless) 

\textbf{Spell Resistance:} Yes (harmless) 

\textbf{To Develop: }441,000 gp; 9 days; 17,640 XP. Seeds: \textit{life }(DC 27), 
\textit{transform }(DC 21). Factors: change creature type to aberration (+5 DC); 
add worm that walks' spell-like, extraordinary, and supernatural abilities (+40 
DC). Mitigating factors: increase casting time by 10 minutes (-20 DC); increase 
casting time by 7 days (-14 DC). 

When this spell is cast, worms or maggots slither and crawl their way inexorably 
toward the corpse touched. Over the course of the next week, they consume the flesh 
of the corpse, and when the last bit of the corpse is devoured, the creature is 
returned to a semblance of life as a worm that walks.

\newpage

\end{document}
