%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{array}
\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

This material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

\subsubsection*{{\LARGE{}EPIC CLASS PROGRESSIONS}}

\vspace{12pt}
\subsubsection*{{\LARGE{}EPIC BARBARIAN }}

\textbf{Hit Die:} d12

\textbf{Skill Points at Each Additional Level:} 4 + Int modifier 

\textbf{Barbarian Rage:} The epic barbarian gains one use of rage per day every 
four levels after 20th. 

\textbf{Trap Sense (Ex):} The epic barbarian's bonus increases by +1 every three 
levels higher than 18th. 

\textbf{Damage Reduction (Ex):} The epic barbarian's damage reduction increases 
by 1 point every three levels higher than 19th. 

\textbf{Bonus Feats:} The epic barbarian gains a bonus feat (selected from the 
list of epic barbarian bonus feats) every four levels higher than 20th. 

\textit{\textbf{Epic Barbarian Bonus Feat List: }}Armor Skin, Chaotic Rage, Damage 
Reduction, Devastating Critical, Dire Charge, Epic Endurance, Epic Prowess, Epic 
Speed, Epic Toughness, Epic Weapon Focus, Fast Healing, Incite Rage, Legendary 
Climber, Legendary Leaper, Legendary Rider, Legendary Tracker, Legendary Wrestler, 
Mighty Rage, Overwhelming Critical, Ruinous Rage, Terrifying Rage, Thundering Rage. 

\textbf{Table: The Epic Barbarian }

\begin{tabular}{|>{\raggedright}p{56pt}|>{\raggedright}p{222pt}|}
\hline
B\textbf{arbarian Level} & \subsection*{S\textbf{pecial }}\tabularnewline
\hline
21st & Trap sense +7\tabularnewline
\hline
22nd & Damage reduction 6/-\tabularnewline
\hline
23rd--- & \tabularnewline
\hline
24th & Rage 7/day, Trap sense +8, bonus feat \tabularnewline
\hline
25th & Damage reduction 7/-\tabularnewline
\hline
26th--- & \tabularnewline
\hline
27th & Trap sense +9\tabularnewline
\hline
28th & Damage reduction 8/-, Rage 8/day, bonus feat \tabularnewline
\hline
29th--- & \tabularnewline
\hline
30th & Trap sense +10\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\section*{{\LARGE{}EPIC BARD}}

\textbf{Hit Die:} d6

\textbf{Skill Points at Each Additional Level:} 6 + Int modifier

\textbf{Spells:} The bard's caster level is equal to his or her class level. The 
bard's number of spells per day does not increase after 20th level. The bard does 
not learn additional spells unless he or she selects the Spell Knowledge feat.

\textbf{Bardic Music:} The bard gains no new bardic music effects from his or her 
Perform ranks. However, he or she may select epic feats that grant new bardic music 
effects

\textbf{Bardic Knowledge:} Add the bard's class level + Intelligence modifier to 
all bardic knowledge checks, as normal. 

\textbf{Bonus Feats:} The epic bard gains a bonus feat (selected from the list 
of epic bard bonus feats) every three levels after 20th. 

\textit{Epic Bard Bonus Feat List: }Augmented Alchemy, Deafening Song, Epic Inspiration, 
Epic Leadership, Epic Reputation, Epic Skill Focus, Group Inspiration, Hindering 
Song, Improved Combat Casting, Improved Metamagic, Improved Spell Capacity, Inspire 
Excellence, Lasting Inspiration, Master Staff, Master Wand, Music of the Gods, 
Permanent Emanation, Polyglot, Ranged Inspiration, Rapid Inspiration, Reactive 
Countersong, Spell Knowledge, Spell Stowaway, Spell Opportunity, Tenacious Magic. 

\textbf{Inspire Courage (Su):} The epic bard's bonus when this ability is used 
increases by +1 every six levels higher than 20\textsuperscript{th}.

\vspace{12pt}
\textbf{Table: The Epic Bard }

\begin{tabular}{|>{\raggedright}p{29pt}|>{\raggedright}p{158pt}|}
\hline
B\textbf{ard Level} & \section*{S\textbf{pecial }}\tabularnewline
\hline
21st--- &  \tabularnewline
\hline
22nd--- &  \tabularnewline
\hline
23rd & Bonus feat \tabularnewline
\hline
24th--- &  \tabularnewline
\hline
25th--- &  \tabularnewline
\hline
26th & Inspire Courage +5, Bonus feat \tabularnewline
\hline
27th--- &  \tabularnewline
\hline
28th--- &  \tabularnewline
\hline
29th & Bonus feat \tabularnewline
\hline
30th--- &  \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\section*{{\LARGE{}EPIC CLERIC }}

\textbf{Hit Die:} d8

\textbf{Skill Points at Each Additional Level:} 2 + Int modifier

\textbf{Spells:} The cleric's caster level is equal to his or her class level. 
The cleric's number of spells per day does not increase after 20th level

\textbf{Turn or Rebuke Undead:} Use the cleric's class level to determine the most 
powerful undead affected by a turn or rebuke check and the turning damage, just 
as normal. 

\textbf{Bonus Feats:} The epic cleric gains a bonus feat (selected from the list 
of epic cleric bonus feats) every three levels after 20th. 

\textit{Epic Cleric Bonus Feat List: }Armor Skin, Automatic Quicken Spell, Automatic 
Silent Spell, Automatic Still Spell, Bonus Domain, Enhance Spell, Epic Spell Focus, 
Epic Spell Penetration, Epic Spellcasting, Ignore Material Components, Improved 
Alignment-Based Casting, Improved Combat Casting, Improved Heighten Spell, Improved 
Metamagic, Improved Spell Capacity, Intensify Spell, Multispell, Negative Energy 
Burst, Permanent Emanation, Planar Turning, Positive Energy Aura, Spectral Strike, 
Spell Stowaway, Spell Opportunity, Spontaneous Domain Access, Spontaneous Spell, 
Tenacious Magic, Undead Mastery, Zone of Animation. 

\vspace{12pt}
\textbf{Table: The Epic Cleric }

\begin{tabular}{|>{\raggedright}p{38pt}|>{\raggedright}p{60pt}|}
\hline
\section*{C\textbf{leric Level}} & \section*{S\textbf{pecial }}\tabularnewline
\hline
21st--- &  \tabularnewline
\hline
22nd--- &  \tabularnewline
\hline
23rd & Bonus feat \tabularnewline
\hline
24th--- & \tabularnewline
\hline
25th--- & \tabularnewline
\hline
26th & Bonus feat \tabularnewline
\hline
27th--- &  \tabularnewline
\hline
28th--- &  \tabularnewline
\hline
29th & Bonus feat \tabularnewline
\hline
30th--- &  \tabularnewline
\hline
\end{tabular}

\vspace{24pt}
\subsubsection*{{\LARGE{}EPIC DRUID }}

\textbf{Hit Die:} d8

\textbf{Skill Points at Each Additional Level:} 4 + Int modifier

\textbf{Spells:} The druid's caster level is equal to his or her class level. The 
druid's number of spells per day does not increase after 20th level. 

\textbf{Animal Companion: }The Druid's animal companion continues to increase in 
power.  Every three levels higher than 20\textsuperscript{th} the animal companion 
gains 2 bonus hit dice, +1 strength and dexterity, and an additional trick. 

\textit{\textbf{Wild Shape:}}\textit{ }The epic druid can use this ability to take 
the form of an animal one additional time per day every four levels higher than 
18th.  The druid's ability to wild shape into an elemental does not improve.

\textbf{Bonus Feats:} The epic druid gains a bonus feat (selected from the list 
of epic druid bonus feats) every four levels after 20th. 

Epic Druid Bonus Feat List: Automatic Quicken Spell, Automatic Silent Spell, Automatic 
Still Spell, Colossal Wild Shape, Diminutive Wild Shape, Dragon Wild Shape, Energy 
Resistance, Enhance Spell, Epic Spell Focus, Epic Spell Penetration, Epic Spellcasting, 
Fast Healing, Fine Wild Shape, Gargantuan Wild Shape, Ignore Material Components, 
Improved Combat Casting, Improved Elemental Wild Shape, Improved Heighten Spell, 
Improved Metamagic, Improved Spell Capacity, Intensify Spell, Magical Beast Companion, 
Magical Beast Wild Shape, Multi-spell, Perfect Health, Permanent Emanation, Plant 
Wild Shape, Spell Stowaway, Spell Opportunity, Spontaneous Spell, Tenacious Magic, 
Vermin Wild Shape. 

\textbf{Table: The Epic Druid }

\begin{tabular}{|>{\raggedright}p{29pt}|>{\raggedright}p{86pt}|}
\hline
\section*{D\textbf{ruid Level}} & \section*{S\textbf{pecial }}\tabularnewline
\hline
21st--- &  \tabularnewline
\hline
22nd & W\textit{ild shape }7/day \tabularnewline
\hline
23rd--- &  \tabularnewline
\hline
24th & Bonus feat \tabularnewline
\hline
25th--- &  \tabularnewline
\hline
26th & W\textit{ild shape }8/day \tabularnewline
\hline
27th--- &  \tabularnewline
\hline
28th & Bonus feat \tabularnewline
\hline
29th--- &  \tabularnewline
\hline
30th & W\textit{ild shape }9/day\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\section*{{\LARGE{}EPIC FIGHTER }}

\textbf{Hit Die:} d10

\textbf{Skill Points at Each Additional Level:} 2 + Int modifier

\textbf{Bonus Feats:} The epic fighter gains a bonus feat (selected from the list 
of epic fighter bonus feats) every two levels after 20th. 

\textit{Epic Fighter Bonus Feat List: }Armor Skin, Combat Archery, Damage Reduction, 
Devastating Critical, Dire Charge, Distant Shot, Energy Resistance, Epic Endurance, 
Epic Leadership, Epic Prowess, Epic Toughness, Epic Weapon Focus, Epic Weapon Specialization, 
Exceptional Deflection, Improved Combat Reflexes, Improved Manyshot, Improved Stunning 
Fist, Improved Whirlwind Attack, Infinite Deflection, Instant Reload, Legendary 
Commander, Legendary Rider, Legendary Wrestler, Overwhelming Critical, Penetrate 
Damage Reduction, Perfect Two-Weapon Fighting, Reflect Arrows, Spellcasting Harrier, 
Storm of Throws, Superior Initiative, Swarm of Arrows, Two-Weapon Rend, Uncanny 
Accuracy. In addition to the feats on this list, the fighter may treat any feat 
designated as a fighter bonus feat, but not listed here, as being on his or her 
bonus feat list. 

\textbf{Table: The Epic Fighter }

\begin{tabular}{|>{\raggedright}p{38pt}|>{\raggedright}p{60pt}|}
\hline
F\textbf{ighter Level} & \section*{S\textbf{pecial }}\tabularnewline
\hline
21st--- &  \tabularnewline
\hline
22nd & Bonus feat \tabularnewline
\hline
23rd--- &  \tabularnewline
\hline
24th & Bonus feat \tabularnewline
\hline
25th--- &  \tabularnewline
\hline
26th & Bonus feat \tabularnewline
\hline
27th--- &  \tabularnewline
\hline
28th & Bonus feat \tabularnewline
\hline
29th--- &  \tabularnewline
\hline
30th & Bonus feat \tabularnewline
\hline
\end{tabular}

\vspace{24pt}
\section*{{\LARGE{}EPIC MONK }}

\textbf{Hit Die:} d8

\textbf{Skill Points at Each Additional Level:} 4 + Int modifier

\textbf{AC Bonus:} The monk's bonus to Armor Class when unarmored increases by 
+1 every five levels higher than 20th. 

\textbf{Unarmed Strike:} The damage for a monk's unarmed strike does not increase 
after 16th level. 

\textbf{Stunning Attack: }If the monk has the Stunning Fist feat use the monk's 
class level when determining the DC to resist this attack, as normal. 

\textbf{Unarmored Speed Bonus:} The epic monk's speed when wearing no armor increases 
by 10 feet every three levels higher than 18th. The unarmored speed of Small and 
dwarven monks increases by 5 feet instead of 10 feet.

\textit{\textbf{Ki }}\textbf{Strike:} The monk's \textit{ki }strike ability does 
not automatically increase with class level after 16th level. 

\textbf{Wholeness of Body (Su):} The epic monk can cure up to twice his or her 
class level in hit points each day, as normal. 

\textbf{Abundant Step (Su):} Use the monk's class level when determining the effective 
caster level of this ability, as normal. 

\textbf{Diamond Soul (Ex):} The epic monk's spell resistance is equal to his or 
her class level +10, as normal. 

\textbf{Quivering Palm (Su):} Use the monk's class level when determining the DC 
to resist this attack, as normal. 

\textbf{Empty Body (Su):} Use the monk's class level when determining the duration 
of this effect, as normal. 

\textbf{Bonus Feats:} The epic monk gains a bonus feat (selected from the list 
of epic monk bonus feats) every five levels higher than 20th. 

\textit{Epic Monk Bonus Feat List: }Armor Skin, Blinding Speed, Damage Reduction, 
Energy Resistance, Epic Prowess, Epic Speed, Epic Toughness, Exceptional Deflection, 
Fast Healing, Improved Combat Reflexes, Improved \textit{Ki }Strike, Improved Spell 
Resistance, Improved Stunning Fist, Infinite Deflection, Keen Strike, Legendary 
Climber, Legendary Wrestler, Reflect Arrows, Righteous Strike, Self-Concealment, 
Shattering Strike, Vorpal Strike. 

\textbf{Table: The Epic Monk }

\begin{tabular}{|>{\raggedright}p{29pt}|>{\raggedright}p{32pt}|>{\raggedright}p{78pt}|>{\raggedright}p{60pt}|}
\hline
\section*{M\textbf{onk Level}} & \section*{A\textbf{C Bonus}} & \section*{U\textbf{narmored 
Speed Bonus}} & \section*{S\textbf{pecial }}\tabularnewline
\hline
21st & +4 & +70 ft. & \tabularnewline
\hline
22nd & +4 & +70 ft. & \tabularnewline
\hline
23rd & +4 & +70 ft. & \tabularnewline
\hline
24th & +4 & +80 ft. & \tabularnewline
\hline
25th & +5 & +80 ft. & Bonus feat \tabularnewline
\hline
26th & +5 & +80 ft. & \tabularnewline
\hline
27th & +5 & +90 ft. & \tabularnewline
\hline
28th & +5 & +90 ft. & \tabularnewline
\hline
29th & +5 & +90 ft. & \tabularnewline
\hline
30th & +6 & +100ft. & Bonus feat \tabularnewline
\hline
\end{tabular}

\vspace{24pt}
\section*{{\LARGE{}EPIC PALADIN }}

\textbf{Hit Die:} d10

\textbf{Skill Points at Each Additional Level:} 2 + Int modifier

\textit{\textbf{Lay on Hands}}\textbf{ (Su)}\textit{\textbf{:}}\textit{ }Each day 
a paladin can cure a total number of hit points equal to his or her Charisma bonus 
(if any) times his or her class level, as normal

\textbf{Smite Evil (Su):} The epic paladin adds his or her class level to damage 
with any smite evil attack, as normal. He or she can smite one additional time 
per day for every five levels higher than 20\textsuperscript{th}.

\textbf{Turn Undead: }The paladin turns undead as a cleric of two levels lower, 
as normal. 

\textbf{Spells:} The paladin's caster level is equal to one-half his or her class 
level, as normal.  The paladin's number of spells per day does not increase after 
20th level. 

\textbf{Special Mount:} The epic paladin's special mount continues to increase 
in power. Every five levels after 20th the special mount gains +2 bonus Hit Dice, 
its natural armor increases by +2, its Strength adjustment increases by +1, and 
its Intelligence increases by +1. The mount's spell resistance equals the paladin's 
class level + 5. 

\textit{\textbf{Remove Disease}}\textbf{ (Sp)}\textit{\textbf{: }}The epic paladin 
can use \textit{remove disease }one additional time per week for every three levels 
higher than 18th. 

\textbf{Bonus Feats:} The epic paladin gains a bonus feat (selected from the list 
of epic paladin feats) every three levels higher than 20th. 

\textit{Epic Paladin Bonus Feat List: }Armor Skin, Devastating Critical, Epic Leadership, 
Epic Prowess, Epic Reputation, Epic Toughness, Epic Weapon Focus, Great Smiting, 
Holy Strike, Improved Aura of Courage, Improved Combat Casting, Improved Spell 
Capacity, Legendary Commander, Legendary Rider, Overwhelming Critical, Perfect 
Health, Permanent Emanation, Planar Turning, Positive Energy Aura, Spectral Strike, 
Spontaneous Spell, Widen Aura of Courage. 

\textbf{Table: The Epic Paladin }

\begin{tabular}{|>{\raggedright}p{38pt}|>{\raggedright}p{123pt}|}
\hline
P\textbf{aladin Level} & \section*{S\textbf{pecial }}\tabularnewline
\hline
21st & R\textit{emove disease }6/week \tabularnewline
\hline
22nd--- &  \tabularnewline
\hline
23rd & Bonus feat \tabularnewline
\hline
24th & R\textit{emove disease }7/week \tabularnewline
\hline
25th & Smite evil 6/day\tabularnewline
\hline
26th & Bonus feat\tabularnewline
\hline
27th & R\textit{emove disease }8/week \tabularnewline
\hline
28th--- &  \tabularnewline
\hline
29th & Bonus feat \tabularnewline
\hline
30th & R\textit{emove disease }9/week \tabularnewline
\hline
\end{tabular}

\vspace{24pt}
\section*{{\LARGE{}EPIC RANGER }}

\textbf{Hit Die:} d8

\textbf{Skill Points at Each Additional Level:} 6 + Int modifier

\textbf{Spells:} The ranger's caster level is equal to one-half his or her class 
level, as normal. The ranger's number of spells per day does not increase after 
20th level. 

\textbf{Favored Enemy (Ex):} The epic ranger gains one additional favored enemy, 
and his or her bonuses against one category of favored enemies goes up by +2, every 
five levels higher than 20th. 

\textbf{Animal companion: } The epic ranger's animal companion continues to increase 
in power.  The ranger's effective druid level is half his or her ranger level.

\textbf{Bonus Feats:} The epic ranger gains a bonus feat (selected from the list 
of epic ranger feats) every three levels after 20th. 

\textit{Epic Ranger Bonus Feat List: }Bane of Enemies, Blinding Speed, Combat Archery, 
Death of Enemies, Distant Shot, Epic Endurance, Epic Prowess, Epic Speed, Epic 
Toughness, Epic Weapon Focus, Fast Healing, Improved Combat Casting, Improved Favored 
Enemy, Improved Manyshot, Improved Spell Capacity, Legendary Climber, Legendary 
Rider, Legendary Tracker, Perfect Health, Perfect Two-Weapon Fighting, Permanent 
Emanation, Spontaneous Spell, Storm of Throws, Swarm of Arrows, Two-Weapon Rend, 
Uncanny Accuracy. 

\textbf{Table: The Epic Ranger }

\begin{tabular}{|>{\raggedright}p{38pt}|>{\raggedright}p{96pt}|}
\hline
R\textbf{anger Level} & S\textbf{pecial }\tabularnewline
\hline
21st--- &  \tabularnewline
\hline
22nd--- &  \tabularnewline
\hline
23rd & Bonus feat \tabularnewline
\hline
24th--- &  \tabularnewline
\hline
25th & 6th favored enemy \tabularnewline
\hline
26th & Bonus feat \tabularnewline
\hline
27th--- &  \tabularnewline
\hline
28th--- &  \tabularnewline
\hline
29th & Bonus feat \tabularnewline
\hline
30th & 7th favored enemy \tabularnewline
\hline
\end{tabular}

\vspace{24pt}
\section*{{\LARGE{}EPIC ROGUE }}

\textbf{Hit Die:} d6. 

\textbf{Skill Points at Each Additional Level:} 8 + Int modifier.

\textbf{Sneak Attack:} The epic rogue's sneak attack damage increases by +1d6 at 
every odd-numbered level. 

\textbf{Special Abilities: }The rogue does not gain additional rogue special abilities 
after 19th level, but can choose one of the rogue class special abilities\textit{ 
}(crippling strike, defensive roll, improved evasion, opportunist, skill mastery, 
or slippery mind) instead of a bonus feat. 

\textbf{Trap Sense (Ex):} The epic rogue's bonus increases by +1 every three levels 
higher than 18th.

\textbf{Bonus Feats:} The epic rogue gains a bonus feat (selected from the list 
of epic rogue feats) every four levels after 20th. 

\textit{Epic Rogue Bonus Feat List: }Blinding Speed, Combat Archery, Dexterous 
Fortitude, Dexterous Will, Epic Dodge, Epic Reputation, Epic Skill Focus, Epic 
Speed, Improved Combat Reflexes, Improved Sneak Attack, Legendary Climber, Lingering 
Damage, Self-Concealment, Sneak Attack of Opportunity, Spellcasting Harrier, Superior 
Initiative, Trap Sense, Uncanny Accuracy. The rogue may choose a special rogue 
ability\textit{ }instead of a bonus feat. 

\textbf{Table: The Epic Rogue }

\begin{tabular}{|>{\raggedright}p{31pt}|>{\raggedright}p{143pt}|}
\hline
R\textbf{ogue Level} & \section*{S\textbf{pecial }}\tabularnewline
\hline
21st & Sneak attack +11d6 , trap sense +7\tabularnewline
\hline
22nd--- &  \tabularnewline
\hline
23rd & Sneak attack +12d6 \tabularnewline
\hline
24th & Trap sense +8, bonus feat \tabularnewline
\hline
25th & Sneak attack +13d6 \tabularnewline
\hline
26th--- & \tabularnewline
\hline
27th & Sneak attack +14d6, trap sense +9\tabularnewline
\hline
28th & Bonus feat \tabularnewline
\hline
29th & Sneak attack +15d6\tabularnewline
\hline
30th & Trap sense +10\tabularnewline
\hline
\end{tabular}

\vspace{24pt}
\section*{{\LARGE{}EPIC SORCERER}}

\textbf{Hit Die:} d4. 

\textbf{Skill Points at Each Additional Level:} 2 + Int modifier. 

\textbf{Spells:} The sorcerer's caster level is equal to his or her class level. 
The sorcerer's number of spells per day does not increase after 20th level. The 
sorcerer does not learn additional spells unless he or she selects the Spell Knowledge 
feat. 

\textbf{Familiar:} The sorcerer's familiar continues to increase in power. Every 
two levels beyond 20th the familiar's natural armor and Intelligence each increase 
by +1. The familiar's spell resistance is equal to the master's class level + 5. 
At 21\textsuperscript{st} level and again every ten levels higher than 21\textsuperscript{st}, 
the familiar gains the benefit of the Familiar Spell epic feat for a spell of its 
master's choice.

\textbf{Bonus Feats:} The epic sorcerer gains a bonus feat (selected from the list 
of epic sorcerer bonus feats) every three levels after 20th. 

\textit{Epic Sorcerer Bonus Feat List: }Augmented Alchemy, Automatic Quicken Spell, 
Automatic Silent Spell, Automatic Still Spell, Energy Resistance, Enhance Spell, 
Epic Spell Focus, Epic Spell Penetration, Epic Spellcasting, Familiar Spell, Ignore 
Material Components, Improved Combat Casting, Improved Heighten Spell, Improved 
Metamagic, Improved Spell Capacity, Intensify Spell, Master Staff, Master Wand, 
Multispell, Permanent Emanation, Spell Knowledge, Spell Stowaway, Spell Opportunity. 

\textbf{Table: The Epic Sorcerer }

\begin{tabular}{|>{\raggedright}p{47pt}|>{\raggedright}p{60pt}|}
\hline
S\textbf{orcerer Level} & \section*{S\textbf{pecial }}\tabularnewline
\hline
21st--- &  \tabularnewline
\hline
22nd--- &  \tabularnewline
\hline
23rd & Bonus feat \tabularnewline
\hline
24th--- &  \tabularnewline
\hline
25th--- &  \tabularnewline
\hline
26th & Bonus feat \tabularnewline
\hline
27th--- &  \tabularnewline
\hline
28th--- &  \tabularnewline
\hline
29th & Bonus feat \tabularnewline
\hline
30th--- & \tabularnewline
\hline
\end{tabular}

\vspace{24pt}
\section*{{\LARGE{}EPIC WIZARD }}

\textbf{Hit Die:} d4. 

\textbf{Skill Points at Each Additional Level:} 2 + Int modifier. 

\textbf{Spells:} The wizard's caster level is equal to her class level. The wizard's 
number of spells per day does not increase after 20th level. Each time the wizard 
achieves a new level, he or she learns two new spells of any spell levels that 
he or she can cast (according to his or her new level). 

\textbf{Familiar:} The wizards's familiar continues to increase in power. Every 
two levels beyond 20th the familiar's natural armor and Intelligence each increase 
by +1. The familiar's spell resistance is equal to the master's class level + 5. 
At 21\textsuperscript{st} level and again every ten levels higher than 21\textsuperscript{st}, 
the familiar gains the benefit of the Familiar Spell epic feat for a spell of its 
master's choice.

\textbf{Bonus Feats:} The epic wizard gains a bonus feat (selected from the list 
of epic wizard feats) every three levels after 20th. 

\textit{Epic Wizard Bonus Feat List: }Augmented Alchemy, Automatic Quicken Spell, 
Automatic Silent Spell, Automatic Still Spell, Combat Casting, Craft Epic Magic 
Arms and Armor, Craft Epic Rod, Craft Epic Staff, Craft Epic Wondrous Item, Efficient 
Item Creation, Enhance Spell, Epic Spell Focus, Epic Spell Penetration, Epic Spellcasting, 
Familiar Spell, Forge Epic Ring, Ignore Material Components, Improved Combat Casting, 
Improved Heighten Spell, Improved Metamagic, Improved Spell Capacity, Intensify 
Spell, Multispell, Permanent Emanation, Scribe Epic Scroll, Spell Focus, Spell 
Knowledge, Spell Mastery, Spell Penetration Spell Stowaway, Spell Opportunity, 
Spontaneous Spell, Tenacious Magic. In addition to the feats on this list, the 
wizard may select any item creation feat or metamagic feat not listed here. 

\textbf{Table: The Epic Wizard }

\begin{tabular}{|>{\raggedright}p{38pt}|>{\raggedright}p{60pt}|}
\hline
W\textbf{izard Level} & \section*{S\textbf{pecial }}\tabularnewline
\hline
21st--- &  \tabularnewline
\hline
22nd--- &  \tabularnewline
\hline
23rd & Bonus feat \tabularnewline
\hline
24th--- &  \tabularnewline
\hline
25th--- &  \tabularnewline
\hline
26th & Bonus feat \tabularnewline
\hline
27th--- &  \tabularnewline
\hline
28th--- &  \tabularnewline
\hline
29th & Bonus feat \tabularnewline
\hline
30th--- &  \tabularnewline
\hline
\end{tabular}

\vspace{24pt}
\section*{FAMILIARS AT EPIC LEVEL }

Special abilities granted to a familiar continue to grow as the level of its master 
increases beyond 20th, as shown on Table: Epic Familiar Special Abilities. Even 
though the table shows advancement only to 42nd level, that is not the limit. The 
patterns in the table continue infinitely. Familiar special abilities gained at 
less than 20th level also continue to improve. 

\textit{Familiar Spell: }The familiar gains the benefit of the Familiar Spell epic 
feat for the spell its master chooses. 

\textbf{Table: Epic Familiar Special Abilities}

\begin{tabular}{|>{\raggedright}p{65pt}|>{\raggedright}p{42pt}|>{\raggedright}p{24pt}|>{\raggedright}p{78pt}|}
\hline
M\textbf{aster Class Level} & N\textbf{atural Armor} & I\textbf{nt} & S\textbf{pecial 
}\tabularnewline
\hline
21st-22nd & +11 & 16 & Familiar Spell \tabularnewline
\hline
23rd-24th & +12 & 17  & \tabularnewline
\hline
25th-26th & +13 & 18 & \tabularnewline
\hline
27th-28th & +14 & 19  & \tabularnewline
\hline
29th-30th & +15 & 20  & \tabularnewline
\hline
31st-32nd & +16 & 21 & Familiar Spell \tabularnewline
\hline
33rd-34th & +17 & 22  & \tabularnewline
\hline
35th-36th & +18 & 23  & \tabularnewline
\hline
37th-38th & +19 & 24  & \tabularnewline
\hline
39th-40th & +20 & 25  & \tabularnewline
\hline
41st-42nd & +21 & 26 & Familiar Spell \tabularnewline
\hline
\end{tabular}

\vspace{36pt}
\subsubsection*{{\LARGE{}EPIC PSIONIC CLASS PROGRESSIONS}}

Epic level psionic characters follow the same epic progression as non-psionics 
epic characters, except as noted below.• 

\parindent=3pt
For manifesters, manifester level continues to increase after 20th level. However, 
a manifester's power points and powers known at each level do not increase automatically 
after 20th level.• 

The powers of psicrystals that are based on the creator's level continue to increase 
as their owner gains levels.• 

\parindent=7pt
Bonus feats for epic psionic classes can be chosen from among any epic feats for 
which the character meets the prerequisites or any psionic, metapsionic, or psionic 
item creation feat, as normal.

\vspace{12pt}
\subsection*{NO EPIC BONUS POWER PROGRESSION}

\parindent=0pt
Epic psionic characters do not gain additional bonus power points above 20th. They 
may still gain more power points as their key ability score increases.

Psionic characters can take the Improved Manifestation feat, which grants them 
additional power points each time they take the feat. Epic manifesters can also 
take Epic Expanded Knowledge or Power Knowledge, which give them access to additional 
powers, and Epic Psionic Focus, which allows manifesters to use more than one metapsionic 
power each time they expend their psionic focus.

\vspace{12pt}
\section*{{\LARGE{}EPIC PSION}}

\textbf{Hit Die:} d4.

\textbf{Skill Points at Each Additional Level:} 2 + Int modifier.

\textbf{Powers: }The psion's manifester level is equal to his class level. The 
psion's base power points do not increase after 20th level. Each time the psion 
attains a new level, he learns two new powers of any level or levels that he can 
manifest.

\textbf{Psicrystal:} The epic psion's psicrystal, if he has one, continues to increase 
in power. At every odd-numbered level higher than 20th (21st, 23rd, and so on) 
the psicrystal's natural armor bonus and Intelligence each increase by 1. The psicrystal's 
power resistance is equal to the master's level + 5. At 21st level and again every 
ten levels higher than 21st, the psicrystal gains the benefit of the Psicrystal 
Power epic feat for a power of its master's choice.

\textbf{Bonus Feats:} The epic psion gains a bonus feat every three levels higher 
than 20th (23rd, 26th, and so on).

\textbf{Table: The Epic Psion }

\begin{tabular}{|>{\raggedright}p{38pt}|>{\raggedright}p{60pt}|}
\hline
\section*{P\textbf{sion Level}} & \section*{S\textbf{pecial }}\tabularnewline
\hline
21st--- &  \tabularnewline
\hline
22nd--- &  \tabularnewline
\hline
23rd & Bonus feat \tabularnewline
\hline
24th--- &  \tabularnewline
\hline
25th--- &  \tabularnewline
\hline
26th & Bonus feat \tabularnewline
\hline
27th--- &  \tabularnewline
\hline
28th--- &  \tabularnewline
\hline
29th & Bonus feat \tabularnewline
\hline
30th--- &  \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\section*{PSICRYSTALS AT EPIC LEVEL }

Special abilities granted to a psycrystal continue to grow as the level of its 
master increases beyond 20th, as shown on Table: Epic Psicrystal Special Abilities. 
Even though the table shows advancement only to 42nd level, that is not the limit. 
The patterns in the table continue infinitely. Psicrystal special abilities gained 
at less than 20th level also continue to improve. 

\textit{Psicrystal Power: }The psicrystal gains the benefit of the Psicrystal Power 
epic feat for the power its master chooses. 

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{47pt}|>{\raggedright}p{51pt}|>{\raggedright}p{33pt}|>{\raggedright}p{70pt}|}
\hline
\multicolumn{4}{|p{202pt}|}{T\textbf{able: Epic Psicrystal Special Abilities}}\tabularnewline
\hline
O\textbf{wner Level} & N\textbf{atural Armor Adj.} & I\textbf{nt Adj.} & \section*{S\textbf{pecial}}\tabularnewline
\hline
21st-22nd & +10 & +10 & Psicrystal Power\tabularnewline
\hline
23rd-24th & +11 & +11 & \tabularnewline
\hline
25th-26th & +12 & +12 & \tabularnewline
\hline
27th-28th & +13 & +13 & \tabularnewline
\hline
29th-30th & +14 & +14 & \tabularnewline
\hline
31st-32nd & +15 & +15 & Psicrystal Power\tabularnewline
\hline
33rd-34th & +16 & +16 & \tabularnewline
\hline
35th-36th & +17 & +17 & \tabularnewline
\hline
37th-38th & +18 & +18 & \tabularnewline
\hline
39th-40th & +19 & +19 & \tabularnewline
\hline
41st-42nd & +20 & +20 & Psicrystal Power\tabularnewline
\hline
\end{tabular}

\vspace{24pt}
\section*{{\LARGE{}EPIC PSYCHIC WARRIOR}}

\textbf{Hit Die:} d8. 

\textbf{Skill Points at Each Additional Level:} 2 + Int modifier. 

\textbf{Powers:} The psychic warrior's manifester level is equal to her class level. 
The psychic warrior's base power points do not increase after 20th level. The epic 
psychic warrior gains one new power of any level that she can manifest at every 
even-numbered level she attains higher than 20th (22nd, 24th, 26th, and so on).

\textbf{Bonus Feats:} The epic psychic warrior gains a bonus feat (selected from 
the list of epic psychic warrior bonus feats) at 21st level and every three levels 
thereafter. 

\textbf{Table: The Epic Psychic Warrior }

\begin{tabular}{|>{\raggedright}p{74pt}|>{\raggedright}p{60pt}|}
\hline
P\textbf{sychic Warrior Level} & S\textbf{pecial }\tabularnewline
\hline
21st & Bonus feat \tabularnewline
\hline
22nd--- &  \tabularnewline
\hline
23rd--- &  \tabularnewline
\hline
24th & Bonus feat \tabularnewline
\hline
25th--- &  \tabularnewline
\hline
26th--- &  \tabularnewline
\hline
27th & Bonus feat \tabularnewline
\hline
28th--- &  \tabularnewline
\hline
29th--- &  \tabularnewline
\hline
30th & Bonus feat\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\section*{{\LARGE{}EPIC SOULKNIFE}}

\textbf{Hit Die:} d10.

\textbf{Skill Points at Each Additional Level:} 4 + Int modifier.

\textbf{Mind Blade: }At 25th level and every five levels thereafter, the soulknife 
gains an additional +1 enhancement bonus on attack rolls and damage rolls (+6 at 
25th, +7 at 30th, and so on).

\textbf{Mind Blade Enhancement: }The epic soulknife's mind blade gains additional 
powers as the soulknife gains epic levels. At 22nd level and every four levels 
thereafter, the soulknife can improve the value of the weapon special abilities 
on his mind blade by 1 (+5 equivalent bonus at 22nd, +6 at 26th, and so on). As 
noted in the soulknife class description, reassigning qualities still takes 8 hours 
of concentration.

\textbf{Psychic Strike: }The epic soulknife's psychic strike progression continues 
past 20th level, continuing to grant an additional die (d8) of damage every four 
levels higher than 19th (+6d8 at 23rd, +7d8 at 27th, and so on).

\textbf{Bonus Feats: }The epic soulknife gains a bonus psionic feat every three 
levels higher than 20th (23rd, 26th, and so on).

\textbf{Table: The Epic Soulknife }

\begin{tabular}{|>{\raggedright}p{70pt}|>{\raggedright}p{164pt}|}
\hline
S\textbf{oulknife Level} & S\textbf{pecial }\tabularnewline
\hline
21st--- &  \tabularnewline
\hline
22nd & Mind blade enhancement +5\tabularnewline
\hline
23rd & Bonus feat\tabularnewline
\hline
24th--- &  \tabularnewline
\hline
25th & +\textit{6 mind blade }\tabularnewline
\hline
26th & Mind blade enhancement +5, bonus feat\tabularnewline
\hline
27th--- &  \tabularnewline
\hline
28th--- &  \tabularnewline
\hline
29th & Bonus feat\tabularnewline
\hline
30th & +\textit{7 mind blade }\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\section*{{\LARGE{}EPIC WILDER}}

\textbf{Hit Die:} d6.

\textbf{Skill Points at Each Additional Level:} 4 + Int modifier.

\textbf{Powers:} The epic wilder's manifester level is equal to her class level. 
The wilder's power points and powers known do not increase after 20th level (except 
through the use of epic feats).

\textbf{Wild Surge: }The epic wilder's wild surge continues to increase after 20th 
level. At 23rd level and every four levels higher than 23rd, the wild surge continues 
to gain an additional +1 boost to manifester level when used (+7 at 23rd, +8 at 
27th, and so on).

\textbf{Surging Euphoria:} The wilder's surging euphoria progression continues 
past 20th level. At 28th level and every eight levels higher than 28th, the morale 
bonus on her attack rolls, damage rolls, and saving throws increases by 1 (+4 at 
28th, +5 at 36th, and so on).

\textbf{Volatile Mind: }The epic wilder's volatile mind ability continues to increase. 
At 21st level and every four levels higher than 21st, the penalty assessed against 
telepathy powers manifested on the wilder is increased by 1 power point (5 points 
at 21st, 6 points at 25th, and so on).

\textbf{Table: The Epic Wilder }

\begin{tabular}{|>{\raggedright}p{38pt}|>{\raggedright}p{127pt}|}
\hline
\section*{W\textbf{ilder Level}} & \section*{S\textbf{pecial }}\tabularnewline
\hline
21st & Volatile mind (5 power points)\tabularnewline
\hline
22nd--- &  \tabularnewline
\hline
23rd &  Wild surge +7\tabularnewline
\hline
24th--- &  \tabularnewline
\hline
25th & Volatile mind (6 power points)\tabularnewline
\hline
26th--- & \tabularnewline
\hline
27th & Wild surge +8\tabularnewline
\hline
28th & Surging euphoria +4 \tabularnewline
\hline
29th & Volatile mind (7 power points)\tabularnewline
\hline
30th--- &  \tabularnewline
\hline
\end{tabular}

\newpage

\end{document}
