%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{array}
\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

This material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

{\LARGE{}MONSTER FEATS}

These feats apply to abilities most commonly found amongst monsters or are related 
to monsters.

\vspace{12pt}
ABILITY FOCUS  [GENERAL]

Choose one of the creature's special attacks.

\textbf{Prerequisite:} Special attack.

\textbf{Benefit:} Add +2 to the DC for all saving throws against the special attack 
on which the creature focuses.

\textbf{Special:} A creature can gain this feat multiple times. Its effects do 
not stack. Each time the creature takes the feat it applies to a different special 
attack.

\vspace{12pt}
AWESOME BLOW [GENERAL, FIGHTER]

\textbf{Prerequisites:} Str 25, Power Attack, Improved Bull Rush, size Large or 
larger.

\textbf{Benefit:} As a standard action, the creature may choose to subtract 4 from 
its melee attack roll and deliver an awesome blow. If the creature hits a corporeal 
opponent smaller than itself with an awesome blow, its opponent must succeed on 
a Reflex save (DC = damage dealt) or be knocked flying 10 feet in a direction of 
the attacking creature's choice and fall prone. The attacking creature can only 
push the opponent in a straight line, and the opponent can't move closer to the 
attacking creature than the square it started in. If an obstacle prevents the completion 
of the opponent's move, the opponent and the obstacle each take 1d6 points of damage, 
and the opponent stops in the space adjacent to the obstacle.

\vspace{12pt}
CRAFT CONSTRUCT [ITEM CREATION]

\textbf{Prerequisites:} Craft Magic Arms and Armor, Craft Wondrous Item.

\textbf{Benefit:} A creature with this feat can create any construct whose prerequisites 
it meets. Enchanting a construct takes one day for each 1,000 gp in its market 
price. To enchant a construct, a spellcaster must spend 1/25 the item's price in 
XP and use up raw materials costing half of this price (see individual construct 
monster entries for details).

A creature with this feat can repair constructs that have taken damage. In one 
day of work, the creature can repair up to 20 points of damage by expending 50 
gp per point of damage repaired.

A newly created construct has average hit points for its Hit Dice.

\vspace{12pt}
EMPOWER SPELL-LIKE ABILITY  [GENERAL]

\textbf{Prerequisite:} Spell-like ability at caster level 6th or higher.

\textbf{Benefit:} Choose one of the creature's spell-like abilities, subject to 
the restrictions below. The creature can use that ability as an empowered spell-like 
ability three times per day (or less, if the ability is normally usable only once 
or twice per day).

When a creature uses an empowered spell-like ability, all variable, numeric effects 
of the spell-like ability are increased by one half. Saving throws and opposed 
rolls are not affected. Spell-like abilities without random variables are not affected.

The creature can only select a spell-like ability duplicating a spell with a level 
less than or equal to half its caster level (round down) -2. For a summary, see 
the table in the description of the Quicken Spell-Like Ability feat. 

\textbf{Special:} This feat can be taken multiple times. Each time it is taken, 
the creature can apply it to a different one of its spell-like abilities.

\vspace{12pt}
FLYBY ATTACK  [GENERAL]

\textbf{Prerequisite:} Fly speed.

\textbf{Benefit:} When flying, the creature can take a move action (including a 
dive) and another standard action at any point during the move. The creature cannot 
take a second move action during a round when it makes a flyby attack.

\textbf{Normal: }Without this feat, the creature takes a standard action either 
before or after its move.

\vspace{12pt}
HOVER  [GENERAL]

\textbf{Prerequisite:} Fly speed.

\textbf{Benefit:} When flying, the creature can halt its forward motion and hover 
in place as a move action. It can then fly in any direction, including straight 
down or straight up, at half speed, regardless of its maneuverability.

If a creature begins its turn hovering, it can hover in place for the turn and 
take a full-round action. A hovering creature cannot make wing attacks, but it 
can attack with all other limbs and appendages it could use in a full attack. The 
creature can instead use a breath weapon or cast a spell instead of making physical 
attacks, if it could normally do so.

If a creature of Large size or larger hovers within 20 feet of the ground in an 
area with lots of loose debris, the draft from its wings creates a hemispherical 
cloud with a radius of 60 feet. The winds so generated can snuff torches, small 
campfires, exposed lanterns, and other small, open flames of non-magical origin. 
Clear vision within the cloud is limited to 10 feet. Creatures have concealment 
at 15 to 20 feet (20\% miss chance). At 25 feet or more, creatures have total concealment 
(50\% miss chance, and opponents cannot use sight to locate the creature).

Those caught in the cloud must succeed on a Concentration check (DC 10 + 1/2 creature's 
HD) to cast a spell.

\textbf{Normal: }Without this feat, a creature must keep moving while flying unless 
it has perfect maneuverability.

\vspace{12pt}
IMPROVED NATURAL ARMOR  [GENERAL]

\textbf{Prerequisites:} Natural armor, Con 13.

\textbf{Benefit:} The creature's natural armor bonus increases by 1.

\textbf{Special:} A creature can gain this feat multiple times. Each time the creature 
takes the feat its natural armor bonus increases by another point.

\vspace{12pt}
IMPROVED NATURAL ATTACK  [GENERAL]

\textbf{Prerequisite:} Natural weapon, base attack bonus +4.

\textbf{Benefit:} Choose one of the creature's natural attack forms. The damage 
for this natural weapon increases by one step, as if the creature's size had increased 
by one category: 1d2, 1d3, 1d4, 1d6, 1d8, 2d6, 3d6, 4d6, 6d6, 8d6, 12d6. 

A weapon or attack that deals 1d10 points of damage increases as follows: 1d10, 
2d8, 3d8, 4d8, 6d8, 8d8, 12d8.

\vspace{12pt}
MULTIATTACK  [GENERAL]

\textbf{Prerequisite:} Three or more natural attacks.

\textbf{Benefit:} The creature's secondary attacks with natural weapons take only 
a -2 penalty.

\textbf{Normal:} Without this feat, the creature's secondary attacks with natural 
weapons take a -5 penalty.

\vspace{12pt}
MULTIWEAPON FIGHTING  [GENERAL]

\textbf{Prerequisites:} Dex 13, three or more hands.

\textbf{Benefit:} Penalties for fighting with multiple weapons are reduced by 2 
with the primary hand and reduced by 6 with off hands.

\textbf{Normal:} A creature without this feat takes a -6 penalty on attacks made 
with its primary hand and a -10 penalty on attacks made with its off hands. (It 
has one primary hand, and all the others are off hands.) See Two-Weapon Fighting\textit{.}

\textbf{Special:} This feat replaces the Two-Weapon Fighting feat for creatures 
with more than two arms.

\vspace{12pt}
QUICKEN SPELL-LIKE ABILITY  [GENERAL]

\textbf{Prerequisite:} Spell-like ability at caster level 10th or higher.

\textbf{Benefit:} Choose one of the creature's spell-like abilities, subject to 
the restrictions described below. The creature can use that ability as a quickened 
spell-like ability three times per day (or less, if the ability is normally usable 
only once or twice per day).

Using a quickened spell-like ability is a free action that does not provoke an 
attack of opportunity. The creature can perform another action---including the 
use of another spell-like ability---in the same round that it uses a quickened 
spell-like ability. The creature may use only one quickened spell-like ability 
per round.

The creature can only select a spell-like ability duplicating a spell with a level 
less than or equal to half its caster level (round down) -4. For a summary, see 
the table below.

In addition, a spell-like ability that duplicates a spell with a casting time greater 
than 1 full round cannot be quickened.

\textbf{Normal: }Normally the use of a spell-like ability requires a standard action 
and provokes an attack of opportunity unless noted otherwise.

\textbf{Special:} This feat can be taken multiple times. Each time it is taken, 
the creature can apply it to a different one of its spell-like abilities.

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{50pt}|>{\raggedright}p{57pt}|>{\raggedright}p{60pt}|}
\hline
\multicolumn{3}{|p{167pt}|}{E\textbf{mpower and Quicken Spell-Like Ability}}\tabularnewline
\hline
S\textbf{pell Level} & C\textbf{aster Level to Empower} & C\textbf{aster Level 
to Quicken}\tabularnewline
\hline
0 & 4th & 8th\tabularnewline
\hline
1st & 6th & 10th\tabularnewline
\hline
2nd & 8th & 12th\tabularnewline
\hline
3rd & 10th & 14th\tabularnewline
\hline
4th & 12th & 16th\tabularnewline
\hline
5th & 14th & 18th\tabularnewline
\hline
6th & 16th & 20th\tabularnewline
\hline
7th & 18th--- & \tabularnewline
\hline
8th & 20th--- & \tabularnewline
\hline
9th--- & --- & \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
SNATCH  [GENERAL]

\textbf{Prerequisite:} Size Huge or larger.

\textbf{Benefits:} The creature can choose to start a grapple when it hits with 
a claw or bite attack, as though it had the improved grab special attack. If the 
creature gets a hold on a creature three or more sizes smaller, it squeezes each 
round for automatic bite or claw damage. A snatched opponent held in the creature's 
mouth is not allowed a Reflex save against the creature's breath weapon, if it 
has one.

The creature can drop a creature it has snatched as a free action or use a standard 
action to fling it aside. A flung creature travels 1d6 x 10 feet, and takes 1d6 
points of damage per 10 feet traveled. If the creature flings a snatched opponent 
while flying, the opponent takes this amount or falling damage, whichever is greater.

\vspace{12pt}
WINGOVER  [GENERAL]

\textbf{Prerequisite:} Fly speed.

\textbf{Benefits:} A flying creature with this feat can change direction quickly 
once each round as a free action. This feat allows it to turn up to 180 degrees 
regardless of its maneuverability, in addition to any other turns it is normally 
allowed. A creature cannot gain altitude during a round when it executes a wingover, 
but it can dive.

The change of direction consumes 10 feet of flying movement.

\newpage

\end{document}
