%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{array}
\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

his material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

{\LARGE{}MONSTERS (H-I)}

\vspace{12pt}
{\LARGE{}HAG}

Although different hags are unique in appearance and mannerism, they have many 
aspects in common. All take the form of crones whose bent shapes belie their fierce 
power and swiftness.

Hags speak Giant and Common.

COMBAT

Hags are tremendously strong. They are naturally resistant to spells and can cast 
magic of their own. Hags often gather to form coveys. A covey, usually containing 
one hag of each type, can use powers beyond those of the individual members.

ANNIS

\begin{tabular}{|>{\raggedright}p{88pt}|>{\raggedright}p{238pt}|}
\hline
  & Large Monstrous Humanoid\tabularnewline
\hline
H\textbf{it Dice:} & 7d8+14 (45 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 20 (-1 size, +1 Dex, +10 natural), touch 10, flat-footed 
19\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +7/+18\tabularnewline
\hline
A\textbf{ttack:} & Claw +13 melee (1d6+7)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +13 melee (1d6+7) and bite +8 melee (1d6+3)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Improved grab, rake 1d6+7, rend 2d6+10, spell-like 
abilities\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 2/bludgeoning, darkvision 60 ft., 
spell resistance 19\tabularnewline
\hline
S\textbf{aves:} & Fort +6, Ref +6, Will +6\tabularnewline
\hline
A\textbf{bilities:} & Str 25, Dex 12, Con 14, Int 13, Wis 13, Cha 10\tabularnewline
\hline
S\textbf{kills:} & Bluff +8, Diplomacy +2, Disguise +0 (+2 acting), Hide +5, Intimidate 
+2, Listen +10, Spot +10\tabularnewline
\hline
F\textbf{eats:} & Alertness, Blind-Fight, Great Fortitude\tabularnewline
\hline
E\textbf{nvironment:} & Cold marshes\tabularnewline
\hline
O\textbf{rganization:} & Solitary or covey (3 hags of any kind plus 1-8 ogres and 
1-4 evil giants)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 6\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & By character class\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

An annis commonly uses its \textit{disguise self }ability to take the form of an 
exceptionally tall human, a fair giant, or an ogre.

An annis stands some 8 feet tall and weighs about 325 pounds.

Combat

Though physically powerful, these hags do not favor simple assaults but try to 
divide and confuse their foes before combat. They love to pose as commoners or 
gentlefolk to lull their victims into a sense of false security before they attack.

\textbf{Improved Grab (Ex):} To use this ability, an annis must hit a Large or 
smaller opponent with a claw attack. It can then attempt to start a grapple as 
a free action without provoking an attack of opportunity.

\textbf{Rake (Ex):} Attack bonus +13 melee, damage 1d6+7. An annis can attack a 
grappled foe with both claws at no penalty.

\textbf{Rend (Ex): }An annis that hits with both claw attacks latches onto the 
opponent's body and tears the flesh. This attack automatically deals an extra 2d6+10 
points of damage.

\textbf{Spell-Like Abilities:} 3/day---\textit{disguise self}, \textit{fog cloud. 
}Caster level 8th.

\vspace{12pt}
GREEN HAG

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{204pt}|}
\hline
  & Medium Monstrous Humanoid\tabularnewline
\hline
H\textbf{it Dice:} & 9d8+9 (49 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares), swim 30 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 22 (+1 Dex, +11 natural), touch 11, flat-footed 21\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +9/+13\tabularnewline
\hline
A\textbf{ttack:} & Claw +13 melee (1d4+4)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +13 melee (1d4+4)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Spell-like abilities, weakness, mimicry\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 90 ft., spell resistance 18\tabularnewline
\hline
S\textbf{aves:} & Fort +6, Ref +7, Will +7\tabularnewline
\hline
A\textbf{bilities:} & Str 19, Dex 12, Con 12, Int 13, Wis 13, Cha 14\tabularnewline
\hline
S\textbf{kills:} & Concentration +7, Craft or Knowledge (any one) +7, Hide +9, 
Listen +11, Spot +11 Swim +12\tabularnewline
\hline
F\textbf{eats:} & Alertness, Blind-Fight, Combat Casting, Great Fortitude\tabularnewline
\hline
E\textbf{nvironment:} & Temperate marshes\tabularnewline
\hline
O\textbf{rganization:} & Solitary or covey (3 hags of any kind plus 1-8 ogres and 
1-4 evil giants) \tabularnewline
\hline
C\textbf{hallenge Rating:} & 5\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & By character class\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

Green hags are found in desolate swamps and dark forests.

A green hag is about the same height and weight as a female human.

Combat

Green hags prefer to attack from hiding, usually after distracting foes. They often 
use darkvision to their advantage by attacking during moonless nights.

\textbf{Spell-Like Abilities:} At will---\textit{dancing lights, disguise self, 
ghost sound }(DC 12), \textit{invisibility, pass without trace, tongues, water 
breathing. }Caster level 9th. The save DC is Charisma-based.

\textbf{Weakness (Su):} A green hag can weaken a foe by making a special touch 
attack. The opponent must succeed on a DC 16 Fortitude save or take 2d4 points 
of Strength damage. The save DC is Charisma-based.

\textbf{Mimicry (Ex):} A green hag can imitate the sounds of almost any animal 
found near its lair.

\textbf{Skills:} A green hag has a +8 racial bonus on any Swim check to perform 
some special action or avoid a hazard. It can always choose to take 10 on a Swim 
check, even if distracted or endangered. It can use the run action while swimming, 
provided it swims in a straight line.

\vspace{12pt}
SEA HAG

\begin{tabular}{|>{\raggedright}p{77pt}|>{\raggedright}p{248pt}|}
\hline
  & Medium Monstrous Humanoid (Aquatic)\tabularnewline
\hline
H\textbf{it Dice:} & 3d8+6 (19 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares), swim 40 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 14 (+1 Dex, +3 natural), touch 11, flat-footed 13\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +3/+7\tabularnewline
\hline
A\textbf{ttack:} & Claw +7 melee (1d4+4)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +7 melee (1d4+4)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Horrific appearance, evil eye\tabularnewline
\hline
S\textbf{pecial Qualities:} & Amphibious, spell resistance 14\tabularnewline
\hline
S\textbf{aves:} & Fort +2, Ref +4, Will +4\tabularnewline
\hline
A\textbf{bilities:} & Str 19, Dex 12, Con 12, Int 10, Wis 13, Cha 14\tabularnewline
\hline
S\textbf{kills:} & Craft or Knowledge (any one) +3, Hide +4, Listen +6, Spot +6, 
Swim +12\tabularnewline
\hline
F\textbf{eats:} & Alertness, Toughness\tabularnewline
\hline
E\textbf{nvironment:} & Temperate aquatic\tabularnewline
\hline
O\textbf{rganization:} & Solitary or covey (3 hags of any kind plus 1-8 ogres and 
1-4 evil giants)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 4\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & By character class\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

The sea hag is found in the water of seas or overgrown lakes.

A sea hag is about the same height and weight as a female human.

Combat

Sea hags are not subtle and prefer a direct approach to combat.

They usually remain in hiding until they can affect as many foes as possible with 
their horrific appearance.

\textbf{Horrific Appearance (Su):} The sight of a sea hag is so revolting that 
anyone (other than another hag) who sets eyes upon one must succeed on a DC 13 
Fortitude save or instantly be weakened, taking 2d6 points of Strength damage. 
This damage cannot reduce a victim's Strength score below 0, but anyone reduced 
to Strength 0 is helpless. Creatures that are affected by this power or that successfully 
save against it cannot be affected again by the same hag's horrific appearance 
for 24 hours. The save DC is Charisma-based.

\textbf{Evil Eye (Su): }Three times per day, a sea hag can cast its dire gaze upon 
any single creature within 30 feet. The target must succeed on a DC 13 Will save 
or be dazed for three days, although \textit{remove curse }or \textit{dispel evil 
}can restore sanity sooner. In addition, an affected creature must succeed on a 
DC 13 Fortitude save or die from fright. Creatures with immunity to fear effects 
are not affected by the sea hag's evil eye. The save DCs are Charisma-based.

\textbf{Amphibious (Ex):} Although sea hags are aquatic, they can survive indefinitely 
on land.

\textbf{Skills:} A sea hag has a +8 racial bonus on any Swim check to perform some 
special action or avoid a hazard. It can always choose to take 10 on a Swim check, 
even if distracted or endangered. It can use the run action while swimming, provided 
it swims in a straight line.

\vspace{12pt}
HAG COVEY

From time to time, a trio of hags gathers as a covey. Usually this foul triune 
includes one hag of each type, but this is not always the case.

Combat

Hags in a covey rely on deception and their boosted magical abilities in combat.

A covey of hags is 80\% likely to be guarded by 1d8 ogres and 1d4 evil giants who 
do their bidding. These minions are often disguised with a \textit{veil }spell 
to appear less threatening and sent forth as spies. Such minions often (60\%) carry 
magic stones known as \textit{hag eyes }(see below).

\textbf{Spell-Like Abilities:} 3/day---\textit{animate dead, bestow curse }(DC 
17)\textit{, control weather, dream, forcecage, mind blank, mirage arcana }(DC 
18)\textit{, polymorph, veil }(DC 19)\textit{, vision. }Caster level 9th. The save 
DCs are based on a Charisma score of 16. To use one of these abilities (which requires 
a full-round action), all three hags must be within 10 feet of one another, and 
all must participate.

Once per month, a covey that does not have a \textit{hag eye }can create one from 
a gem worth at least 20 gp (see below).

\vspace{12pt}
Hag Eye

A \textit{hag eye }is a magic gem created by a covey. It appears to be nothing 
more than a semiprecious stone, but a \textit{gem of seeing }or other such effect 
reveals it as a disembodied eye. Often, a \textit{hag eye }is worn as a ring, brooch, 
or other adornment. Any of the three hags who created the \textit{hag eye }can 
see through it whenever they wish, so long as it is on the same plane of existence 
as the hag. A \textit{hag eye }has hardness 5 and 10 hit points. Destroying a \textit{hag 
eye }deals 1d10 points of damage to each member of the covey and causes the one 
who sustained the greatest damage to be blinded for 24 hours. 

\vspace{12pt}
{\LARGE{}HALF-CELESTIAL}

No matter the form, half-celestials are always comely and delightful to the senses, 
having golden skin, sparkling eyes, angelic wings, or some other sign of their 
higher nature. 

CREATING A HALF-CELESTIAL

``Half-celestial'' is an inherited template that can be added to any living, corporeal 
creature with an Intelligence score of 4 or higher and nonevil alignment (referred 
to hereafter as the base creature).

A half-celestial uses all the base creature's statistics and special abilities 
except as noted here.

\textbf{Size and Type:} The creature's type changes to outsider. Do not recalculate 
the creature's Hit Dice, base attack bonus, or saves. Size is unchanged. Half-celestials 
are normally native outsiders.

\textbf{Speed:} A half-celestial has feathered wings and can fly at twice the base 
creature's base land speed (good maneuverability). If the base creature has a fly 
speed, use that instead.

\textbf{Armor Class:} Natural armor improves by +1 (this stacks with any natural 
armor bonus the base creature has).

\textbf{Special Attacks:} A half-celestial retains all the special attacks of the 
base creature and also gains the following special abilities. 

\textit{Daylight (Su): }Half-celestials can use a \textit{daylight }effect (as 
the spell) at will.

\textit{Smite Evil (Su): }Once per day a half-celestial can make a normal melee 
attack to deal extra damage equal to its HD (maximum of +20) against an evil foe.

\textbf{Spell-Like Abilities:} A half-celestial with an Intelligence or Wisdom 
score of 8 or higher has two or more spell-like abilities, depending on its Hit 
Dice, as indicated on the table below. The abilities are cumulative

Unless otherwise noted, an ability is usable once per day. Caster level equals 
the creature's HD, and the save DC is Charisma-based.

\begin{tabular}{|>{\raggedright}p{23pt}|>{\raggedright}p{157pt}|}
\hline
H\textbf{D} & A\textbf{bilities}\tabularnewline
\hline
1-2 & P\textit{rotection from evil }3/day, \textit{bless}\tabularnewline
\hline
3-4 & A\textit{id, detect evil}\tabularnewline
\hline
5-6 & C\textit{ure serious wounds, neutralize poison}\tabularnewline
\hline
7-8 & H\textit{oly smite, remove disease}\tabularnewline
\hline
9-10 & D\textit{ispel evil}\tabularnewline
\hline
11-12 & H\textit{oly word}\tabularnewline
\hline
13-14 & H\textit{oly aura }3/day, \textit{hallow}\tabularnewline
\hline
15-16 & M\textit{ass charm monster}\tabularnewline
\hline
17-18 & S\textit{ummon monster IX }(celestials only)\tabularnewline
\hline
19-20 & R\textit{esurrection}\tabularnewline
\hline
\end{tabular}

\textbf{Special Qualities:} A half-celestial has all the special qualities of the 
base creature, plus the following special qualities.---

Darkvision out to 60 feet.---

Immunity to disease.---

Resistance to acid 10, cold 10, and electricity 10.---

Damage reduction: 5/magic (if HD 11 or less) or 10/magic (if HD 12 or more).---

A half-celestial's natural weapons are treated as magic weapons for the purpose 
of overcoming damage reduction.---

Spell resistance equal to creature's HD + 10 (maximum 35).--- 

\parindent=3pt
+4 racial bonus on Fortitude saves against poison.

\parindent=0pt
\textbf{Abilities:} Increase from the base creature as follows: Str +4, Dex +2, 
Con +4, Int +2, Wis +4, Cha +4.

\textbf{Skills:} A half-celestial gains skill points as an outsider and has skill 
points equal to (8 + Int modifier) x$ $(HD +3). Do not include Hit Dice from class 
levels in this calculation---the half-celestial gains outsider skill points only 
for its racial Hit Dice, and gains the normal amount of skill points for its class 
levels. Treat skills from the base creature's list as class skills, and other skills 
as cross-class. 

\textbf{Challenge Rating:} HD 5 or less, as base creature +1; HD 6 to 10, as base 
creature +2; HD 11 or more, as base creature +3.

\textbf{Alignment:} Always good (any).

\textbf{Level Adjustment:} Same as base creature +4.

\vspace{12pt}
{\LARGE{}HALF-DRAGON}

Half-dragon creatures are always more formidable than others of their kind that 
do not have dragon blood, and their appearance betrays their nature---scales, elongated 
features, reptilian eyes, and exaggerated teeth and claws. Sometimes they have 
wings.

CREATING A HALF-DRAGON

``Half-dragon'' is an inherited template that can be added to any living, corporeal 
creature (referred to hereafter as the base creature).

A half-dragon uses all the base creature's statistics and special abilities except 
as noted here.

\textbf{Size and Type:} The creature's type changes to dragon. Size is unchanged. 
Do not recalculate base attack bonus or saves.

\textbf{Hit Dice:} Increase base creature's racial HD by one die size, to a maximum 
of d12. Do not increase class HD.

\textbf{Speed:} A half-dragon that is Large or larger has wings and can fly at 
twice its base land speed (maximum 120 ft.) with average maneuverability. A half-dragon 
that is Medium or smaller does not have wings.

\textbf{Armor Class:} Natural armor improves by +4. 

\textbf{Attack: }A half-dragon has two claw attacks and a bite attack, and the 
claws are the primary natural weapon. If the base creature can use weapons, the 
half-dragon retains this ability. A half-dragon fighting without weapons uses a 
claw when making an attack action. When it has a weapon, it usually uses the weapon 
instead.

\textbf{Full Attack:} A half-dragon fighting without weapons uses both claws and 
its bite when making a full attack. If armed with a weapon, it usually uses the 
weapon as its primary attack and its bite as a natural secondary attack. If it 
has a hand free, it uses a claw as an additional natural secondary attack.

\textbf{Damage:} Half-dragons have bite and claw attacks. If the base creature 
does not have these attack forms, use the damage values in the table below. Otherwise, 
use the values below or the base creature's damage values, whichever are greater.

\begin{tabular}{|>{\raggedright}p{45pt}|>{\raggedright}p{53pt}|>{\raggedright}p{58pt}|}
\hline
S\textbf{ize} & B\textbf{ite Damage} & C\textbf{law Damage}\tabularnewline
\hline
Fine & 1--- & \tabularnewline
\hline
Diminutive & 1d2 & 1\tabularnewline
\hline
Tiny & 1d3 & 1d2\tabularnewline
\hline
Small & 1d4 & 1d3\tabularnewline
\hline
Medium & 1d6 & 1d4\tabularnewline
\hline
Large & 1d8 & 1d6\tabularnewline
\hline
Huge & 2d6 & 1d8\tabularnewline
\hline
Gargantuan & 3d6 & 2d6\tabularnewline
\hline
Colossal & 4d6 & 3d6\tabularnewline
\hline
\end{tabular}

\textbf{Special Attacks:} A half-dragon retains all the special attacks of the 
base creature and gains a breath weapon based on the dragon variety (see the table 
below), usable once per day. A half-dragon's breath weapon deals 6d8 points of 
damage. A successful Reflex save (DC 10 + 1/2 half-dragon's racial HD + half-dragon's 
Con modifier) reduces damage by half.

\begin{tabular}{|>{\raggedright}p{70pt}|>{\raggedright}p{146pt}|}
\hline
D\textbf{ragon Variety} & B\textbf{reath Weapon}\tabularnewline
\hline
Black & 60-foot line of acid\tabularnewline
\hline
Blue & 60-foot ine of lightning\tabularnewline
\hline
Green & 30-foot cone of corrosive (acid) gas\tabularnewline
\hline
Red & 30-foot cone of fire\tabularnewline
\hline
White & 30-foot cone of cold\tabularnewline
\hline
Brass & 60-foot line of fire\tabularnewline
\hline
Bronze & 60-foot line of lightning\tabularnewline
\hline
Copper & 60-foot line of acid\tabularnewline
\hline
Gold & 30-foot cone of fire\tabularnewline
\hline
Silver & 30-foot cone of cold\tabularnewline
\hline
\end{tabular}

\textbf{Special Qualities:} A half-dragon has all the special qualities of the 
base creature, plus darkvision out to 60 feet and low-light vision. A half-dragon 
has immunity to \textit{sleep }and paralysis effects, and an additional immunity 
based on its dragon variety.

\begin{tabular}{|>{\raggedright}p{65pt}|>{\raggedright}p{41pt}|>{\raggedright}p{65pt}|>{\raggedright}p{41pt}|}
\hline
D\textbf{ragon Variety} & I\textbf{mmunity} & D\textbf{ragon Variety} & I\textbf{mmunity}\tabularnewline
\hline
Black & Acid & Brass & Fire\tabularnewline
\hline
Blue & Electricity & Bronze & Electricity\tabularnewline
\hline
Green & Acid & Copper & Acid\tabularnewline
\hline
Red & Fire & Gold & Fire\tabularnewline
\hline
White & Cold & Silver & Cold\tabularnewline
\hline
\end{tabular}

\textbf{Abilities:} Increase from the base creature as follows: Str +8, Con +2, 
Int +2, Cha +2.

\textbf{Skills:} A half-dragon gains skill points as a dragon and has skill points 
equal to (6 + Int modifier) x$ $(HD + 3). Do not include Hit Dice from class levels 
in this calculation---the half-dragon gains dragon skill points only for its racial 
Hit Dice, and gains the normal amount of skill points for its class levels. Treat 
skills from the base creature's list as class skills, and other skills as cross-class.

\textbf{Environment:} Same as either the base creature or the dragon variety.

\textbf{Challenge Rating:} Same as the base creature + 2 (minimum 3).

\textbf{Alignment:} Same as the dragon variety.

\textbf{Level Adjustment:} Same as base creature +3.

\vspace{12pt}
{\LARGE{}HALF-FIEND}

No matter its form, a half-fiend is always hideous to behold, having dark scales, 
horns, glowing red eyes, bat wings, a fetid odor, or some other obvious sign that 
it is tainted with evil.

CREATING A HALF-FIEND

``Half-fiend'' is an inherited template that can be added to any living, corporeal 
creature with an Intelligence score of 4 or more and nongood alignment (referred 
to hereafter as the base creature).

A half-fiend uses all the base creature's statistics and special abilities except 
as noted here.

\textbf{Size and Type: }The creature's type changes to outsider. Do not recalculate 
Hit Dice, base attack bonus, or saves. Size is unchanged. Half-fiends are normally 
native outsiders.

\textbf{Speed:} A half-fiend has bat wings. Unless the base creature has a better 
fly speed, the creature can fly at the base creature's base land speed (average 
maneuverability).

\textbf{Armor Class:} Natural armor improves by +1 (this stacks with any natural 
armor bonus the base creature has).

\textbf{Attack:} A half-fiend has two claw attacks and a bite attack, and the claws 
are the primary natural weapon. If the base creature can use weapons, the half-fiend 
retains this ability. A half-fiend fighting without weapons uses a claw when making 
an attack action. When it has a weapon, it usually uses the weapon instead.

\textbf{Full Attack:} A half-fiend fighting without weapons uses both claws and 
its bite when making a full attack. If armed with a weapon, it usually uses the 
weapon as its primary attack and its bite as a natural secondary attack. If it 
has a hand free, it uses a claw as an additional natural secondary attack.

\textbf{Damage: }Half-fiends have bite and claw attacks. If the base creature does 
not have these attack forms, use the damage values in the table below. Otherwise, 
use the values below or the base creature's damage values, whichever are greater.

\begin{tabular}{|>{\raggedright}p{45pt}|>{\raggedright}p{53pt}|>{\raggedright}p{58pt}|}
\hline
S\textbf{ize} & B\textbf{ite Damage} & C\textbf{law Damage}\tabularnewline
\hline
Fine & 1--- & \tabularnewline
\hline
Diminutive & 1d2 & 1\tabularnewline
\hline
Tiny & 1d3 & 1d2\tabularnewline
\hline
Small & 1d4 & 1d3\tabularnewline
\hline
Medium & 1d6 & 1d4\tabularnewline
\hline
Large & 1d8 & 1d6\tabularnewline
\hline
Huge & 2d6 & 1d8\tabularnewline
\hline
Gargantuan & 3d6 & 2d6\tabularnewline
\hline
Colossal & 4d6 & 3d6\tabularnewline
\hline
\end{tabular}

\textbf{Special Attacks: }A half-fiend retains all the special attacks of the base 
creature and gains the following special attack.

\textit{Smite Good (Su): }Once per day the creature can make a normal melee attack 
to deal extra damage equal to its HD (maximum of +20) against a good foe.

Spell-Like Abilities: A half-fiend with an Intelligence or Wisdom score of 8 or 
higher has spell-like abilities depending on its Hit Dice, as indicated on the 
table below. The abilities are cumulative\textit{. }Unless otherwise noted, an 
ability is usable once per day. Caster level equals the creature's HD, and the 
save DC is Charisma-based.

\begin{tabular}{|>{\raggedright}p{23pt}|>{\raggedright}p{133pt}|}
\hline
H\textbf{D} & A\textbf{bilities}\tabularnewline
\hline
1-2 & D\textit{arkness }3/day\tabularnewline
\hline
3-4 & D\textit{esecrate}\tabularnewline
\hline
5-6 & U\textit{nholy blight}\tabularnewline
\hline
7-8 & P\textit{oison }3/day\tabularnewline
\hline
9-10 & C\textit{ontagion}\tabularnewline
\hline
11-12 & B\textit{lasphemy}\tabularnewline
\hline
13-14 & U\textit{nholy aura }3/day, \textit{unhallow}\tabularnewline
\hline
15-16 & H\textit{orrid wilting}\tabularnewline
\hline
17-18 & S\textit{ummon monster IX }(fiends only)\tabularnewline
\hline
19-20 & D\textit{estruction}\tabularnewline
\hline
\end{tabular}

\textbf{Special Qualities: }A half-fiend has all the special qualities of the base 
creature, plus the following special qualities.---

Darkvision out to 60 feet.---

Immunity to poison.---

Resistance to acid 10, cold 10, electricity 10, and fire 10.---

Damage reduction: 5/magic (if HD 11 or less) or 10/magic (if HD 12 or more).---

A half-fiend's natural weapons are treated as magic weapons for the purpose of 
overcoming damage reduction.---

Spell resistance equal to creature's HD + 10 (maximum 35).

\textbf{Abilities:} Increase from the base creature as follows: Str +4, Dex +4, 
Con +2, Int +4, Cha +2.

\textbf{Skills:} A half-fiend gains skill points as an outsider and has skill points 
equal to (8 + Int modifier) x $ $(HD + 3). Do not include Hit Dice from class levels 
in this calculation---the half-fiend gains outsider skill points only for its racial 
Hit Dice, and gains the normal amount of skill points for its class levels. Treat 
skills from the base creature's list as class skills, and other skills as cross-class.

\textbf{Challenge Rating:} HD 4 or less, as base creature +1; HD 5 to 10, as base 
creature +2; HD 11 or more, as base creature +3.

\textbf{Alignment:} Always evil (any).

\textbf{Level Adjustment:} +4.

\vspace{12pt}
{\LARGE{}HALFLING}

\begin{tabular}{|>{\raggedright}p{72pt}|>{\raggedright}p{253pt}|}
\hline
  & H\textbf{alfling, 1st-Level Warrior}\tabularnewline
\hline
  & Small Humanoid (Halfling)\tabularnewline
\hline
H\textbf{it Dice:} & 1d8+1 (5 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 16 (+1 size, +1 Dex, +3 studded leather, +1 light shield), 
touch 12, flat-footed 15\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +1/-3\tabularnewline
\hline
A\textbf{ttack:} & Longsword +3 melee (1d6/19-20) or light crossbow +3 ranged (1d6/19-20)\tabularnewline
\hline
F\textbf{ull Attack:} & Longsword +3 melee (1d6/19-20) or light crossbow +3 ranged 
(1d6/19-20)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Halfling traits\tabularnewline
\hline
S\textbf{pecial Qualities:} & Halfling traits\tabularnewline
\hline
S\textbf{aves:} & Fort +4, Ref +2, Will +0\tabularnewline
\hline
A\textbf{bilities:} & Str 11, Dex 13, Con 12, Int 10, Wis 9, Cha 8\tabularnewline
\hline
S\textbf{kills:} & Climb +2, Hide +4, Jump -4, Listen +3, Move Silently +1\tabularnewline
\hline
F\textbf{eats:} & Weapon Focus (longsword)\tabularnewline
\hline
E\textbf{nvironment:} & Warm plains\linebreak{}
(Deep halfling: Warm hills)\linebreak{}
(Tallfellow: Temperate forests)\tabularnewline
\hline
O\textbf{rganization:} & Company (2-4), squad (11-20 plus 2 3rd-level sergeants 
and 1 leader of 3rd-6th level), or band (30-100 plus 100\% noncombatants plus 1 
3rd-level sergeant per 20 adults, 5 5th-level lieutenants, 3 7th-level captains, 
6-10 dogs, and 2-5 riding dogs)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 1/2\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Usually neutral\tabularnewline
\hline
A\textbf{dvancement:} & By character class\tabularnewline
\hline
L\textbf{evel Adjustment:} & +0\tabularnewline
\hline
\end{tabular}

Halflings stand about 3 feet tall and usually weigh between 30 and 35 pounds. They 
have brown or black eyes. Halfling men often have long sideburns, but beards are 
rare among them and mustaches almost unseen. Halflings prefer simple, comfortable, 
and practical clothes. Unlike members of most races, they prefer actual comfort 
to shows of wealth. Halflings reach adulthood in their early twenties and generally 
live into the middle of their second century. Halflings speak Halfling and Common.

Most halflings encountered outside their home are warriors; the information in 
the statistics block is for one of 1st level.

COMBAT

Halflings prefer to fight defensively, usually hiding and launching ranged attacks 
as the foe approaches. Their tactics are very much like those of elves but place 
more emphasis on cover and concealment and less on mobility.

\textbf{Halfling Traits (Ex):} Halflings possess the following racial traits.--- 

\parindent=3pt
+2 Dexterity, -2 Strength.---

\parindent=0pt
Small size. +1 bonus to Armor Class, +1 bonus on attack rolls, +4 bonus on Hide 
checks, -4 penalty on grapple checks, lifting and carrying limits 3/4 those of 
Medium characters.---

A halfling's base land speed is 20 feet.--- 

\parindent=3pt
+2 racial bonus on Climb, Jump, and Move Silently checks.--- 

+1 racial bonus on all saving throws.--- 

\parindent=7pt
+2 morale bonus on saving throws against fear. This bonus stacks with the halfling's 
+1 bonus on saving throws in general. --- 

\parindent=3pt
+1 racial bonus on attack rolls with thrown weapons and

\parindent=0pt
slings.--- 

\parindent=3pt
+2 racial bonus on Listen checks.---

\parindent=0pt
Automatic Languages: Common, Halfling. Bonus Languages: Dwarven, Elven, Gnome, 
Goblin, Orc.---

Favored Class: Rogue.

The halfling warrior presented here had the following ability scores before racial 
adjustments: Str 13, Dex 11, Con 12, Int 10, Wis 9, Cha 8.

\vspace{12pt}
SUBRACES

The information above is for the lightfoot halfling, the most common halfling variety. 
There are two other major halfling subraces, which differ from lightfoot halflings 
as follows.

\vspace{12pt}
\section*{\textbf{Tallfellow}}

Tallfellows are somewhat rare among halfling folk. Tallfellows are 4 feet tall 
or more and weigh between 30 and 35 pounds. They generally speak Elven in addition 
to Common and Halfling. 

\textbf{Tallfellow Traits (Ex):} These traits are in addition to the lightfoot 
halfling traits, except where noted.--- 

\parindent=3pt
+2 racial bonus on Search, Spot, and Listen checks. Like an elf, a tallfellow who 
merely passes within 5 feet of a secret or concealed door is entitled to a Search 
check as though actively looking for it. This trait replaces the lightfoot's +2 
bonus on Listen checks.---

\parindent=0pt
Tallfellows are less athletic than lightfoot halflings and do not have a racial 
bonus on Climb, Jump, and Move Silently checks.

\vspace{12pt}
\section*{\textbf{Deep Halfling}}

These halflings are shorter and stockier than the more common lightfeet.

Deep halflings are about 2-1/2 tall and weigh between 30 and 35 pounds. Deep halflings 
speak Dwarven fluently.

\textbf{Deep Halfling Traits (Ex):} These traits are in addition to the lightfoot 
halfling traits, except where noted.---

Darkvision out to 60 feet.---

Stonecunning: Like dwarves, deep halflings have a +2 racial bonus on checks to 
notice unusual stonework. Something that isn't stone but that is disguised as stone 
also counts as unusual stonework. A deep halfling who merely comes within 10 feet 
of unusual stonework can make a check as though actively searching and can use 
the Search skill to find stonework traps as a rogue can. A deep halfling can also 
intuit depth, sensing the approximate distance underground as naturally as a human 
can sense which way is up.--- 

\parindent=3pt
+2 racial bonus on Appraise checks and Craft checks that are related to stone or 
metal.---

\parindent=0pt
Deep halflings are less athletic than lightfoot halflings and do not have a racial 
bonus on Climb, Jump, and Move Silently checks.

\vspace{12pt}
{\LARGE{}HARPY}

\begin{tabular}{|>{\raggedright}p{47pt}|>{\raggedright}p{133pt}|>{\raggedright}p{133pt}|}
\hline
 & H\textbf{arpy} & H\textbf{arpy Archer, 7th-Level Fighter}\tabularnewline
\hline
  & Medium Monstrous Humanoid & Medium Monstrous Humanoid\tabularnewline
\hline
H\textbf{it Dice:} & 7d8 (31 hp) & 7d8 + 7d10 + 28 (103 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +2 & +9\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), fly 80 ft. (average) & 20 ft. (4 squares), 
fly 80 ft. (average)\tabularnewline
\hline
A\textbf{rmor Class:} & 13 (+2 Dex, +1 natural), touch 12, flat-footed 11 & AC 
23 (+5 Dex, +1 natural, +6 \textit{+3 studded leather, }+1 \textit{ring of protection 
+1}), touch 16, flat-footed 18\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +7/+7 & +14/+15\tabularnewline
\hline
A\textbf{ttack:} & Club +7 melee (1d6) & +\textit{1 frost composite longbow }(+1 
Str bonus) +22 ranged (1d8+4/19-20/\ensuremath{\times}3 plus 1d6 cold) or claw 
+15 melee (1d3+1)\tabularnewline
\hline
F\textbf{ull Attack:} & Club +7/+2 melee (1d6) and 2 claws +2 melee (1d3) & +\textit{1 
frost composite longbow }(+1 Str bonus) +22/+17/+12 ranged (1d8+4/19-20/\ensuremath{\times}3 
plus 1d6 cold) or 2 claws +15 melee (1d3+1)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Captivating song & Captivating song\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft. & Darkvision 60 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +2, Ref +7, Will +6 & Fort +11, Ref +14, Will +11\tabularnewline
\hline
A\textbf{bilities:} & Str 10, Dex 15, Con 10, Int 7, Wis 12, Cha 17 & Str 12, Dex 
20, Con 14, Int 6, Wis 11, Cha 19\tabularnewline
\hline
S\textbf{kills:} & Bluff +11, Intimidate +7, Listen +7, Perform (oratory) +5, Spot 
+3 & Bluff +11, Intimidate +5, Listen +7, Perform (oratory) +10, Spot +5\tabularnewline
\hline
F\textbf{eats:} & Dodge, Flyby Attack, Persuasive & Alertness, Improved Critical 
(composite longbow), Improved Initiative, Iron Will, Manyshot, Point Blank Shot, 
Rapid Shot, Weapon Focus (composite longbow), Weapon Specialization (composite 
longbow)\tabularnewline
\hline
E\textbf{nvironment:} & Temperate marshes & Temperate marshes\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or flight (7-12) & Solitary\tabularnewline
\hline
C{\small{}\textbf{hallenge Rating:}} & 4 & 11\tabularnewline
\hline
T\textbf{reasure:} & Standard & Standard (including equipment)\tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic evil & Usually chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & By character class & By character class\tabularnewline
\hline
L{\small{}\textbf{evel Adjustment:}} & +{\small{}3} & +3\tabularnewline
\hline
\end{tabular}

Harpies like to entrance hapless travelers with their magical songs and lead them 
to unspeakable torments. Only when a harpy has finished playing with its new ``toys'' 
will it release them from suffering by killing and consuming them.

COMBAT

When a harpy engages in battle, it prefers to use Flyby Attack and strike with 
a melee weapon.

\textbf{Captivating Song (Su):} The most insidious ability of the harpy is its 
song. When a harpy sings, all creatures (other than harpies) within a 300-foot 
spread must succeed on a DC 16 Will save or become captivated. This is a sonic 
mind-affecting charm effect. A creature that successfully saves cannot be affected 
again by the same harpy's song for 24 hours. The save DC is Charisma-based. 

A captivated victim walks toward the harpy, taking the most direct route available. 
If the path leads into a dangerous area (through flame, off a cliff, or the like), 
that creature gets a second saving throw. Captivated creatures can take no actions 
other than to defend themselves. (Thus, a fighter cannot run away or attack but 
takes no defensive penalties.) A victim within 5 feet of the harpy stands there 
and offers no resistance to the monster's attacks. The effect continues for as 
long as the harpy sings and for 1 round thereafter. A bard's countersong ability 
allows the captivated creature to attempt a new Will save.

\textbf{Skills:} Harpies have a +4 racial bonus on Bluff and Listen checks.

\vspace{12pt}
HARPY ARCHER

A cruel hunter and roaming brigand, the harpy archer has trained as a fighter specializing 
in ranged combat. Harpy archers often become mercenaries, selling their services 
to the highest bidder. When not employed, they make ends meet as highway robbers, 
forcing merchant caravans to pay protection money.

COMBAT

\textbf{Captivating Song (Su):} Will DC 17 negates. 

\textit{Possessions: +3 studded leather, +1 frost composite longbow }(+1 Str bonus), 
10 cold iron arrows, 10 silvered arrows, 5 \textit{+2 arrows, lesser bracers of 
archery, potion of cure moderate wounds, potion of cat's grace, cloak of resistance 
+2, ring of protection +1. }(Different harpy archers may have different possessions.)

\vspace{12pt}
{\LARGE{}HELL HOUND}

\begin{tabular}{|>{\raggedright}p{47pt}|>{\raggedright}p{133pt}|>{\raggedright}p{133pt}|}
\hline
 & H\textbf{ellhound} & N\textbf{essian Warhound}\tabularnewline
\hline
  & Medium Outsider (Evil, Extraplanar, Fire, Lawful) & Large Outsider (Evil, Extraplanar, 
Fire, Lawful)\tabularnewline
\hline
H\textbf{it Dice:} & 4d8+4 (22 hp) & 12d8+60 (114 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +5 & +6\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares) & 40 ft. (8 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 16 (+1 Dex, +5 natural), touch 11, flat-footed 15 & 24 
(-1 size, +2 Dex, +7 natural, +6 \textit{+2 chain shirt barding}), touch 11, flat-footed 
22\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +4/+5 & +12/+24\tabularnewline
\hline
A\textbf{ttack:} & Bite +5 melee (1d8+1 plus 1d6 fire) & Bite +20 melee (2d6+12/19-20 
plus 1d8 fire)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +5 melee (1d8+1 plus 1d6 fire) & Bite +20 melee (2d6+12/19-20 
plus 1d8 fire)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Breath weapon, fiery bite & Breath weapon, fiery bite\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., immunity to fire, scent, vulnerability 
to cold & Darkvision 60 ft., immunity to fire, scent, vulnerability to cold\tabularnewline
\hline
S\textbf{aves:} & Fort +5, Ref +5, Will +4 & Fort +13, Ref +10, Will +9\tabularnewline
\hline
A\textbf{bilities:} & Str 13, Dex 13, Con 13, Int 6, Wis 10, Cha 6 & Str 26, Dex 
14, Con 20, Int 4, Wis 12, Cha 6\tabularnewline
\hline
S\textbf{kills:} & Hide +13, Jump +12, Listen +7, Move Silently +13, Spot +7, Survival 
+7* & Hide +17, Jump +19, Listen +18, Move Silently +21, Spot +18, Survival +8*, 
Tumble +3\tabularnewline
\hline
F\textbf{eats:} & Improved Initiative, Run, Track\textsuperscript{B} & Alertness, 
Improved Critical (bite), Improved Initiative, Track, Weapon Focus (bite)\tabularnewline
\hline
E\textbf{nvironment:} & A lawful evil-aligned plane & A lawful evil plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or pack (5-12) & Solitary, pair, or pack 
(1-2 Nessian warhounds and 5-12 hell hounds)\tabularnewline
\hline
C{\small{}\textbf{hallenge Rating:}} & 3 & 9\tabularnewline
\hline
T\textbf{reasure:} & None & +\textit{2 chain shirt barding}\tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil & Always lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & 5-8 HD (Medium); 9-12 HD (Large) & 13-17 HD (Large); 18-24 
HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +3 (cohort) & +4 (cohort)\tabularnewline
\hline
\end{tabular}

A typical hell hound stands 4-1/2 feet high at the shoulder and weighs 120 pounds.

Hell hounds do not speak but understand Infernal.

COMBAT

Hell hounds are efficient hunters. A favorite pack tactic is to surround prey quietly, 
then attack with one or two hounds, driving it toward the rest with their fiery 
breath. If the prey doesn't run, the pack closes in. Hell hounds track fleeing 
prey relentlessly.

A hell hound's natural weapons, as well as any weapons it wields, are treated as 
evil-aligned

and lawful-aligned for the purpose of overcoming damage reduction.

\textbf{Breath Weapon (Su): }10-foot cone, once every 2d4 rounds, damage 2d6 fire, 
Reflex DC 13 half. The save DC is Constitution-based.

\textbf{Fiery Bite (Su):} A hell hound deals an extra 1d6 points of fire damage 
every time it bites an opponent, as if its bite were a flaming weapon.

\textbf{Skills:} Hell hounds have a +5 racial bonus on Hide and Move Silently checks.

*They also receive a +8 racial bonus on Survival checks when tracking by scent, 
due to their keen sense of smell.

\vspace{12pt}
NESSIAN WARHOUND

Nessian warhounds are coalblack mastiffs the size of draft horses, often fitted 
with shirts of infernal chainmail.

Nessian warhounds resemble hell hounds, except as otherwise noted.

\textbf{Breath Weapon (Su): }10-foot cone, once every 2d4 rounds, damage 3d6 fire, 
Reflex DC 21 half. The save DC is Constitution- based.

\textbf{Fiery Bite (Su): }A Nessian warhound deals an extra 1d8 points of fire 
d a m a g e every time it bites an opponent, as if its bite were a flaming weapon. 
 

\vspace{12pt}
{\LARGE{}HIPPOGRIFF}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{226pt}|}
\hline
  & Large Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 3d10+9 (25 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +2\tabularnewline
\hline
S\textbf{peed:} & 50 ft. (10 squares), fly 100 ft. (average)\tabularnewline
\hline
A\textbf{rmor Class:} & 15 (-1 size, +2 Dex, +4 natural), touch 11, flat-footed 
13\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +3/+11\tabularnewline
\hline
A\textbf{ttack:} & Claw +6 melee (1d4+4)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +6 melee (1d4+4) and bite +1 melee (1d8+2)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:}--- & \tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., low-light vision, scent\tabularnewline
\hline
S\textbf{aves:} & Fort +6, Ref +5, Will +2\tabularnewline
\hline
A\textbf{bilities:} & Str 18, Dex 15, Con 16, Int 2, Wis 13, Cha 8\tabularnewline
\hline
S\textbf{kills:} & Listen +4, Spot +8\tabularnewline
\hline
F\textbf{eats:} & Dodge, Wingover\tabularnewline
\hline
E\textbf{nvironment:} & Temperate hills\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or flight (7-12)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & 4-6 HD (Large); 7-9 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

Hippogriffs are aggressive flying creatures that combine features of horses and 
giant eagles. Voracious omnivores, hippogriffs will hunt humanoids as readily as 
any other meal. A typical hippogriff is 9 feet long, has a wingspan of 20 feet, 
and weighs 1,000 pounds.

COMBAT

Hippogriffs dive at their prey and strike with their clawed forelegs. When they 
cannot dive, they slash with claws and beak. Mated pairs and flights of these creatures 
attack in concert, diving repeatedly to drive away or kill intruders. Hippogriffs 
fight to the death to defend their nests and their hatchlings, which are prized 
as aerial mounts and fetch a handsome price in many civilized areas.

\textbf{Skills:} Hippogriffs have a +4 racial bonus on Spot checks.

TRAINING A HIPPOGRIFF

A hippogriff requires training before it can bear a rider in combat.

Training a hippogriff requires six weeks of work and a DC 25 Handle Animal check. 
Riding a hippogriff requires an exotic saddle. A hippogriff can fight while carrying 
a rider, but the rider cannot also attack unless he or she succeeds on a Ride check. 

Hippogriff eggs are worth 2,000 gp apiece on the open market, while young are worth 
3,000 gp each. Professional trainers charge 1,000 gp to rear or train a hippogriff.

\textbf{Carrying Capacity:} A light load for a hippogriff is up to 300 pounds; 
a medium load, 301-600 pounds; and a heavy load, 601-900 pounds.

\vspace{12pt}
{\LARGE{}HOBGOBLIN}

\begin{tabular}{|>{\raggedright}p{78pt}|>{\raggedright}p{247pt}|}
\hline
  & H\textbf{obgoblin, 1st-Level Warrior}\tabularnewline
\hline
  & Medium Humanoid (Goblinoid)\tabularnewline
\hline
H\textbf{it Dice:} & 1d8+2 (6 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 15 (+1 Dex, +3 studded leather, +1 light shield), touch 
11, flat-footed 14\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +1/+2\tabularnewline
\hline
A\textbf{ttack:} & Longsword +2 melee (1d8+1/19-20) or javelin +2 ranged (1d6+1)\tabularnewline
\hline
F\textbf{ull Attack:} & Longsword +2 melee (1d8+1/19-20) or javelin +2 ranged (1d6+1)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:}--- & \tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +4, Ref +1, Will -1\tabularnewline
\hline
A\textbf{bilities:} & Str 13, Dex 13, Con 14, Int 10, Wis 9, Cha 8\tabularnewline
\hline
S\textbf{kills:} & Hide +3, Listen +2, Move Silently +3, Spot +2\tabularnewline
\hline
F\textbf{eats:} & Alertness\tabularnewline
\hline
E\textbf{nvironment:} & Warm hills\tabularnewline
\hline
O\textbf{rganization:} & Gang (4-9), band (10-100 plus 50\% noncombatants plus 
1 3rd-level sergeant per 20 adults and 1 leader of 4th-6th level), warband (10-24), 
or tribe (30-300 plus 50\% noncombatants plus 1 3rd-level sergeant per 20 adults, 
1 or 2 lieutenants of 4th or 5th level, 1 leader of 6th-8th level, 2-4 dire wolves, 
and 1-4 ogres or 1-2 trolls)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 1/2\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Usually lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & By character class\tabularnewline
\hline
L\textbf{evel Adjustment:} & +1\tabularnewline
\hline
\end{tabular}

Hobgoblins are larger cousins of goblins. Hobgoblins' hair color ranges from dark 
reddish-brown to dark gray. They have dark orange or red-orange skin. Large males 
have blue or red noses. Hobgoblins' eyes are yellowish or dark brown, while their 
teeth are yellow. Their garments tend to be brightly colored, often blood red with 
black-tinted leather. Their weaponry is kept polished and in good repair. Hobgoblins 
speak Goblin and Common. Most hobgoblins encountered outside their homes are warriors; 
the information in the statistics block is for one of 1st level.

COMBAT

These creatures have a strong grasp of strategy and tactics and are capable of 
carrying out sophisticated battle plans. Under the leadership of a skilled strategist 
or tactician, their discipline can prove a deciding factor. Hobgoblins hate elves 
and attack them first, in preference to other opponents.

\textbf{Skills:} Hobgoblins have a +4 racial bonus on Move Silently checks.

The hobgoblin warrior presented here had the following ability scores before racial 
adjustments: Str 13, Dex 11, Con 12, Int 10, Wis 9, Cha 8.

HOBGOBLIN CHARACTERS

Hobgoblin characters possess the following racial traits.--- 

\parindent=3pt
+2 Dexterity, +2 Constitution.---

\parindent=0pt
A hobgoblin's base land speed is 30 feet.---

Darkvision out to 60 feet.--- 

\parindent=3pt
+4 racial bonus on Move Silently checks.---

\parindent=0pt
Automatic Languages: Common, Goblin. Bonus Languages: Draconic, Dwarven, Infernal, 
Giant, Orc.---

Favored Class: Fighter.---

Level adjustment +1.

\vspace{12pt}
{\LARGE{}HOMUNCULUS}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{199pt}|}
\hline
  & Tiny Construct\tabularnewline
\hline
H\textbf{it Dice:} & 2d10 (11 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +2\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), fly 50 ft. (good)\tabularnewline
\hline
A\textbf{rmor Class:} & 14 (+2 Dex, +2 size), touch 14, flat-footed 12\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +1/-8\tabularnewline
\hline
A\textbf{ttack:} & Bite +2 melee (1d4-1 plus poison)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +2 melee (1d4-1 plus poison)\tabularnewline
\hline
S\textbf{pace/Reach:} & 2-1/2 ft./0 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Poison\tabularnewline
\hline
S\textbf{pecial Qualities:} & Construct traits, darkvision 60 ft., low-light vision\tabularnewline
\hline
S\textbf{aves:} & Fort +0, Ref +4, Will +1\tabularnewline
\hline
A\textbf{bilities:} & Str 8, Dex 15, Con ---, Int 10, Wis 12, Cha 7\tabularnewline
\hline
S\textbf{kills:} & Hide +14, Listen +4, Spot +4 \tabularnewline
\hline
F\textbf{eats:} & Lightning Reflexes\tabularnewline
\hline
E\textbf{nvironment:} & Any\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 1\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Any (same as creator)\tabularnewline
\hline
A\textbf{dvancement:} & 3-6 HD (Tiny)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A homunculus is a miniature servant created by a wizard. These creatures are weak 
combatants but make effective spies, messengers, and scouts. A homunculus's creator 
determines its precise features. Homunculi are little more than tools designed 
to carry out assigned tasks. They are extensions of their creators, sharing the 
same alignment and basic nature. A homunculus cannot speak, but the process of 
creating one links it telepathically with its creator.

It knows what its master knows and can convey to him or her everything it sees 
and hears, out to a distance of 1,500 feet. A homunculus never travels beyond this 
range willingly, though it can be removed forcibly. If this occurs, the creature 
does everything in its power to regain contact with its master. An attack that 
destroys a homunculus deals 2d10 points of damage to its master.

If the creature's master is slain, the homunculus also dies, and its body swiftly 
melts away into a pool of ichor.

COMBAT

Homunculi land on their victims and bite with their venomous fangs.

\textbf{Poison (Ex):} Injury, Fortitude DC 13, initial damage sleep for 1 minute, 
secondary damage sleep for another 5d6 minutes. The save DC is Constitution-based 
and includes a +2 racial bonus.

CONSTRUCTION

A homunculus is shaped from a mixture of clay, ashes, mandrake root, spring water, 
and one pint of the creator's own blood. The materials cost 50 gp. The creature's 
master may assemble the body or hire someone else to do the job. Creating the body 
requires a DC 12 Craft (sculpting) check or a DC 12 Craft (pottery) check.

After the body is sculpted, it is animated through an extended magical ritual that 
requires a specially prepared laboratory or workroom, similar to an alchemist's 
laboratory and costing 500 gp to establish. If the creator is personally constructing 
the creature's body, the building and ritual can be performed together.

A homunculus with more than 2 Hit Dice can be created, but each additional Hit 
Die adds +2,000 gp to the cost to create. 

Craft Construct, \textit{arcane eye, mirror image, mending, }caster must be at 
least 4th level; Price --- (never sold); Cost 1,050 gp + 78 XP.

\vspace{12pt}
{\LARGE{}HOWLER}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{233pt}|}
\hline
  & Large Outsider (Chaotic, Evil, Extraplanar)\tabularnewline
\hline
H\textbf{it Dice:} & 6d8+12 (39 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +7\tabularnewline
\hline
S\textbf{peed:} & 60 ft. (12 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 17 (-1 size, +3 Dex, +5 natural), touch 12, flat-footed 
14\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +6/+15\tabularnewline
\hline
A\textbf{ttack:} & Bite +10 melee (2d8+5)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +10 melee (2d8+5) and 1d4 quills +5 melee (1d6+2)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Quills, howl\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +7, Ref +8, Will +7\tabularnewline
\hline
A\textbf{bilities:} & Str 21, Dex 17, Con 15, Int 6, Wis 14, Cha 8\tabularnewline
\hline
S\textbf{kills:} & Climb +14, Hide +8, Listen +13, Move Silently +12, Search +7, 
Spot +13, Survival +2 (+4 following tracks)\tabularnewline
\hline
F\textbf{eats:} & Alertness, Combat Reflexes, Improved Initiative\tabularnewline
\hline
E\textbf{nvironment:} & A chaotic-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary, gang (2-4), or pack (6-10)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 3\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 7-9 HD (Large); 11-18 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +3 (cohort)\tabularnewline
\hline
\end{tabular}

Howlers live on planes where chaos and evil hold sway. A howler is about 8 feet 
long and weighs about 2,000 pounds. Although they are surprisingly intelligent, 
howlers do not speak---they only howl. If there is a language within the howls, 
as some have suggested, even spells cannot decipher it. Howlers understand Abyssal.

COMBAT

Howlers attack in groups, for they are cowardly and cruel. They prefer to charge 
into combat, race out, and then charge in again. A howler's natural weapons, as 
well as any weapons it wields, are treated as chaotic-aligned and evil-aligned 
for the purpose of overcoming damage reduction.

\textbf{Quills (Ex):} A howler's neck bristles with long quills. While biting, 
the creature thrashes about, striking with 1d4 of them. An opponent hit by a howler's 
quill attack must succeed on a DC 16 Reflex save or have the quill break off in 
his or her flesh. Lodged quills impose a -1 penalty on attacks, saves, and checks 
per quill. The save DC is Dexterity-based.

A quill can be removed safely with a DC 20 Heal check; otherwise, removing a quill 
deals an extra 1d6 points of damage.

\textbf{Howl (Ex):} All beings other than outsiders that hear the creature's howling 
for an hour or longer are subject to its effect, though it does not help the howler 
in combat. Anyone within hearing range of a howler for a full hour must succeed 
on a DC 12 Will save or take 1 point of Wisdom damage. The save DC is Charisma-based. 
The save must be repeated for each hour of exposure. This is a sonic mind-affecting 
effect.

TRAINING A HOWLER

Although intelligent, a howler requires training before it can bear a rider in 
combat. To be trained, a howler must have a friendly attitude toward the trainer 
(this can be achieved through a successful Diplomacy check). Training a friendly 
howler requires six weeks of work and a DC 25 Handle Animal check. Riding a howler 
requires an exotic saddle. A howler can fight while carrying a rider, but the rider 
cannot also attack unless he or she succeeds on a Ride check.

\textbf{Carrying Capacity:} A light load for a howler is up to 460 pounds; a medium 
load, 461-920 pounds; and a heavy load, 921-1,380 pounds. A howler can drag 6,900 
pounds. 

\vspace{12pt}
{\LARGE{}HYDRA}

\begin{tabular}{|>{\raggedright}p{52pt}|>{\raggedright}p{83pt}|>{\raggedright}p{83pt}|>{\raggedright}p{83pt}|}
\hline
  & F\textbf{ive-Headed Hydra} & S\textbf{ix-Headed Hydra} & S\textbf{even-Headed 
Hydra}\tabularnewline
\hline
  & Huge Magical Beast & Huge Magical Beast & Huge Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 5d10+28 (55 hp) & 6d10+33 (66 hp) & 7d10+38 (77 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1 & +1 & +1\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), swim 20 ft. & 20 ft. (4 squares), swim 20 
ft. & 20 ft. (4 squares), swim 20 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 15 (-2 size, +1 Dex, +6 natural), touch 9, flat-footed 
14 & 16 (-2 size, +1 Dex, +7 natural), touch 9, flat-footed 15 & 17 (-2 size, +1 
Dex, +8 natural), touch 9, flat-footed 16\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +5/+16 & +6/+17 & +7/+19\tabularnewline
\hline
A\textbf{ttack:} & 5 bites +6 melee (1d10+3) & 6 bites +8 melee (1d10+3) & 7 bites 
+10 melee (1d10+4)\tabularnewline
\hline
F\textbf{ull Attack:} & 5 bites +6 melee (1d10+3) & 6 bites +8 melee (1d10+3) & 7 
bites +10 melee (1d10+4)\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./10 ft. & 15 ft./10 ft. & 15 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:}--- & --- & --- & \tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., fast healing 15, low-light vision, 
scent & Darkvision 60 ft., fast healing 16, low-light vision, scent & Darkvision 
60 ft., fast healing 17, low-light vision, scent\tabularnewline
\hline
S\textbf{aves:} & Fort +9, Ref +5, Will +3 & Fort +10, Ref +6, Will +4 & Fort +10, 
Ref +6, Will +4\tabularnewline
\hline
A\textbf{bilities:} & Str 17, Dex 12, Con 20, Int 2, Wis 10, Cha 9 & Str 17, Dex 
12, Con 20, Int 2, Wis 10, Cha 9 & Str 19, Dex 12, Con 20, Int 2, Wis 10, Cha 9\tabularnewline
\hline
S\textbf{kills:} & Listen +6, Spot +6, Swim +11 & Listen +6, Spot +7, Swim +11 & Listen 
+7, Spot +7, Swim +12\tabularnewline
\hline
F\textbf{eats:} & Combat Reflexes\textsuperscript{B}, Iron Will, Toughness & Combat 
Reflexes\textsuperscript{B}, Iron Will, Toughness, Weapon Focus (bite) & Combat 
Reflexes\textsuperscript{B}{\scriptsize{} }, Iron Will, Toughness, Weapon Focus 
(bite)\tabularnewline
\hline
E\textbf{nvironment:} & Temperate marshes\linebreak{}
(Pyro: Warm marshes) \linebreak{}
(Cryo: Cold marshes) & Temperate marshes \linebreak{}
(Pyro: Warm marshes) \linebreak{}
(Cryo: Cold marshes) & Temperate marshes \linebreak{}
(Pyro: Warm marshes) \linebreak{}
(Cryo: Cold marshes)\tabularnewline
\hline
O\textbf{rganization:} & Solitary & Solitary & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 4 (normal); \linebreak{}
6 (pyro- or cryo-) & 5 (normal); \linebreak{}
7 (pyro- or cryo-) & 6 (normal); \linebreak{}
8 (pyro- or cryo-)\tabularnewline
\hline
T\textbf{reasure:} & 1/10 coins; 50\% goods; 50\% items & 1/10 coins; 50\% goods; 
50\% items & 1/10 coins; 50\% goods; 50\% items\tabularnewline
\hline
A\textbf{lignment:} & Usually neutral & Usually neutral & Usually neutral\tabularnewline
\hline
A\textbf{dvancement:}--- & --- & --- & \tabularnewline
\hline
L\textbf{evel Adjustment:}--- & --- & --- & \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{51pt}|>{\raggedright}p{83pt}|>{\raggedright}p{83pt}|>{\raggedright}p{83pt}|}
\hline
  & E\textbf{ight-Headed Hydra } & N\textbf{ine-Headed Hydra} & T\textbf{en-Headed 
Hydra}\tabularnewline
\hline
  & Huge Magical Beast & Huge Magical Beast & Huge Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 8d10+43 (87 hp) & 9d10+48 (97 hp) & 10d10+53 (108 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1 & +1 & +1\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), swim 20 ft. & 20 ft. (4 squares), swim 20 
ft. & 20 ft. (4 squares), swim 20 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 18 (-2 size, +1 Dex, +9 natural), touch 9, flat-footed 
17 & 19 (-2 size, +1 Dex, +10 natural), touch 9, flat-footed 18 & 20 (-2 size, 
+1 Dex, +11 natural), touch 9, flat-footed 19\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +8/+20 & +9/+22 & +10/+23\tabularnewline
\hline
A\textbf{ttack:} & 8 bites +11 melee (1d10+4) & 9 bites +13 melee (1d10+5) & 10 
bites +14 melee (1d10+5)\tabularnewline
\hline
F\textbf{ull Attack:} & 8 bites +11 melee (1d10+4) & 9 bites +13 melee (1d10+5) & 10 
bites +14 melee (1d10+5)\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./10 ft. & 15 ft./10 ft. & 15 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:}--- & --- & --- & \tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., fast healing 18, low-light vision, 
scent & Darkvision 60 ft., fast healing 19, low-light vision, scent & Darkvision 
60 ft., fast healing 20, low-light vision, scent\tabularnewline
\hline
S\textbf{aves:} & Fort +11, Ref +7, Will +4 & Fort +11, Ref +7, Will +5 & Fort 
+12, Ref +8, Will +3\tabularnewline
\hline
A\textbf{bilities:} & Str 19, Dex 12, Con 20, Int 2, Wis 10, Cha 9 & Str 21, Dex 
12, Con 20, Int 2, Wis 10, Cha 9 & Str 21, Dex 12, Con 20, Int 2, Wis 10, Cha 9\tabularnewline
\hline
S\textbf{kills:} & Listen +7, Spot +8, Swim +12 & Listen +8, Spot +8, Swim +13 & Listen 
+8, Spot +9, Swim +13\tabularnewline
\hline
F\textbf{eats:} & Combat Reflexes\textsuperscript{B}{\scriptsize{} }, Iron Will, 
Toughness, Weapon Focus (bite) & Blind-Fight, Combat Reflexes\textsuperscript{B}, 
Iron Will, Toughness, Weapon Focus (bite) & Blind-Fight, Combat Reflexes\textsuperscript{B}, 
Iron Will, Toughness, Weapon Focus (bite)\tabularnewline
\hline
E\textbf{nvironment:} & Temperate marshes \linebreak{}
(Pyro: Warm marshes) \linebreak{}
(Cryo: Cold marshes) & Temperate marshes \linebreak{}
(Pyro: Warm marshes) \linebreak{}
(Cryo: Cold marshes) & Temperate marshes \linebreak{}
(Pyro: Warm marshes) \linebreak{}
(Cryo: Cold marshes)\tabularnewline
\hline
O\textbf{rganization:} & Solitary & Solitary & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 7 (normal); \linebreak{}
9 (pyro- or cryo-) & 8 (normal); \linebreak{}
10 (pyro- or cryo-) & 9 (normal); \linebreak{}
11 (pyro- or cryo-)\tabularnewline
\hline
T\textbf{reasure:} & 1/10 coins; 50\% goods; 50\% items & 1/10 coins; 50\% goods; 
50\% items & 1/10 coins; 50\% goods; 50\% items\tabularnewline
\hline
A\textbf{lignment:} & Usually neutral & Usually neutral & Usually neutral\tabularnewline
\hline
A\textbf{dvancement:}--- & --- & --- & \tabularnewline
\hline
L\textbf{evel Adjustment:}--- & --- & --- & \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{54pt}|>{\raggedright}p{130pt}|>{\raggedright}p{130pt}|}
\hline
  & E\textbf{leven-Headed Hydra} & T\textbf{welve-Headed Hydra}\tabularnewline
\hline
  & Huge Magical Beast & Huge Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 11d10+58 (118 hp) & 12d10+63 (129 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1 & +1\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), swim 20 ft. & 20 ft. (4 squares), swim 20 
ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 21 (-2 size, +1 Dex, +12 natural), touch 9, flat-footed 
20 & 22 (-2 size, +1 Dex, +13 natural), touch 9, flat-footed 21\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +11/+25 & +12/+26\tabularnewline
\hline
A\textbf{ttack:} & 11 bites +16 melee (1d10+6) & 12 bites +17 melee (2d8+6)\tabularnewline
\hline
F\textbf{ull Attack:} & 11 bites +16 melee (1d10+6) & 12 bites +17 melee (2d8+6)\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./10 ft. & 15 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:}--- & --- & \tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., fast healing 21, low-light vision, 
scent & Darkvision 60 ft., fast healing 22, low-light vision, scent\tabularnewline
\hline
S\textbf{aves:} & Fort +12, Ref +8, Will +5 & Fort +13, Ref +9, Will +6\tabularnewline
\hline
A\textbf{bilities:} & Str 23, Dex 12, Con 20, Int 2, Wis 10, Cha 9 & Str 23, Dex 
12, Con 20, Int 2, Wis 10, Cha 9\tabularnewline
\hline
S\textbf{kills:} & Listen +9, Spot +9, Swim +14 & Listen +9, Spot +10, Swim +14\tabularnewline
\hline
F\textbf{eats:} & Blind-Fight, Combat Reflexes\textsuperscript{B}, Iron Will, Toughness, 
Weapon Focus (bite) & Blind-Fight, Combat Reflexes\textsuperscript{B}{\scriptsize{} 
}, Improved Natural Attack (bite), Iron Will, Toughness, Weapon Focus (bite)\tabularnewline
\hline
E\textbf{nvironment:} & Temperate marshes \linebreak{}
(Pyro: Warm marshes) \linebreak{}
(Cryo: Cold marshes) & Temperate marshes \linebreak{}
(Pyro: Warm marshes) \linebreak{}
(Cryo: Cold marshes)\tabularnewline
\hline
O\textbf{rganization:} & Solitary & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 10 (normal); \linebreak{}
12 (pyro- or cryo-) & 11 (normal); \linebreak{}
13 (pyro- or cryo-)\tabularnewline
\hline
T\textbf{reasure:} & 1/10 coins; 50\% goods; 50\% items & 1/10 coins; 50\% goods; 
50\% items\tabularnewline
\hline
A\textbf{lignment:} & Usually neutral & Usually neutral\tabularnewline
\hline
A\textbf{dvancement:}--- & --- & \tabularnewline
\hline
L\textbf{evel Adjustment:}--- & --- & \tabularnewline
\hline
\end{tabular}

Hydras are reptile-like monsters with multiple heads. A hydra is gray-brown to 
dark brown, with a light yellow or tan underbelly. The eyes are amber and the teeth 
are yellow-white. It is about 20 feet long and weighs about 4,000 pounds. Hydras 
do not speak.

COMBAT

Hydras can attack with all their heads at no penalty, even if they move or charge 
during the round.

A hydra can be killed either by severing all its heads or by slaying its body. 
To sever a head, an opponent must make a successful sunder attempt with a slashing 
weapon. (The player should declare where the attack is aimed before making the 
attack roll.) Making a sunder attempt provokes an attack of opportunity unless 
the foe has the Improved Sunder feat. An opponent can strike at a hydra's heads 
from any position in which he could strike at the hydra itself, because the hydra's 
head writhe and whip about in combat. An opponent can ready an action to attempt 
to sunder a hydra's head when the creature bites at him. Each of a hydra's heads 
has hit points equal to the creature's full normal hit point total, divided by 
its original number of heads. Losing a head deals damage to the body equal to half 
the head's full normal hit points. A natural reflex seals the neck shut to prevent 
further blood loss. A hydra can no longer attack with a severed head but takes 
no other penalties.

Each time a head is severed, two new heads spring from the stump in 1d4 rounds. 
A hydra can never have more than twice its original number of heads at any one 
time, and any extra heads it gains beyond its original number wither and die within 
a day. To prevent a severed head from growing back into two heads, at least 5 points 
of fire or acid damage must be dealt to the stump (a touch attack to hit) before 
the new heads appear. A flaming weapon (or similar effect) deals its energy damage 
to the stump in the same blow in which a head is severed. Fire or acid damage from 
an area effect may burn multiple stumps in addition to dealing damage to the hydra's 
body. A hydra does not die from losing its heads until all its heads have been 
cut off and the stumps seared by fire or acid. 

A hydra's body can be slain just like any other creature's, but hydras possess 
fast healing (see below) and are difficult to defeat in this fashion. Any attack 
that is not (or cannot be) an attempt to sunder a head affects the body.

Targeted magical effects cannot sever a hydra's heads (and thus must be directed 
at the body) unless they deal slashing damage and could be used to make sunder 
attempts.

\textbf{Fast Healing (Ex):} Each round, a hydra heals damage equal to 10 + the 
number of its original heads. 

\textbf{Skills: }Hydras have a +2 racial bonus on Listen and Spot checks, thanks 
to their multiple heads.

A hydra has a +8 racial bonus on any Swim check to perform some special action 
or avoid a hazard. It can always choose to take 10 on a Swim check, even if distracted 
or endangered. It can use the run action while swimming, provided it swims in a 
straight line.

\textbf{Feats:} A hydra's Combat Reflexes feat allows it to use all its heads for 
attacks of opportunity.

\vspace{12pt}
PYROHYDRA

Huge Magical Beast (Fire)

These reddish hydras can breathe jets of fire 10 feet high, 10 feet wide, and 20 
feet long. All heads breathe once every 1d4 rounds. Each jet deals 3d6 points of 
fire damage per head. A successful Reflex save halves the damage. The save DC is 
10 + 1/2 hydra's original number of heads + hydra's Con modifier.

Fire attacks cannot prevent a pyrohydra's stump from growing new heads (since a 
pyrohydra has immunity to fire), but 5 points of cold damage does.

\vspace{12pt}
CRYOHYDRA

Huge Magical Beast (Cold)

These purplish hydras can breathe jets of frost 10 feet high, 10 feet wide, and 
20 feet long. All heads breathe once every 1d4 rounds. Each jet deals 3d6 points 
of cold damage per head. A successful Reflex save halves the damage. The save DC 
is 10 + 1/2 hydra's original number of heads + hydra's Con modifier.

Cold attacks cannot prevent a cryohydra's stump from growing new heads (since a 
cryohydra has immunity to cold), but 5 points of fire damage does.

\vspace{12pt}
{\LARGE{}INEVITABLE}

Inevitables are constructs whose sole aim is to enforce the natural laws of the 
universe.

Each type of inevitable is designed to find and punish a particular kind of transgression, 
hunting down a person or group that has violated a fundamental principle. When 
an inevitable is created, it receives its first mission, then finds the transgressors 
and metes out appropriate punishment. The sentence is usually death, although some 
inevitables insist on compensation to the wronged party instead, using \textit{geas 
}and \textit{mark of justice }to ensure compliance. From its first step, an inevitable 
focuses totally on its target. It continues its efforts no matter how cold the 
trail or hopeless the task. Inevitables are single-minded in pursuit of their quarry, 
but they are under orders to leave innocents alone. Accomplices to their prey are 
fair game, however, which sometimes creates conflicts within their programming. 

Inevitables gladly sacrifice themselves to complete a mission, but they aren't 
suicidal. Faced with impending defeat, they are likely to withdraw and seek a way 
to even the odds. They are determined but patient foes. They ally with others if 
that helps accomplish their mission, but they have a hard time keeping allies for 
long.

Inevitables tend to stick out in a crowd while they're in observation mode, but 
they seem oblivious to the attention. Their forms vary, but all inevitables are 
gold-and-silver clockwork creatures, with gears and pistons where muscles would 
be on flesh-and-blood creatures. Their eyes glow with a golden radiance.

Note that unlike most constructs, inevitables have an Intelligence score and can 
think, learn, and remember.

Inevitables speak Abyssal, Celestial, Infernal, and the native language of their 
first target.

COMBAT

Unless their very existence is threatened, inevitables focus completely on the 
transgressor they've been assigned to, ignoring other combatants completely. An 
inevitable might attack anyone who hinders its progress, but it won't tarry beyond 
the point where it can reengage its quarry. Inevitables take self-defense very 
seriously; anyone who attacks an inevitable with what the creature perceives as 
deadly force is met with deadly force in return.

An inevitable's natural weapons, as well as any weapons it wields, are treated 
as lawful-aligned for the purpose of overcoming damage reduction.

\textbf{Fast Healing (Ex): }An inevitable heals a certain amount of damage each 
round (specified in each variety's description) as long as it has at least 1 hit 
point. However, damage dealt by chaotic weapons heals at the normal rate.

KOLYARUT

\begin{tabular}{|>{\raggedright}p{79pt}|>{\raggedright}p{246pt}|}
\hline
  & Medium Construct (Extraplanar, Lawful)\tabularnewline
\hline
Hit Dice: & 13d10+20 (91 hp)\tabularnewline
\hline
Initiative: & +1\tabularnewline
\hline
Speed: & 20 ft. in banded mail (4 squares); base speed 30 ft.\tabularnewline
\hline
Armor Class: & 27 (+1 Dex, +10 natural, +6 banded mail) touch 11, flat-footed 26\tabularnewline
\hline
Base Attack/Grapple: & +9/+11\tabularnewline
\hline
Attack: & Vampiric touch +11 melee touch (5d6) or enervation ray +10 ranged touch 
(as spell) or \textit{+2 longsword }+13 melee (1d8+5/19-20) or slam +11 melee (1d6+3)\tabularnewline
\hline
Full Attack: & Vampiric touch +11/+6 melee touch (5d6) or enervation ray +10 ranged 
touch (as spell) or \textit{+2 longsword }+13/+8 melee (1d8+5/19-20) or slam +11/+6 
melee (1d6+3)\tabularnewline
\hline
Space/Reach: & 5 ft./5 ft.\tabularnewline
\hline
Special Attacks: & Enervation ray, spell-like abilities, vampiric touch\tabularnewline
\hline
Special Qualities: & Construct traits, damage reduction 10/chaotic, darkvision 
60 ft., fast healing 5, low-light vision, spell resistance 22\tabularnewline
\hline
Saves: & Fort +6, Ref +7, Will +7\tabularnewline
\hline
Abilities: & Str 14, Dex 13, Con ---, Int 10, Wis 17, Cha 16\tabularnewline
\hline
Skills: & Diplomacy +5, Disguise +12, Gather Information +12, Listen +11, Search 
+5, Sense Motive +12, Spot +11, Survival +3 (+5 following tracks)\tabularnewline
\hline
Feats: & Alertness, Combat Casting, Great Fortitude, Lightning Reflexes, Quickened 
Spell-Like Ability (\textit{suggestion})\tabularnewline
\hline
Environment: & A lawful-aligned plane\tabularnewline
\hline
Organization: & Solitary\tabularnewline
\hline
Challenge Rating: & 12\tabularnewline
\hline
Treasure: & None\tabularnewline
\hline
Alignment: & Always lawful neutral\tabularnewline
\hline
Advancement: & 14-22 HD (Medium); 23-39 HD (Large)\tabularnewline
\hline
Level Adjustment:--- & \tabularnewline
\hline
\end{tabular}

Kolyaruts mete out punishment to those who break bargains and oaths. 

Before beginning a mission against a deal-breaker, a kolyarut learns as much about 
the contract or oath as possible. It's not interested in those who break deals 
accidentally or against their will--- only those who willingly break contracts 
violate the principle that kolyaruts are created to uphold. If a written contract 
was broken, the kolyarut typically carries a copy of the contract with it.

Kolyaruts are the most talkative of the inevitables, making credible attempts at 
social niceties such as proper greetings before getting down to the matter at hand. 
They can use \textit{disguise self }to appear as almost any kind of humanoid---useful 
if they need to go undercover to catch their quarry.

Combat

Like all inevitables, a kolyarut is patient enough to study a target before striking. 
It has a good idea of the deal-breaker's abilities and defenses before it enters 
battle. When it fights, it tries to get the conflict over as soon as possible, 
minimizing excess bloodshed and mayhem. It doesn't let concern for innocents delay 
or endanger its mission, however.

A kolyarut's favorite tactic is to use \textit{invisibility }or \textit{disguise 
self }to sneak close, then eliminate the quarry with its vampiric touch ability 
before it can react. A kolyarut has no compunctions about using its vampiric touch 
ability on allies to increase its own power, if doing so helps it complete its 
mission.

A kolyarut's natural weapons, as well as any weapons it wields, are treated as 
lawful-aligned for the purpose of overcoming damage reduction.

\textbf{Enervation Ray (Su):} A kolyarut can fire a black enervation ray at targets 
within 200 feet. The effect is identical with the \textit{enervation }spell (caster 
level 13th).

\textbf{Spell-Like Abilities:} At will---\textit{discern lies }(DC 17)\textit{, 
disguise self, fear }(DC 17)\textit{, hold person }(DC 16)\textit{, invisibility, 
locate creature, suggestion }(DC 16)\textit{; }1/day---\textit{hold monster }(DC 
17)\textit{, mark of justice; }1/week--- \textit{geas/quest. }Caster level 13th. 
The save DCs are Charisma-based.

\textbf{Vampiric Touch (Su): }As a melee touch attack, a kolyarut can steal life 
force from its foe, as the \textit{vampiric touch }spell (caster level 13th).

\textbf{Skills: }A kolyarut has a +4 racial bonus on Disguise, Gather Information, 
and Sense Motive checks.

\vspace{12pt}
MARUT

\begin{tabular}{|>{\raggedright}p{75pt}|>{\raggedright}p{251pt}|}
\hline
  & Large Construct (Extraplanar, Lawful)\tabularnewline
\hline
H\textbf{it Dice:} & 15d10+30 (112 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 30 ft. in full plate armor (6 squares); base speed 40 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 34 (-1 size, +1 Dex, +16 natural, +8 full plate armor), 
touch 10, flat-footed 33\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +11/+27\tabularnewline
\hline
A\textbf{ttack:} & Slam +22 melee (2d6+12 plus 3d6 sonic or 3d6 electricity)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 slams +22 melee (2d6+12 plus 3d6 sonic or 3d6 electricity)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Fists of thunder and lightning, spell-like abilities\tabularnewline
\hline
S\textbf{pecial Qualities:} & Construct traits, damage reduction 15/chaotic, darkvision 
60 ft., fast healing 10, low-light vision, spell resistance 25\tabularnewline
\hline
S\textbf{aves:} & Fort +7, Ref +6, Will +8\tabularnewline
\hline
A\textbf{bilities:} & Str 35, Dex 13, Con ---, Int 12, Wis 17, Cha 18\tabularnewline
\hline
S\textbf{kills:} & Concentration +13, Diplomacy +6, Knowledge (religion) +10, Listen 
+16, Search +10, Sense Motive +12, Spot +16, Survival +3 (+5 following tracks)\tabularnewline
\hline
F\textbf{eats:} & Ability Focus (fists), Awesome Blow, Combat Casting, Great Fortitude, 
Improved Bull Rush, Power Attack\tabularnewline
\hline
E\textbf{nvironment:} & A lawful-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 15\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always lawful neutral\tabularnewline
\hline
A\textbf{dvancement:} & 16-28 HD (Large); 29-45 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

Maruts confront those who would try to deny the grave itself.

Any who use unnatural means to extend their life span could be targeted by a marut. 
Those who take extraordinary measures to cheat death in some other way might be 
labeled transgressors as well. Those who use magic to reverse death aren't worthy 
of a marut's attention unless they do so repeatedly or on a massive scale.

When a marut has identified its target, it walks surely and implacably toward the 
foe, never resting.

Combat

Once it has found its target, a marut brings it the death it has been trying to 
avoid. Those who defile death through necromancy may instead receive a \textit{geas 
}and/or \textit{mark of justice }to enforce proper respect. It typically uses \textit{wall 
of force }to shut off any escape routes, then opens up with \textit{chain lightning 
}while it closes to melee range. Once there, it strikes with its massive fists, 
using \textit{circle of death }if beset by numbers of defenders. It hits spellcasting 
opponents with repeated uses of \textit{greater dispel magic, }and it uses \textit{dimension 
door }and \textit{locate creature }to track down foes who flee.

A marut's natural weapons, as well as any weapons it wields, are treated as lawful-aligned 
for the purpose of overcoming damage reduction.

\textbf{Fists of Thunder and Lightning (Su):} A marut's left fist delivers a loud 
thunderclap whenever it hits something, dealing an extra 3d6 points of sonic damage 
and causing the target to be deafened for 2d6 rounds (Fortitude DC 31 negates the 
deafness). Its right fist delivers a shock for an extra 3d6 points of electricity 
damage, and the flash of lightning causes the target to be blinded for 2d6 rounds 
(Fortitude DC 31 negates the blindness). The save DCs are Strength-based and include 
the marut's Ability Focus feat.

\textbf{Spell-Like Abilities:} At will---\textit{air walk, dimension door, fear 
}(DC 18)\textit{, greater command }(DC 19)\textit{, greater dispel magic, mass 
inflict light wounds }(DC 19), \textit{locate creature, true seeing; }1/day---\textit{chain 
lightning }(DC 20)\textit{, circle of death }(DC 20)\textit{, mark of justice, 
wall of force; }1/week--- \textit{earthquake }(DC 22)\textit{, geas/quest, plane 
shift }(DC 21)\textit{. }Caster level 14th. The save DCs are Charisma-based.

\textbf{Skills:} A marut has a +4 racial bonus on Concentration, Listen, and Spot 
checks.

\vspace{12pt}
ZELEKHUT

\begin{tabular}{|>{\raggedright}p{77pt}|>{\raggedright}p{249pt}|}
\hline
  & Large Construct (Extraplanar, Lawful)\tabularnewline
\hline
H\textbf{it Dice:} & 8d10+30 (74 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +0\tabularnewline
\hline
S\textbf{peed:} & 35 ft. in plate barding (7 squares), fly 40 ft. (average) in 
plate barding (8 squares); base speed 50 ft., fly 60 ft. (average)\tabularnewline
\hline
A\textbf{rmor Class:} & 27 (-1 size, +10 natural, +8 plate barding), touch 9, flat-footed 
27\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +6/+15\tabularnewline
\hline
A\textbf{ttack:} & Spiked chain +10 melee (2d6+5 plus 1d6 electricity)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 spiked chains +10 melee (2d6+5 plus 1d6 electricity)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Spell-like abilities\tabularnewline
\hline
S\textbf{pecial Qualities:} & Construct traits, damage reduction 10/chaotic, darkvision 
60 ft., fast healing 5, low-light vision, spell resistance 20\tabularnewline
\hline
S\textbf{aves:} & Fort +4, Ref +2, Will +5\tabularnewline
\hline
A\textbf{bilities:} & Str 21, Dex 11, Con ---, Int 10, Wis 17, Cha 15\tabularnewline
\hline
S\textbf{kills:} & Diplomacy +4, Listen +9, Search +9, Sense Motive +12, Spot +9, 
Survival +3 (+5 following tracks)\tabularnewline
\hline
F\textbf{eats:} & Great Fortitude, Ride-By Attack, Spirited Charge\tabularnewline
\hline
E\textbf{nvironment:} & A lawful-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 9\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always lawful neutral\tabularnewline
\hline
A\textbf{dvancement:} & 9-16 HD (Large); 17-24 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +7\tabularnewline
\hline
\end{tabular}

Zelekhuts are charged with hunting down those who would deny justice---especially 
those who flee to escape punishment. Expert trackers, they use a combination of 
natural skill and magic to find fugitives wherever they may hide.

A zelekhut may initially appear rather non-threatening---but when it is about to 
enter combat, it can cause two spiked chains to spring forth from its forearms 
as a free action. In similar fashion, it can cause a pair of golden metallic wings 
to emerge from its back. Retracting the chains or the wings is also a free action.

Combat

Once it has found its fugitive, a zelekhut uses its speed and its spell-like abilities 
to cover the most likely escape routes. Then it immobilizes any defenders while 
attempting to protect any innocent bystanders. Finally, it apprehends the fugitive 
with its spiked chains, tripping or disarming the foe as needed.

A zelekhut's natural weapons, as well as any weapons it wields, are treated as 
lawful-aligned for the purpose of overcoming damage reduction.

\textbf{Spell-Like Abilities: }At will---\textit{clairaudience/clairvoyance, dimensional 
anchor, dispel magic, fear }(DC 16)\textit{, hold person }(DC 15)\textit{, locate 
creature, true seeing; }3/day---\textit{hold monster }(DC 17)\textit{, mark of 
justice}; 1/week--- \textit{lesser geas }(DC 16)\textit{. }Caster level 8th. The 
save DCs are Charisma-based.

\textbf{Skills: }A zelekhut has a +4 racial bonus on Search and Sense Motive checks.

\textbf{Feats:} Due to its centaur-like construction, a zelekhut qualifies for 
feats as if it had the Mounted Combat feat.

\vspace{12pt}
{\LARGE{}INVISIBLE STALKER}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{231pt}|}
\hline
  & Large Elemental (Air, Extraplanar)\tabularnewline
\hline
H\textbf{it Dice:} & 8d8+16 (52 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +8\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares), fly 30 ft. (perfect)\tabularnewline
\hline
A\textbf{rmor Class:} & 17 (-1 size, +4 Dex, +4 natural), touch 13, flat-footed 
13\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +6/+14\tabularnewline
\hline
A\textbf{ttack:} & Slam +10 melee (2d6+4)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 slams +10 melee (2d6+4)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:}--- & \tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., elemental traits, natural invisibility, 
improved tracking\tabularnewline
\hline
S\textbf{aves:} & Fort +4, Ref +10, Will +4\tabularnewline
\hline
A\textbf{bilities:} & Str 18, Dex 19, Con 14, Int 14, Wis 15, Cha 11\tabularnewline
\hline
S\textbf{kills:} & Listen +13, Move Silently +15, Search +13, Spot +13, Survival 
+2 (+4 following tracks)\tabularnewline
\hline
F\textbf{eats:} & Combat Reflexes, Improved Initiative, Weapon Focus (slam)\tabularnewline
\hline
E\textbf{nvironment:} & Elemental Plane of Air\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 7\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Usually neutral\tabularnewline
\hline
A\textbf{dvancement:} & 9-12 HD (Large); 13-24 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

Invisible stalkers are creatures native to the Elemental Plane of Air. They sometimes 
serve wizards and sorcerers, who summon them to perform specific tasks.

A summoned invisible stalker undertakes whatever task the summoner commands, even 
if the task sends it hundreds or thousands of miles away. The creature follows 
a command until the task is completed and obeys only the summoner. However, it 
resents protracted missions or complex tasks and seeks to pervert its instructions 
accordingly.

Invisible stalkers have an amorphous form. A \textit{see invisibility }spell shows 
only a dim outline of a cloud, while a \textit{true seeing }spell reveals a roiling 
cloud of vapor.

These creatures speak only Auran but can understand Common.

COMBAT

An invisible stalker attacks by using the air itself as a weapon. It creates a 
sudden, intense blast of wind that pounds a single target on the same plane as 
the creature.

An invisible stalker can be killed only when it is on the Elemental Plane of Air. 
When performing a task elsewhere, it automatically returns to its home plane when 
it takes damage sufficient to destroy it.

\textbf{Natural Invisibility (Su):} This ability is constant, allowing a stalker 
to remain invisible even when attacking. This ability is inherent and not subject 
to the \textit{invisibility purge }spell.

\textbf{Improved Tracking (Ex):} An invisible stalker is a consummate tracker and 
makes Spot checks instead of the usual Survival checks to trace a creature's passage.

\newpage

\end{document}
