%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{array}
\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

This material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

{\LARGE{}MONSTERS (T-Z)}

\vspace{12pt}
{\LARGE{}TARRASQUE}

\begin{tabular}{|>{\raggedright}p{87pt}|>{\raggedright}p{238pt}|}
\hline
  & Colossal Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 48d10+594 (858 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +7\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 35 (-8 size, +3 Dex, +30 natural), touch 5, flat-footed 
32\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +48/+81\tabularnewline
\hline
A\textbf{ttack:} & Bite +57 melee (4d8+17/18-20/x3)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +57 melee (4d8+17/18-20/x3) and 2 horns +52 melee 
(1d10+8) and 2 claws +52 melee (1d12+8) and tail slap +52 melee (3d8+8)\tabularnewline
\hline
S\textbf{pace/Reach:} & 30 ft./20 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Augmented critical, frightful presence, improved grab, 
rush, swallow whole\tabularnewline
\hline
S\textbf{pecial Qualities:} & Carapace, damage reduction 15/epic, immunity to fire, 
poison, disease, energy drain, and ability damage, regeneration 40, scent, spell 
resistance 32\tabularnewline
\hline
S\textbf{aves:} & Fort +38, Ref +29, Will +20\tabularnewline
\hline
A\textbf{bilities:} & Str 45, Dex 16, Con 35, Int 3, Wis 14, Cha 14\tabularnewline
\hline
S\textbf{kills:} & Listen +17, Search +9, Spot +17, Survival +14 (+16 following 
tracks)\tabularnewline
\hline
F\textbf{eats:} & Alertness, Awesome Blow, Blind-Fight, Cleave, Combat Reflexes, 
Dodge, Great Cleave, Improved Bull Rush, Improved Initiative, Iron Will, Power 
Attack, Toughness (6)\tabularnewline
\hline
E\textbf{nvironment:} & Any\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 20\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & 49+ HD (Colossal)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

The tarrasque is 70 feet long and 50 feet tall, and it weighs about 130 tons.

The tarrasque cannot speak.

COMBAT

The tarrasque attacks with its claws, teeth, horns, and tail.

The tarrasque's natural weapons are treated as epic weapons for the purpose of 
overcoming damage reduction.

\textbf{Augmented Critical (Ex):} The tarrasque's bite threatens a critical hit 
on a natural attack roll of 18-20, dealing triple damage on a successful critical 
hit.

\textbf{Frightful Presence (Su):} The tarrasque can inspire terror by charging 
or attacking. Affected creatures must succeed on a DC 36 Will save or become shaken, 
remaining in that condition as long as they remain with 60 feet of the tarrasque. 
The save DC is Charisma-based.

\textbf{Improved Grab (Ex):} To use this ability, the tarrasque must hit a Huge 
or smaller opponent with its bite attack. It can then attempt to start a grapple 
as a free action without provoking an attack of opportunity. If it wins the grapple 
check, it establishes a hold and can try to swallow the foe the following round.

\textbf{Rush (Ex):} Once per minute, the normally slow-moving tarrasque can move 
at a speed of 150 feet. 

\textbf{Swallow Whole (Ex):} The tarrasque can try to swallow a grabbed opponent 
of Huge or smaller size by making a successful grapple check. Once inside, the 
opponent takes 2d8+8 points of crushing damage plus 2d8+6 points of acid damage 
per round from the tarrasque's digestive juices. A swallowed creature can cut its 
way out by dealing 50 points of damage to the tarrasque's digestive tract (AC 25). 
Once the creature exits, muscular action closes the hole; another swallowed opponent 
must cut its own way out. The tarrasque's gullet can hold 2 Huge, 8 Large, 32 Medium, 
128 Small, or 512 Tiny or smaller creatures. 

\textbf{Carapace (Ex):} The tarrasque's armorlike carapace is exceptionally tough 
and highly reflective, deflecting all rays, lines, cones, and even \textit{magic 
missil}e spells. There is a 30\% chance of reflecting any such effect back at the 
caster; otherwise, it is merely negated. Check for reflection before rolling to 
overcome the creature's spell resistance.

\textbf{Regeneration (Ex): }No form of attack deals lethal damage to the tarrasque. 
The tarrasque regenerates even if it fails a saving throw against a \textit{disintegrate 
}spell or a death effect. If the tarrasque fails its save against a spell or effect 
that would kill it instantly (such as those mentioned above), the spell or effect 
instead deals nonlethal damage equal to the creature's full normal hit points +10 
(or 868 hp). The tarrasque is immune to effects that produce incurable or bleeding 
wounds, such as mummy rot, a sword with the wounding special ability, or a clay 
golem's cursed wound ability.

The tarrasque can be slain only by raising its nonlethal damage total to its full 
normal hit points +10 (or 868 hit points) and using a \textit{wish }or \textit{miracle 
}spell to keep it dead.

If the tarrasque loses a limb or body part, the lost portion regrows in 1d6 minutes 
(the detached piece dies and decays normally). The creature can reattach the severed 
member instantly by holding it to the stump.

\textbf{Skills:} The tarrasque has a +8 racial bonus on Listen and Spot checks.

\vspace{12pt}
{\LARGE{}TENDRICULOS}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{226pt}|}
\hline
  & Huge Plant\tabularnewline
\hline
H\textbf{it Dice:} & 9d8+54 (94 hp)\tabularnewline
\hline
I\textbf{nitiative:}- & 1\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 16 (-2 size, -1 Dex, +9 natural), touch 7, flat-footed 
16\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +6/+23\tabularnewline
\hline
A\textbf{ttack:} & Bite +13 melee (2d8+9) and 2 tendrils +8 melee (1d6+4)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +13 melee (2d8+9) and 2 tendrils +8 melee (1d6+4)\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./15 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Improved grab, paralysis, swallow whole\tabularnewline
\hline
S\textbf{pecial Qualities:} & Low-light vision, plant traits, regeneration 10\tabularnewline
\hline
S\textbf{aves:} & Fort +12, Ref +2, Will +4\tabularnewline
\hline
A\textbf{bilities:} & Str 28, Dex 9, Con 22, Int 3, Wis 8, Cha 3\tabularnewline
\hline
S\textbf{kills:} & Hide +9, Listen +1, Move Silently +1, Spot +1\tabularnewline
\hline
F\textbf{eats:} & Alertness, Iron Will, Power Attack, Stealthy\tabularnewline
\hline
E\textbf{nvironment:} & Temperate forests\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 6\tabularnewline
\hline
T\textbf{reasure:} & 1/10th coins; 50\% goods; 50\% items\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & 10-16 HD (Huge); 17-27 HD (Gargantuan)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A tendriculos can rear up to a height of 15 feet. It weighs about 3,500 pounds.

Animals and other plant creatures are unnerved by the presence of a tendriculos; 
they avoid it and any place it has been within the last 24 hours.

COMBAT

Prowling deep forests or waiting in vegetated areas (looking like nothing more 
than a small hillock), a tendriculos attacks savagely, showing no fear. It attempts 
to swallow as much flesh as it can, as quickly as it can.

\textbf{Improved Grab (Ex): }To use this ability, a tendriculos must hit a creature 
at least one size smaller than itself with its bite attack. It can then attempt 
to start a grapple as a free action without provoking an attack of opportunity. 
If it wins the grapple check, it establishes a hold and can try to swallow the 
opponent in the following round. A tendriculos can also use its improved grab ability 
on a tendril attack. If it wins the grapple check, it establishes a hold, picks 
up the opponent, and transfers it to the mouth as a free action, automatically 
dealing bite damage.

\textbf{Swallow Whole/Paralysis (Ex): }A tendriculos can try to swallow a grabbed 
opponent by making a successful grapple check. Once inside the plant's mass, the 
opponent must succeed on a DC 20 Fortitude save or be paralyzed for 3d6 rounds 
by the tendriculos's digestive juices, taking 2d6 points of acid damage per round. 
A new save is required each round inside the plant. The save DC is Constitution-based. 
A swallowed creature that avoids paralysis can climb out of the mass with a successful 
grapple check. This returns it to the plant's maw, where another successful grapple 
check is needed to get free. A swallowed creature can also cut its way out by using 
a light slashing or piercing weapon to deal 25 points of damage to the tendriculos's 
interior (AC 14). Once the creature exits, the plant's regenerative capacity closes 
the hole; another swallowed opponent must cut its own way out. A Huge tendriculos's 
interior can hold 2 Large, 8 Medium, 32 Small, 128 Tiny, or 512 Diminutive or smaller 
opponents. 

\textbf{Regeneration (Ex): }Bludgeoning weapons and acid deal normal damage to 
a tendriculos. A tendriculos that loses part of its body mass can regrow it in 
1d6 minutes. Holding the severed portion against the mass enables it to reattach 
instantly. 

\vspace{12pt}
{\LARGE{}THOQQUA}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{213pt}|}
\hline
  & Medium Elemental (Earth, Extraplanar, Fire)\tabularnewline
\hline
H\textbf{it Dice:} & 3d8+3 (16 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares), burrow 20 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 18 (+1 Dex, +7 natural), touch 11, flat-footed 17\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +2/+4\tabularnewline
\hline
A\textbf{ttack:} & Slam +4 melee (1d6+3 plus 2d6 fire)\tabularnewline
\hline
F\textbf{ull Attack:} & Slam +4 melee (1d6+3 plus 2d6 fire)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Heat, burn\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., elemental traits, immunity to 
fire, tremorsense 60 ft., vulnerability to cold\tabularnewline
\hline
S\textbf{aves:} & Fort +4, Ref +2, Will +2\tabularnewline
\hline
A\textbf{bilities:} & Str 15, Dex 13, Con 13, Int 6, Wis 12, Cha 10\tabularnewline
\hline
S\textbf{kills:} & Listen +5, Move Silently +3, Survival +3\tabularnewline
\hline
F\textbf{eats:} & Alertness, Track\tabularnewline
\hline
E\textbf{nvironment:} & Elemental Plane of Fire\tabularnewline
\hline
O\textbf{rganization:} & Solitary or pair\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Usually neutral\tabularnewline
\hline
A\textbf{dvancement:} & 4-9 HD (Large)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A thoqqua is about 1 foot in diameter and 4 to 5 feet long. It weighs about 200 
pounds.

COMBAT

When a thoqqua is disturbed, its first instinct is to attack. Its favored tactic 
is to spring directly at a foe, either by bursting out of the rock or by coiling 
up its body and launching itself like a spring. (Treat this as a charge, even though 
the thoqqua does not need to move 10 feet before attacking.)

\textbf{Heat (Ex): }Merely touching or being touched by a thoqqua automatically 
deals 2d6 fire damage.

\textbf{Burn (Ex):} When a thoqqua hits with its slam attack, the opponent must 
succeed on a DC 12 Reflex save or catch fire. The save DC is Constitution-based. 
The flame burns for 1d4 rounds if not extinguished sooner. A burning creature can 
use a full-round action to put out the flame.

\vspace{12pt}
{\LARGE{}TITAN}

\begin{tabular}{|>{\raggedright}p{73pt}|>{\raggedright}p{252pt}|}
\hline
  & Huge Outsider (Chaotic, Extraplanar)\tabularnewline
\hline
H\textbf{it Dice:} & 20d8+280 (370 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 40 ft. in half-plate armor (8 squares); base speed 60 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 38 (-2 size, +19 natural, +11 \textit{+4 half-plate armor}), 
touch 8, flat-footed 38\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +20/+44\tabularnewline
\hline
A\textbf{ttack:} & Gargantuan \textit{+3 adamantine warhammer }+37 melee (4d6+27/x3) 
or \textit{+3 javelin }+22 ranged (2d6+19) or slam +34 (1d8+16)\tabularnewline
\hline
F\textbf{ull Attack:} & Gargantuan \textit{+3 adamantine warhammer }+37/+32/+27/+22 
melee (4d6+27/3) or \textit{+3 javelin }+22 ranged (2d6+19) or 2 slams +34 (1d8+16)\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./15 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Oversized weapon, spell-like abilities\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 15/lawful, darkvision 60 ft., spell 
resistance 32\tabularnewline
\hline
S\textbf{aves:} & Fort +26, Ref +13, Will +21\tabularnewline
\hline
A\textbf{bilities:} & Str 43, Dex 12, Con 39, Int 21, Wis 28, Cha 24\tabularnewline
\hline
S\textbf{kills:} & Balance +7, Bluff +19, Climb +22, Concentration +37, Craft (any 
one) +28, Diplomacy +11, Disguise +7 (+9 acting), Heal +20, Intimidate +32, Jump 
+38, Knowledge (any one) +28, Listen +32, Perform (oratory) +30, Sense Motive +32, 
Search +28, Spellcraft +17, Spot +32, Survival +9 (+11 following tracks), Swim 
+16\tabularnewline
\hline
F\textbf{eats:} & Awesome Blow, Blind-Fight, Cleave, Improved Bull Rush, Improved 
Sunder, Power Attack, Quicken Spell-Like Ability (\textit{chain lightning})\tabularnewline
\hline
E\textbf{nvironment:} & A chaotic good-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary or pair\tabularnewline
\hline
C\textbf{hallenge Rating:} & 21\tabularnewline
\hline
T\textbf{reasure:} & Double standard plus \textit{+4 half-plate armor }and Gargantuan 
\textit{+3 adamantine warhammer}\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic (any)\tabularnewline
\hline
A\textbf{dvancement:} & 21-30 HD (Huge); 31-60 HD (Gargantuan)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A titan is about 25 feet tall and weighs about 14,000 pounds. 

Titans speak Abyssal, Common, Celestial, Draconic, and Giant.

COMBAT

A titan's natural weapons, as well as any weapons it wields, are treated as chaotic-aligned 
for the purpose of overcoming damage reduction. 

\textbf{Oversized Weapon (Ex):} A titan wields a great, two-handed warhammer (big 
enough for Gargantuan creatures) without penalty.

\textbf{Spell-Like Abilities: }At will---\textit{chain lightning }(DC 23), \textit{charm 
monster }(DC 21), \textit{cure critical wounds }(DC 21), \textit{fire storm }(DC 
24), \textit{greater dispel magic, hold monster }(DC 22), \textit{invisibility, 
invisibility purge, levitate, persistent image }(DC 22), \textit{polymorph }(humanoid 
forms only, duration 1 hour); 3/day---\textit{etherealness, word of chaos }(DC 
22), \textit{summon nature's ally IX}; 1/day---\textit{gate, maze, meteor swarm 
}(DC 26). Caster level 20th. The save DCs are Charisma-based. 

In addition, titans of good or neutral alignment can use the following additional 
spell-like abilities: At will---\textit{daylight, holy smite }(DC 21), \textit{remove 
curse }(DC 21); 1/day---\textit{greater restoration}. Caster level 20th. The save 
DCs are Charisma-based.

Titans of evil alignment can use the following additional spell-like abilities: 
At will---\textit{bestow curse }(DC 21), \textit{deeper darkness, unholy blight 
}(DC 21); 1/day---\textit{crushing hand }(DC 26). Caster level 20th. The save DCs 
are Charisma-based.

TACTICS ROUND-BY-ROUND

Titans enjoy combat and usually close with their foes. If that proves ineffective, 
they swiftly back off and pelt the foe with spell-like abilities and magical effects. 
Because of a titan's Quicken Spell-Like Ability feat, it can use \textit{chain 
lightning }as a free action, and frequently attacks in melee while lashing out 
with this ability at the same time.

\textit{Prior to combat: Invisibility purge }or \textit{invisibility}.

\textit{Round 1: }Charge and attempt to sunder the weapon of the most dangerous 
foe. Hurl \textit{chain lightning }at opponents standing away from the fight.

\textit{Round 2: }Full attack against the disarmed opponent, and hurl \textit{chain 
lightning }at other opponents.

\textit{Round 3: }Back away from first opponent and use \textit{maze }or \textit{meteor 
swarm }on any spellcaster causing trouble.

\textit{Round 4: }Sunder the weapon of the next most effective combatant, or use 
\textit{greater dispel magic }on all nearby opponents.

\textit{Round 5: }Full attack against any nearby opponent, or use \textit{fire 
storm}.

Use another quickened \textit{chain lightning }if foes seem really dangerous.

A titan usually reserves its \textit{gate }and \textit{etherealness }abilities 
to escape a fight that is not going well.

\vspace{12pt}
{\LARGE{}TOJANIDA}

\begin{tabular}{|>{\raggedright}p{47pt}|>{\raggedright}p{84pt}|>{\raggedright}p{84pt}|>{\raggedright}p{84pt}|}
\hline
  & J\textbf{uvenile Tojanida} & A\textbf{dult Tojanida} & E\textbf{lder Tojanida}\tabularnewline
\hline
  & Small Outsider (Extraplanar, Water) & Medium Outsider (Extraplanar, Water) & Large 
Outsider (Extraplanar, Water)\tabularnewline
\hline
H\textbf{it Dice:} & 3d8+6 (19 hp) & 7d8+14 (45 hp) & 15d8+60 (127 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1 & +1 & +1\tabularnewline
\hline
S\textbf{peed:} & 10 ft. (2 squares), swim 90 ft. & 10 ft. (2 squares), swim 90 
ft. & 10 ft. (2 squares), swim 90 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 22 (+1 size, +1 Dex, +10 natural), touch 12, flat-footed 
21 & 23 (+1 Dex, +12 natural), touch 11, flat-footed 22 & 24 (-1 size, +1 Dex, 
+14 natural), touch 10, flat-footed 23\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +3/+1 & +7/+10 & +15/+25\tabularnewline
\hline
A\textbf{ttack:} & Bite +6 melee (2d6+2) & Bite +10 melee (2d8+3) & Bite +20 melee 
(4d6+6)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +6 melee (2d6+2) and 2 claws +1 melee (1d4+1) & Bite 
+10 melee (2d8+3) and 2 claws +5 melee (1d6+1) & Bite +20 melee (4d6+6) and 2 claws 
+15 melee (1d8+3)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. & 5 ft./5 ft. & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Improved grab, ink cloud & Improved grab, ink cloud & Improved 
grab, ink cloud\tabularnewline
\hline
S\textbf{pecial Qualities:} & All-around vision, darkvision 60 ft., immunity to 
acid and cold, resistance to & All-around vision, darkvision 60 ft., immunity to 
acid and cold, electricity 10 and fire 10 & All-around vision, darkvision 60 ft., 
resistance to electricity 10 and fire 10 immunity to acid and cold, resistance 
to electricity 10 and fire 10\tabularnewline
\hline
S\textbf{aves:} & Fort +5, Ref +4, Will +4 & Fort +7, Ref +6, Will +6 & Fort +13, 
Ref +10, Will +10\tabularnewline
\hline
A\textbf{bilities:} & Str 14, Dex 13, Con 15, Int 10, Wis 12, Cha 9 & Str 16, Dex 
13, Con 15, Int 10, Wis 12, Cha 9 & Str 22, Dex 13, Con 19, Int 10, Wis 12, Cha 
9\tabularnewline
\hline
S\textbf{kills:} & Diplomacy +1, Escape Artist +7, Hide +11, Knowledge (the planes) 
+6, Listen +7, Search +6, Spot +9, Sense Motive +7, Survival +1 (+3 other planes 
and following tracks), Swim +10, Use Rope +1 (+3 with bindings) & Diplomacy +1, 
Escape Artist +11, Hide +11, Knowledge (the planes) +6, Listen +11, Search +14, 
Sense Motive +11, Spot +15, Survival +1 (+3 other planes and following tracks), 
Swim +11, Use Rope+1 (+3 with bindings) & Escape Artist +19, Hide +15, Intimidate 
+17, Knowledge (the Planes) +18, Listen +21, Search +22, Sense Motive +17, Spot 
+25, Survival +1 (+3 other planes and following tracks), Swim +14, Use Rope+1 (+3 
with bindings)\tabularnewline
\hline
F\textbf{eats:} & Blind-Fight, Dodge & Blind-Fight, Dodge, Power Attack & Alertness, 
Blind-Fight, Cleave, Dodge, Improved Sunder, Power Attack\tabularnewline
\hline
E\textbf{nvironment:} & Elemental Plane of Water & Elemental Plane of Water & Elemental 
Plane of Water\tabularnewline
\hline
O\textbf{rganization:} & Solitary or clutch (2-4) & Solitary or clutch (2-4) & Solitary 
or clutch (2-4)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 3 & 5 & 9\tabularnewline
\hline
T\textbf{reasure:} & Standard & Standard & Standard\tabularnewline
\hline
A\textbf{lignment:} & Always neutral & Always neutral & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & 4-6 HD (Small) & 8-14 HD (Medium) & 16-24 HD (Large); 25-45 
HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & --- & --- & \tabularnewline
\hline
\end{tabular}

A tojanida's shell is blue-green in color. Inside the shell is a fleshy body from 
which extend seven stalks. Four of these stalks have paddles for locomotion, two 
are tipped with claws, and one bears the creature's head. Eight vents in the shell, 
four at each end, allow the tojanida to thrust its stalks out in whatever configuration 
it finds convenient.

A juvenile tojanida is up to 25 years old. It has a shell about 3 feet long, and 
it weighs about 60 pounds. An adult is aged 26 to 80. Its shell is about 6 feet 
long, and it weighs about 220 pounds. An elder can reach 150 years of age. It has 
a shell about 9 feet long, and it weighs about 500 pounds. 

Tojanidas speak Aquan and can be loquacious, but usually only on the subject of 
food.

COMBAT

\textbf{Improved Grab (Ex): }To use this ability, a tojanida must hit with a bite 
or claw attack. It can then attempt to start a grapple as a free action without 
provoking an attack of opportunity. Underwater, a tojanida can tow a grabbed victim 
of its own size or smaller at its swim speed (but it cannot run). A favorite tactic 
is to grab a single opponent, then withdraw, hauling the opponent away from its 
allies.

\textbf{Ink Cloud (Ex):} A tojanida can emit a spherical cloud of jet-black ink 
with a radius of 30 feet once per minute as a free action. The effect is otherwise 
similar to \textit{fog cloud }cast by an individual of a level equal to the tojanida's 
Hit Dice. Out of water, the ink emerges in a stream up to 30 feet long, which a 
tojanida can squirt into an opponent's eyes. The affected creature must succeed 
on a Reflex save or be blinded for 1 round. The save DC is 13 against a juvenile, 
15 against an adult, and 21 against an elder. The save DCs are Constitution based.

\textbf{All-Around Vision (Ex): }The multiple apertures in a tojanida's shell allow 
it to look in any direction, bestowing a +4 racial bonus on Spot and Search checks. 
Opponents gain no flanking bonuses when attacking a tojanida.

\textbf{Skills:} A tojanida has a +8 racial bonus on any Swim check to perform 
some special action or avoid a hazard. It can always choose to take 10 on a Swim 
check, even if distracted or endangered. It can use the run action while swimming, 
provided it swims in a straight line.

\vspace{12pt}
{\LARGE{}TREANT}

\begin{tabular}{|>{\raggedright}p{88pt}|>{\raggedright}p{237pt}|}
\hline
  & Huge Plant\tabularnewline
\hline
H\textbf{it Dice:} & 7d8+35 (66 hp)\tabularnewline
\hline
I\textbf{nitiative:}- & 1\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 20 (-2 size, -1 Dex, +13 natural), touch 7, flat-footed 
20\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +5/+22\tabularnewline
\hline
A\textbf{ttack:} & Slam +12 melee (2d6+9)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 slams +12 melee (2d6+9)\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./15 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Animate trees, double damage against objects, trample 
2d6+13\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 10/slashing, low-light vision, plant 
traits, vulnerability to fire\tabularnewline
\hline
S\textbf{aves:} & Fort +10, Ref +1, Will +7\tabularnewline
\hline
A\textbf{bilities:} & Str 29, Dex 8, Con 21, Int 12, Wis 16, Cha 12\tabularnewline
\hline
S\textbf{kills:} & Diplomacy +3, Hide -9*, Intimidate +6, Knowledge (nature) +6, 
Listen +8, Sense Motive +8, Spot +8, Survival +8 (+10 aboveground)\tabularnewline
\hline
F\textbf{eats:} & Improved Sunder, Iron Will, Power Attack\tabularnewline
\hline
E\textbf{nvironment:} & Temperate forests\tabularnewline
\hline
O\textbf{rganization:} & Solitary or grove (4-7)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 8\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Usually neutral good\tabularnewline
\hline
A\textbf{dvancement:} & 8-16 HD (Huge); 17-21 HD (Gargantuan)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +5\tabularnewline
\hline
\end{tabular}

A treant's leaves are deep green in the spring and summer. In the fall and winter 
the leaves change to yellow, orange, or red, but they rarely fall out. A treant's 
legs fit together when closed to look like the trunk of a tree, and a motionless 
treant is nearly indistinguishable from a tree.

A treant is about 30 feet tall, with a ``trunk'' about 2 feet in diameter. It weighs 
about 4,500 pounds.

Treants speak their own language, plus Common and Sylvan. Most also can manage 
a smattering of just about all other humanoid tongues---at least enough to say 
``Get away from my trees!''

COMBAT

Treants prefer to watch potential foes carefully before attacking. They often charge 
suddenly from cover to trample the despoilers of forests. If sorely pressed, they 
animate trees as reinforcements.

\textit{\textbf{Animate Trees }}\textbf{(Sp): }A treant can animate trees within 
180 feet at will, controlling up to two trees at a time. It takes 1 full round 
for a normal tree to uproot itself. Thereafter it moves at a speed of 10 feet and 
fights as a treant in all respects. Animated trees lose their ability to move if 
the treant that animated them is incapacitated or moves out of range. The ability 
is otherwise similar to \textit{liveoak }(caster level 12th). Animated trees have 
the same vulnerability to fire that a treant has.

\textbf{Double Damage against Objects (Ex): }A treant or animated tree that makes 
a full attack against an object or structure deals double damage.

\textbf{Trample (Ex):} Reflex DC 22 half. The save DC is Strength-based. Skills: 
*Treants have a +16 racial bonus on Hide checks made in forested areas.

\vspace{12pt}
{\LARGE{}TRITON}

\begin{tabular}{|>{\raggedright}p{79pt}|>{\raggedright}p{246pt}|}
\hline
  & Medium Outsider (Native, Water)\tabularnewline
\hline
H\textbf{it Dice:} & 3d8+3 (16 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +0\tabularnewline
\hline
S\textbf{peed:} & 5 ft. (1 square), swim 40 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 16 (+6 natural), touch 10, flat-footed 16\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +3/+4\tabularnewline
\hline
A\textbf{ttack:} & Trident +4 melee (1d8+1) or heavy crossbow +3 ranged (1d10/19-20)\tabularnewline
\hline
F\textbf{ull Attack:} & Trident +4 melee (1d8+1) or heavy crossbow +3 ranged (1d10/19-20)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Spell-like abilities\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +4, Ref +3, Will +4\tabularnewline
\hline
A\textbf{bilities:} & Str 12, Dex 10, Con 12, Int 13, Wis 13, Cha 11\tabularnewline
\hline
S\textbf{kills:} & Craft (any one) +7, Diplomacy +2, Hide +6, Listen +7, Move Silently 
+6, Ride +6, Search +7, Sense Motive +7, Spot +7, Survival +7 (+9 following tracks), 
Swim +9\tabularnewline
\hline
F\textbf{eats:} & Mounted Combat, Ride-By Attack\tabularnewline
\hline
E\textbf{nvironment:} & Temperate aquatic\tabularnewline
\hline
O\textbf{rganization:} & Company (2-5), squad (6-11), or band (20-80)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Usually neutral good\tabularnewline
\hline
A\textbf{dvancement:} & 4-9 HD (Medium) \tabularnewline
\hline
L\textbf{evel Adjustment:} & +2\tabularnewline
\hline
\end{tabular}

A triton has silvery skin that fades into silver-blue scales on the lower half 
of its body. A triton's hair is deep blue or blue-green.

A triton is about the same size and weight as a human. Tritons speak Common and 
Aquan.

COMBAT

The reclusive tritons prefer to avoid combat, but they fiercely defend their homes. 
They attack with either melee or ranged weapons as the circumstances warrant. When 
encountered outside their lair, they are 90\% likely to be mounted on friendly 
sea creatures such as porpoises.

\textbf{Spell-Like Abilities: }1/day---\textit{summon nature's ally IV. }Caster 
level 7th. Tritons often choose water elementals for their companions.

\textbf{Skills:} A triton has a +8 racial bonus on any Swim check to perform some 
special action or avoid a hazard. It can always choose to take 10 on a Swim check, 
even if distracted or endangered. It can use the run action while swimming, provided 
it swims in a straight line.

\vspace{12pt}
{\LARGE{}TROGLODYTE}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{203pt}|}
\hline
  & Medium Humanoid (Reptilian)\tabularnewline
\hline
H\textbf{it Dice:} & 2d8+4 (13 hp)\tabularnewline
\hline
I\textbf{nitiative:}- & 1\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 15 (-1 Dex, +6 natural), touch 9, flat-footed 15\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +1/+1\tabularnewline
\hline
A\textbf{ttack:} & Club +1 melee (1d6) or claw +1 melee (1d4) or javelin +1 ranged 
(1d6)\tabularnewline
\hline
F\textbf{ull Attack:} & Club +1 melee (1d6) and claw -1 melee (1d4) and bite -1 
melee (1d4); or 2 claws +1 melee (1d4) and bite -1 melee (1d4); or javelin +1 ranged 
(1d6)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Stench\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 90 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +5, Ref -1, Will +0\tabularnewline
\hline
A\textbf{bilities:} & Str 10, Dex 9, Con 14, Int 8, Wis 10, Cha 10\tabularnewline
\hline
S\textbf{kills:} & Hide +5*, Listen +3\tabularnewline
\hline
F\textbf{eats:} & Multiattack\textsuperscript{B}, Weapon Focus (javelin)\tabularnewline
\hline
E\textbf{nvironment:} & Underground\tabularnewline
\hline
O\textbf{rganization:} & Clutch (2-5), squad (6-11 plus 1-2 monitor lizards), or 
band (20-80 plus 20\% noncombatants plus 3-13 monitor lizards)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 1\tabularnewline
\hline
T\textbf{reasure:} & 50\% coins; 50\% goods; 50\% items\tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & By character class\tabularnewline
\hline
L\textbf{evel Adjustment:} & +2\tabularnewline
\hline
\end{tabular}

A troglodyte stands about 5 feet tall and weighs about 150 pounds.

Troglodytes speak Draconic.

COMBAT

Half of a group of troglodytes are armed only with claws and teeth; the rest carry 
one or two javelins and clubs. They normally conceal themselves, launch a volley 
of javelins, then close to attack. If the battle goes against them, they retreat 
and attempt to hide.

Stench (Ex): When a troglodyte is angry or frightened, it secretes an oily, musk-like 
chemical that nearly every form of animal life finds offensive. All living creatures 
(except troglodytes) within 30 feet of a troglodyte must succeed on a DC 13 Fortitude 
save or be sickened for 10 rounds. The save DC is Constitution-based. Creatures 
that successfully save cannot be affected by the same troglodyte's stench for 24 
hours. A \textit{delay poison }or \textit{neutralize poison }spell removes the 
effect from the sickened creature. Creatures with immunity to poison are unaffected, 
and creatures resistant to poison receive their normal bonus on their saving throws.

Skills: The skin of a troglodyte changes color somewhat, allowing it to blend in 
with its surroundings like a chameleon and providing a +4 racial bonus on Hide 
checks. *In rocky or underground settings, this bonus improves to +8.

TROGLODYTE CHARACTERS

Troglodyte characters possess the following racial traits.--- -

\parindent=3pt
2 Dexterity, +4 Constitution, -2 Intelligence.---

\parindent=0pt
Medium size.---

A troglodyte's base land speed is 30 feet.---

Darkvision out to 90 feet.---

Racial Hit Dice: A troglodyte begins with two levels of humanoid, which provide 
2d8 Hit Dice, a base attack bonus of +1, and base saving throw bonuses of Fort 
+3, Ref +0, and Will +0. ---

Racial Skills: A troglodyte's humanoid levels give it skill points equal to 5 x$ $(2 
+ Int modifier, minimum 1). Its class skills are Hide and Listen. Troglodytes have 
a +4 racial bonus on Hide checks (+8 in rocky or underground surroundings).---

Racial Feats: A troglodyte's humanoid levels give it one feat. A troglodyte receives 
Multiattack as a bonus feat.--- 

\parindent=3pt
+6 natural armor bonus.---

\parindent=0pt
Natural Weapons: 2 claws (1d4) and bite (1d4).---

Special Attacks (see above): Stench.---

Automatic Languages: Draconic. Bonus Languages: Common, Giant, Goblin, Orc.---

Favored Class: Cleric.---

Level adjustment +2.

\vspace{12pt}
{\LARGE{}TROLL}

\begin{tabular}{|>{\raggedright}p{52pt}|>{\raggedright}p{131pt}|>{\raggedright}p{131pt}|}
\hline
 & T\textbf{roll} & T\textbf{roll Hunter, 6th-Level Ranger}\tabularnewline
\hline
 & Large Giant & Large Giant\tabularnewline
\hline
H\textbf{it Dice:} & 6d8+36 (63 hp) & 6d8+6d8+72 (130 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +2 & +1\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares) & 30 ft. (6 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 16 (-1 size, +2 Dex, +5 natural), touch 11, flat-footed 
14 & 21 (-1 size, +1 Dex, +6 natural, +5 \textit{+1 chain shirt}), touch 10, flat-footed 
20\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +4/+14 & +10/+21\tabularnewline
\hline
A\textbf{ttack:} & Claw +9 melee (1d6+6) & Claw +16 melee (1d6+7) or \textit{+1 
battleaxe }+17 melee (2d6+8/\ensuremath{\times}3) or javelin +10 ranged (1d8+7)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +9 melee (1d6+6) and bite +4 melee (1d6+3) & 2 
claws +16 melee (1d6+7) and bite +11 melee (1d6+3); or \textit{+1 battleaxe }+17/+12 
melee (2d6+8/\ensuremath{\times}3) and claw +12 melee (1d6+3) and bite +12 (1d6+3); 
or javelin +10 ranged (1d8+7)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft. & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Rend 2d6+9  & Rend 2d6+10, spells\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 90 ft., low-light vision, regeneration 
5, scent & Darkvision 90 ft., favored enemy elves (+4), favored enemy humans (+2), 
low-light vision, regeneration 5, scent, wild empathy\tabularnewline
\hline
S\textbf{aves:} & Fort +11, Ref +4, Will +3 & Fort +16, Ref +8, Will +8\tabularnewline
\hline
A\textbf{bilities:} & Str 23, Dex 14, Con 23, Int 6, Wis 9, Cha 6 & Str 25, Dex 
12, Con 22, Int 10, Wis 15, Cha 10\tabularnewline
\hline
S\textbf{kills:} & Listen +5, Spot +6 & Knowledge (nature) +6, Knowledge (dungeoneering) 
+6, Listen +13, Move Silently +9, Search +6, Spot +13, Survival +11 (+13 above- 
or underground and following tracks)\tabularnewline
\hline
F\textbf{eats:} & Alertness, Iron Will, Track & Alertness, Cleave, Endurance\textsuperscript{B}, 
Improved Natural Armor, Improved Two-Weapon Fighting\textsuperscript{B}, Iron Will, 
Power Attack, Track\textsuperscript{B}, Two-Weapon Fighting\textsuperscript{B}\tabularnewline
\hline
E\textbf{nvironment:} & Cold mountains \linebreak{}
(Scrag: Cold aquatic) & Cold mountains\tabularnewline
\hline
O\textbf{rganization:} & Solitary or gang (2-4) & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 5 & 11\tabularnewline
\hline
T\textbf{reasure:} & Standard & Standard\tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic evil & Usually chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & By character class--- & \tabularnewline
\hline
L\textbf{evel Adjustment:} & +5 & +5\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
Trolls walk upright but hunched forward with sagging shoulders. Their gait is uneven, 
and when they run, their arms dangle and drag along the ground. For all this seeming 
awkwardness, trolls are very agile.

A typical adult troll stands 9 feet tall and weighs 500 pounds. Females are slightly 
larger than males. A troll's rubbery hide is moss green, mottled green and gray, 
or putrid gray. The hair is usually greenish black or iron gray.

Trolls speak Giant.

COMBAT

Trolls have no fear of death: They launch themselves into combat without hesitation, 
flailing wildly at the closest opponent. Even when confronted with fire, they try 
to get around the flames and attack.

\textbf{Rend (Ex): }If a troll hits with both claw attacks, it latches onto the 
opponent's body and tears the flesh. This attack automatically deals an additional 
2d6+9 points of damage.

\textbf{Regeneration (Ex): }Fire and acid deal normal damage to a troll. If a troll 
loses a limb or body part, the lost portion regrows in 3d6 minutes. The creature 
can reattach the severed member instantly by holding it to the stump.

\vspace{12pt}
SCRAG

These cousins of the troll have the aquatic subtype. They dwell in any body of 
water in any climate. They have a base land speed of 20 feet and a swim speed of 
40 feet and are found only in aquatic environments. They regenerate only if mostly 
immersed in water. 

\vspace{12pt}
TROLL HUNTER

Some trolls, more cunning than most, are not satisfied with merely eating civilized 
beings but train to hunt them relentlessly. These troll hunters are fearsome rangers 
who focus on slaying and devouring humanoid prey.

COMBAT

A troll hunter makes full use of its scent ability to track its favored enemies 
and generally prefers to hunt in darkness.

The troll hunter uses its limited repertoire of spells to protect itself from damaging 
forms of energy and to immobilize enemies.

\textit{Typical Ranger Spells Prepared }(2; save DC 12 + spell level): 1st---\textit{entangle, 
resist energy.}

\vspace{12pt}
TROLLS AS CHARACTERS

Troll characters possess the following racial traits.--- 

\parindent=3pt
+12 Strength, +4 Dexterity, +12 Constitution, -4 Intelligence (minimum 3), -2 Wisdom, 
-4 Charisma.---

\parindent=0pt
Large size. -1 penalty to Armor Class, -1 penalty on attack rolls, -4 penalty on 
Hide checks, +4 bonus on grapple checks, lifting and carrying limits double those 
of Medium characters.---

Space/Reach: 10 feet/10 feet.---

A troll's base land speed is 30 feet.---

Darkvision out to 60 feet and low-light vision.---

Racial Hit Dice: A troll begins with six levels of giant, which provide 6d8 Hit 
Dice, a base attack bonus of +4, and base saving throw bonuses of Fort +5, Ref 
+2, and Will +2.---

Racial Skills: A troll's giant levels give it skill points equal to 9 x$ $(2 + 
Int modifier, minimum 1). Its class skills are Listen and Spot.---

Racial Feats: A troll's giant levels give it three feats.--- 

\parindent=3pt
+5 natural armor bonus.---

\parindent=0pt
Natural Weapons: Claw (1d6) and bite (1d6).---

Special Attacks (see above): Rend, damage 2d6 + 1-1/2 times Str modifier.---

Special Qualities: Regeneration 5, scent.---

Automatic Languages: Giant. Bonus Languages: Common, Orc.---

Favored Class: Fighter.---

Level adjustment +5.

\vspace{12pt}
{\LARGE{}UNICORN}

\begin{tabular}{|>{\raggedright}p{47pt}|>{\raggedright}p{133pt}|>{\raggedright}p{133pt}|}
\hline
 & U\textbf{nicorn} & \subsection*{C\textbf{elestial Charger, 7th-Level Cleric}}\tabularnewline
\hline
  & Large Magical Beast & Large Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 4d10+20 (42 hp) & 8d10+7d8+75 (155 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +3 & +4\tabularnewline
\hline
S\textbf{peed:} & 60 ft. (12 squares) & 60 ft. (12 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 18 (-1 size, +3 Dex, +6 natural), touch 12, flat-footed 
15 & 24 (-1 size, +4 Dex, +6 natural, +5 \textit{bracers of armor +5}), touch 13, 
flat-footed 20\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +4/+13 & +13/+24\tabularnewline
\hline
A\textbf{ttack:} & Horn +11 melee (1d8+8) & Horn +22 melee (1d8+10)\tabularnewline
\hline
F\textbf{ull Attack:} & Horn +11 melee (1d8+8) and 2 hooves +3 melee (1d4+2) & Horn 
+22 melee (1d8+10) and 2 hooves +14 melee (1d4+3)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft. & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:}--- &  & Turn undead 13/day, smite evil, spells\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., magic circle against evil, spell-like 
abilities, immunity to poison, charm, and compulsion, low-light vision, scent, 
wild empathy & Damage reduction 10/magic, darkvision 60 ft., immunity to poison, 
charm, and compulsion, low-light vision, magic circle against evil, resistance 
to acid 10, cold 10, and electricity 10, scent, spell-like abilities, spell resistance 
20, wild empathy\tabularnewline
\hline
S\textbf{aves:} & Fort +9, Ref +7, Will +6 & Fort +16, Ref +12, Will +15\tabularnewline
\hline
A\textbf{bilities:} & Str 20, Dex 17, Con 21, Int 10, Wis 21, Cha 24 & Str 24, 
Dex 18, Con 20, Int 13, Wis 27, Cha 22\tabularnewline
\hline
S\textbf{kills:} & Jump +21, Listen +11, Move Silently +9, Spot +11, Survival +8* & Concentration 
+11, Knowledge (nature) +9, Knowledge (religion) +8, Listen +15, Move Silently 
+12, Spellcraft +5, Spot +15, Survival +15 (+17 aboveground)*\tabularnewline
\hline
F\textbf{eats:} & Alertness, Skill Focus (Survival) & Alertness, Combat Casting, 
Extra Turning, Improved Turning, Run, Skill Focus (Survival)\tabularnewline
\hline
E\textbf{nvironment:} & Temperate forests & A chaotic good plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or grace (3-6) & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 3 & 13\tabularnewline
\hline
T\textbf{reasure:} & None & None\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic good & Always chaotic good\tabularnewline
\hline
A\textbf{dvancement:} & 5-8 HD (Large) & By character class\tabularnewline
\hline
L\textbf{evel Adjustment:} & +4 (cohort) & +8 (cohort)\tabularnewline
\hline
\end{tabular}

A unicorn has deep sea-blue, violet, brown, or fiery gold eyes. Males sport a white 
beard.

A typical adult unicorn grows to 8 feet in length, stands 5 feet high at the shoulder, 
and weighs 1,200 pounds. Females are slightly smaller and slimmer than males.

Unicorns speak Sylvan and Common.

COMBAT

Unicorns normally attack only when defending themselves or their forests. They 
either charge, impaling foes with their horns like lances, or strike with their 
hooves. The horn is a +3 magic weapon, though its power fades if removed from the 
unicorn.

\textbf{Magic Circle against Evil (Su): }This ability continuously duplicates the 
effect of the spell. A unicorn cannot suppress this ability.

\textbf{Spell-Like Abilities: }Unicorns can use \textit{detect evil }at will as 
a free action.

Once per day a unicorn can use \textit{greater teleport }to move anywhere within 
its home. It cannot teleport beyond the forest boundaries nor back from outside.

A unicorn can use \textit{cure light wounds }three times per day and \textit{cure 
moderate wounds }once per day (caster level 5th) by touching a wounded creature 
with its horn. Once per day it can use \textit{neutralize poison }(DC 21, caster 
level 8th) with a touch of its horn. The save DC is Charisma-based.

\textbf{Wild Empathy (Ex): }This power works like the druid's wild empathy class 
feature, except that a unicorn has a +6 racial bonus on the check.

\textbf{Skills:} Unicorns have a +4 racial bonus on Move Silently checks. *Unicorns 
have a +3 competence bonus on Survival checks within the boundaries of their forest.

\vspace{12pt}
CELESTIAL CHARGER

The celestial charger described here is an 8 HD celestial unicorn with seven levels 
of cleric.

COMBAT

The save DC for this celestial charger's \textit{neutralize poison }ability (DC 
20) is adjusted for its greater Hit Dice and altered Charisma score. 

A celestial charger's natural weapons are treated as magic weapons for the purpose 
of overcoming damage reduction.

\textbf{Smite Evil (Su): }Once per day a celestial charger can make a normal melee 
attack to deal 15 points of extra damage against an evil foe.

\textit{Cleric Spells Prepared }(6/7/6/5/4; save DC 18 + spell level): 0--- \textit{detect 
magic, detect poison }(2), \textit{light, virtue }(2); 1st---\textit{bless }(2), 
\textit{calm animals*, obscuring mist, remove fear, sanctuary, shield of faith}; 
2nd---\textit{aid}* (2), \textit{animal messenger, lesser restoration, remove paralysis, 
shield other}; 3rd---\textit{prayer, protection from energy, remove curse, searing 
light }(2); 4th---\textit{air walk, divine power, holy smite*, restoration}.

*Domain spell. Domains: Animal and Good.

\vspace{12pt}
{\LARGE{}VAMPIRE}

Vampires appear just as they did in life, although their features are often hardened 
and feral, with the predatory look of wolves.

Like liches, they often embrace finery and decadence and may assume the guise of 
nobility. Despite their human appearance, vampires can be easily recognized, for 
they cast no shadows and throw no reflections in mirrors.

Vampires speak any languages they knew in life.

CREATING A VAMPIRE

``Vampire'' is an acquired template that can be added to any humanoid or monstrous 
humanoid creature (referred to hereafter as the base creature).

A vampire uses all the base creature's statistics and special abilities except 
as noted here.

\textbf{Size and Type: }The creature's type changes to undead (augmented humanoid 
or monstrous humanoid). Do not recalculate base attack bonus, saves, or skill points. 
Size is unchanged.

\textbf{Hit Dice:} Increase all current and future Hit Dice to d12s.

\textbf{Speed: }Same as the base creature. If the base creature has a swim speed, 
the vampire retains the ability to swim and is not vulnerable to immersion in running 
water (see below).

\textbf{Armor Class: }The base creature's natural armor bonus improves by +6.

\textbf{Attack:} A vampire retains all the attacks of the base creature and also 
gains a slam attack if it didn't already have one. If the base creature can use 
weapons, the vampire retains this ability. A creature with natural weapons retains 
those natural weapons. A vampire fighting without weapons uses either its slam 
attack or its primary natural weapon (if it has any). A vampire armed with a weapon 
uses its slam or a weapon, as it desires.

\textbf{Full Attack:} A vampire fighting without weapons uses either its slam attack 
(see above) or its natural weapons (if it has any). If armed with a weapon, it 
usually uses the weapon as its primary attack along with a slam or other natural 
weapon as a natural secondary attack. 

\textbf{Damage:} Vampires have slam attacks. If the base creature does not have 
this attack form, use the appropriate damage value from the table below according 
to the vampire's size. Creatures that have other kinds of natural weapons retain 
their old damage values or use the appropriate value from the table below, whichever 
is better.

\begin{tabular}{|>{\raggedright}p{45pt}|>{\raggedright}p{26pt}|}
\hline
S\textbf{ize} & D{\small{}\textbf{amage}}\tabularnewline
\hline
Fine & 1\tabularnewline
\hline
Diminutive & 1d2\tabularnewline
\hline
Tiny & 1d3\tabularnewline
\hline
Small & 1d4\tabularnewline
\hline
Medium & 1d6\tabularnewline
\hline
Large & 1d8\tabularnewline
\hline
Huge & 2d6\tabularnewline
\hline
Gargantuan & 2d8\tabularnewline
\hline
Colossal & 4d6\tabularnewline
\hline
\end{tabular}

\textbf{Special Attacks:} A vampire retains all the special attacks of the base 
creature and gains those described below. Saves have a DC of 10 + 1/2 vampire's 
HD + vampire's Cha modifier unless noted otherwise.

\textit{\textbf{Blood Drain (Ex):}}\textit{ }A vampire can suck blood from a living 
victim with its fangs by making a successful grapple check. If it pins the foe, 
it drains blood, dealing 1d4 points of Constitution drain each round the pin is 
maintained. On each such successful attack, the vampire gains 5 temporary hit points.

\textit{\textbf{Children of the Night (Su):}}\textit{ }Vampires command the lesser 
creatures of the world and once per day can call forth 1d6+1 rat swarms, 1d4+1 
bat swarms, or a pack of 3d6 wolves as a standard action. (If the base creature 
is not terrestrial, this power might summon other creatures of similar power.) 
These creatures arrive in 2d6 rounds and serve the vampire for up to 1 hour.

\textit{\textbf{Dominate (Su):}}\textit{ }A vampire can crush an opponent's will 
just by looking onto his or her eyes. This is similar to a gaze attack, except 
that the vampire must use a standard action, and those merely looking at it are 
not affected. Anyone the vampire targets must succeed on a Will save or fall instantly 
under the vampire's influence as though by a \textit{dominate person }spell (caster 
level 12th). The ability has a range of 30 feet.

\textit{\textbf{Create Spawn (Su):}}\textit{ }A humanoid or monstrous humanoid 
slain by a vampire's energy drain rises as a vampire spawn (see the Vampire Spawn 
entry) 1d4 days after burial. 

If the vampire instead drains the victim's Constitution to 0 or lower, the victim 
returns as a spawn if it had 4 or less HD and as a vampire if it had 5 or more 
HD. In either case, the new vampire or spawn is under the command of the vampire 
that created it and remains enslaved until its master's destruction. At any given 
time a vampire may have enslaved spawn totaling no more than twice its own Hit 
Dice; any spawn it creates that would exceed this limit are created as free-willed 
vampires or vampire spawn. A vampire that is enslaved may create and enslave spawn 
of its own, so a master vampire can control a number of lesser vampires in this 
fashion. A vampire may voluntarily free an enslaved spawn in order to enslave a 
new spawn, but once freed, a vampire or vampire spawn cannot be enslaved again.

\textit{\textbf{Energy Drain (Su):}}\textit{ }Living creatures hit by a vampire's 
slam attack (or any other natural weapon the vampire might possess) gain two negative 
levels. For each negative level bestowed, the vampire gains 5 temporary hit points. 
A vampire can use its energy drain ability once per round.

\textbf{Special Qualities:} A vampire retains all the special qualities of the 
base creature and gains those described below.

\textit{Alternate Form (Su): }A vampire can assume the shape of a bat, dire bat, 
wolf, or dire wolf as a standard action. This ability is similar to a \textit{polymorph 
}spell cast by a 12th-level character, except that the vampire does not regain 
hit points for changing form and must choose from among the forms mentioned here. 
While in its alternate form, the vampire loses its natural slam attack and dominate 
ability, but it gains the natural weapons and extraordinary special attacks of 
its new form. It can remain in that form until it assumes another or until the 
next sunrise. (If the base creature is not terrestrial, this power might allow 
other forms.)

\textit{Damage Reduction (Su): }A vampire has damage reduction 10/silver and magic. 
A vampire's natural weapons are treated as magic weapons for the purpose of overcoming 
damage reduction.

\textit{Fast Healing (Ex): }A vampire heals 5 points of damage each round so long 
as it has at least 1 hit point. If reduced to 0 hit points in combat, it automatically 
assumes gaseous form and attempts to escape. It must reach its coffin home within 
2 hours or be utterly destroyed. (It can travel up to nine miles in 2 hours.) Any 
additional damage dealt to a vampire forced into gaseous form has no effect. Once 
at rest in its coffin, a vampire is helpless. It regains 1 hit point after 1 hour, 
then is no longer helpless and resumes healing at the rate of 5 hit points per 
round.

\textit{Gaseous Form (Su): }As a standard action, a vampire can assume gaseous 
form at will as the spell (caster level 5th), but it can remain gaseous indefinitely 
and has a fly speed of 20 feet with perfect maneuverability.

\textit{Resistances (Ex): }A vampire has resistance to cold 10 and electricity 
10.

\textit{Spider Climb (Ex): }A vampire can climb sheer surfaces as though with a 
\textit{spider climb }spell.

\textit{Turn Resistance (Ex): }A vampire has +4 turn resistance. 

\textbf{Abilities:} Increase from the base creature as follows: Str +6, Dex +4, 
Int +2, Wis +2, Cha +4. As an undead creature, a vampire has no Constitution score.

\textbf{Skills:} Vampires have a +8 racial bonus on Bluff, Hide, Listen, Move Silently, 
Search, Sense Motive, and Spot checks. Otherwise same as the base creature.

\textbf{Feats: }Vampires gain Alertness, Combat Reflexes, Dodge, Improved Initiative, 
and Lightning Reflexes, assuming the base creature meets the prerequisites and 
doesn't already have these feats.

\textbf{Environment:} Any, usually same as base creature.

\textbf{Organization:} Solitary, pair, gang (3-5), or troupe (1-2 plus 2-5 vampire 
spawn)

\textbf{Challenge Rating:} Same as the base creature +2.

\textbf{Treasure: }Double standard.

\textbf{Alignment: }Always evil (any).

\textbf{Advancement:} By character class.

\textbf{Level Adjustment:} Same as the base creature +8.

\vspace{12pt}
Vampire Weaknesses

For all their power, vampires have a number of weaknesses.

\textbf{Repelling a Vampire:} Vampires cannot tolerate the strong odor of garlic 
and will not enter an area laced with it. Similarly, they recoil from a mirror 
or a strongly presented holy symbol. These things don't harm the vampire---they 
merely keep it at bay. A recoiling vampire must stay at least 5 feet away from 
a creature holding the mirror or holy symbol and cannot touch or make melee attacks 
against the creature holding the item for the rest of the encounter. Holding a 
vampire at bay takes a standard action.

Vampires are also unable to cross running water, although they can be carried over 
it while resting in their coffins or aboard a ship.

They are utterly unable to enter a home or other building unless invited in by 
someone with the authority to do so. They may freely enter public places, since 
these are by definition open to all. 

\textbf{Slaying a Vampire:} Reducing a vampire's hit points to 0 or lower incapacitates 
it but doesn't always destroy it (see the note on fast healing). However, certain 
attacks can slay vampires. Exposing any vampire to direct sunlight disorients it: 
It can take only a single move action or attack action and is destroyed utterly 
in the next round if it cannot escape. Similarly, immersing a vampire in running 
water robs it of one-third of its hit points each round until it is destroyed at 
the end of the third round of immersion. Driving a wooden stake through a vampire's 
heart instantly slays the monster. However, it returns to life if the stake is 
removed, unless the body is destroyed. A popular tactic is to cut off the creature's 
head and fill its mouth with holy wafers (or their equivalent).

\vspace{12pt}
\section*{Vampire Characters}

Vampires are always evil, which causes characters of certain classes to lose some 
class abilities. In addition,  certain classes take additional penalties.

\textbf{Clerics:} Vampire clerics lose their ability to turn undead but gain the 
ability to rebuke undead. This ability does not affect the vampire's controller 
or any other vampires that a master controls. A vampire cleric has access to two 
of the following domains: Chaos, Destruction, Evil, or Trickery.

\textbf{Sorcerers and Wizards:} Vampire sorcerers and wizards retain their class 
abilities, but if a character has a familiar other than a rat or bat, the link 
between them is broken, and the familiar shuns its former companion. The character 
can summon another familiar, but it must be a rat or bat.

\vspace{12pt}
{\LARGE{}VAMPIRE SPAWN}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{213pt}|}
\hline
  & Medium Undead\tabularnewline
\hline
H\textbf{it Dice:} & 4d12+3 (29 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +6\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 15 (+2 Dex, +3 natural), touch 12, flat-footed 13\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +2/+5\tabularnewline
\hline
A\textbf{ttack:} & Slam +5 melee (1d6+4 plus energy drain)\tabularnewline
\hline
F\textbf{ull Attack:} & Slam +5 melee (1d6+4 plus energy drain)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Blood drain, domination, energy drain\tabularnewline
\hline
S\textbf{pecial Qualities:} & +2 turn resistance, damage reduction 5/silver, darkvision 
60 ft., fast healing 2, gaseous form, resistance to cold 10 and electricity 10, 
spider climb, undead traits\tabularnewline
\hline
S\textbf{aves:} & Fort +1, Ref +5, Will +5\tabularnewline
\hline
A\textbf{bilities:} & Str 16, Dex 14, Con ---, Int 13, Wis 13, Cha 14\tabularnewline
\hline
S\textbf{kills:} & Bluff +6, Climb +8, Craft or Profession (any one) +4, Diplomacy 
+4, Hide +10, Jump +8, Listen +11, Move Silently +10, Search +8, Sense Motive +11, 
Spot +11\tabularnewline
\hline
F\textbf{eats:} & Alertness\textsuperscript{B}, Improved Initiative\textsuperscript{B}, 
Lightning Reflexes\textsuperscript{B}, Skill Focus (selected Craft or Profession 
skill), Toughness\tabularnewline
\hline
E\textbf{nvironment:} & Any\tabularnewline
\hline
O\textbf{rganization:} & Solitary or pack (2-5)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 4\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Always evil (any)\tabularnewline
\hline
A\textbf{dvancement:}--- & \tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

Vampire spawn are undead creatures that come into being when vampires slay mortals. 
Like their creators, spawn remain bound to their coffins and to the soil of their 
graves. Vampire spawn appear much as they did in life, although their features 
are often hardened, with a predatory look.

Vampire spawn speak Common.

COMBAT

Vampire spawn use their inhuman strength when engaging mortals, hammering their 
foes with powerful blows and dashing them against rocks or walls. They also use 
their gaseous form and flight abilities to strike where opponents are most vulnerable.

\textbf{Blood Drain (Ex):} A vampire spawn can suck blood from a living victim 
with its fangs by making a successful grapple check. If it pins the foe, it drains 
blood, dealing 1d4 points of Constitution drain each round. On each such successful 
drain attack, the vampire spawn gains 5 temporary hit pints.

\textbf{Domination (Su):} A vampire spawn can crush an opponent's will just by 
looking onto his or her eyes. This is similar to a gaze attack, except that the 
vampire must take a standard action, and those merely looking at it are not affected. 
Anyone the vampire targets must succeed on a DC 14 Will save or fall instantly 
under the vampire's influence as though by a \textit{dominate person }spell from 
a 5thlevel caster. The ability has a range of 30 feet. The save DC is Charisma-based.

\textbf{Energy Drain (Su):} Living creatures hit by a vampire spawn's slam attack 
gain one negative level. The DC is 14 for the Fortitude save to remove a negative 
level. The save DC is Charisma-based. For each such negative level bestowed, the 
vampire spawn gains 5 temporary hit points.

\textbf{Fast Healing (Ex):} A vampire spawn heals 2 points of damage each round 
so long as it has at least 1 hit point. If reduced to 0 hit points in combat, it 
automatically assumes gaseous form and attempts to escape. It must reach its coffin 
home within 2 hours or be utterly destroyed. (It can travel up to nine miles in 
2 hours.) Once at rest in its coffin, it is helpless. It regains 1 hit point after 
1 hour, then is no longer helpless and resumes healing at the rate of 2 hit points 
per round.

\textbf{Gaseous Form (Su):} As a standard action, a vampire spawn can assume \textit{gaseous 
form }at will as the spell (caster level 6th), but it can remain gaseous indefinitely 
and has a fly speed of 20 feet with perfect maneuverability.

\textbf{Spider Climb (Ex): }A vampire spawn can climb sheer surfaces as though 
with a \textit{spider climb }spell.

\textbf{Skills:} Vampire spawn have a +4 racial bonus on Bluff, Hide, Listen, Move 
Silently, Search, Sense Motive, and Spot checks.

VAMPIRE SPAWN WEAKNESSES

Vampire spawn are vulnerable to all attacks and effects that repel or slay vampires. 
For details, see the Vampire entry.

\vspace{12pt}
{\LARGE{}VARGOUILLE}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{222pt}|}
\hline
  & Small Outsider (Evil, Extraplanar)\tabularnewline
\hline
H\textbf{it Dice:} & 1d8+1 (5 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & Fly 30 ft. (good) (6 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 12 (+1 size, +1 Dex), touch 11, flat-footed 11\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +1/-3\tabularnewline
\hline
A\textbf{ttack:} & Bite +3 melee (1d4 plus poison)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +3 melee (1d4 plus poison)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Shriek, kiss, poison\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +3, Ref +3, Will +3\tabularnewline
\hline
A\textbf{bilities:} & Str 10, Dex 13, Con 12, Int 5, Wis 12, Cha 8\tabularnewline
\hline
S\textbf{kills:} & Hide +9, Intimidate +3, Listen +5, Move Silently +5, Spot +5\tabularnewline
\hline
F\textbf{eats:} & Weapon Finesse\tabularnewline
\hline
E\textbf{nvironment:} & An evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Cluster (2-5) or mob (6-11)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral evil\tabularnewline
\hline
A\textbf{dvancement:} & 2-3 HD (Small)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A vargouille is slightly larger than a human head, about 18 inches high, with a 
wingspan of 4 feet. It weighs about 10 pounds.

Vargouilles speak Infernal.

COMBAT

Vargouilles attack by biting with their jagged teeth. Their special attacks make 
them even more dangerous. A vargouille's natural weapons, as well as any weapons 
it wields, are treated as evil-aligned for the purpose of overcoming damage reduction.

\textbf{Shriek (Su):} Instead of biting, a vargouille can open its distended mouth 
and let out a terrible shriek. Those within 60 feet (except other vargouilles) 
who hear the shriek and can clearly see the creature must succeed on a DC 12 Fortitude 
save or be paralyzed with fear for 2d4 rounds or until the monster attacks them, 
goes out of range, or leaves their sight. A paralyzed creature is susceptible to 
the vargouille's kiss (see below). A creature that successfully saves cannot be 
affected again by the same vargouille's shriek for 24 hours. The shriek is a mind-affecting 
fear effect. The save DC is Constitution-based and includes a +1 racial bonus.

\textbf{Kiss (Su):} A vargouille can kiss a paralyzed target with a successful 
melee touch attack. An affected opponent must succeed on a DC 15 Fortitude save 
or begin a terrible transformation that turns the creature into a vargouille within 
24 hours (and often much sooner; roll 1d6 separately for each phase of the transformation).

First, over a period of 1d6 hours, all the victim's hair falls out. Within another 
1d6 hours thereafter, the ears grow into leathery wings, tentacles sprout on the 
chin and scalp, and the teeth become long, pointed fangs. During the next 1d6 hours, 
the victim takes Intelligence drain and Charisma drain equal to 1 point per hour 
(to a minimum of 3). The transformation is complete 1d6 hours later, when the head 
breaks free of the body (which promptly dies) and becomes a vargouille. This transformation 
is interrupted by sunlight, and even a \textit{daylight }spell can delay death, 
but to reverse the transformation requires \textit{remove disease}. The save DC 
is Constitution-based and

includes a +4 racial bonus.

\textbf{Poison (Ex):} Injury, Fortitude DC 12 or be unable to heal the vargouille's 
bite damage naturally or magically. A \textit{neutralize poison }or \textit{heal 
}spell removes the effect, while \textit{delay poison }allows magical healing. 
The save DC is Constitution-based and includes a +1 racial bonus.

\vspace{12pt}
{\LARGE{}WIGHT}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{193pt}|}
\hline
 & Medium Undead\tabularnewline
\hline
H\textbf{it Dice:} & 4d12 (26 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 15 (+1 Dex, +4 natural), touch 11, flat-footed 14\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +2/+3\tabularnewline
\hline
A\textbf{ttack:} & Slam +3 melee (1d4+1 plus energy drain)\tabularnewline
\hline
F\textbf{ull Attack:} & Slam +3 melee (1d4+1 plus energy drain)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Create spawn, energy drain\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., undead traits\tabularnewline
\hline
S\textbf{aves:} & Fort +1, Ref +2, Will +5\tabularnewline
\hline
A\textbf{bilities:} & Str 12, Dex 12, Con ---, Int 11, Wis 13, Cha 15\tabularnewline
\hline
S\textbf{kills:} & Hide +8, Listen +7, Move Silently +16, Spot +7 \tabularnewline
\hline
F\textbf{eats:} & Alertness, Blind-Fight\tabularnewline
\hline
E\textbf{nvironment:} & Any\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, gang (3-5), or pack (6-11)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 3\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & 5-8 HD (Medium)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A wight's appearance is a weird and twisted reflection of the form it had in life. 
A wight is about the height and weight of a human.

Wights speak Common.

COMBAT

Wights attack by hammering with their fists.

\textbf{Create Spawn (Su):} Any humanoid slain by a wight becomes a wight in 1d4 
rounds. Spawn are under the command of the wight that created them and remain enslaved 
until its death. They do not possess any of the abilities they had in life.

\textbf{Energy Drain (Su): }Living creatures hit by a wight's slam attack gain 
one negative level. The DC is 14 for the Fortitude save to remove a negative level. 
The save DC is Charisma-based. For each such negative level bestowed, the wight 
gains 5 temporary hit points.

\textbf{Skills:} Wights have a +8 racial bonus on Move Silently checks.

\vspace{12pt}
{\LARGE{}WILL-O'-WISP}

\begin{tabular}{|>{\raggedright}p{82pt}|>{\raggedright}p{243pt}|}
\hline
  & Small Aberration (Air)\tabularnewline
\hline
H\textbf{it Dice:} & 9d8 (40 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +13\tabularnewline
\hline
S\textbf{peed:} & Fly 50 ft. (perfect) (10 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 29 (+1 size, +9 Dex, +9 deflection), touch 29, flat-footed 
20\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +6/-3\tabularnewline
\hline
A\textbf{ttack:} & Shock +16 melee touch (2d8 electricity)\tabularnewline
\hline
F\textbf{ull Attack:} & Shock +16 melee touch (2d8 electricity)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:}--- & \tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., immunity to magic, natural invisibility\tabularnewline
\hline
S\textbf{aves:} & Fort +3, Ref +12, Will +9\tabularnewline
\hline
A\textbf{bilities:} & Str 1, Dex 29, Con 10, Int 15, Wis 16, Cha 12\tabularnewline
\hline
S\textbf{kills:} & Bluff +13, Diplomacy +3, Disguise +1 (+3 acting), Intimidate 
+3, Listen +17, Search +14, Spot +17, Survival +3 (+5 following tracks)\tabularnewline
\hline
F\textbf{eats:} & Alertness, Blind-Fight, Dodge, Improved Initiative, Weapon Finesse\textsuperscript{B}\tabularnewline
\hline
E\textbf{nvironment:} & Temperate marshes\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or string (3-4)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 6\tabularnewline
\hline
T\textbf{reasure:} & 1/10 coins; 50\% goods; 50\% items\tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 10-18 HD (Small)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

Will-o'-wisps can be yellow, white, green, or blue. They are easily mistaken for 
lanterns, especially in the foggy marshes and swamps where they reside. A will-o'-wisp's 
body is a globe of spongy material about 1 foot across and weighing about 3 pounds, 
and its glowing body sheds as much light as a torch.

Will-o'-wisps speak Common and Auran. They have no vocal apparatus but can vibrate 
to create a voice with a ghostly sound.

COMBAT

Will-o'-wisps usually avoid combat. They prefer to confuse and bewilder adventurers, 
luring them into morasses or other hazardous places. When they are forced to fight, 
they loose small electrical shocks, which act as melee touch attacks.

\textbf{Immunity to Magic (Ex): }A will-o'-wisp is immune to most spells or spell-like 
abilities that allow spell resistance, except \textit{magic missile }and \textit{maze.}

\textbf{Natural Invisibility (Ex): }A startled or frightened will-o'-wisp can extinguish 
its glow, effectively becoming invisible as the spell.

\vspace{12pt}
\subsection*{{\LARGE{}WINTER WOLF}}

\subsection*{\begin{tabular}{|>{\raggedright}p{74pt}|>{\raggedright}p{252pt}|}
\hline
 } & Large Magical Beast (Cold)\tabularnewline
\hline
H\textbf{it Dice:} & 6d10+18 (51 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +5\tabularnewline
\hline
S\textbf{peed:} & 50 ft. (10 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 15 (-1 size, +1 Dex, +5 natural), touch 10, flat-footed 
14\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +6/+14\tabularnewline
\hline
A\textbf{ttack:} & Bite +9 melee (1d8+6 plus 1d6 cold)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +9 melee (1d8+6 plus 1d6 cold)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Breath weapon, freezing bite, trip\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., immunity to cold, low-light vision, 
scent, vulnerability to fire\tabularnewline
\hline
S\textbf{aves:} & Fort +8, Ref +6, Will +3\tabularnewline
\hline
A\textbf{bilities:} & Str 18, Dex 13, Con 16, Int 9, Wis 13, Cha 10\tabularnewline
\hline
S\textbf{kills:} & Hide -1*, Listen +6, Move Silently +7, Spot +6, Survival +1*\tabularnewline
\hline
F\textbf{eats:} & Alertness, Improved Initiative, Track\tabularnewline
\hline
E\textbf{nvironment:} & Cold forests\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or pack (3-5)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 5\tabularnewline
\hline
T\textbf{reasure:} & 1/10 coins; 50\% goods; 50\% items\tabularnewline
\hline
A\textbf{lignment:} & Usually neutral evil\tabularnewline
\hline
A\textbf{dvancement:} & 7-9 HD (Large); 10-18 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +3 (cohort)\tabularnewline
\hline
\end{tabular}

A winter wolf grows about 8 feet long and stands about 4-1/2 feet at the shoulder. 
It weighs about 450 pounds.

Winter wolves can speak Giant and Common.

COMBAT

Winter wolves typically hunt in packs. Their size, cunning, and formidable breath 
weapon allow them to hunt and kill creatures much larger than themselves. A pack 
usually circles an opponent, each wolf attacking in turn to exhaust it. If they're 
in a hurry, white wolves try to pin their foes.

\textbf{Breath Weapon (Su):} 15-foot cone, once every 1d4 rounds, damage 4d6 cold, 
Reflex DC 16 half. The save DC is Constitution-based. 

\textbf{Freezing Bite (Su):} A winter wolf deals an extra 1d6 points of cold damage 
every time it bites an opponent, as if its bite were a frost weapon.

\textbf{Trip (Ex): }A winter wolf that hits with a bite attack can attempt to trip 
the opponent (+8 check modifier) as a free action without making a touch attack 
or provoking an attack of opportunity. If the attempt fails, the opponent cannot 
react to trip the winter wolf.

\textbf{Skills: }Winter wolves have a +1 racial bonus on Listen, Move Silently, 
and Spot checks. Winter wolves have a +2 racial bonus on Hide checks. *Their natural 
coloration grants them a +7 racial bonus on Hide checks in areas of snow and ice. 
A winter wolf has a +4 racial bonus on Survival checks when tracking by scent.

\vspace{12pt}
{\LARGE{}WORG}

\begin{tabular}{|>{\raggedright}p{89pt}|>{\raggedright}p{236pt}|}
\hline
  & Medium Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 4d10+8 (30 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +2\tabularnewline
\hline
S\textbf{peed:} & 50 ft. (10 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 14 (+2 Dex, +2 natural), touch 12, flat-footed 12\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +4/+7\tabularnewline
\hline
A\textbf{ttack:} & Bite +7 melee (1d6+4)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +7 melee (1d6+4)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Trip\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., low-light vision, scent\tabularnewline
\hline
S\textbf{aves:} & Fort +6, Ref +6, Will +3\tabularnewline
\hline
A\textbf{bilities:} & Str 17, Dex 15, Con 15, Int 6, Wis 14, Cha 10\tabularnewline
\hline
S\textbf{kills:} & Hide +4, Listen +6, Move Silently +6, Spot +6, Survival +2*\tabularnewline
\hline
F\textbf{eats:} & Alertness, Track\tabularnewline
\hline
E\textbf{nvironment:} & Temperate plains\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or pack (6-11)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2\tabularnewline
\hline
T\textbf{reasure:} & 1/10 coins; 50\% goods; 50\% items\tabularnewline
\hline
A\textbf{lignment:} & Usually neutral evil\tabularnewline
\hline
A\textbf{dvancement:} & 5-6 HD (Medium); 7-12 HD (Large)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +1 (cohort)\tabularnewline
\hline
\end{tabular}

A typical worg has gray or black fur, grows to 5 feet long and stands 3 feet tall 
at the shoulder. It weighs 300 pounds.

More intelligent than their smaller cousins, worgs speak their own language. Some 
can also speak Common and Goblin.

COMBAT

Mated pairs or packs work together to bring down large game, while lone worgs usually 
chase down creatures smaller than themselves. Both often use hit-and-run tactics 
to exhaust their quarry. A pack usually circles a larger opponent: Each wolf attacks 
in turn, biting and retreating, until the creature is exhausted, at which point 
the pack moves in for the kill. If they get impatient or heavily outnumber the 
opponent, worgs attempt to pin it.

\textbf{Trip (Ex):} A worg that hits with a bite attack can attempt to trip the 
opponent (+3 check modifier) as a free action without making a touch attack or 
provoking an attack of opportunity. If the attempt fails, the opponent cannot react 
to trip the worg.

\textbf{Skills: }A worg has a +1 racial bonus on Listen, Move Silently, and Spot 
checks, and a +2 racial bonus on Hide checks. *A worg has a +4 racial bonus on 
Survival checks when tracking

by scent.

\vspace{12pt}
{\LARGE{}WRAITH}

\begin{tabular}{|>{\raggedright}p{51pt}|>{\raggedright}p{131pt}|>{\raggedright}p{131pt}|}
\hline
 & W\textbf{raith} & D\textbf{read Wraith}\tabularnewline
\hline
  & Medium Undead (Incorporeal) & Large Undead (Incorporeal)\tabularnewline
\hline
H\textbf{it Dice:} & 5d12 (32 hp) & 16d12 (104 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +7 & +13\tabularnewline
\hline
S\textbf{peed:} & Fly 60 ft. (good) (12 squares) & Fly 60 ft. (good) (12 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 15 (+3 Dex, +2 deflection), touch 15, flat-footed 12 & 25 
(-1 size, +9 Dex, +7 deflection), touch 25, flat-footed 16\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +2/--- & +8/---\tabularnewline
\hline
A\textbf{ttack:} & Incorporeal touch +5 melee (1d4 plus 1d6 Constitution drain) & Incorporeal 
touch +16 melee (2d6 plus 1d8 Constitution drain)\tabularnewline
\hline
F\textbf{ull Attack:} & Incorporeal touch +5 melee (1d4 plus 1d6 Constitution drain) & Incorporeal 
touch +16 melee (2d6 plus 1d8 Constitution drain)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Constitution drain, create spawn & Constitution drain, 
create spawn\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., daylight powerlessness, incorporeal 
traits, +2 turn resistance, undead traits, unnatural aura & Darkvision 60 ft., 
daylight powerlessness, incorporeal traits, lifesense 60 ft., undead traits, unnatural 
aura\tabularnewline
\hline
S\textbf{aves:} & Fort +1, Ref +4, Will +6 & Fort +5, Ref +14, Will +14\tabularnewline
\hline
A\textbf{bilities:} & Str ---, Dex 16, Con ---, Int 14, Wis 14, Cha 15 & Str ---, 
Dex 28, Con ---, Int 17, Wis 18, Cha 24\tabularnewline
\hline
S\textbf{kills:} & Diplomacy +6, Hide +11, Intimidate +10, Listen +12, Search +10, 
Sense Motive +8, Spot +12, Survival +2 (+4 following tracks) & Diplomacy +9, Hide 
+24, Intimidate +26, Knowledge (religion) +22, Listen +25, Search +22, Sense Motive 
+23, Spot +25, Survival +4 (+6 following tracks)\tabularnewline
\hline
F\textbf{eats:} & Alertness\textsuperscript{B}, Blind-Fight, Combat Reflexes, Improved 
Initiative\textsuperscript{B} & Alertness\textsuperscript{B}, Blind-Fight, Combat 
Reflexes, Dodge, Improved Initiative\textsuperscript{B}, Improved Natural Attack 
(incorporeal touch), Mobility, Spring Attack\tabularnewline
\hline
E\textbf{nvironment:} & Any & Any\tabularnewline
\hline
O\textbf{rganization:} & Solitary, gang (2-5), or pack (6-11) & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 5 & 11\tabularnewline
\hline
T\textbf{reasure:} & None & None\tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil & Always lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & 6-10 HD (Medium) & 17-32 HD (Large)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & --- & \tabularnewline
\hline
\end{tabular}

Wraiths are incorporeal creatures born of evil and darkness. In some cases, the 
grim silhouette of a wraith might appear armored or outfitted with weapons. This 
appearance does not affect the creature's AC or combat abilities but only reflects 
the shape it had in life.

A wraith is about as tall as a human, while a dread wraith is roughly the size 
of an ogre. Since both are incorporeal, they are weightless.

Wraiths speak Common and Infernal.

COMBAT

Both the wraith and the dread wraith share the following abilities.

\textbf{Unnatural Aura (Su):} Animals, whether wild or domesticated, can sense 
the unnatural presence of a wraith at a distance of 30 feet. They will not willingly 
approach nearer than that and panic if forced to do so; they remain panicked as 
long as they are within that range.

\textbf{Daylight Powerlessness (Ex):} Wraiths are utterly powerless in natural 
sunlight (not merely a \textit{daylight }spell) and flee from it. 

WRAITH

\textbf{Constitution Drain (Su):} Living creatures hit by a wraith's incorporeal 
touch attack must succeed on a DC 14 Fortitude save or take 1d6 points of Constitution 
drain. The save DC is Charisma-based. On each such successful attack, the wraith 
gains 5 temporary hit points.

\textbf{Create Spawn (Su):} Any humanoid slain by a wraith becomes a wraith in 
1d4 rounds. Its body remains intact and inanimate, but its spirit is torn free 
from its corpse and transformed. Spawn are under the command of the wraith that 
created them and remain enslaved until its death. They do not possess any of the 
abilities they had in life.

DREAD WRAITH

The oldest and most malevolent wraiths lurk in the depths of forgotten temples 
and other forsaken places. They can sense the approach of living creatures, and 
hunger for them. Despite its size, the dread wraith possesses unearthly quickness, 
and makes use of its Spring Attack feat and natural reach to strike with deadly 
effect and melt back into the shadows---or the walls.

\textbf{Lifesense (Su):} A dread wraith notices and locates living creatures within 
60 feet, just as if it possessed the blindsight ability. It also senses the strength 
of their life force automatically, as if it had cast \textit{deathwatch.}

\textbf{Constitution Drain (Su):} Living creatures hit by a dread wraith's incorporeal 
touch attack must succeed on a DC 25 Fortitude save or take 1d8 points of Constitution 
drain. The save DC is Charisma-based. On each such successful attack, the dread 
wraith gains 5 temporary hit points.

\textbf{Create Spawn (Su): }Any humanoid slain by a dread wraith becomes a wraith 
in 1d4 rounds. Its body remains intact and inanimate, but its spirit is torn free 
from its corpse and transformed. Spawn are under the command of the wraith that 
created them and remain enslaved until its death. They do not possess any of the 
abilities they had in life.

\vspace{12pt}
{\LARGE{}WYVERN}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{235pt}|}
\hline
  & Large Dragon\tabularnewline
\hline
H\textbf{it Dice:} & 7d12+14 (59 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), fly 60 ft. (poor)\tabularnewline
\hline
A\textbf{rmor Class:} & 18 (-1 size, +1 Dex, +8 natural), touch 10, flat-footed 
17\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +7/+15\tabularnewline
\hline
A\textbf{ttack:} & Sting +10 melee (1d6+4 plus poison) or talon +10 melee (2d6+4) 
or bite +10 melee (2d8+4)\tabularnewline
\hline
F\textbf{ull Attack:} & Sting +10 melee (1d6+4 plus poison) and bite +8 melee (2d8+4) 
and 2 wings +8 melee (1d8+2) and 2 talons +8 melee (2d6+4) \tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Poison, improved grab\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., immunity to \textit{sleep }and 
paralysis, low-light vision, scent\tabularnewline
\hline
S\textbf{aves:} & Fort +7, Ref +6, Will +6\tabularnewline
\hline
A\textbf{bilities:} & Str 19, Dex 12, Con 15, Int 6, Wis 12, Cha 9\tabularnewline
\hline
S\textbf{kills:} & Hide +7, Listen +13, Move Silently +11, Spot +16 \tabularnewline
\hline
F\textbf{eats:} & Ability Focus (poison), Alertness, Flyby Attack, Multiattack\textsuperscript{B}\tabularnewline
\hline
E\textbf{nvironment:} & Warm hills\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or flight (3-6)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 6\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Usually neutral\tabularnewline
\hline
A\textbf{dvancement:} & 8-10 HD (Huge); 11-21 HD (Gargantuan)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A distant cousin to the true dragons, the wyvern is a huge flying lizard with a 
poisonous stinger in its tail.

A wyvern's body is 15 feet long, and dark brown to gray; half that length is tail. 
Its wingspan is about 20 feet. A wyvern weighs about one ton.

Wyverns speak Draconic, but usually don't bother with anything more elaborate than 
a loud hiss or a deep-throated growl much like that of a bull alligator.

COMBAT

Wyverns are rather stupid but always aggressive: They attack nearly anything that 
isn't obviously more powerful than themselves. A wyvern dives from the air, snatching 
the opponent with its talons and stinging it to death. A wyvern can slash with 
its talons only when making a flyby attack.

\textbf{Improved Grab (Ex):} To use this ability, a wyvern must hit with its talons. 
It can then attempt to start a grapple as a free action without provoking an attack 
of opportunity. If it wins the grapple check, it establishes a hold and stings. 

\textbf{Poison (Ex):} Injury, Fortitude DC 17, initial and secondary damage 2d6 
Con. The save DC is Constitution-based.

\textbf{Skills:} Wyverns have a +3 racial bonus on Spot checks.

\vspace{12pt}
{\LARGE{}XILL}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{223pt}|}
\hline
  & Medium Outsider (Extraplanar)\tabularnewline
\hline
H\textbf{it Dice:} & 5d8+10 (32 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +7\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 20 (+3 Dex, +7 natural), touch 13, flat-footed 17\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +5/+7\tabularnewline
\hline
A\textbf{ttack:} & Short sword +7 melee (1d6+2/19-20) or claw +7 melee (1d4+2) 
or longbow +8 ranged (1d8/x3)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 short swords +5 melee (1d6+2/19-20, 1d6+1/19-20) and 
2 claws +5 melee (1d4+1); or 4 claws +5 melee (1d4+2, 1d4+1); or 2 longbows +4 
ranged (1d8/x3)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Implant, improved grab, paralysis\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., planewalk, spell resistance 21\tabularnewline
\hline
S\textbf{aves:} & Fort +6, Ref +7, Will +5\tabularnewline
\hline
A\textbf{bilities:} & Str 15, Dex 16, Con 15, Int 12, Wis 12, Cha 11\tabularnewline
\hline
S\textbf{kills:} & Balance +13, Climb +10, Diplomacy +2, Escape Artist +11, Intimidate 
+8, Listen +9, Move Silently +11, Sense Motive +8, Spot +9, Tumble +11, Use Rope 
+3 (+5 with bindings) \tabularnewline
\hline
F\textbf{eats:} & Improved Initiative, Multiattack\textsuperscript{B}, Multiweapon 
Fighting\tabularnewline
\hline
E\textbf{nvironment:} & Ethereal Plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary or gang (2-5)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 6\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & 6-8 HD (Medium); 9-15 HD (Large)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +4\tabularnewline
\hline
\end{tabular}

A xill stands 4 to 5 feet tall and weighs about 100 pounds. Xills speak Infernal.

COMBAT

Xills are dangerous opponents, attacking with all four limbs. More civilized ones 
use weapons, usually fighting with two at a time so as to leave two claws free 
for grab attacks. Xills typically lie in wait on the Ethereal Plane for suitable 
prey to happen by, then ambush it using their planewalk ability. They make full 
use of their Tumble skill in combat: Usually, one or two distract physically powerful 
enemies by attacking, then assuming a defensive stance while their fellows maneuver 
to advantage. 

\textbf{Implant (Ex): }As a standard action, a xill can lay eggs inside a paralyzed 
creature. The young emerge about 90 days later, literally devouring the host from 
inside. A \textit{remove disease }spell rids a victim of the egg, as does a DC 
25 Heal check. If the check fails, the healer can try again, but each attempt (successful 
or not) deals 1d4 points of damage to the patient.

\textbf{Improved Grab (Ex): }To use this ability, a xill must hit with one or more 
claw attacks. It can then attempt to start a grapple as a free action without provoking 
an attack of opportunity. It receives a +2 bonus on the grapple check for each 
claw that hits. If it wins the grapple check and maintains the hold in the next 
round, it automatically bites the foe at that time. The bite deals no damage but 
injects a paralyzing venom.

\textbf{Paralysis (Ex):} Those bitten by a xill must succeed on a DC 14 Fortitude 
save or be paralyzed for 1d4 hours. The save DC is Constitution-based.

\textbf{Planewalk (Su): }These planar travelers like to slip between the folds 
of space to attack enemies as though from thin air. They can cross from the Ethereal 
Plane with a move action but take 2 rounds to cross back, during which time they 
are immobile. As a xill fades away, it becomes harder to hit: Opponents have a 
20\% miss chance in

the first round and a 50\% miss chance in the second. A xill can planewalk with 
a willing or helpless creature.

\vspace{12pt}
{\LARGE{}XORN}

\begin{tabular}{|>{\raggedright}p{42pt}|>{\raggedright}p{86pt}|>{\raggedright}p{86pt}|>{\raggedright}p{86pt}|}
\hline
  & M\textbf{inor Xorn} & A\textbf{verage Xorn} & E\textbf{lder Xorn}\tabularnewline
\hline
  & Small Outsider (Extraplanar, Earth) & Medium Outsider (Extraplanar, Earth) & Large 
Outsider (Extraplanar, Earth)\tabularnewline
\hline
H\textbf{it Dice:} & 3d8+9 (22 hp) & 7d8+17 (48 hp) & 15d8+63 (130 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +0 & +0 & +0\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), burrow 20 ft. & 20 ft. (4 squares), burrow 
20 ft. & 20 ft. (4 squares), burrow 20 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 23 (+1 size, +12 natural), touch 11, flat-footed 23 & 24 
(+14 natural), touch 10, flat-footed 24 & 25 (-1 size, +16 natural), touch 9, flat-footed 
25\tabularnewline
\hline
A\textbf{ttack:} & Bite +6 melee (2d8+2) & Bite +10 melee (4d6+3) & Bite +21 melee 
(4d8+7)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +6 melee (2d8+2) and 3 claws +4 melee (1d3+1) & Bite 
+10 melee (4d6+3) and 3 claws +8 melee (1d4+1) & Bite +21 melee (4d8+7) and 3 claws 
+19 melee (1d6+3)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. & 5 ft./5 ft. & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:}--- & --- & --- & \tabularnewline
\hline
S\textbf{pecial Qualities:} & All-around vision, earth glide, damage reduction 
5/bludgeoning, darkvision 60 ft., immunity to cold and fire, resistance to electricity 
10, tremorsense 60 ft. & All-around vision, earth glide, damage reduction 5/bludgeoning, 
darkvision 60 ft., immunity to cold and fire, resistance to electricity 10, tremorsense 
60 ft. & All-around vision, earth glide, damage reduction 5/bludgeoning, darkvision 
60 ft., immunity to cold and fire, resistance to electricity 10, tremorsense 60 
ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +5, Ref +3, Will +3 & Fort +7, Ref +5, Will +5 & Fort +13, 
Ref +9, Will +9\tabularnewline
\hline
A\textbf{bilities:} & Str 15, Dex 10, Con 15, Int 10, Wis 11, Cha 10 & Str 17, 
Dex 10, Con 15, & Str 25, Dex 10, Con 19, Int 10, Wis 11, Cha 10 Int 10, Wis 11, 
Cha 10\tabularnewline
\hline
S\textbf{kills:} & Hide +10, Intimidate +3, Knowledge (dungeoneering) +6, Listen 
+6, Move Silently +3, Search +6, Spot +8, Survival +6 (+8 following tracks or underground) & Hide 
+10, Intimidate +10, Knowledge(dungeoneering) +10, Listen +10, Move Silently +10, 
Search +10, Spot +10, Survival+10 (+12 following tracks or underground) & Hide 
+14, Intimidate +18, Knowledge (dungeoneering) +18, Listen +18, Move Silently +18, 
Search +22, Spot +22, Survival+18 (+20 following tracks or underground)\tabularnewline
\hline
F\textbf{eats:} & Multiattack, Toughness & Cleave\textsuperscript{B}, Multiattack, 
Power Attack, Toughness & Awesome Blow, Cleave\textsuperscript{B}, Great Cleave, 
Improved Bull Rush, Multiattack, Power Attack, Toughness\tabularnewline
\hline
E\textbf{nvironment:} & Elemental Plane of Earth & Elemental Plane of Earth & Elemental 
Plane of Earth\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or cluster (3-5) & Solitary, pair, or 
cluster (3-5) & Solitary, pair, or party (6-11)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 3 & 6 & 8\tabularnewline
\hline
T\textbf{reasure:} & None & None & None\tabularnewline
\hline
A\textbf{lignment:} & Usually neutral & Usually neutral & Usually neutral\tabularnewline
\hline
A\textbf{dvancement:} & 4-6 HD (Small) & 8-14 HD (Medium) & 16-21 HD (Large); 22-45 
HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & --- & --- & \tabularnewline
\hline
\end{tabular}

Minor xorns are about 3 feet tall and wide and weigh about 120 pounds. Average 
xorns are about 5 feet tall and wide, weighing about 600 pounds. Elder xorns are 
about 8 feet tall and wide and weigh about 9,000 pounds.

Xorns speak Common and Terran.

COMBAT

Xorns do not attack fleshly beings except to defend themselves or their property, 
since they cannot digest meat. Xorns are indifferent to creatures of the Material 
Plane---with the sole exception of anyone carrying a significant amount of precious 
metals or minerals, which xorns eat. They can smell food up to 20 feet away. A 
xorn can be quite aggressive when seeking food, especially on the Material Plane, 
where such sustenance is harder to find than it is on its native plane.

A xorn's favorite mode of attack is to wait just beneath a stone surface until 
a foe comes within reach, then emerge suddenly. Groups of xorns often send one 
of their number to the surface to negotiate for food while the remainder position 
themselves for a surprise attack.

\textbf{All-Around Vision (Ex): }A xorn's symmetrically placed eyes allow it to 
look in any direction, providing a +4 racial bonus on Spot and Search checks. A 
xorn can't be flanked. 

\textbf{Earth Glide (Ex):} A xorn can glide through stone, dirt, or almost any 
other sort of earth except metal as easily as a fish swims through water. Its burrowing 
leaves behind no tunnel or hole, nor does it create any ripple or other signs of 
its presence. A \textit{move earth }spell cast on an area containing a burrowing 
xorn flings the xorn back 30 feet, stunning the creature for 1 round unless it 
succeeds on a DC 15 Fortitude save.

\vspace{12pt}
{\LARGE{}YETH HOUND}

\begin{tabular}{|>{\raggedright}p{80pt}|>{\raggedright}p{245pt}|}
\hline
  & Medium Outsider (Extraplanar, Evil)\tabularnewline
\hline
H\textbf{it Dice:} & 3d8+6 (19 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +6\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares), fly 60 ft. (good)\tabularnewline
\hline
A\textbf{rmor Class:} & 20 (+2 Dex, +8 natural), touch 12, flat-footed 18\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +3/+6\tabularnewline
\hline
A\textbf{ttack:} & Bite +6 melee (1d8+4)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +6 melee (1d8+4)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Bay, trip\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 10/silver, darkvision 60 ft., flight, 
scent\tabularnewline
\hline
S\textbf{aves:} & Fort +5, Ref +5, Will +5\tabularnewline
\hline
A\textbf{bilities:} & Str 17, Dex 15, Con 15, Int 6, Wis 14, Cha 10\tabularnewline
\hline
S\textbf{kills:} & Listen +11, Spot +11, Search +7, Survival +11 (+13 following 
tracks)*\tabularnewline
\hline
F\textbf{eats:} & Improved Initiative, Track\tabularnewline
\hline
E\textbf{nvironment:} & An evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or pack (6-11)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 3\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral evil\tabularnewline
\hline
A\textbf{dvancement:} & 4-6 HD (Medium); 7-9 HD (Large)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +3 (cohort)\tabularnewline
\hline
\end{tabular}

A yeth hound stands 5 feet tall at the shoulder and weighs about 400 pounds.

Yeth hounds cannot speak, but understand Infernal.

COMBAT

Yeth hounds hunt only at night. They fear the sun and never venture out in daylight, 
even if their lives depend on it. 

A yeth hound's natural weapons, as well as any weapons it wields, are treated as 
evil-aligned for the purpose of overcoming damage reduction.

\textbf{Bay (Su):} When a yeth hound howls or barks, all creatures except other 
evil outsiders within a 300-foot spread must succeed on a DC 11 Will save or become 
panicked for 2d4 rounds. This is a sonic mind-affecting fear effect. Whether or 
not the save is successful, an affected creature is immune to the same hound's 
bay for 24 hours. The save DC is Charisma-based.

\textbf{Trip (Ex): }A yeth hound that hits with its bite attack can attempt to 
trip the opponent (+3 check modifier) as a free action without making a touch attack 
or provoking an attack of opportunity. If the attempt fails, the opponent cannot 
react to trip the yeth hound.

\textbf{Flight (Su):} A yeth hound can cease or resume flight as a free action. 

\textbf{Skills:} *A yeth hound has a +4 racial bonus on Survival checks when tracking 
by scent.

\vspace{12pt}
{\LARGE{}YRTHAK}

\begin{tabular}{|>{\raggedright}p{90pt}|>{\raggedright}p{236pt}|}
\hline
  & Huge Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 12d10+36 (102 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +6\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), fly 60 ft. (average)\tabularnewline
\hline
A\textbf{rmor Class:} & 18 (-2 size, +2 Dex, +8 natural), touch 10, flat-footed 
16\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +12/+25\tabularnewline
\hline
A\textbf{ttack:} & Bite +15 melee (2d8+5) or sonic lance +12 ranged touch (6d6)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +15 melee (2d8+5) and 2 claws +13 melee (1d6+2); or 
sonic lance +12 ranged touch (6d6)\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Sonic lance, explosion\tabularnewline
\hline
S\textbf{pecial Qualities:} & Blindsight 120 ft., immunities, vulnerability to 
sonic\tabularnewline
\hline
S\textbf{aves:} & Fort +11, Ref +10, Will +5\tabularnewline
\hline
A\textbf{bilities:} & Str 20, Dex 14, Con 17, Int 7, Wis 13, Cha 11\tabularnewline
\hline
S\textbf{kills:} & Listen +12, Move Silently +10 \tabularnewline
\hline
F\textbf{eats:} & Endurance, Flyby Attack, Improved Initiative, Multiattack, Snatch\tabularnewline
\hline
E\textbf{nvironment:} & Temperate mountains\tabularnewline
\hline
O\textbf{rganization:} & Solitary or clutch (2-4)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 9\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Often neutral\tabularnewline
\hline
A\textbf{dvancement:} & 13-16 HD (Huge); 17-36 HD (Gargantuan)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A yrthak is blind. It senses sound and movement by means of a special organ on 
its long tongue. It emits powerfully focused beams of sound from the protrusion 
on its head. The creature is a yellowish-green color, with the wings and fin being 
more yellow and the head and body more green. The teeth are yellow.

A yrthak is about 20 feet long, with a wingspan of 40 feet. It weighs about 5,000 
pounds.

Despite their intelligence, yrthaks do not speak.

COMBAT

A yrthak prefers to attack from the air, strafing the ground with sonic attacks 
or snatching up and dropping prey.

\textbf{Sonic Lance (Su): }Once every 2 rounds, a yrthak can focus sonic energy 
in a ray up to 60 feet long. This is a ranged touch attack that deals 6d6 points 
of damage to a single target.

\textbf{Explosion (Su):} A yrthak can fire its sonic lance at the ground, a large 
rock, a stone wall, or the like to create an explosion of shattered stone. This 
attack deals 2d6 points of piercing damage to all within 10 feet of the effect's 
center. This counts as a use of the sonic lance attack and thus is usable only 
once every 2 rounds, and never on the round following a sonic lance attack.

\textbf{Blindsight (Ex): }A yrthak can ascertain all foes within 120 feet. Beyond 
that range it is considered blinded. Yrthaks are invulnerable to gaze attacks, 
visual effects of spells such as illusions, and other attack forms that rely on 
sight. A yrthak whose sense of hearing is impaired is effectively blinded, treating 
all targets as having total concealment.

\textbf{Immunities: }Yrthaks have immunity to gaze attacks, visual effects, illusions, 
and other attack forms that rely on sight.

\textbf{Skills:} Yrthaks have a +4 racial bonus on Listen checks.

\vspace{12pt}
\subsection*{{\LARGE{}ZOMBIE}}

Zombies are corpses reanimated through dark and sinister magic.

Because of their utter lack of intelligence, the instructions given to a newly 
created zombie must be very simple.

CREATING A ZOMBIE

``Zombie'' is an acquired template that can be added to any corporeal creature 
(other than an undead) that has a skeletal system (referred to hereafter as the 
base creature). 

\textbf{Size and Type:} The creature's type changes to undead. It retains any subtypes 
except alignment subtypes and subtypes that indicate kind. It does not gain the 
augmented subtype. It uses all the base creature's statistics and special abilities 
except as noted here.

\textbf{Hit Dice:} Drop any Hit Dice from class levels (to a minimum of 1), double 
the number of Hit Dice left, and raise them to d12s. If the base creature has more 
than 10 Hit Dice (not counting those gained with experience), it can't be made 
into a zombie with the \textit{animate dead }spell.

\textbf{Speed:} If the base creature can fly, its maneuverability rating drops 
to clumsy.

\textbf{Armor Class:} Natural armor bonus increases by a number based on the zombie's 
size:

\begin{tabular}{|>{\raggedright}p{60pt}|>{\raggedright}p{14pt}|}
\hline
Tiny or smaller & +0\tabularnewline
\hline
Small & +1\tabularnewline
\hline
Medium & +2\tabularnewline
\hline
Large & +3\tabularnewline
\hline
Huge & +4\tabularnewline
\hline
Gargantuan & +7\tabularnewline
\hline
Colossal & +11\tabularnewline
\hline
\end{tabular}

\textbf{Base Attack: }A zombie has a base attack bonus equal to 1/2 its Hit Dice.

\textbf{Attacks:} A zombie retains all the natural weapons, manufactured weapon 
attacks, and weapon proficiencies of the base creature. A zombie also gains a slam 
attack.

\textbf{Damage:} Natural and manufactured weapons deal damage normally. A slam 
attack deals damage depending on the zombie's size. (Use the base creature's slam 
damage if it's better.)

\begin{tabular}{|>{\raggedright}p{45pt}|>{\raggedright}p{13pt}|}
\hline
Fine & 1\tabularnewline
\hline
Diminutive & 1d2\tabularnewline
\hline
Tiny & 1d3\tabularnewline
\hline
Small & 1d4\tabularnewline
\hline
Medium & 1d6\tabularnewline
\hline
Large & 1d8\tabularnewline
\hline
Huge & 2d6\tabularnewline
\hline
Gargantuan & 2d8\tabularnewline
\hline
Colossal & 4d6\tabularnewline
\hline
\end{tabular}

\textbf{Special Attacks:} A zombie retains none of the base creature's special 
attacks.

\textbf{Special Qualities:} A zombie loses most special qualities of the base creature. 
It retains any extraordinary special qualities that improve its melee or ranged 
attacks. A zombie gains the following special quality.

\textit{Single Actions Only (Ex): }Zombies have poor reflexes and can perform only 
a single move action or attack action each round. A zombie can move up to its speed 
and attack in the same round, but only if it attempts a charge.

\textbf{Saves:} Base save bonuses are Fort +1/3 HD, Ref +1/3 HD, and Will +1/2 
HD + 2.

\textbf{Abilities:} A zombie's Strength increases by +2, its Dexterity decreases 
by 2, it has no Constitution or Intelligence score, its Wisdom changes to 10, and 
its Charisma changes to 1.

\textbf{Skills:} A zombie has no skills.

\textbf{Feats: }A zombie loses all feats of the base creature and gains Toughness.

\textbf{Environment:} Any land and underground. 

\textbf{Organization:} Any.

\textbf{Challenge Rating:} Depends on Hit Dice, as follows:

\begin{tabular}{|>{\raggedright}p{34pt}|>{\raggedright}p{73pt}|}
\hline
H\textbf{it Dice} & C\textbf{hallenge Rating}\tabularnewline
\hline
1/2 & 1/8\tabularnewline
\hline
1 & 1/4\tabularnewline
\hline
2 & 1/2\tabularnewline
\hline
4 & 1\tabularnewline
\hline
6 & 2\tabularnewline
\hline
8-10 & 3\tabularnewline
\hline
12-14 & 4\tabularnewline
\hline
15-16 & 5\tabularnewline
\hline
18-20 & 6\tabularnewline
\hline
\end{tabular}

\textbf{Treasure: }None.

\textbf{Alignment: }Always neutral evil.

Advancement: As base creature, but double Hit Dice (maximum 20), or --- if the 
base creature advances by character class.

\textbf{Level Adjustment:} ---.

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{50pt}|>{\raggedright}p{84pt}|>{\raggedright}p{84pt}|>{\raggedright}p{84pt}|}
\hline
  & K\textbf{obold Zombie} & H\textbf{uman Commoner Zombie} & T\textbf{roglodyte 
Zombie}\tabularnewline
\hline
  & Small Undead & Medium Undead & Medium Undead\tabularnewline
\hline
H\textbf{it Dice:} & 2d12+3 (16 hp) & 2d12+3 (16 hp) & 4d12+3 (29 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +0- & 1- & 2\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares; can't run) & 30 ft. (6 squares; can't run) & 30 
ft. (6 squares; can't run)\tabularnewline
\hline
A\textbf{rmor Class:} & 13 (+1 size, +2 natural), touch 11, flat-footed 13 & 11 
(-1 Dex, +2 natural), touch 9, flat-footed 11 & 16 (-2 Dex, +8 natural), touch 
8, flat-footed 16\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +1/-4 & +1/+2 & +2/+3\tabularnewline
\hline
A\textbf{ttack:} & Spear+1 melee (1d6-1/x3) or slam +1 melee (1d4-1) or light crossbow 
+2 ranged (1d6/19-20) & Slam+2 melee (1d6+1) or club +2 melee (1d6+1) & Greatclub 
+3 melee (1d10+1) or bite +3 melee (1d4+1) or slam +3 melee (1d6+1) or javelin 
+0 ranged (1d6+1)\tabularnewline
\hline
F\textbf{ull Attack:} & Spear +0 melee (1d6-1/x3) or slam +0 melee (1d4-1) or light 
crossbow +1 ranged (1d6/19-20) & Slam+2 melee, (1d6+1) or club +2 melee (1d6+1) & Greatclub 
+3 melee (1d10+1) or bite +3 melee (1d4+1) or slam +3 melee (1d6+1) or javelin 
+0 ranged (1d6+1)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. & 5 ft./5 ft & 5 ft./5 ft\tabularnewline
\hline
S\textbf{pecial Attacks:}--- & --- & --- & \tabularnewline
\hline
S\textbf{pecial Qualities:} & Single actions only, damage reduction 5/slashing, 
darkvision 60 ft., undead traits & Single actions only, damage reduction 5/slashing, 
darkvision 60 ft., undead traits & Single actions only, damage reduction 5/slashing, 
darkvision 60 ft., undead traits\tabularnewline
\hline
S\textbf{aves:} & Fort +0, Ref +0, Will +3 & Fort +0, Ref -1, Will +3 & Fort +1, 
Ref -1, Will +4\tabularnewline
\hline
A\textbf{bilities:} & Str 8, Dex 11, Con ---, Int ---, Wis 10, Cha 1 & Str 12, 
Dex 8, Con ---, Int ---, Wis 10, Cha 1 & Str 12, Dex 7, Con ---, Int ---, Wis 10, 
Cha 1\tabularnewline
\hline
S\textbf{kills:}--- & --- & --- & \tabularnewline
\hline
F\textbf{eats:} & Toughness & Toughness & Toughness\tabularnewline
\hline
E\textbf{nvironment:} & Temperate forests & Any & Underground\tabularnewline
\hline
O\textbf{rganization:} & Any & Any & Any\tabularnewline
\hline
C\textbf{hallenge Rating:} & 1/2 & 1/2 & 1\tabularnewline
\hline
T\textbf{reasure:} & None & None & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral evil & Always neutral evil & Always neutral 
evil\tabularnewline
\hline
A\textbf{dvancement:} & None & None & None\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & --- & --- & \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{50pt}|>{\raggedright}p{83pt}|>{\raggedright}p{83pt}|>{\raggedright}p{83pt}|}
\hline
  & B\textbf{ugbear Zombie} & O\textbf{gre Zombie} & M\textbf{inotaur Zombie}\tabularnewline
\hline
  & Medium Undead & Large Undead & Large Undead\tabularnewline
\hline
H\textbf{it Dice:} & 6d12+3 (42 hp) & 8d12+3 (55 hp) & 12d8+3 (81 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +0- & 2- & 1\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares; can't run) & 40 ft. (8 squares; can't run) & 30 
ft. (6 squares; can't run)\tabularnewline
\hline
A\textbf{rmor Class:} & 16 (+5 natural, +1 light wooden shield), touch 10, flat-footed 
16 & 15 (-1 size, -2 Dex, +8 natural) touch 7, flat-footed 15 & 16 (-1 size, -1 
Dex, +8 natural) touch 8, flat-footed 16\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +3/+6 & +4/+14 & +6/+15\tabularnewline
\hline
A\textbf{ttack:} & Morningstar +6 melee (1d8+3) or slam +6 melee (1d6+3) or javelin 
+3 ranged (1d6+2) & Greatclub +9 melee (2d8+9) or slam +9 melee (1d8+9) or javelin 
+1 ranged (1d8+6) & Greataxe +10 melee (3d6+7/x3) or gore +10 melee (1d8+5) or 
slam +10 melee (1d8+5)\tabularnewline
\hline
F\textbf{ull Attack:} & Morningstar +6 melee (1d8+3) or slam +6 melee (1d6+3) or 
javelin +3 ranged (1d6+2) & Greatclub +9 melee (2d8+9) or slam +9 melee (1d8+9) 
or javelin +1 ranged (1d8+6) & Greataxe +10 melee (3d6+7/x3) or gore +10 melee 
(1d8+5) or slam +10 melee (1d8+5)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. & 10 ft./10 ft. & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:}--- & --- & --- & \tabularnewline
\hline
S\textbf{pecial Qualities:} & Single actions only, damage reduction 5/slashing, 
darkvision 60 ft., undead traits & Single actions only, damage reduction 5/slashing, 
darkvision 60 ft., undead traits & Single actions only, damage reduction 5/slashing, 
darkvision 60 ft., undead traits\tabularnewline
\hline
S\textbf{aves:} & Fort +2, Ref +2, Will +5 & Fort +2, Ref +0, Will +6 & Fort +4, 
Ref +3, Will +8\tabularnewline
\hline
A\textbf{bilities:} & Str 17, Dex 10, Con ---, Int ---, Wis 10, Cha 1 & Str 23, 
Dex 6, Con ---, Int ---, Wis 10, Cha 1 & Str 21, Dex 8, Con ---, Int ---, Wis 10, 
Cha 1\tabularnewline
\hline
S\textbf{kills:}--- & --- & --- & \tabularnewline
\hline
F\textbf{eats:} & Toughness & Toughness & Toughness\tabularnewline
\hline
E\textbf{nvironment:} & Temperate mountains & Temperate hills & Underground\tabularnewline
\hline
O\textbf{rganization:} & Any & Any & Any\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2 & 3 & 4\tabularnewline
\hline
T\textbf{reasure:} & None & None & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral evil & Always neutral evil & Always neutral 
evil\tabularnewline
\hline
A\textbf{dvancement:} & None & None & None\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & --- & --- & \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{55pt}|>{\raggedright}p{129pt}|>{\raggedright}p{129pt}|}
\hline
  & W\textbf{yvern Zombie} & G\textbf{ray Render Zombie}\tabularnewline
\hline
  & Large Undead & Large Undead\tabularnewline
\hline
H\textbf{it Dice:} & 14d12+3 (94 hp) & 20d8+3 (133 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +0- & 1\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares; can't run), fly 60 ft. (poor) & 30 ft. (6 
squares; can't run)\tabularnewline
\hline
A\textbf{rmor Class:} & 20 (-2 size, +12 natural), touch 8, flat-footed 20 & 16 
(-1 size, -1 Dex, +8 natural) touch 8, flat-footed 16\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +7/+16 & +10/+21\tabularnewline
\hline
A\textbf{ttack:} & Slam +11 melee (2d6+7) or talons +11 melee (2d6+5) & Bite +16 
melee (2d6+7) or slam +16 melee (1d8+10)\tabularnewline
\hline
F\textbf{ull Attack:} & Slam +11 melee (2d6+7) or talons +11 melee (2d6+5) & Bite 
+16 melee (2d6+7) or slam +16 melee (1d8+10)\tabularnewline
\hline
S\textbf{pecial Attacks:}--- & --- & \tabularnewline
\hline
S\textbf{pecial Qualities:} & Single actions only, damage reduction 5/slashing, 
darkvision 60 ft., undead traits & Single actions only, damage reduction 5/slashing, 
darkvision 60 ft., undead traits\tabularnewline
\hline
S\textbf{aves:} & Fort +4, Ref +4, Will +9 & Fort +6, Ref +5, Will +12\tabularnewline
\hline
A\textbf{bilities:} & Str 21, Dex 10, Con ---, Int ---, Wis 10, Cha 1 & Str 25, 
Dex 8, Con ---, Int ---, Wis 10, Cha 1\tabularnewline
\hline
S\textbf{kills:}--- & --- & \tabularnewline
\hline
F\textbf{eats:} & Toughness & Toughness\tabularnewline
\hline
E\textbf{nvironment:} & Warm hills & Temperate marshes\tabularnewline
\hline
O\textbf{rganization:} & Any & Any\tabularnewline
\hline
C\textbf{hallenge Rating:} & 4 & 6\tabularnewline
\hline
T\textbf{reasure:} & None & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral evil & Always neutral evil\tabularnewline
\hline
A\textbf{dvancement:} & 16-20 HD (Huge) & None\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & --- & \tabularnewline
\hline
\end{tabular}

\newpage

\end{document}
