%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{array}
\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

This material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

{\LARGE{}SPELLS (M-O)}

\vspace{12pt}
Mage Armor

Conjuration (Creation) [Force]

\textbf{Level:} Sor/Wiz 1

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 1 hour/level (D)

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} No

An invisible but tangible field of force surrounds the subject of a \textit{mage 
armor }spell, providing a +4 armor bonus to AC.

Unlike mundane armor, \textit{mage armor }entails no armor check penalty, arcane 
spell failure chance, or speed reduction. Since \textit{mage armor }is made of 
force, incorporeal creatures can't bypass it the way they do normal armor.

\textit{Focus: }A piece of cured leather.

\vspace{12pt}
Mage Hand

Transmutation

\textbf{Level:} Brd 0, Sor/Wiz 0

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One nonmagical, unattended object weighing up to 5 lb.

\textbf{Duration:} Concentration

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You point your finger at an object and can lift it and move it at will from a distance. 
As a move action, you can propel the object as far as 15 feet in any direction, 
though the spell ends if the distance between you and the object ever exceeds the 
spell's range.

\vspace{12pt}
Mage's Disjunction

Abjuration

\textbf{Level: }Magic 9, Sor/Wiz 9

\textbf{Components:} V

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Area:} All magical effects and magic items within a 40-ft.-radius burst

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates (object)

\textbf{Spell Resistance:} No

All magical effects and magic items within the radius of the spell, except for 
those that you carry or touch, are disjoined. That is, spells and spell-like effects 
are separated into their individual components (ending the effect as a \textit{dispel 
magic }spell does), and each permanent magic item must make a successful Will save 
or be turned into a normal item. An item in a creature's possession uses its own 
Will save bonus or its possessor's Will save bonus, whichever is higher.

You also have a 1\% chance per caster level of destroying an \textit{antimagic 
field. }If the \textit{antimagic field }survives the \textit{disjunction, }no items 
within it are disjoined.

Even artifacts are subject to \textit{disjunction, }though there is only a 1\% 
chance per caster level of actually affecting such powerful items. Additionally, 
if an artifact is destroyed, you must make a DC 25 Will save or permanently lose 
all spellcasting abilities. (These abilities cannot be recovered by mortal magic, 
not even \textit{miracle }or \textit{wish}.)

\textit{Note: }Destroying artifacts is a dangerous business, and it is 95\% likely 
to attract the attention of some powerful being who has an interest in or connection 
with the device.

\vspace{12pt}
Mage's Faithful Hound

Conjuration (Creation)

\textbf{Level:} Sor/Wiz 5

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} Phantom watchdog

\textbf{Duration:} 1 hour/caster level or until discharged, then 1 round/caster 
level; see text

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You conjure up a phantom watchdog that is invisible to everyone but yourself. It 
then guards the area where it was conjured (it does not move). The hound immediately 
starts barking loudly if any Small or larger creature approaches within 30 feet 
of it. (Those within 30 feet of the hound when it is conjured may move about in 
the area, but if they leave and return, they activate the barking.) The hound sees 
invisible and ethereal creatures. It does not react to figments, but it does react 
to shadow illusions.

If an intruder approaches to within 5 feet of the hound, the dog stops barking 
and delivers a vicious bite (+10 attack bonus, 2d6+3 points of piercing damage) 
once per round. The dog also gets the bonuses appropriate to an invisible creature.

The dog is considered ready to bite intruders, so it delivers its first bite on 
the intruder's turn. Its bite is the equivalent of a magic weapon for the purpose 
of damage reduction. The hound cannot be attacked, but it can be dispelled.

The spell lasts for 1 hour per caster level, but once the hound begins barking, 
it lasts only 1 round per caster level. If you are ever more than 100 feet distant 
from the hound, the spell ends.

\textit{Material Component: }A tiny silver whistle, a piece of bone, and a thread.

\vspace{12pt}
Mage's Lucubration

Transmutation

\textbf{Level: }Wiz 6

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} Instantaneous

You instantly recall any one spell of 5th level or lower that you have used during 
the past 24 hours. The spell must have been actually cast during that period. The 
recalled spell is stored in your mind as through prepared in the normal fashion.

If the recalled spell requires material components, you must provide them. The 
recovered spell is not usable until the material components are available.

\vspace{12pt}
Mage's Magnificent Mansion

Conjuration (Creation)

\textbf{Level:} Sor/Wiz 7

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} Extradimensional mansion, up to three 10-ft. cubes/level (S)

\textbf{Duration:} 2 hours/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You conjure up an extradimensional dwelling that has a single entrance on the plane 
from which the spell was cast. The entry point looks like a faint shimmering in 
the air that is 4 feet wide and 8 feet high. Only those you designate may enter 
the mansion, and the portal is shut and made invisible behind you when you enter. 
You may open it again from your own side at will. Once observers have passed beyond 
the entrance, they are in a magnificent foyer with numerous chambers beyond. The 
atmosphere is clean, fresh, and warm.

You can create any floor plan you desire to the limit of the spell's effect. The 
place is furnished and contains sufficient foodstuffs to serve a nine-course banquet 
to a dozen people per caster level. A staff of near-transparent servants (as many 
as two per caster level), liveried and obedient, wait upon all who enter. The servants 
function as \textit{unseen servant }spells except that they are visible and can 
go anywhere in the mansion.

Since the place can be entered only through its special portal, outside conditions 
do not affect the mansion, nor do conditions inside it pass to the plane beyond.

\textit{Focus: }A miniature portal carved from ivory, a small piece of polished 
marble, and a tiny silver spoon (each item worth 5 gp).

\vspace{12pt}
Mage's Private Sanctum

Abjuration

\textbf{Level:} Sor/Wiz 5

\textbf{Components:} V, S, M

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Area:} 30-ft. cube/level (S)

\textbf{Duration:} 24 hours (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell ensures privacy. Anyone looking into the area from outside sees only 
a dark, foggy mass. Darkvision cannot penetrate it. No sounds, no matter how loud, 
can escape the area, so nobody can eavesdrop from outside. Those inside can see 
out normally.

Divination (scrying) spells cannot perceive anything within the area, and those 
within are immune to \textit{detect thoughts}. The ward prevents speech between 
those inside and those outside (because it blocks sound), but it does not prevent 
other communication, such as a \textit{sending }or \textit{message }spell, or telepathic 
communication, such as that between a wizard and her familiar.

The spell does not prevent creatures or objects from moving into and out of the 
area.

\textit{Mage's private sanctum }can be made permanent with a \textit{permanency 
}spell.

\textit{Material Component: }A thin sheet of lead, a piece of opaque glass, a wad 
of cotton or cloth, and powdered chrysolite.

\vspace{12pt}
Mage's Sword

Evocation [Force]

\textbf{Level:} Sor/Wiz 7

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} One sword

\textbf{Duration:} 1 round/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

This spell brings into being a shimmering, swordlike plane of force. The sword 
strikes at any opponent within its range, as you desire, starting in the round 
that you cast the spell. The sword attacks its designated target once each round 
on your turn. Its attack bonus is equal to your caster level + your Int bonus or 
your Cha bonus (for wizards or sorcerers, respectively) with an additional +3 enhancement 
bonus. As a force effect, it can strike ethereal and incorporeal creatures. It 
deals 4d6+3 points of force damage, with a threat range of 19-20 and a critical 
multiplier of x2.

The sword always strikes from your direction. It does not get a bonus for flanking 
or help a combatant get one. If the sword goes beyond the spell range from you, 
if it goes out of your sight, or if you are not directing it, the sword returns 
to you and hovers.

Each round after the first, you can use a standard action to switch the sword to 
a new target. If you do not, the sword continues to attack the previous round's 
target.

The sword cannot be attacked or harmed by physical attacks, but \textit{dispel 
magic}, \textit{disintegrate}, a \textit{sphere of annihilation}, or a \textit{rod 
of cancellation }affects it. The sword's AC is 13 (10, +0 size bonus for Medium 
object, +3 deflection bonus).

If an attacked creature has spell resistance, the resistance is checked the first 
time \textit{Mage's sword }strikes it. If the sword is successfully resisted, the 
spell is dispelled. If not, the sword has its normal full effect on that creature 
for the duration of the spell.

\textit{Focus: }A miniature platinum sword with a grip and pommel of copper and 
zinc. It costs 250 gp to construct.

\vspace{12pt}
Magic Aura

Illusion (Glamer)

\textbf{Level:} Brd 1, Magic 1, Sor/Wiz 1

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} One touched object weighing up

to 5 lb./level

\textbf{Duration:} One day/level (D)

\textbf{Saving Throw:} None; see text

\textbf{Spell Resistance:} No

You alter an item's aura so that it registers to \textit{detect }spells (and spells 
with similar capabilities) as though it were nonmagical, or a magic item of a kind 
you specify, or the subject of a spell you specify. 

If the object bearing \textit{magic aura }has \textit{identify }cast on it or is 
similarly examined, the examiner recognizes that the aura is false and detects 
the object's actual qualities if he succeeds on a Will save. Otherwise, he believes 
the aura and no amount of testing reveals what the true magic is.

If the targeted item's own aura is exceptionally powerful (if it is an artifact, 
for instance), \textit{magic aura }doesn't work.

\textit{Note: }A magic weapon, shield, or suit of armor must be a masterwork item, 
so a sword of average make, for example, looks suspicious if it has a magical aura.

\textit{Focus: }A small square of silk that must be passed over the object that 
receives the aura.

\vspace{12pt}
Magic Circle against Chaos

Abjuration [Lawful]

\textbf{Level:} Clr 3, Law 3, Pal 3, Sor/Wiz 3

This spell functions like \textit{magic circle against evil, }except that it is 
similar to \textit{protection from chaos }instead of \textit{protection from evil}, 
and it can imprison a nonlawful called creature.

\vspace{12pt}
Magic Circle against Evil

Abjuration [Good]

\textbf{Level:} Clr 3, Good 3, Pal 3, Sor/Wiz 3

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Area:} 10-ft.-radius emanation from touched creature

\textbf{Duration:} 10 min./level

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} No; see text

All creatures within the area gain the effects of a \textit{protection from evil 
}spell, and no nongood summoned creatures can enter the area either. You must overcome 
a creature's spell resistance in order to keep it at bay (as in the third function 
of \textit{protection from evil}), but the deflection and resistance bonuses and 
the protection from mental control apply regardless of enemies' spell resistance.

This spell has an alternative version that you may choose when casting it. A \textit{magic 
circle against evil }can be focused inward rather than outward. When focused inward, 
the spell binds a nongood called creature (such as those called by the \textit{lesser 
planar binding, planar binding, }and \textit{greater planar binding }spells) for 
a maximum of 24 hours per caster level, provided that you cast the spell that calls 
the creature within 1 round of casting the \textit{magic circle}. The creature 
cannot cross the circle's boundaries. If a creature too large to fit into the spell's 
area is the subject of the spell, the spell acts as a normal \textit{protection 
from evil }spell for that creature only.

A \textit{magic circle }leaves much to be desired as a trap. If the circle of powdered 
silver laid down in the process of spellcasting is broken, the effect immediately 
ends. The trapped creature can do nothing that disturbs the circle, directly or 
indirectly, but other creatures can. If the called creature has spell resistance, 
it can test the trap once a day. If you fail to overcome its spell resistance, 
the creature breaks free, destroying the circle. A creature capable of any form 
of dimensional travel (\textit{astral projection, blink, dimension door, etherealness, 
gate, plane shift, shadow walk, teleport, }and similar abilities) can simply leave 
the circle through that means. You can prevent the creature's extradimensional 
escape by casting a \textit{dimensional anchor }spell on it, but you must cast 
the spell before the creature acts. If you are successful, the \textit{anchor }effect 
lasts as long as the \textit{magic circle }does. The creature cannot reach across 
the \textit{magic circle, }but its ranged attacks (ranged weapons, spells, magical 
abilities, and the like) can. The creature can attack any target it can reach with 
its ranged attacks except for the circle itself.

You can add a special diagram (a two-dimensional bounded figure with no gaps along 
its circumference, augmented with various magical sigils) to make the \textit{magic 
circle }more secure. Drawing the diagram by hand takes 10 minutes and requires 
a DC 20 Spellcraft check. You do not know the result of this check. If the check 
fails, the diagram is ineffective. You can take 10 when drawing the diagram if 
you are under no particular time pressure to complete the task. This task also 
takes 10 full minutes. If time is no factor at all, and you devote 3 hours and 
20 minutes to the task, you can take 20.

A successful diagram allows you to cast a \textit{dimensional anchor }spell on 
the \textit{magic circle }during the round before casting any summoning spell. 
The \textit{anchor }holds any called creatures in the \textit{magic circle }for 
24 hours per caster level. A creature cannot use its spell resistance against a 
\textit{magic circle }prepared with a diagram, and none of its abilities or attacks 
can cross the diagram. If the creature tries a Charisma check to break free of 
the trap (see the \textit{lesser planar binding }spell), the DC increases by 5. 
The creature is immediately released if anything disturbs the diagram---even a 
straw laid across it. However, the creature itself cannot disturb the diagram either 
directly or indirectly, as noted above.

This spell is not cumulative with \textit{protection from evil }and vice versa.

\textit{Arcane Material Component: }A little powdered silver with which you trace 
a 3-footdiameter circle on the floor (or ground) around the creature to be warded.

\vspace{12pt}
Magic Circle against Good

Abjuration [Evil]

\textbf{Level:} Clr 3, Evil 3, Sor/Wiz 3

This spell functions like \textit{magic circle against evil, }except that it is 
similar to \textit{protection from good }instead of \textit{protection from evil}, 
and it can imprison a nonevil called creature.

\vspace{12pt}
Magic Circle against Law

Abjuration [Chaotic]

\textbf{Level:} Chaos 3, Clr 3, Sor/Wiz 3

This spell functions like \textit{magic circle against evil, }except that it is 
similar to \textit{protection from law }instead of \textit{protection from evil}, 
and it can imprison a nonchaotic called creature.

\vspace{12pt}
Magic Fang

Transmutation

\textbf{Level:} Drd 1, Rgr 1

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Living creature touched

\textbf{Duration:} 1 min./level

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

\textit{Magic fang }gives one natural weapon of the subject a +1 enhancement bonus 
on attack and damage rolls. The spell can affect a slam attack, fist, bite, or 
other natural weapon. (The spell does not change an unarmed strike's damage from 
nonlethal damage to lethal damage.)

\textit{Magic fang }can be made permanent with a \textit{permanency }spell.

\vspace{12pt}
Magic Fang, Greater

Transmutation

\textbf{Level:} Drd 3, Rgr 3

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One living creature

\textbf{Duration:} 1 hour/level

This spell functions like \textit{magic fang}, except that the enhancement bonus 
on attack and damage rolls is +1 per four caster levels (maximum +5).

Alternatively, you may imbue all of the creature's natural weapons with a +1 enhancement 
bonus (regardless of your caster level).

\textit{Greater magic fang }can be made permanent with a \textit{permanency }spell.

\vspace{12pt}
Magic Jar

Necromancy

\textbf{Level:} Sor/Wiz 5

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Medium (100 ft. + 10 ft./level)

\textbf{Target:} One creature

\textbf{Duration:} 1 hour/level or until you return to your body

\textbf{Saving Throw: }Will negates; see text

\textbf{Spell Resistance:} Yes

By casting \textit{magic jar, }you place your soul in a gem or large crystal (known 
as the \textit{magic jar}), leaving your body lifeless. Then you can attempt to 
take control of a nearby body, forcing its soul into the \textit{magic jar. }You 
may move back to the jar (thereby returning the trapped soul to its body) and attempt 
to possess another body. The spell ends when you send your soul back to your own 
body, leaving the receptacle empty.

To cast the spell, the \textit{magic jar }must be within spell range and you must 
know where it is, though you do not need line of sight or line of effect to it. 
When you transfer your soul upon casting, your body is, as near as anyone can tell, 
dead. 

While in the \textit{magic jar, }you can sense and attack any life force within 
10 feet per caster level (and on the same plane of existence). You do need line 
of effect from the jar to the creatures. You cannot determine the exact creature 
types or positions of these creatures. In a group of life forces, you can sense 
a difference of 4 or more Hit Dice between one creature and another and can determine 
whether a life force is powered by positive or negative energy. (Undead creatures 
are powered by negative energy. Only sentient undead creatures have, or are, souls.)

You could choose to take over either a stronger or a weaker creature, but which 
particular stronger or weaker creature you attempt to possess is determined randomly.

Attempting to possess a body is a full-round action. It is blocked by \textit{protection 
from evil }or a similar ward. You possess the body and force the creature's soul 
into the \textit{magic jar }unless the subject succeeds on a Will save. Failure 
to take over the host leaves your life force in the \textit{magic jar, }and the 
target automatically succeeds on further saving throws if you attempt to possess 
its body again.

If you are successful, your life force occupies the host body, and the host's life 
force is imprisoned in the \textit{magic jar. }You keep your Intelligence, Wisdom, 
Charisma, level, class, base attack bonus, base save bonuses, alignment, and mental 
abilities. The body retains its Strength, Dexterity, Constitution, hit points, 
natural abilities, and automatic abilities. A body with extra limbs does not allow 
you to make more attacks (or more advantageous two-weapon attacks) than normal. 
You can't choose to activate the body's extraordinary or supernatural abilities. 
The creature's spells and spell-like abilities do not stay with the body.

As a standard action, you can shift freely from a host to the \textit{magic jar 
}if within range, sending the trapped soul back to its body. The spell ends when 
you shift from the jar to your own body.

If the host body is slain, you return to the \textit{magic jar, }if within range, 
and the life force of the host departs (it is dead). If the host body is slain 
beyond the range of the spell, both you and the host die. Any life force with nowhere 
to go is treated as slain.

If the spell ends while you are in the \textit{magic jar, }you return to your body 
(or die if your body is out of range or destroyed). If the spell ends while you 
are in a host, you return to your body (or die, if it is out of range of your current 
position), and the soul in the \textit{magic jar }returns to its body (or dies 
if it is out of range). Destroying the receptacle ends the spell, and the spell 
can be dispelled at either the \textit{magic jar }or at the host's location.

\textit{Focus: }A gem or crystal worth at least 100 gp.

\vspace{12pt}
Magic Missile

Evocation [Force]

\textbf{Level:} Sor/Wiz 1

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Targets:} Up to five creatures, no two of which can be more than 15 ft. 
apart

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

A missile of magical energy darts forth from your fingertip and strikes its target, 
dealing 1d4+1 points of force damage.

The missile strikes unerringly, even if the target is in melee combat or has less 
than total cover or total concealment. Specific parts of a creature can't be singled 
out. Inanimate objects are not damaged by the spell.

For every two caster levels beyond 1st, you gain an additional missile---two at 
3rd level, three at 5th, four at 7th, and the maximum of five missiles at 9th level 
or higher. If you shoot multiple missiles, you can have them strike a single creature 
or several creatures. A single missile can strike only one creature. You must designate 
targets before you check for spell resistance or roll damage.

\vspace{12pt}
Magic Mouth

Illusion (Glamer)

\textbf{Level:} Brd 1, Sor/Wiz 2

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One creature or object

\textbf{Duration:} Permanent until discharged

\textbf{Saving Throw: }Will negates (object)

\textbf{Spell Resistance:} Yes (object)

This spell imbues the chosen object or creature with an enchanted mouth that suddenly 
appears and speaks its message the next time a specified event occurs. The message, 
which must be twenty-five or fewer words long, can be in any language known by 
you and can be delivered over a period of 10 minutes. The mouth cannot utter verbal 
components, use command words, or activate magical effects. It does, however, move 
according to the words articulated; if it were placed upon a statue, the mouth 
of the statue would move and appear to speak. Of course, \textit{magic mouth }can 
be placed upon a tree, rock, or any other object or creature.

The spell functions when specific conditions are fulfilled according to your command 
as set in the spell. Commands can be as general or as detailed as desired, although 
only visual and audible triggers can be used. Triggers react to what appears to 
be the case. Disguises and illusions can fool them. Normal darkness does not defeat 
a visual trigger, but magical \textit{darkness }or \textit{invisibility }does. 
Silent movement or magical \textit{silence }defeats audible triggers. Audible triggers 
can be keyed to general types of noises or to a specific noise or spoken word. 
Actions can serve as triggers if they are visible or audible. A \textit{magic mouth 
}cannot distinguish alignment, level, Hit Dice, or class except by external garb.

The range limit of a trigger is 15 feet per caster level, so a 6th-level caster 
can command a \textit{magic mouth }to respond to triggers as far as 90 feet away. 
Regardless of range, the mouth can respond only to visible or audible triggers 
and actions in line of sight or within hearing distance.

\textit{Magic mouth }can be made permanent with a \textit{permanency }spell.

\textit{Material Component: }A small bit of honeycomb and jade dust worth 10 gp.

\vspace{12pt}
Magic Stone

Transmutation

\textbf{Level:} Clr 1, Drd 1, Earth 1

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Targets:} Up to three pebbles touched

\textbf{Duration:} 30 minutes or until discharged

\textbf{Saving Throw: }Will negates (harmless, object)

\textbf{Spell Resistance:} Yes (harmless, object)

You transmute as many as three pebbles, which can be no larger than sling bullets, 
so that they strike with great force when thrown or slung. If hurled, they have 
a range increment of 20 feet. If slung, treat them as sling bullets (range increment 
50 feet). The spell gives them a +1 enhancement bonus on attack and damage rolls. 
The user of the stones makes a normal ranged attack\textit{. }Each stone that hits 
deals 1d6+1 points of damage (including the spell's enhancement bonus), or 2d6+2 
points against undead.

\vspace{12pt}
Magic Vestment

Transmutation

\textbf{Level:} Clr 3, Strength 3, War 3

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Armor or shield touched

\textbf{Duration:} 1 hour/level

\textbf{Saving Throw: }Will negates (harmless, object)

\textbf{Spell Resistance:} Yes (harmless, object)

You imbue a suit of armor or a shield with an enhancement bonus of +1 per four 
caster levels (maximum +5 at 20th level).

An outfit of regular clothing counts as armor that grants no AC bonus for the purpose 
of this spell.

\vspace{12pt}
Magic Weapon

Transmutation

\textbf{Level:} Clr 1, Pal 1, Sor/Wiz 1, War 1

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target: }Weapon touched

\textbf{Duration:} 1 min./level

\textbf{Saving Throw: }Will negates (harmless, object)

\textbf{Spell Resistance:} Yes (harmless, object)

\textit{Magic weapon }gives a weapon a +1 enhancement bonus on attack and damage 
rolls. (An enhancement bonus does not stack with a masterwork weapon's +1 bonus 
on attack rolls.)

You can't cast this spell on a natural weapon, such as an unarmed strike (instead, 
see \textit{magic fang}). A monk's unarmed strike is considered a weapon, and thus 
it can be enhanced by this spell.

\vspace{12pt}
Magic Weapon, Greater

Transmutation

\textbf{Level:} Clr 4, Pal 3, Sor/Wiz 3

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One weapon or fifty projectiles (all of which must be in contact 
with each other at the time of casting)

\textbf{Duration:} 1 hour/level

\textbf{Saving Throw: }Will negates (harmless, object)

\textbf{Spell Resistance:} Yes (harmless, object)

This spell functions like \textit{magic weapon, }except that it gives a weapon 
an enhancement bonus on attack and damage rolls of +1 per four caster levels (maximum 
+5).

Alternatively, you can affect as many as fifty arrows, bolts, or bullets. The projectiles 
must be of the same kind, and they have to be together (in the same quiver or other 
container). Projectiles, but not thrown weapons, lose their transmutation when 
used. (Treat shuriken as projectiles, rather than as thrown weapons, for the purpose 
of this spell.)

\textit{Arcane Material Component: }Powdered lime and carbon.

\vspace{12pt}
Major Creation

Conjuration (Creation)

\textbf{Level:} Sor/Wiz 5

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Duration:} See text

This spell functions like \textit{minor creation, }except that you can also create 
an object of mineral nature: stone, crystal, metal, or the like. The duration of 
the created item varies with its relative hardness and rarity, as indicated on 
the following table.

\begin{tabular}{|>{\raggedright}p{136pt}|>{\raggedright}p{60pt}|}
\hline
\subsection*{H\textbf{ardness and Rarity Examples}} & \subsection*{D\textbf{uration}}\tabularnewline
\hline
Vegetable matter & 2 hr./level\tabularnewline
\hline
Stone, crystal, base metals & 1 hr./level\tabularnewline
\hline
Precious metals & 20 min./level\tabularnewline
\hline
Gems & 10 min./level\tabularnewline
\hline
Rare metal\textsuperscript{\textbf{1}} & 1 round/level\tabularnewline
\hline
\multicolumn{2}{|p{196pt}|}{1 Includes adamantine, alchemical silver, and mithral. 
You can't use major creation to create a cold iron item.}\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
Major Image

Illusion (Figment)

\textbf{Level:} Brd 3, Sor/Wiz 3

\textbf{Duration:} Concentration + 3 rounds

This spell functions like \textit{silent image, }except that sound, smell, and 
thermal illusions are included in the spell effect. While concentrating, you can 
move the image within the range. 

The image disappears when struck by an opponent unless you cause the illusion to 
react appropriately.

\vspace{12pt}
Make Whole

Transmutation

\textbf{Level:} Clr 2

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One object of up to 10 cu. ft./ level

This spell functions like \textit{mending, }except that \textit{make whole }completely 
repairs an object made of any substance, even one with multiple breaks, to be as 
strong as new. The spell does not restore the magical abilities of a broken magic 
item made whole, and it cannot mend broken magic rods, staffs, or wands. The spell 
does not repair items that have been warped, burned, disintegrated, ground to powder, 
melted, or vaporized, nor does it affect creatures (including constructs).

\vspace{12pt}
Mark of Justice

Necromancy

\textbf{Level:} Clr 5, Pal 4

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} Permanent; see text

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

You draw an indelible mark on the subject and state some behavior on the part of 
the subject that will activate the mark. When activated, the mark curses the subject. 
Typically, you designate some sort of criminal behavior that activates the mark, 
but you can pick any act you please. The effect of the mark is identical with the 
effect of \textit{bestow curse}.

Since this spell takes 10 minutes to cast and involves writing on the target, you 
can cast it only on a creature that is willing or restrained.

Like the effect of \textit{bestow curse}, a \textit{mark of justice }cannot be 
dispelled, but it can be removed with a \textit{break enchantment}, \textit{limited 
wish}, \textit{miracle}, \textit{remove curse}, or \textit{wish }spell. \textit{Remove 
curse }works only if its caster level is equal to or higher than your \textit{mark 
of justice }caster level. These restrictions apply regardless of whether the mark 
has activated.

\vspace{12pt}
Mass (Spell Name)

Any spell whose name begins with \textit{mass }is alphabetized in this chapter 
according to the second word of the spell name. Thus, the description of a \textit{mass 
}spell appears near the description of the spell on which it is based. Spell chains 
that have \textit{mass }spells in them include those based on the spells \textit{bear's 
endurance, bull's strength, cat's grace, charm monster, cure critical wounds, cure 
light wounds, cure moderate wounds, cure serious wounds, eagle's splendor, enlarge 
person, fox's cunning, heal, hold monster, hold person, inflict critical wounds, 
inflict light wounds, inflict moderate wounds, inflict serious wounds, invisibility, 
owl's wisdom, reduce person, }and \textit{suggestion.}

\vspace{12pt}
Maze

Conjuration (Teleportation)

\textbf{Level:} Sor/Wiz 8

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One creature

\textbf{Duration:} See text

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

You banish the subject into an extradimensional labyrinth of force planes. Each 
round on its turn, it may attempt a DC 20 Intelligence check to escape the labyrinth 
as a full-round action. If the subject doesn't escape, the maze disappears after 
10 minutes, forcing the subject to leave.

On escaping or leaving the maze, the subject reappears where it had been when the 
\textit{maze }spell was cast. If this location is filled with a solid object, the 
subject appears in the nearest open space. Spells and abilities that move a creature 
within a plane, such as \textit{teleport }and \textit{dimension door, }do not help 
a creature escape a \textit{maze }spell, although a \textit{plane shift }spell 
allows it to exit to whatever plane is designated in that spell. Minotaurs are 
not affected by this spell.

\vspace{12pt}
Meld into Stone

Transmutation [Earth]

\textbf{Level:} Clr 3, Drd 3

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 10 min./level

\textit{Meld into stone }enables you to meld your body and possessions into a single 
block of stone. The stone must be large enough to accommodate your body in all 
three dimensions. When the casting is complete, you and not more than 100 pounds 
of nonliving gear merge with the stone. If either condition is violated, the spell 
fails and is wasted.

While in the stone, you remain in contact, however tenuous, with the face of the 
stone through which you melded. You remain aware of the passage of time and can 
cast spells on yourself while hiding in the stone. Nothing that goes on outside 
the stone can be seen, but you can still hear what happens around you. Minor physical 
damage to the stone does not harm you, but its partial destruction (to the extent 
that you no longer fit within it) expels you and deals you 5d6 points of damage. 
The stone's complete destruction expels you and slays you instantly unless you 
make a DC 18 Fortitude save.

Any time before the duration expires, you can step out of the stone through the 
surface that you entered. If the spell's duration expires or the effect is dispelled 
before you voluntarily exit the stone, you are violently expelled and take 5d6 
points of damage.

The following spells harm you if cast upon the stone that you are occupying: \textit{Stone 
to flesh }expels you and deals you 5d6 points of damage. \textit{Stone shape }deals 
you 3d6 points of damage but does not expel you. \textit{Transmute rock to mud 
}expels you and then slays you instantly unless you make a DC 18 Fortitude save, 
in which case you are merely expelled. Finally, \textit{passwall }expels you without 
damage.

\vspace{12pt}
Mending

Transmutation

\textbf{Level:} Brd 0, Clr 0, Drd 0, Sor/Wiz 0

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 10 ft.

\textbf{Target:} One object of up to 1 lb.

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates (harmless, object)

\textbf{Spell Resistance:} Yes (harmless, object)

\textit{Mending }repairs small breaks or tears in objects (but not warps, such 
as might be caused by a \textit{warp wood }spell). It will weld broken metallic 
objects such as a ring, a chain link, a medallion, or a slender dagger, providing 
but one break exists.

Ceramic or wooden objects with multiple breaks can be invisibly rejoined to be 
as strong as new. A hole in a leather sack or a wineskin is completely healed over 
by \textit{mending. }The spell can repair a magic item, but the item's magical 
abilities are not restored. The spell cannot mend broken magic rods, staffs, or 
wands, nor does it affect creatures (including constructs).

\vspace{12pt}
Message

Transmutation [Language-Dependent]

\textbf{Level:} Brd 0, Sor/Wiz 0

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Targets:} One creature/level

\textbf{Duration:} 10 min./level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You can whisper messages and receive whispered replies with little chance of being 
overheard. You point your finger at each creature you want to receive the message. 
When you whisper, the whispered message is audible to all targeted creatures within 
range. Magical \textit{silence, }1 foot of stone, 1 inch of common metal (or a 
thin sheet of lead), or 3 feet of wood or dirt blocks the spell. The message does 
not have to travel in a straight line. It can circumvent a barrier if there is 
an open path between you and the subject, and the path's entire length lies within 
the spell's range. The creatures that receive the message can whisper a reply that 
you hear. The spell transmits sound, not meaning. It doesn't transcend language 
barriers.

\textit{Note: }To speak a message, you must mouth the words and whisper, possibly 
allowing observers the opportunity to read your lips.

\textit{Focus: }A short piece of copper wire.

\vspace{12pt}
Meteor Swarm

Evocation [Fire]

\textbf{Level:} Sor/Wiz 9

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Area:} Four 40-ft.-radius spreads; see text

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None or Reflex half; see text

\textbf{Spell Resistance:} Yes

\textit{Meteor swarm }is a very powerful and spectacular spell that is similar 
to \textit{fireball }in many aspects. When you cast it, four 2- foot-diameter spheres 
spring from your outstretched hand and streak in straight lines to the spots you 
select. The meteor spheres leave a fiery trail of sparks.

If you aim a sphere at a specific creature, you may make a ranged touch attack 
to strike the target with the meteor. Any creature struck by one of these spheres 
takes 2d6 points of bludgeoning damage (no save) and receives no saving throw against 
the sphere's fire damage (see below). If a targeted sphere misses its target, it 
simply explodes at the nearest corner of the target's space. You may aim more than 
one meteor at the same target.

Once a sphere reaches its destination, it explodes in a 40-foot-radius spread, 
dealing 6d6 points of fire damage to each creature in the area. If a creature is 
within the area of more than one sphere, it must save separately against each. 
(Fire resistance applies to each sphere's damage individually.)

\vspace{12pt}
Mind Blank

Abjuration

\textbf{Level:} Protection 8, Sor/Wiz 8

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One creature

\textbf{Duration:} 24 hours

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

The subject is protected from all devices and spells that detect, influence, or 
read emotions or thoughts. This spell protects against all mind-affecting spells 
and effects as well as information gathering by divination spells or effects. \textit{Mind 
blank }even foils \textit{limited wish}, \textit{miracle}, and \textit{wish }spells 
when they are used in such a way as to affect the subject's mind or to gain information 
about it. In the case of scrying that scans an area the creature is in, such as 
\textit{arcane eye, }the spell works but the creature simply isn't detected. Scrying 
attempts that are targeted specifically at the subject do not work at all.

\vspace{12pt}
Mind Fog

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Brd 5, Sor/Wiz 5

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Effect:} Fog spreads in 20-ft. radius, 20 ft. high

\textbf{Duration:} 30 minutes and 2d6 rounds; see text

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

\textit{Mind fog }produces a bank of thin mist that weakens the mental resistance 
of those caught in it. Creatures in the \textit{mind fog }take a -10 competence 
penalty on Wisdom checks and Will saves. (A creature that successfully saves against 
the fog is not affected and need not make further saves even if it remains in the 
fog.) Affected creatures take the penalty as long as they remain in the fog and 
for 2d6 rounds thereafter. The fog is stationary and lasts for 30 minutes (or until 
dispersed by wind).

A moderate wind (11+ mph) disperses the fog in four rounds; a strong wind (21+ 
mph) disperses the fog in 1 round.

The fog is thin and does not significantly hamper vision.

\vspace{12pt}
Minor Creation

Conjuration (Creation)

\textbf{Level:} Sor/Wiz 4

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 minute

\textbf{Range:} 0 ft.

\textbf{Effect:} Unattended, nonmagical object of nonliving plant matter, up to 
1 cu. ft./level

\textbf{Duration:} 1 hour/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You create a nonmagical, unattended object of nonliving, vegetable matter. The 
volume of the item created cannot exceed 1 cubic foot per caster level. You must 
succeed on an appropriate skill check to make a complex item.

Attempting to use any created object as a material component causes the spell to 
fail.

\textit{Material Component: }A tiny piece of matter of the same sort of item you 
plan to create with \textit{minor creation}.

\vspace{12pt}
Minor Image

Illusion (Figment)

\textbf{Level:} Brd 2, Sor/Wiz 2

\textbf{Duration:} Concentration +2 rounds

This spell functions like \textit{silent image}, except that \textit{minor image 
}includes some minor sounds but not understandable speech.

\vspace{12pt}
Miracle

Evocation

\textbf{Level:} Clr 9, Luck 9

\textbf{Components:} V, S, XP; see text

\textbf{Casting Time:} 1 standard action

\textbf{Range:} See text

\textbf{Target, Effect, or Area: }See text

\textbf{Duration:} See text

\textbf{Saving Throw:} See text

\textbf{Spell Resistance:} Yes

You don't so much cast a \textit{miracle }as request one. You state what you would 
like to have happen and request that your deity (or the power you pray to for spells) 
intercede.

A \textit{miracle }can do any of the following things.• 

Duplicate any cleric spell of 8th level or lower (including spells to which you 
have access because of your domains). • 

Duplicate any other spell of 7th level or lower.• 

Undo the harmful effects of certain spells, such as \textit{feeblemind }or \textit{insanity}.• 

Have any effect whose power level is in line with the above effects.

If the \textit{miracle }has any of the above effects, casting it has no experience 
point cost.

Alternatively, a cleric can make a very powerful request. Casting such a \textit{miracle 
}costs the cleric 5,000 XP because of the powerful divine energies involved. Examples 
of especially powerful \textit{miracles }of this sort could include the following.• 

Swinging the tide of a battle in your favor by raising fallen allies to continue 
fighting.• 

Moving you and your allies, with all your and their gear, from one plane to another 
through planar barriers to a specific locale with no chance of error.• 

Protecting a city from an earthquake, volcanic eruption, flood, or other major 
natural disaster.

In any event, a request that is out of line with the deity's (or alignment's) nature 
is refused.

A duplicated spell allows saving throws and spell resistance as normal, but the 
save DCs are as for a 9th-level spell. When a \textit{miracle }duplicates a spell 
that has an XP cost, you must pay that cost. When a \textit{miracle }spell duplicates 
a spell with a material component that costs more than 100 gp, you must provide 
that component.

\textit{XP Cost: }5,000 XP (for some uses of the \textit{miracle }spell; see above).

\vspace{12pt}
Mirage Arcana

Illusion (Glamer)

\textbf{Level:} Brd 5, Sor/Wiz 5

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Area:} One 20-ft. cube/level (S)

\textbf{Duration:} Concentration +1 hour/ level (D)

This spell functions like \textit{hallucinatory terrain}, except that it enables 
you to make any area appear to be something other than it is. The illusion includes 
audible, visual, tactile, and olfactory elements. Unlike \textit{hallucinatory 
terrain, }the spell can alter the appearance of structures (or add them where none 
are present). Still, it can't disguise, conceal, or add creatures (though creatures 
within the area might hide themselves within the illusion just as they can hide 
themselves within a real location).

\vspace{12pt}
Mirror Image

Illusion (Figment)

\textbf{Level:} Brd 2, Sor/Wiz 2

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal; see text

\textbf{Target:} You

\textbf{Duration:} 1 min./level (D)

Several illusory duplicates of you pop into being, making it difficult for enemies 
to know which target to attack. The figments stay near you and disappear when struck.

\textit{Mirror image }creates 1d4 images plus one image per three caster levels 
(maximum eight images total). These figments separate from you and remain in a 
cluster, each within 5 feet of at least one other figment or you. You can move 
into and through a \textit{mirror image. }When you and the \textit{mirror image 
}separate, observers can't use vision or hearing to tell which one is you and which 
the image. The figments may also move through each other. The figments mimic your 
actions, pretending to cast spells when you cast a spell, drink potions when you 
drink a potion, levitate when you levitate, and so on.

Enemies attempting to attack you or cast spells at you must select from among indistinguishable 
targets. Generally, roll randomly to see whether the selected target is real or 
a figment. Any successful attack against an image destroys it. An image's AC is 
10 + your size modifier + your Dex modifier. Figments seem to react normally to 
area spells (such as looking like they're burned or dead after being hit by a \textit{fireball}).

While moving, you can merge with and split off from figments so that enemies who 
have learned which image is real are again confounded.

An attacker must be able to see the images to be fooled. If you are invisible or 
an attacker shuts his or her eyes, the spell has no effect. (Being unable to see 
carries the same penalties as being blinded.)

\vspace{12pt}
Misdirection

Illusion (Glamer)

\textbf{Level:} Brd 2, Sor/Wiz 2

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One creature or object, up to a 10-ft. cube in size

\textbf{Duration:} 1 hour/level

\textbf{Saving Throw:} None or Will negates; see text

\textbf{Spell Resistance:} No

By means of this spell, you misdirect the information from divination spells that 
reveal auras (\textit{detect evil, detect magic, discern lies, }and the like). 
On casting the spell, you choose another object within range. For the duration 
of the spell, the subject of \textit{misdirection }is detected as if it were the 
other object. (Neither the subject nor the other object gets a saving throw against 
this effect.) Detection spells provide information based on the second object rather 
than on the actual target of the detection unless the caster of the detection succeeds 
on a Will save. For instance, you could make yourself detect as a tree if one were 
within range at casting: not evil, not lying, not magical, neutral in alignment, 
and so forth. This spell does not affect other types of divination magic (\textit{augury, 
detect thoughts, clairaudience/clairvoyance, }and the like).

\vspace{12pt}
Mislead

Illusion (Figment, Glamer)

\textbf{Level:} Brd 5, Luck 6, Sor/Wiz 6, Trickery 6

\textbf{Components:} S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target/Effect:} You/one illusory double

\textbf{Duration:} 1 round/level (D) and concentration + 3 rounds; see text

\textbf{Saving Throw:} None or Will disbelief (if interacted with); see text

\textbf{Spell Resistance:} No

You become invisible (as \textit{improved invisibility}, a glamer), and at the 
same time, an illusory double of you (as \textit{major image, }a figment) appears. 
You are then free to go elsewhere while your double moves away. The double appears 
within range but thereafter moves as you direct it (which requires concentration 
beginning on the first round after the casting). You can make the figment appear 
superimposed perfectly over your own body so that observers don't notice an image 
appearing and you turning invisible. You and the figment can then move in different 
directions. The double moves at your speed and can talk and gesture as if it were 
real, but it cannot attack or cast spells, though it can pretend to do so.

The illusory double lasts as long as you concentrate upon it, plus 3 additional 
rounds. After you cease concentration, the illusory double continues to carry out 
the same activity until the duration expires. The \textit{improved invisibility 
}lasts for 1 round per level, regardless of concentration.

\vspace{12pt}
Mnemonic Enhancer

Transmutation

\textbf{Level: }Wiz 4

\textbf{Components:} V, S, M, F

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} Instantaneous

Casting this spell allows you to prepare additional spells or retain spells recently 
cast. Pick one of these two versions when the spell is cast.

\textit{Prepare: }You prepare up to three additional levels of spells. A cantrip 
counts as 1/2 level for this purpose. You prepare and cast these spells normally.

\textit{Retain: }You retain any spell of 3rd level or lower that you had cast up 
to 1 round before you started casting the \textit{mnemonic enhancer}. This restores 
the previously cast spell to your mind.

In either event, the spell or spells prepared or retained fade after 24 hours (if 
not cast).

\textit{Material Component: }A piece of string, and ink consisting of squid secretion 
with black dragon's blood.

\textit{Focus: }An ivory plaque of at least 50 gp value.

\vspace{12pt}
Modify Memory

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Brd 4

\textbf{Components:} V, S

\textbf{Casting Time:} 1 round; see text

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One living creature

\textbf{Duration:} Permanent

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

You reach into the subject's mind and modify as many as 5 minutes of its memories 
in one of the following ways.• 

Eliminate all memory of an event the subject actually experienced. This spell cannot 
negate \textit{charm, geas/quest, suggestion, }or similar spells.• 

Allow the subject to recall with perfect clarity an event it actually experienced.• 

Change the details of an event the subject actually experienced.• 

Implant a memory of an event the subject never experienced.

Casting the spell takes 1 round. If the subject fails to save, you proceed with 
the spell by spending as much as 5 minutes (a period of time equal to the amount 
of memory time you want to modify) visualizing the memory you wish to modify in 
the subject. If your concentration is disturbed before the visualization is complete, 
or if the subject is ever beyond the spell's range during this time, the spell 
is lost.

A modified memory does not necessarily affect the subject's actions, particularly 
if it contradicts the creature's natural inclinations. An illogical modified memory 
is dismissed by the creature as a bad dream or a memory muddied by too much wine. 

\vspace{12pt}
Moment of Prescience

Divination

\textbf{Level:} Luck 8, Sor/Wiz 8

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 1 hour/level or until discharged

This spell grants you a powerful sixth sense in relation to yourself. Once during 
the spell's duration, you may choose to use its effect. This spell grants you an 
insight bonus equal to your caster level (maximum +25) on any single attack roll, 
opposed ability or skill check, or saving throw. Alternatively, you can apply the 
insight bonus to your AC against a single attack (even if flatfooted). Activating 
the effect doesn't take an action; you can even activate it on another character's 
turn if needed. You must choose to use the \textit{moment of prescience }before 
you make the roll it is to modify. Once used, the spell ends.

You can't have more than one \textit{moment of prescience }active on you at the 
same time.

\vspace{12pt}
Mount

Conjuration (Summoning)

\textbf{Level:} Sor/Wiz 1

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 round

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} One mount

\textbf{Duration:} 2 hours/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You summon a light horse or a pony (your choice) to serve you as a mount. The steed 
serves willingly and well. The mount comes with a bit and bridle and a riding saddle.

\textit{Material Component: }A bit of horse hair.

\vspace{12pt}
Move Earth

Transmutation [Earth]

\textbf{Level:} Drd 6, Sor/Wiz 6

\textbf{Components:} V, S, M

\textbf{Casting Time:} See text

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Area:} Dirt in an area up to 750 ft. square and up to 10 ft. deep (S)

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

\textit{Move earth }moves dirt (clay, loam, sand), possibly collapsing embankments, 
moving hillocks, shifting dunes, and so forth.

However, in no event can rock formations be collapsed or moved. The area to be 
affected determines the casting time. For every 150-foot square (up to 10 feet 
deep), casting takes 10 minutes. The maximum area, 750 feet by 750 feet, takes 
4 hours and 10 minutes to move.

This spell does not violently break the surface of the ground. Instead, it creates 
wavelike crests and troughs, with the earth reacting with glacierlike fluidity 
until the desired result is achieved. Trees, structures, rock formations, and such 
are mostly unaffected except for changes in elevation and relative topography.

The spell cannot be used for tunneling and is generally too slow to trap or bury 
creatures. Its primary use is for digging or filling moats or for adjusting terrain 
contours before a battle.

This spell has no effect on earth creatures.

\textit{Material Component: }A mixture of soils (clay, loam, and sand) in a small 
bag, and an iron blade.

\vspace{12pt}
Neutralize Poison

Conjuration (Healing)

\textbf{Level:} Brd 4, Clr 4, Drd 3, Pal 4, Rgr 3

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature or object of up to 1 cu. ft./level touched

\textbf{Duration:} 10 min./level

\textbf{Saving Throw: }Will negates (harmless, object)

\textbf{Spell Resistance:} Yes (harmless, object)

You detoxify any sort of venom in the creature or object touched. A poisoned creature 
suffers no additional effects from the poison, and any temporary effects are ended, 
but the spell does not reverse instantaneous effects, such as hit point damage, 
temporary ability damage, or effects that don't go away on their own.

The creature is immune to any poison it is exposed to during the duration of the 
spell. Unlike with \textit{delay poison}, such effects aren't postponed until after 
the duration ---the creature need not make any saves against poison effects applied 
to it during the length of the spell.

This spell can instead neutralize the poison in a poisonous creature or object 
for the duration of the spell, at the caster's option.

\textit{Arcane Material Component: }A bit of charcoal.

\vspace{12pt}
Nightmare

Illusion (Phantasm) [Mind-Affecting, Evil]

\textbf{Level:} Brd 5, Sor/Wiz 5

\textbf{Components:} V, S

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Unlimited

\textbf{Target:} One living creature

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates; see text

\textbf{Spell Resistance:} Yes

You send a hideous and unsettling phantasmal vision to a specific creature that 
you name or otherwise specifically designate.

The \textit{nightmare }prevents restful sleep and causes 1d10 points of damage. 
The \textit{nightmare }leaves the subject fatigued and unable to regain arcane 
spells for the next 24 hours.

The difficulty of the save depends on how well you know the subject and what sort 
of physical connection (if any) you have to that creature.

\begin{tabular}{|>{\raggedright}p{227pt}|>{\raggedright}p{88pt}|}
\hline
K\textbf{nowledge} & W\textbf{ill Save Modifier}\tabularnewline
\hline
None\textsuperscript{\textbf{1}} & +10\tabularnewline
\hline
Secondhand (you have heard of the subject) & +5\tabularnewline
\hline
Firsthand (you have met the subject) & +0\tabularnewline
\hline
Familiar (you know the subject well)- & 5\tabularnewline
\hline
\multicolumn{2}{|p{316pt}|}{1 You must have some sort of connection to a creature 
you have no knowledge of.}\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{152pt}|>{\raggedright}p{79pt}|}
\hline
C\textbf{onnection} & W\textbf{ill Save Modifier}\tabularnewline
\hline
Likeness or picture- & 2\tabularnewline
\hline
Possession or garment- & 4\tabularnewline
\hline
Body part, lock of hair, bit of nail, etc.- & 10\tabularnewline
\hline
\end{tabular}

\textit{Dispel evil }cast on the subject while you are casting the spell dispels 
the \textit{nightmare }and causes you to be stunned for 10 minutes per caster level 
of the \textit{dispel evil}.

If the recipient is awake when the spell begins, you can choose to cease casting 
(ending the spell) or to enter a trance until the recipient goes to sleep, whereupon 
you become alert again and complete the casting. If you are disturbed during the 
trance, you must succeed on a Concentration check as if you were in the midst of 
casting a spell or the spell ends.

If you choose to enter a trance, you are not aware of your surroundings or the 
activities around you while in the trance.

You are defenseless, both physically and mentally, while in the trance. (You always 
fail any saving throw, for example.)

Creatures who don't sleep (such as elves, but not half-elves) or dream are immune 
to this spell.

\vspace{12pt}
Nondetection

Abjuration

\textbf{Level:} Rgr 4, Sor/Wiz 3, Trickery 3

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature or object touched

\textbf{Duration:} 1 hour/level

\textbf{Saving Throw: }Will negates (harmless, object)

\textbf{Spell Resistance:} Yes (harmless, object)

The warded creature or object becomes difficult to detect by divination spells 
such as \textit{clairaudience/clairvoyance, locate object, }and \textit{detect 
}spells. \textit{Nondetection }also prevents location by such magic items as \textit{crystal 
ball}s. If a divination is attempted against the warded creature or item, the caster 
of the divination must succeed on a caster level check (1d20 + caster level) against 
a DC of 11 + the caster level of the spellcaster who cast \textit{nondetection}. 
If you cast \textit{nondetection }on yourself or on an item currently in your possession, 
the DC is 15 + your caster level.

If cast on a creature, \textit{nondetection }wards the creature's gear as well 
as the creature itself.

\textit{Material Component: }A pinch of diamond dust worth 50 gp.

\vspace{12pt}
Obscure Object

Abjuration

\textbf{Level:} Brd 1, Clr 3, Sor/Wiz 2

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} One object touched of up to 100 lb./level

\textbf{Duration:} 8 hours (D)

\textbf{Saving Throw: }Will negates (object)

\textbf{Spell Resistance:} Yes (object)

This spell hides an object from location by divination (scrying) effects, such 
as the \textit{scrying }spell or a \textit{crystal ball}. Such an attempt automatically 
fails (if the divination is targeted on the object) or fails to perceive the object 
(if the divination is targeted on a nearby location, object, or person).

\textit{Arcane Material Component: }A piece of chameleon skin.

\vspace{12pt}
Obscuring Mist

Conjuration (Creation)

\textbf{Level:} Air 1, Clr 1, Drd 1, Sor/Wiz 1, Water 1

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 20 ft.

\textbf{Effect:} Cloud spreads in 20-ft. radius from you, 20 ft. high

\textbf{Duration:} 1 min./level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

A misty vapor arises around you. It is stationary once created. The vapor obscures 
all sight, including darkvision, beyond 5 feet. A creature 5 feet away has concealment 
(attacks have a 20\% miss chance). Creatures farther away have total concealment 
(50\% miss chance, and the attacker cannot use sight to locate the target).

A moderate wind (11+ mph), such as from a \textit{gust of wind }spell, disperses 
the fog in 4 rounds. A strong wind (21+ mph) disperses the fog in 1 round. A \textit{fireball, 
flame strike}, or similar spell burns away the fog in the explosive or fiery spell's 
area. A \textit{wall of fire }burns away the fog in the area into which it deals 
damage.

This spell does not function underwater.

\vspace{12pt}
Open/Close

Transmutation

\textbf{Level:} Brd 0, Sor/Wiz 0

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} Object weighing up to 30 lb. or portal that can be opened or closed

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates (object)

\textbf{Spell Resistance:} Yes (object)

You can open or close (your choice) a door, chest, box, window, bag, pouch, bottle, 
barrel, or other container. If anything resists this activity (such as a bar on 
a door or a lock on a chest), the spell fails. In addition, the spell can only 
open and close things weighing 30 pounds or less. Thus, doors, chests, and similar 
objects sized for enormous creatures may be beyond this spell's ability to affect.

\textit{Focus: }A brass key.

\vspace{12pt}
Order's Wrath

Evocation [Lawful]

\textbf{Level:} Law 4

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Area:} Nonlawful creatures within a burst that fills a 30-ft. cube

\textbf{Duration:} Instantaneous (1 round); see text

\textbf{Saving Throw: }Will partial; see text

\textbf{Spell Resistance:} Yes

You channel lawful power to smite enemies. The power takes the form of a three-dimensional 
grid of energy. Only chaotic and neutral (not lawful) creatures are harmed by the 
spell.

The spell deals 1d8 points of damage per two caster levels (maximum 5d8) to chaotic 
creatures (or 1d6 points of damage per caster level, maximum 10d6, to chaotic outsiders) 
and causes them to be dazed for 1 round. A successful Will save reduces the damage 
to half and negates the daze effect.

The spell deals only half damage to creatures who are neither chaotic nor lawful, 
and they are not dazed. They can reduce the damage in half again (down to one-quarter 
of the roll) with a successful Will save.

\vspace{12pt}
Overland Flight

Transmutation

\textbf{Level:} Sor/Wiz 5

\textbf{Components:} V, S

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 1 hour/level

This spell functions like a \textit{fly }spell, except you can fly at a speed of 
40 feet (30 feet if wearing medium or heavy armor, or if carrying a medium or heavy 
load) with average maneuverability. When using this spell for long-distance movement, 
you can hustle without taking nonlethal damage (a forced march still requires Constitution 
checks). This means you can cover 64 miles in an eight-hour period of flight (or 
48 miles at a speed of 30 feet).

\vspace{12pt}
Owl's Wisdom

Transmutation

\textbf{Level:} Clr 2, Drd 2, Pal 2, Rgr 2, Sor/Wiz 2

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 1 min./level

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes

The transmuted creature becomes wiser. The spell grants a +4 enhancement bonus 
to Wisdom, adding the usual benefit to Wisdom-related skills. Clerics, druids, 
paladins, and rangers (and other Wisdom-based spellcasters) who receive \textit{owl's 
wisdom }do not gain any additional bonus spells for the increased Wisdom, but the 
save DCs for their spells increase.

\textit{Arcane Material Component: }A few feathers, or a pinch of droppings, from 
an owl.

\vspace{12pt}
Owl's Wisdom, Mass

Transmutation

\textbf{Level:} Clr 6, Drd 6, Sor/Wiz 6

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One creature/level, no two of which can be more than 30 ft. apart

This spell functions like \textit{owl's wisdom}, except that it affects multiple 
creatures.

\newpage

\end{document}
