%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{array}
\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

This material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

\subsection*{{\LARGE{}MONSTERS (D-De)}}

\vspace{12pt}
\subsection*{{\LARGE{}DARKMANTLE}}

\subsection*{\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{192pt}|}
\hline
 } & Small Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 1d10+1 (6 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +4\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), fly 30 ft. (poor)\tabularnewline
\hline
A\textbf{rmor Class:} & 17 (+1 size, +6 natural), touch 11, flat-footed 17\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +1/+0\tabularnewline
\hline
A\textbf{ttack:} & Slam +5 melee (1d4+4)\tabularnewline
\hline
F\textbf{ull Attack:} & Slam +5 melee (1d4+4)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Darkness, improved grab, constrict 1d4+4\tabularnewline
\hline
S\textbf{pecial Qualities:} & Blindsight 90 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +3, Ref +2, Will +0\tabularnewline
\hline
A\textbf{bilities:} & Str 16, Dex 10, Con 13, Int 2, Wis 10, Cha 10\tabularnewline
\hline
S\textbf{kills:} & Hide +10, Listen +5*, Spot +5*\tabularnewline
\hline
F\textbf{eats:} & Improved Initiative\tabularnewline
\hline
E\textbf{nvironment:} & Underground\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, clutch (3-9), or swarm (6-15)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 1\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & 2-3 HD (Small)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

The darkmantle hangs from a ceiling by a muscular ``foot'' at the top of its body. 
It can look like a stalactite, by holding its tentacles stiffly under itself, or 
like a lump of rock, spreading its tentacles so the membrane between them covers 
its body. Its shell and skin usually resemble limestone, but a darkmantle can change 
its color to match almost any type of stony background.

A darkmantle is about 4 feet long from the tips of its tentacles to the top of 
its head. It weighs about 30 pounds.

COMBAT

A darkmantle attacks by dropping onto its prey and wrapping its tentacles around 
the opponent's head. Once attached, it squeezes and tries to suffocate the foe. 
A darkmantle that misses its initial attack often flies up and tries to drop on 
the opponent again.

\textbf{Darkness (Su): }Once per day a darkmantle can cause darkness as the \textit{darkness 
}spell (caster level 5th). It most often uses this ability just before attacking.

\textbf{Improved Grab (Ex): }To use this ability, a darkmantle must hit a Large 
or smaller creature with its slam attack. It can then attempt to start a grapple 
as a free action without provoking an attack of opportunity.

If it wins the grapple check, it attaches to the opponent's head and can constrict.

\textbf{Constrict (Ex): }A darkmantle deals 1d4+4 points of damage with a successful 
grapple check.

\textbf{Blindsight (Ex): }A darkmantle can ``see'' by emitting high-frequency sounds, 
inaudible to most other creatures, that allows it to ascertain objects and creatures 
within 90 feet. A \textit{silence }spell negates this ability and effectively blinds 
the darkmantle.

\textbf{Skills: }A darkmantle has a +4 racial bonus on Listen and Spot checks.

These bonuses are lost if its blindsight is negated. The creature's variable coloration 
gives it a +4 racial bonus on Hide checks.

\vspace{12pt}
{\LARGE{}DELVER}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{231pt}|}
\hline
  & Huge Aberration\tabularnewline
\hline
H\textbf{it Dice:} & 15d8+78 (145 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +5\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares), burrow 10 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 24 (-2 size, +1 Dex, +15 natural), touch 9, flat-footed 
23\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +11/+27\tabularnewline
\hline
A\textbf{ttack:} & Slam +17 melee (1d6+8 plus 2d6 acid)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 slams +17 melee (1d6+8 plus 2d6 acid)\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Corrosive slime\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., immunity to acid, stone shape, 
tremorsense 60 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +12, Ref +6, Will +11\tabularnewline
\hline
A\textbf{bilities:} & Str 27, Dex 13, Con 21, Int 14, Wis 14, Cha 12\tabularnewline
\hline
S\textbf{kills:} & Knowledge (dungeoneering) +14, Knowledge (nature) +4, Listen 
+20, Move Silently +17, Spot +20, Survival +14 (+16 underground)\tabularnewline
\hline
F\textbf{eats:} & Alertness, Blind-Fight, Great Fortitude, Improved Initiative, 
Power Attack, Toughness\tabularnewline
\hline
E\textbf{nvironment:} & Underground\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 9\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Usually neutral\tabularnewline
\hline
A\textbf{dvancement:} & 16-30 HD (Huge); 31-45 HD (Gargantuan)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A delver is about 15 feet long and 10 feet wide. It weighs about 6,000 pounds.

Delvers speak Terran and Undercommon.

COMBAT

A delver prefers to fight from its tunnel, which it uses to protect its flanks 
while lashing out with its flippers.

A delver expecting trouble may honeycomb an area with tunnels, leaving most closed 
with layers of stone 1 or 2 inches thick.

The delver can quickly dissolve the stone cover and pop up to attack unexpectedly.

\textbf{Corrosive Slime (Ex): }A delver produces a mucuslike slime that contains 
a highly corrosive substance. The slime is particularly effective against stone. 
A delver's mere touch deals 2d6 points of acid damage to organic creatures or objects. 
Against metallic creatures or objects, a delver's slime deals 4d8 points of damage, 
and against stony creatures (including earth elementals) or objects it deals 8d10 
points of damage. A slam attack by a delver leaves a patch of slime that deals 
2d6 points of damage on contact and another 2d6 points of damage in each of the 
next 2 rounds. A large quantity (at least a quart) of water or weak acid, such 
as vinegar, washes off the slime. An opponent's armor and clothing dissolve and 
become useless immediately unless the wearer succeeds on a DC 22 Reflex save. Weapons 
that strike a delver also dissolve immediately unless the wielder succeeds on a 
DC 22 Reflex save. A creature attacking a delver with natural weapons takes damage 
from its slime each time an attack hits unless the creature succeeds on a DC 22 
Reflex save. These save DCs are Constitution-based.

\textbf{Stone Shape (Ex):} A delver can alter its slime to temporarily soften stone 
instead of dissolving it. Once every 10 minutes, a delver can soften and shape 
up to 25 cubic feet of stone, as a \textit{stone shape }spell (caster level 15th).

\vspace{12pt}
{\LARGE{}DEMON}

Demons are a race of creatures native to chaotic evil-aligned planes. They are 
ferocity personified and will attack any creature just for the sheer fun of it---even 
other demons.

Demon Traits: Most demons possess the following traits (unless otherwise noted 
in a creature's entry).---

Immunity to electricity and poison.---

Resistance to acid 10, cold 10, and fire 10.---

\textit{Summon }(Sp): Many demons share the ability to summon others of their kind 
(the success chance and type of demon summoned are noted in each monster description). 
Demons are often reluctant to use this power until in obvious peril or extreme 
circumstances.---

Telepathy.

Except where otherwise noted, demons speak Abyssal, Celestial, and Draconic.

\vspace{12pt}
BABAU

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{213pt}|}
\hline
  & Medium Outsider (Chaotic, Extraplanar, Evil)\tabularnewline
\hline
H\textbf{it Dice:} & 7d8+35 (66 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 19 (+1 Dex, +8 natural), touch 11, flat-footed 18\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +7/+12\tabularnewline
\hline
A\textbf{ttack:} & Claw +12 melee (1d6+5)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +12 melee (1d6+5) and bite +7 melee (1d6+2)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Sneak attack +2d6, spell-like abilities, \textit{summon 
demon}\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 10/cold iron or good, darkvision 
60 ft., immunity to electricity and poison, protective slime, resistance to acid 
10, cold 10, and fire 10, spell resistance 14, telepathy 100 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +10, Ref +6, Will +6\tabularnewline
\hline
A\textbf{bilities:} & Str 21, Dex 12, Con 20, Int 14, Wis 13, Cha 16\tabularnewline
\hline
S\textbf{kills:} & Climb +15, Disable Device +12, Disguise +13, Escape Artist +11, 
Hide +19, Listen +19, Move Silently +19, Open Lock +11, Search +20, Sleight of 
Hand +11, Survival +1 (+3 following tracks), Use Rope +1 (+3 with bindings)\tabularnewline
\hline
F\textbf{eats:} & Cleave, Multiattack, Power Attack\tabularnewline
\hline
E\textbf{nvironment:} & A chaotic evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary or gang (3-6)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 6\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 8-14 HD (Large); 15-21 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A babau is about 6 feet tall and weighs about 140 pounds.

Combat

Babaus are sneaky and sly. They attack the most powerful foe first, hoping to eliminate 
the true threats quickly and then toy with the rest. When ambushing their opponents, 
they make excellent use of the combination of multiple attacks and sneak attacks.

A babau's natural weapons, as well as any weapons it wields, are treated as chaotic-aligned 
and evil-aligned for the purpose of overcoming damage reduction.

\textbf{Sneak Attack (Ex): }A babau can make a sneak attack like a rogue, dealing 
an extra 2d6 points of damage whenever a foe is denied his or her Dexterity bonus, 
or when the babau is flanking.

\textbf{Spell-Like Abilities:} At will---\textit{darkness, dispel magic, see invisibility}, 
\textit{greater teleport }(self plus 50 pounds of objects only). Caster level 7th.

\textbf{Protective Slime (Su):} A slimy red jelly coats the babau's skin. Any weapon 
that touches it takes 1d8 points of acid damage from the corrosive goo, and the 
weapon's hardness does not reduce this damage. A magic weapon may attempt a DC 
18 Reflex save to avoid taking this damage. A creature who strikes the babau with 
an unarmed attack, unarmed strike, touch spell, or natural weapon takes this damage 
as well but can negate the damage with a DC 18 Reflex save. The save DCs are Constitution-based.

\textit{\textbf{Summon Demon }}\textbf{(Sp):}Once per day a babau can attempt to 
summon 1 babau with a 40\% chance of success. This ability is the equivalent of 
a 3rd-level spell.

\textbf{Skills:} Babaus have a +8 racial bonus on Hide, Listen, Move Silently, 
and Search checks.

\vspace{12pt}
BALOR

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{235pt}|}
\hline
  & Large Outsider (Chaotic, Extraplanar, Evil)\tabularnewline
\hline
H\textbf{it Dice:} & 20d8+200 (290 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +11\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares), fly 90 ft. (good)\tabularnewline
\hline
A\textbf{rmor Class:} & 35 (-1 size, +7 Dex, +19 natural), touch 16, flat-footed 
28\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +20/+36\tabularnewline
\hline
A\textbf{ttack:} & +\textit{1 vorpal longsword }+33 melee (2d6+8/19-20)\tabularnewline
\hline
F\textbf{ull Attack:} & +\textit{1 vorpal longsword }+31/+26/+21/+16 melee (2d6+8/19-20) 
and \textit{+1 flaming whip }+30/+25 melee (1d4+4 plus 1d6 fire plus entangle); 
or 2 slams +31 melee (1d10+7)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft. (20 ft. with \textit{+1 flaming whip})\tabularnewline
\hline
S\textbf{pecial Attacks:} & Death throes, entangle, spell-like abilities, \textit{summon 
demon}, vorpal sword\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 15/cold iron and good, darkvision 
60 ft., flaming body, immunity to electricity, fire, and poison, resistance to 
acid 10 and cold 10, spell resistance 28, telepathy 100 ft., true seeing\tabularnewline
\hline
S\textbf{aves:} & Fort +22, Ref +19, Will +19\tabularnewline
\hline
A\textbf{bilities:} & Str 35, Dex 25, Con 31, Int 24, Wis 24, Cha 26\tabularnewline
\hline
S\textbf{kills:} & Bluff +31, Concentration +33, Diplomacy +35, Disguise +8 (+10 
acting), Hide +26, Intimidate +33, Knowledge (any two) +30, Listen +38, Move Silently 
+30, Search +30, Sense Motive +30, Spellcraft +30 (+32 scrolls), Spot +38, Survival 
+7 (+9 following tracks), Use Magic Device +31 (+33 scrolls)\tabularnewline
\hline
F\textbf{eats:} & Cleave, Improved Initiative, Improved Two-Weapon Fighting, Power 
Attack, Quicken Spell-Like Ability (\textit{telekinesis}), Two-Weapon Fighting, 
Weapon Focus (longsword)\tabularnewline
\hline
E\textbf{nvironment:} & A chaotic evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary or troupe (1 balor, 1 marilith, and 2-5 hezrous)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 20\tabularnewline
\hline
T\textbf{reasure:} & Standard coins; double goods; standard items, plus \textit{+1 
vorpal greatsword }and \textit{+1 flaming whip}\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 21-30 HD (Large); 31-60 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A balor stands about 12 feet tall. Its skin is usually dark red. It weighs about 
4,500 pounds.

\textbf{Combat}

Balors love to join battle armed with their swords and whips. If they face stiff 
resistance, they may teleport away to loose a few spell-like effects at the foe.

A balor's \textit{+1 flaming whip }is a long, flexible weapon with many tails tipped 
with hooks, spikes, and balls. The weapon deals bludgeoning and slashing damage, 
in addition to fire damage.

A balor's natural weapons, as well as any weapons it wields, are treated as chaotic-aligned 
and evil-aligned for the purpose of overcoming damage reduction.

\textbf{Death Throes (Ex):}When killed, a balor explodes in a blinding flash of 
light that deals 100 points of damage to anything within 100 feet (Reflex DC 30 
half ). This explosion automatically destroys any weapons the balor is holding. 
The save DC is Constitution-based.

\textbf{Entangle (Ex):} A balor's \textit{+1 flaming whip }entangles foes much 
like an attack with a net. The whip has 20 hit points. The whip needs no folding. 
If it hits, the target and the balor immediately make opposed Strength checks; 
if the balor wins, it drags the target against its flaming body (see below). The 
target remains anchored against the balor's body until it escapes the whip.

\textbf{Spell-Like Abilities:} At will--- \textit{blasphemy }(DC 25), \textit{dominate 
monster }(DC 27), \textit{greater dispel magic, greater teleport }(self plus 50 
pounds of objects only), \textit{insanity }(DC 25), \textit{power word stun, telekinesis 
}(DC 23), \textit{unholy aura }(DC 26); 1/day---\textit{fire storm }(DC 26), \textit{implosion 
}(DC 27). Caster level 20th. The save DCs are Charisma-based.

\textbf{Vorpal Sword (Su):} Every balor carries a \textit{+1 vorpal longsword }that 
looks like a flame or a bolt of lightning.

\textit{\textbf{Summon Demon }}\textbf{(Sp):} Once per day a balor can automatically 
summon 4d10 dretches, 1d4 hezrous, or one nalfeshnee, glabrezu, marilith, or balor. 
This ability is the equivalent of a 9th-level spell.

\textbf{Flaming Body (Su):} The body of a balor is wreathed in flame. 

Anyone grappling a balor takes 6d6 points of fire damage each round.

\textbf{True Seeing (Su):} Balors have a continuous true seeing ability, as the 
spell (caster level 20th).

\textbf{Skills:} Balors have a +8 racial bonus on Listen and Spot checks.

\textbf{Tactics Round-by-Round}

The balor is most effective as a ranged combatant, using its spell-like abilities 
to attack from a distance.

\textit{Prior to combat: Unholy aura.}

\textit{Round 1: Fire storm }or \textit{implosion }and quickened \textit{telekinesis, 
}or summon additional demons. If the balor does not deem itself seriously threatened, 
it conserves abilities usable only once per day and uses \textit{blasphemy }instead.

\textit{Round 2: Insanity }or \textit{power word stun.}

\textit{Round 3: }Full melee attack with weapons, including entangle with whip.

\textit{Round 4: Teleport }or fly away with entangled foe to reestablish range; 
repeat round 1 and continue.

A balor who wants to drive off or neutralize a party without slaying its foes avoids 
lethal attacks.

\textit{Prior to combat: Unholy aura.}

\textit{Round 1: Dominate monster.}

\textit{Round 2: Power word stun.}

\textit{Round 3: Insanity }or \textit{telekinesis }to incapacitate or repel a dangerous 
opponent.

\textit{Round 4: Teleport }or fly away to reestablish range; repeat round 1 and 
continue.

\vspace{12pt}
BEBILITH

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{231pt}|}
\hline
  & Huge Outsider (Chaotic, Extraplanar, Evil)\tabularnewline
\hline
H\textbf{it Dice:} & 12d8+96 (150 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +5\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares), climb 20 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 22 (-2 size, +1 Dex, +13 natural), touch 9, flat-footed 
21\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +12/+29\tabularnewline
\hline
A\textbf{ttack:} & Bite +19 melee (2d6+9 plus poison) or web +11 ranged\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +19 melee (2d6+9 plus poison) and 2 claws +14 melee 
(2d4+4); or web +11 ranged\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Poison, rend armor, web\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 10/good, darkvision 60 ft., plane 
shift, scent, telepathy 100 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +16, Ref +9, Will +9\tabularnewline
\hline
A\textbf{bilities:} & Str 28, Dex 12, Con 26, Int 11, Wis 13, Cha 13\tabularnewline
\hline
S\textbf{kills:} & Climb +24, Diplomacy +3, Hide +16, Jump +28, Listen +16, Move 
Silently +16, Search +15, Sense Motive +16, Spot +16, Survival +1 (+3 following 
tracks)\tabularnewline
\hline
F\textbf{eats:} & Cleave, Improved Initiative, Improved Grapple, Power Attack, 
Track\tabularnewline
\hline
E\textbf{nvironment:} & A chaotic evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 10\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 13-18 HD (Huge); 19-36 HD (Gargantuan)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

Bebiliths are enormous, predatory, arachnid demons that hunt other demons. A bebilith 
has a body the size of a plow horse, with legs spanning more than 14 feet. It weighs 
more than two tons.

Bebiliths understand but do not speak Abyssal. Their telepathy allows them to communicate 
silently with one another.

COMBAT

A bebilith attacks any creature it sees. It usually picks one target and concentrates 
its attacks on that opponent, using its webs to isolate the target from its comrades. 
Should the bebilith become overwhelmed by tougher opponents, it often attempts 
to bite one or more of its victims and retreats, allowing its poison to do its 
work.

A bebilith's natural weapons, as well as any weapons it wields, are treated as 
chaotic-aligned and evil-aligned for the purpose of overcoming damage reduction.

\textbf{Poison (Ex):} Injury, Fortitude DC 24, initial damage 1d6 Con, secondary 
damage 2d6 Con. Bebilith venom is highly perishable, losing its potency and becoming 
inert, foul-smelling goo almost as soon as it comes into contact with air. The 
save DC is Constitution-based.

\textbf{Rend Armor (Ex): }If a bebilith hits with both claw attacks, it pulls apart 
any armor worn by its foe. This attack deals 4d6+18 points of damage to the opponent's 
armor. Creatures not wearing armor are unaffected by this special attack. Armor 
reduced to 0 hit points is destroyed. Damaged armor may be repaired with a successful 
Craft (armorsmithing) check.

\textbf{Web (Ex): }A bebilith can throw a web up to four times per day. This is 
similar to an attack with a net but has a maximum range of 30 feet, with a range 
increment of 10 feet. This attack is effective against targets of up to Gargantuan 
size. The web anchors the target in place, allowing no movement.

An entangled creature can escape with a DC 24 Escape Artist check or burst the 
web with a DC 24 Strength check. The check DCs are Constitution-based. The web 
has 14 hit points and hardness 0. There is a 75\% chance that the webbing will 
not burn if any sort of fire is applied to it (check each round).

\textbf{Plane Shift (Su):} This ability affects only the bebilith. It is otherwise 
similar to the spell (caster level 12th).

\textbf{Skills:} A bebilith has mottled coloration that gives it a +8 racial bonus 
on Hide checks.

\vspace{12pt}
DRETCH

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{200pt}|}
\hline
  & Small Outsider (Chaotic, Extraplanar, Evil)\tabularnewline
\hline
H\textbf{it Dice:} & 2d8+4 (13 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +0\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 16 (+1 size, +5 natural), touch 11, flat-footed 16\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +2/-1\tabularnewline
\hline
A\textbf{ttack:} & Claw +4 melee (1d6+1)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +4 melee (1d6+1) and bite +2 melee (1d4)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Spell-like abilities, \textit{summon demon}\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 5/cold iron or good, darkvision 
60 ft., immunity to electricity and poison, resistance to acid 10, cold 10, and 
fire 10, telepathy 100 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +5, Ref +3, Will +3\tabularnewline
\hline
A\textbf{bilities:} & Str 12, Dex 10, Con 14, Int 5, Wis 11, Cha 11\tabularnewline
\hline
S\textbf{kills:} & Hide +9, Listen +5, Move Silently +5, Spot +5, Search +2, Survival 
+0 (+2 following tracks)\tabularnewline
\hline
F\textbf{eats:} & Multiattack\tabularnewline
\hline
E\textbf{nvironment:} & A chaotic evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, gang (3-5), crowd (6-15), or mob (10-40)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 3-6 HD (Small)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +2\tabularnewline
\hline
\end{tabular}

A dretch is about 4 feet tall and weighs about 60 pounds.

Dretches cannot speak but can communicate telepathically.

Combat

Dretches are slow, stupid, and not very effective combatants. In one-on-one combat, 
they rely on their damage reduction to keep them alive. In groups, they depend 
on sheer numbers to overcome foes and immediately summon other dretches to improve 
the odds in battle. They flee at the first sign of adversity unless more powerful 
demons are present to intimidate them into fighting. Dretches' fear of their greater 
kin is stronger then even their fear of death.

A dretch's natural weapons, as well as any weapons it wields, are treated as chaotic-aligned 
and evil-aligned for the purpose of overcoming damage reduction.

\textbf{Spell-Like Abilities:} 1/day---\textit{scare }(DC 12), \textit{stinking 
cloud }(DC 13). Caster level 2nd. The save DCs are Charisma-based.

\textit{\textbf{Summon Demon }}\textbf{(Sp):} Once per day a dretch can attempt 
to summon another dretch with a 35\% chance of success. This ability is the equivalent 
of a 1st-level spell.

\textbf{Telepathy (Su): }Dretches can communicate telepathically with creatures 
within 100 feet that speak Abyssal.

\vspace{12pt}
GLABREZU

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{203pt}|}
\hline
  & Huge Outsider (Chaotic, Extraplanar, Evil)\tabularnewline
\hline
H\textbf{it Dice:} & 12d8+120 (174 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +0\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 27 (-2 size, +19 natural) touch 8, flat-footed 27\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +12/+30\tabularnewline
\hline
A\textbf{ttack:} & Pincers +20 melee (2d8+10)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 pincers +20 melee (2d8+10) and 2 claws +18 melee (1d6+5) 
and bite +18 melee (1d8+5)\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./15 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Improved grab, spell-like abilities, \textit{summon 
demon}\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 10/good, darkvision 60 ft., immunity 
to electricity and poison, resistance to acid 10, cold 10, and fire 10, spell resistance 
21, telepathy 100 ft., true seeing\tabularnewline
\hline
S\textbf{aves:} & Fort +18, Ref +8, Will +11\tabularnewline
\hline
A\textbf{bilities:} & Str 31, Dex 10, Con 31, Int 16, Wis 16, Cha 20\tabularnewline
\hline
S\textbf{kills:} & Bluff +22, Concentration +25, Diplomacy +9, Disguise +5 (+7 
acting), Intimidate +24, Knowledge (any two) +18, Listen +26, Move Silently +18, 
Search +18, Sense Motive +18, Spellcraft +18, Spot +26, Survival +3 (+5 following 
tracks)\tabularnewline
\hline
F\textbf{eats:} & Cleave, Great Cleave, Multiattack, Persuasive, Power Attack\tabularnewline
\hline
E\textbf{nvironment:} & A chaotic evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary or troupe (1 glabrezu, 1 succubus, and 2-5 vrocks)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 13\tabularnewline
\hline
T\textbf{reasure:} & Standard coins; double goods; standard items\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 13-18 HD (Huge); 19-36 HD (Gargantuan)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

Like succubi, glabrezu tempt victims into ruin, but they lure their prey with power 
or wealth rather than passion.

Glabrezu have penetrating violet eyes, and their skin color ranges from deep russet 
to pitch black. A glabrezu stands about 15 feet tall and weighs about 5,500 pounds.

Combat

Glabrezu prefer subterfuge to combat. However, if their attempts to entice or deceive 
fail, these enormous demons attack with a vengeance. They follow a \textit{confusion 
}attack with melee attacks, hoping to finish off wounded foes with \textit{chaos 
hammer }or \textit{unholy blight.}

A glabrezu's natural weapons, as well as any weapons it wields, are treated as 
chaotic-aligned and evil-aligned for the purpose of overcoming damage reduction.

\textbf{Improved Grab (Ex):} To use this ability, a glabrezu must hit a Medium 
or smaller opponent with a pincer attack. It can then attempt to start a grapple 
as a free action without provoking an attack of opportunity.

\textbf{Spell-Like Abilities:} At will---\textit{chaos hammer }(DC 19), \textit{confusion 
}(DC 19), \textit{dispel magic, mirror image, reverse gravity }(DC 22), \textit{greater 
teleport }(self plus 50 pounds of objects only), \textit{unholy blight }(DC 19); 
1/day---\textit{power word stun}. Caster level 14th. The save DCs are Charisma-based.

Once per month, a glabrezu can fulfill a \textit{wish }for a mortal humanoid. The 
demon can use this ability to offer a mortal whatever he or she desires---but unless 
the \textit{wish }is used to create pain and suffering in the world, the glabrezu 
demands either terrible evil acts or great sacrifice as compensation.

\textit{\textbf{Summon Demon }}\textbf{(Sp): }Once per day a glabrezu can attempt 
to summon 4d10 dretches or 1d2 vrocks with a 50\% chance of success, or another 
glabrezu with a 20\% chance of success. This ability is the equivalent of a 4th-level 
spell.

\textbf{True Seeing (Su): }Glabrezu continuously use true seeing as the spell (caster 
level 14th).

\textbf{Skills: }Glabrezu have a +8 racial bonus on Listen and Spot checks.

\vspace{12pt}
HEZROU

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{231pt}|}
\hline
  & Large Outsider (Chaotic, Extraplanar, Evil)\tabularnewline
\hline
H\textbf{it Dice:} & 10d8+93 (138 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +0\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 23 (-1 size, +14 natural) touch 9, flat-footed 23\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +10/+19\tabularnewline
\hline
A\textbf{ttack:} & Bite +14 melee (4d4+5)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +14 melee (4d4+5) and 2 claws +9 melee (1d8+2)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Spell-like abilities, stench, improved grab, \textit{summon 
demon}\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 10/good, darkvision 60 ft., immunity 
to electricity and poison, resistance to acid 10, cold 10, and fire 10, spell resistance 
19, telepathy 100 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +16, Ref +7, Will +9\tabularnewline
\hline
A\textbf{bilities:} & Str 21, Dex 10, Con 29, Int 14, Wis 14, Cha 18\tabularnewline
\hline
S\textbf{kills:} & Climb +18, Concentration +22, Hide +13, Escape Artist +13, Intimidate 
+17, Listen +23, Move Silently +13, Search +15, Spellcraft +15, Spot +23, Survival 
+2 (+4 following tracks), Use Rope +0 (+2 with bindings)\tabularnewline
\hline
F\textbf{eats:} & Blind-Fight, Cleave, Power Attack, Toughness\tabularnewline
\hline
E\textbf{nvironment:} & A chaotic evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary or gang (2-4)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 11\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 11-15 HD (Large); 16-30 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +9\tabularnewline
\hline
\end{tabular}

A hezrou can walk both upright and on all fours, but it always fights standing 
up

A hezrou is about 8 feet tall and weighs about 750 pounds.

Combat

Hezrous enjoy melee combat even more than vrocks do. They eagerly press an attack 
deep into the heart of enemy forces, so their stench can take effect as quickly 
as possible. They enter most battles by using \textit{blasphemy, }and follow it 
with an occasional \textit{chaos hammer }or \textit{unholy blight, }depending on 
the alignment of their opponents.

A hezrou's natural weapons, as well as any weapons it wields, are treated as chaotic-aligned 
and evil-aligned for the purpose of overcoming damage reduction.

\textbf{Improved Grab (Ex): }To use this ability, a hezrou must hit with both claw 
attacks. It can then attempt to start a grapple as a free action without provoking 
an attack of opportunity.

\textbf{Spell-Like Abilities:} At will---\textit{chaos hammer }(DC 18), \textit{greater 
teleport }(self plus 50 pounds of objects only), \textit{unholy blight }(DC 18); 
3/day---\textit{blasphemy }(DC 21), \textit{gaseous form. }Caster level 13th. The 
save DCs are Charisma-based.

\textbf{Stench (Ex): }A hezrou's skin produces a foul-smelling, toxic liquid whenever 
it fights. Any living creature (except other demons) within 10 feet must succeed 
on a DC 24 Fortitude save or be nauseated for as long as it remains within the 
affected area and for 1d4 rounds afterward. Creatures that successfully save are 
sickened for as long as they remain in the area. A creature that successfully saves 
cannot be affected again by the same hezrou's stench for 24 hours. A \textit{delay 
poison }or \textit{neutralize poison }spell removes either condition from one creature. 
Creatures that have immunity to poison are unaffected, and creatures resistant 
to poison receive their normal bonus on their saving throws.. The save DC is Constitution-based.

\textit{\textbf{Summon Demon }}\textbf{(Sp): }Once per day a hezrou can attempt 
to summon 4d10 dretches or another hezrou with a 35\% chance of success. This ability 
is the equivalent of a 4th-level spell.

\textbf{Skills:} Hezrous have a +8 racial bonus on Listen and Spot checks.

\vspace{12pt}
MARILITH

\begin{tabular}{|>{\raggedright}p{90pt}|>{\raggedright}p{236pt}|}
\hline
  & Large Outsider (Chaotic, Extraplanar, Evil)\tabularnewline
\hline
H\textbf{it Dice:} & 16d8+144 (216 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +4\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 29 (-1 size, +4 Dex, +16 natural), touch 13, flat-footed 
25\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +16/+29\tabularnewline
\hline
A\textbf{ttack:} & Longsword +25 melee (2d6+9/19-20) or slam +24 melee (1d8+9) 
or tail slap +24 melee (4d6+9)\tabularnewline
\hline
F\textbf{ull Attack:} & Primary longsword +25/+20/+15/+10 melee (2d6+9/19-20) and 
5 longswords +25 melee (2d6+4/19-20) and tail slap +22 melee (4d6+4); or 6 slams 
+24 melee (1d8+9) and tail slap +22 melee (4d6+4)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Constrict 4d6+13, improved grab, spell-like abilities, 
\textit{summon demon}\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 10/good and cold iron, darkvision 
60 ft., immunity to electricity and poison, resistance to acid 10, cold 10, and 
fire 10, spell resistance 25, telepathy 100 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +19, Ref +14, Will +14\tabularnewline
\hline
A\textbf{bilities:} & Str 29, Dex 19, Con 29, Int 18, Wis 18, Cha 24\tabularnewline
\hline
S\textbf{kills:} & Bluff +26, Concentration +28, Diplomacy +30, Disguise +7 (+9 
acting), Hide +19, Intimidate +28, Listen +31, Move Silently +23, Search 23, Sense 
Motive +23, Spellcraft +23 (+25 scrolls), Spot +31, Survival +4 (+6 following tracks), 
Use Magic Device +26 (+28 scrolls)\tabularnewline
\hline
F\textbf{eats:} & Combat Expertise, Combat Reflexes, Multiattack, Multiweapon Fighting, 
Power Attack, Weapon Focus (longsword)\tabularnewline
\hline
E\textbf{nvironment:} & A chaotic evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary or pair\tabularnewline
\hline
C\textbf{hallenge Rating:} & 17\tabularnewline
\hline
T\textbf{reasure:} & Standard coins; double goods; standard items, plus 1d4 magic 
weapons\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 17-20 HD (Large); 21-48 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A marilith usually holds a longsword in each of its six hands and wears many bangles 
and jewels.

A marilith stands about 9 feet tall and measures about 20 feet from head to tip 
of tail. It weighs about 4,000 pounds.

Combat

Though mariliths thrive on grand strategy and army-level tactics, they love physical 
combat and never pass up an opportunity to fight. Each of a marilith's six arms 
can wield a weapon, and the creature gets an additional three weapon attacks with 
its primary arm. Mariliths seldom rush headlong into battle, however, preferring 
to hang back and size up the situation first. They always seek to gain the best 
possible advantage from the local terrain, obstacles, and any vulnerability or 
weakness in their opponents.

A marilith's natural weapons, as well as any weapons it wields, are treated as 
chaotic-aligned and evil-aligned for the purpose of overcoming damage reduction.

\textbf{Constrict (Ex): }A marilith deals 4d6+13 points of damage with a successful 
grapple check. The constricted creature must succeed on a DC 27 Fortitude save 
or lose consciousness for as long as it remains in the coils and for 2d4 rounds 
thereafter. The save DC is Strength-based.

\textbf{Improved Grab (Ex): }To use this ability, a marilith must hit with its 
tail slap attack. It can then attempt to start a grapple as a free action without 
provoking an attack of opportunity. If it succeeds on the grapple check, it can 
constrict.

\textbf{Spell-Like Abilities: }At will---\textit{align weapon, blade barrier }(DC 
23), \textit{magic weapon, project image }(DC 23), \textit{polymorph, see invisibility, 
telekinesis }(DC 22), \textit{greater teleport }(self plus 50 pounds of objects 
only), \textit{unholy aura }(DC 25). Caster level 16th. The save DCs are Charisma-based.

\textit{\textbf{Summon Demon }}\textbf{(Sp):} Once per day a marilith can attempt 
to summon 4d10 dretches, 1d4 hezrou, or one nalfeshnee with a 50\% chance of success, 
or one glabrezu or another marilith with a 20\% chance of success. This ability 
is the equivalent of a 5th-level spell.

\textbf{True Seeing (Su):} Mariliths continuously use this ability, as the spell 
(caster level 16th).

\textbf{Skills:} Mariliths have a +8 racial bonus on Listen and Spot checks.

\textbf{Feats: }In combination with its natural abilities, a marilith's Multiweapon 
Fighting feat allows it to attack with all its arms at no penalty.

\vspace{12pt}
NALFESHNEE

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{224pt}|}
\hline
  & Huge Outsider (Chaotic, Extraplanar, Evil)\tabularnewline
\hline
H\textbf{it Dice:} & 14d8+112 (175 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares), fly 40 ft. (poor)\tabularnewline
\hline
A\textbf{rmor Class:} & 27 (-2 size, +1 Dex, +18 natural) touch 9, flat-footed 
26\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +14/+29\tabularnewline
\hline
A\textbf{ttack:} & Bite +20 melee (2d8+7)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +20 melee (2d8+7) and 2 claws +17 melee (1d8+3)\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./15 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Smite, spell-like abilities, \textit{summon demon}\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 10/good, darkvision 60 ft., immunity 
to electricity and poison, resistance to acid 10, cold 10, and fire 10, spell resistance 
22, telepathy 100 ft., true seeing\tabularnewline
\hline
S\textbf{aves:} & Fort +17, Ref +10, Will +15\tabularnewline
\hline
A\textbf{bilities:} & Str 25, Dex 13, Con 27, Int 22, Wis 22, Cha 20\tabularnewline
\hline
S\textbf{kills:} & Bluff +22, Concentration +25, Diplomacy +26, Disguise +5 (+7 
acting), Hide +10, Intimidate +22, Knowledge (arcana) +23, Listen +31, Move Silently 
+18, Search +23, Sense Motive +23, Spellcraft +25 (+27 scrolls), Spot +31, Survival 
+6 (+8 following tracks), Use Magic Device +22 (+24 scrolls)\tabularnewline
\hline
F\textbf{eats:} & Cleave, Improved Bull Rush, Multiattack, Power Attack, Weapon 
Focus (bite)\tabularnewline
\hline
E\textbf{nvironment:} & A chaotic evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary or troupe (1 nalfeshnee, 1 hezrou, and 2-5 vrocks)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 14\tabularnewline
\hline
T\textbf{reasure:} & Standard coins; double goods; standard items\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 15-20 HD (Huge); 21-42 HD (Gargantuan)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A nalfeshnee can fly despite its small wings.

A nalfeshnee is more than 20 feet tall and weighs 8,000 pounds.

Combat

When fulfilling their duties in the underworld, nalfeshnees usually disdain combat 
as being beneath them. Given the opportunity, they succumb to blood lust and do 
battle. They disable opponents with their smite ability and slaughter them while 
they can't fight back.

A nalfeshnee's natural weapons, as well as any weapons it wields, are treated as 
chaotic-aligned and evil-aligned for the purpose of overcoming damage reduction.

\textbf{Smite (Su): }Three times per day a nalfeshnee can create a nimbus of unholy 
light. When the demon triggers the ability, rainbow-colored beams play around its 
body. One round later they burst in a 60-foot radius. Any creature within this 
area must succeed on a DC 22 Will save or be dazed for 1d10 rounds as visions of 
its worst fears hound it. The creature receives its full Dexterity and shield bonuses 
to AC if attacked but can take no actions. Other demons are immune to this effect. 
The save DC is Charisma-based.

\textbf{Spell-Like Abilities: }At will---\textit{call lightning }(DC 18), \textit{feeblemind 
}(DC 20), \textit{greater dispel magic, slow }(DC 18), \textit{greater teleport 
}(self plus 50 pounds of objects only), \textit{unholy aura }(DC 23). Caster level 
12th. The save DCs are Charisma-based.

\textit{\textbf{Summon Demon }}\textbf{(Sp): }Twice per day a nalfeshnee can attempt 
to summon 1d4 vrocks, 1d4 hezrous, or one glabrezu with a 50\% chance of success, 
or another nalfeshnee with a 20\% chance of success. This ability is the equivalent 
of a 5th-level spell.

\textbf{True Seeing (Su): }Nalfeshnees continuously use true seeing, as the spell 
(caster level 14th).

\textbf{Skills: }Nalfeshnees have a +8 racial bonus on Listen and Spot checks.

\vspace{12pt}
QUASIT

\begin{tabular}{|>{\raggedright}p{88pt}|>{\raggedright}p{237pt}|}
\hline
  & Tiny Outsider (Chaotic, Extraplanar, Evil)\tabularnewline
\hline
H\textbf{it Dice:} & 3d8 (13 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +7\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), fly 50 ft. (perfect)\tabularnewline
\hline
A\textbf{rmor Class:} & 18 (+2 size, +3 Dex, +3 natural), touch 15, flat-footed 
15\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +3/-6\tabularnewline
\hline
A\textbf{ttack:} & Claw +8 melee (1d3-1 plus poison)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +8 melee (1d3-1 plus poison) and bite +3 melee 
(1d4-1)\tabularnewline
\hline
S\textbf{pace/Reach:} & 2-1/2 ft./0 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Poison, spell-like abilities\tabularnewline
\hline
S\textbf{pecial Qualities:} & Alternate form, damage reduction 5/cold iron or good, 
darkvision 60 ft., fast healing 2, immunity to poison, resistance to fire 10\tabularnewline
\hline
S\textbf{aves:} & Fort +3, Ref +6, Will +4\tabularnewline
\hline
A\textbf{bilities:} & Str 8, Dex 17, Con 10, Int 10, Wis 12, Cha 10\tabularnewline
\hline
S\textbf{kills:} & Bluff +6, Diplomacy +2, Disguise +0 (+2 acting), Hide +17, Intimidate 
+2, Knowledge (any one) +6, Listen +7, Move Silently +9, Search +6, Spellcraft 
+6, Spot +6 \tabularnewline
\hline
F\textbf{eats:} & Improved Initiative, Weapon Finesse\tabularnewline
\hline
E\textbf{nvironment:} & A chaotic evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 4-6 HD (Tiny)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- &  (Improved Familiar)\tabularnewline
\hline
\end{tabular}

In its natural form, a quasit stands about 1-1/2 feet tall and weighs about 8 pounds.

Quasits speak Common and Abyssal.

Combat

Although quasits thirst for victory and power as other demons do, they are cowards 
at heart. They typically attack from ambush, using their alternate form ability 
and \textit{invisibility }to get within reach, then try to scuttle away. When retreating, 
they use their \textit{cause fear }ability to deter pursuit. 

A quasit's natural weapons, as well as any weapons it wields, are treated as chaotic-aligned 
and evil-aligned for the purpose of overcoming damage reduction.

\textbf{Poison (Ex): }Injury, Fortitude DC 13, initial damage 1d4 Dex, secondary 
damage 2d4 Dex. The save DC is Constitution-based and includes a +2 racial bonus.

\textbf{Spell-Like Abilities: }At will---\textit{detect good, detect magic, }and 
\textit{invisibility }(self only); 1/day---\textit{cause fear }(as the spell, except 
that its area is a 30-foot radius from the quasit, save DC 11). Caster level 6th.

The save DCs are Charisma-based.

Once per week a quasit can use \textit{commune }to ask six questions. The ability 
otherwise works as the spell (caster level 12th).

\textbf{Alternate Form (Su): }A quasit can assume other forms at will as a standard 
action. This ability functions as a \textit{polymorph }spell cast on itself (caster 
level 12th), except that a quasit does not regain hit points for changing form, 
and any individual quasit can assume only one or two forms no larger than Medium. 
Common forms include bat, monstrous centipede, toad, and wolf. A quasit in alternate 
form loses its poison attack.

\vspace{12pt}
RETRIEVER

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{231pt}|}
\hline
  & Huge Construct (Extraplanar)\tabularnewline
\hline
H\textbf{it Dice:} & 10d10+80 (135 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +3\tabularnewline
\hline
S\textbf{peed:} & 50 ft. (10 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 21 (-2 size, +3 Dex, +10 natural), touch 11, flat-footed 
18\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +7/+25\tabularnewline
\hline
A\textbf{ttack:} & Claw +15 melee (2d6+10) and eye ray +8 ranged touch\tabularnewline
\hline
F\textbf{ull Attack:} & 4 claws +15 melee (2d6+10) and bite +10 melee (1d8+5) and 
eye ray +8 ranged touch\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Eye rays, \textit{find target, }improved grab\tabularnewline
\hline
S\textbf{pecial Qualities:} & Construct traits, darkvision 60 ft., fast healing 
5, low-light vision\tabularnewline
\hline
S\textbf{aves:} & Fort +3, Ref +6, Will +3\tabularnewline
\hline
A\textbf{bilities:} & Str 31, Dex 17, Con ---, Int ---, Wis 11, Cha 1\tabularnewline
\hline
S\textbf{kills:}--- & \tabularnewline
\hline
F\textbf{eats:}--- & \tabularnewline
\hline
E\textbf{nvironment:} & A chaotic evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 11\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 11-15 HD (Huge); 16-30 HD (Gargantuan)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A retriever specializes in recovering lost or desired objects, runaway slaves, 
and enemies and bringing them back to its master. A retriever has a body the size 
of an ox, with legs spanning more than 14 feet. It weighs about 6,500 pounds.

Combat

Retrievers attack with four claws, but their eye rays are far more deadly.

\textbf{Eye Rays (Su): }A retriever's eyes can produce four different magical rays 
with a range of 100 feet. Each round, it can fire one ray as a free action. A particular 
ray is usable only once every 4 rounds. A retriever can fire an eye ray in the 
same round that it makes physical attacks. The save DC for all rays is 18. The 
save DC is Dexterity-based.

The four eye effects are:

\textit{Fire: }Deals 12d6 points of fire damage to the target (Reflex half ).

\textit{Cold: }Deals 12d6 points of cold damage to the target (Reflex half ).

\textit{Electricity: }Deals 12d6 points of electricity damage to the target (Reflex 
half ).

\textit{Petrification: }The target must succeed on a Fortitude save or turn to 
stone permanently.

\textit{\textbf{Find Target }}\textbf{(Sp):} When ordered to find an item or a 
creature, a retriever does so unerringly, as though guided by \textit{discern location.}

The being giving the order must have seen (or must have an item belonging to) the 
creature to be found, or must have touched the object to be located. This ability 
is the equivalent of an 8th-level spell.

\textbf{Improved Grab (Ex): }To use this ability, a retriever must hit with its 
bite attack. It can then attempt to start a grapple as a free action without provoking 
an attack of opportunity. If it wins the grapple check, it establishes a hold and 
grips the opponent fast in its mouth. This is the method by which it usually ``retrieves'' 
things.

\vspace{12pt}
SUCCUBUS

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{215pt}|}
\hline
  & Medium Outsider (Chaotic, Extraplanar, Evil)\tabularnewline
\hline
H\textbf{it Dice:} & 6d8+6 (33 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares), fly 50 ft. (average)\tabularnewline
\hline
A\textbf{rmor Class:} & 20 (+1 Dex, +9 natural), touch 11, flat-footed 19\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +6/+7\tabularnewline
\hline
A\textbf{ttack:} & Claw +7 melee (1d6+1)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +7 melee (1d6+1)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Energy drain, spell-like abilities, \textit{summon 
demon}\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 10/cold iron or good, darkvision 
60 ft., immunity to electricity and poison, resistance to acid 10, cold 10, and 
fire 10, spell resistance 18, telepathy 100 ft., tongues\tabularnewline
\hline
S\textbf{aves:} & Fort +6, Ref +6, Will +7\tabularnewline
\hline
A\textbf{bilities:} & Str 13, Dex 13, Con 13, Int 16, Wis 14, Cha 26\tabularnewline
\hline
S\textbf{kills:} & Bluff +19, Concentration +10, Diplomacy +12, Disguise +17* (+19 
acting), Escape Artist +10, Hide +10, Intimidate +19, Knowledge (any one) +12, 
Listen +19, Move Silently +10, Search +12, Spot +19, Survival +2 (+4 following 
tracks), Use Rope +1 (+3 with bindings)\tabularnewline
\hline
F\textbf{eats:} & Dodge, Mobility, Persuasive\tabularnewline
\hline
E\textbf{nvironment:} & A chaotic evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 7\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 7-12 HD (Medium)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +6\tabularnewline
\hline
\end{tabular}

A succubus is 6 feet tall in its natural form and weighs about 125 pounds.

Combat

Succubi are not warriors. They flee combat whenever they can. If forced to fight, 
they can attack with their claws, but they prefer to turn foes against one another. 
Succubi use their \textit{polymorph }ability to assume humanoid guise, and can 
maintain this deception indefinitely. Their preferred tactic when dealing with 
heroes is to feign friendship and create an opportunity to be alone with one of 
them, whereupon the succubus applies her life-draining kiss. Succubi are not above 
taking on the role of a damsel in distress when encountered within a dungeon.

A succubus's natural weapons, as well as any weapons it wields, are treated as 
chaotic-aligned and evil-aligned for the purpose of overcoming damage reduction.

\textbf{Energy Drain (Su): }A succubus drains energy from a mortal it lures into 
some act of passion, or by simply planting a kiss on the victim. If the target 
is not willing to be kissed, the succubus must start a grapple, which provokes 
an attack of opportunity. The succubus's kiss or embrace bestows one negative level. 
The kiss also has the effect of a \textit{suggestion }spell, asking the victim 
to accept another kiss from the succubus. The victim must succeed on a DC 21 Will 
save to negate the effect of the \textit{suggestion}. The DC is 21 for the Fortitude 
save to remove a negative level. These save DCs are Charisma-based.

\textbf{Spell-Like Abilities: }At will---\textit{charm monster }(DC 22), \textit{detect 
good, detect thoughts }(DC 20), \textit{ethereal jaunt }(self plus 50 pounds of 
objects only), \textit{polymorph }(humanoid form only, no limit on duration), \textit{suggestion 
}(DC 21), \textit{greater teleport }(self plus 50 pounds of objects only). Caster 
level 12th. The save DCs are Charisma-based.

\textit{\textbf{Summon Demon }}\textbf{(Sp):} Once per day a succubus can attempt 
to summon 1 vrock with a 30\% chance of success. This ability is the equivalent 
of a 3rd-level spell.

\textbf{Tongues (Su): }A succubus has a permanent tongues ability (as the spell, 
caster level 12th). Succubi usually use verbal communication with mortals.

\textbf{Skills:} Succubi have a +8 racial bonus on Listen and Spot checks.

*While using her \textit{polymorph }ability, a succubus gains a +10 circumstance 
bonus on Disguise checks.

\vspace{12pt}
VROCK

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{231pt}|}
\hline
  & Large Outsider (Chaotic, Extraplanar, Evil)\tabularnewline
\hline
H\textbf{it Dice:} & 10d8+70 (115 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +2\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares), fly 50 ft. (average)\tabularnewline
\hline
A\textbf{rmor Class:} & 22 (-1 size, +2 Dex, +11 natural), touch 11, flat-footed 
20\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +10/+20\tabularnewline
\hline
A\textbf{ttack:} & Claw +15 melee (2d6+6)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +15 melee (2d6+6) and bite +13 melee (1d8+3) and 
2 talons +13 melee (1d6+3)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Dance of ruin, spell-like abilities, spores, stunning 
screech, \textit{summon demon}\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 10/good, darkvision 60 ft., immunity 
to electricity and poison, resistance to acid 10, cold 10, and fire 10, spell resistance 
17, telepathy 100 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +14, Ref +9, Will +10\tabularnewline
\hline
A\textbf{bilities:} & Str 23, Dex 15, Con 25, Int 14, Wis 16, Cha 16\tabularnewline
\hline
S\textbf{kills:} & Concentration +20, Diplomacy +5, Hide +11, Intimidate +16, Knowledge 
(any one) +15, Listen +24, Move Silently +15, Search +15, Sense Motive +16, Spellcraft 
+15, Spot +24, Survival +3 (+5 following tracks)\tabularnewline
\hline
F\textbf{eats:} & Cleave, Combat Reflexes, Multiattack, Power Attack\tabularnewline
\hline
E\textbf{nvironment:} & A chaotic evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, gang (3-5), or squad (6-10)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 9\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 11-14 HD (Large); 15-30 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +8\tabularnewline
\hline
\end{tabular}

A vrock is about 8 feet tall and weighs about 500 pounds.

Combat

Vrocks are vicious fighters who like to fly down into the enemy and cause as much 
damage as possible. They prance about in battle, taking briefly to the air and 
bringing their clawed feet into play. Despite their advantage in mobility, the 
vrocks' deep love of battle frequently leads them into melee combats against heavy 
odds.

A vrock's natural weapons, as well as any weapons it wields, are treated as chaotic-aligned 
and evil-aligned for the purpose of overcoming damage reduction.

\textbf{Dance of Ruin (Su):} To use this ability, a group of at least three vrocks 
must join hands in a circle, dancing wildly and chanting. 

At the end of 3 rounds of dancing, a wave of crackling energy flashes outward in 
a 100-foot radius. All creatures except for demons within the radius take 20d6 
points of damage (Reflex DC 18 half ). Stunning, paralyzing, or slaying one of 
the vrocks stops the dance. The save DC is Charisma-based.

\textbf{Spell-Like Abilities:} At will---\textit{mirror image, telekinesis }(DC 
18), \textit{greater teleport }(self plus 50 pounds of objects only); 1/day---\textit{heroism. 
}Caster level 12th. The save DCs are Charisma-based.

\textbf{Spores (Ex): }A vrock can release masses of spores from its body once every 
3 rounds as a free action. The spores automatically deal 1d8 points of damage to 
all creatures adjacent to the vrock. They then penetrate the skin and grow, dealing 
an additional 1d4 points of damage each round for 10 rounds. At the end of this 
time, the victim is covered with a tangle of viny growths. (The vines are harmless 
and wither away in 1d4 days.) A \textit{delay poison }spell stops the spores' growth 
for its duration. \textit{Bless, neutralize poison, }or \textit{remove disease 
}kills the spores, as does sprinkling the victim with a vial of holy water.

\textbf{Stunning Screech (Su):} Once per hour a vrock can emit a piercing screech. 
All creatures except for demons within a 30-foot radius must succeed on a DC 22 
Fortitude save or be stunned for 1 round. The save DC is Constitution-based.

\textit{\textbf{Summon Demon }}\textbf{(Sp): }Once per day a vrock can attempt 
to summon 2d10 dretches or another vrock with a 35\% chance of success. This ability 
is the equivalent of a 3rd-level spell.

\textbf{Skills:} Vrocks have a +8 racial bonus on Listen and Spot checks.

\vspace{12pt}
{\LARGE{}DERRO}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{204pt}|}
\hline
  & Small Monstrous Humanoid\tabularnewline
\hline
H\textbf{it Dice:} & 3d8+3 (16 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +6\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 19 (+1 size, +2 Dex, +2 natural, +3 studded leather armor, 
+1 buckler), touch 13, flat-footed 17\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +3/-1\tabularnewline
\hline
A\textbf{ttack:} & Short sword +4 melee (1d4/19-20) or repeating light crossbow 
+6 ranged (1d6/19-20 plus poison)\tabularnewline
\hline
F\textbf{ull Attack:} & Short sword +4 melee (1d4/19-20) or repeating light crossbow 
+6 ranged (1d6/19-20 plus poison)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Poison use, spell-like abilities, sneak attack +1d6\tabularnewline
\hline
S\textbf{pecial Qualities:} & Madness, spell resistance 15 vulnerability to sunlight\tabularnewline
\hline
S\textbf{aves:} & Fort +2, Ref +5, Will +6\tabularnewline
\hline
A\textbf{bilities:} & Str 11, Dex 14, Con 13, Int 10, Wis 5*, Cha 16*\tabularnewline
\hline
S\textbf{kills:} & Bluff +5, Hide +10, Listen +1, Move Silently +8\tabularnewline
\hline
F\textbf{eats:} & Blind-Fight, Improved Initiative\tabularnewline
\hline
E\textbf{nvironment:} & Underground\tabularnewline
\hline
O\textbf{rganization:} & Team (2-4), squad (5-8 plus 1 3rd-level sorcerer), or 
band (11-20 plus 30\% noncombatants plus 3 3rd-level sorcerers and 1 sorcerer of 
5th-8th level)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 3\tabularnewline
\hline
T\textbf{reasure:} & Standard coins; double goods; standard items\tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & By character class\tabularnewline
\hline
L\textbf{evel Adjustment:}--- &  (+2 if sane)\tabularnewline
\hline
\end{tabular}

Derro are afflicted by a form of racial madness, which most often manifests as 
delusions of grandeur coupled with an overpowering urge to inflict torment on other 
creatures. Derro are capable of holding their murderous impulses in check for short 
periods of time in order to cooperate with creatures of other races, but such arrangements 
rarely last more than a few weeks. Of course, no derro is capable of recognizing 
that he is out of his mind.

COMBAT

Derro are stealthy and bloodthirsty. They like to carefully arrange cruel traps 
and deadly ambushes, and strike savagely from hiding. They delight in taking captives 
who can be tortured to death later, and favor traps and poisons that disable without 
killing.

\textbf{Madness (Ex):} Derro use their Charisma modifier on Will saves instead 
of their Wisdom modifier, and have immunity to \textit{confusion }and \textit{insanity 
}effects. A derro cannot be restored to sanity by any means short of a \textit{miracle 
}or \textit{wish }spell.

*The racial madness of the derro provides a +6 bonus to their Charisma scores and 
a -6 penalty to their Wisdom scores. A derro restored to sanity gains 6 points 
of Wisdom and loses 6 points of Charisma.

\textbf{Poison Use (Ex): }Derro typically carry 2d4 doses of greenblood oil or 
Medium monstrous spider venom, applying it to their crossbow bolts. Derro are not 
at risk of poisoning themselves when handling poison.

\textbf{Sneak Attack (Ex):} Any time a derro's opponent is denied his Dexterity 
bonus to AC, or if a derro flanks his opponent, he deals an extra 1d6 points of 
damage. This ability is just like the rogue's sneak attack and subject to the same 
limitations.

\textbf{Spell-Like Abilities: }At will---\textit{darkness, ghost sound; }1/day--- 
\textit{daze }(DC 13)\textit{, sound burst }(DC 15). Caster level 3rd. The save 
DCs are Charisma-based.

\textbf{Vulnerability to Sunlight (Ex):} A derro takes 1 point of Constitution 
damage for every hour it is exposed to sunlight, and it dies if its Constitution 
score reaches 0. Lost Constitution points are recovered at the rate of 1 per every 
24-hour period spent underground or otherwise sheltered from the sun.

\textbf{Skills:} Derro have a +4 racial bonus on Hide and Move Silently checks.

\vspace{12pt}
{\LARGE{}DESTRACHAN}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{226pt}|}
\hline
  & Large Aberration\tabularnewline
\hline
H\textbf{it Dice:} & 8d8+24 (60 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +5\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 18 (-1 size, +1 Dex, +8 natural), touch 10, flat-footed 
17\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +6/+14\tabularnewline
\hline
A\textbf{ttack:} & Claw +9 melee (1d6+4)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +9 melee (1d6+4)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Destructive harmonics\tabularnewline
\hline
S\textbf{pecial Qualities:} & Blindsight 100 ft., immunities, protection from sonics\tabularnewline
\hline
S\textbf{aves:} & Fort +5, Ref +5, Will +10\tabularnewline
\hline
A\textbf{bilities:} & Str 18, Dex 12, Con 16, Int 12, Wis 18, Cha 12\tabularnewline
\hline
S\textbf{kills:} & Hide +8, Listen +25, Move Silently +7, Survival +9\tabularnewline
\hline
F\textbf{eats:} & Dodge, Improved Initiative, Lightning Reflexes\tabularnewline
\hline
E\textbf{nvironment:} & Underground\tabularnewline
\hline
O\textbf{rganization:} & Solitary or pack (3-5)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 8\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Usually neutral evil\tabularnewline
\hline
A\textbf{dvancement:} & 9-16 HD (Large); 17-24 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

The dungeon-dwelling destrachan looks like some bizarre, nonintelligent beast, 
but it's an incredibly evil and crafty sadist. A destrachan has a pair of complex, 
three-part ears that it can adjust to be more or less sensitive to various sounds. 
It is blind, yet hunts with a sense of hearing more precise than most creatures' 
sight.

From its tubular mouth a destrachan emits carefully focused harmonics, producing 
sonic energy so powerful it can shatter a stone wall. So skilled is a destrachan 
at controlling the sounds it emits that it can choose what type of material to 
affect with its attack.

A destrachan is about 10 feet long from its mouth to the tip of the tail and weighs 
about 4,000 pounds.

A destrachan speaks no language but understands Common. If a destrachan must communicate, 
it does so through action.

COMBAT

A destrachan uses its claws only as a last resort or to finish off foes weakened 
by its sonic attacks. It often enters battle with surprise if possible. It first 
focuses on destroying metal armor and weapons and then changes to harmonics that 
disrupt flesh.

\textbf{Destructive Harmonics (Su): }A destrachan can blast sonic energy in a cone 
up to 80 feet long. It can also use this attack to affect any creatures or objects 
within a 30-foot radius. It can tune the harmonics of this destructive power to 
affect different types of targets. All save DCs are Charisma-based.

\textit{Flesh: }Disrupting tissue and rending bone, this attack deals 4d6 points 
of damage to all within the area (Reflex DC 15 half ).

\textit{Nerves: }A destrachan can focus its harmonics to knock out foes rather 
than slay them. This attack deals 6d6 points of nonlethal damage to all within 
the area (Reflex DC 15 half ).

\textit{Material: }When using this form of harmonics, a destrachan chooses wood, 
stone, metal, or glass. All objects made of that material within the area must 
succeed on a DC 15 Fortitude save or shatter. Objects (or portions of objects) 
that have up to 30 hit points are potentially affected by this attack.

\textbf{Blindsight (Ex): }A destrachan can use hearing to ascertain all foes within 
100 feet as a sighted creature would.

\textbf{Immunities:} Destrachans have immunity to gaze attacks, visual effects, 
illusions, and other attack forms that rely on sight.

\textbf{Protection from Sonics (Ex):} While they can be affected by loud noises 
and sonic spells (such as \textit{ghost sound }or \textit{silence}), destrachans 
are less vulnerable to sonic attacks (+4 circumstance bonus on all saves) because 
they can protect their ears. A destrachan whose sense of hearing is impaired is 
effectively blinded, and all targets are treated as having total concealment.

\textbf{Skills: }A destrachan has a +10 racial bonus on Listen checks.

\vspace{12pt}
{\LARGE{}DEVIL}

Devils are fiends from lawful evil-aligned planes. 

Many devils are surrounded by a fear aura, which they use to break up powerful 
groups and defeat opponents piecemeal. Devils with spell-like abilities use their 
illusion abilities to delude and confuse foes as much as possible. A favorite trick 
is to create illusory reinforcements; enemies can never be entirely sure if a threat 
is only a figment or real summoned devils joining the fray.

\textbf{Devil Traits: }Most devils possesses the following traits (unless otherwise 
noted in a creature's entry).---

Immunity to fire and poison.---

Resistance to acid 10 and cold 10.---

See in Darkness (Su): Some devils can see perfectly in darkness of any kind, even 
that created by a \textit{deeper darkness }spell.---

\textit{Summon }(Sp): Some devils share the ability to summon others of their kind 
(the success chance and type of devils summoned are noted in each monster description).---

Telepathy. 

Except when otherwise noted, devils speak Infernal, Celestial, and Draconic.

\vspace{12pt}
BARBED DEVIL (HAMATULA)

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{209pt}|}
\hline
  & Medium Outsider (Evil, Extraplanar, Lawful)\tabularnewline
\hline
H\textbf{it Dice:} & 12d8+72 (126 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +6\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 29 (+6 Dex, +13 natural) touch 16, flat-footed 23\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +12/+22\tabularnewline
\hline
A\textbf{ttack:} & Claw +18 melee (2d8+6 plus fear)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +18 melee (2d8+6 plus fear)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Fear, improved grab, impale 3d8+9, \textit{summon devil}\tabularnewline
\hline
S\textbf{pecial Qualities:} & Barbed defense, damage reduction 10/good, darkvision 
60 ft., immunity to fire and poison, resistance to acid 10 and cold 10, see in 
darkness, spell resistance 23, spell-like abilities, telepathy 100 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +14, Ref +14, Will +12\tabularnewline
\hline
A\textbf{bilities:} & Str 23, Dex 23, Con 23, Int 12, Wis 14, Cha 18\tabularnewline
\hline
S\textbf{kills:} & Concentration +21, Diplomacy +6, Hide +21, Intimidate +19, Knowledge 
(any one) +16, Listen +19, Move Silently +21, Search +16, Sense Motive +17, Spot 
+19, Survival +2 (+4 following tracks)\tabularnewline
\hline
F\textbf{eats:} & Alertness, Cleave, Improved Grapple, Iron Will, Power Attack\tabularnewline
\hline
E\textbf{nvironment:} & A lawful evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, team (3-5), or squad (6-10)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 11\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & 13-24 (Medium); 25-36 HD (Large)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A barbed devil is about 7 feet tall and weighs about 300 pounds.

Combat

Barbed devils eagerly fight with their claws, trying to impale their opponents. 
They use \textit{hold person }to immobilize those who avoid their hug attacks.

A barbed devil's natural weapons, as well as any weapons it wields, are treated 
as evil-aligned and lawful-aligned for the purpose of overcoming damage reduction.

\textbf{Fear (Su):} A creature hit by a barbed devil must succeed on a DC 20 Will 
save or be affected as though by \textit{fear }(caster level 9th). 

Whether or not the save is successful, that creature cannot be affected by that 
same barbed devil's fear ability for 24 hours. The save DC is Charisma-based.

\textbf{Impale (Ex):} A barbed devil deals 3d8+9 points of piercing damage to a 
grabbed opponent with a successful grapple check. 

\textbf{Improved Grab (Ex):} To use this ability, a barbed devil must hit with 
a claw attack. It can then attempt to start a grapple as a free action without 
provoking an attack of opportunity. If it wins the grapple check, it establishes 
a hold and can impale the opponent on its barbed body.

\textit{\textbf{Summon Devil }}\textbf{(Sp):} Once per day a barbed devil can attempt 
to summon 1d6 bearded devils or another barbed devil with a 35\% chance of success. 
This ability is the equivalent of a 4th-level spell.

\textbf{Barbed Defense (Su):} Any creature striking a barbed devil with handheld 
weapons or natural weapons takes 1d8+6 points of piercing and slashing damage from 
the devil's barbs. Note that weapons with exceptional reach, such as longspears, 
do not endanger their users in this way.

\textbf{Spell-Like Abilities:} At will---\textit{greater teleport }(self plus 50 
pounds of objects only), \textit{hold person }(DC 16), \textit{major image }(DC 
17), \textit{scorching ray }(2 rays only). 1/day---\textit{order's wrath }(DC 18), 
\textit{unholy blight }(DC 18). Caster level 12th. The save DCs are Charisma-based.

\vspace{12pt}
BEARDED DEVIL (BARBAZU)

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{210pt}|}
\hline
  & Medium Outsider (Evil, Extraplanar, Lawful)\tabularnewline
\hline
H\textbf{it Dice:} & 6d8+18 (45 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +6\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 19 (+2 Dex, +7 natural) touch 12, flat-footed 17\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +6/+8\tabularnewline
\hline
A\textbf{ttack:} & Glaive +9 melee (1d10+3 plus infernal wound) or claw +8 melee 
(1d6+2)\tabularnewline
\hline
F\textbf{ull Attack:} & Glaive +9/+4 melee (1d10+3 plus infernal wound) or 2 claws 
+8 melee (1d6+2)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. (10 ft. with glaive)\tabularnewline
\hline
S\textbf{pecial Attacks:} & Infernal wound, beard, battle frenzy, \textit{summon 
devil}\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 5/silver or good, darkvision 60 
ft., immunity to fire and poison, resistance to acid 10 and cold 10, see in darkness, 
spell resistance 17, telepathy 100 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +8, Ref +7, Will +5\tabularnewline
\hline
A\textbf{bilities:} & Str 15, Dex 15, Con 17, Int 6, Wis 10, Cha 10\tabularnewline
\hline
S\textbf{kills:} & Climb +11, Diplomacy +2, Hide +11, Listen +9, Move Silently 
+9, Sense Motive +9, Spot +9\tabularnewline
\hline
F\textbf{eats:} & Improved Initiative, Power Attack, Weapon Focus (glaive)\tabularnewline
\hline
E\textbf{nvironment:} & A lawful evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, team (3-5), or squad (6-10)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 5\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & 7-9 HD (Medium); 10-18 HD (Large)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +6\tabularnewline
\hline
\end{tabular}

Every bearded devil carries a sawtoothed glaive.

A bearded devil stands 6 feet tall and weighs about 225 pounds.

Combat

Bearded devils are aggressive and love to fight. They revel in their battle frenzy, 
spreading mayhem among their foes. 

A bearded devil's natural weapons, as well as any weapons it wields, are treated 
as evil-aligned and lawful-aligned for the purpose of overcoming damage reduction.

\textbf{Spell-Like Abilities: }At will--- \textit{greater teleport }(self plus 
50 pounds of objects only). Caster level 12th.

\textbf{Infernal Wound (Su):} The damage a bearded devil deals with its glaive 
causes a persistent wound. An injured creature loses 2 additional hit points each 
round. The wound does not heal naturally and resists healing spells. The continuing 
hit point loss can be stopped by a DC 16 Heal check, a \textit{cure }spell, or 
a \textit{heal }spell. However, a character attempting to cast a \textit{cure }spell 
or a \textit{heal }spell on a creature damaged by a bearded devil's glaive must 
succeed on a DC 16 caster level check, or the spell has no effect on the injured 
character.

A successful Heal check automatically stops the continuing hit point loss as well 
as restoring hit points. The infernal wound is a supernatural ability of the bearded 
devil, not of the weapon. The check DC is Constitution-based.

\textbf{Beard (Ex):} If a bearded devil hits a single opponent with both claw attacks, 
it automatically hits with its beard. The affected creature takes 1d8+2 points 
of damage and must succeed on a DC 16 Fortitude save or be infected with a vile 
disease known as devil chills (incubation period 1d4 days, damage 1d4 Str). Damage 
is dealt each day until the afflicted creature succeeds on three consecutive Fortitude 
saves, the disease is cured magically, or the creature dies. The save DC is Constitution-based.

\textbf{Battle Frenzy (Ex):} Twice per day, a bearded devil can work itself into 
a battle frenzy similar to the barbarian's rage (+4 Strength, +4 Constitution, 
+2 morale bonus on Will saves, -2 AC penalty). The frenzy lasts for 6 rounds, and 
the bearded devil suffers no ill effects afterward.

\textit{\textbf{Summon Devil }}\textbf{(Sp): }Once per day a bearded devil can 
attempt to summon 2d10 lemures with a 50\% chance of success, or another bearded 
devil with a 35\% chance of success. This ability is the equivalent of a 3rd-level 
spell.

\vspace{12pt}
BONE DEVIL (OSYLUTH)

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{231pt}|}
\hline
  & Large Outsider (Evil, Extraplanar, Lawful)\tabularnewline
\hline
H\textbf{it Dice:} & 10d8+50 (95 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +9\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 25 (-1 size, +5 Dex, +11 natural) touch 14, flat-footed 
20\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +10/+19\tabularnewline
\hline
A\textbf{ttack:} & Bite +14 melee (1d8+5)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +14 melee (1d8+5) and 2 claws +12 melee (1d4+2) and 
sting +12 melee (3d4+2 plus poison)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Spell-like abilities, fear aura, poison, \textit{summon 
devil}\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 10/good, darkvision 60 ft., immunity 
to fire and poison, resistance to acid 10 and cold 10, see in darkness, spell resistance 
21, telepathy 100 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +12, Ref +12, Will +11\tabularnewline
\hline
A\textbf{bilities:} & Str 21, Dex 21, Con 21, Int 14, Wis 14, Cha 14\tabularnewline
\hline
S\textbf{kills:} & Bluff +15, Concentration +18, Diplomacy +6, Disguise +2 (+4 
acting), Hide +14, Intimidate +17, Knowledge (any one) +15, Listen +17, Move Silently 
+18, Search +15, Sense Motive +15, Spot +17, Survival +2 (+4 following tracks)\tabularnewline
\hline
F\textbf{eats:} & Alertness, Improved Initiative, Iron Will, Multiattack\tabularnewline
\hline
E\textbf{nvironment:} & A lawful evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary, team (2-4), or squad (6-10)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 9\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & 11-20 HD (Large); 21-30 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

Bone devils stand about 9 feet tall and weigh about 500 pounds.

Combat

Bone devils hate all other creatures and attack ruthlessly. They freely use \textit{wall 
of ice }to keep the enemy divided.

A bone devil's natural weapons, as well as any weapons it wields, are treated as 
evil-aligned and lawful-aligned for the purpose of overcoming damage reduction.

\textbf{Fear Aura (Su):} Bone devils can radiate a 5-foot-radius fear aura as a 
free action. Affected creatures must succeed on a DC 17 Will save or be affected 
as though by a \textit{fear }spell (caster level 7th). A creature that successfully 
saves cannot be affected again by the same bone devil's aura for 24 hours. Other 
devils are immune to the aura. The save DC is Charisma-based.

\textbf{Poison (Ex):} Injury, Fortitude DC 20, initial damage 1d6 Str, secondary 
damage 2d6 Str. The save DC

is Constitution-based.

\textbf{Spell-Like Abilities:} At will---\textit{greater teleport }(self plus 50 
pounds of objects only), \textit{dimensional anchor, fly, invisibility }(self only), 
\textit{major image }(DC 15), \textit{wall of ice}. Caster level 12th. The save 
DC is Charisma-based.

\textit{\textbf{Summon Devil }}\textbf{(Sp): }Once per day a bone devil can attempt 
to summon 2d10 lemures with a 50\% chance of success, or another bone devil with 
a 35\% chance of success. This ability is the equivalent of a 4th-level spell.

\vspace{12pt}
CHAIN DEVIL (KYTON)

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{195pt}|}
\hline
  & Medium Outsider (Evil, Extraplanar, Lawful)\tabularnewline
\hline
H\textbf{it Dice:} & 8d8+16 (52 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +6\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 20 (+2 Dex, +8 natural), touch 12, flat-footed 18\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +8/+10\tabularnewline
\hline
A\textbf{ttack:} & Chain +10 melee (2d4+2/19-20)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 chains +10 melee (2d4+2/19-20)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. (10 ft. with chains)\tabularnewline
\hline
S\textbf{pecial Attacks:} & Dancing chains, unnerving gaze\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 5/silver or good, darkvision 60 
ft., immunity to cold, regeneration 2, spell resistance 18\tabularnewline
\hline
S\textbf{aves:} & Fort +8, Ref +8, Will +6\tabularnewline
\hline
A\textbf{bilities:} & Str 15, Dex 15, Con 15, Int 6, Wis 10, Cha 12\tabularnewline
\hline
S\textbf{kills:} & Climb +13, Craft (blacksmithing) +17, Escape Artist +13, Intimidate 
+12, Listen +13, Spot +13, Use Rope +2 (+4 with bindings)\tabularnewline
\hline
F\textbf{eats:} & Alertness, Improved Critical (chain), Improved Initiative\tabularnewline
\hline
E\textbf{nvironment:} & A lawful evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary, gang (2-4), band (6-10), or mob (11-20)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 6\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & 9-16 HD (Medium)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +6\tabularnewline
\hline
\end{tabular}

Kytons are humanlike devils, wrapped in chains instead of clothing. A chain devil 
is 6 feet tall and weighs about 300 pounds, chains included.

Chain devils speak Infernal and Common.

Combat

A chain devil attacks by flailing away with the spiked chains that serve as its 
clothing, armor, and weapons. 

A chain devil's natural weapons, as well as any weapons it wields, are treated 
as evil-aligned and lawful-aligned

for the purpose of overcoming damage reduction.

\textbf{Dancing Chains (Su): }A chain devil's most awesome attack is its ability 
to control up to four chains within 20 feet as a standard action, making the chains 
dance or move as it wishes. In addition, a chain devil can increase these chains' 
length by up to 15 feet and cause them to sprout razor-edged barbs. These chains 
attack as effectively as the devil itself. If a chain is in another creature's 
possession, the creature can attempt a DC 15 Will save to break the chain devil's 
power over that chain. If the save is successful, the kyton cannot attempt to control 
that particular chain again for 24 hours or until the chain leaves the creature's 
possession. The save DC is Charisma-based.

A chain devil can climb chains it controls at its normal speed without making Climb 
checks.

\textbf{Unnerving Gaze (Su):} Range 30 ft., Will DC 15 negates. A chain devil can 
make its face resemble one of an opponent's departed loved ones or bitter enemies. 
Those who fail their saves take a -2 penalty on attack rolls for 1d3 rounds. The 
save DC is Charisma-based.

\textbf{Regeneration (Ex): }Chain devils take normal damage from silvered weapons, 
good-aligned weapons, and spells or effects with the good descriptor. A chain devil 
that loses a piece of its body regrows it in 2d6x10 minutes. Holding the severed 
member against the stump enables it to reattach instantly.

\textbf{Skills:} Chain devils have a +8 racial bonus on Craft checks involving 
metalwork.

\vspace{12pt}
ERINYES

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{218pt}|}
\hline
  & Medium Outsider (Evil, Extraplanar, Lawful)\tabularnewline
\hline
H\textbf{it Dice:} & 9d8+45 (85 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +5\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares), fly 50 ft. (good)\tabularnewline
\hline
A\textbf{rmor Class:} & 23 (+5 Dex, +8 natural) touch 15, flat-footed 18\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +9/+14\tabularnewline
\hline
A\textbf{ttack:} & Longsword +14 melee (1d8+5/19-20) or \textit{+1 flaming composite 
longbow }(+5 Str bonus) +15 ranged (1d8+6/x3 plus 1d6 fire) or rope +14 ranged 
(entangle)\tabularnewline
\hline
F\textbf{ull Attack:} & Longsword +14/+9 melee (1d8+5/19-20) or \textit{+1 flaming 
composite longbow }(+5 Str bonus) +15/+10 ranged (1d8+6/x3 plus 1d6 fire) or rope 
+14 ranged (entangle)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Entangle, spell-like abilities, \textit{summon devil}\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 5/good, darkvision 60 ft., immunity 
to fire and poison, resistance to acid 10 and cold 10, see in darkness, spell resistance 
20, telepathy 100 ft., true seeing\tabularnewline
\hline
S\textbf{aves:} & Fort +11, Ref +11, Will +10\tabularnewline
\hline
A\textbf{bilities:} & Str 21, Dex 21, Con 21, Int 14, Wis 18, Cha 20\tabularnewline
\hline
S\textbf{kills:} & Concentration +17, Diplomacy +7, Escape Artist +17, Hide +17, 
Knowledge (any two) +14, Listen +16, Move Silently +17, Search +14, Sense Motive 
+16, Spot +16, Survival +4 (+6 following tracks), Use Rope +5 (+7 with bindings)\tabularnewline
\hline
F\textbf{eats:} & Dodge\textsuperscript{B}, Mobility\textsuperscript{B}, Point 
Blank Shot, Precise Shot, Rapid Shot, Shot on the Run\tabularnewline
\hline
E\textbf{nvironment:} & A lawful evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 8\tabularnewline
\hline
T\textbf{reasure:} & Standard, plus rope and \textit{+1 flaming composite longbow 
}(+5 Str bonus)\tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & 10-18 HD (Medium)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +7\tabularnewline
\hline
\end{tabular}

Unlike other devils, erinyes appear attractive to humans, resembling very comely 
women or men. An erinyes stands about 6 feet tall and weighs about 150 pounds.

Erinyes speak Infernal, Celestial, and Draconic.

Combat

Erinyes prefer to engage in combat from a distance. They use \textit{charm monster 
}to distract or disorganize their opponents, then rain down fiery arrows from above.

An erinyes's natural weapons, as well as any weapons it wields, are treated as 
evil-aligned and lawful-aligned for the purpose of overcoming damage reduction.

\textbf{Entangle (Ex):} Each erinyes carries a stout rope some 50 feet long that 
entangles opponents of any size as an \textit{animate rope }spell (caster level 
16th). An erinyes can hurl its rope 30 feet with no range penalty. Typically, an 
erinyes entangles a foe, lifts it into the air, and drops it from a great height.

\textbf{Spell-Like Abilities: }At will---\textit{greater teleport }(self plus 50 
pounds of objects only), \textit{charm monster }(DC 19), \textit{minor image }(DC 
17), \textit{unholy blight }(DC 19). Caster level 12th. The save DCs are Charisma-based.

\textit{\textbf{Summon Devil }}\textbf{(Sp): }Once per day an erinyes can attempt 
to summon 2d10 lemures or 1d4 bearded devils with a 50\% chance of success. This 
ability is the equivalent of a 3rd-level spell.

\textbf{True Seeing (Su):} Erinyes continuously use true seeing, as the spell (caster 
level 14th).

\vspace{12pt}
HELLCAT (BEZEKIRA)

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{231pt}|}
\hline
  & Large Outsider (Evil, Extraplanar, Lawful)\tabularnewline
\hline
H\textbf{it Dice:} & 8d8+24 (60 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +9\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 21 (-1 size, +5 Dex, +7 natural), touch 14, flat-footed 
16\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +8/+18\tabularnewline
\hline
A\textbf{ttack:} & Claw +13 melee (1d8+6)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +13 melee (1d8+6) and bite +8 melee (2d8+3)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Improved grab, pounce, rake 1d8+3\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 5/good, darkvision 60 ft., invisible 
in light, resistance to fire 10, scent, spell resistance 19, telepathy 100 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +9, Ref +11, Will +8\tabularnewline
\hline
A\textbf{bilities:} & Str 23, Dex 21, Con 17, Int 10, Wis 14, Cha 10\tabularnewline
\hline
S\textbf{kills:} & Balance +16, Climb +17, Hide +13, Jump +21, Listen +17, Move 
Silently +20, Spot +13, Swim +17\tabularnewline
\hline
F\textbf{eats:} & Dodge, Improved Initiative, Track\tabularnewline
\hline
E\textbf{nvironment:} & A lawful evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or pride (6-10)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 7\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & 9-10 HD (Large); 11-24 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

While its appearance suggests it is incorporeal, a hellcats has a corporeal body 
and can be harmed by physical attacks.

Hellcats use a natural telepathy to communicate with one another and those they 
encounter. A hellcat measures about 9 feet long and weighs about 900 pounds.

Combat

A hellcat can hold its own in combat thanks to sharp claws and wicked fangs. It 
prefers to leap upon opponents, just as a lion does.

A hellcat's natural weapons, as well as any weapons it wields, are treated as evil-aligned 
and lawful-aligned for the purpose of overcoming damage reduction.

\textbf{Improved Grab (Ex):} To use this ability, a hellcat must hit with its bite 
attack. It can then attempt to start a grapple as a free action without provoking 
an attack of opportunity. If it wins the grapple check, it establishes a hold and 
can rake.

\textbf{Pounce (Ex):} If a hellcat charges, it can make a full attack, including 
two rake attacks.

\textbf{Rake (Ex):} Attack bonus +13 melee, damage 1d8+3.

\textbf{Invisible in Light (Ex):} A hellcat is invisible in any area lit well enough 
for a human to see. In a darkened area, it shows up as a faintly glowing outline 
visible up to 30 feet away (60 feet if the viewer has low-light vision). Magical 
darkness smothers the glow and conceals the outline.

\textbf{Skills:} Hellcats have a +4 racial bonus on Listen and Move Silently checks.

\vspace{12pt}
HORNED DEVIL (CORNUGON)

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{231pt}|}
\hline
  & Large Outsider (Evil, Extraplanar, Lawful)\tabularnewline
\hline
H\textbf{it Dice:} & 15d8+105 (172 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +7\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), fly 50 ft. (average)\tabularnewline
\hline
A\textbf{rmor Class:} & 35 (-1 size, +7 Dex, +19 natural) touch 16, flat-footed 
28\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +15/+29\tabularnewline
\hline
A\textbf{ttack:} & Spiked chain +25 melee (2d6+15 plus stun) or claw +24 melee 
(2d6+10) or tail +24 melee (2d6+10 plus infernal wound)\tabularnewline
\hline
F\textbf{ull Attack:} & Spiked chain +25/+20/+15 melee (2d6+15 plus stun) and bite 
+22 melee (2d8+5) and tail +22 melee (2d6+5 plus infernal wound); or 2 claws +24 
melee (2d6+10) and bite + 22 melee (2d8+5) and tail +22 melee (2d6+5 plus infernal 
wound)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft. (20 ft. with spiked chain)\tabularnewline
\hline
S\textbf{pecial Attacks:} & Fear aura, infernal wound, spell-like abilities, stun, 
\textit{summon devil}\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 10/good and silver, darkvision 60 
ft., immunity to fire and poison, resistance to acid 10 and cold 10, regeneration 
5, see in darkness, spell resistance 28, telepathy 100 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +16, Ref +16, Will +15\tabularnewline
\hline
A\textbf{bilities:} & Str 31, Dex 25, Con 25, Int 14, Wis 18, Cha 22\tabularnewline
\hline
S\textbf{kills:} & Bluff +24, Climb +28, Concentration +24, Diplomacy +10, Disguise 
+6 (+8 acting), Hide +21, Intimidate +26, Listen +22, Move Silently +23, Search 
+20, Sense Motive +22, Spot +22, Survival +4 (+6 following tracks)\tabularnewline
\hline
F\textbf{eats:} & Cleave, Improved Sunder, Iron Will, Multiattack, Power Attack, 
Weapon Focus (spiked chain)\tabularnewline
\hline
E\textbf{nvironment:} & A lawful evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary, team (2-4), or squad (6-10)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 16\tabularnewline
\hline
T\textbf{reasure:} & Standard coins; double goods; standard items\tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & 16-20 HD (Large); 21-45 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A horned devil is 9 feet tall and weighs about 600 pounds.

Combat

Horned devils are bold fighters. They rarely retreat, even against overwhelming 
odds. They love to fight with their spiked chains, usually singling out the most 
powerful foes to stun and eliminate quickly.

A horned devil's natural weapons, as well as any weapons it wields, are treated 
as evil-aligned and lawful-aligned for the purpose of overcoming damage reduction.

\textbf{Spell-Like Abilities:} At will---\textit{dispel chaos }(DC 21), \textit{dispel 
good }(DC 21), \textit{magic circle against good, greater teleport }(self plus 
50 pounds of objects only); \textit{persistent image }(DC 21) 3/day---\textit{fireball 
}(DC 19), \textit{lightning bolt }(DC 19). Caster level 15th. The save DCs are 
Charisma-based.

\textbf{Fear Aura (Su): }A horned devil can radiate a 5-foot-radius fear aura as 
a free action. A creature in the area must succeed on a DC 23 Will save or be affected 
as though by a \textit{fear }spell (caster level 15th). A creature that successfully 
saves cannot be affected again by the same horned devil's aura for 24 hours. Other 
devils are immune to the aura. The save DC is Charisma-based.

\textbf{Stun (Su):} Whenever a horned devil hits with a spiked chain attack, the 
opponent must succeed on a DC 27 Fortitude save or be stunned for 1d4 rounds. The 
save DC is Strength-based. This ability is a function of the horned devil, not 
of the spiked chain.

\textbf{Infernal Wound (Su):} The damage a horned devil deals with its tail attack 
causes a persistent wound. An injured creature loses 2 additional hit points each 
round. The wound does not heal naturally and resists healing spells. The continuing 
hit point loss can be stopped by a DC 24 Heal check, a \textit{cure }spell, or 
a \textit{heal }spell. However, a character attempting to cast a \textit{cure }spell 
or a \textit{heal }spell on a creature damaged by a horned devil's tail must succeed 
on a DC 24 caster level check, or the spell has no effect on the injured character. 
A successful Heal check automatically stops the continuing hit point loss as well 
as restoring hit points. The check DC is Constitution-based.

\textit{\textbf{Summon Devil }}\textbf{(Sp):} Once per day a horned devil can attempt 
to summon 2d10 lemures or 1d6 bearded devils with a 50\% chance of success, 1d6 
barbed devils with a 35\% chance of success, or another horned devil with a 20\% 
chance of success. This ability is the equivalent of a 6th-level spell.

\textbf{Regeneration (Ex):} A horned devil takes normal damage from good-aligned 
silvered weapons, and from spells or effects with the good descriptor.

\vspace{12pt}
ICE DEVIL (GELUGON)

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{222pt}|}
\hline
  & Large Outsider (Evil, Extraplanar, Lawful)\tabularnewline
\hline
H\textbf{it Dice:} & 14d8+84 (147 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +5\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 32 (-1 size, +5 Dex, +18 natural) touch 14, flat-footed 
27\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +14/+24\tabularnewline
\hline
A\textbf{ttack:} & Spear +20 melee (2d6+9/x3 plus slow) or claw +19 melee (1d10+6)\tabularnewline
\hline
F\textbf{ull Attack:} & Spear +20/+15/+10 melee (2d6+9/x3 plus slow) and bite +14 
melee (2d6+3) and tail +14 melee (3d6+3 plus slow); or 2 claws +19 melee (1d10+6) 
and bite +14 melee (2d6+3) and tail +14 melee (3d6+3 plus slow)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Fear aura, slow, spell-like abilities, \textit{summon 
devil }\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 10/good, darkvision 60 ft., immunity 
to fire and poison, resistance to acid 10 and cold 10, regeneration 5, see in darkness, 
spell resistance 25, telepathy 100 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +15, Ref +14, Will +15\tabularnewline
\hline
A\textbf{bilities:} & Str 23, Dex 21, Con 23, Int 22, Wis 22, Cha 20\tabularnewline
\hline
S\textbf{kills:} & Bluff +22, Climb +23, Concentration +23, Diplomacy +9, Disguise 
+5 (+7 acting), Intimidate +24, Jump +27, Knowledge (any three) +23, Listen +25, 
Move Silently +22, Search +23, Sense Motive +23, Spellcraft +23, Spot +25, Survival 
+6 (+8 following tracks)\tabularnewline
\hline
F\textbf{eats:} & Alertness, Cleave, Combat Reflexes, Power Attack, Weapon Focus 
(spear)\tabularnewline
\hline
E\textbf{nvironment:} & A lawful evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary, team (2-4), squad (6-10), or troupe (1-2 ice 
devils, 7-12 bearded devils, and 1-4 bone devils)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 13\tabularnewline
\hline
T\textbf{reasure:} & Standard coins; double goods; standard items\tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & 15-28 HD (Large); 29-42 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

An ice devil is about 12 feet tall and weighs about 700 pounds. 

Combat

An ice devil prefers to fight only when doing so serves its mission, but it never 
hesitates to attack when it deems a battle necessary---or likely to end in its 
victory.

An ice devil's natural weapons, as well as any weapons it wields, are treated as 
evil-aligned and lawful-aligned for the purpose of overcoming damage reduction.

\textbf{Fear Aura (Su):} An ice devil can radiate a 10-foot-radius fear aura as 
a free action. A creature in the area must succeed on a DC 22 Will save or be affected 
as though by a \textit{fear }spell (caster level 13th). A creature that successfully 
saves cannot be affected again by the same ice devil's aura for 24 hours. Other 
devils are immune to the aura. The save DC is Charisma-based.

\textbf{Slow (Su): }A hit from an ice devil's tail or spear induces numbing cold. 
The opponent must succeed on a DC 23 Fortitude save or be affected as though by 
a \textit{slow }spell for 1d6 rounds. The save DC is Constitution-based.

\textbf{Spell-Like Abilities:} At will---\textit{cone of cold }(DC 20), \textit{fly, 
ice storm }(DC 19), \textit{greater teleport }(self plus 50 pounds of objects only), 
\textit{persistent image }(DC 20), \textit{unholy aura }(DC 23), \textit{wall of 
ice }(DC 19). Caster level 13th. The save DCs are Charisma-based.

\textit{\textbf{Summon Devil }}\textbf{(Sp):} Once per day an ice devil can attempt 
to summon 2d10 lemures or 1d6 bearded devils, 2d4 bone devils with a 50\% chance 
of success, or another ice devil with a 20\% chance of success. This ability is 
the equivalent of a 4th-level spell.

\textbf{Regeneration (Ex):} An ice devil takes normal damage from good-aligned 
weapons and from spells or effects with the good descriptor.

\vspace{12pt}
IMP

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{231pt}|}
\hline
  & Tiny Outsider (Evil, Extraplanar, Lawful)\tabularnewline
\hline
H\textbf{it Dice:} & 3d8 (13 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +3\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), fly 50 ft. (perfect)\tabularnewline
\hline
A\textbf{rmor Class:} & 20 (+2 size, +3 Dex, +5 natural), touch 15, flat-footed 
17\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +3/-5\tabularnewline
\hline
A\textbf{ttack:} & Sting +8 melee (1d4 plus poison)\tabularnewline
\hline
F\textbf{ull Attack:} & Sting +8 melee (1d4 plus poison)\tabularnewline
\hline
S\textbf{pace/Reach:} & 2-1/2 ft./0 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Poison, spell-like abilities\tabularnewline
\hline
S\textbf{pecial Qualities:} & Alternate form, damage reduction 5/good or silver, 
darkvision 60 ft., fast healing 2, immunity to poison, resistance to fire 5\tabularnewline
\hline
S\textbf{aves:} & Fort +3, Ref +6, Will +4\tabularnewline
\hline
A\textbf{bilities:} & Str 10, Dex 17, Con 10, Int 10, Wis 12, Cha 14\tabularnewline
\hline
S\textbf{kills:} & Diplomacy +8, Hide +17, Knowledge (any one) +6, Listen +7, Move 
Silently +9, Search +6, Spellcraft +6, Spot +7, Survival +1 (+3 following tracks)\tabularnewline
\hline
F\textbf{eats:} & Dodge, Weapon Finesse\tabularnewline
\hline
E\textbf{nvironment:} & A lawful evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & 4-6 HD (Tiny)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- &  (Improved Familiar)\tabularnewline
\hline
\end{tabular}

In its natural form, an imp stands almost 2 feet tall and weighs about 8 pounds.

Combat

Imps are craven, but not so timid as to pass up an opportunity for a surprise attack 
using their \textit{invisibility }and alternate form ability. In its natural form, 
an imp attacks with the wicked stinger on its tail. It quickly flies out of reach 
if a foe manages to strike back effectively.

An imp's natural weapons, as well as any weapons it wields, are treated as evil-aligned 
and lawful-aligned for the purpose of overcoming damage reduction.

\textbf{Poison (Ex):} Injury, Fortitude DC 13, initial damage 1d4 Dex, secondary 
damage 2d4 Dex. The save DC is Constitution-based and includes a +2 racial bonus.

\textbf{Spell-Like Abilities:} At will---\textit{detect good, detect magic, invisibility 
}(self only); 1/day---\textit{suggestion }(DC 15). Caster level 6th. The save DC 
is Charisma-based.

Once per week an imp can use \textit{commune }to ask six questions. The ability 
otherwise works as the spell (caster level 12th).

\textbf{Alternate Form (Su):} An imp can assume other forms at will as a standard 
action. This ability functions as a \textit{polymorph }spell cast on itself (caster 
level 12th), except that an imp does not regain hit points for changing form, and 
an individual imp can assume only one or two forms no larger than Medium. Common 
forms include monstrous spider, raven, rat, and boar.

\vspace{12pt}
LEMURE

\begin{tabular}{|>{\raggedright}p{85pt}|>{\raggedright}p{206pt}|}
\hline
  & Medium Outsider (Evil, Extraplanar, Lawful)\tabularnewline
\hline
Hit Dice: & 2d8 (9 hp)\tabularnewline
\hline
Initiative: & +0\tabularnewline
\hline
Speed: & 20 ft. (4 squares)\tabularnewline
\hline
Armor Class: & 14 (+4 natural) touch 10, flat-footed 14\tabularnewline
\hline
Base Attack/Grapple: & +2/+2\tabularnewline
\hline
Attack: & Claw +2 melee (1d4)\tabularnewline
\hline
Full Attack: & 2 claws +2 melee (1d4)\tabularnewline
\hline
Space/Reach: & 5 ft./5 ft.\tabularnewline
\hline
Special Attacks:--- & \tabularnewline
\hline
Special Qualities: & Damage reduction 5/good or silver, darkvision 60 ft., immunity 
to fire and poison, mindless, resistance to acid 10 and cold 10, mindless, see 
in darkness\tabularnewline
\hline
Saves: & Fort +3, Ref +3, Will +3\tabularnewline
\hline
Abilities: & Str 10, Dex 10, Con 10, Int ---, Wis 11, Cha 5\tabularnewline
\hline
Environment: & A lawful evil-aligned plane\tabularnewline
\hline
Organization: & Solitary, pair, gang (3-5), swarm (6-15), or mob (10-40)\tabularnewline
\hline
Challenge Rating: & 1\tabularnewline
\hline
Treasure: & None\tabularnewline
\hline
Alignment: & Always lawful evil\tabularnewline
\hline
Advancement: & 3-6 HD (Medium)\tabularnewline
\hline
Level Adjustment:--- & \tabularnewline
\hline
\end{tabular}

A lemure is about 5 feet tall and weighs about 100 pounds.

Lemures are mindless and cannot communicate, but they are sensitive to telepathic 
messages from other devils, typically obeying a devil's mental commands.

Combat

Lemures surge toward anything they meet and try to claw it apart. Only a telepathic 
command from other devils or the complete destruction of the lemures can make them 
stop.

A lemure's natural weapons, as well as any weapons it wields, are treated as evil-aligned 
and lawful-aligned for the purpose of overcoming damage reduction.

\textbf{Mindless (Ex):} Immunity to mind-affecting effects (charms, compulsions, 
phantasms, patterns, and morale effects).

\vspace{12pt}
PIT FIEND

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{231pt}|}
\hline
  & Large Outsider (Evil, Extraplanar, Lawful)\tabularnewline
\hline
H\textbf{it Dice:} & 18d8+144 (225 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +12\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares), fly 60 ft. (average)\tabularnewline
\hline
A\textbf{C:} & 40 (-1 size, +8 Dex, +23 natural) touch 17, flat-footed 32\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +18/+35\tabularnewline
\hline
A\textbf{ttack:} & Claw +30 melee (2d8+13)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +30 melee (2d8+13) and 2 wings +28 melee (2d6+6) 
and bite +28 melee (4d6+6 plus poison plus disease) and tail slap +28 melee (2d8+6)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Constrict 2d8+26, fear aura, improved grab, spell-like 
abilities, \textit{summon devil}\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 15/good and silver, darkvision 60 
ft., immunity to fire and poison, resistance to acid 10 and cold 10, regeneration 
5, see in darkness, spell resistance 32, telepathy 100 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +19, Ref +19, Will +21\tabularnewline
\hline
A\textbf{bilities:} & Str 37, Dex 27, Con 27, Int 26, Wis 26, Cha 26\tabularnewline
\hline
S\textbf{kills:} & Balance +10, Bluff +29, Climb +34, Concentration +29, Diplomacy 
+10, Disguise +29 (+31 acting), Hide +25, Intimidate +31, Jump +40, Knowledge (arcana) 
+29, Knowledge (nature) +10, Knowledge (the planes) +29, Knowledge (religion) +29, 
Listen +29, Move Silently +29, Search +29, Spellcraft +31, Spot +29, Survival +8 
(+10 on other planes, +10 when tracking), Tumble +31\tabularnewline
\hline
F\textbf{eats:} & Cleave, Great Cleave, Improved Initiative, Iron Will, Multiattack, 
Power Attack, Quicken Spell-Like Ability (\textit{fireball})\tabularnewline
\hline
E\textbf{nvironment:} & A lawful evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, team (3-4), or troupe (1-2 pit fiends, 
2-5 horned devils, and 2-5 barbed devils)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 20\tabularnewline
\hline
T\textbf{reasure:} & Standard coins; double goods; standard items\tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & 19-36 HD (Large); 37-54 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A pit fiend often wraps its wings around itself like a grotesque cloak, and appears 
wreathed in flames.

A pit fiend is 12 feet tall and weighs 800 pounds.

Combat

Pit fiends are wily and resourceful fighters, using \textit{invisibility }to gain 
the upper hand and biting at foes seemingly able to see them. They don't hesitate 
to blanket an area with \textit{fireballs}, and can call down the wrath of an inferno 
with \textit{meteor swarm.}

A pit fiend's natural weapons, as well as any weapons it wields, are treated as 
evil-aligned and lawful-aligned for the purpose of overcoming damage reduction.

\textbf{Constrict (Ex): }A pit fiend deals 2d8+26 points of damage with a successful 
grapple check.

\textbf{Disease (Su):} A creature struck by a pit fiend's bite attack must succeed 
on a DC 27 Fortitude save or be infected with a vile disease known as devil chills 
(incubation period 1d4 days, damage 1d4 Str). The save DC is Constitution-based.

\textbf{Fear Aura (Su): }A pit fiend can radiate a 20-foot-radius fear aura as 
a free action. A creature in the area must succeed on a DC 27 Will save or be affected 
as though by a \textit{fear }spell (caster level 18th). A creature that successfully 
saves cannot be affected again by the same pit fiend's aura for 24 hours. Other 
devils are immune to the aura. The save DC is Charisma-based.

\textbf{Improved Grab (Ex):} To use this ability, a pit fiend must hit with its 
tail slap attack. It can then attempt to start a grapple as a free action without 
provoking an attack of opportunity.

If it wins the grapple check, it establishes a hold and can constrict.

\textbf{Poison (Ex): }Injury, Fortitude DC 27, initial damage 1d6 Con, secondary 
damage death. The save DC is Constitution-based.

\textbf{Spell-Like Abilities:} At will---\textit{blasphemy }(DC 25), \textit{create 
undead, fireball }(DC 21), \textit{greater dispel magic, greater teleport }(self 
plus 50 pounds of objects only), \textit{invisibility, magic circle against good, 
mass hold monster }(DC 27), \textit{persistent image }(DC 23), \textit{power word 
stun, unholy aura }(DC 26); 1/day---\textit{meteor swarm }(DC 27). Caster level 
18th. The save DCs are Charisma-based.

Once per year a pit fiend can use \textit{wish }as the spell (caster level 20th).

\textit{\textbf{Summon Devil }}\textbf{(Sp):} Twice per day a pit fiend can automatically 
summon 2 lemures, bone devils, or bearded devils, or 1 erinyes, horned devil, or 
ice devil. This ability is the equivalent of an 8th-level spell. 

\textbf{Regeneration (Ex):} A pit fiend takes normal damage from good-aligned silvered 
weapons, and from spells or effects with the good descriptor.

Tactics Round-by-Round

A pit fiend typically opens combat by using its spell-like abilities, attempting 
to neutralize dangerous opponents before entering melee.

\textit{Prior to combat: Unholy aura}; activate fear aura, \textit{summon devil}.

\textit{Round 1: }Quickened \textit{fireball }and \textit{mass hold monster }if 
facing three or more visible, active opponents; otherwise \textit{power word stun 
}against unarmored opponent (preferably a spellcaster).

\textit{Round 2: Meteor swarm }against as many foes as possible, approach worst-injured 
enemy.

\textit{Round 3: }Full attack against injured enemy.

\textit{Round 4: }Continue melee against injured enemy, or \textit{power word stun 
}against annoying spellcaster.

\textit{Round 5: }Repeat from round 1, or \textit{greater teleport }to safety if 
endangered.

\vspace{12pt}
{\LARGE{}DEVOURER}

\begin{tabular}{|>{\raggedright}p{86pt}|>{\raggedright}p{239pt}|}
\hline
  & Large Undead (Extraplanar)\tabularnewline
\hline
H\textbf{it Dice:} & 12d12 (78 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +4\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 24 (-1 size, +15 natural), touch 9, flat-footed 24\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +6/+19\tabularnewline
\hline
A\textbf{ttack:} & Claw +15 melee (1d6+9)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +15 melee (1d6+9)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Energy drain, trap essence, spell-like abilities\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., spell deflection, spell resistance 
21, undead traits\tabularnewline
\hline
S\textbf{aves:} & Fort +4, Ref +4, Will +11\tabularnewline
\hline
A\textbf{bilities:} & Str 28, Dex 10, Con ---, Int 16, Wis 16, Cha 17\tabularnewline
\hline
S\textbf{kills:} & Climb +24, Concentration +18, Diplomacy +5, Jump +24, Listen 
+18, Move Silently +15, Search +10, Sense Motive +11, Spot +18, Survival +3 (+5 
following tracks)\tabularnewline
\hline
F\textbf{eats:} & Blind-Fight, Combat Casting, Combat Expertise, Improved Initiative, 
Weapon Focus (claw)\tabularnewline
\hline
E\textbf{nvironment:} & Any\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 11\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral evil\tabularnewline
\hline
A\textbf{dvancement:} & 13-24 HD (Large); 25-36 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

Devourers lurk on the Ethereal Plane and the Astral Plane, stalking both natives 
and travelers with equal sadistic glee.

A devourer is about 9 feet tall and weighs 500 pounds.

Devourers speak Common.

COMBAT

Even if it had no special abilities, a devourer would be a terrible opponent, for 
its bony claws can flay enemies alive.

\textbf{Energy Drain (Su):} Living creatures hit by a devourer's claw attack or 
\textit{spectral hand }ability gain one negative level. The DC is 19 for the Fortitude 
save to remove a negative level. The save DC is Charisma-based.

\textbf{Trap Essence (Su)}: The devourer is named for its ability to consume an 
enemy's life essence. To do so, it must forgo its normal melee attacks and make 
a trap essence attack. This requires a normal attack roll but deals no damage. 
The he save DC is Charisma-based. A slain creature's essence is trapped within 
the devourer's ribs, and the tiny figure takes on that victim's features. The trapped 
essence cannot be raised or resurrected, but a \textit{limited wish, miracle, }or 
\textit{wish }spell frees it, as does destroying the devourer. A devourer can hold 
only one essence at a time.

The trapped essence provides a devourer with enough power to use five spell-like 
abilities for each Hit Die or level of the trapped creature. As this energy is 
expended, the twisted soul fades away until it evaporates completely. The trapped 
essence gains one negative level for every five times the devourer uses one of 
its spell-like abilities. When the essence's number of negative levels equals the 
creature's total Hit Dice or level, the essence is destroyed. If an essence is 
freed, the restored creature must succeed on a DC 19 Fortitude save for each negative 
level or lose that level permanently.

\textbf{Spell-Like Abilities}: At the start of any encounter, the trapped essence 
within a devourer is assumed to have 3d4+3 levels (enough fuel for thirty to seventy-five 
uses). Once per round, a devourer can use one of the following abilities: \textit{confusion 
}(DC 17), \textit{control undead }(DC 20), \textit{ghoul touch }(DC 15), \textit{lesser 
planar ally, ray of enfeeblement }(DC 14), \textit{spectral hand, suggestion }(DC 
16), \textit{true seeing}. Caster level 18th. The save DCs are Charisma-based.

\textbf{Spell Deflection (Su):} The trapped essence provides a measure of magical 
protection. If any of the following spells are cast at the devourer and overcome 
its spell resistance, they affect the imprisoned essence instead: \textit{banishment, 
chaos hammer, confusion, crushing despair, detect thoughts, dispel evil, dominate 
person, fear, geas/quest, holy word, hypnosis, imprisonment, magic jar, maze, suggestion, 
trap the soul, }or any form of charm or compulsion. In many cases, this deflection 
effectively neutralizes the spell. Some of these effects might eliminate the trapped 
essence, depriving the devourer of its spell-like abilities until it can consume 
another victim. 

\newpage

\end{document}
