%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

This material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

{\LARGE{}SPELL LIST II}

\vspace{12pt}
{\LARGE{}DRUID SPELLS}

0-LEVEL DRUID SPELLS (ORISONS)

\textbf{Create Water:} Creates 2 gallons/level of pure water.

\textbf{Cure Minor Wounds:} Cures 1 point of damage.

\textbf{Detect Magic:} Detects spells and magic items within 60 ft.

\textbf{Detect Poison:} Detects poison in one creature or object.

\textbf{Flare:} Dazzles one creature (-1 penalty on attack rolls).

\textbf{Guidance:} +1 on one attack roll, saving throw, or skill check.

\textbf{Know Direction:} You discern north.

\textbf{Light:} Object shines like a torch.

\textbf{Mending:} Makes minor repairs on an object.

\textbf{Purify Food and Drink:} Purifies 1 cu. ft./level of food or water.

\textbf{Read Magic:} Read scrolls and spellbooks.

\textbf{Resistance:} Subject gains +1 bonus on saving throws.

\textbf{Virtue:} Subject gains 1 temporary hp.

1ST-LEVEL DRUID SPELLS

\textbf{Calm Animals:} Calms (2d4 + level) HD of animals.

\textbf{Charm Animal:} Makes one animal your friend.

\textbf{Cure Light Wounds:} Cures 1d8 damage +1/level (max +5).

\textbf{Detect Animals or Plants:} Detects kinds of animals or plants.

\textbf{Detect Snares and Pits:} Reveals natural or primitive traps.

\textbf{Endure Elements:} Exist comfortably in hot or cold environments.

\textbf{Entangle:} Plants entangle everyone in 40-ft.-radius.

\textbf{Faerie Fire:} Outlines subjects with light, canceling \textit{blur}, concealment, 
and the like.

\textbf{Goodberry:} 2d4 berries each cure 1 hp (max 8 hp/24 hours).

\textbf{Hide from Animals:} Animals can't perceive one subject/level.

\textbf{Jump:} Subject gets bonus on Jump checks.

\textbf{Longstrider:} Your speed increases by 10 ft.

\textbf{Magic Fang:} One natural weapon of subject creature gets +1 on attack and 
damage rolls.

\textbf{Magic Stone:} Three stones gain +1 on attack rolls, deal 1d6+1 damage.

\textbf{Obscuring Mist:} Fog surrounds you.

\textbf{Pass without Trace:} One subject/level leaves no tracks.

\textbf{Produce Flame:} 1d6 damage +1/level, touch or thrown.

\textbf{Shillelagh:} Cudgel or quarterstaff becomes +1 weapon (1d10 damage) for 
1 min./level.

\textbf{Speak with Animals:} You can communicate with animals.

\textbf{Summon Nature's Ally I:} Calls creature to fight.

2ND-LEVEL DRUID SPELLS

\textbf{Animal Messenger:} Sends a Tiny animal to a specific place.

\textbf{Animal Trance:} Fascinates 2d6 HD of animals.

\textbf{Barkskin:} Grants +2 (or higher) enhancement to natural armor.

\textbf{Bear's Endurance:} Subject gains +4 to Con for 1 min./level.

\textbf{Bull's Strength:} Subject gains +4 to Str for 1 min./level.

\textbf{Cat's Grace:} Subject gains +4 to Dex for 1 min./level.

\textbf{Chill Metal:} Cold metal damages those who touch it.

\textbf{Delay Poison:} Stops poison from harming subject for 1 hour/level.

\textbf{Fire Trap }\textsuperscript{\textbf{M}}\textbf{:} Opened object deals 1d4 
+1/level damage.

\textbf{Flame Blade:} Touch attack deals 1d8 +1/two levels damage.

\textbf{Flaming Sphere:} Creates rolling ball of fire, 2d6 damage, lasts 1 round/level.

\textbf{Fog Cloud:} Fog obscures vision.

\textbf{Gust of Wind:} Blows away or knocks down smaller creatures.

\textbf{Heat Metal:} Make metal so hot it damages those who touch it.

\textbf{Hold Animal:} Paralyzes one animal for 1 round/level.

\textbf{Owl's Wisdom:} Subject gains +4 to Wis for 1 min./level.

\textbf{Reduce Animal:} Shrinks one willing animal.

\textbf{Resist Energy:} Ignores 10 (or more) points of damage/attack from specified 
energy type.

\textbf{Restoration, Lesser:} Dispels magical ability penalty or repairs 1d4 ability 
damage.

\textbf{Soften Earth and Stone:} Turns stone to clay or dirt to sand or mud.

\textbf{Spider Climb:} Grants ability to walk on walls and ceilings.

\textbf{Summon Nature's Ally II:} Calls creature to fight.

\textbf{Summon Swarm:} Summons swarm of bats, rats, or spiders.

\textbf{Tree Shape:} You look exactly like a tree for 1 hour/level.

\textbf{Warp Wood:} Bends wood (shaft, handle, door, plank).

\textbf{Wood Shape:} Rearranges wooden objects to suit you.

3RD-LEVEL DRUID SPELLS

\textbf{Call Lightning:} Calls down lightning bolts (3d6 per bolt) from sky.

\textbf{Contagion:} Infects subject with chosen disease.

\textbf{Cure Moderate Wounds:} Cures 2d8 damage +1/level (max +10).

\textbf{Daylight:} 60-ft. radius of bright light.

\textbf{Diminish Plants:} Reduces size or blights growth of normal plants.

\textbf{Dominate Animal:} Subject animal obeys silent mental commands.

\textbf{Magic Fang, Greater:} One natural weapon of subject creature gets +1/four 
levels on attack and damage rolls (max +5).

\textbf{Meld into Stone:} You and your gear merge with stone.

\textbf{Neutralize Poison:} Immunizes subject against poison, detoxifies venom 
in or on subject.

\textbf{Plant Growth:} Grows vegetation, improves crops.

\textbf{Poison:} Touch deals 1d10 Con damage, repeats in 1 min.

\textbf{Protection from Energy:} Absorb 12 points/level of damage from one kind 
of energy.

\textbf{Quench:} Extinguishes nonmagical fires or one magic item.

\textbf{Remove Disease:} Cures all diseases affecting subject.

\textbf{Sleet Storm:} Hampers vision and movement.

\textbf{Snare:} Creates a magic booby trap.

\textbf{Speak with Plants:} You can talk to normal plants and plant creatures.

\textbf{Spike Growth:} Creatures in area take 1d4 damage, may be \textit{slowed.}

\textbf{Stone Shape:} Sculpts stone into any shape.

\textbf{Summon Nature's Ally III:} Calls creature to fight.

\textbf{Water Breathing:} Subjects can breathe underwater.

\textbf{Wind Wall:} Deflects arrows, smaller creatures, and gases.

4TH-LEVEL DRUID SPELLS

\textbf{Air Walk:} Subject treads on air as if solid (climb at 45-degree angle).

\textbf{Antiplant Shell:} Keeps animated plants at bay.

\textbf{Blight:} Withers one plant or deals 1d6/level damage to plant creature.

\textbf{Command Plants:} Sway the actions of one or more plant creatures.

\textbf{Control Water:} Raises or lowers bodies of water.

\textbf{Cure Serious Wounds:} Cures 3d8 damage +1/level (max +15).

\textbf{Dispel Magic:} Cancels spells and magical effects.

\textbf{Flame Strike:} Smite foes with divine fire (1d6/level damage).

\textbf{Freedom of Movement:} Subject moves normally despite impediments.

\textbf{Giant Vermin:} Turns centipedes, scorpions, or spiders into giant vermin.

\textbf{Ice Storm:} Hail deals 5d6 damage in cylinder 40 ft. across.

\textbf{Reincarnate:} Brings dead subject back in a random body.

\textbf{Repel Vermin:} Insects, spiders, and other vermin stay 10 ft. away.

\textbf{Rusting Grasp:} Your touch corrodes iron and alloys.

\textbf{Scrying }\textsuperscript{\textbf{F}}\textbf{:} Spies on subject from a 
distance.

\textbf{Spike Stones:} Creatures in area take 1d8 damage, may be \textit{slowed.}

\textbf{Summon Nature's Ally IV:} Calls creature to fight.

5TH-LEVEL DRUID SPELLS

\textbf{Animal Growth:} One animal/two levels doubles in size.

\textbf{Atonement:} Removes burden of misdeeds from subject.

\textbf{Awaken }\textsuperscript{\textbf{X}}\textbf{:} Animal or tree gains human 
intellect.

\textbf{Baleful Polymorph:} Transforms subject into harmless animal.

\textbf{Call Lightning Storm:} As \textit{call lightning, }but 5d6 damage per bolt.

\textbf{Commune with Nature:} Learn about terrain for 1 mile/level.

\textbf{Control Winds:} Change wind direction and speed.

\textbf{Cure Critical Wounds:} Cures 4d8 damage +1/level (max +20).

\textbf{Death Ward:} Grants immunity to all death spells and negative energy effects.

\textbf{Hallow }\textsuperscript{\textbf{M}}\textbf{:} Designates location as holy.

\textbf{Insect Plague:} Locust swarms attack creatures.

\textbf{Stoneskin }\textsuperscript{\textbf{M}}\textbf{:} Ignore 10 points of damage 
per attack.

\textbf{Summon Nature's Ally V:} Calls creature to fight.

\textbf{Transmute Mud to Rock:} Transforms two 10-ft. cubes per level.

\textbf{Transmute Rock to Mud:} Transforms two 10-ft. cubes per level.

\textbf{Tree Stride:} Step from one tree to another far away.

\textbf{Unhallow }\textsuperscript{\textbf{M}}\textbf{:} Designates location as 
unholy.

\textbf{Wall of Fire:} Deals 2d4 fire damage out to 10 ft. and 1d4 out to 20 ft. 
Passing through wall deals 2d6 damage +1/level.

\textbf{Wall of Thorns:} Thorns damage anyone who tries to pass.

6TH-LEVEL DRUID SPELLS

\textbf{Antilife Shell:} 10-ft.-radius field hedges out living creatures.

\textbf{Bear's Endurance, Mass:} As \textit{bear's endurance}, affects one subject/ 
level.

\textbf{Bull's Strength, Mass:} As \textit{bull's strength}, affects one subject/level.

\textbf{Cat's Grace, Mass:} As \textit{cat's grace}, affects one subject/level.

\textbf{Cure Light Wounds, Mass:} Cures 1d8 damage +1/level for many creatures.

\textbf{Dispel Magic, Greater:} As \textit{dispel magic, }but +20 on check.

\textbf{Find the Path:} Shows most direct way to a location.

\textbf{Fire Seeds:} Acorns and berries become grenades and bombs.

\textbf{Ironwood:} Magic wood is strong as steel.

\textbf{Liveoak:} Oak becomes treant guardian.

\textbf{Move Earth:} Digs trenches and builds hills.

\textbf{Owl's Wisdom, Mass:} As \textit{owl's wisdom}, affects one subject/level.

\textbf{Repel Wood:} Pushes away wooden objects.

\textbf{Spellstaff:} Stores one spell in wooden quarterstaff.

\textbf{Stone Tell:} Talk to natural or worked stone.

\textbf{Summon Nature's Ally VI:} Calls creature to fight.

\textbf{Transport via Plants:} Move instantly from one plant to another of the 
same kind.

\textbf{Wall of Stone:} Creates a stone wall that can be shaped.

7TH-LEVEL DRUID SPELLS

\textbf{Animate Plants:} One or more plants animate and fight for you.

\textbf{Changestaff:} Your staff becomes a treant on command.

\textbf{Control Weather:} Changes weather in local area.

\textbf{Creeping Doom:} Swarms of centipedes attack at your command.

\textbf{Cure Moderate Wounds, Mass:} Cures 2d8 damage +1/level for many creatures.

\textbf{Fire Storm:} Deals 1d6/level fire damage.

\textbf{Heal:} Cures 10 points/level of damage, all diseases and mental conditions.

\textbf{Scrying, Greater:} As \textit{scrying}, but faster and longer.

\textbf{Summon Nature's Ally VII:} Calls creature to fight.

\textbf{Sunbeam:} Beam blinds and deals 4d6 damage.

\textbf{Transmute Metal to Wood: }Metal within 40 ft. becomes wood.

\textbf{True Seeing }\textsuperscript{\textbf{M}}\textbf{:} Lets you see all things 
as they really are.

\textbf{Wind Walk:} You and your allies turn vaporous and travel fast.

8TH-LEVEL DRUID SPELLS

\textbf{Animal Shapes:} One ally/level polymorphs into chosen animal.

\textbf{Control Plants:} Control actions of one or more plant creatures.

\textbf{Cure Serious Wounds, Mass:} Cures 3d8 damage +1/level for many creatures.

\textbf{Earthquake:} Intense tremor shakes 80-ft.-radius.

\textbf{Finger of Death:} Kills one subject.

\textbf{Repel Metal or Stone:} Pushes away metal and stone.

\textbf{Reverse Gravity:} Objects and creatures fall upward.

\textbf{Summon Nature's Ally VIII:} Calls creature to fight.

\textbf{Sunburst:} Blinds all within 10 ft., deals 6d6 damage.

\textbf{Whirlwind:} Cyclone deals damage and can pick up creatures.

\textbf{Word of Recall:} Teleports you back to designated place.

9TH-LEVEL DRUID SPELLS

\textbf{Antipathy:} Object or location affected by spell repels certain creatures.

\textbf{Cure Critical Wounds, Mass:} Cures 4d8 damage +1/level for many creatures.

\textbf{Elemental Swarm:} Summons multiple elementals.

\textbf{Foresight:} ``Sixth sense'' warns of impending danger.

\textbf{Regenerate:} Subject's severed limbs grow back, cures 4d8 damage +1/level 
(max +35).

\textbf{Shambler:} Summons 1d4+2 shambling mounds to fight for you.

\textbf{Shapechange }\textsuperscript{\textbf{F}}\textbf{:} Transforms you into 
any creature, and change forms once per round.

\textbf{Storm of Vengeance:} Storm rains acid, lightning, and hail.

\textbf{Summon Nature's Ally IX:} Calls creature to fight.

\textbf{Sympathy }\textsuperscript{\textbf{M}}\textbf{:} Object or location attracts 
certain creatures.

\vspace{12pt}
{\LARGE{}PALADIN SPELLS}

1ST-LEVEL PALADIN SPELLS

\textbf{Bless:} Allies gain +1 on attack rolls and +1 on saves against fear.

\textbf{Bless Water:} Makes holy water.

\textbf{Bless Weapon:} Weapon strikes true against evil foes.

\textbf{Create Water:} Creates 2 gallons/level of pure water.

\textbf{Cure Light Wounds:} Cures 1d8 damage +1/level (max +5).

\textbf{Detect Poison:} Detects poison in one creature or small object.

\textbf{Detect Undead:} Reveals undead within 60 ft.

\textbf{Divine Favor:} You gain +1 per three levels on attack and damage rolls.

\textbf{Endure Elements:} Exist comfortably in hot or cold environments.

\textbf{Magic Weapon:} Weapon gains +1 bonus.

\textbf{Protection from Chaos/Evil:} +2 to AC and saves, counter mind control, 
hedge out elementals and outsiders.

\textbf{Read Magic:} Read scrolls and spellbooks.

\textbf{Resistance:} Subject gains +1 on saving throws.

\textbf{Restoration, Lesser:} Dispels magical ability penalty or repairs 1d4 ability 
damage.

\textbf{Virtue:} Subject gains 1 temporary hp.

2ND-LEVEL PALADIN SPELLS

\textbf{Bull's Strength:} Subject gains +4 to Str for 1 min./level.

\textbf{Delay Poison:} Stops poison from harming subject for 1 hour/level.

\textbf{Eagle's Splendor:} Subject gains +4 to Cha for 1 min./level.

\textbf{Owl's Wisdom:} Subject gains +4 to Wis for 1 min./level.

\textbf{Remove Paralysis:} Frees one or more creatures from paralysis or \textit{slow 
}effect.

\textbf{Resist Energy:} Ignores 10 (or more) points of damage/attack from specified 
energy type.

\textbf{Shield Other }\textsuperscript{\textbf{F}}\textbf{:} You take half of subject's 
damage.

\textbf{Undetectable Alignment:} Conceals alignment for 24 hours.

\textbf{Zone of Truth:} Subjects within range cannot lie.

3RD-LEVEL PALADIN SPELLS

\textbf{Cure Moderate Wounds:} Cures 2d8 damage +1/level (max +10).

\textbf{Daylight:} 60-ft. radius of bright light.

\textbf{Discern Lies:} Reveals deliberate falsehoods.

\textbf{Dispel Magic:} Cancels spells and magical effects.

\textbf{Heal Mount:} As \textit{heal }on warhorse or other special mount.

\textbf{Magic Circle against Chaos:} As \textit{protection from chaos}, but 10-ft. 
radius and 10 min./level.

\textbf{Magic Circle against Evil:} As \textit{protection from evil}, but 10-ft. 
radius and 10 min./level.

\textbf{Magic Weapon, Greater:} +1 bonus/four levels (max +5).

\textbf{Prayer:} Allies +1 bonus on most rolls, enemies -1 penalty.

\textbf{Remove Blindness/Deafness:} Cures normal or magical conditions.

\textbf{Remove Curse:} Frees object or person from curse.

4TH-LEVEL PALADIN SPELLS

\textbf{Break Enchantment:} Frees subjects from enchantments, alterations, curses, 
and petrification.

\textbf{Cure Serious Wounds:} Cures 3d8 damage +1/level (max +15).

\textbf{Death Ward:} Grants immunity to death spells and negative energy effects.

\textbf{Dispel Chaos:} +4 bonus against attacks by chaotic creatures.

\textbf{Dispel Evil:} +4 bonus against attacks by evil creatures.

\textbf{Holy Sword:} Weapon becomes +5, deals +2d6 damage against evil.

\textbf{Mark of Justice:} Designates action that will trigger \textit{curse }on 
subject.

\textbf{Neutralize Poison:} Immunizes subject against poison, detoxifies venom 
in or on subject.

\textbf{Restoration }\textsuperscript{\textbf{M}}\textbf{:} Restores level and 
ability score drains.

\vspace{12pt}
{\LARGE{}RANGER SPELLS}

1ST-LEVEL RANGER SPELLS

\textbf{Alarm:} Wards an area for 2 hours/level.

\textbf{Animal Messenger:} Sends a Tiny animal to a specific place.

\textbf{Calm Animals:} Calms (2d4 + level) HD of animals.

\textbf{Charm Animal:} Makes one animal your friend.

\textbf{Delay Poison:} Stops poison from harming subject for 1 hour/level.

\textbf{Detect Animals or Plants:} Detects kinds of animals or plants.

\textbf{Detect Poison:} Detects poison in one creature or object.

\textbf{Detect Snares and Pits:} Reveals natural or primitive traps.

\textbf{Endure Elements:} Exist comfortably in hot or cold environments.

\textbf{Entangle:} Plants entangle everyone in 40-ft.-radius circle.

\textbf{Hide from Animals:} Animals can't perceive one subject/level.

\textbf{Jump:} Subject gets bonus on Jump checks.

\textbf{Longstrider:} Increases your speed.

\textbf{Magic Fang:} One natural weapon of subject creature gets +1 on attack and 
damage rolls.

\textbf{Pass without Trace:} One subject/level leaves no tracks.

\textbf{Read Magic:} Read scrolls and spellbooks.

\textbf{Resist Energy:} Ignores 10 (or more) points of damage/attack from specified 
energy type.

\textbf{Speak with Animals:} You can communicate with animals.

\textbf{Summon Nature's Ally I:} Calls animal to fight for you.

2ND-LEVEL RANGER SPELLS

\textbf{Barkskin:} Grants +2 (or higher) enhancement to natural armor.

\textbf{Bear's Endurance:} Subject gains +4 to Con for 1 min./level.

\textbf{Cat's Grace:} Subject gains +4 to Dex for 1 min./level.

\textbf{Cure Light Wounds:} Cures 1d8 damage +1/level (max +5).

\textbf{Hold Animal:} Paralyzes one animal for 1 round/level.

\textbf{Owl's Wisdom:} Subject gains +4 to Wis for 1 min./level.

\textbf{Protection from Energy:} Absorb 12 points/level of damage from one kind 
of energy.

\textbf{Snare:} Creates a magic booby trap.

\textbf{Speak with Plants:} You can talk to normal plants and plant creatures.

\textbf{Spike Growth:} Creatures in area take 1d4 damage, may be \textit{slowed.}

\textbf{Summon Nature's Ally II:} Calls animal to fight for you.

\textbf{Wind Wall:} Deflects arrows, smaller creatures, and gases.

3RD-LEVEL RANGER SPELLS

\textbf{Command Plants:} Sway the actions of one or more plant creatures.

\textbf{Cure Moderate Wounds:} Cures 2d8 damage +1/level (max +10).

\textbf{Darkvision:} See 60 ft. in total darkness.

\textbf{Diminish Plants:} Reduces size or blights growth of normal plants.

\textbf{Magic Fang, Greater:} One natural weapon of subject creature gets +1/three 
caster levels on attack and damage rolls (max +5).

\textbf{Neutralize Poison:} Immunizes subject against poison, detoxifies venom 
in or on subject.

\textbf{Plant Growth:} Grows vegetation, improves crops.

\textbf{Reduce Animal:} Shrinks one willing animal.

\textbf{Remove Disease:} Cures all diseases affecting subject.

\textbf{Repel Vermin:} Insects, spiders, and other vermin stay 10 ft. away.

\textbf{Summon Nature's Ally III:} Calls animal to fight for you.

\textbf{Tree Shape:} You look exactly like a tree for 1 hour/level.

\textbf{Water Walk:} Subject treads on water as if solid.

4TH-LEVEL RANGER SPELLS

\textbf{Animal Growth:} One animal/two levels doubles in size.

\textbf{Commune with Nature:} Learn about terrain for 1 mile/level.

\textbf{Cure Serious Wounds:} Cures 3d8 damage +1/level (max +15).

\textbf{Freedom of Movement:} Subject moves normally despite impediments.

\textbf{Nondetection }\textsuperscript{\textbf{M}}\textbf{:} Hides subject from 
divination, scrying.

\textbf{Summon Nature's Ally IV:} Calls animal to fight for you.

\textbf{Tree Stride:} Step from one tree to another far away.

\vspace{12pt}
{\LARGE{}SORCERER/WIZARD SPELLS}

0-LEVEL SORCERER/WIZARD SPELLS (CANTRIPS)

Abjur

\textbf{Resistance:} Subject gains +1 on saving throws.

Conj

\textbf{Acid Splash:} Orb deals 1d3 acid damage.

Div

\textbf{Detect Poison:} Detects poison in one creature or small object.

\textbf{Detect Magic:} Detects spells and magic items within 60 ft.

\textbf{Read Magic:} Read scrolls and spellbooks.

Ench

\textbf{Daze:} Humanoid creature of 4 HD or less loses next action.

Evoc

\textbf{Dancing Lights:} Creates torches or other lights.

\textbf{Flare:} Dazzles one creature (-1 on attack rolls).

\textbf{Light:} Object shines like a torch.

\textbf{Ray of Frost:} Ray deals 1d3 cold damage.

Illus

\textbf{Ghost Sound:} Figment sounds.

Necro

\textbf{Disrupt Undead:} Deals 1d6 damage to one undead.

\textbf{Touch of Fatigue:} Touch attack fatigues target.

Trans

\textbf{Mage Hand:} 5-pound telekinesis.

\textbf{Mending:} Makes minor repairs on an object.

\textbf{Message:} Whispered conversation at distance.

\textbf{Open/Close:} Opens or closes small or light things.

Univ

\textbf{Arcane Mark:} Inscribes a personal rune (visible or invisible).

\textbf{Prestidigitation:} Performs minor tricks.

1ST-LEVEL SORCERER/WIZARD SPELLS

Abjur

\textbf{Alarm:} Wards an area for 2 hours/level.

\textbf{Endure Elements:} Exist comfortably in hot or cold environments.

\textbf{Hold Portal:} Holds door shut.

\textbf{Protection from Chaos/Evil/Good/Law:} +2 to AC and saves, counter mind 
control, hedge out elementals and outsiders.

\textbf{Shield:} Invisible disc gives +4 to AC, blocks \textit{magic missiles.}

Conj

\textbf{Grease:} Makes 10-ft. square or one object slippery.

\textbf{Mage Armor:} Gives subject +4 armor bonus.

\textbf{Mount:} Summons riding horse for 2 hours/level.

\textbf{Obscuring Mist:} Fog surrounds you.

\textbf{Summon Monster I:} Calls extraplanar creature to fight for you.

\textbf{Unseen Servant:} Invisible force obeys your commands.

Div

\textbf{Comprehend Languages:} You understand all spoken and written languages.

\textbf{Detect Secret Doors:} Reveals hidden doors within 60 ft.

\textbf{Detect Undead:} Reveals undead within 60 ft.

\textbf{Identify }\textsuperscript{\textbf{M}}\textbf{:} Determines properties 
of magic item.

\textbf{True Strike:} +20 on your next attack roll.

Ench

\textbf{Charm Person: }Makes one person your friend.

\textbf{Hypnotism:} Fascinates 2d4 HD of creatures.

\textbf{Sleep:} Puts 4 HD of creatures into magical slumber.

Evoc

\textbf{Burning Hands:} 1d4/level fire damage (max 5d4).

\textbf{Floating Disk:} Creates 3-ft.-diameter horizontal disk that holds 100 lb./level.

\textbf{Magic Missile:} 1d4+1 damage; +1 missile per two levels above 1st (max 
5).

\textbf{Shocking Grasp:} Touch delivers 1d6/level electricity damage (max 5d6).

Illus

\textbf{Color Spray:} Knocks unconscious, blinds, and/or stuns weak creatures.

\textbf{Disguise Self:} Changes your appearance.

\textbf{Magic Aura:} Alters object's magic aura.

\textbf{Silent Image:} Creates minor illusion of your design.

\textbf{Ventriloquism:} Throws voice for 1 min./level.

Necro

\textbf{Cause Fear:} One creature of 5 HD or less flees for 1d4 rounds.

\textbf{Chill Touch:} One touch/level deals 1d6 damage and possibly 1 Str damage.

\textbf{Ray of Enfeeblement:} Ray deals 1d6 +1 per two levels Str damage.

Trans

\textbf{Animate Rope: }Makes a rope move at your command.

\textbf{Enlarge Person:} Humanoid creature doubles in size.

\textbf{Erase: }Mundane or magical writing vanishes.

\textbf{Expeditious Retreat:} Your speed increases by 30 ft.

\textbf{Feather Fall:} Objects or creatures fall slowly.

\textbf{Jump:} Subject gets bonus on Jump checks.

\textbf{Magic Weapon: }Weapon gains +1 bonus.

\textbf{Reduce Person:} Humanoid creature halves in size.

2ND-LEVEL SORCERER/WIZARD SPELLS

Abjur

\textbf{Arcane Lock }\textsuperscript{\textbf{M}}\textbf{:} Magically locks a portal 
or chest.

\textbf{Obscure Object:} Masks object against scrying.

\textbf{Protection from Arrows:} Subject immune to most ranged attacks.

\textbf{Resist Energy:} Ignores first 10 (or more) points of damage/attack from 
specified energy type.

Conj

\textbf{Acid Arrow:} Ranged touch attack; 2d4 damage for 1 round +1 round/three 
levels.

\textbf{Fog Cloud:} Fog obscures vision.

\textbf{Glitterdust:} Blinds creatures, outlines invisible creatures.

\textbf{Summon Monster II:} Calls extraplanar creature to fight for you.

\textbf{Summon Swarm:} Summons swarm of bats, rats, or spiders.

\textbf{Web:} Fills 20-ft.-radius spread with sticky spiderwebs.

Div

\textbf{Detect Thoughts:} Allows ``listening'' to surface thoughts.

\textbf{Locate Object:} Senses direction toward object (specific or type).

\textbf{See Invisibility:} Reveals invisible creatures or objects.

Ench

\textbf{Daze Monster:} Living creature of 6 HD or less loses next action.

\textbf{Hideous Laughter:} Subject loses actions for 1 round/level.

\textbf{Touch of Idiocy:} Subject takes 1d6 points of Int, Wis, and Cha damage.

Evoc

\textbf{Continual Flame }\textsuperscript{\textbf{M}}\textbf{:} Makes a permanent, 
heatless torch.

\textbf{Darkness:} 20-ft. radius of supernatural shadow.

\textbf{Flaming Sphere:} Creates rolling ball of fire, 2d6 damage, lasts 1 round/level.

\textbf{Gust of Wind:} Blows away or knocks down smaller creatures.

\textbf{Scorching Ray:} Ranged touch attack deals 4d6 fire damage, +1 ray/four 
levels (max 3).

\textbf{Shatter:} Sonic vibration damages objects or crystalline creatures.

Illus

\textbf{Blur:} Attacks miss subject 20\% of the time.

\textbf{Hypnotic Pattern:} Fascinates (2d4 + level) HD of creatures.

\textbf{Invisibility:} Subject is invisible for 1 min./level or until it attacks.

\textbf{Magic Mouth }\textsuperscript{\textbf{M}}\textbf{:} Speaks once when triggered.

\textbf{Minor Image:} As \textit{silent image}, plus some sound.

\textbf{Mirror Image:} Creates decoy duplicates of you (1d4 +1 per three levels, 
max 8).

\textbf{Misdirection:} Misleads divinations for one creature or object.

\textbf{Phantom Trap }\textsuperscript{\textbf{M}}\textbf{: }Makes item seem trapped.

Necro

\textbf{Blindness/Deafness:} Makes subject blinded or deafened.

\textbf{Command Undead:} Undead creature obeys your commands.

\textbf{False Life:} Gain 1d10 temporary hp +1/level (max +10).

\textbf{Ghoul Touch:} Paralyzes one subject, which exudes stEnch

that makes those nearby sickened.

\textbf{Scare:} Panics creatures of less than 6 HD.

\textbf{Spectral Hand:} Creates disembodied glowing hand to deliver touch attacks.

Trans

\textbf{Alter Self:} Assume form of a similar creature.

\textbf{Bear's Endurance:} Subject gains +4 to Con for 1 min./level.

\textbf{Bull's Strength:} Subject gains +4 to Str for 1 min./level.

\textbf{Cat's Grace:} Subject gains +4 to Dex for 1 min./level.

\textbf{Darkvision:} See 60 ft. in total darkness.

\textbf{Eagle's Splendor:} Subject gains +4 to Cha for 1 min./level.

\textbf{Fox's Cunning:} Subject gains +4 Int for 1 min./level.

\textbf{Knock:} Opens locked or magically sealed door.

\textbf{Levitate:} Subject moves up and down at your direction.

\textbf{Owl's Wisdom:} Subject gains +4 to Wis for 1 min./level.

\textbf{Pyrotechnics:} Turns fire into blinding light or choking smoke.

\textbf{Rope Trick:} As many as eight creatures hide in extradimensional space.

\textbf{Spider Climb:} Grants ability to walk on walls and ceilings.

\textbf{Whispering Wind:} Sends a short message 1 mile/level.

3RD-LEVEL SORCERER/WIZARD SPELLS

Abjur

\textbf{Dispel Magic:} Cancels magical spells and effects.

\textbf{Explosive Runes:} Deals 6d6 damage when read.

\textbf{Magic Circle against Chaos/Evil/Good/Law:} As \textit{protection }spells, 
but 10-ft. radius and 10 min./level.

\textbf{Nondetection }\textsuperscript{\textbf{M}}\textbf{:} Hides subject from 
divination, scrying.

\textbf{Protection from Energy:} Absorb 12 points/level of damage from one kind 
of energy.

Conj

\textbf{Phantom Steed:} Magic horse appears for 1 hour/level.

\textbf{Sepia Snake Sigil }\textsuperscript{\textbf{M}}\textbf{:} Creates text 
symbol that immobilizes reader.

\textbf{Sleet Storm:} Hampers vision and movement.

\textbf{Stinking Cloud:} Nauseating vapors, 1 round/level.

\textbf{Summon Monster III:} Calls extraplanar creature to fight for you.

Div

\textbf{Arcane Sight:} Magical auras become visible to you.

\textbf{Clairaudience/Clairvoyance:} Hear or see at a distance for 1 min./level.

\textbf{Tongues:} Speak any language.

Ench

\textbf{Deep Slumber:} Puts 10 HD of creatures to sleep.

\textbf{Heroism:} Gives +2 bonus on attack rolls, saves, skill checks.

\textbf{Hold Person:} Paralyzes one humanoid for 1 round/level.

\textbf{Rage:} Subjects gains +2 to Str and Con, +1 on Will saves, -2 to AC.

\textbf{Suggestion:} Compels subject to follow stated course of action.

Evoc

\textbf{Daylight:} 60-ft. radius of bright light.

\textbf{Fireball:} 1d6 damage per level, 20-ft. radius.

\textbf{Lightning Bolt:} Electricity deals 1d6/level damage.

\textbf{Tiny Hut:} Creates shelter for ten creatures.

\textbf{Wind Wall:} Deflects arrows, smaller creatures, and gases.

Illus

\textbf{Displacement:} Attacks miss subject 50\%.

\textbf{Illusory Script }\textsuperscript{\textbf{M}}\textbf{:} Only intended reader 
can decipher.

\textbf{Invisibility Sphere:} Makes everyone within 10 ft. invisible.

\textbf{Major Image:} As \textit{silent image}, plus sound, smell and thermal effects.

Necro

\textbf{Gentle Repose:} Preserves one corpse.

\textbf{Halt Undead:} Immobilizes undead for 1 round/level.

\textbf{Ray of Exhaustion:} Ray makes subject exhausted.

\textbf{Vampiric Touch:} Touch deals 1d6/two levels damage; caster gains damage 
as hp.

Trans

\textbf{Blink:} You randomly vanish and reappear for 1 round/level.

\textbf{Flame Arrow:} Arrows deal +1d6 fire damage.

\textbf{Fly:} Subject flies at speed of 60 ft.

\textbf{Gaseous Form:} Subject becomes insubstantial and can fly slowly.

\textbf{Haste:} One creature/level moves faster, +1 on attack rolls, AC, and Reflex 
saves.

\textbf{Keen Edge:} Doubles normal weapon's threat range.

\textbf{Magic Weapon, Greater:} +1/four levels (max +5).

\textbf{Secret Page:} Changes one page to hide its real content.

\textbf{Shrink Item:} Object shrinks to one-sixteenth size.

\textbf{Slow:} One subject/level takes only one action/round, -1 to AC, reflex 
saves, and attack rolls.

\textbf{Water Breathing:} Subjects can breathe underwater.

4TH-LEVEL SORCERER/WIZARD SPELLS

Abjur

\textbf{Dimensional Anchor:} Bars extradimensional movement.

\textbf{Fire Trap }\textsuperscript{\textbf{M}}\textbf{:} Opened object deals 1d4 
damage +1/level.

\textbf{Globe of Invulnerability, Lesser:} Stops 1st- through 3rd-level spell effects.

\textbf{Remove Curse:} Frees object or person from curse.

\textbf{Stoneskin }\textsuperscript{\textbf{M}}\textbf{:} Ignore 10 points of damage 
per attack.

Conj

\textbf{Black Tentacles:} Tentacles grapple all within 20 ft. spread.

\textbf{Dimension Door:} Teleports you short distance.

\textbf{Minor Creation:} Creates one cloth or wood object.

\textbf{Secure Shelter:} Creates sturdy cottage.

\textbf{Solid Fog:} Blocks vision and slows movement.

\textbf{Summon Monster IV:} Calls extraplanar creature to fight for you.

Div

\textbf{Arcane Eye:} Invisible floating eye moves 30 ft./round.

\textbf{Detect Scrying:} Alerts you of magical eavesdropping.

\textbf{Locate Creature:} Indicates direction to familiar creature.

\textbf{Scrying }\textsuperscript{\textbf{F}}\textbf{:} Spies on subject from a 
distance.

Ench

\textbf{Charm Monster:}Makes monster believe it is your ally.

\textbf{Confusion:} Subjects behave oddly for 1 round/level.

\textbf{Crushing Despair:} Subjects take -2 on attack rolls, damage rolls, saves, 
and checks.

\textbf{Geas, Lesser:} Commands subject of 7 HD or less.

Evoc

\textbf{Fire Shield:} Creatures attacking you take fire damage; you're protected 
from heat or cold.

\textbf{Ice Storm:} Hail deals 5d6 damage in cylinder 40 ft. across.

\textbf{Resilient Sphere:} Force globe protects but traps one subject.

\textbf{Shout:} Deafens all within cone and deals 5d6 sonic damage.

\textbf{Wall of Fire:} Deals 2d4 fire damage out to 10 ft. and 1d4 out to 20 ft. 
Passing through wall deals 2d6 damage +1/level.

\textbf{Wall of Ice:} \textit{Ice plane }creates wall with 15 hp +1/level, or \textit{hemisphere 
}can trap creatures inside.

Illus

\textbf{Hallucinatory Terrain:} Makes one type of terrain appear like another (field 
into forest, or the like).

\textbf{Illusory Wall:} Wall, floor, or ceiling looks real, but anything can pass 
through.

\textbf{Invisibility, Greater:} As \textit{invisibility, }but subject can attack 
and stay invisible.

\textbf{Phantasmal Killer:} Fearsome illusion kills subject or deals 3d6 damage.

\textbf{Rainbow Pattern:} Lights fascinate 24 HD of creatures.

\textbf{Shadow Conjuration:} Mimics conjuration below 4th level, but only 20\% 
real.

Necro

\textbf{Animate Dead }\textsuperscript{\textbf{M}}\textbf{:} Creates undead skeletons 
and zombies.

\textbf{Bestow Curse:} -6 to an ability score; -4 on attack rolls, saves, and checks; 
or 50\% chance of losing each action.

\textbf{Contagion:} Infects subject with chosen disease.

\textbf{Enervation:} Subject gains 1d4 negative levels.

\textbf{Fear:} Subjects within cone flee for 1 round/level.

Trans

\textbf{Enlarge Person, Mass:} Enlarges several creatures.

\textbf{Mnemonic Enhancer }\textsuperscript{\textbf{F}}\textbf{:} \textit{Wizard 
only. }Prepares extra spells or retains one just cast.

\textbf{Polymorph:} Gives one willing subject a new form.

\textbf{Reduce Person, Mass:} Reduces several creatures.

\textbf{Stone Shape:} Sculpts stone into any shape.

5TH-LEVEL SORCERER/WIZARD SPELLS

Abjur

\textbf{Break Enchantment:} Frees subjects from enchantments, alterations, curses, 
and petrification.

\textbf{Dismissal:} Forces a creature to return to native plane.

\textbf{Mage's Private Sanctum:} Prevents anyone from viewing or scrying an area 
for 24 hours.

Conj

\textbf{Cloudkill:} Kills 3 HD or less; 4-6 HD save or die, 6+ HD take Con damage.

\textbf{Mage's Faithful Hound:} Phantom dog can guard, attack.

\textbf{Major Creation:} As \textit{minor creation, }plus stone and metal.

\textbf{Planar Binding, Lesser:} Traps extraplanar creature of 6 HD or less until 
it performs a task.

\textbf{Secret Chest }\textsuperscript{\textbf{F}}\textbf{:} Hides expensive chest 
on Ethereal Plane; you retrieve it at will.

\textbf{Summon Monster V:} Calls extraplanar creature to fight for you.

\textbf{Teleport:} Instantly transports you as far as 100 miles/level.

\textbf{Wall of Stone:} Creates a stone wall that can be shaped.

Div

\textbf{Contact Other Plane:} Lets you ask question of extraplanar entity.

\textbf{Prying Eyes:} 1d4 +1/level floating eyes scout for you.

\textbf{Telepathic Bond:} Link lets allies communicate.

Ench

\textbf{Dominate Person:} Controls humanoid telepathically.

\textbf{Feeblemind:} Subject's Int and Cha drop to 1.

\textbf{Hold Monster:} As \textit{hold person}, but any creature.

\textbf{Mind Fog:} Subjects in fog get -10 to Wis and Will checks.

\textbf{Symbol of Sleep }\textsuperscript{\textbf{M}}\textbf{:} Triggered rune 
puts nearby creatures into catatonic slumber.

Evoc

\textbf{Cone of Cold:} 1d6/level cold damage.

\textbf{Interposing Hand:} Hand provides cover against one opponent.

\textbf{Sending:} Delivers short message anywhere, instantly.

\textbf{Wall of Force:}Wall is immune to damage.

Illus

\textbf{Dream:} Sends message to anyone sleeping.

\textbf{False Vision }\textsuperscript{\textbf{M}}\textbf{:} Fools scrying with 
an illusion.

\textbf{Mirage Arcana:} As \textit{hallucinatory terrain, }plus structures.

\textbf{Nightmare:} Sends vision dealing 1d10 damage, fatigue.

\textbf{Persistent Image:} As \textit{major image}, but no concentration required.

\textbf{Seeming:} Changes appearance of one person per two levels.

\textbf{Shadow Evocation:} Mimics evocation below 5th level, but only 20\% real.

Necro

\textbf{Blight:} Withers one plant or deals 1d6/level damage to plant creature.

\textbf{Magic Jar }\textsuperscript{\textbf{F}}\textbf{:} Enables possession of 
another creature.

\textbf{Symbol of Pain }\textsuperscript{\textbf{M}}\textbf{:} Triggered rune wracks 
nearby creatures with pain.

\textbf{Waves of Fatigue:} Several targets become fatigued.

Trans

\textbf{Animal Growth:} One animal/two levels doubles in size.

\textbf{Baleful Polymorph:} Transforms subject into harmless animal.

\textbf{Fabricate:} Transforms raw materials into finished items.

\textbf{Overland Flight:} You fly at a speed of 40 ft. and can hustle over long 
distances.

\textbf{Passwall:} Creates passage through wood or stone wall.

\textbf{Telekinesis:} Moves object, attacks creature, or hurls object or creature.

\textbf{Transmute Mud to Rock:} Transforms two 10-ft. cubes per level.

\textbf{Transmute Rock to Mud:} Transforms two 10-ft. cubes per level.

Univ

\textbf{Permanency }\textsuperscript{\textbf{X}}\textbf{:}Makes certain spells 
permanent.

6TH-LEVEL SORCERER/WIZARD SPELLS

Abjur

\textbf{Antimagic Field:} Negates magic within 10 ft.

\textbf{Dispel Magic, Greater:} As \textit{dispel magic, }but +20 on check.

\textbf{Globe of Invulnerability:} As \textit{lesser globe of invulnerability}, 
plus 4th-level spell effects.

\textbf{Guards and Wards:} Array of magic effects protect area.

\textbf{Repulsion:} Creatures can't approach you.

Conj

\textbf{Acid Fog:} Fog deals acid damage.

\textbf{Planar Binding:} As \textit{lesser planar binding, }but up to 12 HD.

\textbf{Summon Monster VI:} Calls extraplanar creature to fight for you.

\textbf{Wall of Iron }\textsuperscript{\textbf{M}}\textbf{:} 30 hp/four levels; 
can topple onto foes.

Div

\textbf{Analyze Dweomer }\textsuperscript{\textbf{F}}\textbf{:} Reveals magical 
aspects of subject.

\textbf{Legend Lore }\textsuperscript{\textbf{M}}{\scriptsize{}\textbf{ }}\textsuperscript{\textbf{F}}\textbf{:} 
Lets you learn tales about a person, place, or thing.

\textbf{True Seeing }\textsuperscript{\textbf{M}}\textbf{:} Lets you see all things 
as they really are.

Ench

\textbf{Geas/Quest:} As \textit{lesser geas, }plus it affects any creature.

\textbf{Heroism, Greater:} Gives +4 bonus on attack rolls, saves, skill checks; 
immunity to fear; temporary hp.

\textbf{Suggestion, Mass:} As \textit{suggestion, }plus one subject/level.

\textbf{Symbol of Persuasion }\textsuperscript{\textbf{M}}\textbf{:} Triggered 
rune charms nearby creatures.

Evoc

\textbf{Chain Lightning:} 1d6/level damage; 1 secondary bolt/level each deals half 
damage.

\textbf{Contingency }\textsuperscript{\textbf{F}}\textbf{:} Sets trigger condition 
for another spell.

\textbf{Forceful Hand:} Hand pushes creatures away.

\textbf{Freezing Sphere:} Freezes water or deals cold damage.

Illus

\textbf{Mislead:} Turns you invisible and creates illusory double.

\textbf{Permanent Image:} Includes sight, sound, and smell.

\textbf{Programmed Image }\textsuperscript{\textbf{M}}\textbf{:} As \textit{major 
image}, plus triggered by event.

\textbf{Shadow Walk:} Step into shadow to travel rapidly.

\textbf{Veil:} Changes appearance of group of creatures.

Necro

\textbf{Circle of Death }\textsuperscript{\textbf{M}}\textbf{:} Kills 1d4/level 
HD of creatures.

\textbf{Create Undead }\textsuperscript{\textbf{M}}\textbf{:} Creates ghouls, ghasts, 
mummies, or mohrgs.

\textbf{Eyebite:} Target becomes panicked, sickened, and comatose.

\textbf{Symbol of Fear }\textsuperscript{\textbf{M}}\textbf{:} Triggered rune panics 
nearby creatures.

\textbf{Undeath to Death }\textsuperscript{\textbf{M}}\textbf{:} Destroys 1d4/level 
HD of undead (max 20d4).

Trans

\textbf{Bear's Endurance, Mass:} As \textit{bear's endurance}, affects one subject/level.

\textbf{Bull's Strength, Mass:} As \textit{bull's strength}, affects one subject/ 
level.

\textbf{Cat's Grace, Mass:} As \textit{cat's grace}, affects one subject/level.

\textbf{Control Water:} Raises or lowers bodies of water.

\textbf{Disintegrate:}Makes one creature or object vanish.

\textbf{Eagle's Splendor, Mass:} As \textit{eagle's splendor}, affects one subject/level.

\textbf{Flesh to Stone:} Turns subject creature into statue.

\textbf{Fox's Cunning, Mass:} As \textit{fox's cunning}, affects one subject/ level.

\textbf{Mage's Lucubration:} \textit{Wizard only. }Recalls spell of 5th level or 
lower.

\textbf{Move Earth:} Digs trenches and build hills.

\textbf{Owl's Wisdom, Mass:} As \textit{owl's wisdom}, affects one subject/ level.

\textbf{Stone to Flesh:} Restores petrified creature.

\textbf{Transformation }\textsuperscript{\textbf{M}}\textbf{:} You gain combat 
bonuses.

7TH-LEVEL SORCERER/WIZARD SPELLS

Abjur

\textbf{Banishment:} Banishes 2 HD/level of extraplanar creatures.

\textbf{Sequester:} Subject is invisible to sight and scrying; renders creature 
comatose.

\textbf{Spell Turning:} Reflect 1d4+6 spell levels back at caster.

Conj

\textbf{Instant Summons }\textsuperscript{\textbf{M}}\textbf{:} Prepared object 
appears in your hand.

\textbf{Mage's Magnificent Mansion }\textsuperscript{\textbf{F}}\textbf{:} Door 
leads to extradimensional mansion.

\textbf{Phase Door:} Creates an invisible passage through wood or stone.

\textbf{Plane Shift }\textsuperscript{\textbf{F}}\textbf{:} As many as eight subjects 
travel to another plane.

\textbf{Summon Monster VII:} Calls extraplanar creature to fight for you.

\textbf{Teleport, Greater:} As \textit{teleport, }but no range limit and no off-target 
arrival.

\textbf{Teleport Object:} As \textit{teleport, }but affects a touched object.

Div

\textbf{Arcane Sight, Greater:} As \textit{arcane sight}, but also reveals magic 
effects on creatures and objects.

\textbf{Scrying, Greater:} As \textit{scrying}, but faster and longer.

\textbf{Vision }\textsuperscript{\textbf{M}}{\scriptsize{}\textbf{ }}\textsuperscript{\textbf{X}}\textbf{:} 
As \textit{legend lore}, but quicker and strenuous.

Ench

\textbf{Hold Person, Mass:} As \textit{hold person}, but all within 30 ft.

\textbf{Insanity:} Subject suffers continuous \textit{confusion}.

\textbf{Power Word Blind:} Blinds creature with 200 hp or less.

\textbf{Symbol of Stunning }\textsuperscript{\textbf{M}}\textbf{:} Triggered rune 
stuns nearby creatures.

Evoc

\textbf{Delayed Blast Fireball:} 1d6/level fire damage; you can postpone blast 
for 5 rounds.

\textbf{Forcecage }\textsuperscript{\textbf{M}}\textbf{:} Cube or cage of force 
imprisons all inside.

\textbf{Grasping Hand:} Hand provides cover, pushes, or grapples.

\textbf{Mage's Sword }\textsuperscript{\textbf{F}}\textbf{:} Floating magic blade 
strikes opponents.

\textbf{Prismatic Spray:} Rays hit subjects with variety of effects.

Illus

\textbf{Invisibility, Mass:} As \textit{invisibility}, but affects all in range.

\textbf{Project Image:} Illusory double can talk and cast spells.

\textbf{Shadow Conjuration, Greater:} As \textit{shadow conjuration, }but up to 
6th level and 60\% real.

\textbf{Simulacrum }\textsuperscript{\textbf{M}}{\scriptsize{}\textbf{ }}\textsuperscript{\textbf{X}}\textbf{:} 
Creates partially real double of a creature.

Necro

\textbf{Control Undead:} Undead don't attack you while under your command.

\textbf{Finger of Death:} Kills one subject.

\textbf{Symbol of Weakness }\textsuperscript{\textbf{M}}\textbf{:} Triggered rune 
weakens nearby creatures.

\textbf{Waves of Exhaustion:} Several targets become exhausted.

Trans

\textbf{Control Weather:} Changes weather in local area.

\textbf{Ethereal Jaunt:} You become ethereal for 1 round/level.

\textbf{Reverse Gravity:} Objects and creatures fall upward.

\textbf{Statue:} Subject can become a statue at will.

Univ

\textbf{Limited Wish }\textsuperscript{\textbf{X}}\textbf{:} Alters reality---within 
spell limits.

8TH-LEVEL SORCERER/WIZARD SPELLS

Abjur

\textbf{Dimensional Lock:} Teleportation and interplanar travel blocked for one 
day/level.

\textbf{Mind Blank:} Subject is immune to mental/emotional magic and scrying.

\textbf{Prismatic Wall: }Wall's colors have array of effects.

\textbf{Protection from Spells }\textsuperscript{\textbf{M}}{\scriptsize{}\textbf{ 
}}\textsuperscript{\textbf{F}}\textbf{:} Confers +8 resistance bonus.

Conj

\textbf{Incendiary Cloud:} Cloud deals 4d6 fire damage/round.

\textbf{Maze:} Traps subject in extradimensional maze.

\textbf{Planar Binding, Greater:} As \textit{lesser planar binding, }but up to 
18 HD.

\textbf{Summon Monster VIII:} Calls extraplanar creature to fight for you.

\textbf{Trap the Soul }\textsuperscript{\textbf{M}}{\scriptsize{}\textbf{ }}\textsuperscript{\textbf{F}}\textbf{:} 
Imprisons subject within gem.

Div

\textbf{Discern Location:} Reveals exact location of creature or object.

\textbf{Moment of Prescience:} You gain insight bonus on single attack roll, check, 
or save.

\textbf{Prying Eyes, Greater:} As \textit{prying eyes}, but eyes have \textit{true 
seeing}.

Ench

\textbf{Antipathy:} Object or location affected by spell repels certain creatures.

\textbf{Binding }\textsuperscript{\textbf{M}}\textbf{:} Utilizes an array of techniques 
to imprison a creature.

\textbf{Charm Monster, Mass:} As \textit{charm monster, }but all within 30 ft.

\textbf{Demand:} As \textit{sending, }plus you can send \textit{suggestion.}

\textbf{Irresistible Dance:} Forces subject to dance.

\textbf{Power Word Stun:} Stuns creature with 150 hp or less.

\textbf{Symbol of Insanity }\textsuperscript{\textbf{M}}\textbf{:} Triggered rune 
renders nearby creatures insane.

\textbf{Sympathy }\textsuperscript{\textbf{F}}\textbf{:} Object or location attracts 
certain creatures.

Evoc

\textbf{Clenched Fist:} Large hand provides cover, pushes, or attacks your foes.

\textbf{Polar Ray:} Ranged touch attack deals 1d6/level cold damage.

\textbf{Shout, Greater:} Devastating yell deals 10d6 sonic damage; stuns creatures, 
damages objects.

\textbf{Sunburst:} Blinds all within 10 ft., deals 6d6 damage.

Illus

\textbf{Scintillating Pattern:} Twisting colors \textit{confuse, }stun, or render 
unconscious.

\textbf{Screen:} Illusion hides area from vision, scrying.

\textbf{Shadow Evocation, Greater:} As \textit{shadow evocation, }but up to 7th 
level and 60\% real.

\textbf{Telekinetic Sphere:} As\textit{ resilient sphere, }but you move sphere 
telekinetically.

Necro

\textbf{Clone }\textsuperscript{\textbf{M}}{\scriptsize{}\textbf{ }}\textsuperscript{\textbf{F}}\textbf{:} 
Duplicate awakens when original dies.

\textbf{Create Greater Undead }\textsuperscript{\textbf{M}}\textbf{:} Create shadows, 
wraiths, spectres, or devourers.

\textbf{Horrid Wilting:} Deals 1d6/level damage within 30 ft.

\textbf{Symbol of Death }\textsuperscript{\textbf{M}}\textbf{:} Triggered rune 
slays nearby creatures.

Trans

\textbf{Iron Body:} Your body becomes living iron.

\textbf{Polymorph Any Object:} Changes any subject into anything else.

\textbf{Temporal Stasis }\textsuperscript{\textbf{M}}\textbf{:} Puts subject into 
suspended animation.

9TH-LEVEL SORCERER/WIZARD SPELLS

Abjur

\textbf{Freedom:} Releases creature from \textit{imprisonment.}

\textbf{Imprisonment:} Entombs subject beneath the earth.

\textbf{Mage's Disjunction:} Dispels magic, disenchants magic items.

\textbf{Prismatic Sphere:} As \textit{prismatic wall}, but surrounds on all sides.

Conj

\textbf{Gate }\textsuperscript{\textbf{X}}\textbf{:} Connects two planes for travel 
or summoning.

\textbf{Refuge }\textsuperscript{\textbf{M}}\textbf{:} Alters item to transport 
its possessor to you.

\textbf{Summon Monster IX:} Calls extraplanar creature to fight for you.

\textbf{Teleportation Circle }\textsuperscript{\textbf{M}}\textbf{:} Circle teleports 
any creature inside to designated spot.

Div

\textbf{Foresight:} ``Sixth sense'' warns of impending danger.

Ench

\textbf{Dominate Monster:} As \textit{dominate person, }but any creature.

\textbf{Hold Monster, Mass:} As \textit{hold monster, }but all within 30 ft.

\textbf{Power Word Kill:} Kills one creature with 100 hp or less.

Evoc

\textbf{Crushing Hand:} Large hand provides cover, pushes, or crushes your foes.

\textbf{Meteor Swarm:} Four exploding spheres each deal 6d6 fire damage.

Illus

\textbf{Shades:} As \textit{shadow conjuration}, but up to 8th level and 80\% real.

\textbf{Weird:} As \textit{phantasmal killer, }but affects all within 30 ft.

Necro

\textbf{Astral Projection }\textsuperscript{\textbf{M}}\textbf{:} Projects you 
and companions onto Astral Plane.

\textbf{Energy Drain:} Subject gains 2d4 negative levels.

\textbf{Soul Bind }\textsuperscript{\textbf{F}}\textbf{:} Traps newly dead soul 
to prevent \textit{resurrection}.

\textbf{Wail of the Banshee:} Kills one creature/level.

Trans

\textbf{Etherealness:} Travel to Ethereal Plane with companions.

\textbf{Shapechange }\textsuperscript{\textbf{F}}\textbf{:} Transforms you into 
any creature, and change forms once per round.

\textbf{Time Stop:} You act freely for 1d4+1 rounds.

Univ

\textbf{Wish }\textsuperscript{\textbf{X}}\textbf{:} As \textit{limited wish, }but 
with fewer limits.

\newpage

\end{document}
