%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{array}
\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

This material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

{\LARGE{}SPELLS (S)}

\vspace{12pt}
Sanctuary

Abjuration

\textbf{Level:} Clr 1, Protection 1

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 1 round/level

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} No

Any opponent attempting to strike or otherwise directly attack the warded creature, 
even with a targeted spell, must attempt a Will save. If the save succeeds, the 
opponent can attack normally and is unaffected by that casting of the spell. If 
the save fails, the opponent can't follow through with the attack, that part of 
its action is lost, and it can't directly attack the warded creature for the duration 
of the spell. Those not attempting to attack the subject remain unaffected. This 
spell does not prevent the warded creature from being attacked or affected by area 
or effect spells. The subject cannot attack without breaking the spell but may 
use nonattack spells or otherwise act.

\vspace{12pt}
Scare

Necromancy [Fear, Mind-Affecting]

\textbf{Level:} Brd 2, Sor/Wiz 2

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Targets:} One living creature per three levels, no two of which can be 
more than 30 ft. apart

\textbf{Duration:} 1 round/level or 1 round; see text for \textit{cause fear}

\textbf{Saving Throw: }Will partial

\textbf{Spell Resistance:} Yes

This spell functions like \textit{cause fear, }except that it causes all targeted 
creatures of less than 6 HD to become frightened.

\textit{Material Component: }A bit of bone from an undead skeleton, zombie, ghoul, 
ghast, or mummy.

\vspace{12pt}
Scintillating Pattern

Illusion (Pattern) [Mind-Affecting]

\textbf{Level:} Sor/Wiz 8

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} Colorful lights in a 20-ft.-radius spread

\textbf{Duration:} Concentration + 2 rounds

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

A twisting pattern of discordant, coruscating colors weaves through the air, affecting 
creatures within it. The spell affects a total number of Hit Dice of creatures 
equal to your caster level (maximum 20). Creatures with the fewest HD are affected 
first; and, among creatures with equal HD, those who are closest to the spell's 
point of origin are affected first. Hit Dice that are not sufficient to affect 
a creature are wasted. The spell affects each subject according to its Hit Dice.

\textit{6 or less: }Unconscious for 1d4 rounds, then stunned for 1d4 rounds, and 
then \textit{confused }for 1d4 rounds. (Treat an unconscious result as stunned 
for nonliving creatures.)

\textit{7 to 12: }Stunned for 1d4 rounds, then \textit{confused }for 1d4 rounds. 

\textit{13 or more: Confused }for 1d4 rounds. 

Sightless creatures are not affected by \textit{scintillating pattern}.

\textit{Material Component: }A small crystal prism.

\vspace{12pt}
Scorching Ray

Evocation [Fire]

\textbf{Level:} Sor/Wiz 2

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} One or more rays

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

You blast your enemies with fiery rays. You may fire one ray, plus one additional 
ray for every four levels beyond 3rd (to a maximum of three rays at 11th level). 
Each ray requires a ranged touch attack to hit and deals 4d6 points of fire damage.

The rays may be fired at the same or different targets, but all bolts must be aimed 
at targets within 30 feet of each other and fired simultaneously.

\vspace{12pt}
Screen

Illusion (Glamer)

\textbf{Level:} Sor/Wiz 8, Trickery 7

\textbf{Components:} V, S

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Area:} 30-ft. cube/level (S)

\textbf{Duration:} 24 hours

\textbf{Saving Throw:} None or Will disbelief (if interacted with); see text

\textbf{Spell Resistance:} No

This spell combines several elements to create a powerful protection from scrying 
and direct observation. When casting the spell, you dictate what will and will 
not be observed in the spell's area. The illusion created must be stated in general 
terms. Once the conditions are set, they cannot be changed.

Attempts to scry the area automatically detect the image stated by you with no 
save allowed. Sight and sound are appropriate to the illusion created. 

Direct observation may allow a save (as per a normal illusion), if there is cause 
to disbelieve what is seen. Even entering the area does not cancel the illusion 
or necessarily allow a save, assuming that hidden beings take care to stay out 
of the way of those affected by the illusion.

\vspace{12pt}
Scrying

Divination (Scrying)

\textbf{Level:} Brd 3, Clr 5, Drd 4, Sor/Wiz 4

\textbf{Components:} V, S, M/DF, F

\textbf{Casting Time:} 1 hour

\textbf{Range:} See text

\textbf{Effect }Magical sensor

\textbf{Duration:} 1 min./level

\textbf{Saving Throw:} Will negates

\textbf{Spell Resistance:} Yes

You can see and hear some creature, which may be at any distance. If the subject 
succeeds on a Will save, the scrying attempt simply fails. The difficulty of the 
save depends on how well you know the subject and what sort of physical connection 
(if any) you have to that creature. Furthermore, if the subject is on another plane, 
it gets a +5 bonus on its Will save.

\begin{tabular}{|>{\raggedright}p{215pt}|>{\raggedright}p{100pt}|}
\hline
\subsection*{K\textbf{nowledge}} & \subsection*{W\textbf{ill Save Modifier}}\tabularnewline
\hline
None\textsuperscript{\textbf{1}} & +10\tabularnewline
\hline
Secondhand (you have heard of the subject) & +5\tabularnewline
\hline
Firsthand (you have met the subject) & +0\tabularnewline
\hline
Familiar (you know the subject well)- & 5\tabularnewline
\hline
\multicolumn{2}{|p{316pt}|}{1 You must have some sort of connection to a creature 
you have no knowledge of.}\tabularnewline
\hline
C\textbf{onnection} & \subsubsection*{W\textbf{ill Save Modifier}}\tabularnewline
\hline
Likeness or picture- & 2\tabularnewline
\hline
Possession or garment- & 4\tabularnewline
\hline
Body part, lock of hair, bit of nail, etc.- & 10\tabularnewline
\hline
\end{tabular}

If the save fails, you can see and hear the subject and the subject's immediate 
surroundings (approximately 10 feet in all directions of the subject). If the subject 
moves, the sensor follows at a speed of up to 150 feet.

As with all divination (scrying) spells, the sensor has your full visual acuity, 
including any magical effects. In addition, the following spells have a 5\% chance 
per caster level of operating through the sensor: \textit{detect chaos}, \textit{detect 
evil}, \textit{detect good}, \textit{detect law}, \textit{detect magic}, and \textit{message}.

If the save succeeds, you can't attempt to scry on that subject again for at least 
24 hours.

\textit{Arcane Material Component: }The eye of a hawk, an eagle, or a roc, plus 
nitric acid, copper, and zinc.

\textit{Wizard, Sorcerer, or Bard Focus: }A mirror of finely wrought and highly 
polished silver costing not less than 1,000 gp. The mirror must be at least 2 feet 
by 4 feet.

\textit{Cleric Focus: }A holy water font costing not less than 100 gp.

\textit{Druid Focus: }A natural pool of water. 

\vspace{12pt}
Scrying, Greater

Divination (Scrying)

\textbf{Level:} Brd 6, Clr 7, Drd 7, Sor/Wiz 7

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Duration:} 1 hour/level

This spell functions like \textit{scrying}, except as noted above. Additionally, 
all of the following spells function reliably through the sensor: \textit{detect 
chaos}, \textit{detect evil}, \textit{detect good}, \textit{detect law}, \textit{detect 
magic}, \textit{message}, \textit{read magic}, and \textit{tongues}.

\vspace{12pt}
Sculpt Sound

Transmutation

\textbf{Level:} Brd 3

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} One creature or object/level, no two of which can be more than 
30 ft. apart

\textbf{Duration:} 1 hour/level (D)

\textbf{Saving Throw: }Will negates (object)

\textbf{Spell Resistance:} Yes (object)

You change the sounds that creatures or objects make. You can create sounds where 
none exist, deaden sounds, or transform sounds into other sounds. All affected 
creatures or objects must be transmuted in the same way. Once the transmutation 
is made, you cannot change it.

You can change the qualities of sounds but cannot create words with which you are 
unfamiliar yourself.

A spellcaster whose voice is changed dramatically is unable to cast spells with 
verbal components.

\vspace{12pt}
Searing Light

Evocation

\textbf{Level:} Clr 3, Sun 3

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Effect:} Ray

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

Focusing divine power like a ray of the sun, you project a blast of light from 
your open palm. You must succeed on a ranged touch attack to strike your target. 
A creature struck by this ray of light takes 1d8 points of damage per two caster 
levels (maximum 5d8). An undead creature takes 1d6 points of damage per caster 
level (maximum 10d6), and an undead creature particularly vulnerable to bright 
light takes 1d8 points of damage per caster level (maximum 10d8). A construct or 
inanimate object takes only 1d6 points of damage per two caster levels (maximum 
5d6).

\vspace{12pt}
Secret Chest

Conjuration (Summoning)

\textbf{Level:} Sor/Wiz 5

\textbf{Components:} V, S, F

\textbf{Casting Time:} 10 minutes

\textbf{Range:} See text

\textbf{Target:} One chest and up to 1 cu. ft. of goods/caster level

\textbf{Duration:} Sixty days or until discharged

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You hide a chest on the Ethereal Plane for as long as sixty days and can retrieve 
it at will. The chest can contain up to 1 cubic foot of material per caster level 
(regardless of the chest's actual size, which is about 3 feet by 2 feet by 2 feet). 
If any living creatures are in the chest, there is a 75\% chance that the spell 
simply fails. Once the chest is hidden, you can retrieve it by concentrating (a 
standard action), and it appears next to you.

The chest must be exceptionally well crafted and expensive, constructed for you 
by master crafters. The cost of such a chest is never less than 5,000 gp. Once 
it is constructed, you must make a tiny replica (of the same materials and perfect 
in every detail), so that the miniature of the chest appears to be a perfect copy. 
(The replica costs 50 gp.) You can have but one pair of these chests at any given 
time---even a \textit{wish }spell does not allow more. The chests are nonmagical 
and can be fitted with locks, wards, and so on, just as any normal chest can be.

To hide the chest, you cast the spell while touching both the chest and the replica. 
The chest vanishes into the Ethereal Plane. You need the replica to recall the 
chest. After sixty days, there is a cumulative chance of 5\% per day that the chest 
is irretrievably lost. If the miniature of the chest is lost or destroyed, there 
is no way, not even with a \textit{wish }spell, that the large chest can be summoned 
back, although an extraplanar expedition might be mounted to find it.

Living things in the chest eat, sleep, and age normally, and they die if they run 
out of food, air, water, or whatever they need to survive.

\textit{Focus: }The chest and its replica.

\vspace{12pt}
Secret Page

Transmutation

\textbf{Level:} Brd 3, Sor/Wiz 3

\textbf{Components:} V, S, M

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Touch

\textbf{Target:} Page touched, up to 3 sq. ft. in size

\textbf{Duration:} Permanent

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

S\textit{ecret page }alters the contents of a page so that they appear to be something 
entirely different. The text of a spell can be changed to show even another spell. 
\textit{Explosive runes }or \textit{sepia snake sigil }can be cast upon the \textit{secret 
page.}

A \textit{comprehend languages }spell alone cannot reveal a \textit{secret page'}s 
contents. You are able to reveal the original contents by speaking a special word. 
You can then peruse the actual page, and return it to its \textit{secret page }form 
at will. You can also remove the spell by double repetition of the special word. 
A \textit{detect magic }spell reveals dim magic on the page in question but does 
not reveal its true contents. T\textit{rue seeing }reveals the presence of the 
hidden material but does not reveal the contents unless cast in combination with 
\textit{comprehend languages. }A \textit{secret page }spell can be dispelled, and 
the hidden writings can be destroyed by means of an \textit{erase }spell.

\textit{Material Component: }Powdered herring scales and will-o'-wisp essence.

\vspace{12pt}
Secure Shelter

Conjuration (Creation)

\textbf{Level:} Brd 4, Sor/Wiz 4

\textbf{Components:} V, S, M, F; see text

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} 20-ft.-square structure

\textbf{Duration:} 2 hours/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You conjure a sturdy cottage or lodge made of material that is common in the area 
where the spell is cast. The floor is level, clean, and dry. In all respects the 
lodging resembles a normal cottage, with a sturdy door, two shuttered windows, 
and a small fireplace.

The shelter has no heating or cooling source (other than natural insulation qualities). 
Therefore, it must be heated as a normal dwelling, and extreme heat adversely affects 
it and its occupants. The dwelling does, however, provide considerable security 
otherwise---it is as strong as a normal stone building, regardless of its material 
composition. The dwelling resists flames and fire as if it were stone. It is impervious 
to normal missiles (but not the sort cast by siege engines or giants).

The door, shutters, and even chimney are secure against intrusion, the former two 
being \textit{arcane locked }and the latter secured by an iron grate at the top 
and a narrow flue. In addition, these three areas are protected by an \textit{alarm 
}spell. Finally, an \textit{unseen servant }is conjured to provide service to you 
for the duration of the shelter.

The \textit{secure shelter }contains rude furnishings ---eight bunks, a trestle 
table, eight stools, and a writing desk.

\textit{Material Component: }A square chip of stone, crushed lime, a few grains 
of sand, a sprinkling of water, and several splinters of wood. These must be augmented 
by the components of the \textit{unseen servant }spell (string and a bit of wood) 
if this benefit is to be included.

\textit{Focus: }The focus of the \textit{alarm }spell (silver wire and a tiny bell) 
if this benefit is to be included.

\vspace{12pt}
See Invisibility

Divination

\textbf{Level:} Brd 3, Sor/Wiz 2

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 10 min./level (D)

You can see any objects or beings that are invisible within your range of vision, 
as well as any that are ethereal, as if they were normally visible. Such creatures 
are visible to you as translucent shapes, allowing you easily to discern the difference 
between visible, invisible, and ethereal creatures.

The spell does not reveal the method used to obtain invisibility. It does not reveal 
illusions or enable you to see through opaque objects. It does not reveal creatures 
who are simply hiding, concealed, or otherwise hard to see.

\textit{See invisibility }can be made permanent with a \textit{permanency }spell.

\textit{Material Component: }A pinch of talc and a small sprinkling of powdered 
silver.

\vspace{12pt}
Seeming

Illusion (Glamer)

\textbf{Level:} Brd 5, Sor/Wiz 5

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} One creature per two levels, no two of which can be more than 
30 ft. apart

\textbf{Duration:} 12 hours (D)

\textbf{Saving Throw: }Will negates or Will disbelief (if interacted with)

\textbf{Spell Resistance:} Yes or No; see text

This spell functions like \textit{disguise self, }except that you can change the 
appearance of other people as well. Affected creatures resume their normal appearances 
if slain.

Unwilling targets can negate the spell's effect on them by making Will saves or 
with spell resistance.

\vspace{12pt}
Sending

Evocation

\textbf{Level:} Clr 4, Sor/Wiz 5

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 10 minutes

\textbf{Range:} See text

\textbf{Target:} One creature

\textbf{Duration:} 1 round; see text

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You contact a particular creature with which you are familiar and send a short 
message of twenty-five words or less to the subject. The subject recognizes you 
if it knows you. It can answer in like manner immediately. A creature with an Intelligence 
score as low as 1 can understand the \textit{sending, }though the subject's ability 
to react is limited as normal by its Intelligence score. Even if the \textit{sending 
}is received, the subject is not obligated to act upon it in any manner.

If the creature in question is not on the same plane of existence as you are, there 
is a 5\% chance that the \textit{sending }does not arrive. (Local conditions on 
other planes may worsen this chance considerably.)

\textit{Arcane Material Component: }A short piece of fine copper wire.

\vspace{12pt}
Sepia Snake Sigil

Conjuration (Creation) [Force]

\textbf{Level:} Brd 3, Sor/Wiz 3

\textbf{Components:} V, S, M

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Touch

\textbf{Target:} One touched book or written work

\textbf{Duration:} Permanent or until discharged; until released or 1d4 days + 
one day/level; see text

\textbf{Saving Throw:} Reflex negates

\textbf{Spell Resistance:} No

When you cast \textit{sepia snake sigil, }a small symbol appears in the text of 
one written work such as a book, scroll, or map. The text containing the symbol 
must be at least twenty-five words long. When anyone reads the text containing 
the symbol, the \textit{sepia snake }springs into being and strikes the reader, 
provided there is line of effect between the symbol and the reader.

Simply seeing the enspelled text is not sufficient to trigger the spell; the subject 
must deliberately read it. The target is entitled to a save to evade the snake's 
strike. If it succeeds, the \textit{sepia snake }dissipates in a flash of brown 
light accompanied by a puff of dun-colored smoke and a loud noise. If the target 
fails its save, it is engulfed in a shimmering amber field of force and immobilized 
until released, either at your command or when 1d4 days + one day per caster level 
have elapsed.

While trapped in the amber field of force, the subject does not age, breathe, grow 
hungry, sleep, or regain spells. It is preserved in a state of suspended animation, 
unaware of its surroundings. It can be damaged by outside forces (and perhaps even 
killed), since the field provides no protection against physical injury. However, 
a dying subject does not lose hit points or become stable until the spell ends.

The hidden sigil cannot be detected by normal observation, and \textit{detect magic 
}reveals only that the entire text is magical.

A \textit{dispel magic }can remove the sigil. An \textit{erase }spell destroys 
the entire page of text.

\textit{Sepia snake sigil }can be cast in combination with other spells that hide 
or garble text, such as \textit{secret page.}

\textit{Material Component: }500 gp worth of powdered amber, a scale from any snake, 
and a pinch of mushroom spores.

\vspace{12pt}
Sequester

Abjuration

\textbf{Level:} Sor/Wiz 7

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} One willing creature or object (up to a 2-ft. cube/level) touched

\textbf{Duration:} One day/level (D)

\textbf{Saving Throw:} None or Will negates (object)

\textbf{Spell Resistance:} No or Yes (object)

When cast, this spell not only prevents divination spells from working to detect 
or locate the creature or object affected by \textit{sequester, }it also renders 
the affected creature or object invisible to any form of sight or seeing (as the 
\textit{invisibility }spell). The spell does not prevent the subject from being 
discovered through tactile means or through the use of devices. Creatures affected 
by \textit{sequester }become comatose and are effectively in a state of suspended 
animation until the spell wears off or is dispelled.

\textit{Note: }The Will save prevents an attended or magical object from being 
\textit{sequestered. }There is no save to see the \textit{sequestered }creature 
or object or to detect it with a divination spell.

\textit{Material Component: }A basilisk eyelash, gum arabic, and a dram of whitewash.

\vspace{12pt}
Shades

Illusion (Shadow)

\textbf{Level:} Sor/Wiz 9

This spell functions like \textit{shadow conjuration, }except that it mimics sorcerer 
and wizard conjuration spells of 8th level or lower. The illusory conjurations 
created deal four-fifths (80\%) damage to nonbelievers, and nondamaging effects 
are 80\% likely to work against nonbelievers.

\vspace{12pt}
Shadow Conjuration

Illusion (Shadow)

\textbf{Level:} Brd 4, Sor/Wiz 4

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} See text

\textbf{Effect:} See text

\textbf{Duration:} See text

\textbf{Saving Throw: }Will disbelief (if interacted with); varies; see text

\textbf{Spell Resistance:} Yes; see text

You use material from the Plane of Shadow to shape quasi-real illusions of one 
or more creatures, objects, or forces. \textit{Shadow conjuration }can mimic any 
sorcerer or wizard conjuration (summoning) or conjuration (creation) spell of 3rd 
level or lower.

\textit{Shadow conjurations }are actually one-fifth (20\%) as strong as the real 
things, though creatures who believe the \textit{shadow conjurations }to be real 
are affected by them at full strength.

Any creature that interacts with the conjured object, force, or creature can make 
a Will save to recognize its true nature.

Spells that deal damage have normal effects unless the affected creature succeeds 
on a Will save. Each disbelieving creature takes only one-fifth (20\%) damage from 
the attack. If the disbelieved attack has a special effect other than damage, that 
effect is only 20\% likely to occur. Regardless of the result of the save to disbelieve, 
an affected creature is also allowed any save that the spell being simulated allows, 
but the save DC is set according to \textit{shadow conjuration'}s level (4th) rather 
than the spell's normal level. In addition, any effect created by \textit{shadow 
conjuration }allows spell resistance, even if the spell it is simulating does not. 
Shadow objects or substances have normal effects except against those who disbelieve 
them.

Against disbelievers, they are 20\% likely to work.

A shadow creature has one-fifth the hit points of a normal creature of its kind 
(regardless of whether it's recognized as shadowy). It deals normal damage and 
has all normal abilities and weaknesses. Against a creature that recognizes it 
as a shadow creature, however, the shadow creature's damage is one-fifth (20\%) 
normal, and all special abilities that do not deal lethal damage are only 20\% 
likely to work. (Roll for each use and each affected character separately.) Furthermore, 
the shadow creature's AC bonuses are one-fifth as large.

A creature that succeeds on its save sees the \textit{shadow conjurations }as transparent 
images superimposed on vague, shadowy forms.

Objects automatically succeed on their Will saves against this spell.

\vspace{12pt}
Shadow Conjuration, Greater

Illusion (Shadow)

\textbf{Level:} Sor/Wiz 7

This spell functions like \textit{shadow conjuration, }except that it can duplicate 
any sorcerer or wizard conjuration (summoning) or conjuration (creation) spell 
of 6th level or lower. The illusory conjurations created deal three-fifths (60\%) 
damage to nonbelievers, and nondamaging effects are 60\% likely to work against 
nonbelievers.

\vspace{12pt}
Shadow Evocation

Illusion (Shadow)

\textbf{Level:} Brd 5, Sor/Wiz 5

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} See text

\textbf{Effect:} See text

\textbf{Duration:} See text

\textbf{Saving Throw: }Will disbelief (if interacted with)

\textbf{Spell Resistance:} Yes

You tap energy from the Plane of Shadow to cast a quasi-real, illusory version 
of a sorcerer or wizard evocation spell of 4th level or lower. (For a spell with 
more than one level, use the best one applicable to you.)

Spells that deal damage have normal effects unless an affected creature succeeds 
on a Will save. Each disbelieving creature takes only one-fifth damage from the 
attack. If the disbelieved attack has a special effect other than damage, that 
effect is one-fifth as strong (if applicable) or only 20\% likely to occur. If 
recognized as a \textit{shadow evocation, }a damaging spell deals only one-fifth 
(20\%) damage. Regardless of the result of the save to disbelieve, an affected 
creature is also allowed any save (or spell resistance) that the spell being simulated 
allows, but the save DC is set according to \textit{shadow evocation'}s level (5th) 
rather than the spell's normal level.

Nondamaging effects have normal effects except against those who disbelieve them. 
Against disbelievers, they have no effect.

Objects automatically succeed on their Will saves against this spell.

\vspace{12pt}
Shadow Evocation, Greater

Illusion (Shadow)

\textbf{Level:} Sor/Wiz 8

This spell functions like \textit{shadow evocation, }except that it enables you 
to create partially real, illusory versions of sorcerer or wizard evocation spells 
of 7th level or lower. If recognized as a \textit{greater shadow evocation}, a 
damaging spell deals only three-fifths (60\%) damage.

\vspace{12pt}
Shadow Walk

Illusion (Shadow)

\textbf{Level:} Brd 5, Sor/Wiz 6

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Targets:} Up to one touched creature/ level

\textbf{Duration:} 1 hour/level (D)

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

To use the \textit{shadow walk }spell, you must be in an area of shadowy illumination. 
You and any creature you touch are then transported along a coiling path of shadowstuff 
to the edge of the Material Plane where it borders the Plane of Shadow. The effect 
is largely illusory, but the path is quasi-real. You can take more than one creature 
along with you (subject to your level limit), but all must be touching each other.

In the region of shadow, you move at a rate of 50 miles per hour, moving normally 
on the borders of the Plane of Shadow but much more rapidly relative to the Material 
Plane. Thus, you can use this spell to travel rapidly by stepping onto the Plane 
of Shadow, moving the desired distance, and then stepping back onto the Material 
Plane.

Because of the blurring of reality between the Plane of Shadow and the Material 
Plane, you can't make out details of the terrain or areas you pass over during 
transit, nor can you predict perfectly where your travel will end. It's impossible 
to judge distances accurately, making the spell virtually useless for scouting 
or spying. Furthermore, when the spell effect ends, you are shunted 1d10x100 feet 
in a random horizontal direction from your desired endpoint. If this would place 
you within a solid object, you are shunted 1d10x1,000 feet in the same direction. 
If this would still place you within a solid object, you (and any creatures with 
you) are shunted to the nearest empty space available, but the strain of this activity 
renders each creature fatigued (no save).

\textit{Shadow walk }can also be used to travel to other planes that border on 
the Plane of Shadow, but this usage requires the transit of the Plane of Shadow 
to arrive at a border with another plane of reality. The transit of the Plane of 
Shadow requires 1d4 hours.

Any creatures touched by you when \textit{shadow walk }is cast also make the transition 
to the borders of the Plane of Shadow.

They may opt to follow you, wander off through the plane, or stumble back into 
the Material Plane (50\% chance for either of the latter results if they are lost 
or abandoned by you). Creatures unwilling to accompany you into the Plane of Shadow 
receive a Will saving throw, negating the effect if successful.

\vspace{12pt}
Shambler

Conjuration (Creation)

\textbf{Level:} Drd 9, Plant 9

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Effect:} Three or more shambling mounds, no two of which can be more than 
30 ft. apart; see text

\textbf{Duration:} Seven days or seven months (D); see text

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

The \textit{shambler }spell creates 1d4+2 shambling mounds with 11 HD each. The 
creatures willingly aid you in combat or battle, perform a specific mission, or 
serve as bodyguards. The creatures remain with you for seven days unless you dismiss 
them. If the shamblers are created only for guard duty, however, the duration of 
the spell is seven months. In this case, the shamblers can only be ordered to guard 
a specific site or location. Shamblers summoned to guard duty cannot move outside 
the spell's range, which is measured from the point where each first appeared. 

The shamblers have resistance to fire as normal shambling mounds do only if the 
terrain is rainy, marshy, or damp.

\vspace{12pt}
Shapechange

Transmutation

\textbf{Level:} Animal 9, Drd 9, Sor/Wiz 9

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 10 min./level (D)

This spell functions like \textit{polymorph, }except that it enables you to assume 
the form of any single nonunique creature (of any type) from Fine to Colossal size. 
The assumed form cannot have more than your caster level in Hit Dice (to a maximum 
of 25 HD). Unlike \textit{polymorph, }this spell allows incorporeal or gaseous 
forms to be assumed.

You gain all extraordinary and supernatural abilities (both attacks and qualities) 
of the assumed form, but you lose your own supernatural abilities. You also gain 
the type of the new form in place of your own. The new form does not disorient 
you. Parts of your body or pieces of equipment that are separated from you do not 
revert to their original forms.

You can become just about anything you are familiar with. You can change form once 
each round as a free action. The change takes place either immediately before your 
regular action or immediately after it, but not during the action. If you use this 
spell to create a disguise, you get a +10 bonus on your Disguise check.

\textit{Focus: }A jade circlet worth no less than 1,500 gp, which you must place 
on your head when casting the spell. (The focus melds into your new form when you 
change shape.)

\vspace{12pt}
Shatter

Evocation [Sonic]

\textbf{Level:} Brd 2, Chaos 2, Clr 2, Destruction 2, Sor/Wiz 2

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Area or Target:} 5-ft.-radius spread; or one solid object or one crystalline 
creature

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates (object); Will negates (object) or Fortitude 
half; see text

\textbf{Spell Resistance:} Yes (object)

\textit{Shatter }creates a loud, ringing noise that breaks brittle, nonmagical 
objects; sunders a single solid, nonmagical object; or damages a crystalline creature.

Used as an area attack, \textit{shatter }destroys nonmagical objects of crystal, 
glass, ceramic, or porcelain. All such objects within a 5-foot radius of the point 
of origin are smashed into dozens of pieces by the spell. Objects weighing more 
than 1 pound per your level are not affected, but all other objects of the appropriate 
composition are shattered.

Alternatively, you can target \textit{shatter }against a single solid object, regardless 
of composition, weighing up to 10 pounds per caster level. Targeted against a crystalline 
creature (of any weight), \textit{shatter }deals 1d6 points of sonic damage per 
caster level (maximum 10d6), with a Fortitude save for half damage.

\textit{Arcane Material Component: }A chip of mica.

\vspace{12pt}
Shield

Abjuration [Force]

\textbf{Level:} Sor/Wiz 1

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 1 min./level (D)

\textit{Shield }creates an invisible, tower shield-sized mobile disk of force that 
hovers in front of you. It negates \textit{magic missile }attacks directed at you. 
The disk also provides a +4 shield bonus to AC. This bonus applies against incorporeal 
touch attacks, since it is a force effect. The \textit{shield }has no armor check 
penalty or arcane spell failure chance. Unlike with a normal tower shield, you 
can't use the \textit{shield }spell for cover.

\vspace{12pt}
Shield of Faith

Abjuration

\textbf{Level:} Clr 1

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 1 min./level

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

This spell creates a shimmering, magical field around the touched creature that 
averts attacks. The spell grants the subject a +2 deflection bonus to AC, with 
an additional +1 to the bonus for every six levels you have (maximum +5 deflection 
bonus at 18th level).

\textit{Material Component: }A small parchment with a bit of holy text written 
upon it.

\vspace{12pt}
Shield of Law

Abjuration [Lawful]

\textbf{Level:} Clr 8, Law 8

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 20 ft.

\textbf{Targets:} One creature/level in a 20-ft.-radius burst centered on you

\textbf{Duration:} 1 round/level (D)

\textbf{Saving Throw:} See text

\textbf{Spell Resistance:} Yes (harmless)

A dim, blue glow surrounds the subjects, protecting them from attacks, granting 
them resistance to spells cast by chaotic creatures, and \textit{slowing }chaotic 
creatures when they strike the subjects. This abjuration has four effects.

First, each warded creature gains a +4 deflection bonus to AC and a +4 resistance 
bonus on saves. Unlike \textit{protection from chaos}, this benefit applies against 
all attacks, not just against attacks by chaotic creatures.

Second, a warded creature gains spell resistance 25 against chaotic spells and 
spells cast by chaotic creatures.

Third, the abjuration blocks possession and mental influence, just as \textit{protection 
from chaos }does.

Finally, if a chaotic creature succeeds on a melee attack against a warded creature, 
the attacker is \textit{slowed }(Will save negates, as the \textit{slow }spell, 
but against \textit{shield of law'}s save DC).

\textit{Focus: }A tiny reliquary containing some sacred relic, such as a scrap 
of parchment from a lawful text. The reliquary costs at least 500 gp.

\vspace{12pt}
Shield Other

Abjuration

\textbf{Level:} Clr 2, Pal 2, Protection 2

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One creature

\textbf{Duration:} 1 hour/level (D)

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

This spell wards the subject and creates a mystic connection between you and the 
subject so that some of its wounds are transferred to you. The subject gains a 
+1 deflection bonus to AC and a +1 resistance bonus on saves. Additionally, the 
subject takes only half damage from all wounds and attacks (including that dealt 
by special abilities) that deal hit point damage. The amount of damage not taken 
by the warded creature is taken by you. Forms of harm that do not involve hit points, 
such as \textit{charm }effects, temporary ability damage, level draining, and death 
effects, are not affected. If the subject suffers a reduction of hit points from 
a lowered Constitution score, the reduction is not split with you because it is 
not hit point damage. When the spell ends, subsequent damage is no longer divided 
between the subject and you, but damage already split is not reassigned to the 
subject.

If you and the subject of the spell move out of range of each other, the spell 
ends.

\textit{Focus: }A pair of platinum rings (worth at least 50 gp each) worn by both 
you and the warded creature.

\vspace{12pt}
Shillelagh

Transmutation

\textbf{Level:} Drd 1

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} One touched nonmagical oak club or quarterstaff

\textbf{Duration:} 1 min./level

\textbf{Saving Throw: }Will negates (object)

\textbf{Spell Resistance:} Yes (object)

Your own nonmagical club or quarterstaff becomes a weapon with a +1 enhancement 
bonus on attack and damage rolls. (A quarterstaff gains this enhancement for both 
ends of the weapon.) It deals damage as if it were two size categories larger (a 
Small club or quarterstaff so transmuted deals 1d8 points of damage, a Medium 2d6, 
and a Large 3d6), +1 for its enhancement bonus. These effects only occur when the 
weapon is wielded by you. If you do not wield it, the weapon behaves as if unaffected 
by this spell.

\vspace{12pt}
Shocking Grasp

Evocation [Electricity]

\textbf{Level:} Sor/Wiz 1

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature or object touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

Your successful melee touch attack deals 1d6 points of electricity damage per caster 
level (maximum 5d6). When delivering the jolt, you gain a +3 bonus on attack rolls 
if the opponent is wearing metal armor (or made out of metal, carrying a lot of 
metal, or the like).

\vspace{12pt}
Shout

Evocation [Sonic]

\textbf{Level:} Brd 4, Sor/Wiz 4

\textbf{Components:} V

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 30 ft.

\textbf{Area:} Cone-shaped burst

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} Fortitude partial or Reflex negates (object); see text

\textbf{Spell Resistance:} Yes (object)

You emit an ear-splitting yell that deafens and damages creatures in its path. 
Any creature within the area is deafened for 2d6 rounds and takes 5d6 points of 
sonic damage. A successful save negates the deafness and reduces the damage by 
half. Any exposed brittle or crystalline object or crystalline creature takes 1d6 
points of sonic damage per caster level (maximum 15d6). An affected creature is 
allowed a Fortitude save to reduce the damage by half, and a creature holding fragile 
objects can negate damage to them with a successful Reflex save.

A \textit{shout }spell cannot penetrate a \textit{silence }spell.

\vspace{12pt}
Shout, Greater

Evocation [Sonic]

\textbf{Level:} Brd 6, Sor/Wiz 8

\textbf{Components:} V, S, F

\textbf{Range:} 60 ft.

\textbf{Saving Throw:} Fortitude partial or Reflex negates (object); see text

This spell functions like \textit{shout, }except that the cone deals 10d6 points 
of sonic damage (or 1d6 points of sonic damage per caster level, maximum 20d6, 
against exposed brittle or crystalline objects or crystalline creatures). It also 
causes creatures to be stunned for 1 round and deafened for 4d6 rounds. A creature 
in the area of the cone can negate the stunning and halve both the damage and the 
duration of the deafness with a successful Fortitude save. A creature holding vulnerable 
objects can attempt a Reflex save to negate the damage to those objects.

\textit{Arcane Focus: }A small metal or ivory horn.

\vspace{12pt}
Shrink Item

Transmutation

\textbf{Level:} Sor/Wiz 3

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} One touched object of up to 2 cu. ft./level

\textbf{Duration:} One day/level; see text

\textbf{Saving Throw: }Will negates (object)

\textbf{Spell Resistance:} Yes (object)

You are able to shrink one nonmagical item (if it is within the size limit) to 
1/16 of its normal size in each dimension (to about 1/4,000 the original volume 
and mass). This change effectively reduces the object's size by four categories. 
Optionally, you can also change its now shrunken composition to a clothlike one. 
Objects changed by a \textit{shrink item }spell can be returned to normal composition 
and size merely by tossing them onto any solid surface or by a word of command 
from the original caster. Even a burning fire and its fuel can be shrunk by this 
spell. Restoring the shrunken object to its normal size and composition ends the 
spell.

\textit{Shrink item }can be made permanent with a \textit{permanency }spell, in 
which case the affected object can be shrunk and expanded an indefinite number 
of times, but only by the original caster.

\vspace{12pt}
Silence

Illusion (Glamer)

\textbf{Level:} Brd 2, Clr 2

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Area:} 20-ft.-radius emanation centered on a creature, object, or point 
in space

\textbf{Duration:} 1 min./level (D)

\textbf{Saving Throw: }Will negates; see text or none (object)

\textbf{Spell Resistance:} Yes; see text or no (object)

Upon the casting of this spell, complete silence prevails in the affected area. 
All sound is stopped: Conversation is impossible, spells with verbal components 
cannot be cast, and no noise whatsoever issues from, enters, or passes through 
the area. The spell can be cast on a point in space, but the effect is stationary 
unless cast on a mobile object. The spell can be centered on a creature, and the 
effect then radiates from the creature and moves as it moves. An unwilling creature 
can attempt a Will save to negate the spell and can use spell resistance, if any. 
Items in a creature's possession or magic items that emit sound receive the benefits 
of saves and spell resistance, but unattended objects and points in space do not. 
This spell provides a defense against sonic or language-based attacks.

\vspace{12pt}
Silent Image

Illusion (Figment)

\textbf{Level:} Brd 1, Sor/Wiz 1

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Effect:} Visual figment that cannot extend beyond four 10-ft. cubes + one 
10-ft. cube/level (S)

\textbf{Duration:} Concentration

\textbf{Saving Throw: }Will disbelief (if interacted with)

\textbf{Spell Resistance:} No

This spell creates the visual illusion of an object, creature, or force, as visualized 
by you. The illusion does not create sound, smell, texture, or temperature. You 
can move the image within the limits of the size of the effect.

\textit{Focus: }A bit of fleece.

\vspace{12pt}
Simulacrum

Illusion (Shadow)

\textbf{Level:} Sor/Wiz 7

\textbf{Components:} V, S, M, XP

\textbf{Casting Time:} 12 hours

\textbf{Range:} 0 ft.

\textbf{Effect:} One duplicate creature

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

\textit{Simulacrum }creates an illusory duplicate of any creature. The duplicate 
creature is partially real and formed from ice or snow. It appears to be the same 
as the original, but it has only one-half of the real creature's levels or Hit 
Dice (and the appropriate hit points, feats, skill ranks, and special abilities 
for a creature of that level or HD). You can't create a simulacrum of a creature 
whose Hit Dice or levels exceed twice your caster level. You must make a Disguise 
check when you cast the spell to determine how good the likeness is. A creature 
familiar with the original might detect the ruse with a successful Spot check (opposed 
by the caster's Disguise check) or a DC 20 Sense Motive check.

At all times the simulacrum remains under your absolute command. No special telepathic 
link exists, so command must be exercised in some other manner. A simulacrum has 
no ability to become more powerful. It cannot increase its level or abilities. 
If reduced to 0 hit points or otherwise destroyed, it reverts to snow and melts 
instantly into nothingness. A complex process requiring at least 24 hours, 100 
gp per hit point, and a fully equipped magical laboratory can repair damage to 
a simulacrum.

\textit{Material Component: }The spell is cast over the rough snow or ice form, 
and some piece of the creature to be duplicated (hair, nail, or the like) must 
be placed inside the snow or ice. Additionally, the spell requires powdered ruby 
worth 100 gp per HD of the simulacrum to be created.

\textit{XP Cost: }100 XP per HD of the simulacrum to be created (minimum 1,000 
XP).

\vspace{12pt}
Slay Living

Necromancy [Death]

\textbf{Level:} Clr 5, Death 5

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Living creature touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} Fortitude partial

\textbf{Spell Resistance:} Yes

You can slay any one living creature. You must succeed on a melee touch attack 
to touch the subject, and it can avoid death with a successful Fortitude save. 
If it succeeds, it instead takes 3d6 points of damage +1 point per caster level.

\vspace{12pt}
Sleep

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Brd 1, Sor/Wiz 1

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 round

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Area:} One or more living creatures within a 10-ft.-radius burst

\textbf{Duration:} 1 min./level

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

A \textit{sleep }spell causes a magical slumber to come upon 4 Hit Dice of creatures. 
Creatures with the fewest HD are affected first.

Among creatures with equal HD, those who are closest to the spell's point of origin 
are affected first. Hit Dice that are not sufficient to affect a creature are wasted.

Sleeping creatures are helpless. Slapping or wounding awakens an affected creature, 
but normal noise does not. Awakening a creature is a standard action (an application 
of the aid another action).

\textit{Sleep }does not target unconscious creatures, constructs, or undead creatures.

\textit{Material Component: }A pinch of fine sand, rose petals, or a live cricket.

\vspace{12pt}
Sleet Storm

Conjuration (Creation) [Cold]

\textbf{Level:} Drd 3, Sor/Wiz 3

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Area:} Cylinder (40-ft. radius, 20 ft. high)

\textbf{Duration:} 1 round/level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

Driving sleet blocks all sight (even darkvision) within it and causes the ground 
in the area to be icy. A creature can walk within or through the area of sleet 
at half normal speed with a DC 10 Balance check. Failure means it can't move in 
that round, while failure by 5 or more means it falls (see the Balance skill for 
details).

The sleet extinguishes torches and small fires.

\textit{Arcane Material Component: }A pinch of dust and a few drops of water.

\vspace{12pt}
Slow

Transmutation

\textbf{Level:} Brd 3, Sor/Wiz 3

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} One creature/level, no two of which can be more than 30 ft. apart

\textbf{Duration:} 1 round/level

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

An affected creature moves and attacks at a drastically slowed rate. A \textit{slowed 
}creature can take only a single move action or standard action each turn, but 
not both (nor may it take full-round actions). Additionally, it takes a -1 penalty 
on attack rolls, AC, and Reflex saves. A \textit{slowed }creature moves at half 
its normal speed (round down to the next 5-foot increment), which affects the creature's 
jumping distance as normal for decreased speed.

Multiple \textit{slow }effects don't stack. \textit{Slow }counters and dispels 
\textit{haste}.

\textit{Material Component: }A drop of molasses.

\vspace{12pt}
Snare

Transmutation

\textbf{Level:} Rgr 2, Drd 3

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 3 rounds

\textbf{Range:} Touch

\textbf{Target:} Touched nonmagical circle of vine, rope, or thong with a 2 ft. 
diameter + 2 ft./level

\textbf{Duration:} Until triggered or broken

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell enables you to make a snare that functions as a magic trap. The snare 
can be made from any supple vine, a thong, or a rope. When you cast \textit{snare 
}upon it, the cordlike object blends with its surroundings (Search DC 23 for a 
character with the trapfinding ability to locate). One end of the snare is tied 
in a loop that contracts around one or more of the limbs of any creature stepping 
inside the circle.

If a strong and supple tree is nearby, the snare can be fastened to it. The spell 
causes the tree to bend and then straighten when the loop is triggered, dealing 
1d6 points of damage to the creature trapped and lifting it off the ground by the 
trapped limb or limbs. If no such tree is available, the cordlike object tightens 
around the creature, dealing no damage but causing it to be entangled.

The snare is magical. To escape, a trapped creature must make a DC 23 Escape Artist 
check or a DC 23 Strength check that is a full-round action. The snare has AC 7 
and 5 hit points. A successful escape from the snare breaks the loop and ends the 
spell.

\vspace{12pt}
Soften Earth and Stone

Transmutation [Earth]

\textbf{Level:} Drd 2, Earth 2

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Area:} 10-ft. square/level; see text

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

When this spell is cast, all natural, undressed earth or stone in the spell's area 
is softened. Wet earth becomes thick mud, dry earth becomes loose sand or dirt, 
and stone becomes soft clay that is easily molded or chopped. You affect a 10-footsquare 
area to a depth of 1 to 4 feet, depending on the toughness or resilience of the 
ground at that spot. Magical, enchanted, dressed, or worked stone cannot be affected. 
Earth or stone creatures are not affected.

A creature in mud must succeed on a Reflex save or be caught for 1d2 rounds and 
unable to move, attack, or cast spells. A creature that succeeds on its save can 
move through the mud at half speed, and it can't run or charge.

Loose dirt is not as troublesome as mud, but all creatures in the area can move 
at only half their normal speed and can't run or charge over the surface.

Stone softened into clay does not hinder movement, but it does allow characters 
to cut, shape, or excavate areas they may not have been able to affect before.

While \textit{soften earth and stone }does not affect dressed or worked stone, 
cavern ceilings or vertical surfaces such as cliff faces can be affected. Usually, 
this causes a moderate collapse or landslide as the loosened material peels away 
from the face of the wall or roof and falls.

A moderate amount of structural damage can be dealt to a manufactured structure 
by softening the ground beneath it, causing it to settle. However, most well-built 
structures will only be damaged by this spell, not destroyed.

\vspace{12pt}
Solid Fog

Conjuration (Creation)

\textbf{Level:} Sor/Wiz 4

\textbf{Components:} V, S, M

\textbf{Duration:} 1 min./level

\textbf{Spell Resistance:} No

This spell functions like \textit{fog cloud}, but in addition to obscuring sight, 
the \textit{solid fog }is so thick that any creature attempting to move through 
it progresses at a speed of 5 feet, regardless of its normal speed, and it takes 
a -2 penalty on all melee attack and melee damage rolls. The vapors prevent effective 
ranged weapon attacks (except for magic rays and the like). A creature or object 
that falls into \textit{solid fog }is slowed, so that each 10 feet of vapor that 
it passes through reduces falling damage by 1d6. A creature can't take a 5-foot 
step while in \textit{solid fog.}

However, unlike normal fog, only a severe wind (31+ mph) disperses these vapors, 
and it does so in 1 round.

\textit{Solid fog }can be made permanent with a \textit{permanency }spell. A permanent 
\textit{solid fog }dispersed by wind reforms in 10 minutes.

\textit{Material Component: }A pinch of dried, powdered peas combined with powdered 
animal hoof.

\vspace{12pt}
Song of Discord

Enchantment (Compulsion) [Mind-Affecting, Sonic]

\textbf{Level:} Brd 5

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Area:} Creatures within a 20-ft.-radius spread

\textbf{Duration:} 1 round/level

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

This spell causes those within the area to turn on each other rather than attack 
their foes. Each affected creature has a 50\% chance to attack the nearest target 
each round. (Roll to determine each creature's behavior every round at the beginning 
of its turn.) A creature that does not attack its nearest neighbor is free to act 
normally for that round.

Creatures forced by a \textit{song of discord }to attack their fellows employ all 
methods at their disposal, choosing their deadliest spells and most advantageous 
combat tactics. They do not, however, harm targets that have fallen unconscious.

\vspace{12pt}
Soul Bind

Necromancy

\textbf{Level:} Clr 9, Sor/Wiz 9

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} Corpse

\textbf{Duration:} Permanent

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} No

You draw the soul from a newly dead body and imprison it in a black sapphire gem. 
The subject must have been dead no more than 1 round per caster level. The soul, 
once trapped in the gem, cannot be returned through \textit{clone, raise dead, 
reincarnation, resurrection, true resurrection, }or even a \textit{miracle }or 
a \textit{wish. }Only by destroying the gem or dispelling the spell on the gem 
can one free the soul (which is then still dead).

\textit{Focus: }A black sapphire of at least 1,000 gp value for every Hit Die possessed 
by the creature whose soul is to be bound. If the gem is not valuable enough, it 
shatters when the binding is attempted. (While creatures have no concept of level 
or Hit Dice as such, the value of the gem needed to trap an individual can be researched. 
Remember that this value can change over time as creatures gain more Hit Dice.)

\vspace{12pt}
Sound Burst

Evocation [Sonic]

\textbf{Level:} Brd 2, Clr 2

\textbf{Components:} V, S, F/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Area:} 10-ft.-radius spread

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} Fortitude partial

\textbf{Spell Resistance:} Yes

You blast an area with a tremendous cacophony. Every creature in the area takes 
1d8 points of sonic damage and must succeed on a Fortitude save to avoid being 
stunned for 1 round. 

Creatures that cannot hear are not stunned but are still damaged.

\textit{Arcane Focus: }A musical instrument.

\vspace{12pt}
Speak with Animals

Divination

\textbf{Level:} Brd 3, Drd 1, Rgr 1

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 1 min./level

You can comprehend and communicate with animals. You are able to ask questions 
of and receive answers from animals, although the spell doesn't make them any more 
friendly or cooperative than normal. Furthermore, wary and cunning animals are 
likely to be terse and evasive, while the more stupid ones make inane comments. 
If an animal is friendly toward you, it may do some favor or service for you.

\vspace{12pt}
Speak with Dead

Necromancy [Language-Dependent]

\textbf{Level:} Clr 3

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 10 minutes

\textbf{Range:} 10 ft.

\textbf{Target:} One dead creature

\textbf{Duration:} 1 min./level

\textbf{Saving Throw: }Will negates; see text

\textbf{Spell Resistance:} No

You grant the semblance of life and intellect to a corpse, allowing it to answer 
several questions that you put to it. You may ask one question per two caster levels. 
Unasked questions are wasted if the duration expires. The corpse's knowledge is 
limited to what the creature knew during life, including the languages it spoke 
(if any). Answers are usually brief, cryptic, or repetitive. If the creature's 
alignment was different from yours, the corpse gets a Will save to resist the spell 
as if it were alive.

If the corpse has been subject to \textit{speak with dead }within the past week, 
the new spell fails. You can cast this spell on a corpse that has been deceased 
for any amount of time, but the body must be mostly intact to be able to respond. 
A damaged corpse may be able to give partial answers or partially correct answers, 
but it must at least have a mouth in order to speak at all. 

This spell does not let you actually speak to the person (whose soul has departed). 
It instead draws on the imprinted knowledge stored in the corpse. The partially 
animated body retains the imprint of the soul that once inhabited it, and thus 
it can speak with all the knowledge that the creature had while alive. The corpse, 
however, cannot learn new information.

Indeed, it can't even remember being questioned.

This spell does not affect a corpse that has been turned into an undead creature.

\vspace{12pt}
Speak with Plants

Divination

\textbf{Level:} Brd 4, Drd 3, Rgr 2

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 1 min./level

You can comprehend and communicate with plants, including both normal plants and 
plant creatures. You are able to ask questions of and receive answers from plants. 
A regular plant's sense of its surroundings is limited, so it won't be able to 
give (or recognize) detailed descriptions of creatures or answer questions about 
events outside its immediate vicinity.

The spell doesn't make plant creatures any more friendly or cooperative than normal. 
Furthermore, wary and cunning plant creatures are likely to be terse and evasive, 
while the more stupid ones may make inane comments. If a plant creature is friendly 
toward you, it may do some favor or service for you.

\vspace{12pt}
Spectral Hand

Necromancy

\textbf{Level:} Sor/Wiz 2

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Effect:} One spectral hand

\textbf{Duration:} 1 min./level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

A ghostly, glowing hand shaped from your life force materializes and moves as you 
desire, allowing you to deliver low-level, touch range spells at a distance. On 
casting the spell, you lose 1d4 hit points that return when the spell ends (even 
if it is dispelled), but not if the hand is destroyed. (The hit points can be healed 
as normal.) For as long as the spell lasts, any touch range spell of 4th level 
or lower that you cast can be delivered by the \textit{spectral hand}. The spell 
gives you a +2 bonus on your melee touch attack roll, and attacking with the hand 
counts normally as an attack. The hand always strikes from your direction. The 
hand cannot flank targets like a creature can. After it delivers a spell, or if 
the hand goes beyond the spell range, goes out of your sight, the hand returns 
to you and hovers.

The hand is incorporeal and thus cannot be harmed by normal weapons. It has improved 
evasion (half damage on a failed Reflex save and no damage on a successful save), 
your save bonuses, and an AC of at least 22. Your Intelligence modifier applies 
to the hand's AC as if it were the hand's Dexterity modifier. The hand has 1 to 
4 hit points, the same number that you lost in creating it.

\vspace{12pt}
Spell Immunity

Abjuration

\textbf{Level:} Clr 4, Protection 4, Strength 4

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 10 min./level

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

The warded creature is immune to the effects of one specified spell for every four 
levels you have. The spells must be of 4th level or lower. The warded creature 
effectively has unbeatable spell resistance regarding the specified spell or spells. 
Naturally, that immunity doesn't protect a creature from spells for which spell 
resistance doesn't apply. \textit{Spell immunity }protects against spells, spell-like 
effects of magic items, and innate spell-like abilities of creatures. It does not 
protect against supernatural or extraordinary abilities, such as breath weapons 
or gaze attacks.

Only a particular spell can be protected against, not a certain domain or school 
of spells or a group of spells that are similar in effect. 

A creature can have only one \textit{spell immunity }or \textit{greater spell immunity 
}spell in effect on it at a time.

\vspace{12pt}
Spell Immunity, Greater

Abjuration

\textbf{Level:} Clr 8

This spell functions like \textit{spell immunity}, except the immunity applies 
to spells of 8th level or lower.

A creature can have only one \textit{spell immunity }or \textit{greater spell immunity 
}spell in effect on it at a time.

\vspace{12pt}
Spell Resistance

Abjuration

\textbf{Level:} Clr 5, Magic 5, Protection 5

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 1 min./level

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

The creature gains spell resistance equal to 12 + your caster level.

\vspace{12pt}
Spellstaff

Transmutation

\textbf{Level:} Drd 6

\textbf{Components:} V, S, F

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Touch

\textbf{Target: }Wooden quarterstaff touched

\textbf{Duration:} Permanent until discharged (D)

\textbf{Saving Throw: }Will negates (object)

\textbf{Spell Resistance:} Yes (object)

You store one spell that you can normally cast in a wooden quarterstaff. Only one 
such spell can be stored in a staff at a given time, and you cannot have more than 
one \textit{spellstaff }at any given time. You can cast a spell stored within a 
staff just as though it were among those you had prepared, but it does not count 
against your normal allotment for a given day. You use up any applicable material 
components required to cast the spell when you store it in the \textit{spellstaff.}

\textit{Focus: }The staff that stores the spell.

\vspace{12pt}
Spell Turning

Abjuration

\textbf{Level:} Luck 7, Magic 7, Sor/Wiz 7

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} Until expended or 10 min./level

Spells and spell-like effects targeted on you are turned back upon the original 
caster. The abjuration turns only spells that have you as a target. Effect and 
area spells are not affected. \textit{Spell turning }also fails to stop touch range 
spells. 

From seven to ten (1d4+6) spell levels are affected by the turning. The exact number 
is rolled secretly.

When you are targeted by a spell of higher level than the amount of spell turning 
you have left, that spell is partially turned. The subtract the amount of spell 
turning left from the spell level of the incoming spell, then divide the result 
by the spell level of the incoming spell to see what fraction of the effect gets 
through. For damaging spells, you and the caster each take a fraction of the damage. 
For nondamaging spells, each of you has a proportional chance to be affected.

If you and a spellcasting attacker are both warded by \textit{spell turning }effects 
in operation, a resonating field is created.

Roll randomly to determine the result.

\begin{tabular}{|>{\raggedright}p{28pt}|>{\raggedright}p{257pt}|}
\hline
d\textbf{\%} & E\textbf{ffect}\tabularnewline
\hline
01-70 & Spell drains away without effect.\tabularnewline
\hline
71-80 & Spell affects both of you equally at full effect.\tabularnewline
\hline
81-97 & Both turning effects are rendered nonfunctional for 1d4 minutes.\tabularnewline
\hline
98-100 & Both of you go through a rift into another plane.\tabularnewline
\hline
\end{tabular}

\textit{Arcane Material Component: }A small silver mirror.

\vspace{12pt}
Spider Climb

Transmutation

\textbf{Level:} Drd 2, Sor/Wiz 2

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 10 min./level

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

The subject can climb and travel on vertical surfaces or even traverse ceilings 
as well as a spider does. The affected creature must have its hands free to climb 
in this manner. The subject gains a climb speed of 20 feet; furthermore, it need 
not make Climb checks to traverse a vertical or horizontal surface (even upside 
down). A \textit{spider climbing }creature retains its Dexterity bonus to Armor 
Class (if any) while climbing, and opponents get no special bonus to their attacks 
against it. It cannot, however, use the run action while climbing.

\textit{Material Component: }A drop of bitumen and a live spider, both of which 
must be eaten by the subject.

\vspace{12pt}
Spike Growth

Transmutation

\textbf{Level:} Drd 3, Rgr 2

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Area:} One 20-ft. square/level

\textbf{Duration:} 1 hour/level (D)

\textbf{Saving Throw:} Reflex partial

\textbf{Spell Resistance:} Yes

Any ground-covering vegetation in the spell's area becomes very hard and sharply 
pointed without changing its appearance.

In areas of bare earth, roots and rootlets act in the same way. Typically, \textit{spike 
growth }can be cast in any outdoor setting except open water, ice, heavy snow, 
sandy desert, or bare stone. Any creature moving on foot into or through the spell's 
area takes 1d4 points of piercing damage for each 5 feet of movement through the 
spiked area.

Any creature that takes damage from this spell must also succeed on a Reflex save 
or suffer injuries to its feet and legs that slow its land speed by one-half. This 
speed penalty lasts for 24 hours or until the injured creature receives a \textit{cure 
}spell (which also restores lost hit points). Another character can remove the 
penalty by taking 10 minutes to dress the injuries and succeeding on a Heal check 
against the spell's save DC.

\textit{Spike growth }can't be disabled with the Disable Device skill.

\textit{Note: }Magic traps such as \textit{spike growth }are hard to detect. A 
rogue (only) can use the Search skill to find a \textit{spike growth}. The DC is 
25 + spell level, or DC 28 for \textit{spike growth }(or DC 27 for \textit{spike 
growth }cast by a ranger).

\vspace{12pt}
Spike Stones

Transmutation [Earth]

\textbf{Level:} Drd 4, Earth 4

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Area:} One 20-ft. square/level

\textbf{Duration:} 1 hour/level (D)

\textbf{Saving Throw:} Reflex partial

\textbf{Spell Resistance:} Yes

Rocky ground, stone floors, and similar surfaces shape themselves into long, sharp 
points that blend into the background.

\textit{Spike stones }impede progress through an area and deal damage. Any creature 
moving on foot into or through the spell's area moves at half speed.

In addition, each creature moving through the area takes 1d8 points of piercing 
damage for each 5 feet of movement through the spiked area.

Any creature that takes damage from this spell must also succeed on a Reflex save 
to avoid injuries to its feet and legs. A failed save causes the creature's speed 
to be reduced to half normal for 24 hours or until the injured creature receives 
a \textit{cure }spell (which also restores lost hit points). Another character 
can remove the penalty by taking 10 minutes to dress the injuries and succeeding 
on a Heal check against the spell's save DC.

\textit{Spike stones }is a magic trap that can't be disabled with the Disable Device 
skill.

\textit{Note: }Magic traps such as \textit{spike stones }are hard to detect. A 
rogue (only) can use the Search skill to find \textit{spike stones}. The DC is 
25 + spell level, or DC 29 for \textit{spike stones}.

\vspace{12pt}
Spiritual Weapon

Evocation [Force]

\textbf{Level:} Clr 2, War 2

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Effect: }Magic weapon of force

\textbf{Duration:} 1 round/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

A weapon made of pure force springs into existence and attacks opponents at a distance, 
as you direct it, dealing 1d8 force damage per hit, +1 point per three caster levels 
(maximum +5 at 15th level). The weapon takes the shape of a weapon favored by your 
deity or a weapon with some spiritual significance or symbolism to you (see below) 
and has the same threat range and critical multipliers as a real weapon of its 
form. It strikes the opponent you designate, starting with one attack in the round 
the spell is cast and continuing each round thereafter on your turn. It uses your 
base attack bonus (possibly allowing it multiple attacks per round in subsequent 
rounds) plus your Wisdom modifier as its attack bonus. It strikes as a spell, not 
as a weapon, so, for example, it can damage creatures that have damage reduction. 
As a force effect, it can strike incorporeal creatures without the normal miss 
chance associated with incorporeality. The weapon always strikes from your direction. 
It does not get a flanking bonus or help a combatant get one. Your feats or combat 
actions do not affect the weapon. If the weapon goes beyond the spell range, if 
it goes out of your sight, or if you are not directing it, the weapon returns to 
you and hovers.

Each round after the first, you can use a move action to redirect the weapon to 
a new target. If you do not, the weapon continues to attack the previous round's 
target. On any round that the weapon switches targets, it gets one attack. Subsequent 
rounds of attacking that target allow the weapon to make multiple attacks if your 
base attack bonus would allow it to. Even if the \textit{spiritual weapon }is a 
ranged weapon, use the spell's range, not the weapon's normal range increment, 
and switching targets still is a move action.

A \textit{spiritual weapon }cannot be attacked or harmed by physical attacks, but 
\textit{dispel magic}, \textit{disintegrate}, a \textit{sphere of annihilation}, 
or a \textit{rod of cancellation }affects it. A \textit{spiritual weapon'}s AC 
against touch attacks is 12 (10 + size bonus for Tiny object).

If an attacked creature has spell resistance, you make a caster level check (1d20 
+ caster level) against that spell resistance the first time the \textit{spiritual 
weapon }strikes it. If the weapon is successfully resisted, the spell is dispelled. 
If not, the weapon has its normal full effect on that creature for the duration 
of the spell.

The weapon that you get is often a force replica of your deity's own personal weapon. 
A cleric without a deity gets a weapon based on his alignment. A neutral cleric 
without a deity can create a \textit{spiritual weapon }of any alignment, provided 
he is acting at least generally in accord with that alignment at the time. The 
weapons associated with each alignment are as follows.

\textit{Chaos: }Battleaxe

\textit{Evil: }Light flail

\textit{Good: }Warhammer

\textit{Law: }Longsword, 

\vspace{12pt}
Statue

Transmutation

\textbf{Level:} Sor/Wiz 7

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 round

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 1 hour/level (D)

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

A \textit{statue }spell turns the subject to solid stone, along with any garments 
and equipment worn or carried. In statue form, the subject gains hardness 8. The 
subject retains its own hit points.

The subject can see, hear, and smell normally, but it does not need to eat or breathe. 
Feeling is limited to those sensations that can affect the granite-hard substance 
of the individual's body. Chipping is equal to a mere scratch, but breaking off 
one of the statue's arms constitutes serious damage.

The subject of a \textit{statue }spell can return to its normal state, act, and 
then return instantly to the statue state (a free action) if it so desires, as 
long as the spell duration is in effect.

\textit{Material Component: }Lime, sand, and a drop of water stirred by an iron 
bar, such as a nail or spike.

\vspace{12pt}
Status

Divination

\textbf{Level:} Clr 2

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Targets:} One living creature touched per three levels

\textbf{Duration:} 1 hour/level

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

When you need to keep track of comrades who may get separated, \textit{status }allows 
you to mentally monitor their relative positions and general condition. You are 
aware of direction and distance to the creatures and any conditions affecting them: 
unharmed, wounded, disabled, staggered, unconscious, dying, nauseated, panicked, 
stunned, poisoned, diseased, \textit{confused, }or the like. Once the spell has 
been cast upon the subjects, the distance between them and the caster does not 
affect the spell as long as they are on the same plane of existence. If a subject 
leaves the plane, or if it dies, the spell ceases to function for it.

\vspace{12pt}
Stinking Cloud

Conjuration (Creation)

\textbf{Level:} Sor/Wiz 3

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Effect:} Cloud spreads in 20-ft. radius, 20 ft. high

\textbf{Duration:} 1 round/level

\textbf{Saving Throw:} Fortitude negates; see text

\textbf{Spell Resistance:} No

\textit{Stinking cloud }creates a bank of fog like that created by \textit{fog 
cloud}, except that the vapors are nauseating. Living creatures in the cloud become 
nauseated. This condition lasts as long as the creature is in the cloud and for 
1d4+1 rounds after it leaves. (Roll separately for each nauseated character.) Any 
creature that succeeds on its save but remains in the cloud must continue to save 
each round on your turn.

\textit{Stinking cloud }can be made permanent with a \textit{permanency }spell. 
A permanent \textit{stinking cloud }dispersed by wind reforms in 10 minutes.

\textit{Material Component: }A rotten egg or several skunk cabbage leaves.

\vspace{12pt}
Stone Shape

Transmutation [Earth]

\textbf{Level:} Clr 3, Drd 3, Earth 3, Sor/Wiz 4

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Stone or stone object touched, up to 10 cu. ft. + 1 cu. ft./level

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You can form an existing piece of stone into any shape that suits your purpose. 
While it's possible to make crude coffers, doors, and so forth with \textit{stone 
shape, }fine detail isn't possible. There is a 30\% chance that any shape including 
moving parts simply doesn't work.

\textit{Arcane Material Component: }Soft clay, which must be worked into roughly 
the desired shape of the stone object and then touched to the stone while the verbal 
component is uttered.

\vspace{12pt}
Stoneskin

Abjuration

\textbf{Level:} Drd 5, Earth 6, Sor/Wiz 4, Strength 6

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 10 min./level or until discharged

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

The warded creature gains resistance to blows, cuts, stabs, and slashes. The subject 
gains damage reduction 10/adamantine. (It ignores the first 10 points of damage 
each time it takes damage from a weapon, though an adamantine weapon bypasses the 
reduction.) Once the spell has prevented a total of 10 points of damage per caster 
level (maximum 150 points), it is discharged.

\textit{Material Component: }Granite and 250 gp worth of diamond dust sprinkled 
on the target's skin.

\vspace{12pt}
Stone Tell

Divination

\textbf{Level:} Drd 6

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 1 min./level

You gain the ability to speak with stones, which relate to you who or what has 
touched them as well as revealing what is covered or concealed behind or under 
them. The stones relate complete descriptions if asked. A stone's perspective, 
perception, and knowledge may prevent the stone from providing the details you 
are looking for.

You can speak with natural or worked stone.

\vspace{12pt}
Stone to Flesh

Transmutation

\textbf{Level:} Sor/Wiz 6

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Target:} One petrified creature or a cylinder of stone from 1 ft. to 3 
ft. in diameter and up to 10 ft. long

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} Fortitude negates (object); see text

\textbf{Spell Resistance:} Yes

This spell restores a petrified creature to its normal state, restoring life and 
goods. The creature must make a DC 15 Fortitude save to survive the process. Any 
petrified creature, regardless of size, can be restored.

The spell also can convert a mass of stone into a fleshy substance. Such flesh 
is inert and lacking a vital life force unless a life force or magical energy is 
available. (For example, this spell would turn a stone golem into a flesh golem, 
but an ordinary statue would become a corpse.) You can affect an object that fits 
within a cylinder from 1 foot to 3 feet in diameter and up to 10 feet long or a 
cylinder of up to those dimensions in a larger mass of stone.

\textit{Material Component: }A pinch of earth and a drop of blood.

\vspace{12pt}
Storm of Vengeance

Conjuration (Summoning)

\textbf{Level:} Drd 9, Clr 9

\textbf{Components:} V, S

\textbf{Casting Time:} 1 round

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Effect:} 360-ft.-radius storm cloud

\textbf{Duration:} Concentration (maximum 10 rounds) (D)

\textbf{Saving Throw:} See text

\textbf{Spell Resistance:} Yes

This spell creates an enormous black storm cloud. Lightning and crashing claps 
of thunder appear within the storm. Each creature beneath the cloud must succeed 
on a Fortitude save or be deafened for 1d4x10 minutes.

If you do not maintain concentration on the spell after casting it, the spell ends. 
 If you continue to concentrate, the spell generates additional effects in each 
following round, as noted below. Each effect occurs during your turn.

\textit{2nd Round: }Acid rains down in the area, dealing 1d6 points of acid damage 
(no save).

\textit{3rd Round: }You call six bolts of lightning down from the cloud. You decide 
where the bolts strike. No two bolts may be directed at the same target. Each bolt 
deals 10d6 points of electricity damage. A creature struck can attempt a Reflex 
save for half damage.

\textit{4th Round: }Hailstones rain down in the area, dealing 5d6 points of bludgeoning 
damage (no save).

\textit{5th through 10th Rounds: }Violent rain and wind gusts reduce visibility. 
The rain obscures all sight, including darkvision, beyond 5 feet. A creature 5 
feet away has concealment (attacks have a 20\% miss chance). Creatures farther 
away have total concealment (50\% miss chance, and the attacker cannot use sight 
to locate the target). Speed is reduced by three-quarters.

Ranged attacks within the area of the storm are impossible. Spells cast within 
the area are disrupted unless the caster succeeds on a Concentration check against 
a DC equal to the \textit{storm of vengeance'}s save DC + the level of the spell 
the caster is trying to cast.

\vspace{12pt}
Suggestion

Enchantment (Compulsion) [Language-Dependent, Mind-Affecting]

\textbf{Level:} Brd 2, Sor/Wiz 3

\textbf{Components:} V, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One living creature

\textbf{Duration:} 1 hour/level or until completed

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

You influence the actions of the target creature by suggesting a course of activity 
(limited to a sentence or two). The \textit{suggestion }must be worded in such 
a manner as to make the activity sound reasonable. Asking the creature to do some 
obviously harmful act automatically negates the effect of the spell. 

The suggested course of activity can continue for the entire duration. If the suggested 
activity can be completed in a shorter time, the spell ends when the subject finishes 
what it was asked to do. You can instead specify conditions that will trigger a 
special activity during the duration. If the condition is not met before the spell 
duration expires, the activity is not performed.

A very reasonable \textit{suggestion }causes the save to be made with a penalty 
(such as -1 or -2).

\textit{Material Component: }A snake's tongue and either a bit of honeycomb or 
a drop of sweet oil.

\vspace{12pt}
Suggestion, Mass

Enchantment (Compulsion) [Language-Dependent, Mind-Affecting]

\textbf{Level:} Brd 5, Sor/Wiz 6

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Targets:} One creature/level, no two of which can be more than 30 ft. apart

This spell functions like \textit{suggestion}, except that it can affect more creatures. 
The same \textit{suggestion }applies to all these creatures.

\vspace{12pt}
Summon Instrument

Conjuration (Summoning)

\textbf{Level:} Brd 0

\textbf{Components:} V, S

\textbf{Casting Time:} 1 round

\textbf{Range:} 0 ft.

\textbf{Effect:} One summoned handheld musical instrument

\textbf{Duration:} 1 min./level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell summons one handheld musical instrument of your choice. This instrument 
appears in your hands or at your feet (your choice). The instrument is typical 
for its type. Only one instrument appears per casting, and it will play only for 
you. You can't summon an instrument too large to be held in two hands.

\vspace{12pt}
Summon Monster I

Conjuration (Summoning) [see text]

\textbf{Level:} Brd 1, Clr 1, Sor/Wiz 1

\textbf{Components:} V, S, F/DF

\textbf{Casting Time:} 1 round

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} One summoned creature

\textbf{Duration:} 1 round/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell summons an extraplanar creature (typically an outsider, elemental, or 
magical beast native to another plane). It appears where you designate and acts 
immediately, on your turn. It attacks your opponents to the best of its ability. 
If you can communicate with the creature, you can direct it not to attack, to attack 
particular enemies, or to perform other actions.

The spell conjures one of the creatures from the 1st-level list on the accompanying 
Summon Monster table. You choose which kind of creature to summon, and you can 
change that choice each time you cast the spell.

A summoned monster cannot summon or otherwise conjure another creature, nor can 
it use any teleportation or planar travel abilities. Creatures cannot be summoned 
into an environment that cannot support them.

When you use a summoning spell to summon an air, chaotic, earth, evil, fire, good, 
lawful, or water creature, it is a spell of that type.

\textit{Arcane Focus: }A tiny bag and a small (not necessarily lit) candle.

\vspace{12pt}
Summon Monster II

Conjuration (Summoning) [see text for \textit{summon monster I}]

\textbf{Level:} Brd 2, Clr 2, Sor/Wiz 2

\textbf{Effect:} One or more summoned creatures, no two of which can be more than 
30 ft. apart

This spell functions like \textit{summon monster I, }except that you can summon 
one creature from the 2nd-level list or 1d3 creatures of the same kind from the 
1st-level list.

\vspace{12pt}
Summon Monster III

Conjuration (Summoning) [see text for \textit{summon monster I}]

\textbf{Level:} Brd 3, Clr 3, Sor/Wiz 3

\textbf{Effect:} One or more summoned creatures, no two of which can be more than 
30 ft. apart

This spell functions like \textit{summon monster I, }except that you can summon 
one creature from the 3rd-level list, 1d3 creatures of the same kind from the 2nd-level 
list, or 1d4+1 creatures of the same kind from the 1st-level list.

\vspace{12pt}
Summon Monster IV

Conjuration (Summoning) [see text for \textit{summon monster I}]

\textbf{Level:} Brd 4, Clr 4, Sor/Wiz 4

\textbf{Effect:} One or more summoned creatures, no two of which can be more than 
30 ft. apart

This spell functions like \textit{summon monster I, }except that you can summon 
one creature from the 4th-level list, 1d3 creatures of the same kind from the 3rd-level 
list, or 1d4+1 creatures of the same kind from a lower-level list.

\vspace{12pt}
Summon Monster V

Conjuration (Summoning) [see text for \textit{summon monster I}]

\textbf{Level:} Brd 5, Clr 5, Sor/Wiz 5

\textbf{Effect:} One or more summoned creatures, no two of which can be more than 
30 ft. apart

This spell functions like \textit{summon monster I, }except that you can summon 
one creature from the 5th-level list, 1d3 creatures of the same kind from the 4th-level 
list, or 1d4+1 creatures of the same kind from a lower-level list.

\vspace{12pt}
Summon Monster VI

Conjuration (Summoning) [see text for \textit{summon monster I}]

\textbf{Level:} Brd 6, Clr 6, Sor/Wiz 6

\textbf{Effect:} One or more summoned creatures, no two of which can be more than 
30 ft. apart

This spell functions like \textit{summon monster I, }except you can summon one 
creature from the 6th-level list, 1d3 creatures of the same kind from the 5th-level 
list, or 1d4+1 creatures of the same kind from a lower-level list.

\vspace{12pt}
Summon Monster VII

Conjuration (Summoning) [see text for \textit{summon monster I}]

\textbf{Level:} Clr 7, Sor/Wiz 7

This spell functions like \textit{summon monster I, }except that you can summon 
one creature from the 7th-level list, 1d3 creatures of the same kind from the 6th-level 
list, or 1d4+1 creatures of the same kind from a lower-level list.

\vspace{12pt}
Summon Monster VIII

Conjuration (Summoning) [see text for \textit{summon monster I}]

\textbf{Level:} Clr 8, Sor/Wiz 8

This spell functions like \textit{summon monster I, }except that you can summon 
one creature from the 8th-level list, 1d3 creatures of the same kind from the 7th-level 
list, or 1d4+1 creatures of the same kind from a lower-level list.

\vspace{12pt}
Summon Monster IX

Conjuration (Summoning) [see text for \textit{summon monster I}]

\textbf{Level:} Chaos 9, Clr 9, Evil 9, Good 9, Law 9, Sor/Wiz 9

This spell functions like \textit{summon monster I, }except that you can summon 
one creature from the 9th-level list, 1d3 creatures of the same kind from the 8th-level 
list, or 1d4+1 creatures of the same kind from a lower-level list.

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{227pt}|>{\raggedright}p{21pt}|}
\hline
S\textbf{ummon Monster} & \tabularnewline
\hline
1\textbf{st Level} & \tabularnewline
\hline
Celestial dog & LG\tabularnewline
\hline
Celestial owl & LG\tabularnewline
\hline
Celestial giant fire beetle & NG\tabularnewline
\hline
Celestial porpoise\textsuperscript{\textbf{1}} & NG\tabularnewline
\hline
Celestial badger & CG\tabularnewline
\hline
Celestial monkey & CG\tabularnewline
\hline
Fiendish dire rat & LE\tabularnewline
\hline
Fiendish raven & LE\tabularnewline
\hline
Fiendish monstrous centipede, Medium & NE\tabularnewline
\hline
Fiendish monstrous scorpion, Small & NE\tabularnewline
\hline
Fiendish hawk & CE\tabularnewline
\hline
Fiendish monstrous spider, Small & CE\tabularnewline
\hline
Fiendish octopus\textsuperscript{\textbf{1}} & CE\tabularnewline
\hline
Fiendish snake, Small viper & CE\tabularnewline
\hline
2\textbf{nd Level} & \tabularnewline
\hline
Celestial giant bee & LG\tabularnewline
\hline
Celestial giant bombardier beetle & NG\tabularnewline
\hline
Celestial riding dog & NG\tabularnewline
\hline
Celestial eagle & CG\tabularnewline
\hline
Lemure (devil) & LE\tabularnewline
\hline
Fiendish squid\textsuperscript{\textbf{1}} & LE\tabularnewline
\hline
Fiendish wolf & LE\tabularnewline
\hline
Fiendish monstrous centipede, Large & NE\tabularnewline
\hline
Fiendish monstrous scorpion, Medium & NE\tabularnewline
\hline
Fiendish shark, Medium\textsuperscript{\textbf{1}} & NE\tabularnewline
\hline
Fiendish monstrous spider, Medium & CE\tabularnewline
\hline
Fiendish snake, Medium viper & CE\tabularnewline
\hline
3\textbf{rd Level} & \tabularnewline
\hline
Celestial black bear & LG\tabularnewline
\hline
Celestial bison & NG\tabularnewline
\hline
Celestial dire badger & CG\tabularnewline
\hline
Celestial hippogriff & CG\tabularnewline
\hline
Elemental, Small (any) & N\tabularnewline
\hline
Fiendish ape & LE\tabularnewline
\hline
Fiendish dire weasel & LE\tabularnewline
\hline
Hell hound & LE\tabularnewline
\hline
Fiendish snake, constrictor  & LE\tabularnewline
\hline
Fiendish boar & NE\tabularnewline
\hline
Fiendish dire bat & NE\tabularnewline
\hline
Fiendish monstrous centipede, Huge & NE\tabularnewline
\hline
Fiendish crocodile & CE\tabularnewline
\hline
Dretch (demon) & CE\tabularnewline
\hline
Fiendish snake, Large viper & CE\tabularnewline
\hline
Fiendish wolverine & CE\tabularnewline
\hline
4\textbf{th Level} & \tabularnewline
\hline
Archon, lantern & LG\tabularnewline
\hline
Celestial giant owl & LG\tabularnewline
\hline
Celestial giant eagle & CG\tabularnewline
\hline
Celestial lion & CG\tabularnewline
\hline
Mephit (any) & N\tabularnewline
\hline
Fiendish dire wolf & LE\tabularnewline
\hline
Fiendish giant wasp & LE\tabularnewline
\hline
Fiendish giant praying mantis & NE\tabularnewline
\hline
Fiendish shark, Large\textsuperscript{\textbf{1}} & NE\tabularnewline
\hline
Yeth hound & NE\tabularnewline
\hline
Fiendish monstrous spider, Large & CE\tabularnewline
\hline
Fiendish snake, Huge viper & CE\tabularnewline
\hline
Howler & CE\tabularnewline
\hline
5\textbf{th Level} & \tabularnewline
\hline
Archon, hound & LG\tabularnewline
\hline
Celestial brown bear & LG\tabularnewline
\hline
Celestial giant stag beetle & NG\tabularnewline
\hline
Celestial sea cat\textsuperscript{\textbf{1}} & NG\tabularnewline
\hline
Celestial griffon & CG\tabularnewline
\hline
Elemental, Medium (any) & N\tabularnewline
\hline
Achaierai & LE\tabularnewline
\hline
Devil, bearded & LE\tabularnewline
\hline
Fiendish deinonychus & LE\tabularnewline
\hline
Fiendish dire ape & LE\tabularnewline
\hline
Fiendish dire boar & NE\tabularnewline
\hline
Fiendish shark, Huge & NE\tabularnewline
\hline
Fiendish monstrous scorpion, Large & NE\tabularnewline
\hline
Shadow mastiff & NE\tabularnewline
\hline
Fiendish dire wolverine & CE\tabularnewline
\hline
Fiendish giant crocodile & CE\tabularnewline
\hline
Fiendish tiger & CE\tabularnewline
\hline
6\textbf{th Level} & \tabularnewline
\hline
Celestial polar bear & LG\tabularnewline
\hline
Celestial orca whale\textsuperscript{\textbf{1}} & NG\tabularnewline
\hline
Bralani (eladrin) & CG\tabularnewline
\hline
Celestial dire lion & CG\tabularnewline
\hline
Elemental, Large (any) & N\tabularnewline
\hline
Janni (genie) & N\tabularnewline
\hline
Chaos beast & CN\tabularnewline
\hline
Devil, chain & LE\tabularnewline
\hline
Xill & LE\tabularnewline
\hline
Fiendish monstrous centipede, Gargantuan & NE\tabularnewline
\hline
Fiendish rhinoceros & NE\tabularnewline
\hline
Fiendish elasmosaurus\textsuperscript{\textbf{1}} & CE\tabularnewline
\hline
Fiendish monstrous spider, Huge & CE\tabularnewline
\hline
Fiendish snake, giant constrictor & CE\tabularnewline
\hline
7\textbf{th Level} & \tabularnewline
\hline
Celestial elephant & LG\tabularnewline
\hline
Avoral (guardinal) & NG\tabularnewline
\hline
Celestial baleen whale\textsuperscript{\textbf{1}} & NG\tabularnewline
\hline
Djinni (genie) & CG\tabularnewline
\hline
Elemental, Huge (any) & N\tabularnewline
\hline
Invisible stalker & N\tabularnewline
\hline
Devil, bone & LE\tabularnewline
\hline
Fiendish megaraptor & LE\tabularnewline
\hline
Fiendish monstrous scorpion, Huge & NE\tabularnewline
\hline
Babau (demon) & CE\tabularnewline
\hline
Fiendish giant octopus\textsuperscript{\textbf{1}} & CE\tabularnewline
\hline
Fiendish girallon & CE\tabularnewline
\hline
8\textbf{th Level} & \tabularnewline
\hline
Celestial dire bear & LG\tabularnewline
\hline
Celestial cachalot whale\textsuperscript{\textbf{1}} & NG\tabularnewline
\hline
Celestial triceratops & NG\tabularnewline
\hline
Lillend & CG\tabularnewline
\hline
Elemental, greater (any) & N\tabularnewline
\hline
Fiendish giant squid\textsuperscript{\textbf{1}} & LE\tabularnewline
\hline
Hellcat & LE\tabularnewline
\hline
Fiendish monstrous centipede, Colossal & NE\tabularnewline
\hline
Fiendish dire tiger & CE\tabularnewline
\hline
Fiendish monstrous spider, Gargantuan & CE\tabularnewline
\hline
Fiendish tyrannosaurus & CE\tabularnewline
\hline
Vrock (demon) & CE\tabularnewline
\hline
9\textbf{th Level} & \tabularnewline
\hline
Couatl & LG\tabularnewline
\hline
Leonal (guardinal) & NG\tabularnewline
\hline
Celestial roc & CG\tabularnewline
\hline
Elemental, elder (any) & N\tabularnewline
\hline
Devil, barbed & LE\tabularnewline
\hline
Fiendish dire shark\textsuperscript{\textbf{1}} & NE\tabularnewline
\hline
Fiendish monstrous scorpion, Gargantuan & NE\tabularnewline
\hline
Night hag & NE\tabularnewline
\hline
Bebilith (demon) & CE\tabularnewline
\hline
Fiendish monstrous spider, Colossal & CE\tabularnewline
\hline
Hezrou (demon) & CE\tabularnewline
\hline
\multicolumn{2}{|p{248pt}|}{1 May be summoned only into an aquatic or watery environment.}\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
Summon Nature's Ally I

Conjuration (Summoning)

\textbf{Level:} Drd 1, Rgr 1

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 round

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} One summoned creature

\textbf{Duration:} 1 round/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell summons a natural creature. It appears where you designate and acts 
immediately, on your turn. It attacks your opponents to the best of its ability. 
If you can communicate with the creature, you can direct it not to attack, to attack 
particular enemies, or to perform other actions.

A summoned monster cannot summon or otherwise conjure another creature, nor can 
it use any teleportation or planar travel abilities. Creatures cannot be summoned 
into an environment that cannot support them.

The spell conjures one of the creatures from the 1st-level list on the accompanying 
Summon Nature's Ally table. You choose which kind of creature to summon, and you 
can change that choice each time you cast the spell. All the creatures on the table 
are neutral unless otherwise noted.

\vspace{12pt}
Summon Nature's Ally II

Conjuration (Summoning)

\textbf{Level:} Drd 2, Rgr 2

\textbf{Effect:} One or more creatures, no two of which can be more than 30 ft. 
apart

This spell functions like \textit{summon nature's ally I, }except that you can 
summon one 2nd-level creature or 1d3 1st-level creatures of the same kind.

\vspace{12pt}
Summon Nature's Ally III

Conjuration (Summoning) [see text]

\textbf{Level:} Drd 3, Rgr 3

\textbf{Effect:} One or more creatures, no two of which can be more than 30 ft. 
apart

This spell functions like \textit{summon nature's ally I, }except that you can 
summon one 3rd-level creature, 1d3 2nd-level creatures of the same kind, or 1d4+1 
1st-level creatures of the same kind.

When you use a summoning spell to summon an air, chaotic, earth, evil, fire, good, 
lawful, or water creature, it is a spell of that type. 

Summon Nature's Ally IV

Conjuration (Summoning) [see text]

\textbf{Level:} Animal 4, Drd 4, Rgr 4

\textbf{Effect:} One or more creatures, no two of which can be more than 30 ft. 
apart

This spell functions like \textit{summon nature's ally I, }except that you can 
summon one 4th-level creature, 1d3 3rd-level creatures of the same kind, or 1d4+1 
lower-level creatures of the same kind.

When you use a summoning spell to summon an air, chaotic, earth, evil, fire, good, 
lawful, or water creature, it is a spell of that type.

\vspace{12pt}
Summon Nature's Ally V

Conjuration (Summoning) [see text]

\textbf{Level:} Drd 5

\textbf{Effect:} One or more creatures, no two of which can be more than 30 ft. 
apart

This spell functions like \textit{summon nature's ally I, }except that you can 
summon one 5th-level creature, 1d3 4th-level creatures of the same kind, or 1d4+1 
lower-level creatures of the same kind.

When you use a summoning spell to summon an air, chaotic, earth, evil, fire, good, 
lawful, or water creature, it is a spell of that type.

\vspace{12pt}
Summon Nature's Ally VI

Conjuration (Summoning) [see text]

\textbf{Level:} Drd 6

\textbf{Effect:} One or more creatures, no two of which can be more than 30 ft. 
apart

This spell functions like \textit{summon nature's ally I, }except that you can 
summon one 6th-level creature, 1d3 5th-level creatures of the same kind, or 1d4+1 
lower-level creatures of the same kind.

When you use a summoning spell to summon an air, chaotic, earth, evil, fire, good, 
lawful, or water creature, it is a spell of that type.

\vspace{12pt}
Summon Nature's Ally VII

Conjuration (Summoning) [see text]

\textbf{Level:} Drd 7

\textbf{Effect:} One or more creatures, no two of which can be more than 30 ft. 
apart

This spell functions like \textit{summon nature's ally I, }except that you can 
summon one 7th-level creature, 1d3 6th-level creatures of the same kind, or 1d4+1 
lower-level creatures of the same kind.

When you use a summoning spell to summon an air, chaotic, earth, evil, fire, good, 
lawful, or water creature, it is a spell of that type.

\vspace{12pt}
Summon Nature's Ally VIII

Conjuration (Summoning) [see text]

\textbf{Level:} Animal 8, Drd 8

\textbf{Effect:} One or more creatures, no two of which can be more than 30 ft. 
apart

This spell functions like \textit{summon nature's ally I, }except that you can 
summon one 8th-level creature, 1d3 7th-level creatures of the same kind, or 1d4+1 
lower-level creatures of the same kind.

When you use a summoning spell to summon an air, chaotic, earth, evil, fire, good, 
lawful, or water creature, it is a spell of that type.

\vspace{12pt}
Summon Nature's Ally IX

Conjuration (Summoning) [see text]

\textbf{Level:} Drd 9

\textbf{Effect:} One or more creatures, no two of which can be more than 30 ft. 
apart

This spell functions like \textit{summon nature's ally I, }except that you can 
summon one 9th-level creature, 1d3 8th-level creatures of the same kind, or 1d4+1 
lower-level creatures of the same kind.

When you use a summoning spell to summon an air, chaotic, earth, evil, fire, good, 
lawful, or water creature, it is a spell of that type.

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{260pt}|}
\hline
S\textbf{ummon Nature's Ally}\tabularnewline
\hline
1\textbf{st Level}\tabularnewline
\hline
Dire rat\tabularnewline
\hline
Eagle (animal)\tabularnewline
\hline
Monkey (animal)\tabularnewline
\hline
Octopus\textsuperscript{\textbf{1}}\textbf{ }(animal)\tabularnewline
\hline
Owl (animal)\tabularnewline
\hline
Porpoise\textsuperscript{\textbf{1}}\textbf{ }(animal)\tabularnewline
\hline
Snake, Small viper (animal)\tabularnewline
\hline
Wolf (animal)\tabularnewline
\hline
2\textbf{nd Level}\tabularnewline
\hline
Bear, black (animal)\tabularnewline
\hline
Crocodile (animal)\tabularnewline
\hline
Dire badger\tabularnewline
\hline
Dire bat\tabularnewline
\hline
Elemental, Small (any)\tabularnewline
\hline
Hippogriff\tabularnewline
\hline
Shark, Medium\textsuperscript{\textbf{1}}\textbf{ }(animal)\tabularnewline
\hline
Snake, Medium viper (animal)\tabularnewline
\hline
Squid\textsuperscript{\textbf{1}}\textbf{ }(animal)\tabularnewline
\hline
Wolverine (animal)\tabularnewline
\hline
3\textbf{rd Level}\tabularnewline
\hline
Ape (animal)\tabularnewline
\hline
Dire weasel\tabularnewline
\hline
Dire wolf\tabularnewline
\hline
Eagle, giant [NG]\tabularnewline
\hline
Lion\tabularnewline
\hline
Owl, giant [NG]\tabularnewline
\hline
Satyr [CN; without pipes]\tabularnewline
\hline
Shark, Large\textsuperscript{\textbf{1}}\textbf{ }(animal)\tabularnewline
\hline
Snake, constrictor (animal)\tabularnewline
\hline
Snake, Large viper (animal)\tabularnewline
\hline
Thoqqua\tabularnewline
\hline
4\textbf{th Level}\tabularnewline
\hline
Arrowhawk, juvenile\tabularnewline
\hline
Bear, brown (animal)\tabularnewline
\hline
Crocodile, giant (animal)\tabularnewline
\hline
Deinonychus (dinosaur)\tabularnewline
\hline
Dire ape\tabularnewline
\hline
Dire boar\tabularnewline
\hline
Dire wolverine\tabularnewline
\hline
Elemental, Medium (any)\tabularnewline
\hline
Salamander, flamebrother [NE]\tabularnewline
\hline
Sea cat\textsuperscript{\textbf{1}}\tabularnewline
\hline
Shark, Huge\textsuperscript{\textbf{1}}\textbf{ }(animal)\tabularnewline
\hline
Snake, Huge viper (animal)\tabularnewline
\hline
Tiger (animal)\tabularnewline
\hline
Tojanida, juvenile\textsuperscript{\textbf{1}}\tabularnewline
\hline
Unicorn [CG]\tabularnewline
\hline
Xorn, minor\tabularnewline
\hline
5\textbf{th Level}\tabularnewline
\hline
Arrowhawk, adult\tabularnewline
\hline
Bear, polar (animal)\tabularnewline
\hline
Dire lion\tabularnewline
\hline
Elasmosaurus\textsuperscript{\textbf{1}}\textbf{ }(dinosaur)\tabularnewline
\hline
Elemental, Large (any)\tabularnewline
\hline
Griffon\tabularnewline
\hline
Janni (genie)\tabularnewline
\hline
Rhinoceros (animal)\tabularnewline
\hline
Satyr [CN; with pipes]\tabularnewline
\hline
Snake, giant constrictor (animal)\tabularnewline
\hline
Nixie (sprite)\tabularnewline
\hline
Tojanida, adult\textsuperscript{\textbf{1}}\tabularnewline
\hline
Whale, orca\textsuperscript{\textbf{1}}\textbf{ }(animal)\tabularnewline
\hline
6\textbf{th Level}\tabularnewline
\hline
Dire bear\tabularnewline
\hline
Elemental, Huge (any)\tabularnewline
\hline
Elephant (animal)\tabularnewline
\hline
Girallon\tabularnewline
\hline
Megaraptor (dinosaur)\tabularnewline
\hline
Octopus, giant\textsuperscript{\textbf{1}}\textbf{ }(animal)\tabularnewline
\hline
Pixie* (sprite) [NG; no special arrows]\tabularnewline
\hline
Salamander, average [NE]\tabularnewline
\hline
Whale, baleen\textsuperscript{\textbf{1}}\tabularnewline
\hline
Xorn, average\tabularnewline
\hline
*Can't cast \textit{irresistible dance}\tabularnewline
\hline
7\textbf{th Level}\tabularnewline
\hline
Arrowhawk, elder\tabularnewline
\hline
Dire tiger\tabularnewline
\hline
Elemental, greater (any)\tabularnewline
\hline
Djinni (genie) [NG]\tabularnewline
\hline
Invisible stalker\tabularnewline
\hline
Pixie* (sprite) [NG; with sleep arrows]\tabularnewline
\hline
Squid, giant\textsuperscript{\textbf{1}}\textbf{ }(animal)\tabularnewline
\hline
Triceratops (dinosaur)\tabularnewline
\hline
Tyrannosaurus (dinosaur)\tabularnewline
\hline
Whale, cachalot\textsuperscript{\textbf{1}}\textbf{ }(animal)\tabularnewline
\hline
Xorn, elder\tabularnewline
\hline
*Can't cast \textit{irresistible dance}\tabularnewline
\hline
8\textbf{th Level}\tabularnewline
\hline
Dire shark\textsuperscript{\textbf{1}}\tabularnewline
\hline
Roc\tabularnewline
\hline
Salamander, noble [NE]\tabularnewline
\hline
Tojanida, elder\tabularnewline
\hline
9\textbf{th Level}\tabularnewline
\hline
Elemental, elder\tabularnewline
\hline
Grig [NG; with fiddle] (sprite)\tabularnewline
\hline
Pixie* (sprite) [NG; with sleep and memory loss arrows]\tabularnewline
\hline
Unicorn, celestial charger\tabularnewline
\hline
*Can cast \textit{irresistible dance}\tabularnewline
\hline
1 May be summoned only into an aquatic or watery environment.\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
Summon Swarm

Conjuration (Summoning)

\textbf{Level:} Brd 2, Drd 2, Sor/Wiz 2

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 round

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} One swarm of bats, rats, or spiders

\textbf{Duration:} Concentration + 2 rounds

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You summon a swarm of bats, rats, or spiders (your choice), which attacks all other 
creatures within its area. (You may summon the swarm so that it shares the area 
of other creatures.) If no living creatures are within its area, the swarm attacks 
or pursues the nearest creature as best it can. The caster has no control over 
its target or direction of travel.

\textit{Arcane Material Component: }A square of red cloth.

\vspace{12pt}
Sunbeam

Evocation [Light]

\textbf{Level:} Drd 7, Sun 7

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 60 ft.

\textbf{Area:} Line from your hand

\textbf{Duration:} 1 round/level or until all beams are exhausted

\textbf{Saving Throw:} Reflex negates and Reflex half; see text

\textbf{Spell Resistance:} Yes

For the duration of this spell, you can use a standard action to evoke a dazzling 
beam of intense light each round. You can call forth one beam per three caster 
levels (maximum six beams at 18th level). The spell ends when its duration runs 
out or your allotment of beams is exhausted.

Each creature in the beam is blinded and takes 4d6 points of damage. Any creatures 
to which sunlight is harmful or unnatural take double damage. A successful Reflex 
save negates the blindness and reduces the damage by half.

An undead creature caught within the beam takes 1d6 points of damage per caster 
level (maximum 20d6), or half damage if a Reflex save is successful. In addition, 
the beam results in the destruction of any undead creature specifically harmed 
by bright light if it fails its save.

The ultraviolet light generated by the spell deals damage to fungi, mold, oozes, 
and slimes just as if they were undead creatures.

\vspace{12pt}
Sunburst

Evocation [Light]

\textbf{Level:} Drd 8, Sor/Wiz 8, Sun 8

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Area:} 80-ft.-radius burst

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} Reflex partial; see text

\textbf{Spell Resistance:} Yes

\textit{Sunburst }causes a globe of searing radiance to explode silently from a 
point you select. All creatures in the globe are blinded and take 6d6 points of 
damage. A creature to which sunlight is harmful or unnatural takes double damage. 
A successful Reflex save negates the blindness and reduces the damage by half.

An undead creature caught within the globe takes 1d6 points of damage per caster 
level (maximum 25d6), or half damage if a Reflex save is successful. In addition, 
the burst results in the destruction of any undead creature specifically harmed 
by bright light if it fail its save.

The ultraviolet light generated by the spell deals damage to fungi, mold, oozes, 
and slimes just as if they were undead creatures.

\textit{Sunburst }dispels any darkness spells of lower than 9th level within its 
area.

\textit{Arcane Material Component: }A piece of sunstone and a naked flame.

\vspace{12pt}
Symbol of Death

Necromancy [Death]

\textbf{Level:} Clr 8, Sor/Wiz 8

\textbf{Components:} V, S, M

\textbf{Casting Time:} 10 minutes

\textbf{Range:} 0 ft.; see text

\textbf{Effect:} One symbol

\textbf{Duration:} See text

\textbf{Saving Throw:} Fortitude negates

\textbf{Spell Resistance:} Yes

This spell allows you to scribe a potent rune of power upon a surface. When triggered, 
a \textit{symbol of death }slays one or more creatures within 60 feet of the symbol 
(treat as a burst) whose combined total current hit points do not exceed 150. The 
\textit{symbol of death }affects the closest creatures first, skipping creatures 
with too many hit points to affect. Once triggered, the \textit{symbol }becomes 
active and glows, lasting for 10 minutes per caster level or until it has affected 
150 hit points' worth of creatures, whichever comes first. Any creature that enters 
the area while the \textit{symbol of death }is active is subject to its effect, 
whether or not that creature was in the area when it was triggered. A creature 
need save against the \textit{symbol }only once as long as it remains within the 
area, though if it leaves the area and returns while the \textit{symbol }is still 
active, it must save again.

Until it is triggered, the \textit{symbol of death }is inactive (though visible 
and legible at a distance of 60 feet). To be effective, a \textit{symbol of death 
}must always be placed in plain sight and in a prominent location. Covering or 
hiding the rune renders the \textit{symbol of death }ineffective, unless a creature 
removes the covering, in which case the \textit{symbol of death }works normally.

As a default, a \textit{symbol of death }is triggered whenever a creature does 
one or more of the following, as you select: looks at the rune; reads the rune; 
touches the rune; passes over the rune; or passes through a portal bearing the 
rune. Regardless of the trigger method or methods chosen, a creature more than 
60 feet from a symbol of death can't trigger it (even if it meets one or more of 
the triggering conditions, such as reading the rune). Once the spell is cast, a 
\textit{symbol of death'}s triggering conditions cannot be changed.

In this case, ``reading'' the rune means any attempt to study it, identify it, 
or fathom its meaning. Throwing a cover over a \textit{symbol of death }to render 
it inoperative triggers it if the symbol reacts to touch. You can't use a \textit{symbol 
of death }offensively; for instance, a touch-triggered symbol of death remains 
untriggered if an item bearing the \textit{symbol of death }is used to touch a 
creature. Likewise, a \textit{symbol of death }cannot be placed on a weapon and 
set to activate when the weapon strikes a foe.

You can also set special triggering limitations of your own. These can be as simple 
or elaborate as you desire. Special conditions for triggering a \textit{symbol 
of death }can be based on a creature's name, identity, or alignment, but otherwise 
must be based on observable actions or qualities. Intangibles such as level, class, 
Hit Dice, and hit points don't qualify. 

When scribing a \textit{symbol of death}, you can specify a password or phrase 
that prevents a creature using it from triggering the effect\textit{. }Anyone using 
the password remains immune to that particular rune's effects so long as the creature 
remains within 60 feet of the rune. If the creature leaves the radius and returns 
later, it must use the password again.

You also can attune any number of creatures to the \textit{symbol of death, }but 
doing this can extend the casting time. Attuning one or two creatures takes negligible 
time, and attuning a small group (as many as ten creatures) extends the casting 
time to 1 hour. Attuning a large group (as many as twenty-five creatures) takes 
24 hours. Attuning larger groups takes proportionately longer. Any creature attuned 
to a \textit{symbol of death }cannot trigger it and is immune to its effects, even 
if within its radius when triggered. You are automatically considered attuned to 
your own \textit{symbols of death}, and thus always ignore the effects and cannot 
inadvertently trigger them.

\textit{Read magic }allows you to identify a \textit{symbol of death }with a DC 
19 Spellcraft check. Of course, if the \textit{symbol of death }is set to be triggered 
by reading it, this will trigger the symbol.

A \textit{symbol of death }can be removed by a successful \textit{dispel magic 
}targeted solely on the rune. An \textit{erase }spell has no effect on a \textit{symbol 
of death. }Destruction of the surface where a \textit{symbol of death }is inscribed 
destroys the \textit{symbol }but also triggers it.

\textit{Symbol of death }can be made permanent with a \textit{permanency }spell. 
A permanent \textit{symbol of death }that is disabled or that has affected its 
maximum number of hit points becomes inactive for 10 minutes, then can be triggered 
again as normal.

\textit{Note: }Magic traps such as \textit{symbol of death }are hard to detect 
and disable. A rogue (only) can use the Search skill to find a \textit{symbol of 
death }and Disable Device to thwart it. The DC in each case is 25 + spell level, 
or 33 for \textit{symbol of death}.

\textit{Material Component: }Mercury and phosphorus, plus powdered diamond and 
opal with a total value of at least 5,000 gp each.

\vspace{12pt}
Symbol of Fear

Necromancy [Fear, Mind-Affecting]

\textbf{Level:} Clr 6, Sor/Wiz 6

\textbf{Saving Throw: }Will negates

This spell functions like \textit{symbol of death}, except that all creatures within 
60 feet of the \textit{symbol of fear }instead become panicked for 1 round per 
caster level.

\textit{Note: }Magic traps such as \textit{symbol of fear }are hard to detect and 
disable. A rogue (only) can use the Search skill to find a \textit{symbol of fear 
}and Disable Device to thwart it. The DC in each case is 25 + spell level, or 31 
for \textit{symbol of fear}.

\textit{Material Component: }Mercury and phosphorus, plus powdered diamond and 
opal with a total value of at least 1,000 gp.

\vspace{12pt}
Symbol of Insanity

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Clr 8, Sor/Wiz 8

\textbf{Saving Throw: }Will negates

This spell functions like \textit{symbol of death}, except that all creatures within 
the radius of the \textit{symbol of insanity }instead become permanently insane 
(as the \textit{insanity }spell).

Unlike \textit{symbol of death}, symbol of insanity has no hit point limit; once 
triggered, a \textit{symbol of insanity }simply remains active for 10 minutes per 
caster level.

\textit{Note: }Magic traps such as \textit{symbol of insanity }are hard to detect 
and disable. A rogue (only) can use the Search skill to find a \textit{symbol of 
insanity }and Disable Device to thwart it. The DC in each case is 25 + spell level, 
or 33 for \textit{symbol of insanity}.

\textit{Material Component: }Mercury and phosphorus, plus powdered diamond and 
opal with a total value of at least 5,000 gp.

\vspace{12pt}
Symbol of Pain

Necromancy [Evil]

\textbf{Level:} Clr 5, Sor/Wiz 5

This spell functions like \textit{symbol of death}, except that each creature within 
the radius of a \textit{symbol of pain }instead suffers wracking pains that impose 
a -4 penalty on attack rolls, skill checks, and ability checks. These effects last 
for 1 hour after the creature moves farther than 60 feet from the symbol.

Unlike \textit{symbol of death}, \textit{symbol of pain }has no hit point limit; 
once triggered, a \textit{symbol of pain }simply remains active for 10 minutes 
per caster level.

\textit{Note: }Magic traps such as \textit{symbol of pain }are hard to detect and 
disable. A rogue (only) can use the Search skill to find a \textit{symbol of pain 
}and Disable Device to thwart it. The DC in each case is 25 + spell level, or 30 
for \textit{symbol of pain}.

\textit{Material Component: }Mercury and phosphorus, plus powdered diamond and 
opal with a total value of at least 1,000 gp.

\vspace{12pt}
Symbol of Persuasion

Enchantment (Charm) [Mind-Affecting]

\textbf{Level:} Clr 6, Sor/Wiz 6

\textbf{Saving Throw: }Will negates

This spell functions like \textit{symbol of death}, except that all creatures within 
the radius of a \textit{symbol of persuasion }instead become \textit{charmed }by 
the caster (as the \textit{charm monster }spell) for 1 hour per caster level.

Unlike \textit{symbol of death}, \textit{symbol of persuasion }has no hit point 
limit; once triggered, a \textit{symbol of persuasion }simply remains active for 
10 minutes per caster level.

\textit{Note: }Magic traps such as \textit{symbol of persuasion }are hard to detect 
and disable. A rogue (only) can use the Search skill to find a \textit{symbol of 
persuasion }and Disable Device to thwart it. The DC in each case is 25 + spell 
level, or 31 for \textit{symbol of persuasion}.

\textit{Material Component: }Mercury and phosphorus, plus powdered diamond and 
opal with a total value of at least 5,000 gp.

\vspace{12pt}
Symbol of Sleep

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Clr 5, Sor/Wiz 5

\textbf{Saving Throw: }Will negates

This spell functions like \textit{symbol of death}, except that all creatures of 
10 HD or less within 60 feet of the \textit{symbol of sleep }instead fall into 
a catatonic slumber for 3d6x10 minutes. Unlike with the \textit{sleep }spell, sleeping 
creatures cannot be awakened by nonmagical means before this time expires.

Unlike \textit{symbol of death}, \textit{symbol of sleep }has no hit point limit; 
once triggered, a \textit{symbol of sleep }simply remains active for 10 minutes 
per caster level.

\textit{Note: }Magic traps such as \textit{symbol of sleep }are hard to detect 
and disable. A rogue (only) can use the Search skill to find a \textit{symbol of 
sleep }and Disable Device to thwart it. The DC in each case is 25 + spell level, 
or 30 for \textit{symbol of sleep}.

\textit{Material Component: }Mercury and phosphorus, plus powdered diamond and 
opal with a total value of at least 1,000 gp.

\vspace{12pt}
Symbol of Stunning

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Clr 7, Sor/Wiz 7

\textbf{Saving Throw: }Will negates

This spell functions like \textit{symbol of death}, except that all creatures within 
60 feet of a \textit{symbol of stunning }instead become stunned for 1d6 rounds.

\textit{Note: }Magic traps such as \textit{symbol of stunning }are hard to detect 
and disable. A rogue (only) can use the Search skill to find a \textit{symbol of 
stunning }and Disable Device to thwart it. The DC in each case is 25 + spell level, 
or 32 for \textit{symbol of stunning}.

\textit{Material Component: }Mercury and phosphorus, plus powdered diamond and 
opal with a total value of at least 5,000 gp.

\vspace{12pt}
Symbol of Weakness

Necromancy

\textbf{Level:} Clr 7, Sor/Wiz 7

This spell functions like \textit{symbol of death}, except that every creature 
within 60 feet of a \textit{symbol of weakness }instead suffers crippling weakness 
that deals 3d6 points of Strength damage.

Unlike \textit{symbol of death}, \textit{symbol of weakness }has no hit point limit; 
once triggered, a \textit{symbol of weakness }simply remains active for 10 minutes 
per caster level.

\textit{Note: }Magic traps such as \textit{symbol of weakness }are hard to detect 
and disable. A rogue (only) can use the Search skill to find a \textit{symbol of 
weakness }and Disable Device to thwart it. The DC in each case is 25 + spell level, 
or 32 for \textit{symbol of weakness}.

\textit{Material Component: }Mercury and phosphorus, plus powdered diamond and 
opal with a total value of at least 5,000 gp.

\vspace{12pt}
Sympathetic Vibration

Evocation [Sonic]

\textbf{Level:} Brd 6

\textbf{Components:} V, S, F

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Touch

\textbf{Target:} One freestanding structure

\textbf{Duration:} Up to 1 round/level

\textbf{Saving Throw:} None; see text

\textbf{Spell Resistance:} Yes

By attuning yourself to a freestanding structure such you can create a damaging 
vibration within it. Once it begins, the vibration deals 2d10 points of damage 
per round to the target structure. (Hardness has no effect on the spell's damage.) 
You can choose at the time of casting to limit the duration of the spell; otherwise 
it lasts for 1 round/ level. If the spell is cast upon a target that is not freestanding 
the surrounding stone dissipates the effect and no damage occurs.

\textit{Sympathetic vibration }cannot affect creatures (including constructs). 
Since a structure is an unattended object, it gets no saving throw to resist the 
effect.

\textit{Focus: }A tuning fork.

\vspace{12pt}
Sympathy

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Drd 9, Sor/Wiz 8

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 hour

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One location (up to a 10-ft. cube/level) or one object

\textbf{Duration:} 2 hours/level (D)

\textbf{Saving Throw: }Will negates; see text

\textbf{Spell Resistance:} Yes

You cause an object or location to emanate magical vibrations that attract either 
a specific kind of intelligent creature or creatures of a particular alignment, 
as defined by you. The particular kind of creature to be affected must be named 
specifically. A creature subtype is not specific enough. Likewise, the specific 
alignment must be named.

Creatures of the specified kind or alignment feel elated and pleased to be in the 
area or desire to touch or to possess the object. The compulsion to stay in the 
area or touch the object is overpowering. If the save is successful, the creature 
is released from the enchantment, but a subsequent save must be made 1d6x10 minutes 
later. If this save fails, the affected creature attempts to return to the area 
or object.

\textit{Sympathy }counters and dispels \textit{antipathy}.

\textit{Material Component: }1,500 gp worth of crushed pearls and a drop of honey.

\newpage

\end{document}
