%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{array}
\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

This material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

{\LARGE{}SPELLS (H-L)}

\vspace{12pt}
Hallow

Evocation [Good]

\textbf{Level:} Clr 5, Drd 5

\textbf{Components:} V, S, M, DF

\textbf{Casting Time:} 24 hours

\textbf{Range:} Touch

\textbf{Area:} 40-ft. radius emanating from the touched point

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} See text

\textbf{Spell Resistance:} See text

\textit{Hallow }makes a particular site, building, or structure a holy site. This 
has four major effects.

First, the site or structure is guarded by a \textit{magic circle against evil 
}effect.

Second, all Charisma checks made to turn undead gain a +4 sacred bonus, and Charisma 
checks to command undead take a -4 penalty. Spell resistance does not apply to 
this effect. (This provision does not apply to the druid version of the spell.)

Third, any dead body interred in a \textit{hallowed }site cannot be turned into 
an undead creature.

Finally, you may choose to fix a single spell effect to the \textit{hallowed }site. 
The spell effect lasts for one year and functions throughout the entire site, regardless 
of the normal duration and area or effect. You may designate whether the effect 
applies to all creatures, creatures who share your faith or alignment, or creatures 
who adhere to another faith or alignment. At the end of the year, the chosen effect 
lapses, but it can be renewed or replaced simply by casting \textit{hallow }again.

Spell effects that may be tied to a \textit{hallowed }site include \textit{aid, 
bane, bless, cause fear, darkness, daylight, death ward, deeper darkness, detect 
evil, detect magic, dimensional anchor, discern lies, dispel magic, endure elements, 
freedom of movement, invisibility purge, protection from energy, remove fear, resist 
energy, silence, tongues, }and \textit{zone of truth}. Saving throws and spell 
resistance might apply to these spells' effects. (See the individual spell descriptions 
for details.)

An area can receive only one \textit{hallow }spell (and its associated spell effect) 
at a time. \textit{Hallow }counters but does not dispel \textit{unhallow}.

\textit{Material Component: }Herbs, oils, and incense worth at least 1,000 gp, 
plus 1,000 gp per level of the spell to be included in the \textit{hallowed }area.

\vspace{12pt}
Hallucinatory Terrain

Illusion (Glamer)

\textbf{Level:} Brd 4, Sor/Wiz 4

\textbf{Components:} V, S, M

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Area:} One 30-ft. cube/level (S)

\textbf{Duration:} 2 hours/level (D)

\textbf{Saving Throw: }Will disbelief (if interacted with)

\textbf{Spell Resistance:} No

You make natural terrain look, sound, and smell like some other sort of natural 
terrain. Structures, equipment, and creatures within the area are not hidden or 
changed in appearance.

\textit{Material Component: }A stone, a twig, and a bit of green plant.

\vspace{12pt}
Halt Undead

Necromancy

\textbf{Level:} Sor/Wiz 3

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Targets:} Up to three undead creatures, no two of which can be more than 
30 ft. apart

\textbf{Duration:} 1 round/level

\textbf{Saving Throw: }Will negates (see text)

\textbf{Spell Resistance:} Yes

This spell renders as many as three undead creatures immobile. A nonintelligent 
undead creature gets no saving throw; an intelligent undead creature does. If the 
spell is successful, it renders the undead creature immobile for the duration of 
the spell (similar to the effect of \textit{hold person }on a living creature). 
The effect is broken if the \textit{halted }creatures are attacked or take damage.

\textit{Material Component: }A pinch of sulfur and powdered garlic.

\vspace{12pt}
Harm

Necromancy

\textbf{Level:} Clr 6, Destruction 6

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will half; see text

\textbf{Spell Resistance:} Yes

\textit{Harm }charges a subject with negative energy that deals 10 points of damage 
per caster level (to a maximum of 150 points at 15th level). If the creature successfully 
saves, \textit{harm }deals half this amount, but it cannot reduce the target's 
hit points to less than 1.

If used on an undead creature, \textit{harm }acts like \textit{heal}.

\vspace{12pt}
Haste

Transmutation

\textbf{Level:} Brd 3, Sor/Wiz 3

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} One creature/level, no two of which can be more than 30 ft. apart

\textbf{Duration:} 1 round/level

\textbf{Saving Throw:} Fortitude negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

The transmuted creatures move and act more quickly than normal. This extra speed 
has several effects.

When making a full attack action, a hasted creature may make one extra attack with 
any weapon he is holding. The attack is made using the creature's full base attack 
bonus, plus any modifiers appropriate to the situation. (This effect is not cumulative 
with similar effects, such as that provided by a weapon of speed, nor does it actually 
grant an extra action, so you can't use it to cast a second spell or otherwise 
take an extra action in the round.)

A \textit{hasted }creature gains a +1 bonus on attack rolls and a +1 dodge bonus 
to AC and Reflex saves. Any condition that makes you lose your Dexterity bonus 
to Armor Class (if any) also makes you lose dodge bonuses.

All of the \textit{hasted }creature's modes of movement (including land movement, 
burrow, climb, fly, and swim) increase by 30 feet, to a maximum of twice the subject's 
normal speed using that form of movement. This increase counts as an enhancement 
bonus, and it affects the creature's jumping distance as normal for increased speed.

Multiple \textit{haste }effects don't stack. \textit{Haste }dispels and counters 
\textit{slow.}

\textit{Material Component: }A shaving of licorice root.

\vspace{12pt}
Heal

Conjuration (Healing)

\textbf{Level:} Clr 6, Drd 7, Healing 6

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

\textit{Heal }enables you to channel positive energy into a creature to wipe away 
injury and afflictions. It immediately ends any and all of the following adverse 
conditions affecting the Target: ability damage, blinded, \textit{confused, }dazed, 
dazzled, deafened, diseased, exhausted, fatigued, \textit{feebleminded, }insanity, 
nauseated, sickened, stunned, and poisoned. It also cures 10 hit points of damage 
per level of the caster, to a maximum of 150 points at 15th level.

\textit{Heal }does not remove negative levels, restore permanently drained levels, 
or restore permanently drained ability score points.

If used against an undead creature, \textit{heal }instead acts like \textit{harm}.

\vspace{12pt}
Heal, Mass

Conjuration (Healing)

\textbf{Level:} Clr 9, Healing 9

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} One or more creatures, no two of which can be more than 30 ft. 
apart

This spell functions like \textit{heal, }except as noted above. The maximum number 
of hit points restored to each creature is 250.

\vspace{12pt}
Heal Mount

Conjuration (Healing)

\textbf{Level:} Pal 3

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Your mount touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

This spell functions like \textit{heal, }but it affects only the paladin's special 
mount (typically a warhorse).

\vspace{12pt}
Heat Metal

Transmutation [Fire]

\textbf{Level:} Drd 2, Sun 2

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target: }Metal equipment of one creature per two levels, no two of which 
can be more than 30 ft. apart; or 25 lb. of metal/level, all of which must be within 
a 30-ft. circle

\textbf{Duration:} 7 rounds

\textbf{Saving Throw: }Will negates (object)

\textbf{Spell Resistance:} Yes (object)

\textit{Heat metal }makes metal extremely warm. Unattended, nonmagical metal gets 
no saving throw. Magical metal is allowed a saving throw against the spell. An 
item in a creature's possession uses the creature's saving throw bonus unless its 
own is higher.

A creature takes fire damage if its equipment is heated. It takes full damage if 
its armor is affected or if it is holding, touching, wearing, or carrying metal 
weighing one-fifth of its weight. The creature takes minimum damage (1 point or 
2 points; see the table) if it's not wearing metal armor and the metal that it's 
carrying weighs less than one-fifth of its weight.

On the first round of the spell, the metal becomes warm and uncomfortable to touch 
but deals no damage. The same effect also occurs on the last round of the spell's 
duration. During the second (and also the next-to-last) round, intense heat causes 
pain and damage. In the third, fourth, and fifth rounds, the metal is searing hot, 
causing more damage, as shown on the table below.

\begin{tabular}{|>{\raggedright}p{27pt}|>{\raggedright}p{82pt}|>{\raggedright}p{40pt}|}
\hline
R\textbf{ound} & M\textbf{etal Temperature} & D\textbf{amage}\tabularnewline
\hline
1 & Warm & None\tabularnewline
\hline
2 & Hot & 1d4 points\tabularnewline
\hline
3-5 & Searing & 2d4 points\tabularnewline
\hline
6 & Hot & 1d4 points\tabularnewline
\hline
7 & Warm & None\tabularnewline
\hline
\end{tabular}

Any cold intense enough to damage the creature negates fire damage from the spell 
(and vice versa) on a point-for-point basis. If cast underwater, \textit{heat metal 
}deals half damage and boils the surrounding water.

\textit{Heat metal }counters and dispels \textit{chill metal}.

\vspace{12pt}
Helping Hand

Evocation

\textbf{Level:} Clr 3

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 5 miles

\textbf{Effect:} Ghostly hand

\textbf{Duration:} 1 hour/level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You create the ghostly image of a hand, which you can send to find a creature within 
5 miles. The hand then beckons to that creature and leads it to you if the creature 
is willing to follow.

When the spell is cast, the hand appears in front of you. You then specify a person 
(or any creature) by physical description, which can include race, gender, and 
appearance but not ambiguous factors such as level, alignment, or class. When the 
description is complete, the hand streaks off in search of a subject that fits 
the description. The amount of time it takes to find the subject depends on how 
far away she is.

\begin{tabular}{|>{\raggedright}p{53pt}|>{\raggedright}p{63pt}|}
\hline
D\textbf{istance} & T\textbf{ime to Locate}\tabularnewline
\hline
100 ft. or less & 1 round\tabularnewline
\hline
1,000 ft. & 1 minute\tabularnewline
\hline
1 mile & 10 minutes\tabularnewline
\hline
2 miles & 1 hour\tabularnewline
\hline
3 miles & 2 hours\tabularnewline
\hline
4 miles & 3 hours\tabularnewline
\hline
5 miles & 4 hours\tabularnewline
\hline
\end{tabular}

Once the hand locates the subject, it beckons the creature to follow it. If the 
subject does so, the hand points in your direction, indicating the most direct 
feasible route. The hand hovers 10 feet in front of the subject, moving before 
it at a speed of as much as 240 feet per round. Once the hand leads the subject 
back to you, it disappears.

The subject is not compelled to follow the hand or act in any particular way toward 
you. If the subject chooses not to follow, the hand continues to beckon for the 
duration of the spell, then disappears. If the spell expires while the subject 
is en route to you, the hand disappears; the subject must then rely on her own 
devices to locate you.

If more than one subject in a 5-mile radius meets the description, the hand locates 
the closest creature. If that creature refuses to follow the hand, the hand does 
not seek out a second subject.

If, at the end of 4 hours of searching, the hand has found no subject that matches 
the description within 5 miles, it returns to you, displays an outstretched palm 
(indicating that no such creature was found), and disappears.

The ghostly hand has no physical form. It is invisible to anyone except you and 
a potential subject. It cannot engage in combat or execute any other task aside 
from locating a subject and leading it back to you. The hand can't pass through 
solid objects but can ooze through small cracks and slits. The hand cannot travel 
more than 5 miles from the spot it appeared when you cast the spell.

\vspace{12pt}
Heroes' Feast

Conjuration [Creation]

\textbf{Level:} Brd 6, Clr 6

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} Feast for one creature/level

\textbf{Duration:} 1 hour plus 12 hours; see text

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You bring forth a great feast, including a magnificent table, chairs, service, 
and food and drink. The feast takes 1 hour to consume, and the beneficial effects 
do not set in until this hour is over. Every creature partaking of the feast is 
cured of all diseases, sickness, and nausea; becomes immune to poison for 12 hours; 
and gains 1d8 temporary hit points +1 point per two caster levels (maximum +10) 
after imbibing the nectar-like beverage that is part of the feast. The ambrosial 
food that is consumed grants each creature that partakes a +1 morale bonus on attack 
rolls and Will saves and immunity to fear effects for 12 hours.

If the feast is interrupted for any reason, the spell is ruined and all effects 
of the spell are negated.

\vspace{12pt}
Heroism

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Brd 2, Sor/Wiz 3

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 10 min./level

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

This spell imbues a single creature with great bravery and morale in battle. The 
target gains a +2 morale bonus on attack rolls, saves, and skill checks.

\vspace{12pt}
Heroism, Greater

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Brd 5, Sor/Wiz 6

\textbf{Duration:} 1 min./level

This spell functions like \textit{heroism, }except the creature gains a +4 morale 
bonus on attack rolls, saves, and skill checks, immunity to fear effects, and temporary 
hit points equal to your caster level (maximum 20).

\vspace{12pt}
Hide from Animals

Abjuration

\textbf{Level:} Drd 1, Rgr 1

\textbf{Components:} S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Targets:} One creature touched/level

\textbf{Duration:} 10 min./level (D)

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes

Animals cannot see, hear, or smell the warded creatures. Even extraordinary or 
supernatural sensory capabilities, such as blindsense, blindsight, scent, and tremorsense, 
cannot detect or locate warded creatures. Animals simply act as though the warded 
creatures are not there. If a warded character touches an animal or attacks any 
creature, even with a spell, the spell ends for all recipients.

\vspace{12pt}
Hide from Undead

Abjuration

\textbf{Level:} Clr 1

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Targets:} One touched creature/level

\textbf{Duration:} 10 min./level (D)

\textbf{Saving Throw: }Will negates (harmless); see text

\textbf{Spell Resistance:} Yes

Undead cannot see, hear, or smell the warded creatures. Even extraordinary or supernatural 
sensory capabilities, such as blindsense, blindsight, scent, and tremorsense, cannot 
detect or locate warded creatures. Nonintelligent undead creatures are automatically 
affected and act as though the warded creatures are not there. An intelligent undead 
creature gets a single Will saving throw. If it fails, the subject can't see any 
of the warded creatures. However, if it has reason to believe unseen opponents 
are present, it can attempt to find or strike them. If a warded creature attempts 
to turn or command undead, touches an undead creature, or attacks any creature 
(even with a spell), the spell ends for all recipients.

\vspace{12pt}
Hideous Laughter

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Brd 1, Sor/Wiz 2

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One creature; see text

\textbf{Duration:} 1 round/level

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

This spell afflicts the subject with uncontrollable laughter. It collapses into 
gales of manic laughter, falling prone. The subject can take no actions while laughing, 
but is not considered helpless. After the spell ends, it can act normally.

A creature with an Intelligence score of 2 or lower is not affected. A creature 
whose type is different from the caster's receives a +4 bonus on its saving throw, 
because humor doesn't ``translate'' well.

\textit{Material Component: }Tiny tarts that are thrown at the target and a feather 
that is waved in the air.

\vspace{12pt}
Hold Animal

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Animal 2, Drd 2, Rgr 2

\textbf{Components:} V, S

\textbf{Target:} One animal

This spell functions like \textit{hold person, }except that it affects an animal 
instead of a humanoid.

\vspace{12pt}
Hold Monster

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Brd 4, Law 6, Sor/Wiz 5

\textbf{Components:} V, S, M/DF

\textbf{Target:} One living creature

This spell functions like \textit{hold person, }except that it affects any living 
creature that fails its Will save.

\textit{Arcane Material Component: }One hard metal bar or rod, which can be as 
small as a three-penny nail.

\vspace{12pt}
Hold Monster, Mass

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Sor/Wiz 9

\textbf{Targets:} One or more creatures, no two of which can be more than 30 ft. 
apart

This spell functions like \textit{hold person, }except that it affects multiple 
creatures and holds any living creature that fails its Will save.

\vspace{12pt}
Hold Person

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Brd 2, Clr 2, Sor/Wiz 3

\textbf{Components:} V, S, F/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Target:} One humanoid creature

\textbf{Duration:} 1 round/level (D); see text

\textbf{Saving Throw: }Will negates; see text

\textbf{Spell Resistance:} Yes

The subject becomes paralyzed and freezes in place. It is aware and breathes normally 
but cannot take any actions, even speech. Each round on its turn, the subject may 
attempt a new saving throw to end the effect. (This is a full-round action that 
does not provoke attacks of opportunity.)

A winged creature who is paralyzed cannot flap its wings and falls. A swimmer can't 
swim and may drown.

\textit{Arcane Focus: }A small, straight piece of iron.

\vspace{12pt}
Hold Person, Mass

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Sor/Wiz 7

\textbf{Targets:} One or more humanoid creatures, no two of which can be more than 
30 ft. apart

This spell functions like \textit{hold person, }except as noted above.

\vspace{12pt}
Hold Portal

Abjuration

\textbf{Level:} Sor/Wiz 1

\textbf{Component:} V

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Target:} One portal, up to 20 sq. ft./level

\textbf{Duration:} 1 min./level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell magically holds shut a door, gate, window, or shutter of wood, metal, 
or stone. The magic affects the portal just as if it were securely closed and normally 
locked. A \textit{knock }spell or a successful \textit{dispel magic }spell can 
negate a \textit{hold portal }spell.

For a portal affected by this spell, add 5 to the normal DC for forcing open the 
portal.

\vspace{12pt}
Holy Aura

Abjuration [Good]

\textbf{Level:} Clr 8, Good 8

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 20 ft.

\textbf{Targets:} One creature/level in a 20-ft.-radius burst centered on you

\textbf{Duration:} 1 round/level (D)

\textbf{Saving Throw:} See text

\textbf{Spell Resistance:} Yes (harmless)

A brilliant divine radiance surrounds the subjects, protecting them from attacks, 
granting them resistance to spells cast by evil creatures, and causing evil creatures 
to become blinded when they strike the subjects. This abjuration has four effects.

First, each warded creature gains a +4 deflection bonus to AC and a +4 resistance 
bonus on saves. Unlike \textit{protection from evil}, this benefit applies against 
all attacks, not just against attacks by evil creatures.

Second, each warded creature gains spell resistance 25 against evil spells and 
spells cast by evil creatures.

Third, the abjuration blocks possession and mental influence, just as \textit{protection 
from evil }does.

Finally, if an evil creature succeeds on a melee attack against a warded creature, 
the offending attacker is blinded (Fortitude save negates, as \textit{blindness/deafness}, 
but against \textit{holy aura's }save DC).

\textit{Focus: }A tiny reliquary containing some sacred relic. The reliquary costs 
at least 500 gp.

\vspace{12pt}
Holy Smite

Evocation [Good]

\textbf{Level:} Good 4

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Area:} 20-ft.-radius burst

\textbf{Duration:} Instantaneous (1 round); see text

\textbf{Saving Throw: }Will partial; see text

\textbf{Spell Resistance:} Yes

You draw down holy power to smite your enemies. Only evil and neutral creatures 
are harmed by the spell; good creatures are unaffected.

The spell deals 1d8 points of damage per two caster levels (maximum 5d8) to each 
evil creature in the area (or 1d6 points of damage per caster level, maximum 10d6, 
to an evil outsider) and causes it to become blinded for 1 round. A successful 
Will saving throw reduces damage to half and negates the blinded effect.

The spell deals only half damage to creatures who are neither good nor evil, and 
they are not blinded. Such a creature can reduce that damage by half (down to one-quarter 
of the roll) with a successful Will save.

\vspace{12pt}
Holy Sword

Evocation [Good]

\textbf{Level:} Pal 4

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target: }Melee weapon touched

\textbf{Duration:} 1 round/level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell allows you to channel holy power into your sword, or any other melee 
weapon you choose. The weapon acts as a \textit{+5 holy weapon }(+5 enhancement 
bonus on attack and damage rolls, extra 2d6 damage against evil opponents). It 
also emits a \textit{magic circle against evil }effect (as the spell). If the \textit{magic 
circle }ends, the sword creates a new one on your turn as a free action. The spell 
is automatically canceled 1 round after the weapon leaves your hand. You cannot 
have more than one \textit{holy sword }at a time.

If this spell is cast on a magic weapon, the powers of the spell supersede any 
that the weapon normally has, rendering the normal enhancement bonus and powers 
of the weapon inoperative for the duration of the spell. This spell is not cumulative 
with \textit{bless weapon }or any other spell that might modify the weapon in any 
way.

This spell does not work on artifacts.

\textit{Note: }A masterwork weapon's bonus to attack does not stack with an enhancement 
bonus to attack.

\vspace{12pt}
Holy Word

Evocation [Good, Sonic]

\textbf{Level:} Clr 7, Good 7

\textbf{Components:} V

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 40 ft.

\textbf{Area:} Nongood creatures in a 40-ft.-radius spread centered on you

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None or Will negates; see text

\textbf{Spell Resistance:} Yes

Any nongood creature within the area that hears the \textit{holy word }suffers 
the following ill effects.

\begin{tabular}{|>{\raggedright}p{80pt}|>{\raggedright}p{161pt}|}
\hline
H\textbf{D} & E\textbf{ffect}\tabularnewline
\hline
Equal to caster level & Deafened\tabularnewline
\hline
Up to caster level- & 1 Blinded, deafened\tabularnewline
\hline
Up to caster level- & 5 Paralyzed, blinded, deafened\tabularnewline
\hline
Up to caster level- & 10 Killed, paralyzed, blinded, deafened\tabularnewline
\hline
\end{tabular}

The effects are cumulative and concurrent. No saving throw is allowed against these 
effects.

\textit{Deafened: }The creature is deafened for 1d4 rounds.

\textit{Blinded: }The creature is blinded for 2d4 rounds.

\textit{Paralyzed: }The creature is paralyzed and helpless for 1d10 minutes.

\textit{Killed: }Living creatures die. Undead creatures are destroyed.

Furthermore, if you are on your home plane when you cast this spell, nongood extraplanar 
creatures within the area are instantly banished back to their home planes. Creatures 
so banished cannot return for at least 24 hours. This effect takes place regardless 
of whether the creatures hear the \textit{holy word. }The banishment effect allows 
a Will save (at a -4 penalty) to negate.

Creatures whose HD exceed your caster level are unaffected by \textit{holy word.}

\vspace{12pt}
Horrid Wilting

Necromancy

\textbf{Level:} Sor/Wiz 8, Water 8

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Targets:} Living creatures, no two of which can be more than 60 ft. apart

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} Fortitude half

\textbf{Spell Resistance:} Yes

This spell evaporates moisture from the body of each subject living creature, dealing 
1d6 points of damage per caster level (maximum 20d6). This spell is especially 
devastating to water elementals and plant creatures, which instead take 1d8 points 
of damage per caster level (maximum 20d8).

\textit{Arcane Material Component: }A bit of sponge.

\vspace{12pt}
Hypnotic Pattern

Illusion (Pattern) [Mind-Affecting]

\textbf{Level:} Brd 2, Sor/Wiz 2

\textbf{Components:} V (Brd only), S, M; see text

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Effect:} Colorful lights in a 10-ft.-radius spread

\textbf{Duration:} Concentration + 2 rounds

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

A twisting pattern of subtle, shifting colors weaves through the air, fascinating 
creatures within it. Roll 2d4 and add your caster level (maximum 10) to determine 
the total number of Hit Dice of creatures affected. Creatures with the fewest HD 
are affected first; and, among creatures with equal HD, those who are closest to 
the spell's point of origin are affected first. Hit Dice that are not sufficient 
to affect a creature are wasted. Affected creatures become fascinated by the pattern 
of colors. Sightless creatures are not affected.

A wizard or sorcerer need not utter a sound to cast this spell, but a bard must 
sing, play music, or recite a rhyme as a verbal component.

\textit{Material Component: }A glowing stick of incense or a crystal rod filled 
with phosphorescent material.

\vspace{12pt}
Hypnotism

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Brd 1, Sor/Wiz 1

\textbf{Components:} V, S

\textbf{Casting Time:} 1 round

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Area:} Several living creatures, no two of which may be more than 30 ft. 
apart

\textbf{Duration:} 2d4 rounds (D)

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

Your gestures and droning incantation fascinate nearby creatures, causing them 
to stop and stare blankly at you. In addition, you can use their rapt attention 
to make your suggestions and requests seem more plausible. Roll 2d4 to see how 
many total Hit Dice of creatures you affect. Creatures with fewer HD are affected 
before creatures with more HD. Only creatures that can see or hear you are affected, 
but they do not need to understand you to be fascinated.

If you use this spell in combat, each target gains a +2 bonus on its saving throw. 
If the spell affects only a single creature not in combat at the time, the saving 
throw has a penalty of -2.

While the subject is fascinated by this spell, it reacts as though it were two 
steps more friendly in attitude. This allows you to make a single request of the 
affected creature (provided you can communicate with it). The request must be brief 
and reasonable. Even after the spell ends, the creature retains its new attitude 
toward you, but only with respect to that particular request.

A creature that fails its saving throw does not remember that you enspelled it.

\vspace{12pt}
Ice Storm

Evocation [Cold]

\textbf{Level:} Drd 4, Sor/Wiz 4, Water 5

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Area:} Cylinder (20-ft. radius, 40 ft. high)

\textbf{Duration:} 1 full round

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

Great magical hailstones pound down for 1 full round, dealing 3d6 points of bludgeoning 
damage and 2d6 points of cold damage to every creature in the area. A -4 penalty 
applies to each Listen check made within the \textit{ice storm'}s effect, and all 
land movement within its area is at half speed. At the end of the duration, the 
hail disappears, leaving no aftereffects (other than the damage dealt).

\textit{Arcane Material Component: }A pinch of dust and a few drops of water.

\vspace{12pt}
Identify

Divination

\textbf{Level:} Brd 1, Magic 2, Sor/Wiz 1

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 hour

\textbf{Range:} Touch

\textbf{Targets:} One touched object

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

The spell determines all magic properties of a single magic item, including how 
to activate those functions (if appropriate), and how many charges are left (if 
any).

\textit{Identify }does not function when used on an artifact.

\textit{Arcane Material Component: }A pearl of at least 100 gp value, crushed and 
stirred into wine with an owl feather; the infusion must be drunk prior to spellcasting.

\vspace{12pt}
Illusory Script

Illusion (Phantasm) [Mind-Affecting]

\textbf{Level:} Brd 3, Sor/Wiz 3

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 minute or longer; see text

\textbf{Range:} Touch

\textbf{Target:} One touched object weighing no more than 10 lb.

\textbf{Duration:} One day/level (D)

\textbf{Saving Throw: }Will negates; see text

\textbf{Spell Resistance:} Yes

You write instructions or other information on parchment, paper, or any suitable 
writing material. The \textit{illusory script }appears to be some form of foreign 
or magical writing. Only the person (or people) designated by you at the time of 
the casting are able to read the writing; it's unintelligible to any other character, 
although an illusionist recognizes it as \textit{illusory script.}

Any unauthorized creature attempting to read the script triggers a potent illusory 
effect and must make a saving throw. A successful saving throw means the creature 
can look away with only a mild sense of disorientation. Failure means the creature 
is subject to a suggestion implanted in the script by you at the time the illusory 
script spell was cast. The suggestion lasts only 30 minutes. Typical suggestions 
include ``Close the book and leave,'' ``Forget the existence of the book,'' and 
so forth. If successfully dispelled by \textit{dispel magic, }the \textit{illusory 
script }and its secret message disappear. The hidden message can be read by a combination 
of the \textit{true seeing }spell with the \textit{read magic }or \textit{comprehend 
languages }spell.

The casting time depends on how long a message you wish to write, but it is always 
at least 1 minute.

\textit{Material Component: }A lead-based ink (cost of not less than 50 gp).

\vspace{12pt}
Illusory Wall

Illusion (Figment)

\textbf{Level:} Sor/Wiz 4

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} Image 1 ft. by 10 ft. by 10 ft.

\textbf{Duration:} Permanent

\textbf{Saving Throw: }Will disbelief (if interacted with)

\textbf{Spell Resistance:} No

This spell creates the illusion of a wall, floor, ceiling, or similar surface. 
It appears absolutely real when viewed, but physical objects can pass through it 
without difficulty. When the spell is used to hide pits, traps, or normal doors, 
any detection abilities that do not require sight work normally. Touch or a probing 
search reveals the true nature of the surface, though such measures do not cause 
the illusion to disappear.

\vspace{12pt}
Imbue with Spell Ability

Evocation

\textbf{Level:} Clr 4, Magic 4

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Touch

\textbf{Target:} Creature touched; see text

\textbf{Duration:} Permanent until discharged (D)

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

You transfer some of your currently prepared spells, and the ability to cast them, 
to another creature. Only a creature with an Intelligence score of at least 5 and 
a Wisdom score of at least 9 can receive this bestowal. Only cleric spells from 
the schools of abjuration, divination, and conjuration (healing) can be transferred. 
The number and level of spells that the subject can be granted depends on its Hit 
Dice; even multiple castings of \textit{imbue with spell ability }can't exceed 
this limit.

\begin{tabular}{|>{\raggedright}p{67pt}|>{\raggedright}p{201pt}|}
\hline
H\textbf{D of Recipient} & S\textbf{pells Imbued}\tabularnewline
\hline
2 or lower & One 1st-level spell\tabularnewline
\hline
3-4 & One or two 1st-level spells\tabularnewline
\hline
5 or higher & One or two 1st-level spells and one 2nd-level spell\tabularnewline
\hline
\end{tabular}

The transferred spell's variable characteristics (range, duration, area, and the 
like) function according to your level, not the level of the recipient.

Once you cast \textit{imbue with spell ability, }you cannot prepare a new 4th-level 
spell to replace it until the recipient uses the imbued spells or is slain, or 
until you dismiss the \textit{imbue with spell ability }spell. In the meantime, 
you remain responsible to your deity or your principles for the use to which the 
spell is put. If the number of 4th-level spells you can cast decreases, and that 
number drops below your current number of active \textit{imbue with spell ability 
}spells, the more recently cast imbued spells are dispelled.

To cast a spell with a verbal component, the subject must be able to speak. To 
cast a spell with a somatic component, it must have humanlike hands. To cast a 
spell with a material component or focus, it must have the materials or focus.

\vspace{12pt}
Implosion

Evocation

\textbf{Level:} Clr 9, Destruction 9

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} One corporeal creature/round

\textbf{Duration:} Concentration (up to 4 rounds)

\textbf{Saving Throw:} Fortitude negates

\textbf{Spell Resistance:} Yes

You create a destructive resonance in a corporeal creature's body. For each round 
you concentrate, you cause one creature to collapse in on itself, killing it. (This 
effect, being instantaneous, cannot be dispelled.)

You can target a particular creature only once with each casting of the spell.

\textit{Implosion }has no effect on creatures in gaseous form or on incorporeal 
creatures.

\vspace{12pt}
Imprisonment

Abjuration

\textbf{Level:} Sor/Wiz 9

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates; see text

\textbf{Spell Resistance:} Yes

When you cast \textit{imprisonment }and touch a creature, it is entombed in a state 
of suspended animation (see the \textit{temporal stasis }spell) in a small sphere 
far beneath the surface of the earth. The subject remains there unless a \textit{freedom 
}spell is cast at the locale where the imprisonment took place. Magical search 
by a \textit{crystal ball, }a \textit{locate object }spell, or some other similar 
divination does not reveal the fact that a creature is imprisoned, but \textit{discern 
location }does. A \textit{wish }or \textit{miracle }spell will not free the recipient, 
but will reveal where it is entombed. If you know the target's name and some facts 
about its life, the target takes a -4 penalty on its save.

\vspace{12pt}
Incendiary Cloud

Conjuration (Creation) [Fire]

\textbf{Level:} Fire 8, Sor/Wiz 8

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Effect:} Cloud spreads in 20-ft. radius, 20 ft. high

\textbf{Duration:} 1 round/level

\textbf{Saving Throw:} Reflex half; see text

\textbf{Spell Resistance:} No

An \textit{incendiary cloud }spell creates a cloud of roiling smoke shot through 
with white-hot embers. The smoke obscures all sight as a \textit{fog cloud }does. 
In addition, the white-hot embers within the cloud deal 4d6 points of fire damage 
to everything within the cloud on your turn each round. All targets can make Reflex 
saves each round to take half damage.

As with a \textit{cloudkill }spell, the smoke moves away from you at 10 feet per 
round. Figure out the smoke's new spread each round based on its new point of origin, 
which is 10 feet farther away from where you were when you cast the spell. By concentrating, 
you can make the cloud (actually its point of origin) move as much as 60 feet each 
round. Any portion of the cloud that would extend beyond your maximum range dissipates 
harmlessly, reducing the remainder's spread thereafter.

As with \textit{fog cloud}, wind disperses the smoke, and the spell can't be cast 
underwater.

\vspace{12pt}
Inflict Critical Wounds

Necromancy

\textbf{Level:} Clr 4, Destruction 4

This spell functions like \textit{inflict light wounds}, except that you deal 4d8 
points of damage +1 point per caster level (maximum +20).

\vspace{12pt}
Inflict Critical Wounds, Mass

Necromancy

\textbf{Level:} Clr 8

This spell functions like \textit{mass inflict light wounds}, except that it deals 
4d8 points of damage +1 point per caster level (maximum +40).

\vspace{12pt}
Inflict Light Wounds

Necromancy

\textbf{Level:} Clr 1, Destruction 1

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will half

\textbf{Spell Resistance:} Yes

When laying your hand upon a creature, you channel negative energy that deals 1d8 
points of damage +1 point per caster level (maximum +5).

Since undead are powered by negative energy, this spell cures such a creature of 
a like amount of damage, rather than harming it.

\vspace{12pt}
Inflict Light Wounds, Mass

Necromancy

\textbf{Level:} Clr 5, Destruction 5

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One creature/level, no two of which can be more than 30 ft. apart

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will half

\textbf{Spell Resistance:} Yes

Negative energy spreads out in all directions from the point of origin, dealing 
1d8 points of damage +1 point per caster level (maximum +25) to nearby living enemies.

Like other \textit{inflict }spells, \textit{mass inflict light wounds }cures undead 
in its area rather than damaging them. A cleric capable of spontaneously casting 
\textit{inflict }spells can also spontaneously cast \textit{mass inflict }spells.

\vspace{12pt}
Inflict Minor Wounds

Necromancy

\textbf{Level:} Clr 0

\textbf{Saving Throw: }Will negates

This spell functions like \textit{inflict light wounds, }except that you deal 1 
point of damage and a Will save negates the damage instead of halving it.

\vspace{12pt}
Inflict Moderate Wounds

Necromancy

\textbf{Level:} Clr 2

This spell functions like \textit{inflict light wounds}, except that you deal 2d8 
points of damage +1 point per caster level (maximum +10).

\vspace{12pt}
Inflict Moderate Wounds, Mass

Necromancy

\textbf{Level:} Clr 6

This spell functions like \textit{mass inflict light wounds}, except that it deals 
2d8 points of damage +1 point per caster level (maximum +30).

\vspace{12pt}
Inflict Serious Wounds

Necromancy

\textbf{Level:} Clr 3

This spell functions like \textit{inflict light wounds}, except that you deal 3d8 
points of damage +1 point per caster level (maximum +15).

\vspace{12pt}
Inflict Serious Wounds, Mass

Necromancy

\textbf{Level:} Clr 7

This spell functions like \textit{mass inflict light wounds}, except that it deals 
3d8 points of damage +1 point per caster level (maximum +35).

\vspace{12pt}
Insanity

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Sor/Wiz 7

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Target:} One living creature

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

The affected creature suffers from a continuous \textit{confusion }effect, as the 
spell.

\textit{Remove curse }does not remove \textit{insanity}. \textit{Greater restoration}, 
\textit{heal}, \textit{limited wish}, \textit{miracle}, or \textit{wish }can restore 
the creature.

\vspace{12pt}
Insect Plague

Conjuration (Summoning)

\textbf{Level:} Clr 5, Drd 5

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 round

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Effect:} One swarm of locusts per three levels, each of which must be adjacent 
to at least one other swarm

\textbf{Duration:} 1 min./level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You summon a number of swarms of locusts (one per three levels, to a maximum of 
six swarms at 18th level). The swarms must be summoned so that each one is adjacent 
to at least one other swarm (that is, the swarms must fill one contiguous area). 
You may summon the locust swarms so that they share the area of other creatures. 
Each swarm attacks any creatures occupying its area. The swarms are stationary 
after being summoned, and won't pursue creatures that flee.

\vspace{12pt}
Instant Summons

Conjuration (Summoning)

\textbf{Level:} Sor/Wiz 7

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} See text

\textbf{Target:} One object weighing 10 lb. or less whose longest dimension is 
6 ft. or less

\textbf{Duration:} Permanent until discharged

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You call some nonliving item from virtually any location directly to your hand.

First, you must place your \textit{arcane mark} on the item. Then you cast this 
spell, which magically and invisibly inscribes the name of the item on a sapphire 
worth at least 1,000 gp. Thereafter, you can summon the item by speaking a special 
word (set by you when the spell is cast) and crushing the gem. The item appears 
instantly in your hand. Only you can use the gem in this way.

If the item is in the possession of another creature, the spell does not work, 
but you know who the possessor is and roughly where that creature is located when 
the summons occurs.

The inscription on the gem is invisible. It is also unreadable, except by means 
of a \textit{read magic }spell, to anyone but you.

The item can be summoned from another plane, but only if no other creature has 
claimed ownership of it.

\textit{Material Component: }A sapphire worth at least 1,000 gp.

\vspace{12pt}
Interposing Hand

Evocation [Force]

\textbf{Level:} Sor/Wiz 5

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Effect:} 10-ft. hand

\textbf{Duration:} 1 round/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

\textit{Interposing hand }creates a Large magic hand that appears between you and 
one opponent. This floating, disembodied hand then moves to remain between the 
two of you, regardless of where you move or how the opponent tries to get around 
it, providing cover (+4 AC) for you against that opponent. Nothing can fool the 
hand---it sticks with the selected opponent in spite of darkness, invisibility, 
polymorphing, or any other attempt at hiding or disguise. The hand does not pursue 
an opponent, however.

An \textit{interposing hand }is 10 feet long and about that wide with its fingers 
outstretched. It has as many hit points as you do when you're undamaged, and its 
AC is 20 (-1 size, +11 natural). It takes damage as a normal creature, but most 
magical effects that don't cause damage do not affect it.

The hand never provokes attacks of opportunity from opponents. It cannot push through 
a \textit{wall of force }or enter an \textit{antimagic field}, but it suffers the 
full effect of a \textit{prismatic wall }or \textit{prismatic sphere. }The hand 
makes saving throws as its caster.

\textit{Disintegrate }or a successful \textit{dispel magic }destroys it.

Any creature weighing 2,000 pounds or less that tries to push past the hand is 
slowed to half its normal speed. The hand cannot reduce the speed of a creature 
weighing more than 2,000 pounds, but it still affects the creature's attacks.

Directing the spell to a new target is a move action.

\textit{Focus: }A soft glove.

\vspace{12pt}
Invisibility

Illusion (Glamer)

\textbf{Level:} Brd 2, Sor/Wiz 2, Trickery 2

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal or touch

\textbf{Target:} You or a creature or object weighing no more than 100 lb./level

\textbf{Duration:} 1 min./level (D)

\textbf{Saving Throw: }Will negates (harmless) or Will negates (harmless, object)

\textbf{Spell Resistance:} Yes (harmless) or Yes (harmless, object)

The creature or object touched becomes invisible, vanishing from sight, even from 
darkvision. If the recipient is a creature carrying gear, that vanishes, too. If 
you cast the spell on someone else, neither you nor your allies can see the subject, 
unless you can normally see invisible things or you employ magic to do so.

Items dropped or put down by an invisible creature become visible; items picked 
up disappear if tucked into the clothing or pouches worn by the creature. Light, 
however, never becomes invisible, although a source of light can become so (thus, 
the effect is that of a light with no visible source). Any part of an item that 
the subject carries but that extends more than 10 feet from it becomes visible.

Of course, the subject is not magically \textit{silenced, }and certain other conditions 
can render the recipient detectable (such as stepping in a puddle). The spell ends 
if the subject attacks any creature. For purposes of this spell, an attack includes 
any spell targeting a foe or whose area or effect includes a foe. (Exactly who 
is a foe depends on the invisible character's perceptions.) Actions directed at 
unattended objects do not break the spell. Causing harm indirectly is not an attack. 
Thus, an invisible being can open doors, talk, eat, climb stairs, summon monsters 
and have them attack, cut the ropes holding a rope bridge while enemies are on 
the bridge, remotely trigger traps, open a portcullis to release attack dogs, and 
so forth. If the subject attacks directly, however, it immediately becomes visible 
along with all its gear. Spells such as \textit{bless }that specifically affect 
allies but not foes are not attacks for this purpose, even when they include foes 
in their area.

\textit{Invisibility }can be made permanent (on objects only) with a \textit{permanency 
}spell.

\textit{Arcane Material Component: }An eyelash encased in a bit of gum arabic.

\vspace{12pt}
Invisibility, Greater

Illusion (Glamer)

\textbf{Level:} Brd 4, Sor/Wiz 4

\textbf{Components:} V, S

\textbf{Target:} You or creature touched

\textbf{Duration:} 1 round/level (D)

\textbf{Saving Throw: }Will negates (harmless)

This spell functions like \textit{invisibility}, except that it doesn't end if 
the subject attacks.

\vspace{12pt}
Invisibility, Mass

Illusion (Glamer)

\textbf{Level:} Sor/Wiz 7

\textbf{Components:} V, S, M

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Targets:} Any number of creatures, no two of which can be more than 180 
ft. apart

This spell functions like \textit{invisibility, }except that the effect is mobile 
with the group and is broken when anyone in the group attacks. Individuals in the 
group cannot see each other. The spell is broken for any individual who moves more 
than 180 feet from the nearest member of the group. (If only two individuals are 
affected, the one moving away from the other one loses its invisibility. If both 
are moving away from each other, they both become visible when the distance between 
them exceeds 180 feet.)

\textit{Material Component: }An eyelash encased in a bit of gum arabic.

\vspace{12pt}
Invisibility Purge

Evocation

\textbf{Level:} Clr 3

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 1 min./level (D)

You surround yourself with a sphere of power with a radius of 5 feet per caster 
level that negates all forms of invisibility.

Anything invisible becomes visible while in the area.

\vspace{12pt}
Invisibility Sphere

Illusion (Glamer)

\textbf{Level:} Brd 3, Sor/Wiz 3

\textbf{Components:} V, S, M

\textbf{Area:} 10-ft.-radius emanation around the creature or object touched

This spell functions like \textit{invisibility, }except that this spell confers 
invisibility upon all creatures within 10 feet of the recipient. The center of 
the effect is mobile with the recipient.

Those affected by this spell can see each other and themselves as if unaffected 
by the spell. Any affected creature moving out of the area becomes visible, but 
creatures moving into the area after the spell is cast do not become invisible. 
Affected creatures (other than the recipient) who attack negate the invisibility 
only for themselves. If the spell recipient attacks, the \textit{invisibility sphere 
}ends.

\vspace{12pt}
Iron Body

Transmutation

\textbf{Level:} Earth 8, Sor/Wiz 8

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 1 min./level (D)

This spell transforms your body into living iron, which grants you several powerful 
resistances and abilities.

You gain damage reduction 15/adamantine. You are immune to blindness, critical 
hits, ability score damage, deafness, disease, drowning, electricity, poison, stunning, 
and all spells or attacks that affect your physiology or respiration, because you 
have no physiology or respiration while this spell is in effect. You take only 
half damage from acid and fire of all kinds. However, you also become vulnerable 
to all special attacks that affect iron golems.

You gain a +6 enhancement bonus to your Strength score, but you take a -6 penalty 
to Dexterity as well (to a minimum Dexterity score of 1), and your speed is reduced 
to half normal. You have an arcane spell failure chance of 50\% and a -8 armor 
check penalty, just as if you were clad in full plate armor. You cannot drink (and 
thus can't use potions) or play wind instruments.

Your unarmed attacks deal damage equal to a club sized for you (1d4 for Small characters 
or 1d6 for Medium characters), and you are considered armed when making unarmed 
attacks.

Your weight increases by a factor of ten, causing you to sink in water like a stone. 
However, you could survive the crushing pressure and lack of air at the bottom 
of the ocean---at least until the spell duration expires.

\textit{Arcane Material Component: }A small piece of iron that was once part of 
either an iron golem, a hero's armor, or a war machine.

\vspace{12pt}
Ironwood

Transmutation

\textbf{Level:} Drd 6

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 minute/lb. created

\textbf{Range:} 0 ft.

\textbf{Effect:} An \textit{ironwood }object weighing up to 5 lb./level

\textbf{Duration:} One day/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

\textit{Ironwood }is a magical substance created by druids from normal wood. While 
remaining natural wood in almost every way, \textit{ironwood }is as strong, heavy, 
and resistant to fire as steel. Spells that affect metal or iron do not function 
on \textit{ironwood}. Spells that affect wood do affect \textit{ironwood}, although 
\textit{ironwood }does not burn. Using this spell with \textit{wood shape }or a 
wood-related Craft check, you can fashion wooden items that function as steel items. 
Thus, wooden plate armor and wooden swords can be created that are as durable as 
their normal steel counterparts. These items are freely usable by druids.

Further, if you make only half as much \textit{ironwood }as the spell would normally 
allow, any weapon, shield, or suit of armor so created is treated as a magic item 
with a +1 enhancement bonus.

\textit{Material Component: }Wood shaped into the form of the intended \textit{ironwood 
}object.

\vspace{12pt}
Irresistible Dance

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Brd 6, Sor/Wiz 8

\textbf{Components:} V

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Living creature touched

\textbf{Duration:} 1d4+1 rounds

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

The subject feels an undeniable urge to dance and begins doing so, complete with 
foot shuffling and tapping. The spell effect makes it impossible for the subject 
to do anything other than caper and prance in place. The effect imposes a -4 penalty 
to Armor Class and a -10 penalty on Reflex saves, and it negates any AC bonus granted 
by a shield the target holds. The dancing subject provokes attacks of opportunity 
each round on its turn.

\vspace{12pt}
Jump

Transmutation

\textbf{Level:} Drd 1, Rgr 1, Sor/Wiz 1

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 1 min./level (D)

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes

The subject gets a +10 enhancement bonus on Jump checks. The enhancement bonus 
increases to +20 at caster level 5th, and to +30 (the maximum) at caster level 
9th.

\textit{Material Component: }A grasshopper's hind leg, which you break when the 
spell is cast.

\vspace{12pt}
Keen Edge

Transmutation

\textbf{Level:} Sor/Wiz 3

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} One weapon or fifty projectiles, all of which must be in contact 
with each other at the time of casting

\textbf{Duration:} 10 min./level

\textbf{Saving Throw: }Will negates (harmless, object)

\textbf{Spell Resistance:} Yes (harmless, object)

This spell makes a weapon magically keen, improving its ability to deal telling 
blows. This transmutation doubles the threat range of the weapon. A threat range 
of 20 becomes 19-20, a threat range of 19-20 becomes 17-20, and a threat range 
of 18-20 becomes 15-20. The spell can be cast only on piercing or slashing weapons. 
If cast on arrows or crossbow bolts, the \textit{keen edge }on a particular projectile 
ends after one use, whether or not the missile strikes its intended target. (Treat 
shuriken as arrows, rather than as thrown weapons, for the purpose of this spell.)

Multiple effects that increase a weapon's threat range (such as the \textit{keen 
edge }spell and the Improved Critical feat) don't stack. You can't cast this spell 
on a natural weapon, such as a claw.

\vspace{12pt}
Knock

Transmutation

\textbf{Level:} Sor/Wiz 2

\textbf{Components:} V

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Target:} One door, box, or chest with an area of up to 10 sq. ft./level

\textbf{Duration:} Instantaneous; see text

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

The \textit{knock }spell opens stuck, barred, locked, \textit{held, }or \textit{arcane 
locked }doors. It opens secret doors, as well as locked or trick-opening boxes 
or chests. It also loosens welds, shackles, or chains (provided they serve to hold 
closures shut). If used to open a \textit{arcane locked }door, the spell does not 
remove the \textit{arcane lock }but simply suspends its functioning for 10 minutes. 
In all other cases, the door does not relock itself or become stuck again on its 
own. \textit{Knock }does not raise barred gates or similar impediments (such as 
a portcullis), nor does it affect ropes, vines, and the like. The effect is limited 
by the area. Each spell can undo as many as two means of preventing egress. 

\vspace{12pt}
Know Direction

Divination

\textbf{Level:} Brd 0, Drd 0

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} Instantaneous

You instantly know the direction of north from your current position. The spell 
is effective in any environment in which ``north'' exists, but it may not work 
in extraplanar settings. Your knowledge of north is correct at the moment of casting, 
but you can get lost again within moments if you don't find some external reference 
point to help you keep track of direction.

\vspace{12pt}
Legend Lore

Divination

\textbf{Level:} Brd 4, Knowledge 7, Sor/Wiz 6

\textbf{Components:} V, S, M, F

\textbf{Casting Time:} See text

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} See text

\textit{Legend lore }brings to your mind legends about an important person, place, 
or thing. If the person or thing is at hand, or if you are in the place in question, 
the casting time is only 1d4x10 minutes. If you have only detailed information 
on the person, place, or thing, the casting time is 1d10 days, and the resulting 
lore is less complete and specific (though it often provides enough information 
to help you find the person, place, or thing, thus allowing a better \textit{legend 
lore }result next time). If you know only rumors, the casting time is 2d6 weeks, 
and the resulting lore is vague and incomplete (though it often directs you to 
more detailed information, thus allowing a better \textit{legend lore }result next 
time).

During the casting, you cannot engage in other than routine activities: eating, 
sleeping, and so forth. When completed, the divination brings legends (if any) 
about the person, place, or things to your mind. These may be legends that are 
still current, legends that have been forgotten, or even information that has never 
been generally known. If the person, place, or thing is not of legendary importance, 
you gain no information. As a rule of thumb, characters who are 11th level and 
higher are ``legendary,'' as are the sorts of creatures they contend with, the 
major magic items they wield, and the places where they perform their key deeds.

\textit{Material Component: }Incense worth at least 250 gp.

\textit{Focus: }Four strips of ivory (worth 50 gp each) formed into a rectangle.

\vspace{12pt}
Lesser (Spell Name)

Any spell whose name begins with \textit{lesser }is alphabetized in this chapter 
according to the second word of the spell name. Thus, the description of a \textit{lesser 
}spell appears near the description of the spell on which it is based. Spell chains 
that have \textit{lesser }spells in them include those based on the spells \textit{confusion, 
geas, globe of invulnerability, planar ally, planar binding, }and \textit{restoration.}

\vspace{12pt}
Levitate

Transmutation

\textbf{Level:} Sor/Wiz 2

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal or close (25 ft. + 5 ft./2 levels)

\textbf{Target:} You or one willing creature or one object (total weight up to 
100 lb./level)

\textbf{Duration:} 1 min./level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

\textit{Levitate }allows you to move yourself, another creature, or an object up 
and down as you wish. A creature must be willing to be \textit{levitated, }and 
an object must be unattended or possessed by a willing creature. You can mentally 
direct the recipient to move up or down as much as 20 feet each round; doing so 
is a move action. You cannot move the recipient horizontally, but the recipient 
could clamber along the face of a cliff, for example, or push against a ceiling 
to move laterally (generally at half its base land speed).

A \textit{levitating }creature that attacks with a melee or ranged weapon finds 
itself increasingly unstable; the first attack has a -1 penalty on attack rolls, 
the second -2, and so on, to a maximum penalty of -5. A full round spent stabilizing 
allows the creature to begin again at -1.

\textit{Focus: }Either a small leather loop or a piece of golden wire bent into 
a cup shape with a long shank on one end.

\vspace{12pt}
Light

Evocation [Light]

\textbf{Level:} Brd 0, Clr 0, Drd 0, Sor/Wiz 0

\textbf{Components:} V, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Object touched

\textbf{Duration:} 10 min./level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell causes an object to glow like a torch, shedding bright light in a 20-foot 
radius (and dim light for an additional 20 feet) from the point you touch. The 
effect is immobile, but it can be cast on a movable object. Light taken into an 
area of magical \textit{darkness }does not function.

A light spell (one with the light descriptor) counters and dispels a darkness spell 
(one with the darkness descriptor) of an equal or lower level.

\textit{Arcane Material Component: }A firefly or a piece of phosphorescent moss.

\vspace{12pt}
Lightning Bolt

Evocation [Electricity]

\textbf{Level:} Sor/Wiz 3

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 120 ft.

\textbf{Area:} 120-ft. line

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} Reflex half

\textbf{Spell Resistance:} Yes

You release a powerful stroke of electrical energy that deals 1d6 points of electricity 
damage per caster level (maximum 10d6) to each creature within its area. The bolt 
begins at your fingertips.

The \textit{lightning bolt }sets fire to combustibles and damages objects in its 
path. It can melt metals with a low melting point, such as lead, gold, copper, 
silver, or bronze. If the damage caused to an interposing barrier shatters or breaks 
through it, the bolt may continue beyond the barrier if the spell's range permits; 
otherwise, it stops at the barrier just as any other spell effect does.

\textit{Material Component: }A bit of fur and an amber, crystal, or glass rod.

\vspace{12pt}
Limited Wish

Universal

\textbf{Level:} Sor/Wiz 7

\textbf{Components:} V, S, XP

\textbf{Casting Time:} 1 standard action

\textbf{Range:} See text

\textbf{Target, Effect, or Area:} See text

\textbf{Duration:} See text

\textbf{Saving Throw:} None; see text

\textbf{Spell Resistance:} Yes

A \textit{limited wish }lets you create nearly any type of effect. For example, 
a \textit{limited wish }can do any of the following things. 

Duplicate any sorcerer/wizard spell of 6th level or lower, provided the spell is 
not of a school prohibited to you. 

Duplicate any other spell of 5th level or lower, provided the spell is not of a 
school prohibited to you. 

Duplicate any sorcerer/wizard spell of 5th level or lower, even if it's of a prohibited 
school. 

Duplicate any other spell of 4th level or lower, even if it's of a prohibited school. 

Undo the harmful effects of many spells, such as \textit{geas/quest }or \textit{insanity}. 

Produce any other effect whose power level is in line with the above effects, such 
as a single creature automatically hitting on its next attack or taking a -7 penalty 
on its next saving throw.

A duplicated spell allows saving throws and spell resistance as normal (but the 
save DC is for a 7th-level spell). When a \textit{limited wish }duplicates a spell 
that has an XP cost, you must pay that cost or 300 XP, whichever is more. When 
a \textit{limited wish }spell duplicates a spell with a material component that 
costs more than 1,000 gp, you must provide that component.

\textit{XP Cost: }300 XP or more (see above).

\vspace{12pt}
Liveoak

Transmutation

\textbf{Level:} Drd 6

\textbf{Components:} V, S

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Touch

\textbf{Target:} Tree touched

\textbf{Duration:} One day/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell turns an oak tree into a protector or guardian. The spell can be cast 
on only a single tree at a time; while \textit{liveoak }is in effect, you can't 
cast it again on another tree. The tree on which the spell is cast must be within 
10 feet of your dwelling place, within a place sacred to you, or within 300 feet 
of something that you wish to guard or protect.

\textit{Liveoak }must be cast on a healthy, Huge oak. A triggering phrase of up 
to one word per caster level is placed on the targeted oak. The \textit{liveoak 
}spell triggers the tree into animating as a treant. 

If \textit{liveoak }is dispelled, the tree takes root immediately, wherever it 
happens to be. If released by you, the tree tries to return to its original location 
before taking root.

\vspace{12pt}
Locate Creature

Divination

\textbf{Level:} Brd 4, Sor/Wiz 4

\textbf{Components:} V, S, M

\textbf{Duration:} 10 min./level

This spell functions like \textit{locate object, }except this spell locates a known 
or familiar creature.

You slowly turn and sense when you are facing in the direction of the creature 
to be located, provided it is within range. You also know in which direction the 
creature is moving, if any.

The spell can locate a creature of a specific kind or a specific creature known 
to you. It cannot find a creature of a certain type. To find a kind of creature, 
you must have seen such a creature up close (within 30 feet) at least once.

Running water blocks the spell. It cannot detect objects. It can be fooled by \textit{mislead}, 
\textit{nondetection}, and \textit{polymorph }spells.

\textit{Material Component: }A bit of fur from a bloodhound.

\vspace{12pt}
Locate Object

Divination

\textbf{Level:} Brd 2, Clr 3, Sor/Wiz 2, Travel 2

\textbf{Components:} V, S, F/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Area:} Circle, centered on you, with a radius of 400 ft. + 40 ft./level

\textbf{Duration:} 1 min./level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You sense the direction of a well-known or clearly visualized object. You can search 
for general items, in which case you locate the nearest one of its kind if more 
than one is within range. Attempting to find a certain item requires a specific 
and accurate mental image; if the image is not close enough to the actual object, 
the spell fails. You cannot specify a unique item unless you have observed that 
particular item firsthand (not through divination).

The spell is blocked by even a thin sheet of lead. Creatures cannot be found by 
this spell. \textit{Polymorph any object }fools it.

\textit{Arcane Focus: }A forked twig.

\vspace{12pt}
Longstrider

Transmutation

\textbf{Level:} Drd 1, Rgr 1, Travel 1

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 1 hour/level (D)

This spell increases your base land speed by 10 feet. (This adjustment counts as 
an enhancement bonus.) It has no effect on other modes of movement, such as burrow, 
climb, fly, or swim.

\textit{Material Component: }A pinch of dirt.

\vspace{12pt}
Lullaby

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Brd 0

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Area:} Living creatures within a 10-ft.-radius burst

\textbf{Duration:} Concentration + 1 round/level (D)

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

Any creature within the area that fails a Will save becomes drowsy and inattentive, 
taking a -5 penalty on Listen and Spot checks and a -2 penalty on Will saves against 
\textit{sleep }effects while the \textit{lullaby }is in effect. \textit{Lullaby 
}lasts for as long as the caster concentrates, plus up to 1 round per caster level 
thereafter.

\newpage

\end{document}
