%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{array}
\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

This material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

{\LARGE{}SPELLS (C)}

\vspace{12pt}
Call Lightning

Evocation [Electricity]

\textbf{Level:} Drd 3

\textbf{Components:} V, S

\textbf{Casting Time:} 1 round

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Effect:} One or more 30-ft.-long vertical lines of lightning

\textbf{Duration:} 1 min./level

\textbf{Saving Throw:} Reflex half

\textbf{Spell Resistance:} Yes

Immediately upon completion of the spell, and once per round thereafter, you may 
call down a 5-foot-wide, 30-foot-long, vertical bolt of lightning that deals 3d6 
points of electricity damage. The bolt of lightning flashes down in a vertical 
stroke at whatever target point you choose within the spell's range (measured from 
your position at the time). Any creature in the target square or in the path of 
the bolt is affected.

You need not call a bolt of lightning immediately; other actions, even spellcasting, 
can be performed. However, each round after the first you may use a standard action 
(concentrating on the spell) to call a bolt. You may call a total number of bolts 
equal to your caster level (maximum 10 bolts).

If you are outdoors and in a stormy area---a rain shower, clouds and wind, hot 
and cloudy conditions, or even a tornado (including a whirlwind formed by a djinni 
or an air elemental of at least Large size)---each bolt deals 3d10 points of electricity 
damage instead of 3d6.

This spell functions indoors or underground but not underwater.

\vspace{12pt}
Call Lightning Storm

Evocation [Electricity]

\textbf{Level:} Drd 5

\textbf{Range:} Long (400 ft. + 40 ft./level)

This spell functions like \textit{call lightning}, except that each bolt deals 
5d6 points of electricity damage (or 5d10 if created outdoors in a stormy area), 
and you may call a maximum of 15 bolts.

\vspace{12pt}
Calm Animals

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Animal 1, Drd 1, Rgr 1

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} Animals within 30 ft. of each other

\textbf{Duration:} 1 min./level

\textbf{Saving Throw: }Will negates; see text

\textbf{Spell Resistance:} Yes

This spell soothes and quiets animals, rendering them docile and harmless. Only 
ordinary animals (those with Intelligence scores of 1 or 2) can be affected by 
this spell. All the subjects must be of the same kind, and no two may be more than 
30 feet apart. The maximum number of Hit Dice of animals you can affect is equal 
to 2d4 + caster level. A dire animal or an animal trained to attack or guard is 
allowed a saving throw; other animals are not.

The affected creatures remain where they are and do not attack or flee. They are 
not helpless and defend themselves normally if attacked. Any threat breaks the 
spell on the threatened creatures.

\vspace{12pt}
Calm Emotions

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Brd 2, Clr 2, Law 2

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Area:} Creatures in a 20-ft.-radius spread

\textbf{Duration:} Concentration, up to 1 round/level (D)

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

This spell calms agitated creatures. You have no control over the affected creatures, 
but \textit{calm emotions }can stop raging creatures from fighting or joyous ones 
from reveling. Creatures so affected cannot take violent actions (although they 
can defend themselves) or do anything destructive. Any aggressive action against 
or damage dealt to a calmed creature immediately breaks the spell on all calmed 
creatures.

This spell automatically suppresses (but does not dispel) any morale bonuses granted 
by spells such as \textit{bless, good hope, }and \textit{rage, }as well as negating 
a bard's ability to inspire courage or a barbarian's rage ability. It also suppresses 
any fear effects and removes the \textit{confused }condition from all targets\textit{. 
}While the spell lasts, a suppressed spell or effect has no effect. When the \textit{calm 
emotions }spell ends, the original spell or effect takes hold of the creature again, 
provided that its duration has not expired in the meantime.

\vspace{12pt}
Cat's Grace

Transmutation

\textbf{Level:} Brd 2, Drd 2, Rgr 2, Sor/Wiz 2

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 1 min./level

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes

The transmuted creature becomes more graceful, agile, and coordinated. The spell 
grants a +4 enhancement bonus to Dexterity, adding the usual benefits to AC, Reflex 
saves, and other uses of the Dexterity modifier.

\textit{Material Component: }A pinch of cat fur.

\vspace{12pt}
Cat's Grace, Mass

Transmutation

\textbf{Level:} Brd 6, Drd 6, Sor/Wiz 6

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} One creature/level, no two of which can be more than 30 ft. apart

This spell functions like \textit{cat's grace}, except that it affects multiple 
creatures.

\vspace{12pt}
Cause Fear

Necromancy [Fear, Mind-Affecting]

\textbf{Level:} Brd 1, Clr 1, Death 1, Sor/Wiz 1

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One living creature with 5 or fewer HD

\textbf{Duration:} 1d4 rounds or 1 round; see text

\textbf{Saving Throw: }Will partial

\textbf{Spell Resistance:} Yes

The affected creature becomes frightened. If the subject succeeds on a Will save, 
it is shaken for 1 round. Creatures with 6 or more Hit Dice are immune to this 
effect.

\textit{Cause fear }counters and dispels \textit{remove fear}.

\vspace{12pt}
Chain Lightning

Evocation [Electricity]

\textbf{Level:} Air 6, Sor/Wiz 6

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Targets:} One primary target, plus one secondary target/level (each of 
which must be within 30 ft. of the primary target)

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} Reflex half

\textbf{Spell Resistance:} Yes

This spell creates an electrical discharge that begins as a single stroke commencing 
from your fingertips. Unlike \textit{lightning bolt}, \textit{chain lightning }strikes 
one object or creature initially, then arcs to other targets.

The bolt deals 1d6 points of electricity damage per caster level (maximum 20d6) 
to the primary target. After it strikes, lightning can arc to a number of secondary 
targets equal to your caster level (maximum 20). The secondary bolts each strike 
one target and deal half as much damage as the primary one did (rounded down).

Each target can attempt a Reflex saving throw for half damage. You choose secondary 
targets as you like, but they must all be within 30 feet of the primary target, 
and no target can be struck more than once. You can choose to affect fewer secondary 
targets than the maximum.

\textit{Focus: }A bit of fur; a piece of amber, glass, or a crystal rod; plus one 
silver pin for each of your caster levels.

\vspace{12pt}
Changestaff

Transmutation

\textbf{Level:} Drd 7

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 round

\textbf{Range:} Touch

\textbf{Target:} Your touched staff

\textbf{Duration:} 1 hour/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You change a specially prepared quarterstaff into a Huge treantlike creature, about 
24 feet tall. When you plant the end of the staff in the ground and speak a special 
command to conclude the casting of the spell, your staff turns into a creature 
that looks and fights just like a treant. The staff-treant defends you and obeys 
any spoken commands. However, it is by no means a true treant; it cannot converse 
with actual treants or control trees. If the staff-treant is reduced to 0 or fewer 
hit points, it crumbles to powder and the staff is destroyed. Otherwise, the staff 
returns to its normal form when the spell duration expires (or when the spell is 
dismissed), and it can be used as the focus for another casting of the spell. The 
staff-treant is always at full strength when created, despite any wounds it may 
have incurred the last time it appeared.

\textit{Focus: }The quarterstaff, which must be specially prepared. The staff must 
be a sound limb cut from an ash, oak, or yew, then cured, shaped, carved, and polished 
(a process requiring twenty-eight days).

You cannot adventure or engage in other strenuous activity during the shaping and 
carving of the staff.

\vspace{12pt}
Chaos Hammer

Evocation [Chaotic]

\textbf{Level:} Chaos 4

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Area:} 20-ft.-radius burst

\textbf{Duration:} Instantaneous (1d6 rounds); see text

\textbf{Saving Throw: }Will partial; see text

\textbf{Spell Resistance:} Yes

You unleash chaotic power to smite your enemies. The power takes the form of a 
multicolored explosion of leaping, ricocheting energy. Only lawful and neutral 
(not chaotic) creatures are harmed by the spell.

The spell deals 1d8 points of damage per two caster levels (maximum 5d8) to lawful 
creatures (or 1d6 points of damage per caster level, maximum 10d6, to lawful outsiders) 
and slows them for 1d6 rounds (see the \textit{slow} spell). A successful Will 
save reduces the damage by half and negates the slow effect.

The spell deals only half damage against creatures who are neither lawful nor chaotic, 
and they are not slowed. Such a creature can reduce the damage by half again (down 
to one-quarter) with a successful Will save.

\vspace{12pt}
Charm Animal

Enchantment (Charm) [Mind-Affecting]

\textbf{Level:} Drd 1, Rgr 1

\textbf{Target:} One animal

This spell functions like \textit{charm person}, except that it affects a creature 
of the animal type.

\vspace{12pt}
Charm Monster

Enchantment (Charm) [Mind-Affecting]

\textbf{Level:} Brd 3, Sor/Wiz 4

\textbf{Target:} One living creature

\textbf{Duration:} One day/level

This spell functions like \textit{charm person}, except that the effect is not 
restricted by creature type or size.

\vspace{12pt}
Charm Monster, Mass

Enchantment (Charm) [Mind-Affecting]

\textbf{Level:} Brd 6, Sor/Wiz 8

\textbf{Components:} V

\textbf{Targets:} One or more creatures, no two of which can be more than 30 ft. 
apart

\textbf{Duration:} One day/level

This spell functions like \textit{charm monster, }except that \textit{mass charm 
monster }affects a number of creatures whose combined HD do not exceed twice your 
level, or at least one creature regardless of HD. If there are more potential targets 
than you can affect, you choose them one at a time until you choose a creature 
with too many HD.

\vspace{12pt}
Charm Person

Enchantment (Charm) [Mind-Affecting]

\textbf{Level:} Brd 1, Sor/Wiz 1

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One humanoid creature

\textbf{Duration:} 1 hour/level

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

This charm makes a humanoid creature regard you as its trusted friend and ally 
(treat the target's attitude as friendly). If the creature is currently being threatened 
or attacked by you or your allies, however, it receives a +5 bonus on its saving 
throw.

The spell does not enable you to control the \textit{charmed }person as if it were 
an automaton, but it perceives your words and actions in the most favorable way. 
You can try to give the subject orders, but you must win an opposed Charisma check 
to convince it to do anything it wouldn't ordinarily do. (Retries are not allowed.) 
An affected creature never obeys suicidal or obviously harmful orders, but it might 
be convinced that something very dangerous is worth doing.  Any act by you or your 
apparent allies that threatens the \textit{charmed }person breaks the spell. You 
must speak the person's language to communicate your commands, or else be good 
at pantomiming.

\vspace{12pt}
Chill Metal

Transmutation [Cold]

\textbf{Level:} Drd 2

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target: }Metal equipment of one creature per two levels, no two of which 
can be more than 30 ft. apart; or 25 lb. of metal/level, none of which can be more 
than 30 ft. away from any of the rest

\textbf{Duration:} 7 rounds

\textbf{Saving Throw: }Will negates (object)

\textbf{Spell Resistance:} Yes (object)

\textit{Chill metal }makes metal extremely cold. Unattended, nonmagical metal gets 
no saving throw. Magical metal is allowed a saving throw against the spell. An 
item in a creature's possession uses the creature's saving throw bonus unless its 
own is higher.

A creature takes cold damage if its equipment is chilled. It takes full damage 
if its armor is affected or if it is holding, touching, wearing, or carrying metal 
weighing one-fifth of its weight. The creature takes minimum damage (1 point or 
2 points; see the table) if it's not wearing metal armor and the metal that it's 
carrying weighs less than one-fifth of its weight.

On the first round of the spell, the metal becomes chilly and uncomfortable to 
touch but deals no damage. The same effect also occurs on the last round of the 
spell's duration. During the second (and also the next-to-last) round, icy coldness 
causes pain and damage. In the third, fourth, and fifth rounds, the metal is freezing 
cold, causing more damage, as shown on the table below.

\begin{tabular}{|>{\raggedright}p{27pt}|>{\raggedright}p{82pt}|>{\raggedright}p{40pt}|}
\hline
\subsection*{R\textbf{ound}} & \subsection*{M\textbf{etal Temperature}} & \subsection*{D\textbf{amage}}\tabularnewline
\hline
1 & Cold & None\tabularnewline
\hline
2 & Icy & 1d4 points\tabularnewline
\hline
3-5 & Freezing & 2d4 points\tabularnewline
\hline
6 & Icy & 1d4 points\tabularnewline
\hline
7 & Cold & None\tabularnewline
\hline
\end{tabular}

Any heat intense enough to damage the creature negates cold damage from the spell 
(and vice versa) on a point-for-point basis. Underwater, \textit{chill metal }deals 
no damage, but ice immediately forms around the affected metal, making it more 
buoyant.

\textit{Chill metal }counters and dispels \textit{heat metal}.

\vspace{12pt}
Chill Touch

Necromancy

\textbf{Level:} Sor/Wiz 1

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Targets:} Creature or creatures touched (up to one/level)

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} Fortitude partial or Will negates; see text

\textbf{Spell Resistance:} Yes

A touch from your hand, which glows with blue energy, disrupts the life force of 
living creatures. Each touch channels negative energy that deals 1d6 points of 
damage. The touched creature also takes 1 point of Strength damage unless it makes 
a successful Fortitude saving throw. You can use this melee touch attack up to 
one time per level.

An undead creature you touch takes no damage of either sort, but it must make a 
successful Will saving throw or flee as if panicked for 1d4 rounds +1 round per 
caster level.

\vspace{12pt}
Circle of Death

Necromancy [Death]

\textbf{Level:} Sor/Wiz 6

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Area:} Several living creatures within a 40-ft.-radius burst

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} Fortitude negates

\textbf{Spell Resistance:} Yes

A \textit{circle of death }snuffs out the life force of living creatures, killing 
them instantly.

The spell slays 1d4 HD worth of living creatures per caster level (maximum 20d4). 
Creatures with the fewest HD are affected first; among creatures with equal HD, 
those who are closest to the burst's point of origin are affected first. No creature 
of 9 or more HD can be affected, and Hit Dice that are not sufficient to affect 
a creature are wasted.

\textit{Material Component: }The powder of a crushed black pearl with a minimum 
value of 500 gp.

\vspace{12pt}
Clairaudience/Clairvoyance

Divination (Scrying)

\textbf{Level:} Brd 3, Knowledge 3, Sor/Wiz 3

\textbf{Components:} V, S, F/DF

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Effect: }Magical sensor

\textbf{Duration:} 1 min./level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

\textit{Clairaudience/clairvoyance }creates an invisible magical sensor at a specific 
location that enables you to hear or see (your choice) almost as if you were there. 
You don't need line of sight or line of effect, but the locale must be known---a 
place familiar to you or an obvious one. Once you have selected the locale, the 
sensor doesn't move, but you can rotate it in all directions to view the area as 
desired. Unlike other scrying spells, this spell does not allow magically or supernaturally 
enhanced senses to work through it. If the chosen locale is magically dark, you 
see nothing. If it is naturally pitch black, you can see in a 10- foot radius around 
the center of the spell's effect. \textit{Clairaudience/clairvoyance }functions 
only on the plane of existence you are currently occupying.

\textit{Arcane Focus: }A small horn (for hearing) or a glass eye (for seeing).

\vspace{12pt}
Clenched Fist

Evocation [Force]

\textbf{Level:} Sor/Wiz 8, Strength 8

\textbf{Components:} V, S, F/DF

This spell functions like \textit{interposing hand, }except that the hand can interpose 
itself, push, or strike one opponent that you select. The floating hand can move 
as far as 60 feet and can attack in the same round. Since this hand is directed 
by you, its ability to notice or attack invisible or concealed creatures is no 
better than yours.

The hand attacks once per round, and its attack bonus equals your caster level 
+ your Intelligence, Wisdom, or Charisma modifier (for a wizard, cleric, or sorcerer, 
respectively), +11 for the hand's Strength score (33), -1 for being Large. The 
hand deals 1d8+11 points of damage on each attack, and any creature struck must 
make a Fortitude save (against this spell's save DC) or be stunned for 1 round. 
Directing the spell to a new target is a move action.

The \textit{clenched fist }can also interpose itself as \textit{interposing hand 
}does, or it can bull rush an opponent as \textit{forceful hand }does, but at a 
+15 bonus on the Strength check.

Clerics who cast this spell name it for their deities.

\textit{Arcane Focus: }A leather glove.

\vspace{12pt}
Cloak of Chaos

Abjuration [Chaotic]

\textbf{Level:} Chaos 8, Clr 8

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 20 ft.

\textbf{Targets:} One creature/level in a 20-ft.-radius burst centered on you

\textbf{Duration:} 1 round/level (D)

\textbf{Saving Throw:} See text

\textbf{Spell Resistance:} Yes (harmless)

A random pattern of color surrounds the subjects, protecting them from attacks, 
granting them resistance to spells cast by lawful creatures, and causing lawful 
creatures that strike the subjects to become \textit{confused. }This abjuration 
has four effects.

First, each warded creature gains a +4 deflection bonus to AC and a +4 resistance 
bonus on saves. Unlike \textit{protection from law}, the benefit of this spell 
applies against all attacks, not just against attacks by lawful creatures.

Second, each warded creature gains spell resistance 25 against lawful spells and 
spells cast by lawful creatures.

Third, the abjuration blocks possession and mental influence, just as \textit{protection 
from law }does.

Finally, if a lawful creature succeeds on a melee attack against a warded creature, 
the offending attacker is \textit{confused }for 1 round (Will save negates, as 
with the \textit{confusion }spell, but against the save DC of \textit{cloak of 
chaos}).

\textit{Focus: }A tiny reliquary containing some sacred relic, such as a scrap 
of parchment from a chaotic text. The reliquary costs at least 500 gp.

\vspace{12pt}
Clone

Necromancy

\textbf{Level:} Sor/Wiz 8

\textbf{Components:} V, S, M, F

\textbf{Casting Time:} 10 minutes

\textbf{Range:} 0 ft.

\textbf{Effect:} One clone

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell makes an inert duplicate of a creature. If the original individual has 
been slain, its soul immediately transfers to the clone, creating a replacement 
(provided that the soul is free and willing to return). The original's physical 
remains, should they still exist, become inert and cannot thereafter be restored 
to life. If the original creature has reached the end of its natural life span 
(that is, it has died of natural causes), any cloning attempt fails.

To create the duplicate, you must have a piece of flesh (not hair, nails, scales, 
or the like) with a volume of at least 1 cubic inch that was taken from the original 
creature's living body. The piece of flesh need not be fresh, but it must be kept 
from rotting. Once the spell is cast, the duplicate must be grown in a laboratory 
for 2d4 months.

When the clone is completed, the original's soul enters it immediately, if that 
creature is already dead. The clone is physically identical with the original and 
possesses the same personality and memories as the original. In other respects, 
treat the clone as if it were the original character raised from the dead, including 
the loss of one level or 2 points of Constitution (if the original was a 1st-level 
character). If this Constitution adjustment would give the clone a Constitution 
score of 0, the spell fails. If the original creature has lost levels since the 
flesh sample was taken and died at a lower level than the clone would otherwise 
be, the clone is one level below the level at which the original died.

The spell duplicates only the original's body and mind, not its equipment.

A duplicate can be grown while the original still lives, or when the original soul 
is unavailable, but the resulting body is merely a soulless bit of inert flesh, 
which rots if not preserved.

\textit{Material Component: }The piece of flesh and various laboratory supplies 
(cost 1,000 gp).

\textit{Focus: }Special laboratory equipment (cost 500 gp).

\vspace{12pt}
Cloudkill

Conjuration (Creation)

\textbf{Level:} Sor/Wiz 5

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Effect:} Cloud spreads in 20-ft. radius, 20 ft. high

\textbf{Duration:} 1 min./level

\textbf{Saving Throw:} Fortitude partial; see text

\textbf{Spell Resistance:} No

This spell generates a bank of fog, similar to a \textit{fog cloud, }except that 
its vapors are yellowish green and poisonous. These vapors automatically kill any 
living creature with 3 or fewer HD (no save). A living creature with 4 to 6 HD 
is slain unless it succeeds on a Fortitude save (in which case it takes 1d4 points 
of Constitution damage on your turn each round while in the cloud).

A living creature with 6 or more HD takes 1d4 points of Constitution damage on 
your turn each round while in the cloud (a successful Fortitude save halves this 
damage). Holding one's breath doesn't help, but creatures immune to poison are 
unaffected by the spell.

Unlike a \textit{fog cloud}, the \textit{cloudkill }moves away from you at 10 feet 
per round, rolling along the surface of the ground.

Figure out the cloud's new spread each round based on its new point of origin, 
which is 10 feet farther away from the point of origin where you cast the spell.

Because the vapors are heavier than air, they sink to the lowest level of the land, 
even pouring down den or sinkhole openings. It cannot penetrate liquids, nor can 
it be cast underwater.

\vspace{12pt}
Color Spray

Illusion (Pattern) [Mind-Affecting]

\textbf{Level:} Sor/Wiz 1

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 15 ft.

\textbf{Area:} Cone-shaped burst

\textbf{Duration:} Instantaneous; see text

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

A vivid cone of clashing colors springs forth from your hand, causing creatures 
to become stunned, perhaps also blinded, and possibly knocking them unconscious.

Each creature within the cone is affected according to its Hit Dice.

\textit{2 HD or less: }The creature is unconscious, blinded, and stunned for 2d4 
rounds, then blinded and stunned for 1d4 rounds, and then stunned for 1 round. 
(Only living creatures are knocked unconscious.)

\textit{3 or 4 HD: }The creature is blinded and stunned for 1d4 rounds, then stunned 
for 1 round.

\textit{5 or more HD: }The creature is stunned for 1 round.

Sightless creatures are not affected by \textit{color spray}.

\textit{Material Component: }A pinch each of powder or sand that is colored red, 
yellow, and blue.

\vspace{12pt}
Command

Enchantment (Compulsion) [Language-Dependent, Mind-Affecting]

\textbf{Level:} Clr 1

\textbf{Components:} V

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One living creature

\textbf{Duration:} 1 round

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

You give the subject a single command, which it obeys to the best of its ability 
at its earliest opportunity. You may select from the following options.

\textit{Approach: }On its turn, the subject moves toward you as quickly and directly 
as possible for 1 round. The creature may do nothing but move during its turn, 
and it provokes attacks of opportunity for this movement as normal.

\textit{Drop: }On its turn, the subject drops whatever it is holding. It can't 
pick up any dropped item until its next turn.

\textit{Fall: }On its turn, the subject falls to the ground and remains prone for 
1 round. It may act normally while prone but takes any appropriate penalties.

\textit{Flee: }On its turn, the subject moves away from you as quickly as possible 
for 1 round. It may do nothing but move during its turn, and it provokes attacks 
of opportunity for this movement as normal.

\textit{Halt: }The subject stands in place for 1 round. It may not take any actions 
but is not considered helpless.

If the subject can't carry out your command on its next turn, the spell automatically 
fails.

\vspace{12pt}
Command, Greater

Enchantment (Compulsion) [Language-Dependent, Mind-Affecting]

\textbf{Level:} Clr 5

\textbf{Targets:} One creature/level, no two of which can be more than 30 ft. apart

\textbf{Duration:} 1 round/level

This spell functions like \textit{command, }except that up to one creature per 
level may be affected, and the activities continue beyond 1 round. At the start 
of each commanded creature's action after the first, it gets another Will save 
to attempt to break free from the spell. Each creature must receive the same command.

\vspace{12pt}
Command Plants

Transmutation

\textbf{Level:} Drd 4, Plant 4, Rgr 3

\textbf{Components:} V

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} Up to 2 HD/level of plant creatures, no two of which can be more 
than 30 ft. apart

\textbf{Duration:} One day/level

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

This spell allows you some degree of control over one or more plant creatures. 
Affected plant creatures can understand you, and they perceive your words and actions 
in the most favorable way (treat their attitude as friendly). They will not attack 
you while the spell lasts. You can try to give a subject orders, but you must win 
an opposed Charisma check to convince it to do anything it wouldn't ordinarily 
do. (Retries are not allowed.) A commanded plant never obeys suicidal or obviously 
harmful orders, but it might be convinced that something very dangerous is worth 
doing.

You can affect a number of plant creatures whose combined level or HD do not exceed 
twice your level.

\vspace{12pt}
Command Undead

Necromancy

\textbf{Level:} Sor/Wiz 2

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} One undead creature

\textbf{Duration:} One day/level

\textbf{Saving Throw: }Will negates; see text

\textbf{Spell Resistance:} Yes

This spell allows you some degree of control over an undead creature. Assuming 
the subject is intelligent, it perceives your words and actions in the most favorable 
way (treat its attitude as friendly). It will not attack you while the spell lasts. 
You can try to give the subject orders, but you must win an opposed Charisma check 
to convince it to do anything it wouldn't ordinarily do. (Retries are not allowed.) 
An intelligent commanded undead never obeys suicidal or obviously harmful orders, 
but it might be convinced that something very dangerous is worth doing.

A nonintelligent undead creature gets no saving throw against this spell. When 
you control a mindless being, you can communicate only basic commands, such as 
``come here,'' ``go there,'' ``fight,'' ``stand still,'' and so on. Nonintelligent 
undead won't resist suicidal or obviously harmful orders.

Any act by you or your apparent allies that threatens the commanded undead (regardless 
of its Intelligence) breaks the spell.

Your commands are not telepathic. The undead creature must be able to hear you.

\textit{Material Component: }A shred of raw meat and a splinter of bone.

\vspace{12pt}
Commune

Divination

\textbf{Level:} Clr 5

\textbf{Components:} V, S, M, DF, XP

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 1 round/level

You contact your deity---or agents thereof ---and ask questions that can be answered 
by a simple yes or no. (A cleric of no particular deity contacts a philosophically 
allied deity.) You are allowed one such question per caster level. The answers 
given are correct within the limits of the entity's knowledge. ``Unclear'' is a 
legitimate answer, because powerful beings of the Outer Planes are not necessarily 
omniscient. In cases where a one-word answer would be misleading or contrary to 
the deity's interests, a short phrase (five words or less) may be given as an answer 
instead.

The spell, at best, provides information to aid character decisions. The entities 
contacted structure their answers to further their own purposes. If you lag, discuss 
the answers, or go off to do anything else, the spell ends.

\textit{Material Component: }Holy (or unholy) water and incense.

\textit{XP Cost: }100 XP.

\vspace{12pt}
Commune with Nature

Divination

\textbf{Level:} Animal 5, Drd 5, Rgr 4

\textbf{Components:} V, S

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} Instantaneous

You become one with nature, attaining knowledge of the surrounding territory. You 
instantly gain knowledge of as many as three facts from among the following subjects: 
the ground or terrain, plants, minerals, bodies of water, people, general animal 
population, presence of woodland creatures, presence of powerful unnatural creatures, 
or even the general state of the natural setting.

In outdoor settings, the spell operates in a radius of 1 mile per caster level. 
In natural underground settings---caves, caverns, and the like---the radius is 
limited to 100 feet per caster level. The spell does not function where nature 
has been replaced by construction or settlement, such as in dungeons and towns.

\vspace{12pt}
Comprehend Languages

Divination

\textbf{Level:} Brd 1, Clr 1, Sor/Wiz 1

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 10 min./level

You can understand the spoken words of creatures or read otherwise incomprehensible 
written messages. In either case, you must touch the creature or the writing. The 
ability to read does not necessarily impart insight into the material, merely its 
literal meaning. The spell enables you to understand or read an unknown language, 
not speak or write it.

Written material can be read at the rate of one page (250 words) per minute. Magical 
writing cannot be read, though the spell reveals that it is magical. This spell 
can be foiled by certain warding magic (such as the \textit{secret page }and \textit{illusory 
script }spells). It does not decipher codes or reveal messages concealed in otherwise 
normal text.

\textit{Comprehend languages }can be made permanent with a \textit{permanency }spell.

\textit{Arcane Material Component: }A pinch of soot and a few grains of salt.

\vspace{12pt}
Cone of Cold

Evocation [Cold]

\textbf{Level:} Sor/Wiz 5, Water 6

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 60 ft.

\textbf{Area:} Cone-shaped burst

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} Reflex half

\textbf{Spell Resistance:} Yes

\textit{Cone of cold }creates an area of extreme cold, originating at your hand 
and extending outward in a cone. It drains heat, dealing 1d6 points of cold damage 
per caster level (maximum 15d6).

\textit{Arcane Material Component: }A very small crystal or glass cone.

\vspace{12pt}
Confusion

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Brd 3, Sor/Wiz 4, Trickery 4

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Targets:} All creatures in a 15-ft. radius burst

\textbf{Duration:} 1 round/level

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

This spell causes the targets to become \textit{confused, }making them unable to 
independently determine what they will do.

Roll on the following table at the beginning of each subject's turn each round 
to see what the subject does in that round.

\begin{tabular}{|>{\raggedright}p{25pt}|>{\raggedright}p{301pt}|}
\hline
d\textbf{\%} & B\textbf{ehavior}\tabularnewline
\hline
01-10 & Attack caster with melee or ranged weapons (or close with caster if attack 
is not possible).\tabularnewline
\hline
11-20 & Act normally.\tabularnewline
\hline
21-50 & Do nothing but babble incoherently.\tabularnewline
\hline
51-70 & Flee away from caster at top possible speed.\tabularnewline
\hline
71-100 & Attack nearest creature (for this purpose, a familiar counts as part of 
the subject's self).\tabularnewline
\hline
\end{tabular}

A \textit{confused }character who can't carry out the indicated action does nothing 
but babble incoherently. Attackers are not at any special advantage when attacking 
a \textit{confused }character. Any \textit{confused }character who is attacked 
automatically attacks its attackers on its next turn, as long as it is still \textit{confused 
}when its turn comes. Note that a \textit{confused }character will not make attacks 
of opportunity against any creature that it is not already devoted to attacking 
(either because of its most recent action or because it has just been attacked).

\textit{Arcane Material Component: }A set of three nut shells.

\vspace{12pt}
Confusion, Lesser

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Brd 1

\textbf{Components:} V, S, DF

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One living creature

\textbf{Duration:} 1 round

This spell causes a single creature to become \textit{confused }for 1 round. See 
the \textit{confusion }spell, above, to determine the exact effect on the subject.

\vspace{12pt}
Consecrate

Evocation [Good]

\textbf{Level:} Clr 2

\textbf{Components:} V, S, M, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Area:} 20-ft.-radius emanation

\textbf{Duration:} 2 hours/level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell blesses an area with positive energy. Each Charisma check made to turn 
undead within this area gains a +3 sacred bonus. Every undead creature entering 
a \textit{consecrated }area suffers minor disruption, giving it a -1 penalty on 
attack rolls, damage rolls, and saves. Undead cannot be created within or summoned 
into a \textit{consecrated }area.

If the \textit{consecrated }area contains an altar, shrine, or other permanent 
fixture dedicated to your deity, pantheon, or aligned higher power, the modifiers 
given above are doubled (+6 sacred bonus on turning checks, -2 penalties for undead 
in the area). You cannot consecrate an area with a similar fixture of a deity other 
than your own patron.

If the area does contain an altar, shrine, or other permanent fixture of a deity, 
pantheon, or higher power other than your patron, the \textit{consecrate }spell 
instead curses the area, cutting off its connection with the associated deity or 
power. This secondary function, if used, does not also grant the bonuses and penalties 
relating to undead, as given above.

\textit{Consecrate }counters and dispels \textit{desecrate}.

\textit{Material Component: }A vial of holy water and 25 gp worth (5 pounds) of 
silver dust, all of which must be sprinkled around the area.

\vspace{12pt}
Contact Other Plane

Divination

\textbf{Level:} Sor/Wiz 5

\textbf{Components:} V

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} Concentration

You send your mind to another plane of existence (an Elemental Plane or some plane 
farther removed) in order to receive advice and information from powers there. 
(See the accompanying table for possible consequences and results of the attempt.) 
The powers reply in a language you understand, but they resent such contact and 
give only brief answers to your questions. (All questions are answered with ``yes,'' 
``no,'' ``maybe,'' ``never,'' ``irrelevant,'' or some other one-word answer.)

You must concentrate on maintaining the spell (a standard action) in order to ask 
questions at the rate of one per round. A question is answered by the power during 
the same round. For every two caster levels, you may ask one question.

Contact with minds far removed from your home plane increases the probability that 
you will incur a decrease to Intelligence and Charisma, but the chance of the power 
knowing the answer, as well as the probability of the entity answering correctly, 
are likewise increased by moving to distant planes.

Once the Outer Planes are reached, the power of the deity contacted determines 
the effects. (Random results obtained from the table are subject to the personalities 
of individual deities.)

On rare occasions, this divination may be blocked by an act of certain deities 
or forces.

\begin{tabular}{|>{\raggedright}p{82pt}|>{\raggedright}p{64pt}|>{\raggedright}p{35pt}|>{\raggedright}p{32pt}|>{\raggedright}p{19pt}|>{\raggedright}p{45pt}|}
\hline
P\textbf{lane Contacted} & A\textbf{void Int/Cha Decrease} & T\textbf{rue Answer} & D\textbf{on't 
Know} & L\textbf{ie} & R\textbf{andom Answer}\tabularnewline
\hline
Elemental Plane & DC 7/1 week & 01-34 & 35-62 & 63-83 & 84-100\tabularnewline
\hline
(appropriate) & (DC 7/1 week) & (01-68) & (69-75) & (76-98) & (99-100)\tabularnewline
\hline
Positive/Negative Energy Plane & DC 8/1 week & 01-39 & 40-65 & 66-86 & 87-100\tabularnewline
\hline
Astral Plane & DC 9/1 week & 01-44 & 45-67 & 68-88 & 89-100\tabularnewline
\hline
Outer Plane, demideity & DC 10/2 weeks & 01-49 & 50-70 & 71-91 & 92-100\tabularnewline
\hline
Outer Plane, lesser deity & DC 12/3 weeks & 01-60 & 61-75 & 76-95 & 96-100\tabularnewline
\hline
Outer Plane, intermediate deity & DC 14/4 weeks & 01-73 & 74-81 & 82-98 & 99-100\tabularnewline
\hline
Outer Plane, greater deity & DC 16/5 weeks & 01-88 & 89-90 & 91-99 & 100\tabularnewline
\hline
\end{tabular}

\textit{Avoid Int/Cha Decrease: }You must succeed on an Intelligence check against 
this DC to avoid a decrease in Intelligence and Charisma. If the check fails, your 
Intelligence and Charisma scores each fall to 8 for the stated duration, and you 
become unable to cast arcane spells. If you lose Intelligence and Charisma, the 
effect strikes as soon as the first question is asked, and no answer is received. 
(The entries in parentheses are for questions that pertain to the appropriate Elemental 
Plane.)

\textbf{Results of a Successful Contact:} d\% is rolled for the result shown on 
the table:

\textit{True Answer: }You get a true, one-word answer. Questions that cannot be 
answered in this way are answered randomly.

\textit{Don't Know: }The entity tells you that it doesn't know.

\textit{Lie: }The entity intentionally lies to you.

\textit{Random Answer: }The entity tries to lie but doesn't know the answer, so 
it makes one up.

\vspace{12pt}
Contagion

Necromancy [Evil]

\textbf{Level:} Clr 3, Destruction 3, Drd 3, Sor/Wiz 4

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Living creature touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} Fortitude negates

\textbf{Spell Resistance:} Yes

The subject contracts a disease selected from the table below, which strikes immediately 
(no incubation period). The DC noted is for the subsequent saves (use \textit{contagion'}s 
normal save DC for the initial saving throw).

\begin{tabular}{|>{\raggedright}p{143pt}|>{\raggedright}p{26pt}|>{\raggedright}p{87pt}|}
\hline
\subsection*{D\textbf{isease}} & \subsection*{D\textbf{C}} & \subsection*{D\textbf{amage}}\tabularnewline
\hline
Blinding sickness & 16 & 1d4 Str\textsuperscript{\textbf{1}}\tabularnewline
\hline
Cackle fever & 16 & 1d6 Wis\tabularnewline
\hline
Filth fever & 12 & 1d3 Dex and 1d3 Con\tabularnewline
\hline
Mindfire & 12 & 1d4 Int\tabularnewline
\hline
Red ache & 15 & 1d6 Str\tabularnewline
\hline
Shakes & 13 & 1d8 Dex\tabularnewline
\hline
Slimy doom & 14 & 1d4 Con\tabularnewline
\hline
\multicolumn{3}{|p{257pt}|}{1 Each time a victim takes 2 or more points of Strength 
damage from blinding sickness, he or she must make another Fortitude save (using 
the disease's save DC) or be permanently blinded.}\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
Contingency

Evocation

\textbf{Level:} Sor/Wiz 6

\textbf{Components:} V, S, M, F

\textbf{Casting Time:} At least 10 minutes; see text

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} One day/level (D) or until discharged

You can place another spell upon your person so that it comes into effect under 
some condition you dictate when casting \textit{contingency}. The \textit{contingency 
}spell and the companion spell are cast at the same time. The 10-minute casting 
time is the minimum total for both castings; if the companion spell has a casting 
time longer than 10 minutes, use that instead.

The spell to be brought into effect by the \textit{contingency }must be one that 
affects your person and be of a spell level no higher than one-third your caster 
level (rounded down, maximum 6th level).

The conditions needed to bring the spell into effect must be clear, although they 
can be general. In all cases, the \textit{contingency }immediately brings into 
effect the companion spell, the latter being ``cast'' instantaneously when the 
prescribed circumstances occur. If complicated or convoluted conditions are prescribed, 
the whole spell combination (\textit{contingency }and the companion magic) may 
fail when called on. The companion spell occurs based solely on the stated conditions, 
regardless of whether you want it to.

You can use only one \textit{contingency }spell at a time; if a second is cast, 
the first one (if still active) is dispelled.

\textit{Material Component: }That of the companion spell, plus quicksilver and 
an eyelash of an ogre mage, rakshasa, or similar spell-using creature.

\textit{Focus: }A statuette of you carved from elephant ivory and decorated with 
gems (worth at least 1,500 gp). You must carry the focus for the \textit{contingency 
}to work.

\vspace{12pt}
Continual Flame

Evocation [Light]

\textbf{Level:} Clr 3, Sor/Wiz 2

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Object touched

\textbf{Effect: }Magical, heatless flame

\textbf{Duration:} Permanent

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

A flame, equivalent in brightness to a torch, springs forth from an object that 
you touch. The effect looks like a regular flame, but it creates no heat and doesn't 
use oxygen. A \textit{continual flame }can be covered and hidden but not smothered 
or quenched.

Light spells counter and dispel darkness spells of an equal or lower level.

\textit{Material Component: }You sprinkle ruby dust (worth 50 gp) on the item that 
is to carry the flame.

\vspace{12pt}
Control Plants

Transmutation

\textbf{Level:} Drd 8, Plant 8

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} Up to 2 HD/level of plant creatures, no two of which can be more 
than 30 ft. apart

\textbf{Duration:} 1 min./level

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} No

This spell enables you to control the actions of one or more plant creatures for 
a short period of time. You command the creatures by voice and they understand 
you, no matter what language you speak. Even if vocal communication is impossible 
 the controlled plants do not attack you. At the end of the spell, the subjects 
revert to their normal behavior.

Suicidal or self-destructive commands are simply ignored.

\vspace{12pt}
Control Undead

Necromancy

\textbf{Level:} Sor/Wiz 7

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} Up to 2 HD/level of undead creatures, no two of which can be 
more than 30 ft. apart

\textbf{Duration:} 1 min./level

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

This spell enables you to command undead creatures for a short period of time. 
You command them by voice and they understand you, no matter what language you 
speak. Even if vocal communication is impossible the controlled undead do not attack 
you. At the end of the spell, the subjects revert to their normal behavior.

Intelligent undead creatures remember that you controlled them.

\textit{Material Component: }A small piece of bone and a small piece of raw meat.

\vspace{12pt}
Control Water

Transmutation [Water]

\textbf{Level:} Clr 4, Drd 4, Sor/Wiz 6, Water 4

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Area: }Water in a volume of 10 ft./level by 10 ft./level by 2 ft./level 
(S)

\textbf{Duration:} 10 min./level (D)

\textbf{Saving Throw:} None; see text

\textbf{Spell Resistance:} No

Depending on the version you choose, the \textit{control water }spell raises or 
lowers water.

\textit{Lower Water: }This causes water or similar liquid to reduce its depth by 
as much as 2 feet per caster level (to a minimum depth of 1 inch). The water is 
lowered within a squarish depression whose sides are up to caster level x 10 feet 
long. In extremely large and deep bodies of water, such as a deep ocean, the spell 
creates a whirlpool that sweeps ships and similar craft downward, putting them 
at risk and rendering them unable to leave by normal movement for the duration 
of the spell. When cast on water elementals and other water-based creatures, this 
spell acts as a \textit{slow }spell (Will negates). The spell has no effect on 
other creatures.

\textit{Raise Water: }This causes water or similar liquid to rise in height, just 
as the \textit{lower water }version causes it to lower. Boats raised in this way 
slide down the sides of the hump that the spell creates. If the area affected by 
the spell includes riverbanks, a beach, or other land nearby, the water can spill 
over onto dry land.

With either version, you may reduce one horizontal dimension by half and double 
the other horizontal dimension.

\textit{Arcane Material Component: }A drop of water (for \textit{raise water}) 
or a pinch of dust (for \textit{lower water}).

\vspace{12pt}
Control Weather

Transmutation

\textbf{Level:} Air 7, Clr 7, Drd 7, Sor/Wiz 7

\textbf{Components:} V, S

\textbf{Casting Time:} 10 minutes; see text

\textbf{Range:} 2 miles

\textbf{Area:} 2-mile-radius circle, centered on you; see text

\textbf{Duration:} 4d12 hours; see text

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You change the weather in the local area. It takes 10 minutes to cast the spell 
and an additional 10 minutes for the effects to manifest.\textit{ }You can call 
forth weather appropriate to the climate and season of the area you are in.

\begin{tabular}{|>{\raggedright}p{44pt}|>{\raggedright}p{205pt}|}
\hline
S\textbf{eason} & P\textbf{ossible Weather}\tabularnewline
\hline
Spring & Tornado, thunderstorm, sleet storm, or hot weather\tabularnewline
\hline
Summer & Torrential rain, heat wave, or hailstorm\tabularnewline
\hline
Autumn & Hot or cold weather, fog, or sleet\tabularnewline
\hline
Winter & Frigid cold, blizzard, or thaw\tabularnewline
\hline
Late winter & Hurricane-force winds or early spring (coastal area)\tabularnewline
\hline
\end{tabular}

You control the general tendencies of the weather, such as the direction and intensity 
of the wind. You cannot control specific applications of the weather---where lightning 
strikes, for example, or the exact path of a tornado. When you select a certain 
weather condition to occur, the weather assumes that condition 10 minutes later 
(changing gradually, not abruptly). The weather continues as you left it for the 
duration, or until you use a standard action to designate a new kind of weather 
(which fully manifests itself 10 minutes later). Contradictory conditions are not 
possible simultaneously.

\textit{Control weather }can do away with atmospheric phenomena (naturally occurring 
or otherwise) as well as create them.

A druid casting this spell doubles the duration and affects a circle with a 3-mile 
radius.

\vspace{12pt}
Control Winds

Transmutation [Air]

\textbf{Level:} Air 5, Drd 5

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 40 ft./level

\textbf{Area:} 40 ft./level radius cylinder 40 ft. high

\textbf{Duration:} 10 min./level

\textbf{Saving Throw:} Fortitude negates

\textbf{Spell Resistance:} No

You alter wind force in the area surrounding you. You can make the wind blow in 
a certain direction or manner, increase its strength, or decrease its strength. 
The new wind direction and strength persist until the spell ends or until you choose 
to alter your handiwork, which requires concentration. You may create an ``eye'' 
of calm air up to 80 feet in diameter at the center of the area if you so desire, 
and you may choose to limit the area to any cylindrical area less than your full 
limit.

\textit{Wind Direction: }You may choose one of four basic wind patterns to function 
over the spell's area. 

A downdraft blows from the center outward in equal strength in all directions. 

An updraft blows from the outer edges in toward the center in equal strength from 
all directions, veering upward before impinging on the eye in the center. 

A rotation causes the winds to circle the center in clockwise or counterclockwise 
fashion. 

A blast simply causes the winds to blow in one direction across the entire area 
from one side to the other.

\textit{Wind Strength: }For every three caster levels, you can increase or decrease 
wind strength by one level. Each round on your turn, a creature in the wind must 
make a Fortitude save or suffer the effect of being in the windy area.

Strong winds (21+ mph) make sailing difficult.

A severe wind (31+ mph) causes minor ship and building damage.

A windstorm (51+ mph) drives most flying creatures from the skies, uproots small 
trees, knocks down light wooden structures, tears off roofs, and endangers ships.

Hurricane force winds (75+ mph) destroy wooden buildings, sometimes uproot even 
large trees, and cause most ships to founder.

A tornado (175+ mph) destroys all nonfortified buildings and often uproots large 
trees.

\vspace{12pt}
Create Food and Water

Conjuration (Creation)

\textbf{Level:} Clr 3

\textbf{Components:} V, S

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} Food and water to sustain three humans or one horse/level for 
24 hours

\textbf{Duration:} 24 hours; see text

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

The food that this spell creates is simple fare of your choice---highly nourishing, 
if rather bland. Food so created decays and becomes inedible within 24 hours, although 
it can be kept fresh for another 24 hours by casting a \textit{purify food and 
water }spell on it. The water created by this spell is just like clean rain water, 
and it doesn't go bad as the food does.

\vspace{12pt}
Create Greater Undead

Necromancy [Evil]

\textbf{Level:} Clr 8, Death 8, Sor/Wiz 8

This spell functions like create undead, except that you can create more powerful 
and intelligent sorts of undead: shadows, wraiths, spectres, and devourers. The 
type or types of undead you can create is based on your caster level, as shown 
on the table below.

\begin{tabular}{|>{\raggedright}p{55pt}|>{\raggedright}p{69pt}|}
\hline
C\textbf{aster Level} & U\textbf{ndead Created}\tabularnewline
\hline
15th or lower & Shadow\tabularnewline
\hline
16th-17th & Wraith\tabularnewline
\hline
18th-19th & Spectre\tabularnewline
\hline
20th or higher & Devourer\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
Create Undead

Necromancy [Evil]

\textbf{Level:} Clr 6, Death 6, Evil 6, Sor/Wiz 6

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 hour

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One corpse

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

A much more potent spell than \textit{animate dead, }this evil spell allows you 
to create more powerful sorts of undead: ghouls, ghasts, mummies, and mohrgs. The 
type or types of undead you can create is based on your caster level, as shown 
on the table below.

\begin{tabular}{|>{\raggedright}p{55pt}|>{\raggedright}p{69pt}|}
\hline
\subsection*{C\textbf{aster Level}} & \subsection*{U\textbf{ndead Created}}\tabularnewline
\hline
11th or lower & Ghoul\tabularnewline
\hline
12th-14th & Ghast\tabularnewline
\hline
15th-17th & Mummy\tabularnewline
\hline
18th or higher & Mohrg\tabularnewline
\hline
\end{tabular}

You may create less powerful undead than your level would allow if you choose. 
Created undead are not automatically under the control of their animator. If you 
are capable of commanding undead, you may attempt to command the undead creature 
as it forms.

This spell must be cast at night.

\textit{Material Component: }A clay pot filled with grave dirt and another filled 
with brackish water. The spell must be cast on a dead body. You must place a black 
onyx gem worth at least 50 gp per HD of the undead to be created into the mouth 
or eye socket of each corpse. The magic of the spell turns these gems into worthless 
shells.

\vspace{12pt}
Create Water

Conjuration (Creation) [Water]

\textbf{Level:} Clr 0, Drd 0, Pal 1

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} Up to 2 gallons of water/level

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell generates wholesome, drinkable water, just like clean rain water. Water 
can be created in an area as small as will actually contain the liquid, or in an 
area three times as large---possibly creating a downpour or filling many small 
receptacles.

\textit{Note: }Conjuration spells can't create substances or objects within a creature. 
Water weighs about 8 pounds per gallon. One cubic foot of water contains roughly 
8 gallons and weighs about 60 pounds.

\vspace{12pt}
Creeping Doom

Conjuration (Summoning)

\textbf{Level:} Drd 7

\textbf{Components:} V, S

\textbf{Casting Time:} 1 round

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)/ 100 ft.; see text

\textbf{Effect:} One swarm of centipedes per two levels

\textbf{Duration:} 1 min./level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

When you utter the spell of \textit{creeping doom, }you call forth a mass of centipede 
swarms (one per two caster levels, to a maximum of ten swarms at 20th level), which 
need not appear adjacent to one another.

You may summon the centipede swarms so that they share the area of other creatures. 
The swarms remain stationary, attacking any creatures in their area, unless you 
command the creeping doom to move (a standard action). As a standard action, you 
can command any number of the swarms to move toward any prey within 100 feet of 
you. You cannot command any swarm to move more than 100 feet away from you, and 
if you move more than 100 feet from any swarm, that swarm remains stationary, attacking 
any creatures in its area (but it can be commanded again if you move within 100 
feet).

\vspace{12pt}
Crushing Despair

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Brd 3, Sor/Wiz 4

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 30 ft.

\textbf{Area:} Cone-shaped burst

\textbf{Duration:} 1 min./level

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

An invisible cone of despair causes great sadness in the subjects. Each affected 
creature takes a -2 penalty on attack rolls, saving throws, ability checks, skill 
checks, and weapon damage rolls.

\textit{Crushing despair }counters and dispels \textit{good hope}.

\textit{Material Component: }A vial of tears.

\vspace{12pt}
Crushing Hand

Evocation [Force]

\textbf{Level:} Sor/Wiz 9, Strength 9

\textbf{Components:} V, S, M, F/DF

This spell functions like \textit{interposing hand, }except that the hand can interpose 
itself, push, or crush one opponent that you select.

The \textit{crushing hand }can grapple an opponent like \textit{grasping hand }does. 
Its grapple bonus equals your caster level + your Intelligence, Wisdom, or Charisma 
modifier (for a wizard, cleric, or sorcerer, respectively), +12 for the hand's 
Strength score (35), +4 for being Large. The hand deals 2d6+12 points of damage 
(lethal, not nonlethal) on each successful grapple check against an opponent.

The \textit{crushing hand }can also interpose itself as \textit{interposing hand 
}does, or it can bull rush an opponent as \textit{forceful hand }does, but at a 
+18 bonus.

Directing the spell to a new target is a move action.

Clerics who cast this spell name it for their deities.

\textit{Arcane Material Component: }The shell of an egg.

\textit{Arcane Focus: }A glove of snakeskin.

\vspace{12pt}
Cure Critical Wounds

Conjuration (Healing)

\textbf{Level:} Brd 4, Clr 4, Drd 5, Healing 4

This spell functions like \textit{cure light wounds}, except that it cures 4d8 
points of damage +1 point per caster level (maximum +20).

\vspace{12pt}
Cure Critical Wounds, Mass

Conjuration (Healing)

\textbf{Level:} Clr 8, Drd 9, Healing 8

This spell functions like \textit{mass cure light wounds}, except that it cures 
4d8 points of damage +1 point per caster level (maximum +40).

\vspace{12pt}
Cure Light Wounds

Conjuration (Healing)

\textbf{Level:} Brd 1, Clr 1, Drd 1, Healing 1, Pal 1, Rgr 2

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will half (harmless); see text

\textbf{Spell Resistance:} Yes (harmless); see text

When laying your hand upon a living creature, you channel positive energy that 
cures 1d8 points of damage +1 point per caster level (maximum +5).

Since undead are powered by negative energy, this spell deals damage to them instead 
of curing their wounds. An undead creature can apply spell resistance, and can 
attempt a Will save to take half damage.

\vspace{12pt}
Cure Light Wounds, Mass

Conjuration (Healing)

\textbf{Level:} Brd 5, Clr 5, Drd 6, Healing 5

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One creature/level, no two of which can be more than 30 ft. apart

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will half (harmless) or Will half; see text

\textbf{Spell Resistance:} Yes (harmless) or Yes; see text

You channel positive energy to cure 1d8 points of damage +1 point per caster level 
(maximum +25) in each selected creature.

Like other \textit{cure }spells, \textit{mass cure light wounds }deals damage to 
undead in its area rather than curing them. Each affected undead may attempt a 
Will save for half damage.

\vspace{12pt}
Cure Minor Wounds

Conjuration (Healing)

\textbf{Level:} Clr 0, Drd 0

This spell functions like \textit{cure light wounds}, except that it cures only 
1 point of damage.

\vspace{12pt}
Cure Moderate Wounds

Conjuration (Healing)

\textbf{Level:} Brd 2, Clr 2, Drd 3, Healing 2, Pal 3, Rgr 3

This spell functions like \textit{cure light wounds}, except that it cures 2d8 
points of damage +1 point per caster level (maximum +10).

\vspace{12pt}
Cure Moderate Wounds, Mass

Conjuration (Healing)

\textbf{Level:} Brd 6, Clr 6, Drd 7

This spell functions like \textit{mass cure light wounds}, except that it cures 
2d8 points of damage +1 point per caster level (maximum +30).

\vspace{12pt}
Cure Serious Wounds

Conjuration (Healing)

\textbf{Level:} Brd 3, Clr 3, Drd 4, Pal 4, Rgr 4, Healing 3

This spell functions like \textit{cure light wounds}, except that it cures 3d8 
points of damage +1 point per caster level (maximum +15).

\vspace{12pt}
Cure Serious Wounds, Mass

Conjuration (Healing)

\textbf{Level:} Clr 7, Drd 8

This spell functions like \textit{mass cure light wounds}, except that it cures 
3d8 points of damage +1 point per caster level (maximum +35).

\vspace{12pt}
Curse Water

Necromancy [Evil]

\textbf{Level:} Clr 1

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 minute

\textbf{Range:} Touch

\textbf{Target:} Flask of water touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates (object)

\textbf{Spell Resistance:} Yes (object)

This spell imbues a flask (1 pint) of water with negative energy, turning it into 
unholy water. Unholy water damages good outsiders the way holy water damages undead 
and evil outsiders.

\textit{Material Component: }5 pounds of powdered silver (worth 25 gp).

\newpage

\end{document}
