%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{array}
\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

This material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

{\LARGE{}SPELLS (A-B)}

\vspace{12pt}
Acid Arrow

Conjuration (Creation) [Acid]

\textbf{Level:} Sor/Wiz 2

\textbf{Components:} V, S, M, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Effect:} One arrow of acid

\textbf{Duration:} 1 round + 1 round per three levels

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

A magical arrow of acid springs from your hand and speeds to its target. You must 
succeed on a ranged touch attack to hit your target. The arrow deals 2d4 points 
of acid damage with no splash damage. For every three caster levels (to a maximum 
of 18th), the acid, unless somehow neutralized, lasts for another round, dealing 
another 2d4 points of damage in that round.

\textit{Material Component: }Powdered rhubarb leaf and an adder's stomach.

\textit{Focus: }A dart.

\vspace{12pt}
Acid Fog

Conjuration (Creation) [Acid]

\textbf{Level:} Sor/Wiz 6, Water 7

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Effect:} Fog spreads in 20-ft. radius, 20 ft. high

\textbf{Duration:} 1 round/level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

\textit{Acid fog }creates a billowing mass of misty vapors similar to that produced 
by a \textit{solid fog }spell. In addition to slowing creatures down and obscuring 
sight, this spell's vapors are highly acidic. Each round on your turn, starting 
when you cast the spell, the fog deals 2d6 points of acid damage to each creature 
and object within it.

\textit{Arcane Material Component:} A pinch of dried, powdered peas combined with 
powdered animal hoof.

\vspace{12pt}
Acid Splash

Conjuration (Creation) [Acid]

\textbf{Level:} Sor/Wiz 0

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} One missile of acid

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You fire a small orb of acid at the target. You must succeed on a ranged touch 
attack to hit your target. The orb deals 1d3 points of acid damage.

\vspace{12pt}
Aid

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Clr 2, Good 2, Luck 2

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Living creature touched

\textbf{Duration:} 1 min./level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes (harmless)

\textit{Aid }grants the target a +1 morale bonus on attack rolls and saves against 
fear effects, plus temporary hit points equal to 1d8 + caster level (to a maximum 
of 1d8+10 temporary hit points at caster level 10th).

\vspace{12pt}
Air Walk

Transmutation [Air]

\textbf{Level:} Air 4, Clr 4, Drd 4

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature (Gargantuan or smaller) touched

\textbf{Duration:} 10 min./level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes (harmless)

The subject can tread on air as if walking on solid ground. Moving upward is similar 
to walking up a hill. The maximum upward or downward angle possible is 45 degrees, 
at a rate equal to one-half the air walker's normal speed.

A strong wind (21+ mph) can push the subject along or hold it back. At the end 
of its turn each round, the wind blows the air walker 5 feet for each 5 miles per 
hour of wind speed. The creature may be subject to additional penalties in exceptionally 
strong or turbulent winds, such as loss of control over movement or physical damage 
from being buffeted about.

Should the spell duration expire while the subject is still aloft, the magic fails 
slowly. The subject floats downward 60 feet per round for 1d6 rounds. If it reaches 
the ground in that amount of time, it lands safely. If not, it falls the rest of 
the distance, taking 1d6 points of damage per 10 feet of fall. Since dispelling 
a spell effectively ends it, the subject also descends in this way if the \textit{air 
walk }spell is dispelled, but not if it is negated by an \textit{antimagic field}.

You can cast \textit{air walk }on a specially trained mount so it can be ridden 
through the air. You can train a mount to move with the aid of \textit{air walk 
}(counts as a trick; see Handle Animal skill) with one week of work and a DC 25 
Handle Animal check.

\vspace{12pt}
Alarm

Abjuration

\textbf{Level:} Brd 1, Rgr 1, Sor/Wiz 1

\textbf{Components:} V, S, F/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Area:} 20-ft.-radius emanation centered on a point in space

\textbf{Duration:} 2 hours/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

A\textit{larm }sounds a mental or audible alarm each time a creature of Tiny or 
larger size enters the warded area or touches it. A creature that speaks the password 
(determined by you at the time of casting) does not set off the \textit{alarm}. 
You decide at the time of casting whether the \textit{alarm }will be mental or 
audible.

\textit{Mental Alarm: }A mental \textit{alarm }alerts you (and only you) so long 
as you remain within 1 mile of the warded area. You note a single mental ``ping'' 
that awakens you from normal sleep but does not otherwise disturb concentration. 
A \textit{silence }spell has no effect on a mental \textit{alarm}.

\textit{Audible Alarm: }An audible \textit{alarm }produces the sound of a hand 
bell, and anyone within 60 feet of the warded area can hear it clearly. Reduce 
the distance by 10 feet for each interposing closed door and by 20 feet for each 
substantial interposing wall.

In quiet conditions, the ringing can be heard faintly as far as 180 feet away. 
The sound lasts for 1 round. Creatures within a \textit{silence }spell cannot hear 
the ringing.

Ethereal or astral creatures do not trigger the \textit{alarm}.

\textit{Alarm }can be made permanent with a \textit{permanency }spell.

\textit{Arcane Focus:} A tiny bell and a piece of very fine silver wire

\vspace{12pt}
Align Weapon

Transmutation [see text]

\textbf{Level:} Clr 2

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target: }Weapon touched or fifty projectiles (all of which must be in contact 
with each other at the time of casting)

\textbf{Duration:} 1 min./level

\textbf{Saving Throw: }Will negates (harmless, object)

\textbf{Spell Resistance:} Yes (harmless, object)

\textit{Align weapon }makes a weapon good, evil, lawful, or chaotic, as you choose. 
A weapon that is aligned can bypass the damage reduction of certain creatures. 
This spell has no effect on a weapon that already has an alignment.

You can't cast this spell on a natural weapon, such as an unarmed strike.

When you make a weapon good, evil, lawful, or chaotic, \textit{align weapon }is 
a good, evil, lawful, or chaotic spell, respectively.

\vspace{12pt}
Alter Self

Transmutation

\textbf{Level:} Brd 2, Sor/Wiz 2

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 10 min./level (D)

You assume the form of a creature of the same type as your normal form. The new 
form must be within one size category of your normal size. The maximum HD of an 
assumed form is equal to your caster level, to a maximum of 5 HD at 5th level. 
You can change into a member of your own kind or even into yourself.

You retain your own ability scores. Your class and level, hit points, alignment, 
base attack bonus, and base save bonuses all remain the same. You retain all supernatural 
and spell-like special attacks and qualities of your normal form, except for those 
requiring a body part that the new form does not have (such as a mouth for a breath 
weapon or eyes for a gaze attack).

You keep all extraordinary special attacks and qualities derived from class levels, 
but you lose any from your normal form that are not derived from class levels.

If the new form is capable of speech, you can communicate normally. You retain 
any spellcasting ability you had in your original form, but the new form must be 
able to speak intelligibly (that is, speak a language) to use verbal components 
and must have limbs capable of fine manipulation to use somatic or material components.

You acquire the physical qualities of the new form while retaining your own mind. 
Physical qualities include natural size, mundane movement capabilities (such as 
burrowing, climbing, walking, swimming, and flight with wings, to a maximum speed 
of 120 feet for flying or 60 feet for nonflying movement), natural armor bonus, 
natural weapons (such as claws, bite, and so on), racial skill bonuses, racial 
bonus feats, and any gross physical qualities (presence or absence of wings, number 
of extremities, and so forth). A body with extra limbs does not allow you to make 
more attacks (or more advantageous two-weapon attacks) than normal.

You do not gain any extraordinary special attacks or special qualities not noted 
above under physical qualities, such as darkvision, low-light vision, blindsense, 
blindsight, fast healing, regeneration, scent, and so forth.

You do not gain any supernatural special attacks, special qualities, or spell-like 
abilities of the new form. Your creature type and subtype (if any) remain the same 
regardless of your new form. You cannot take the form of any creature with a template, 
even if that template doesn't change the creature type or subtype.

You can freely designate the new form's minor physical qualities (such as hair 
color, hair texture, and skin color) within the normal ranges for a creature of 
that kind. The new form's significant physical qualities (such as height, weight, 
and gender) are also under your control, but they must fall within the norms for 
the new form's kind. You are effectively disguised as an average member of the 
new form's race. If you use this spell to create a disguise, you get a +10 bonus 
on your Disguise check.

When the change occurs, your equipment, if any, either remains worn or held by 
the new form (if it is capable of wearing or holding the item), or melds into the 
new form and becomes nonfunctional. When you revert to your true form, any objects 
previously melded into the new form reappear in the same location on your body 
they previously occupied and are once again functional. Any new items you wore 
in the assumed form and can't wear in your normal form fall off and land at your 
feet; any that you could wear in either form or carry in a body part common to 
both forms at the time of reversion are still held in the same way. Any part of 
the body or piece of equipment that is separated from the whole reverts to its 
true form.

\vspace{12pt}
Analyze Dweomer

Divination

\textbf{Level:} Brd 6, Sor/Wiz 6

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} One object or creature per caster level

\textbf{Duration:} 1 round/level (D)

\textbf{Saving Throw:} None or Will negates; see text

\textbf{Spell Resistance:} No

You discern all spells and magical properties present in a number of creatures 
or objects. Each round, you may examine a single creature or object that you can 
see as a free action. In the case of a magic item, you learn its functions, how 
to activate its functions (if appropriate), and how many charges are left (if it 
uses charges). In the case of an object or creature with active spells cast upon 
it, you learn each spell, its effect, and its caster level.

An attended object may attempt a Will save to resist this effect if its holder 
so desires. If the save succeeds, you learn nothing about the object except what 
you can discern by looking at it. An object that makes its save cannot be affected 
by any other \textit{analyze dweomer }spells for 24 hours.

\textit{Analyze dweomer }does not function when used on an artifact.

\textit{Focus: }A tiny lens of ruby or sapphire set in a small golden loop. The 
gemstone must be worth at least 1,500 gp.

\vspace{12pt}
Animal Growth

Transmutation

\textbf{Level:} Drd 5, Rgr 4, Sor/Wiz 5

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Targets:} Up to one animal (Gargantuan or smaller) per two levels, no two 
of which can be more than 30 ft. apart

\textbf{Duration:} 1 min./level

\textbf{Saving Throw:} Fortitude negates

\textbf{Spell Resistance:} Yes

A number of animals grow to twice their normal size and eight times their normal 
weight. This alteration changes each animal's size category to the next largest, 
grants it a +8 size bonus to Strength and a +4 size bonus to Constitution (and 
thus an extra 2 hit points per HD), and imposes a -2 size penalty to Dexterity. 
The creature's existing natural armor bonus increases by 2. The size change also 
affects the animal's modifier to AC and attack rolls and its base damage. The animal's 
space and reach change as appropriate to the new size, but its speed does not change.

The spell also grants each subject damage reduction 10/magic and a +4 resistance 
bonus on saving throws. If insufficient room is available for the desired growth, 
the creature attains the maximum possible size and may make a Strength check (using 
its increased Strength) to burst any enclosures in the process. If it fails, it 
is constrained without harm by the materials enclosing it--- the spell cannot be 
used to crush a creature by increasing its size.

All equipment worn or carried by an animal is similarly enlarged by the spell, 
though this change has no effect on the magical properties of any such equipment.

Any enlarged item that leaves the enlarged creature's possession instantly returns 
to its normal size.

The spell gives no means of command or influence over the enlarged animals.

Multiple magical effects that increase size do not stack.

\vspace{12pt}
Animal Messenger

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Brd 2, Drd 2, Rgr 1

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One Tiny animal

\textbf{Duration:} One day/level

\textbf{Saving Throw:} None; see text

\textbf{Spell Resistance:} Yes

You compel a Tiny animal to go to a spot you designate. The most common use for 
this spell is to get an animal to carry a message to your allies. The animal cannot 
be one tamed or trained by someone else, including such creatures as familiars 
and animal companions.

Using some type of food desirable to the animal as a lure, you call the animal 
to you. It advances and awaits your bidding. You can mentally impress on the animal 
a certain place well known to you or an obvious landmark. The directions must be 
simple, because the animal depends on your knowledge and can't find a destination 
on its own. You can attach some small item or note to the messenger. The animal 
then goes to the designated location and waits there until the duration of the 
spell expires, whereupon it resumes its normal activities.

During this period of waiting, the messenger allows others to approach it and remove 
any scroll or token it carries. The intended recipient gains no special ability 
to communicate with the animal or read any attached message (if it's written in 
a language he or she doesn't know, for example).

\textit{Material Component: }A morsel of food the animal likes.

\vspace{12pt}
Animal Shapes

Transmutation

\textbf{Level:} Animal 7, Drd 8

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} Up to one willing creature per level, all within 30 ft. of each 
other

\textbf{Duration:} 1 hour/level (D)

\textbf{Saving Throw:} None; see text

\textbf{Spell Resistance:} Yes (harmless)

As \textit{polymorph, }except you polymorph up to one willing creature per caster 
level into an animal of your choice; the spell has no effect on unwilling creatures. 
All creatures must take the same kind of animal form. Recipients remain in the 
animal form until the spell expires or until you dismiss it for all recipients. 
In addition, an individual subject may choose to resume its normal form as a full-round 
action; doing so ends the spell for that subject alone. The maximum HD of an assumed 
form is equal to the subject's HD or your caster level, whichever is lower, to 
a maximum of 20 HD at 20th level.

\vspace{12pt}
Animal Trance

Enchantment (Compulsion) [Mind-Affecting, Sonic]

\textbf{Level:} Brd 2, Drd 2

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} Animals or magical beasts with Intelligence 1 or 2

\textbf{Duration:} Concentration

\textbf{Saving Throw: }Will negates; see text

\textbf{Spell Resistance:} Yes

Your swaying motions and music (or singing, or chanting) compel animals and magical 
beasts to do nothing but watch you. Only a creature with an Intelligence score 
of 1 or 2 can be fascinated by this spell. Roll 2d6 to determine the total number 
of HD worth of creatures that you fascinate. The closest targets are selected first 
until no more targets within range can be affected.

A magical beast, a dire animal, or an animal trained to attack or guard is allowed 
a saving throw; an animal not trained to attack or guard is not.

\vspace{12pt}
Animate Dead

Necromancy [Evil]

\textbf{Level:} Clr 3, Death 3, Sor/Wiz 4

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Targets:} One or more corpses touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell turns the bones or bodies of dead creatures into undead skeletons or 
zombies that follow your spoken commands.

The undead can follow you, or they can remain in an area and attack any creature 
(or just a specific kind of creature) entering the place. They remain animated 
until they are destroyed. (A destroyed skeleton or zombie can't be animated again.)

Regardless of the type of undead you create with this spell, you can't create more 
HD of undead than twice your caster level with a single casting of \textit{animate 
dead}. (The \textit{desecrate }spell doubles this limit)

The undead you create remain under your control indefinitely. No matter how many 
times you use this spell, however, you can control only 4 HD worth of undead creatures 
per caster level. If you exceed this number, all the newly created creatures fall 
under your control, and any excess undead from previous castings become uncontrolled. 
(You choose which creatures are released.) If you are a cleric, any undead you 
might command by virtue of your power to command or rebuke undead do not count 
toward the limit.

\textit{Skeletons: }A skeleton can be created only from a mostly intact corpse 
or skeleton. The corpse must have bones. If a skeleton is made from a corpse, the 
flesh falls off the bones. 

\textit{Zombies: }A zombie can be created only from a mostly intact corpse. The 
corpse must be that of a creature with a true anatomy.

\textit{Material Component: }You must place a black onyx gem worth at least 25 
gp per Hit Die of the undead into the mouth or eye socket of each corpse you intend 
to animate. The magic of the spell turns these gems into worthless, burned-out 
shells.

\vspace{12pt}
Animate Objects

Transmutation

\textbf{Level:} Brd 6, Chaos 6, Clr 6

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Targets:} One Small object per caster level; see text

\textbf{Duration:} 1 round/level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You imbue inanimate objects with mobility and a semblance of life. Each such animated 
object then immediately attacks whomever or whatever you initially designate.

An animated object can be of any nonmagical material. You may animate one Small 
or smaller object or an equivalent number of larger objects per caster level. A 
Medium object counts as two Small or smaller objects, a Large object as four, a 
Huge object as eight, a Gargantuan object as sixteen, and a Colossal object as 
thirty-two. You can change the designated target or targets as a move action, as 
if directing an active spell.

This spell cannot animate objects carried or worn by a creature.

\textit{Animate objects }can be made permanent with a \textit{permanency }spell.

\vspace{12pt}
Animate Plants

Transmutation

\textbf{Level:} Drd 7, Plant 7

\textbf{Components:} V

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} One Large plant per three caster levels or all plants within 
range; see text

\textbf{Duration:} 1 round/level or 1 hour/level; see text

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You imbue inanimate plants with mobility and a semblance of life. Each animated 
plant then immediately attacks whomever or whatever you initially designate as 
though it were an animated object of the appropriate size category. You may animate 
one Large or smaller plant, or an equivalent number of larger plants, per three 
caster levels. A Huge plant counts as two Large or smaller plants, a Gargantuan 
plant as four, and a Colossal plant as eight. You can change the designated target 
or targets as a move action, as if directing an active spell.

Use the statistics for animated objects\textit{, }except that plants smaller than 
Large usually don't have hardness.

\textit{Animate plants }cannot affect plant creatures, nor does it affect nonliving 
vegetable material.

\textit{Entangle: }Alternatively, you may imbue all plants within range with a 
degree of mobility, which allows them to entwine around creatures in the area. 
This usage of the spell duplicates the effect of an \textit{entangle }spell. Spell 
resistance does not keep creatures from being entangled. This effect lasts 1 hour 
per caster level.

\vspace{12pt}
Animate Rope

Transmutation

\textbf{Level:} Brd 1, Sor/Wiz 1

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Target:} One ropelike object, length up to 50 ft. + 5 ft./level; see text

\textbf{Duration:} 1 round/level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You can animate a nonliving ropelike object. The maximum length assumes a rope 
with a 1-inch diameter.

Reduce the maximum length by 50\% for every additional inch of thickness, and increase 
it by 50\% for each reduction of the rope's diameter by half.

The possible commands are ``coil'' (form a neat, coiled stack), ``coil and knot,'' 
``loop,'' ``loop and knot,'' ``tie and knot,'' and the opposites of all of the 
above (``uncoil,'' and so forth). You can give one command each round as a move 
action, as if directing an active spell.

The rope can enwrap only a creature or an object within 1 foot of it---it does 
not snake outward---so it must be thrown near the intended target. Doing so requires 
a successful ranged touch attack roll (range increment 10 feet). A typical 1- inch-diameter 
hempen rope has 2 hit points, AC 10, and requires a DC 23 Strength check to burst 
it. The rope does not deal damage, but it can be used as a trip line or to cause 
a single opponent that fails a Reflex saving throw to become entangled. A creature 
capable of spellcasting that is bound by this spell must make a DC 15 Concentration 
check to cast a spell. An entangled creature can slip free with a DC 20 Escape 
Artist check.

The rope itself and any knots tied in it are not magical.

This spell grants a +2 bonus on any Use Rope checks you make when using the transmuted 
rope.

The spell cannot animate objects carried or worn by a creature.

\vspace{12pt}
Antilife Shell

Abjuration

\textbf{Level:} Animal 6, Clr 6, Drd 6

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 round

\textbf{Range:} 10 ft.

\textbf{Area:} 10-ft.-radius emanation, centered on you

\textbf{Duration:} 10 min./level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

You bring into being a mobile, hemispherical energy field that prevents the entrance 
of most types of living creatures.

The effect hedges out animals, aberrations, dragons, fey, giants, humanoids, magical 
beasts, monstrous humanoids, oozes, plants, and vermin, but not constructs, elementals, 
outsiders, or undead.

This spell may be used only defensively, not aggressively. Forcing an abjuration 
barrier against creatures that the spell keeps at bay collapses the barrier.

\vspace{12pt}
Antimagic Field

Abjuration

\textbf{Level:} Clr 8, Magic 6, Protection 6, Sor/Wiz 6

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 10 ft.

\textbf{Area:} 10-ft.-radius emanation, centered on you

\textbf{Duration:} 10 min./level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} See text

An invisible barrier surrounds you and moves with you. The space within this barrier 
is impervious to most magical effects, including spells, spell-like abilities, 
and supernatural abilities. Likewise, it prevents the functioning of any magic 
items or spells within its confines.

An \textit{antimagic field }suppresses any spell or magical effect used within, 
brought into, or cast into the area, but does not dispel it. Time spent within 
an \textit{antimagic field }counts against the suppressed spell's duration.

Summoned creatures of any type and incorporeal undead wink out if they enter an 
\textit{antimagic field. }They reappear in the same spot once the field goes away. 
Time spent winked out counts normally against the duration of the conjuration that 
is maintaining the creature. If you cast \textit{antimagic field }in an area occupied 
by a summoned creature that has spell resistance, you must make a caster level 
check (1d20 + caster level) against the creature's spell resistance to make it 
wink out. (The effects of instantaneous conjurations\textit{ }are not affected 
by an \textit{antimagic field }because the conjuration itself is no longer in effect, 
only its result.)

A normal creature can enter the area, as can normal missiles. Furthermore, while 
a magic sword does not function magically within the area, it is still a sword 
(and a masterwork sword at that). The spell has no effect on golems and other constructs 
that are imbued with magic during their creation process and are thereafter self-supporting 
(unless they have been summoned, in which case they are treated like any other 
summoned creatures). Elementals, corporeal undead, and outsiders are likewise unaffected 
unless summoned. These creatures' spell-like or supernatural abilities, however, 
may be temporarily nullified by the field. \textit{Dispel magic }does not remove 
the field.

Two or more \textit{antimagic fields }sharing any of the same space have no effect 
on each other. Certain spells, such as \textit{wall of force, prismatic sphere}, 
and \textit{prismatic wall, }remain unaffected by \textit{antimagic field }(see 
the individual spell descriptions). Artifacts and deities are unaffected by mortal 
magic such as this. 

Should a creature be larger than the area enclosed by the barrier, any part of 
it that lies outside the barrier is unaffected by the field.

\textit{Arcane Material Component: }A pinch of powdered iron or iron filings.

\vspace{12pt}
Antipathy

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Drd 9, Sor/Wiz 8

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 hour

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One location (up to a 10-ft. cube/level) or one object

\textbf{Duration:} 2 hours/level (D)

\textbf{Saving Throw: }Will partial

\textbf{Spell Resistance:} Yes

You cause an object or location to emanate magical vibrations that repel either 
a specific kind of intelligent creature or creatures of a particular alignment, 
as defined by you. The kind of creature to be affected must be named specifically. 
A creature subtype is not specific enough. Likewise, the specific alignment to 
be repelled must be named.

Creatures of the designated kind or alignment feel an overpowering urge to leave 
the area or to avoid the affected item.

A compulsion forces them to abandon the area or item, shunning it and never willingly 
returning to it while the spell is in effect. A creature that makes a successful 
saving throw can stay in the area or touch the item but feels uncomfortable doing 
so. This distracting discomfort reduces the creature's Dexterity score by 4 points.

\textit{Antipathy }counters and dispels \textit{sympathy}.

\textit{Arcane Material Component: }A lump of alum soaked in vinegar.

\vspace{12pt}
Antiplant Shell

Abjuration

\textbf{Level:} Drd 4

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 10 ft.

\textbf{Area:} 10-ft.-radius emanation, centered on you

\textbf{Duration:} 10 min./level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

The \textit{antiplant shell }spell creates an invisible, mobile barrier that keeps 
all creatures within the shell protected from attacks by plant creatures or animated 
plants. As with many abjuration spells, forcing the barrier against creatures that 
the spell keeps at bay strains and collapses the field.

\vspace{12pt}
Arcane Eye

Divination (Scrying)

\textbf{Level:} Sor/Wiz 4

\textbf{Components:} V, S, M

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Unlimited

\textbf{Effect: }Magical sensor

\textbf{Duration:} 1 min./level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You create an invisible magical sensor that sends you visual information. You can 
create the \textit{arcane eye }at any point you can see, but it can then travel 
outside your line of sight without hindrance. An \textit{arcane eye }travels at 
30 feet per round (300 feet per minute) if viewing an area ahead as a human would 
(primarily looking at the floor) or 10 feet per round (100 feet per minute) if 
examining the ceiling and walls as well as the floor ahead. It sees exactly as 
you would see if you were there.

The eye can travel in any direction as long as the spell lasts. Solid barriers 
block its passage, but it can pass through a hole or space as small as 1 inch in 
diameter. The eye can't enter another plane of existence, even through a \textit{gate 
}or similar magical portal.

You must concentrate to use an \textit{arcane eye. }If you do not concentrate, 
the eye is inert until you again concentrate.

\textit{Material Component: }A bit of bat fur.

\vspace{12pt}
Arcane Lock

Abjuration

\textbf{Level:} Sor/Wiz 2

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} The door, chest, or portal touched, up to 30 sq. ft./level in 
size

\textbf{Duration:} Permanent

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

An \textit{arcane lock }spell cast upon a door, chest, or portal magically locks 
it. You can freely pass your own \textit{arcane lock }without affecting it; otherwise, 
a door or object secured with this spell can be opened only by breaking in or with 
a successful \textit{dispel magic }or \textit{knock }spell. Add 10 to the normal 
DC to break open a door or portal affected by this spell. (A \textit{knock }spell 
does not remove an \textit{arcane lock; }it only suppresses the effect for 10 minutes.)

\textit{Material Component: }Gold dust worth 25 gp.

\vspace{12pt}
Arcane Mark

Universal

\textbf{Level:} Sor/Wiz 0

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 0 ft.

\textbf{Effect:} One personal rune or mark, all of which must fit within 1 sq. 
ft.

\textbf{Duration:} Permanent

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell allows you to inscribe your personal rune or mark, which can consist 
of no more than six characters. The writing can be visible or invisible. An \textit{arcane 
mark }spell enables you to etch the rune upon any substance without harm to the 
material upon which it is placed. If an invisible mark is made, a \textit{detect 
magic }spell causes it to glow and be visible, though not necessarily understandable.

\textit{See invisibility, true seeing}, a \textit{gem of seeing}, or a \textit{robe 
of eyes }likewise allows the user to see an invisible \textit{arcane mark. }A \textit{read 
magic }spell reveals the words, if any. The mark cannot be dispelled, but it can 
be removed by the caster or by an \textit{erase }spell.

If an \textit{arcane mark }is placed on a living being, normal wear gradually causes 
the effect to fade in about a month.

\textit{Arcane mark }must be cast on an object prior to casting \textit{instant 
summons }on the same object (see that spell description for details).

\vspace{12pt}
Arcane Sight

Divination

\textbf{Level:} Sor/Wiz 3

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 1 min./level (D)

This spell makes your eyes glow blue and allows you to see magical auras within 
120 feet of you. The effect is similar to that of a \textit{detect magic }spell, 
but \textit{arcane sight }does not require concentration and discerns aura location 
and power more quickly.

You know the location and power of all magical auras within your sight. An aura's 
power depends on a spell's functioning level or an item's caster level, as noted 
in the description of the \textit{detect magic }spell. If the items or creatures 
bearing the auras are in line of sight, you can make Spellcraft skill checks to 
determine the school of magic involved in each. (Make one check per aura; DC 15 
+ spell level, or 15 + one-half caster level for a nonspell effect.)

If you concentrate on a specific creature within 120 feet of you as a standard 
action, you can determine whether it has any spellcasting or spell-like abilities, 
whether these are arcane or divine (spell-like abilities register as arcane), and 
the strength of the most powerful spell or spell-like ability the creature currently 
has available for use.

\textit{Arcane sight }can be made permanent with a \textit{permanency }spell.

\vspace{12pt}
Arcane Sight, Greater

Divination

\textbf{Level:} Sor/Wiz 7

This spell functions like \textit{arcane sight, }except that you automatically 
know which spells or magical effects are active upon any individual or object you 
see.

\textit{Greater arcane sight }doesn't let you identify magic items.

Unlike \textit{arcane sight}, this spell cannot be made permanent with a \textit{permanency 
}spell.

\vspace{12pt}
Astral Projection

Necromancy

\textbf{Level:} Clr 9, Sor/Wiz 9, Travel 9

\textbf{Components:} V, S, M

\textbf{Casting Time:} 30 minutes

\textbf{Range:} Touch

\textbf{Targets:} You plus one additional willing creature touched per two caster 
levels

\textbf{Duration:} See text

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

By freeing your spirit from your physical body, this spell allows you to project 
an astral body onto another plane altogether.

You can bring the astral forms of other willing creatures with you, provided that 
these subjects are linked in a circle with you at the time of the casting. These 
fellow travelers are dependent upon you and must accompany you at all times. If 
something happens to you during the journey, your companions are stranded wherever 
you left them.

You project your astral self onto the Astral Plane, leaving your physical body 
behind on the Material Plane in a state of suspended animation. The spell projects 
an astral copy of you and all you wear or carry onto the Astral Plane. Since the 
Astral Plane touches upon other planes, you can travel astrally to any of these 
other planes as you will. To enter one, you leave the Astral Plane, forming a new 
physical body (and equipment) on the plane of existence you have chosen to enter.

While you are on the Astral Plane, your astral body is connected at all times to 
your physical body by a silvery cord. If the cord is broken, you are killed, astrally 
and physically. Luckily, very few things can destroy a silver cord. When a second 
body is formed on a different plane, the incorporeal silvery cord remains invisibly 
attached to the new body. If the second body or the astral form is slain, the cord 
simply returns to your body where it rests on the Material Plane, thereby reviving 
it from its state of suspended animation. Although astral projections are able 
to function on the Astral Plane, their actions affect only creatures existing on 
the Astral Plane; a physical body must be materialized on other planes.

You and your companions may travel through the Astral Plane indefinitely. Your 
bodies simply wait behind in a state of suspended animation until you choose to 
return your spirits to them. The spell lasts until you desire to end it, or until 
it is terminated by some outside means, such as \textit{dispel magic }cast upon 
either the physical body or the astral form, the breaking of the silver cord, or 
the destruction of your body back on the Material Plane (which kills you).

\textit{Material Component: }A jacinth worth at least 1,000 gp, plus a silver bar 
worth 5 gp for each person to be affected.

\vspace{12pt}
Atonement

Abjuration

\textbf{Level:} Clr 5, Drd 5

\textbf{Components:} V, S, M, F, DF, XP

\textbf{Casting Time:} 1 hour

\textbf{Range:} Touch

\textbf{Target:} Living creature touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

This spell removes the burden of evil acts or misdeeds from the subject. The creature 
seeking atonement must be truly repentant and desirous of setting right its misdeeds. 
If the atoning creature committed the evil act unwittingly or under some form of 
compulsion, \textit{atonement }operates normally at no cost to you. However, in 
the case of a creature atoning for deliberate misdeeds and acts of a knowing and 
willful nature, you must intercede with your deity (requiring you to expend 500 
XP) in order to expunge the subject's burden. Many casters first assign a subject 
of this sort a quest (see \textit{geas/quest}) or similar penance to determine 
whether the creature is truly contrite before casting the \textit{atonement }spell 
on its behalf.

\textit{Atonement }may be cast for one of several purposes, depending on the version 
selected.

\textit{Reverse Magical Alignment Change: }If a creature has had its alignment 
magically changed, \textit{atonement }returns its alignment to its original status 
at no cost in experience points.

\textit{Restore Class: }A paladin who has lost her class features due to committing 
an evil act may have her paladinhood restored to her by this spell.

\textit{Restore Cleric or Druid Spell Powers: }A cleric or druid who has lost the 
ability to cast spells by incurring the anger of his or her deity may regain that 
ability by seeking \textit{atonement }from another cleric of the same deity or 
another druid. If the transgression was intentional, the casting cleric loses 500 
XP for his intercession. If the transgression was unintentional, he does not lose 
XP.

\textit{Redemption or Temptation: }You may cast this spell upon a creature of an 
opposing alignment in order to offer it a chance to change its alignment to match 
yours. The prospective subject must be present for the entire casting process. 
Upon completion of the spell, the subject freely chooses whether it retains its 
original alignment or acquiesces to your offer and changes to your alignment. No 
duress, compulsion, or magical influence can force the subject to take advantage 
of the opportunity offered if it is unwilling to abandon its old alignment. This 
use of the spell does not work on outsiders or any creature incapable of changing 
its alignment naturally.

Though the spell description refers to evil acts, \textit{atonement }can also be 
used on any creature that has performed acts against its alignment, whether those 
acts are evil, good, chaotic, or lawful.

\textit{Note: }Normally, changing alignment is up to the player. This use of \textit{atonement 
}simply offers a believable way for a character to change his or her alignment 
drastically, suddenly, and definitively.

\textit{Material Component: }Burning incense.

\textit{Focus: }In addition to your holy symbol or normal divine focus, you need 
a set of prayer beads (or other prayer device, such as a prayer wheel or prayer 
book) worth at least 500 gp.

\textit{XP Cost: }When cast for the benefit of a creature whose guilt was the result 
of deliberate acts, the cost to you is 500 XP per casting (see above).

\vspace{12pt}
Augury

Divination

\textbf{Level:} Clr 2

\textbf{Components:} V, S, M, F

\textbf{Casting Time:} 1 minute

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} Instantaneous

An \textit{augury }can tell you whether a particular action will bring good or 
bad results for you in the immediate future.

The base chance for receiving a meaningful reply is 70\% + 1\% per caster level, 
to a maximum of 90\%; this roll is made secretly. A question may be so straightforward 
that a successful result is automatic, or so vague as to have no chance of success. 
If the \textit{augury }succeeds, you get one of four results:• 

Weal (if the action will probably bring good results).• 

Woe (for bad results).• 

Weal and woe (for both).• 

Nothing (for actions that don't have especially good or bad results).

If the spell fails, you get the ``nothing'' result. A cleric who gets the ``nothing'' 
result has no way to tell whether it was the consequence of a failed or successful 
\textit{augury.}

The \textit{augury }can see into the future only about half an hour, so anything 
that might happen after that does not affect the result. Thus, the result might 
not take into account the long-term consequences of a contemplated action. All 
\textit{auguries }cast by the same person about the same topic use the same dice 
result as the first casting.

\textit{Material Component: }Incense worth at least 25 gp.

\textit{Focus: }A set of marked sticks, bones, or similar tokens of at least 25 
gp value.

\vspace{12pt}
Awaken

Transmutation

\textbf{Level:} Drd 5

\textbf{Components:} V, S, DF, XP

\textbf{Casting Time:} 24 hours

\textbf{Range:} Touch

\textbf{Target:} Animal or tree touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

You awaken a tree or animal to humanlike sentience. To succeed, you must make a 
Will save (DC 10 + the animal's current HD, or the HD the tree will have once awakened).

The \textit{awakened }animal or tree is friendly toward you. You have no special 
empathy or connection with a creature you awaken, although it serves you in specific 
tasks or endeavors if you communicate your desires to it.

An \textit{awakened }tree has characteristics as if it were an animated object, 
except that it gains the plant type and its Intelligence, Wisdom, and Charisma 
scores are each 3d6. An \textit{awakened }plant gains the ability to move its limbs, 
roots, vines, creepers, and so forth, and it has senses similar to a human's.

An \textit{awakened }animal gets 3d6 Intelligence, +1d3 Charisma, and +2 HD. Its 
type becomes magical beast (augmented animal). An awakened animal can't serve as 
an animal companion, familiar, or special mount.

An \textit{awakened }tree or animal can speak one language that you know, plus 
one additional language that you know per point of Intelligence bonus (if any).

\textit{XP Cost: }250 XP.

\vspace{12pt}
Baleful Polymorph

Transmutation

\textbf{Level:} Drd 5, Sor/Wiz 5

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One creature

\textbf{Duration:} Permanent

\textbf{Saving Throw:} Fortitude negates, Will partial; see text

\textbf{Spell Resistance:} Yes

As \textit{polymorph, }except that you change the subject into a Small or smaller 
animal of no more than 1 HD. If the new form would prove fatal to the creature 
the subject gets a +4 bonus on the save.

If the spell succeeds, the subject must also make a Will save. If this second save 
fails, the creature loses its extraordinary, supernatural, and spell-like abilities, 
loses its ability to cast spells (if it had the ability), and gains the alignment, 
special abilities, and Intelligence, Wisdom, and Charisma scores of its new form 
in place of its own. It still retains its class and level (or HD), as well as all 
benefits deriving therefrom (such as base attack bonus, base save bonuses, and 
hit points). It retains any class features (other than spellcasting) that aren't 
extraordinary, supernatural, or spell-like abilities.

Incorporeal or gaseous creatures are immune to being \textit{polymorphed}, and 
a creature with the shapechanger subtype can revert to its natural form as a standard 
action.

\vspace{12pt}
Bane

Enchantment (Compulsion) [Fear, Mind-Affecting]

\textbf{Level:} Clr 1

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 50 ft.

\textbf{Area:} All enemies within 50 ft.

\textbf{Duration:} 1 min./level

\textbf{Saving Throw:}Will negates

\textbf{Spell Resistance:} Yes

\textit{Bane }fills your enemies with fear and doubt. Each affected creature takes 
a -1 penalty on attack rolls and a -1 penalty on saving throws against fear effects.

\textit{Bane }counters and dispels \textit{bless}.

\vspace{12pt}
Banishment

Abjuration

\textbf{Level:} Clr 6, Sor/Wiz 7

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} One or more extraplanar creatures, no two of which can be more 
than 30 ft. apart

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

A \textit{banishment }spell is a more powerful version of the \textit{dismissal 
}spell. It enables you to force extraplanar creatures out of your home plane. As 
many as 2 Hit Dice of creatures per caster level can be banished.

You can improve the spell's chance of success by presenting at least one object 
or substance that the target hates, fears, or otherwise opposes. For each such 
object or substance, you gain a +1 bonus on your caster level check to overcome 
the target's spell resistance (if any), the saving throw DC increases by 2. 

Certain rare items might work twice as well as a normal item for the purpose of 
the bonuses (each providing a +2 bonus on the caster level check against spell 
resistance and increasing the save DC by 4).

\textit{Arcane Focus: }Any item that is distasteful to the subject (optional, see 
above).

\vspace{12pt}
Barkskin

Transmutation

\textbf{Level:} Drd 2, Rgr 2, Plant 2

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Living creature touched

\textbf{Duration:} 10 min./level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes (harmless)

\textit{Barkskin }toughens a creature's skin. The effect grants a +2 enhancement 
bonus to the creature's existing natural armor bonus. This enhancement bonus increases 
by 1 for every three caster levels above 3rd, to a maximum of +5 at caster level 
12th.

The enhancement bonus provided by \textit{barkskin }stacks with the target's natural 
armor bonus, but not with other enhancement bonuses to natural armor. A creature 
without natural armor has an effective natural armor bonus of +0.

\vspace{12pt}
Bear's Endurance

Transmutation

\textbf{Level:} Clr 2, Drd 2, Rgr 2, Sor/Wiz 2

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 1 min./level

\textbf{Saving Throw:}Will negates (harmless)

\textbf{Spell Resistance:} Yes

The affected creature gains greater vitality and stamina. The spell grants the 
subject a +4 enhancement bonus to Constitution, which adds the usual benefits to 
hit points, Fortitude saves, Constitution checks, and so forth.

Hit points gained by a temporary increase in Constitution score are not temporary 
hit points. They go away when the subject's Constitution drops back to normal. 
They are not lost first as temporary hit points are.

\vspace{12pt}
Bear's Endurance, Mass

Transmutation

\textbf{Level:} Clr 6, Drd 6, Sor/Wiz 6

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} One creature/level, no two of which can be more than 30 ft. apart

\textit{Mass bear 's endurance }works like \textit{bear 's endurance}, except that 
it affects multiple creatures.

\vspace{12pt}
Bestow Curse

Necromancy

\textbf{Level:} Clr 3, Sor/Wiz 4

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} Permanent

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

You place a curse on the subject. Choose one of the following three effects.• 
-

6 decrease to an ability score (minimum 1).• -

4 penalty on attack rolls, saves, ability checks, and skill checks.• 

Each turn, the target has a 50\% chance to act normally; otherwise, it takes no 
action.

You may also invent your own curse, but it should be no more powerful than those 
described above.

The \textit{curse }bestowed by this spell cannot be dispelled, but it can be removed 
with a \textit{break enchantment}, \textit{limited wish}, \textit{miracle}, \textit{remove 
curse}, or \textit{wish }spell.

\textit{Bestow curse }counters \textit{remove curse}.

\vspace{12pt}
Binding

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Sor/Wiz 8

\textbf{Components:} V, S, M

\textbf{Casting Time:} One minute

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One living creature

\textbf{Duration:} See text (D)

\textbf{Saving Throw: }Will negates; see text

\textbf{Spell Resistance:} Yes

A \textit{binding }spell creates a magical restraint to hold a creature. The target 
gets an initial saving throw only if its Hit Dice equal at least one-half your 
caster level.

You may have as many as six assistants help you with the spell. For each assistant 
who casts \textit{suggestion}, your caster level for this casting of \textit{binding 
}increases by 1. For each assistant who casts \textit{dominate animal, dominate 
person, }or \textit{dominate monster, }your caster level for this casting of \textit{binding 
}increases by a number equal to one-third of that assistant's level, provided that 
the spell's target is appropriate for a \textit{binding }spell. Since the assistants' 
spells are cast simply to improve your caster level for the purpose of the \textit{binding 
}spell, saving throws and spell resistance against the assistants' spells are irrelevant. 
Your caster level determines whether the target gets an initial Will saving throw 
and how long the \textit{binding }lasts. All \textit{binding }spells are dismissible.

Regardless of the version of \textit{binding }you cast, you can specify triggering 
conditions that end the spell and release the creature whenever they occur. These 
triggers can be as simple or elaborate as you desire, but the condition must be 
reasonable and have a likelihood of coming to pass. The conditions can be based 
on a creature's name, identity, or alignment but otherwise must be based on observable 
actions or qualities. Intangibles such as level, class, Hit Dice, or hit points 
don't qualify. Once the spell is cast, its triggering conditions cannot be changed. 
Setting a release condition increases the save DC (assuming a saving throw is allowed) 
by 2.

If you are casting any of the first three versions of \textit{binding }(those with 
limited durations), you may cast additional \textit{binding }spells to prolong 
the effect, since the durations overlap. If you do so, the target gets a saving 
throw at the end of the first spell's duration, even if your caster level was high 
enough to disallow an initial saving throw. If the creature succeeds on this save, 
all the \textit{binding }spells it has received are broken.

The \textit{binding }spell has six versions. Choose one of the following versions 
when you cast the spell.

\textit{Chaining: }The subject is confined by restraints that generate an \textit{antipathy 
}spell affecting all creatures who approach the subject, except you. The duration 
is one year per caster level. The subject of this form of \textit{binding }is confined 
to the spot it occupied when it received the spell.

\textit{Slumber: }This version causes the subject to become comatose for as long 
as one year per caster level. The subject does not need to eat or drink while \textit{slumbering}, 
nor does it age. This form of \textit{binding }is more difficult to cast than \textit{chaining}, 
making it slightly easier to resist. Reduce the spell's save DC by 1.

\textit{Bound Slumber: }This combination of \textit{chaining }and \textit{slumber 
}lasts for as long as one month per caster level. Reduce the save DC by 2.

\textit{Hedged Prison: }The subject is transported to or otherwise brought within 
a confined area from which it cannot wander by any means. The effect is permanent. 
Reduce the save DC by 3.

\textit{Metamorphosis: }The subject assumes gaseous form, except for its head or 
face. It is held harmless in a jar or other container, which may be transparent 
if you so choose. The creature remains aware of its surroundings and can speak, 
but it cannot leave the container, attack, or use any of its powers or abilities. 
The \textit{binding }is permanent. The subject does not need to breathe, eat, or 
drink while \textit{metamorphosed}, nor does it age. Reduce the save DC by 4.

\textit{Minimus Containment: }The subject is shrunk to a height of 1 inch or even 
less and held within some gem, jar, or similar object. The \textit{binding }is 
permanent. The subject does not need to breathe, eat, or drink while \textit{contained}, 
nor does it age. Reduce the save DC by 4.

You can't dispel a \textit{binding }spell with \textit{dispel magic }or a similar 
effect, though an \textit{antimagic field }or \textit{Mage's disjunction }affects 
it normally. A bound extraplanar creature cannot be sent back to its home plane 
due to \textit{dismissal, banishment, }or a similar effect.

\textit{Components: }The components for a \textit{binding }spell vary according 
to the version of the spell, but they always include a continuous chanting utterance 
read from the scroll or spellbook page containing the spell, somatic gestures, 
and materials appropriate to the form of \textit{binding }used. These components 
can include such items as miniature chains of special metals, soporific herbs of 
the rarest sort (for \textit{slumber }bindings), a bell jar of the finest crystal, 
and the like.

In addition to the specially made props suited to the specific type of \textit{binding 
}(cost 500 gp), the spell requires opals worth at least 500 gp for each HD of the 
target and a vellum depiction or carved statuette of the subject to be captured.

\vspace{12pt}
Black Tentacles

Conjuration (Creation)

\textbf{Level:} Sor/Wiz 4

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Area:} 20-ft.-radius spread

\textbf{Duration:} 1 round/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell conjures a field of rubbery black tentacles, each 10 feet long. These 
waving members seem to spring forth from the earth, floor, or whatever surface 
is underfoot---including water. They grasp and entwine around creatures that enter 
the area, holding them fast and crushing them with great strength.

Every creature within the area of the spell must make a grapple check, opposed 
by the grapple check of the tentacles. Treat the tentacles attacking a particular 
target as a Large creature with a base attack bonus equal to your caster level 
and a Strength score of 19. Thus, its grapple check modifier is equal to your caster 
level +8. The tentacles are immune to all types of damage.

Once the tentacles grapple an opponent, they may make a grapple check each round 
on your turn to deal 1d6+4 points of bludgeoning damage. The tentacles continue 
to crush the opponent until the spell ends or the opponent escapes.

Any creature that enters the area of the spell is immediately attacked by the tentacles. 
Even creatures who aren't grappling with the tentacles may move through the area 
at only half normal speed.

\textit{Material Component: }A piece of tentacle from a giant octopus or a giant 
squid.

\vspace{12pt}
Blade Barrier

Evocation [Force]

\textbf{Level:} Clr 6, Good 6, War 6

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Effect: }Wall of whirling blades up to 20 ft. long/ level, or a ringed 
wall of whirling blades with a radius of up to 5 ft. per two levels; either form 
20 ft. high

\textbf{Duration:} 1 min./level (D)

\textbf{Saving Throw:} Reflex half or Reflex negates; see text

\textbf{Spell Resistance:} Yes

An immobile, vertical curtain of whirling blades shaped of pure force springs into 
existence. Any creature passing through the wall takes 1d6 points of damage per 
caster level (maximum 15d6), with a Reflex save for half damage.

If you evoke the barrier so that it appears where creatures are, each creature 
takes damage as if passing through the wall. Each such creature can avoid the wall 
(ending up on the side of its choice) and thus take no damage by making a successful 
Reflex save.

A \textit{blade barrier }provides cover (+4 bonus to AC, +2 bonus on Reflex saves) 
against attacks made through it.

\vspace{12pt}
Blasphemy

Evocation [Evil, Sonic]

\textbf{Level:} Clr 7, Evil 7

\textbf{Components:} V

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 40 ft.

\textbf{Area:} Nonevil creatures in a 40-ft.-radius spread centered on you

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None or Will negates; see text

\textbf{Spell Resistance:} Yes

Any nonevil creature within the area of a \textit{blasphemy }spell suffers the 
following ill effects.

\begin{tabular}{|>{\raggedright}p{86pt}|>{\raggedright}p{140pt}|}
\hline
\subsection*{H\textbf{D}} & E\textbf{ffect}\tabularnewline
\hline
Equal to caster level & Dazed\tabularnewline
\hline
Up to caster level -1 & Weakened, dazed\tabularnewline
\hline
Up to caster level -5 & Paralyzed, weakened, dazed\tabularnewline
\hline
Up to caster level -10 & Killed, paralyzed, weakened, dazed\tabularnewline
\hline
\end{tabular}

The effects are cumulative and concurrent.

No saving throw is allowed against these effects.

\textit{Dazed: }The creature can take no actions for 1 round, though it defends 
itself normally.

\textit{Weakened: }The creature's Strength score decreases by 2d6 points for 2d4 
rounds.

\textit{Paralyzed: }The creature is paralyzed and helpless for 1d10 minutes.

\textit{Killed: }Living creatures die. Undead creatures are destroyed.

Furthermore, if you are on your home plane when you cast this spell, nonevil extraplanar 
creatures within the area are instantly banished back to their home planes. Creatures 
so banished cannot return for at least 24 hours. This effect takes place regardless 
of whether the creatures hear the \textit{blasphemy. }The banishment effect allows 
a Will save (at a -4 penalty) to negate.

Creatures whose Hit Dice exceed your caster level are unaffected by \textit{blasphemy.}

\vspace{12pt}
Bless

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Clr 1, Pal 1

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 50 ft.

\textbf{Area:} The caster and all allies within a 50-ft. burst, centered on the 
caster

\textbf{Duration:} 1 min./level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes (harmless)

\textit{Bless }fills your allies with courage. Each ally gains a +1 morale bonus 
on attack rolls and on saving throws against fear effects.

\textit{Bless }counters and dispels \textit{bane}.

\vspace{12pt}
Bless Water

Transmutation [Good]

\textbf{Level:} Clr 1, Pal 1

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 minute

\textbf{Range:} Touch

\textbf{Target:} Flask of water touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates (object)

\textbf{Spell Resistance:} Yes (object)

This transmutation imbues a flask (1 pint) of water with positive energy, turning 
it into holy water.

\textit{Material Component: }5 pounds of powdered silver (worth 25 gp).

\vspace{12pt}
Bless Weapon

Transmutation

\textbf{Level:} Pal 1

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target: }Weapon touched

\textbf{Duration:} 1 min./level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This transmutation makes a weapon strike true against evil foes. The weapon is 
treated as having a +1 enhancement bonus for the purpose of bypassing the damage 
reduction of evil creatures or striking evil incorporeal creatures (though the 
spell doesn't grant an actual enhancement bonus). The weapon also becomes good, 
which means it can bypass the damage reduction of certain creatures. (This effect 
overrides and suppresses any other alignment the weapon might have.) Individual 
arrows or bolts can be transmuted, but affected projectile weapons (such as bows) 
don't confer the benefit to the projectiles they shoot.

In addition, all critical hit rolls against evil foes are automatically successful, 
so every threat is a critical hit. This last effect does not apply to any weapon 
that already has a magical effect related to critical hits, such as a keen weapon 
or a vorpal sword.

\vspace{12pt}
Blight

Necromancy

\textbf{Level:} Drd 4, Sor/Wiz 5

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} Fortitude half; see text

\textbf{Spell Resistance:} Yes

This spell withers a single plant of any size. An affected plant creature takes 
1d6 points of damage per level (maximum 15d6) and may attempt a Fortitude saving 
throw for half damage. A plant that isn't a creature doesn't receive a save and 
immediately withers and dies.

This spell has no effect on the soil or surrounding plant life.

\vspace{12pt}
Blindness/Deafness

Necromancy

\textbf{Level:} Brd 2, Clr 3, Sor/Wiz 2

\textbf{Components:} V

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Target:} One living creature

\textbf{Duration:} Permanent (D)

\textbf{Saving Throw:} Fortitude negates

\textbf{Spell Resistance:} Yes

You call upon the powers of unlife to render the subject blinded or deafened, as 
you choose.

\vspace{12pt}
Blink

Transmutation

\textbf{Level:} Brd 3, Sor/Wiz 3

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 1 round/level (D)

You ``blink'' back and forth between the Material Plane and the Ethereal Plane. 
You look as though you're winking in and out of reality very quickly and at random.

\textit{Blinking }has several effects, as follows.

Physical attacks against you have a 50\% miss chance, and the Blind-Fight feat 
doesn't help opponents, since you're ethereal and not merely invisible. If the 
attack is capable of striking ethereal creatures, the miss chance is only 20\% 
(for concealment).

If the attacker can see invisible creatures, the miss chance is also only 20\%. 
(For an attacker who can both see and strike ethereal creatures, there is no miss 
chance.) Likewise, your own attacks have a 20\% miss chance, since you sometimes 
go ethereal just as you are about to strike.

Any individually targeted spell has a 50\% chance to fail against you while you're 
\textit{blinking }unless your attacker can target invisible, ethereal creatures. 
Your own spells have a 20\% chance to activate just as you go ethereal, in which 
case they typically do not affect the Material Plane.

While \textit{blinking}, you take only half damage from area attacks (but full 
damage from those that extend onto the Ethereal Plane). You strike as an invisible 
creature (with a +2 bonus on attack rolls), denying your target any Dexterity bonus 
to AC.

You take only half damage from falling, since you fall only while you are material.

While \textit{blinking}, you can step through (but not see through) solid objects. 
For each 5 feet of solid material you walk through, there is a 50\% chance that 
you become material. If this occurs, you are shunted off to the nearest open space 
and take 1d6 points of damage per 5 feet so traveled. You can move at only three-quarters 
speed (because movement on the Ethereal Plane is at half speed, and you spend about 
half your time there and half your time material.)

Since you spend about half your time on the Ethereal Plane, you can see and even 
attack ethereal creatures. You interact with ethereal creatures roughly the same 
way you interact with material ones.

An ethereal creature is invisible, incorporeal, and capable of moving in any direction, 
even up or down. As an incorporeal creature, you can move through solid objects, 
including living creatures.

An ethereal creature can see and hear the Material Plane, but everything looks 
gray and insubstantial. Sight and hearing on the Material Plane are limited to 
60 feet.

Force effects and abjurations affect you normally. Their effects extend onto the 
Ethereal Plane from the Material Plane, but not vice versa. An ethereal creature 
can't attack material creatures, and spells you cast while ethereal affect only 
other ethereal things. Certain material creatures or objects have attacks or effects 
that work on the Ethereal Plane. Treat other ethereal creatures and objects as 
material.

\vspace{12pt}
Blur

Illusion (Glamer)

\textbf{Level:} Brd 2,Sor/Wiz 2

\textbf{Components:} V

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 1 min./level (D)

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

The subject's outline appears blurred, shifting and wavering. This distortion grants 
the subject concealment (20\% miss chance).

A \textit{see invisibility }spell does not counteract the \textit{blur }effect, 
but a \textit{true seeing }spell does.

Opponents that cannot see the subject ignore the spell's effect (though fighting 
an unseen opponent carries penalties of its own).

\vspace{12pt}
Break Enchantment

Abjuration

\textbf{Level:} Brd 4, Clr 5, Luck 5, Pal 4, Sor/Wiz 5

\textbf{Components:} V, S

\textbf{Casting Time:} 1 minute

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} Up to one creature per level, all within 30 ft. of each other

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} See text

\textbf{Spell Resistance:} No

This spell frees victims from enchantments, transmutations, and curses. \textit{Break 
enchantment }can reverse even an instantaneous effect. For each such effect, you 
make a caster level check (1d20 + caster level, maximum +15) against a DC of 11 
+ caster level of the effect. Success means that the creature is free of the spell, 
curse, or effect. For a cursed magic item, the DC is 25.

If the spell is one that cannot be dispelled by \textit{dispel magic}, \textit{break 
enchantment }works only if that spell is 5th level or lower. 

If the effect comes from some permanent magic item \textit{break enchantment }does 
not remove the curse from the item, but it does frees the victim from the item's 
effects. 

\vspace{12pt}
Bull's Strength

Transmutation

\textbf{Level:} Clr 2, Drd 2, Pal 2, Sor/Wiz 2, Strength 2

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 1 min./level

\textbf{Saving Throw:}Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

The subject becomes stronger. The spell grants a +4 enhancement bonus to Strength, 
adding the usual benefits to melee attack rolls, melee damage rolls, and other 
uses of the Strength modifier.

\textit{Arcane Material Component: }A few hairs, or a pinch of dung, from a bull.

\vspace{12pt}
Bull's Strength, Mass

Transmutation

\textbf{Level:} Clr 6, Drd 6, Sor/Wiz 6

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} One creature/level, no two of which can be more than 30 ft. apart

This spell functions like \textit{bull's strength}, except that it affects multiple 
creatures.

\vspace{12pt}
Burning Hands

Evocation [Fire]

\textbf{Level:} Fire 1, Sor/Wiz 1

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 15 ft.

\textbf{Area:} Cone-shaped burst

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} Reflex half

\textbf{Spell Resistance:} Yes

A cone of searing flame shoots from your fingertips. Any creature in the area of 
the flames takes 1d4 points of fire damage per caster level (maximum 5d4). Flammable 
materials burn if the flames touch them. A character can extinguish burning items 
as a full-round action.

\newpage

\end{document}
