%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{array}
\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

This material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

{\LARGE{}MONSTERS (S)}

\vspace{12pt}
{\LARGE{}SAHUAGIN}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{212pt}|}
\hline
  & Medium Monstrous Humanoid (Aquatic)\tabularnewline
\hline
H\textbf{it Dice:} & 2d8+2 (11 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares), swim 60 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 16 (+1 Dex, +5 natural), touch 11, flat-footed 15\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +2/+4\tabularnewline
\hline
A\textbf{ttack:} & Talon +4 melee (1d4+2) or trident +4 melee (1d8+3) or heavy 
crossbow +3 ranged (1d10/19-20)\tabularnewline
\hline
F\textbf{ull Attack:} & Trident +4 melee (1d8+3) and bite +2 melee (1d4+1); or 
2 talons +4 melee (1d4+2) and bite +2 melee (1d4+1); or heavy crossbow +3 ranged 
(1d10/19-20)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Blood frenzy, rake 1d4+1\tabularnewline
\hline
S\textbf{pecial Qualities:} & Blindsense 30 ft., darkvision 60 ft., freshwater 
sensitivity, light blindness, speak with sharks, water dependent\tabularnewline
\hline
S\textbf{aves:} & Fort +3, Ref +4, Will +4\tabularnewline
\hline
A\textbf{bilities:} & Str 14, Dex 13, Con 12, Int 14, Wis 13, Cha 9\tabularnewline
\hline
S\textbf{kills:} & Handle Animal +4*, Hide +6*, Listen +6*, Profession (hunter) 
+1*, Ride +3, Spot +6*, Survival +1*\tabularnewline
\hline
F\textbf{eats:} & Great Fortitude, Multiattack\textsuperscript{B}\tabularnewline
\hline
E\textbf{nvironment:} & Warm aquatic\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, team (5-8), patrol (11-20 plus 1 3rd-level 
lieutenant and 1-2 sharks), band (20-80 plus 100\% noncombatants plus 1 3rd-level 
lieutenant and 1 4th-level chieftain per 20 adults plus 1-2 sharks), or tribe (70-160 
plus 100\% noncombatants plus 1 3rd-level lieutenant per 20 adults, 1 4th-level 
chieftain per 40 adults, 9 4th-level guards, 1-4 underpriestesses of 3rd-6th level, 
1 7th-level priestess, and 1 baron of 6th-8th level plus 5-8 sharks)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Usually lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & 3-5 HD (Medium), 6-10 HD (Large), or by character class\tabularnewline
\hline
L\textbf{evel Adjustment:} & +2 (+3 if four-armed)\tabularnewline
\hline
\end{tabular}

Most sahuagin feature green coloration, darker along the back and lighter on the 
belly. Many have dark stripes, bands, or spots, but these tend to fade with age. 
An adult male sahuagin stands roughly 6 feet tall and weighs about 200 pounds. 

Sahuagin are the natural enemy of aquatic elves. The two cannot coexist peacefully: 
Wars between them are prolonged, bloody affairs that sometimes interfere with shipping 
and maritime trade. Sahuagin have an only slightly less vehement hatred for tritons.

Sahuagin speak their own language, Sahuagin. Thanks to their high Intelligence 
scores, most sahuagin also speak two bonus languages, usually Common and Aquan.

COMBAT

Sahuagin are savage fighters, asking for and giving no quarter. When swimming, 
a sahuagin tears with its feet as it strikes with its talons or a weapon. About 
half of any group of sahuagin are also armed with nets.

\textbf{Blindsense (Ex): }A sahuagin can locate creatures underwater within a 30-foot 
radius. This ability works only when the sahuagin is underwater.

\textbf{Blood Frenzy: }Once per day a sahuagin that takes damage in combat can 
fly into a frenzy in the following round, clawing and biting madly until either 
it or its opponent is dead. It gains +2 Constitution and +2 Strength, and takes 
a -2 penalty to Armor Class. A sahuagin cannot end its frenzy voluntarily. 

\textbf{Rake (Ex): }Attack bonus +2 melee, damage 1d4+1. A sahuagin also gains 
two rake attacks when it attacks while swimming.

\textbf{Freshwater Sensitivity (Ex): }A sahuagin fully immersed in fresh water 
must succeed on a DC 15 Fortitude save or become fatigued. Even on a success, it 
must repeat the save attempt every 10 minutes it remains immersed.

\textbf{Light Blindness (Ex): }Abrupt exposure to bright light (such as sunlight 
or a \textit{daylight }spell) blinds sahuagin for round. On subsequent rounds, 
they are dazzled while operating in bright light.

\textbf{Speak with Sharks (Ex): }Sahuagin can communicate telepathically with sharks 
up to 150 feet away. The communication is limited to fairly simple concepts such 
as ``food,'' ``danger,'' and ``enemy.'' Sahuagin can use the Handle Animal skill 
to befriend and train sharks.

\textbf{Water Dependent (Ex):} Sahuagin can survive out of the water for 1 hour 
per 2 points of Constitution (after that, refer to the drowning rules on page 304 
of the \textit{Dungeon Master 's Guide}).

\textbf{Skills:} A sahuagin has a +8 racial bonus on any Swim check to perform 
some special action or avoid a hazard. It can always choose to take 10 on a Swim 
check, even if distracted or endangered. It can use the run action while swimming, 
provided it swims in a straight line. 

*Underwater, a sahuagin has a +4 racial bonus on Hide, Listen, and Spot checks.

*A sahuagin has a +4 racial bonus on Survival and Profession (hunter) checks within 
50 miles of its home.

*A sahuagin has a +4 racial bonus on Handle Animal checks when working with sharks.

SAHUAGIN MUTANTS

About one in two hundred sahuagin has four arms. Such creatures can make four claw 
attacks or use extra weapons, in addition to the claw and bite attacks.

If a community of aquatic elves is located within 100 miles of a sahuagin community, 
about one in one hundred sahuagin looks just like an aquatic elf. These creatures, 
called malenti, have a swim speed of 40 feet, can remain out of water for 1 hour 
per point of Constitution, and have freshwater sensitivity and light sensitivity 
(dazzled in bright light). Malenti have no natural attacks. They are otherwise 
identical with sahuagin.

\vspace{12pt}
{\LARGE{}SALAMANDER}

\begin{tabular}{|>{\raggedright}p{48pt}|>{\raggedright}p{84pt}|>{\raggedright}p{84pt}|>{\raggedright}p{84pt}|}
\hline
 & F\textbf{lamebrother Salamander} & A\textbf{verage Salamander} & N\textbf{oble 
Salamander}\tabularnewline
\hline
  & Small Outsider (Extraplanar, Fire) & Medium Outsider (Extraplanar, Fire) & Large 
Outsider (Extraplanar, Fire)\tabularnewline
\hline
H\textbf{it Dice:} & 4d8+8 (26 hp) & 9d8+18 (58 hp) & 15d8+45 (112 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1 & +1 & +1\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares) & 20 ft. (4 squares) & 20 ft. (4 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 19 (+1 size, +1 Dex, +7 natural), touch 12, flat-footed 
18 & 18 (+1 Dex, +7 natural), touch 11, flat-footed 17 & 18 (-1 size, +1 Dex, +8 
natural), touch 10, flat-footed 17\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +4/+1 & +9/+11 & +15/+25\tabularnewline
\hline
A\textbf{ttack:} & Spear +6 melee (1d6+1/x3 plus 1d6 fire) & Spear +11 melee (1d8+3/x3 
plus 1d6 fire) & +\textit{3 longspear }+23 melee (1d8+9/x3 plus 1d8 fire)\tabularnewline
\hline
F\textbf{ull Attack:} & Spear +6 melee (1d6+1/x3 plus 1d6 fire) and tail slap +4 
melee (1d4 plus 1d6 fire) & Spear +11/+6 melee (1d8+3/x3 plus 1d6 fire) and tail 
slap +9 melee (2d6+1 plus 1d6 fire) & +\textit{3 longspear }+23/+18/+13 melee (1d8+9/x3 
plus 1d8 fire) and tail slap +18 melee (2d8+3 plus 1d8 fire)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. & 5 ft./5 ft. (10 ft. with tail) & 10 ft./10 
ft. (20 ft. with tail or longspear)\tabularnewline
\hline
S\textbf{pecial Attacks:} & Constrict 1d4 plus 1d6 fire, heat, improved grab & Constrict 
2d6+1 plus 1d6 fire, heat, improved grab & Constrict 2d8+3 plus 1d8 fire, heat, 
improved grab, spell-like abilities\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., immunity to fire, vulnerability 
to cold & Damage reduction 10/magic, darkvision 60 ft., immunity to fire, vulnerability 
to cold & Damage reduction 15/magic, darkvision 60 ft., immunity to fire, vulnerability 
to cold\tabularnewline
\hline
S\textbf{aves:} & Fort +6, Ref +5, Will +6 & Fort +8, Ref +7, Will +8 & Fort +12, 
Ref +10, Will +11\tabularnewline
\hline
A\textbf{bilities:} & Str 12, Dex 13, Con 14, Int 14, Wis 15, Cha 13 & Str 14, 
Dex 13, Con 14, Int 14, Wis 15, Cha 13 & Str 22, Dex 13, Con 16, Int 16, Wis 15, 
Cha 15\tabularnewline
\hline
S\textbf{kills:} & Craft (blacksmithing) +8, Hide +12, Listen +11, Move Silently 
+6, Spot +11 & Bluff +11, Craft (blacksmithing) +19, Diplomacy +3, Disguise +1 
(+3 acting), Hide +11, Intimidate +3, Listen +8, Move Silently +11,Search +12, 
Spot +8 & Bluff +19, Craft (blacksmithing) +25, Diplomacy +4, Hide +15, Intimidate 
+4, Listen +13, Move Silently +17, Spot +13\tabularnewline
\hline
F\textbf{eats:} & Alertness, Multiattack & Alertness, Multiattack, Power Attack & Alertness, 
Cleave, Great Cleave, Multiattack, Power Attack, Skill Focus (Craft [blacksmithing])\tabularnewline
\hline
E\textbf{nvironment:} & Elemental Plane of Fire & Elemental Plane of Fire & Elemental 
Plane of Fire\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or cluster (3-5) & Solitary, pair, or 
cluster (3-5) & Solitary, pair, or noble party (9-14)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 3 & 6 & 10\tabularnewline
\hline
T\textbf{reasure:} & Standard (nonflammables only) & Standard (nonflammables only) & Double 
standard (nonflammables only) and \textit{+3 longspear}\tabularnewline
\hline
A\textbf{lignment:} & Usually evil (any) & Usually evil (any) & Usually evil (any)\tabularnewline
\hline
A\textbf{dvancement:} & 4-6 HD (Small) & 8-14 HD (Medium) & 16-21 HD (Large); 22-45 
HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +4 & +5--- & \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
Salamanders speak Ignan. Some average salamanders and all nobles also speak Common.

COMBAT

If a salamander has damage reduction, its natural weapons are treated as magic 
weapons for the purpose of overcoming damage reduction.

\textbf{Constrict (Ex): }A salamander deals automatic tail slap damage (including 
fire damage) with a successful grapple check. A noble salamander can constrict 
multiple creatures simultaneously, provided they are all at least two sizes smaller 
than it.

\textbf{Heat (Ex):} A salamander generates so much heat that its mere touch deals 
additional fire damage. Salamanders' metallic weapons also conduct this heat.

\textbf{Improved Grab (Ex):} To use this ability, a salamander must hit a creature 
of up to one size larger than itself with its tail slap attack. It can then attempt 
to start a grapple as a free action without provoking an attack of opportunity. 
If it wins the grapple check, it establishes a hold and can constrict.

\textbf{Spell-Like Abilities:} (Noble salamanders only) 3/day---\textit{burning 
hands }(DC 13), \textit{fireball }(DC 15), \textit{flaming sphere }(DC 14), \textit{wall 
of fire }(DC 16); 1/day---\textit{dispel magic, summon monster VII }(Huge fire 
elemental). Caster level 15th. The save DCs are Charisma-based.

\textbf{Skills:} Salamanders have a +4 racial bonus on Craft (blacksmithing) checks.

\textbf{Feats:} Salamanders have the Multiattack feat even though they do not have 
the requisite three natural weapons.

\vspace{12pt}
{\LARGE{}SATYR}

\begin{tabular}{|>{\raggedright}p{89pt}|>{\raggedright}p{236pt}|}
\hline
  & Medium Fey\tabularnewline
\hline
H\textbf{it Dice:} & 5d6+5 (22 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 15 (+1 Dex, +4 natural), touch 11, flat-footed 14\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +2/+2\tabularnewline
\hline
A\textbf{ttack:} & Head butt +2 melee (1d6) or shortbow +3 ranged (1d6/x3)\tabularnewline
\hline
F\textbf{ull Attack:} & Head butt +2 melee (1d6) and dagger -3 melee (1d4/19-20); 
or shortbow +3 ranged (1d6/x3)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Pipes\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 5/cold iron, low-light vision\tabularnewline
\hline
S\textbf{aves:} & Fort +2, Ref +5, Will +5\tabularnewline
\hline
A\textbf{bilities:} & Str 10, Dex 13, Con 12, Int 12, Wis 13, Cha 13\tabularnewline
\hline
S\textbf{kills:} & Bluff +9, Diplomacy +3, Disguise +1 (+3 acting), Hide +13, Intimidate 
+3, Knowledge (nature) +9, Listen +15, Move Silently +13, Perform (wind instruments) 
+9, Spot +15, Survival +1 (+3 aboveground)\tabularnewline
\hline
F\textbf{eats:} & Alertness\textsuperscript{B}, Dodge, Mobility\tabularnewline
\hline
E\textbf{nvironment:} & Temperate forests\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, band (3-5), or troop (6-11)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2 (without pipes) or 4 (with pipes)\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic neutral\tabularnewline
\hline
A\textbf{dvancement:} & 6-10 HD (Medium)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +2\tabularnewline
\hline
\end{tabular}

A satyr's hair is red or chestnut brown, while its hooves and horns are jet black. 
A satyr is about as tall and heavy as a half-elf.

Satyrs speak Sylvan, and most also speak Common.

COMBAT

The keen senses of a satyr make it almost impossible to surprise one in the wild. 
Conversely, with their own natural grace and agility, satyrs can sneak up on travelers 
who are not carefully watching the surrounding wilderness. Once engaged in battle, 
an unarmed satyr attacks with a powerful head butt. A satyr expecting trouble is 
likely to be armed with a bow and a dagger and typically looses arrows from hiding, 
weakening an enemy before closing. 

\textbf{Pipes (Su):} Satyrs can play a variety of magical tunes on their pan pipes. 
Usually, only one satyr in a group carries pipes. When it plays, all creatures 
within a 60-foot spread (except satyrs) must succeed on a DC 13 Will save or be 
affected by \textit{charm person, sleep, }or \textit{fear }(caster level 10th; 
the satyr chooses the tune and its effect).

In the hands of other beings, these pipes have no special powers. A creature that 
successfully saves against any of the pipe's effects cannot be affected by the 
same set of pipes for 24 hours. The save DC is Charisma-based.

\textbf{Skills:} Satyrs have a +4 racial bonus on Hide, Listen, Move Silently, 
Perform, and Spot checks.

SATYRS AS CHARACTERS

Satyr characters possess the following racial traits.--- 

\parindent=3pt
+2 Dexterity, +2 Constitution, +2 Intelligence, +2 Wisdom, +2 Charisma.---

\parindent=0pt
Medium size.---

A satyr's base land speed is 40 feet.---

Low-light vision.---

Racial Hit Dice: A satyr begins with five levels of fey, which provide 5d8 Hit 
Dice, a base attack bonus of +2, and base saving throw bonuses of Fort +1, Ref 
+4, and Will +4.---

Racial Skills: A satyr's fey levels give it skill points equal to 8 x$ $(6 + Int 
modifier). Its class skills are Bluff, Hide, Knowledge (nature), Listen, Move Silently, 
Perform, and Spot. Satyrs have a +4 racial bonus on Hide, Listen, Move Silently, 
Perform, and Spot checks.---

Racial Feats: A satyr's fey levels give it two feats. A satyr receives Alertness 
as a bonus feat.--- 

\parindent=3pt
+4 natural armor bonus.---

\parindent=0pt
Natural Weapons: Head butt (1d6).---

Special Attacks (see above): Pipes.---

Special Qualities (see above): Damage reduction 5/cold iron. ---

Automatic Languages: Sylvan. Bonus Languages: Common, Elven, Gnome.---

Favored Class: Bard.---

Level adjustment +2.

\vspace{12pt}
{\LARGE{}SEA CAT}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{226pt}|}
\hline
  & Large Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 6d10+18 (51 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 10 ft. (2 squares), swim 40 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 18 (-1 size, +1 Dex, +8 natural), touch 10, flat-footed 
17\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +6/+14\tabularnewline
\hline
A\textbf{ttack:} & Claw +9 melee (1d6+4)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +9 melee (1d6+4) and bite +4 melee (1d8+2)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Rend 2d6+6\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., hold breath, low-light vision, 
scent\tabularnewline
\hline
S\textbf{aves:} & Fort +8, Ref +6, Will +5\tabularnewline
\hline
A\textbf{bilities:} & Str 19, Dex 12, Con 17, Int 2, Wis 13, Cha 10\tabularnewline
\hline
S\textbf{kills:} & Listen +8, Spot +7, Swim +12\tabularnewline
\hline
F\textbf{eats:} & Alertness, Endurance, Iron Will\tabularnewline
\hline
E\textbf{nvironment:} & Temperate aquatic\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or pride (5-12)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 4\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & 7-9 HD (Large); 10-18 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A typical sea cat is 12 feet long and weighs 800 pounds.

COMBAT

Sea cats attack on sight, either for food or to defend their territory, and use 
both claws and teeth to grab and rend their prey. They display tremendous courage, 
always fighting to the death, even against creatures many times their size. Pairs 
and prides of sea cats attack in concert, trying to wear the opponent down until 
one beast can dispatch it.

\textbf{Hold Breath (Ex): }A sea cat can hold its breath for a number of rounds 
equal to 6 $ $its Constitution score before it risks drowning. 

\textbf{Rend (Ex): }A sea cat that hits with both claw attacks latches onto the 
opponent's body and tears the flesh. This automatically deals an extra 2d6+6 points 
of damage.

\textbf{Skills:} A sea cat has a +8 racial bonus on any Swim check to perform some 
special action or avoid a hazard. It can always choose to take 10 on a Swim check, 
even if distracted or endangered. It can use the run action while swimming, provided 
it swims in a straight line.

\vspace{12pt}
{\LARGE{}SHADOW}

\begin{tabular}{|>{\raggedright}p{47pt}|>{\raggedright}p{135pt}|>{\raggedright}p{132pt}|}
\hline
 & S\textbf{hadow} & \subsection*{G\textbf{reater Shadow}}\tabularnewline
\hline
 & Medium Undead (Incorporeal) & Medium Undead (Incorporeal)\tabularnewline
\hline
H\textbf{it Dice:} & 3d12 (19 hp) & 9d12 (58 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +2 & +2\tabularnewline
\hline
S\textbf{peed:} & Fly 40 ft. (good) (8 squares) & Fly 40 ft. (good) (8 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 13 (+2 Dex, +1 deflection), touch 13, flat-footed 11 & 14 
(+2 Dex, +2 deflection), touch 14, flat-footed 12 \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +1/--- & +4/---\tabularnewline
\hline
A\textbf{ttack:} & Incorporeal touch +3 melee (1d6 Str) & Incorporeal touch +6 
melee (1d8 Str)\tabularnewline
\hline
F\textbf{ull Attack:} & Incorporeal touch +3 melee (1d6 Str) & Incorporeal touch 
+6 melee (1d8 Str)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.  & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Create spawn, strength damage  & Create spawn, strength 
damage\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., incorporeal traits, +2 turn resistance, 
undead traits  & Darkvision 60 ft., incorporeal traits, +2 turn resistance, undead 
traits\tabularnewline
\hline
S\textbf{aves:} & Fort +1, Ref +3, Will +4  & Fort +3, Ref +5, Will +7\tabularnewline
\hline
A\textbf{bilities:} & Str ---, Dex 14, Con ---,Int 6, Wis 12, Cha 13  & Str ---, 
Dex 15, Con ---, Int 6, Wis 12, Cha 14\tabularnewline
\hline
S\textbf{kills:} & Hide +8*, Listen +7, Search +4, Spot +7  & Hide +14*, Listen 
+9, Search +6, Spot +9\tabularnewline
\hline
F\textbf{eats:} & Alertness, Dodge  & Alertness, Dodge, Mobility, Spring Attack\tabularnewline
\hline
E\textbf{nvironment:} & Any  & Any\tabularnewline
\hline
O\textbf{rganization:} & Solitary, gang (2-5), or swarm (6-11)  & Solitary\tabularnewline
\hline
C{\small{}\textbf{hallenge Rating:}} & 3 & 8\tabularnewline
\hline
T\textbf{reasure:} & None  & None\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil  & Always chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 4-9 HD (Medium)--- & \tabularnewline
\hline
L\textbf{evel Adjustment:}--- & --- & \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
A shadow can be difficult to see in dark or gloomy areas but stands out starkly 
in brightly illuminated places.

A shadow is 5 to 6 feet tall and is weightless. Shadows cannot speak intelligibly.

COMBAT

Shadows lurk in dark places, waiting for living prey to happen by.

\textbf{Strength Damage (Su):} The touch of a shadow deals 1d6 points of Strength 
damage to a living foe. A creature reduced to Strength 0 by a shadow dies. This 
is a negative energy effect.

\textbf{Create Spawn (Su):} Any humanoid reduced to Strength 0 by a shadow becomes 
a shadow under the control of its killer within 1d4 rounds.

\textbf{Skills:} Shadows have a +2 racial bonus on Listen and Spot checks and a 
+4 racial bonus on Search checks. *A shadow gains a +4 racial bonus on Hide checks 
in areas of shadowy illumination. In brightly lit areas, it takes a -4 penalty 
on Hide checks.

\vspace{12pt}
GREATER SHADOW

Although no more intelligent than an average shadow, a greater shadow is more fearsome 
because of its increased damage and its hit-and-run tactics.

COMBAT

\textbf{Strength Damage (Su):} The touch of a greater shadow deals 1d8 points of 
Strength damage to a living foe.

\vspace{12pt}
{\LARGE{}SHADOW MASTIFF}

\begin{tabular}{|>{\raggedright}p{89pt}|>{\raggedright}p{236pt}|}
\hline
  & Medium Outsider (Extraplanar)\tabularnewline
\hline
H\textbf{it Dice:} & 4d8+12 (30 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +5\tabularnewline
\hline
S\textbf{peed:} & 50 ft. (10 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 14 (+1 Dex, +3 natural), touch 11, flat-footed 13\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +4/+7\tabularnewline
\hline
A\textbf{ttack:} & Bite +7 melee (1d6+4)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +7 melee (1d6+4)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Bay, trip\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., shadow blend, scent\tabularnewline
\hline
S\textbf{aves:} & Fort +7, Ref +5, Will +5\tabularnewline
\hline
A\textbf{bilities:} & Str 17, Dex 13, Con 17, Int 4, Wis 12, Cha 13\tabularnewline
\hline
S\textbf{kills:} & Hide +8, Listen +8, Move Silently +8, Spot +8, Survival +8*\tabularnewline
\hline
F\textbf{eats:} & Dodge, Improved Initiative, Track\textsuperscript{B}\tabularnewline
\hline
E\textbf{nvironment:} & Plane of Shadow\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or pack (5-12)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 5\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral evil\tabularnewline
\hline
A\textbf{dvancement:} & 5-6 HD (Medium); 7-12 HD (Large)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +3 (cohort)\tabularnewline
\hline
\end{tabular}

This creature has the body of a large dog, with a smooth black coat and a mouth 
full of sharp teeth. 

A shadow mastiff is slightly more than 2 feet high at the shoulder and weighs about 
200 pounds. 

Shadow mastiffs cannot speak, but they understand Common.

COMBAT

Shadow mastiffs prefer fighting in shadows or dark conditions, which gives them 
a great advantage.

If a magical light source negates the shadows around them, shadow mastiffs are 
cunning enough to either move out of the light or back off and break up the opposition 
with their baying. They have been known to seize and carry off items enspelled 
with \textit{daylight }spells.

\textbf{Bay (Su):} When a shadow mastiff howls or barks, all creatures except evil 
outsiders within a 300-foot spread must succeed on a DC 13 Will save or become 
panicked for 2d4 rounds. This is a sonic mind-affecting fear effect. Whether or 
not the save is successful, an affected creature is immune to the same mastiff 
's bay for 24 hours. The save DC is Charisma-based.

\textbf{Trip (Ex):} A shadow mastiff that hits with its bite attack can attempt 
to trip the opponent as a free action (+3 check modifier) without making a touch 
attack or provoking an attack of opportunity. If the attempt fails, the opponent 
cannot react to trip the shadow mastiff. 

\textbf{Shadow Blend (Su):} In any condition of illumination other than full daylight, 
a shadow mastiff can disappear into the shadows, giving it total concealment. Artificial 
illumination, even a \textit{light }or \textit{continual flame }spell, does not 
negate this ability. A \textit{daylight }spell, however, will.

\textbf{Skills: }*A shadow mastiff has a +4 racial bonus on Survival checks when 
tracking by scent.

\vspace{12pt}
{\LARGE{}SHAMBLING MOUND}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{213pt}|}
\hline
  & Large Plant\tabularnewline
\hline
H\textbf{it Dice:} & 8d8+24 (60 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +0\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), swim 20 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 20 (-1 size, +11 natural), touch 9, flat-footed 20\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +6/+15\tabularnewline
\hline
A\textbf{ttack:} & Slam +11 melee (2d6+5)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 slams +11 melee (2d6+5)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Improved grab, constrict 2d6+7\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., immunity to electricity, low-light 
vision, plant traits, resistance to fire 10\tabularnewline
\hline
S\textbf{aves:} & Fort +9, Ref +2, Will +4\tabularnewline
\hline
A\textbf{bilities:} & Str 21, Dex 10, Con 17, Int 7, Wis 10, Cha 9\tabularnewline
\hline
S\textbf{kills:} & Hide +3*, Listen +8, Move Silently +8\tabularnewline
\hline
F\textbf{eats:} & Iron Will, Power Attack, Weapon Focus (slam)\tabularnewline
\hline
E\textbf{nvironment:} & Temperate marshes\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 6\tabularnewline
\hline
T\textbf{reasure:} & 1/10th coins; 50\% goods; 50\% items\tabularnewline
\hline
A\textbf{lignment:} & Usually neutral\tabularnewline
\hline
A\textbf{dvancement:} & 9-12 HD (Large); 13-24 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +6\tabularnewline
\hline
\end{tabular}

Shambling mounds, also called shamblers, appear to be heaps of rotting vegetation. 
They are actually intelligent, carnivorous plants.

A shambler's brain and sensory organs are located in its upper body.

A shambler's body has an 8-foot girth and is about 6 feet tall when the creature 
stands erect. It weighs about 3,800 pounds.

COMBAT

A shambling mound batters or constricts its opponents with two huge, armlike appendages.

\textbf{Improved Grab (Ex): }To use this ability, a shambler must hit with both 
slam attacks. It can then attempt to start a grapple as a free action without provoking 
an attack of opportunity. If it wins the grapple check, it establishes a hold and 
can constrict. 

\textbf{Constrict (Ex):} A shambler deals 2d6+7 points of damage with a successful 
grapple check.

\textbf{Immunity to Electricity (Ex): }Shamblers take no damage from electricity. 
Instead, any electricity attack used against a shambler temporarily grants it 1d4 
points of Constitution. The shambler loses these points at the rate of 1 per hour.

Skills: Shamblers have a +4 racial bonus on Hide, Listen, and Move Silently checks. 
*They have a +12 racial bonus on Hide checks when in a swampy or forested area.

\vspace{12pt}
{\LARGE{}SHIELD GUARDIAN}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{233pt}|}
\hline
  & Large Construct\tabularnewline
\hline
H\textbf{it Dice:} & 15d10+30 (112 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +0\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 24 (-1 size, +15 natural), touch 9, flat-footed 24\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +11/+21\tabularnewline
\hline
A\textbf{ttack:} & Slam +16 melee (1d8+6)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 slams +16 melee (1d8+6)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:}--- & \tabularnewline
\hline
S\textbf{pecial Qualities:} & Construct traits, darkvision 60 ft., fast healing 
5, find master, guard, low-light vision, \textit{shield other, spell storing}\tabularnewline
\hline
S\textbf{aves:} & Fort +5, Ref +5, Will +5\tabularnewline
\hline
A\textbf{bilities:} & Str 22, Dex 10, Con ---, Int ---, Wis 10, Cha 1\tabularnewline
\hline
E\textbf{nvironment:} & Any\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 8\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & 16-24 HD (Large); 25-45 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

Created by spellcasters to be bodyguards, shield guardians are constructs that 
protect their masters with spells and stamina. When it is fashioned, a shield guardian 
is keyed to a particular magical amulet. Henceforth, it regards the wearer of that 
amulet to be its master, protecting and following that individual everywhere (unless 
specifically commanded not to do so).

A shield guardian obeys its master's verbal commands to the best of its ability, 
although it is not good for much beyond combat and possibly simple manual labor. 
It can also be keyed to perform specific tasks at specific times or when certain 
conditions are met. The wearer of the amulet can call the shield guardian from 
any distance, and it will come as long as it is on the same plane.

A shield guardian is some 9 feet tall and weighs more than 1,200 pounds.

Shield guardians cannot speak, but they understand commands given in any language.

COMBAT

Shield guardians are straightforward in battle, bashing with their heavy stone 
fists. They are made for defense and are not particularly impressive on offense.

\textbf{Find Master (Su):} As long as a shield guardian and its amulet are on the 
same plane, the shield guardian can find the amulet wearer (or just the amulet, 
if it is removed after the guardian is called).

\textbf{Guard (Ex):} If ordered to do so, a shield guardian moves swiftly to defend 
the wearer of its amulet, blocking blows and disrupting foes. All attacks against 
the amulet wearer take a -2 penalty when the shield guardian is adjacent to its 
master.

\textit{\textbf{Shield Other }}\textbf{(Sp):} The wearer of a shield guardian's 
amulet can activate this defensive ability if within 100 feet of the shield guardian. 
Just as the spell of the same name, this ability transfers to the shield guardian 
half the damage that would be dealt to the amulet wearer (this ability does not 
provide the spell's AC or save bonuses).

\textit{\textbf{Spell Storing }}\textbf{(Sp):} A shield guardian can store one 
spell of 4th level or lower that is cast into it by another creature. It ``casts'' 
this spell when commanded to do so or when a predefined situation arises. Once 
this spell is used, the shield guardian can store another spell (or the same spell 
again).

CONSTRUCTION

A shield guardian is built from wood, bronze, stone, and steel. The materials cost 
5,000 gp.

The creature's master may assemble the body or hire someone else to do the job. 
Creating the body requires a DC 16 Craft (blacksmithing) or Craft (carpentry) check. 
The keyed amulet is fashioned at the same time, and its cost (20,000 gp) is included 
in the cost of the guardian. After the body is sculpted, the shield guardian is 
animated through an extended magical ritual that requires a specially prepared 
laboratory or workroom, similar to an alchemist's laboratory and costing 500 gp 
to establish. If the creator is personally constructing the creature's body, the 
building and the ritual can be performed together. A shield guardian with more 
than 15 Hit Dice can be created, but each additional Hit Die adds +5,000 gp to 
the market price, and the price increases by +20,000 gp if the creature's size 
increases to Huge, modifying the cost to create accordingly. 

CL 15th; Craft Construct, \textit{limited wish, discern location, shield, shield 
other, }caster must be at least 15th level; Price 120,000 gp; Cost 65,000 gp + 
4,600 XP.

AMULET

If a shield guardian's amulet is destroyed, the guardian ceases to function until 
a new one is created. If the wearer dies but the amulet is intact, the shield guardian 
carries out the last command it was given.

\vspace{12pt}
{\LARGE{}SHOCKER LIZARD}

\begin{tabular}{|>{\raggedright}p{77pt}|>{\raggedright}p{249pt}|}
\hline
  & Small Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 2d10+2 (13 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +6\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares), climb 20 ft., swim 20 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 16 (+1 size, +2 Dex, +3 natural), touch 13, flat-footed 
14\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +2/-2\tabularnewline
\hline
A\textbf{ttack:} & Bite +3 melee (1d4)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +3 melee (1d4)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Stunning shock, lethal shock\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., electricity sense, immunity to 
electricity, low-light vision\tabularnewline
\hline
S\textbf{aves:} & Fort +4, Ref +5, Will +1\tabularnewline
\hline
A\textbf{bilities:} & Str 10, Dex 15, Con 13, Int 2, Wis 12, Cha 6\tabularnewline
\hline
S\textbf{kills:} & Climb +11, Hide +11, Jump +7, Listen +4, Spot +4, Swim +10\tabularnewline
\hline
F\textbf{eats:} & Improved Initiative\tabularnewline
\hline
E\textbf{nvironment:} & Warm marshes\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, clutch (3-5), or colony (6-11)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2\tabularnewline
\hline
T\textbf{reasure:} & 1/10 coins; 50\% goods; 50\% items\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & 3-4 HD (Small); 5-6 HD (Medium)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A shocker lizard has a pale gray or blue underside, shading to a darker hue on 
its back. It has blue-black markings along its back and tail. A shocker lizard 
is about 1 foot tall at the shoulder and weighs about 25 pounds.

COMBAT

A shocker lizard relies on its electricity abilities in combat. A lizard tends 
to bite only after its shock has rendered an opponent unconscious or when the shock 
seems to have no effect at all. A solitary lizard flees once it delivers its shocks, 
but if other shocker lizards are nearby, they all home in on their comrade's discharges 
and attempt to administer their shocks to the foe. 

\textbf{Stunning Shock (Su):} Once per round, a shocker lizard can deliver an electrical 
shock to a single opponent within 5 feet. This attack deals 2d8 points of nonlethal 
damage to living opponents (Reflex DC 12 half ). The save DC is Constitution-based.

\textbf{Lethal Shock (Su):} Whenever two or more shocker lizards are within 20 
feet of each other, they can work together to create a lethal shock. This effect 
has a radius of 20 feet, centered on any one contributing lizard. The shock deals 
2d8 points of electricity damage for each lizard contributing to it, to a maximum 
of 12d8. A Reflex save (DC 10 + number of lizards contributing) reduces the damage 
by half.

\textbf{Electricity Sense (Ex):} Shocker lizards automatically detect any electrical 
discharges within 100 feet.

\textbf{Skills:} Shocker lizards have a +4 racial bonus on Hide checks due to their 
coloration. Shocker lizards have a +2 racial bonus on Listen and Spot checks.

Shocker lizards use their Dexterity modifier instead of their Strength modifier 
for Climb and Jump checks.

A shocker lizard has a +8 racial bonus on Climb checks and can always choose to 
take 10 on a Climb check, even if rushed or threatened.

A shocker lizard has a +8 racial bonus on any Swim check to perform some special 
action or avoid a hazard. It can always choose to take 10 on a Swim check, even 
if distracted or endangered. It can use the run action while swimming, provided 
it swims in a straight line.

\vspace{12pt}
{\LARGE{}SKELETON}

Skeletons are the animated bones of the dead, mindless automatons that obey the 
orders of their evil masters.

A skeleton is seldom garbed in anything more than the rotting remnants of any clothing 
or armor it was wearing when slain. A skeleton does only what it is ordered to 
do. It can draw no conclusions of its own and takes no initiative. Because of this 
limitation, its instructions must always be simple. A skeleton attacks until destroyed.

CREATING A SKELETON

``Skeleton'' is an acquired template that can be added to any corporeal creature 
(other than an undead) that has a skeletal system (referred to hereafter as the 
base creature).

\textbf{Size and Type:} The creature's type changes to undead. It retains any subtype 
except for alignment subtypes (such as good) and subtypes that indicate kind. It 
does not gain the augmented subtype. It uses all the base creature's statistics 
and special abilities except as noted here.

\textbf{Hit Dice: }Drop any Hit Dice gained from class levels (to a minimum of 
1) and raise remaining Hit Dice to d12s. If the creature has more than 20 Hit Dice, 
it can't be made into a skeleton by the \textit{animate dead }spell.

\textbf{Speed: }Winged skeletons can't use their wings to fly. If the base creature 
flew magically, so can the skeleton.

\textbf{Armor Class:} Natural armor bonus changes to a number based on the skeleton's 
size:

\begin{tabular}{|>{\raggedright}p{69pt}|>{\raggedright}p{14pt}|}
\hline
Tiny or smaller & +0\tabularnewline
\hline
Small & +1\tabularnewline
\hline
Medium or Large & +2\tabularnewline
\hline
Huge & +3\tabularnewline
\hline
Gargantuan & +6\tabularnewline
\hline
Colossal & +10\tabularnewline
\hline
\end{tabular}

\textbf{Attacks:} A skeleton retains all the natural weapons, manufactured weapon 
attacks, and weapon proficiencies of the base creature, except for attacks that 
can't work without flesh. A creature with hands gains one claw attack per hand; 
the skeleton can strike with each of its claw attacks at its full attack bonus. 
A skeleton's base attack bonus is equal to 1/2 its Hit Dice.

\textbf{Damage:} Natural and manufactured weapons deal damage normally. A claw 
attack deals damage depending on the skeleton's size. (If the base creature already 
had claw attacks with its hands, use the skeleton claw damage only if it's better.)

\begin{tabular}{|>{\raggedright}p{75pt}|>{\raggedright}p{13pt}|}
\hline
Diminutive or Fine & 1\tabularnewline
\hline
Tiny & 1d2\tabularnewline
\hline
Small & 1d3\tabularnewline
\hline
Medium & 1d4\tabularnewline
\hline
Large & 1d6\tabularnewline
\hline
Huge & 1d8\tabularnewline
\hline
Gargantuan & 2d6\tabularnewline
\hline
Colossal & 2d8\tabularnewline
\hline
\end{tabular}

\textbf{Special Attacks: }A skeleton retains none of the base creature's special 
attacks.

\textbf{Special Qualities:} A skeleton loses most special qualities of the base 
creature. It retains any extraordinary special qualities that improve its melee 
or ranged attacks. A skeleton gains the following special qualities.

\textit{Immunity to Cold (Ex): }Skeletons are not affected by cold.

\textit{Damage Reduction 5/Bludgeoning: }Skeletons lack flesh or internal organs.

\textbf{Saves:} Base save bonuses are Fort +1/3 HD, Ref +1/3 HD, and Will +1/2 
HD + 2.

\textbf{Abilities:} A skeleton's Dexterity increases by +2, it has no Constitution 
or Intelligence score, its Wisdom changes to 10, and its Charisma changes to 1.

\textbf{Skills:} A skeleton has no skills.

\textbf{Feats: }A skeleton loses all feats of the base creature and gains Improved 
Initiative.

\textbf{Environment}: Any, usually same as base creature.

\textbf{Organization:} Any.

\textbf{Challenge Rating:} Depends on Hit Dice, as follows:

\begin{tabular}{|>{\raggedright}p{34pt}|>{\raggedright}p{73pt}|}
\hline
H\textbf{it Dice} & C\textbf{hallenge Rating}\tabularnewline
\hline
1/2 & 1/6\tabularnewline
\hline
1 & 1/3\tabularnewline
\hline
2-3 & 1\tabularnewline
\hline
4-5 & 2\tabularnewline
\hline
6-7 & 3\tabularnewline
\hline
8-9 & 4\tabularnewline
\hline
10-11 & 5\tabularnewline
\hline
12-14 & 6\tabularnewline
\hline
15-17 & 7\tabularnewline
\hline
18-20 & 8\tabularnewline
\hline
\end{tabular}

\textbf{Treasure: }None.

\textbf{Alignment: }Always neutral evil.

\textbf{Advancement: }As base creature (or --- if the base creature advances by 
character class).

\textbf{Level Adjustment:} ---.

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{69pt}|>{\raggedright}p{77pt}|>{\raggedright}p{77pt}|>{\raggedright}p{77pt}|}
\hline
  & H\textbf{uman Warrior Skeleton } & W\textbf{olf Skeleton } & O\textbf{wlbear 
Skeleton}\tabularnewline
\hline
  & Medium Undead  & Medium Undead  & Large Undead\tabularnewline
\hline
H\textbf{it Dice: } & 1d12 (6 hp)  & 2d12 (13 hp)  & 5d12 (32 hp)\tabularnewline
\hline
I\textbf{nitiative: } & +5  & +7  & +6\tabularnewline
\hline
S\textbf{peed: } & 30 ft. (6 squares)  & 50 ft. (10 squares)  & 30 ft. (6 squares)\tabularnewline
\hline
A\textbf{rmor Class: } & 15 (+1 Dex, +2 natural, +2 heavy steel shield), touch 
11, flat-footed 14 & 15 (+3 Dex, +2 natural), touch 13, flat-footed 12 & 13 (-1 
size, +2 Dex, +2 natural), touch 11, flat-footed 11\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +0/+1 & +1/+2 & +2/+11\tabularnewline
\hline
A\textbf{ttack:} & Scimitar +1 melee (1d6+1/18-20) or claw +1 melee (1d4+1) & Bite 
+2 melee (1d6+1) & Claw +6 melee (1d6+5)\tabularnewline
\hline
F\textbf{ull Attack:} & Scimitar +1 melee (1d6+1/18-20) or 2 claws +1 melee (1d4+1) & Bite 
+2 melee (1d6+1) & 2 claws +6 melee (1d6+5) and bite +1 melee (1d8+2)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. & 5 ft./5 ft. & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks: ---}--- & --- &  & \tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 5/bludgeoning, darkvision 60 ft., 
immunity to cold, undead traits & Damage reduction 5/bludgeoning, darkvision 60 
ft., immunity to cold, undead traits & Damage reduction 5/bludgeoning, darkvision 
60 ft., immunity to cold, undead traits\tabularnewline
\hline
S\textbf{aves:} & Fort +0, Ref +1, Will +2 & Fort +0, Ref +3, Will +3 & Fort +1, 
Ref +3, Will +4\tabularnewline
\hline
A\textbf{bilities: Str 13, Dex 13, Con ---, Int ---, Wis 10, Cha 1} & Str 13, Dex 
17, Con ---, Int ---, Wis 10, Cha 1 & Str 21, Dex 14, Con ---, Int ---, Wis 10, 
Cha 1 & \tabularnewline
\hline
F\textbf{eats:} & Improved Initiative & Improved Initiative & Improved Initiative\tabularnewline
\hline
E\textbf{nvironment:} & Temperate plains & Temperate forests & Temperate forests\tabularnewline
\hline
O\textbf{rganization:} & Any & Any & Any\tabularnewline
\hline
C\textbf{hallenge Rating:} & 1/3 & 1 & 2\tabularnewline
\hline
T\textbf{reasure:} & None & None & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral evil & Always neutral evil & Always neutral 
evil\tabularnewline
\hline
A\textbf{dvancement:}--- &  & 3 HD (Medium); 4-6 HD (Large) & 6-8 HD (Large); 9-15 
HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & --- & --- & \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{49pt}|>{\raggedright}p{84pt}|>{\raggedright}p{84pt}|>{\raggedright}p{84pt}|}
\hline
  & T\textbf{roll Skeleton} & C\textbf{himera Skeleton} & E\textbf{ttin Skeleton}\tabularnewline
\hline
  & Large Undead & Large Undead & Large Undead\tabularnewline
\hline
H\textbf{it Dice:} & 6d12 (39 hp) & 9d12 (58 hp) & 10d12 (65 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +7 & +6 & +4\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares) & 30 ft. (6 squares) & 40 ft. (8 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 14 (+3 Dex, -1 size, +2 natural), touch 12, flat-footed 
11 & 13 (+2 Dex., -1 size, +2 natural), touch 11, flat-footed 11 & 11 (-1 size, 
+2 natural), touch 9, flat-footed 11\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +3/+13 & +4/+12 & +5/+15\tabularnewline
\hline
A\textbf{ttack:} & Claw +8 melee (1d6+6) & Bite +7 melee (2d6+4) & Morningstar 
+10 melee (2d6+6) or claw +10 melee (1d6+6) or javelin +4 ranged (1d8+6)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +8 melee (1d6+6) and bite +3 melee (1d6+3) & Bite 
+7 melee (2d6+4) and bite+7 melee (1d8+4) and gore +7 melee (1d8+4) and 2 claws 
+2 melee (1d6+2) & 2 morningstars +10 melee (2d6+6) or 2 claws +10 melee (1d6+6) 
or 2 javelins +4 ranged (1d8+6)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft & 10 ft./5 ft & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:}--- & --- & --- & \tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 5/bludgeoning, darkvision 60 ft., 
immunity to cold, undead traits & Damage reduction 5/bludgeoning, darkvision 60 
ft., immunity to cold, undead traits & Damage reduction 5/bludgeoning, darkvision 
60 ft., darkvision 60 ft., immunity to cold, superior twoweapon\tabularnewline
\hline
S\textbf{aves:} & Fort +2, Ref +5, Will +5 & Fort +3, Ref +5, Will +6 & Fort +3, 
Ref +3, Will +7\tabularnewline
\hline
A\textbf{bilities:} & Str 23, Dex 16, Con ---, Int ---, Wis 10, Cha 1 & Str 19, 
Dex 15, Con ---, Int ---, Wis 10, Cha 1 & Str 23, Dex 10, Con ---, Int ---, Wis 
10, Cha 1\tabularnewline
\hline
F\textbf{eats:} & Improved Initiative & Improved Initiative & Improved Initiative\tabularnewline
\hline
E\textbf{nvironment:} & Cold mountains & Temperate hills & Cold hills\tabularnewline
\hline
O\textbf{rganization:} & Any & Any & Any\tabularnewline
\hline
C\textbf{hallenge Rating:} & 3 & 4 & 5\tabularnewline
\hline
T\textbf{reasure:} & None & None & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral evil & Always neutral evil & Always neutral 
evil\tabularnewline
\hline
A\textbf{dvancement:}--- &  & 10-13 HD (Large); 14-27 HD (Huge)--- & \tabularnewline
\hline
L\textbf{evel Adjustment:}--- & --- & --- & \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{49pt}|>{\raggedright}p{84pt}|>{\raggedright}p{84pt}|>{\raggedright}p{84pt}|}
\hline
  & A\textbf{dvanced Megaraptor Skeleton} & C\textbf{loud Giant Skeleton} & Y\textbf{oung 
Adult Red Dragon Skeleton}\tabularnewline
\hline
  & Huge Undead & Huge Undead & Huge Undead (Fire)\tabularnewline
\hline
H\textbf{it Dice:} & 12d12 (78 hp) & 17d12 (110 hp) & 19d12 (123 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +7 & +6 & +5\tabularnewline
\hline
S\textbf{peed:} & 60 ft. (12 squares) & 50 ft. (10 squares) & 40 ft. (8 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 14 (-2 size, +3 Dex, +3 natural), touch 11, flat-footed 
11 & 13 (-2 size, +2 Dex, +3 natural), touch 10, flat-footed 11 & 12 (-2 size, 
+1 Dex, +3 natural), touch 9, flat-footed 11\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +6/+19 & +8/+28 & +9/+27\tabularnewline
\hline
A\textbf{ttack:} & Talons +9 melee (2d8+5) & Gargantuan morningstar +18 melee (4d6+18) 
or claw +18 melee (1d8+12) or rock +8 ranged (2d8+12) & Bite +17 melee (2d8+10)\tabularnewline
\hline
F\textbf{ull Attack:} & Talons +9 melee (2d8+5) and 2 claws +4 melee (1d8+2) and 
bite +4 melee (2d6+2) & Gargantuan morningstar +18/+13 melee (4d6+18) or 2 claws 
+18 melee (1d8+12) or rock +8 ranged (2d8+12) & Bite +17 melee (2d8+10) and 2 claws 
+12 melee (2d6+5) and 2 wings +12 melee (1d8+5) and tail slap +12 melee (2d6+15)\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./10 ft. & 15 ft./15 ft. & 15 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:}--- & --- & --- & \tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 5/bludgeoning, darkvision 60 ft., 
immunity to cold, undead traits & Damage reduction 5/bludgeoning, darkvision 60 
ft., immunity to cold, oversize weapon, undead traits & Damage reduction 5/bludgeoning, 
darkvision 60 ft., immunity to cold and fire, undead traits\tabularnewline
\hline
S\textbf{aves:} & Fort +4, Ref +7, Will +8 & Fort +5, Ref +7, Will +10 & Fort +6, 
Ref +7, Will +8\tabularnewline
\hline
A\textbf{bilities:} & Str 21, Dex 17, Con ---, Int ---, Wis 10, Cha 1 & Str 35, 
Dex 15, Con ---, Int ---, Wis 10, Cha 1 & Str 31, Dex 12, Con ---, Int ---, Wis 
10, Cha 1\tabularnewline
\hline
F\textbf{eats:} & Improved Initiative & Improved Initiative & Improved Initiative\tabularnewline
\hline
E\textbf{nvironment:} & Warm forests & Temperate mountains & Warm mountains\tabularnewline
\hline
O\textbf{rganization:} & Any & Any & Any\tabularnewline
\hline
C\textbf{hallenge Rating:} & 6 & 7 & 8\tabularnewline
\hline
T\textbf{reasure:} & None & None & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral evil & Always neutral evil & Always neutral 
evil\tabularnewline
\hline
A\textbf{dvancement:} & 13-16 HD (Huge); 17-20 HD (Gargantuan)--- &  & 20 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & --- & --- & \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
{\LARGE{}SKUM}

\begin{tabular}{|>{\raggedright}p{88pt}|>{\raggedright}p{237pt}|}
\hline
  & Medium Aberration (Aquatic)\tabularnewline
\hline
H\textbf{it Dice:} & 2d8+2 (11 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), swim 40 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 13 (+1 Dex, +2 natural), touch 11, flat-footed 12\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +1/+5\tabularnewline
\hline
A\textbf{ttack:} & Bite +5 melee (2d6+4)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +5 melee (2d6+4) and 2 claws +0 melee (1d4+2)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Rake 1d6+2\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., amphibious\tabularnewline
\hline
S\textbf{aves:} & Fort +1, Ref +1, Will +3\tabularnewline
\hline
A\textbf{bilities:} & Str 19, Dex 13, Con 13, Int 10, Wis 10, Cha 6\tabularnewline
\hline
S\textbf{kills:} & Hide +6*, Listen +7*, Move Silently +6, Spot +7*, Swim +12\tabularnewline
\hline
F\textbf{eats:} & Alertness\tabularnewline
\hline
E\textbf{nvironment:} & Underground\tabularnewline
\hline
O\textbf{rganization:} & Brood (2-5) or pack (6-15)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Usually lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & 3-4 HD (Medium); 5-6 HD (Large)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +3\tabularnewline
\hline
\end{tabular}

A skum is about the same height and weight as a human. Skum speak Aquan.

COMBAT

\textbf{Rake (Ex):} Attack bonus +0 melee, damage 1d6+2. A skum also gains two 
rake attacks when it attacks while swimming.

\textbf{Skills:} *Skum have a +4 racial bonus on Hide, Listen, and Spot checks 
underwater.

A skum has a +8 racial bonus on any Swim check to perform some special action or 
avoid a hazard. It can always choose to take 10 on a Swim check, even if distracted 
or endangered. It can use the run action while swimming, provided it swims in a 
straight line.

\vspace{12pt}
{\LARGE{}SPECTRE}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{227pt}|}
\hline
  & Medium Undead (Incorporeal)\tabularnewline
\hline
H\textbf{it Dice:} & 7d12 (45 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +7\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares), fly 80 ft. (perfect)\tabularnewline
\hline
A\textbf{rmor Class:} & 15 (+3 Dex, +2 deflection), touch 15, flat-footed 13\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +3/---\tabularnewline
\hline
A\textbf{ttack:} & Incorporeal touch +6 melee (1d8 plus energy drain)\tabularnewline
\hline
F\textbf{ull Attack:} & Incorporeal touch +6 melee (1d8 plus energy drain)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Energy drain, create spawn\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., incorporeal traits, +2 turn resistance, 
sunlight powerlessness, undead traits, unnatural aura\tabularnewline
\hline
S\textbf{aves:} & Fort +2, Ref +5, Will +7\tabularnewline
\hline
A\textbf{bilities:} & Str ---, Dex 16, Con ---, Int 14, Wis 14, Cha 15\tabularnewline
\hline
S\textbf{kills:} & Hide +13, Intimidate +12, Knowledge (religion) +12, Listen +14, 
Search +12, Spot +14, Survival +2 (+4 following tracks)\tabularnewline
\hline
F\textbf{eats:} & Alertness, Blind-Fight, Improved Initiative\tabularnewline
\hline
E\textbf{nvironment:} & Any land and underground\tabularnewline
\hline
O\textbf{rganization:} & Solitary, gang (2-4), or swarm (6-11)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 7\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & 8-14 HD (Medium)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A spectre looks much as it did in life and can be easily recognized by those who 
knew the individual or have seen the individual's face in a painting or a drawing. 
In many cases, the evidence of a violent death is visible on its body. A spectre 
is roughly human-sized and is weightless.

COMBAT

In close combat a spectre attacks with its numbing, life-draining touch. It makes 
full use of its incorporeal nature, moving through walls, ceilings, and floors 
as it attacks.

\textbf{Energy Drain (Su):} Living creatures hit by a spectre's incorporeal touch 
attack gain two negative levels. The DC is 15 for the Fortitude save to remove 
a negative level. The save DC is Charisma-based. For each such negative level bestowed, 
the spectre gains 5 temporary hit points.

\textbf{Create Spawn (Su):} Any humanoid slain by a spectre becomes a spectre in 
1d4 rounds. Spawn are under the command of the spectre that created them and remain 
enslaved until its death. They do not possess any of the abilities they had in 
life.

\textbf{Unnatural Aura (Su):} Animals, whether wild or domesticated, can sense 
the unnatural presence of a spectre at a distance of 30 feet. They do not willingly 
approach nearer than that and panic if forced to do so; they remain panicked as 
long as they are within that range.

\textbf{Sunlight Powerlessness (Ex): }Spectres are powerless in natural sunlight 
(not merely a \textit{daylight }spell) and flee from it. A spectre caught in sunlight 
cannot attack and can take only a single move or attack action in a round.

\vspace{12pt}
{\LARGE{}SPHINX}

Sphinxes are enigmatic creatures with great, feathery wings and leonine bodies. 
All sphinxes are territorial, but the more intelligent ones can differentiate between 
deliberate intrusion and temporary or inadvertent trespass.

A typical sphinx is about 10 feet long and weighs about 800 pounds.

Sphinxes speak Sphinx, Common, and Draconic.

Combat

Most sphinxes fight on the ground, using their wings to help them pounce much as 
lions do. If outnumbered by earthbound creatures, a sphinx takes wing and attacks 
on the fly.

\textbf{Pounce (Ex):} If a sphinx charges a foe, it can make a full attack, including 
two rake attacks.

\textbf{Rake (Ex):} A sphinx that pounces onto a creature can make two rake attacks 
with its hind legs. Each sphinx's description provides its attack bonus and damage.

\vspace{12pt}
ANDROSPHINX

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{204pt}|}
\hline
  & Large Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 12d10+48 (114 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +0\tabularnewline
\hline
S\textbf{peed:} & 50 ft. (10 squares), fly 80 ft. (poor)\tabularnewline
\hline
A\textbf{rmor Class:} & 22 (-1 size, +13 natural), touch 9, flat-footed 22\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +12/+23\tabularnewline
\hline
A\textbf{ttack:} & Claw +18 melee (2d4+7)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +18 melee (2d4+7)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Pounce, rake 2d4+3, roar, spells\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., low-light vision\tabularnewline
\hline
S\textbf{aves:} & Fort +12, Ref +8, Will +7\tabularnewline
\hline
A\textbf{bilities:} & Str 25, Dex 10, Con 19, Int 16, Wis 17, Cha 17\tabularnewline
\hline
S\textbf{kills:} & Intimidate +17, Knowledge (any one) +18, Listen +18, Spot +18, 
Survival +18\tabularnewline
\hline
F\textbf{eats:} & Alertness, Cleave, Great Cleave, Flyby Attack, Power Attack, 
Track\tabularnewline
\hline
E\textbf{nvironment:} & Warm deserts\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 9\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic good\tabularnewline
\hline
A\textbf{dvancement:} & 13-18 HD (Large); 19-36 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +5 (cohort)\tabularnewline
\hline
\end{tabular}

These sphinxes are always male. Androsphinxes are clever and generally good-natured, 
but they can be savage opponents. 

Combat

In battle, an androsphinx rips apart enemies with its razor-sharp claws. It relies 
on its natural weapons in a fight, employing its spells for defense or healing.

\textbf{Rake (Ex): }Attack bonus +18 melee, damage 2d4+3.

\textbf{Roar (Su): }Three times per day an androsphinx can loose a mighty roar. 
The first time it does this, all creatures within 500 feet must succeed on a DC 
19 Will save or be affected as though by a \textit{fear }spell for 2d6 rounds.

If the sphinx roars a second time during the same encounter, all creatures within 
250 feet must succeed on a DC 19 Fortitude save or be paralyzed for 1d4 rounds, 
and all those within 90 feet are deafened for 2d6 rounds (no save). If it roars 
a third time during the same encounter, all those within 250 feet must succeed 
on a DC 19 Fortitude save or take 2d4 points of Strength damage for 2d4 rounds. 
In addition, any Medium or smaller creature within 90 feet must succeed on a DC 
19 Fortitude save or be thrown to the ground and take 2d8 points of damage. The 
force of this roar is so great that it deals 50 points of damage to any stone or 
crystalline object within 90 feet. Magic items and held or carried items can avoid 
damage with a DC 19 Reflex save. Other androsphinxes are immune to these effects. 
The save DCs are Charisma-based.

\textbf{Spells:} An androsphinx casts divine spells as a 6th-level cleric from 
the cleric spell list and from the Good, Healing, and Protection domains. 

\textit{Typical Cleric Spells Prepared }(5/5/5/4; save DC 13 + spell level): 0---\textit{cure 
minor wounds, detect magic, guidance, light, resistance; }1st---\textit{divine 
favor, protection from evil*, shield of faith, remove fear, summon monster I; }2nd---\textit{bull's 
strength, remove paralysis, resist energy, shield other*, summon monster II; }3rd---\textit{cure 
serious wounds*, daylight, invisibility purge, searing light.}

*Domain spell. Domains: Good and Healing.

\vspace{12pt}
CRIOSPHINX

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{227pt}|}
\hline
  & Large Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 10d10+30 (85 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +0\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares), fly 60 ft. (poor)\tabularnewline
\hline
A\textbf{rmor Class:} & 20 (-1 size, +11 natural), touch 9, flat-footed 20\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +10/+20\tabularnewline
\hline
A\textbf{ttack:} & Gore +15 melee (2d6+6)\tabularnewline
\hline
F\textbf{ull Attack:} & Gore +15 melee (2d6+6) and 2 claws +10 melee (1d6+3)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Pounce, rake 1d6+3\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., low-light vision\tabularnewline
\hline
S\textbf{aves:} & Fort +10, Ref +7, Will +3\tabularnewline
\hline
A\textbf{bilities:} & Str 23, Dex 10, Con 17, Int 10, Wis 11, Cha 11\tabularnewline
\hline
S\textbf{kills:} & Intimidate +8, Listen +11, Spot +1\tabularnewline
\hline
F\textbf{eats:} & Alertness, Cleave, Flyby Attack, Power Attack\tabularnewline
\hline
E\textbf{nvironment:} & Warm deserts\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 7\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & 11-15 HD (Large); 16-30 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +3 (cohort)\tabularnewline
\hline
\end{tabular}

These sphinxes are always male. Neither good nor evil, they lack the intelligence 
of the androsphinx.

Combat

Criosphinxes attack with their claws, as do their kin, but they can also butt with 
their horns. They don't cast spells and employ only the most simple battle tactics.

\textbf{Rake (Ex):} Attack bonus +15 melee, damage 1d6+3.

\vspace{12pt}
GYNOSPHINX

\begin{tabular}{|>{\raggedright}p{86pt}|>{\raggedright}p{239pt}|}
\hline
  & Large Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 8d10+8 (52 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +5\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares), fly 60 ft. (poor)\tabularnewline
\hline
A\textbf{rmor Class:} & 21 (-1 size, +1 Dex, +11 natural), touch 10, flat-footed 
20\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +8/+16\tabularnewline
\hline
A\textbf{ttack:} & Claw +11 melee (1d6+4)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +11 melee (1d6+4)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Pounce, rake 1d6+2, spell-like abilities\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., low-light vision\tabularnewline
\hline
S\textbf{aves:} & Fort +7, Ref +7, Will +8\tabularnewline
\hline
A\textbf{bilities:} & Str 19, Dex 12, Con 13, Int 18, Wis 19, Cha 19\tabularnewline
\hline
S\textbf{kills:} & Bluff +15, Concentration +12, Diplomacy +8, Disguise +4 (+6 
acting), Intimidate +13, Listen +17, Sense Motive +15, Spot +17\tabularnewline
\hline
F\textbf{eats:} & Combat Casting, Improved Initiative, Iron Will\tabularnewline
\hline
E\textbf{nvironment:} & Warm deserts\tabularnewline
\hline
O\textbf{rganization:} & Solitary or covey (2-4)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 8\tabularnewline
\hline
T\textbf{reasure:} & Double standard\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & 9-12 HD (Large); 13-24 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +4 (cohort)\tabularnewline
\hline
\end{tabular}

These sphinxes are the female counterparts of androsphinxes. 

Combat

In close combat, gynosphinxes use their powerful claws to flay the flesh from their 
enemies. Despite their deadly nature, they prefer to avoid combat whenever possible.

\textbf{Rake (Ex):} Attack bonus +11 melee, damage 1d6+2.

\textbf{Spell-Like Abilities:} 3/day---\textit{clairaudience/clairvoyance, detect 
magic, read magic, see invisibility}; 1/day---\textit{comprehend languages, locate 
object, dispel magic, remove curse }(DC 18), \textit{legend lore}. Caster level 
14th. The save DC is Charisma-based.

Once per week a gynosphinx can create a \textit{symbol of death, }a \textit{symbol 
of fear, }a \textit{symbol of insanity, }a \textit{symbol of pain, }a \textit{symbol 
of persuasion, }a \textit{symbol of sleep, }and a \textit{symbol of stunning }as 
the spells (caster level 18th), except that all save DCs are 22 and each symbol 
remains a maximum of one week once scribed. The save DCs are Charisma-based.

\vspace{12pt}
HIERACOSPHINX

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{226pt}|}
\hline
  & Large Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 9d10+18 (67 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +2\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares), fly 90 ft. (poor)\tabularnewline
\hline
A\textbf{rmor Class:} & 19 (-1 size, +2 Dex, +8 natural), touch 11, flat-footed 
17\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +9/+18\tabularnewline
\hline
A\textbf{ttack:} & Bite +13 melee (1d10+5)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +13 melee (1d10+5) and 2 claws +8 melee (1d6+2)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Pounce, rake 1d6+2\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., lowlight vision\tabularnewline
\hline
S\textbf{aves:} & Fort +8, Ref +8, Will +5\tabularnewline
\hline
A\textbf{bilities:} & Str 21, Dex 14, Con 15, Int 6, Wis 15, Cha 10\tabularnewline
\hline
S\textbf{kills:} & Listen +10, Spot +14 \tabularnewline
\hline
F\textbf{eats:} & Alertness, Cleave, Flyby Attack, Power Attack\tabularnewline
\hline
E\textbf{nvironment:} & Warm deserts\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or flock (4-7)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 5\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 10-14 HD (Large); 15-27 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +3 (cohort)\tabularnewline
\hline
\end{tabular}

Combat

Hieracosphinxes can make short work of even the most dangerous opponents with their 
claws. They are not particularly intelligent, but are cunning enough to dive at 
their enemies from above with their flying ability.

\textbf{Rake (Ex):} Attack bonus +13 melee, damage 1d6+2.

\textbf{Skills:} Hieracosphinxes have a +4 racial bonus on Spot checks.

\vspace{12pt}
{\LARGE{}SPIDER EATER}

\begin{tabular}{|>{\raggedright}p{87pt}|>{\raggedright}p{238pt}|}
\hline
  & Large Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 4d10+20 (42 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares), fly 60 ft. (good)\tabularnewline
\hline
A\textbf{rmor Class:} & 14 (-1 size, +1 Dex, +4 natural), touch 10, flatfooted 
13\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +4/+13\tabularnewline
\hline
A\textbf{ttack:} & Sting +8 melee (1d8+5 plus poison)\tabularnewline
\hline
F\textbf{ull Attack:} & Sting +8 melee (1d8+5 plus poison) and bite +3 melee (1d8+2)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Implant, poison\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., freedom of movement, lowlight 
vision, scent\tabularnewline
\hline
S\textbf{aves:} & Fort +9, Ref +5, Will +2\tabularnewline
\hline
A\textbf{bilities:} & Str 21, Dex 13, Con 21, Int 2, Wis 12, Cha 10\tabularnewline
\hline
S\textbf{kills:} & Listen +10, Spot +11\tabularnewline
\hline
F\textbf{eats:} & Alertness, Dodge\tabularnewline
\hline
E\textbf{nvironment:} & Temperate forests\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 5\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & 5-12 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A spider eater is about 10 feet long and 4 feet high, and has a wingspan of about 
20 feet. It weighs about 4,000 pounds.

COMBAT

A spider eater attacks with its venomous sting and powerful mandibles. Its usual 
tactic is to deliver a sting, then back off, hovering out of reach until the venom 
takes effect.

\textbf{Implant (Ex): }Female spider eaters lay their eggs inside paralyzed creatures 
of Large or larger size. The young emerge about six weeks later, literally devouring 
the host from inside.

\textbf{Poison (Ex):} Injury, Fortitude DC 17, initial damage none, secondary damage 
paralysis for 1d8+5 weeks. The save DC is Constitution-based.

\textbf{Freedom of Movement (Su): }Spider eaters have a continuous freedom of movement 
ability as the spell (caster level 12th). When the spider eater serves as a mount, 
this effect does not extend to its rider.

\textbf{Skills:} Spider eaters have a +4 racial bonus on Listen and Spot checks.

TRAINING A SPIDER EATER

A spider eater requires training before it can bear a rider in combat.

Training a spider eater requires six weeks of work and a DC 25 Handle Animal check. 
Riding a spider eater requires an exotic saddle. A spider eater can fight while 
carrying a rider, but the rider cannot also attack unless he or she succeeds on 
a Ride check. 

Spider eater eggs are worth 2,000 gp apiece on the open market, while young are 
worth 3,000 gp each. Professional trainers charge 3,000 gp to rear or train a spider 
eater.

\textbf{Carrying Capacity: }A light load for a spider eater is up to 306 pounds; 
a medium load, 307-612 pounds; and a heavy load, 613-920 pounds.

\vspace{12pt}
{\LARGE{}SPRITE}

Sprites are reclusive fey. They go out of their way to fight evil and ugliness 
and to protect their homelands. Combat

Sprites fight their opponents with spell-like abilities and pintsized weaponry. 
They prefer ambushes and other trickery over direct confrontation.

\textbf{Skills:} All sprites have a +2 racial bonus on Search, Spot, and Listen 
checks.

\vspace{12pt}
GRIG

\begin{tabular}{|>{\raggedright}p{78pt}|>{\raggedright}p{247pt}|}
\hline
  & Tiny Fey\tabularnewline
\hline
H\textbf{it Dice:} & 1/2 d6+1 (2 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +4\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), fly 40 ft. (poor)\tabularnewline
\hline
A\textbf{rmor Class:} & 18 (+2 size, +4 Dex, +2 natural), touch 16, flat-footed 
16\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +0/-11\tabularnewline
\hline
A\textbf{ttack:} & Short sword +6 melee (1d3-3/19-20) or longbow +6 ranged (1d4-3/x3)\tabularnewline
\hline
F\textbf{ull Attack:} & Short sword +6 melee (1d3-3/19-20) or longbow +6 ranged 
(1d4-3/x3)\tabularnewline
\hline
S\textbf{pace/Reach:} & 2-1/2 ft./0 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Spell-like abilities, fiddle\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 5/cold iron, low-light vision, spell 
resistance 17\tabularnewline
\hline
S\textbf{aves:} & Fort +1, Ref +6, Will +3\tabularnewline
\hline
A\textbf{bilities:} & Str 5, Dex 18, Con 13, Int 10, Wis 13, Cha 14\tabularnewline
\hline
S\textbf{kills:} & Craft (any one) +4, Escape Artist +8, Hide +16, Jump +3, Listen 
+3, Move Silently +8*, Perform (string instruments) +6, Search +2, Spot +3\tabularnewline
\hline
F\textbf{eats:} & Dodge\textsuperscript{B}, Weapon Finesse\tabularnewline
\hline
E\textbf{nvironment:} & Temperate forests\tabularnewline
\hline
O\textbf{rganization:} & Gang (2-4), band (6-11), or tribe (20-80)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 1\tabularnewline
\hline
T\textbf{reasure:} & No coins; 50\% goods; 50\% items\tabularnewline
\hline
A\textbf{lignment:} & Always neutral good\tabularnewline
\hline
A\textbf{dvancement:} & 1-3 HD (Tiny)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +3\tabularnewline
\hline
\end{tabular}

Grigs can leap great distances. They have light blue skin, forest-green hair, and 
brown hairy legs, and usually wear tunics or brightly colored vests with buttons 
made from tiny gems. A grig stands 1-1/2 feet tall and weighs about 1 pound. 

Grigs speak Sylvan. Some also speak Common.

Combat

Grigs are fierce by sprite standards, attacking opponents fearlessly with bow and 
dagger.

\textbf{Spell-Like Abilities:} 3/day---\textit{disguise self, entangle }(DC 13)\textit{, 
invisibility }(self only), \textit{pyrotechnics }(DC 14)\textit{, ventriloquism 
}(DC 13)\textit{. }Caster level 9th. The save DCs are Charisma-based.

\textbf{Fiddle (Su): }One grig in each band carries a tiny, grig-sized fiddle. 
When the fiddler plays, any nonsprite within 30 feet of the instrument must succeed 
on a DC 12 Will save or be affected as though by \textit{irresistible dance }for 
as long as the playing continues. The save DC is Charisma-based.

\textbf{Skills:} Grigs have a +8 racial bonus on Jump checks. *They also have a 
+5 racial bonus on Move Silently checks in a forest setting.

\vspace{12pt}
NIXIE

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{195pt}|}
\hline
  & Small Fey (Aquatic)\tabularnewline
\hline
H\textbf{it Dice:} & 1d6 (3 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +3\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), swim 30 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 14 (+1 size, +3 Dex), touch 14, flat-footed 11\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +0/-6\tabularnewline
\hline
A\textbf{ttack:} & Short sword +4 melee (1d4-2/19-20) or light crossbow +4 ranged 
(1d6/19-20)\tabularnewline
\hline
F\textbf{ull Attack:} & Short sword +4 melee (1d4-2/19-20) or light crossbow +4 
ranged (1d6/19-20)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & C\textit{harm person}\tabularnewline
\hline
S\textbf{pecial Qualities:} & Amphibious, damage reduction 5/cold iron, low-light 
vision, spell resistance 16, \textit{water breathing, }wild empathy\tabularnewline
\hline
S\textbf{aves:} & Fort +0, Ref +5, Will +3\tabularnewline
\hline
A\textbf{bilities:} & Str 7, Dex 16, Con 11, Int 12, Wis 13, Cha 18\tabularnewline
\hline
S\textbf{kills:} & Bluff +8, Craft (any one) +5, Escape Artist +6, Handle Animal 
+8, Hide +7*, Listen +6, Perform (sing) +7, Search +3, Sense Motive +5, Spot +6, 
Swim +6\tabularnewline
\hline
F\textbf{eats:} & Dodge\textsuperscript{B}, Weapon Finesse\tabularnewline
\hline
E\textbf{nvironment:} & Temperate aquatic\tabularnewline
\hline
O\textbf{rganization:} & Gang (2-4), band (6-11), or tribe (20-80)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 1\tabularnewline
\hline
T\textbf{reasure:} & No coins; 50\% goods (metal or stone only); 50\% items (no 
scrolls)\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & 2-3 HD (Small)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +3\tabularnewline
\hline
\end{tabular}

Most nixies are slim and comely, with lightly scaled, pale green skin and dark 
green hair. Females often twine shells and pearl strings in their hair and dress 
in wraps woven from colorful seaweed. Males wear loincloths of the same materials. 
Nixies prefer not to leave their lakes.

A nixie stands about 4 feet tall and weighs about 45 pounds.

Nixies speak Aquan and Sylvan. Some also speak Common.

Combat

Nixies rely on their \textit{charm person }ability to deter enemies, entering combat 
only to protect themselves and their territory.

\textit{\textbf{Charm Person }}\textbf{(Sp): }A nixie can use \textit{charm person 
}three times per day as the spell (caster level 4th). Those affected must succeed 
on a DC 15 Will save or be \textit{charmed }for 24 hours. Most \textit{charmed 
}creatures are used to perform heavy labor, guard duty, and other onerous tasks 
for the nixie community. Shortly before the effect wears off, the nixie escorts 
the \textit{charmed }creature away and orders it to keep walking. The save DC is 
Charisma-based.

\textbf{Amphibious (Ex): }Although nixies are aquatic, they can survive indefinitely 
on land.

\textit{\textbf{Water Breathing }}\textbf{(Sp): }Once per day a nixie can use \textit{water 
breathing }as the spell (caster level 12th). Nixies usually bestow this effect 
on those they have \textit{charmed.}

\textbf{Wild Empathy (Ex): }This ability works like the druid's wild empathy class 
feature, except that a nixie has a +6 racial bonus on the check.

\textbf{Skills:} A nixie has a +8 racial bonus on any Swim check to perform some 
special action or avoid a hazard. It can always choose to take 10 on a Swim check, 
even if distracted or endangered. It can use the run action while swimming, provided 
it swims in a straight line. *Nixies have a +5 racial bonus on Hide checks when 
in the

water.

\vspace{12pt}
PIXIE

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{235pt}|}
\hline
  & Small Fey\tabularnewline
\hline
H\textbf{it Dice:} & 1d6 (3 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +4\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), fly 60 ft. (good)\tabularnewline
\hline
A\textbf{rmor Class:} & 16 (+1 size, +4 Dex, +1 natural), touch 15, flat-footed 
12\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +0/-6\tabularnewline
\hline
A\textbf{ttack:} & Short sword +5 melee (1d4-2/19-20) or longbow +5 ranged (1d6-2/3)\tabularnewline
\hline
F\textbf{ull Attack:} & Short sword +5 melee (1d4-2/19-20) or longbow +5 ranged 
(1d6-2)/x3\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Spell-like abilities, special arrows \tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 10/cold iron, greater invisibility, 
low-light vision, spell resistance 15\tabularnewline
\hline
S\textbf{aves:} & Fort +0, Ref +6, Will +4\tabularnewline
\hline
A\textbf{bilities:} & Str 7, Dex 18, Con 11, Int 16, Wis 15, Cha 16\tabularnewline
\hline
S\textbf{kills:} & Bluff +7, Concentration +4, Escape Artist +8, Hide +8, Listen 
+8, Move Silently +8, Ride +8, Search +9, Sense Motive +6, Spot +8\tabularnewline
\hline
F\textbf{eats:} & Dodge\textsuperscript{B}, Weapon Finesse\tabularnewline
\hline
E\textbf{nvironment:} & Temperate forests\tabularnewline
\hline
O\textbf{rganization:} & Gang (2-4), band (6-11), or tribe (20-80)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 4 (5 with \textit{irresistible dance})\tabularnewline
\hline
T\textbf{reasure:} & No coins; 50\% goods; 50\% items\tabularnewline
\hline
A\textbf{lignment:} & Always neutral good\tabularnewline
\hline
A\textbf{dvancement:} & 2-3 HD (Small)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +4 (+6 with \textit{irresistible dance})\tabularnewline
\hline
\end{tabular}

Pixies wear bright clothing, often including a cap and shoes with curled and pointed 
toes.

A pixie stands about 2-1/2 feet tall and weighs about 30 pounds.

Pixies speak Sylvan and Common, and may know other languages as well.

Combat

The normally carefree pixies ferociously attack evil creatures and unwanted intruders. 
They take full advantage of their invisibility and other abilities to harass and 
drive away opponents.

\textbf{Greater Invisibility (Su):} A pixie remains invisible even when it attacks. 
This ability is constant, but the pixie can suppress or resume it as a free action.

\textbf{Spell-Like Abilities:} 1/day---\textit{lesser confusion }(DC 14), \textit{dancing 
lights, detect chaos, detect good, detect evil, detect law, detect thoughts }(DC 
15), \textit{dispel magic, entangle }(DC 14), \textit{permanent image }(DC 19; 
visual and auditory elements only), \textit{polymorph }(self only). Caster level 
8th. The save DCs are Charisma-based.

One pixie in ten can use \textit{irresistible dance }(caster level 8th) once per 
day.

\textbf{Special Arrows (Ex): }Pixies sometimes employ arrows that deal no damage 
but can erase memory or put a creature to sleep. 

\textit{Memory Loss: }An opponent struck by this arrow must succeed on a DC 15 
Will save or lose all memory. The save DC is Charisma-based and includes a +2 racial 
bonus. The subject retains skills, languages, and class abilities but forgets everything 
else until he or she receives a \textit{heal }spell or memory restoration with 
\textit{limited wish, wish, }or \textit{miracle}.

\textit{Sleep: }Any opponent struck by this arrow, regardless of Hit Dice, must 
succeed on a DC 15 Fortitude save or be affected as though by a \textit{sleep }spell. 
The save DC is Charisma-based and includes a +2 racial bonus.

Pixies as Characters

A pixie character exchanges its 1 HD of fey for its first class level.

Pixie characters possess the following racial traits.--- -

\parindent=3pt
4 Strength, +8 Dexterity, +6 Intelligence, +4 Wisdom, +6 Charisma.---

\parindent=0pt
Small size. +1 bonus to Armor Class, +1 bonus on attack rolls, +4 bonus on Hide 
checks, -4 penalty on grapple checks, lifting and carrying limits 3/4 those of 
Medium characters.---

A pixie's base land speed is 20 feet. It also has a fly speed of 60 feet (good).---

Low-light vision.---

Skills: Pixies have a +2 racial bonus on Listen, Search, and Spot checks.---

Racial Feats: A pixie receives Dodge as a bonus feat.--- 

\parindent=3pt
+1 natural armor bonus.---

\parindent=0pt
Special Attacks (see above): Spell-like abilities.---

Special Qualities (see above): Damage reduction 10/cold iron, greater invisibility, 
spell resistance equal to 15 + class levels.---

Automatic Languages: Common, Sylvan. Bonus Languages: Elven, Gnome, Halfling.---

Favored Class: Sorcerer.---

Level adjustment +4 (+6 if the pixie can use \textit{irresistible dance).}

\vspace{12pt}
{\LARGE{}STIRGE}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{180pt}|}
\hline
  & Tiny Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 1d10 (5 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +4\tabularnewline
\hline
S\textbf{peed:} & 10 ft (2 squares), fly 40 ft. (average)\tabularnewline
\hline
A\textbf{rmor Class:} & 16 (+2 size, +4 Dex), touch 16, flat-footed 12\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +1/-11 (+1 when attached)\tabularnewline
\hline
A\textbf{ttack:} & Touch +7 melee (attach)\tabularnewline
\hline
F\textbf{ull Attack:} & Touch +7 melee (attach)\tabularnewline
\hline
S\textbf{pace/Reach:} & 2-1/2 ft./0 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Attach, blood drain\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., low-light vision\tabularnewline
\hline
S\textbf{aves:} & Fort +2, Ref +6, Will +1\tabularnewline
\hline
A\textbf{bilities:} & Str 3, Dex 19, Con 10, Int 1, Wis 12, Cha 6\tabularnewline
\hline
S\textbf{kills:} & Hide +14, Listen +4, Spot +4\tabularnewline
\hline
F\textbf{eats:} & Alertness, Weapon Finesse \textsuperscript{B}\tabularnewline
\hline
E\textbf{nvironment:} & Warm marshes\tabularnewline
\hline
O\textbf{rganization:} & Colony (2-4), flock (5-8), or storm (9-14)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 1/2\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:}--- & \tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A stirge's coloration ranges from rust-red to reddish-brown, with a dirty yellow 
underside. The proboscis is pink at the tip, fading to gray at its base.

A stirge's body is about 1 foot long, with a wingspan of about 2 feet. It weighs 
about 1 pound.

COMBAT

A stirge attacks by landing on a victim, finding a vulnerable spot, and plunging 
its proboscis into the flesh. This is a touch attack and can target only Small 
or larger creatures.

\textbf{Attach (Ex):} If a stirge hits with a touch attack, it uses its eight pincers 
to latch onto the opponent's body. An attached stirge is effectively grappling 
its prey. The stirge loses its Dexterity bonus to AC and has an AC of 12, but holds 
on with great tenacity. Stirges have a +12 racial bonus on grapple checks (already 
figured into the Base Attack/Grapple entry above).

An attached stirge can be struck with a weapon or grappled itself. To remove an 
attached stirge through grappling, the opponent must achieve a pin against the 
stirge.

\textbf{Blood Drain (Ex):} A stirge drains blood, dealing 1d4 points of Constitution 
damage in any round when it begins its turn attached to a victim. Once it has dealt 
4 points of Constitution damage, it detaches and flies off to digest the meal. 
If its victim dies before the stirge's appetite has been sated, the stirge detaches 
and seeks a new target.

\vspace{12pt}
{\LARGE{}SWARM}

Swarms are dense masses of Fine, Diminutive, or Tiny creatures that would not be 
particularly dangerous in small groups, but can be terrible foes when gathered 
in sufficient numbers. For game purposes a swarm is defined as a single creature 
with a space of 10 feet---gigantic hordes are actually composed of dozens of swarms 
in close proximity. A swarm has a single pool of Hit Dice and hit points, a single 
initiative modifier, a single speed, and a single Armor Class. It makes saving 
throws as a single creature.

Many different creatures can mass as swarms; bat swarms, centipede swarms, hellwasp 
swarms, locust swarms, rat swarms, and spider swarms are described here. The swarm's 
type varies with the nature of the component creature (most are animals or vermin), 
but all swarms have the swarm subtype.

A swarm of Tiny creatures consists of 300 nonflying creatures or 1,000 flying creatures. 
A swarm of Diminutive creatures consists of 1,500 nonflying creatures or 5,000 
flying creatures. A swarm of Fine creatures consists of 10,000 creatures, whether 
they are flying or not. Swarms of nonflying creatures include many more creatures 
than could normally fit in a 10-foot square based on their normal space, because 
creatures in a swarm are packed tightly together and generally crawl over each 
other and their prey when moving or attacking. Larger swarms are represented by 
multiples of single swarms. A large swarm is completely shapeable, though it usually 
remains contiguous.

COMBAT

In order to attack, a single swarm moves into opponents' spaces, which provokes 
an attack of opportunity. It can occupy the same space as a creature of any size, 
since it crawls all over its prey, but remains a creature with a 10-foot space. 
Swarms never make attacks of opportunity, but they can provoke attacks of opportunity.

Unlike other creatures with a 10-foot space, a swarm is shapeable. It can occupy 
any four contiguous squares, and it can squeeze through any space large enough 
to contain one of its component creatures.

Vulnerabilities of Swarms

Swarms are extremely difficult to fight with physical attacks. However, they have 
a few special vulnerabilities,

as follows:

A lit torch swung as an improvised weapon deals 1d3 points of fire damage per hit.

A weapon with a special ability such as flaming or frost deals its full energy 
damage with each hit, even if the weapon's normal damage can't affect the swarm.

A lit lantern can be used as a thrown weapon, dealing 1d4 points of fire damage 
to all creatures in squares adjacent to where it breaks.

\vspace{12pt}
BAT SWARM

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{213pt}|}
\hline
  & Diminutive Animal (Swarm)\tabularnewline
\hline
H\textbf{it Dice:} & 3d8 (13 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +2\tabularnewline
\hline
S\textbf{peed:} & 5 ft. (1 square), fly 40 ft. (good)\tabularnewline
\hline
A\textbf{rmor Class:} & 16 (+4 size, +2 Dex), touch 14, flat-footed 12\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +2/---\tabularnewline
\hline
A\textbf{ttack:} & Swarm (1d6)\tabularnewline
\hline
F\textbf{ull Attack:} & Swarm (1d6)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./0 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Distraction, wounding\tabularnewline
\hline
S\textbf{pecial Qualities:} & Blindsense 20 ft., half damage from slashing and 
piercing, low-light vision, swarm traits\tabularnewline
\hline
S\textbf{aves:} & Fort +3, Ref +7, Will +3\tabularnewline
\hline
A\textbf{bilities:} & Str 3, Dex 15, Con 10, Int 2, Wis 14, Cha 4\tabularnewline
\hline
S\textbf{kills:} & Listen +11, Spot +11\tabularnewline
\hline
F\textbf{eats:} & Alertness, Lightning Reflexes\tabularnewline
\hline
E\textbf{nvironment:} & Temperate deserts\tabularnewline
\hline
O\textbf{rganization:} & Solitary, flight (2-4 swarms), or colony (11-20 swarms)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & None\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A bat swarm is nocturnal, and is never found aboveground in daylight.

Combat

A bat swarm seeks to surround and attack any warm-blooded prey it encounters. The 
swarm deals 1d6 points of damage to any creature whose space it occupies at the 
end of its move.

\textbf{Distraction (Ex): }Any living creature that begins its turn with a swarm 
in its space must succeed on a DC 11 Fortitude save or be nauseated for 1 round. 
The save DC is Constitution-based.

\textbf{Wounding (Ex):} Any living creature damaged by a bat swarm continues to 
bleed, losing 1 hit point per round thereafter. Multiple wounds do not result in 
cumulative bleeding loss. The bleeding can be stopped by a DC 10 Heal check or 
the application of a \textit{cure }spell or some other healing magic.

\textbf{Blindsense (Ex):} A bat swarm notices and locates creatures within 20 feet. 
Opponents still have total concealment against the bat swarm (but swarm attacks 
ignore concealment). 

\textbf{Skills:} A bat swarm has a +4 racial bonus on Listen and Spot checks. These 
bonuses are lost if its blindsense is negated. 

\vspace{12pt}
CENTIPEDE SWARM

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{204pt}|}
\hline
  & Diminutive Vermin (Swarm)\tabularnewline
\hline
H\textbf{it Dice:} & 9d8-9 (31 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +4\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), climb 20 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 18 (+4 size, +4 Dex), touch 18, flat-footed 14\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +6/---\tabularnewline
\hline
A\textbf{ttack:} & Swarm (2d6 plus poison)\tabularnewline
\hline
F\textbf{ull Attack:} & Swarm (2d6 plus poison)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./0 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Distraction, poison\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., immune to weapon damage, swarm 
traits, tremorsense 30 ft., vermin traits\tabularnewline
\hline
S\textbf{aves:} & Fort +5, Ref +7, Will +3\tabularnewline
\hline
A\textbf{bilities:} & Str 1, Dex 19, Con 8, Int ---, Wis 10, Cha 2\tabularnewline
\hline
S\textbf{kills:} & Climb +12, Spot +4\tabularnewline
\hline
F\textbf{eats:} & Weapon Finesse\textsuperscript{B}\tabularnewline
\hline
E\textbf{nvironment:} & Underground\tabularnewline
\hline
O\textbf{rganization:} & Solitary, tangle (2-4 swarms), or colony (7-12 swarms)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 4\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & None\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

Combat

A centipede swarm seeks to surround and attack any living prey it encounters. A 
swarm deals 2d6 points of damage to any creature whose space it occupies at the 
end of its move.

\textbf{Distraction (Ex): }Any living creature that begins its turn with a centipede 
swarm in its space must succeed on a DC 13 Fortitude save or be nauseated for 1 
round. The save DC is Constitution-based.

\textbf{Poison (Ex):} Injury, Fortitude DC 13, initial and secondary damage 1d4 
Dex. The save DC is Constitution-based.

\textbf{Skills:} A centipede swarm has a +4 racial bonus on Spot checks and a +8 
racial bonus on Climb checks, and uses its Dexterity modifier instead of its Strength 
modifier for Climb checks. A centipede swarm has a +8 racial bonus on Climb checks 
and can always choose to take 10 on a Climb check, even if rushed or threatened.

\vspace{12pt}
HELLWASP SWARM

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{229pt}|}
\hline
  & Diminutive Magical Beast (Extraplanar, Evil, Swarm)\tabularnewline
\hline
H\textbf{it Dice:} & 12d10+27 (93 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +10\tabularnewline
\hline
S\textbf{peed:} & 5 ft. (1 square), fly 40 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 20 (+4 size, +6 Dex), touch 20, flat-footed 14\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +12/---\tabularnewline
\hline
A\textbf{ttack:} & Swarm (3d6 plus poison)\tabularnewline
\hline
F\textbf{ull Attack:} & Swarm (3d6 plus poison)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./0 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Distraction, inhabit, poison\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 10/magic, darkvision 60 ft., hive 
mind, immune to weapon damage, resistance to fire 10, low-light vision, swarm traits\tabularnewline
\hline
S\textbf{aves:} & Fort +10, Ref +14, Will +7\tabularnewline
\hline
A\textbf{bilities:} & Str 1, Dex 22, Con 14, Int 6, Wis 13, Cha 9\tabularnewline
\hline
S\textbf{kills:} & Hide +19, Listen +10, Spot +10\tabularnewline
\hline
F\textbf{eats:} & Ability Focus (poison), Alertness, Improved Initiative, Iron 
Will, Toughness\tabularnewline
\hline
E\textbf{nvironment:} & A evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary, fright (2-4 swarms), or terror (5-8 swarms)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 8\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & None\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A single hellwasp resembles a thumb-sized normal wasp, except its carapace is gleaming 
black with ruby-red stripes, and its compound eyes are an iridescent green. In 
swarms, hellwasps form a collective hive mind intelligence with infernal cunning 
and bloodlust.

Combat

Like any swarm, a hellwasp swarm seeks to surround and attack any living prey it 
encounters. A swarm deals 3d6 points of damage to any creature whose space it occupies 
at the end of its move. The swarm can take over the bodies of its prey and infest 
both the living and the dead, using them as horrible living (or unliving) puppets 
to accomplish acts of wickedness that a swarm of insects could never attempt.

A hellwasp swarm's attack is treated as an evil-aligned weapon and a magic weapon 
for the purpose of overcoming damage reduction.

\textbf{Distraction (Ex): }Any living creature that begins its turn with a hellwasp 
swarm in its space must succeed on a DC 18 Fortitude save or be nauseated for 1 
round. The save DC is Constitution-based.

\textbf{Inhabit (Ex): }A hellwasp swarm can enter the body of a helpless or dead 
creature by crawling into its mouth and other orifices. Inhabiting requires 1 minute, 
and the victim must be Small, Medium, or Large (although four swarms working together 
can  inhabit a Huge creature). The swarm can abandon the body at any time, although 
doing this takes 1 full round. Any attack against the host deals half damage to 
the hellwasp swarm as well, although the swarm's resistances and immunities may 
negate some or all of this damage.

If a hellwasp swarm inhabits a dead body, it can restore animation to the creature 
and control its movements, effectively transforming it into a zombie of the appropriate 
size for as long as the swarm remains inside. If a hellwasp swarm inhabits a living 
victim, it can neutralize the effects of its own poison and control the victim's 
movement and actions as if using \textit{dominate monster }on the victim. The hellwasps 
quickly consume a living victim, dealing 2d4 points of Constitution damage per 
hour they inhabit a body. A body reduced to Constitution 0 is dead.

A hellwasp-inhabited creature is relatively easy to spot, since its skin crawls 
with the forms of insects inside. The swarm is intelligent enough to attempt to 
hide beneath loose clothing or a large cloak to keep its presence from being detected. 
The swarm can attempt a Disguise check to conceal its inhabitation of a host, with 
a -4 penalty if currently inhabiting a Small host.  

A \textit{remove disease }or \textit{heal }spell cast on an inhabited victim forces 
the hellwasp swarm to abandon its host.

\textbf{Poison (Ex):} Injury, Fortitude DC 18, initial and secondary damage 1d6 
Dex. The save DC is Constitution-based.

\textbf{Hive Mind (Ex): }Any hellwasp swarm with at least 1 hit point per Hit Die 
(or 12 hit points, for a standard hellwasp swarm) forms a hive mind, giving it 
an Intelligence of 6. When a hellwasp swarm is reduced below this hit point threshold, 
it becomes mindless. 

\vspace{12pt}
LOCUST SWARM

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{213pt}|}
\hline
  & Diminutive Vermin (Swarm)\tabularnewline
\hline
H\textbf{it Dice:} & 6d8-6 (21 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +4\tabularnewline
\hline
S\textbf{peed:} & 10 ft. (2 squares), fly 30 ft. (poor)\tabularnewline
\hline
A\textbf{rmor Class:} & 18 (+4 size, +4 Dex), touch 18, flat-footed 14\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +4/---\tabularnewline
\hline
A\textbf{ttack:} & Swarm (2d6)\tabularnewline
\hline
F\textbf{ull Attack:} & Swarm (2d6)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./0 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Distraction\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., immune to weapon damage, swarm 
traits, vermin traits\tabularnewline
\hline
S\textbf{aves:} & Fort +4, Ref +6, Will +2\tabularnewline
\hline
A\textbf{bilities:} & Str 1, Dex 19, Con 8, Int ---, Wis 10, Cha 2\tabularnewline
\hline
S\textbf{kills:} & Listen +4, Spot +4\tabularnewline
\hline
F\textbf{eats:}--- & \tabularnewline
\hline
E\textbf{nvironment:} & Temperate plains\tabularnewline
\hline
O\textbf{rganization:} & Solitary, cloud (2-7 swarms), or plague (11-20 swarms)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 3\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & None\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A locust swarm is a cloud of thousands of winged vermin that devours any organic 
material in its path.

Combat

A locust swarm surrounds and attacks any living prey it encounters. A swarm deals 
2d6 points of damage to any creature whose space it occupies at the end of its 
move.

\textbf{Distraction (Ex): }Any living creature that begins its turn with a locust 
swarm in its space must succeed on a DC 12 Fortitude save or be nauseated for 1 
round. The save DC is Constitution-based.

\textbf{Skills:} A locust swarm has a +4 racial bonus on Listen and Spot checks.

\vspace{12pt}
RAT SWARM

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{204pt}|}
\hline
  & Tiny Animal (Swarm)\tabularnewline
\hline
H\textbf{it Dice:} & 4d8 (13 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +2\tabularnewline
\hline
S\textbf{peed:} & 15 ft. (3 squares), climb 15 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 14 (+2 size, +2 Dex), touch 14, flat-footed 12\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +3/---\tabularnewline
\hline
A\textbf{ttack:} & Swarm (1d6 plus disease)\tabularnewline
\hline
F\textbf{ull Attack:} & Swarm (1d6 plus disease)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./0 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Disease, distraction\tabularnewline
\hline
S\textbf{pecial Qualities:} & Half damage from slashing and piercing, low-light 
vision, scent, swarm traits\tabularnewline
\hline
S\textbf{aves:} & Fort +4, Ref +6, Will +2\tabularnewline
\hline
A\textbf{bilities:} & Str 2, Dex 15, Con 10, Int 2, Wis 12, Cha 2\tabularnewline
\hline
S\textbf{kills:} & Balance +10, Climb +10, Hide +14, Listen +6, Spot +7, Swim +10\tabularnewline
\hline
F\textbf{eats:} & Alertness, Weapon Finesse\tabularnewline
\hline
E\textbf{nvironment:} & Any\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pack (2-4 swarms), or infestation (7-12 swarms)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & None\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

Combat

A rat swarm seeks to surround and attack any warm-blooded prey it encounters. A 
swarm deals 1d6 points of damage to any creature whose space it occupies at the 
end of its move.

\textbf{Disease (Ex):} Filth fever---swarm attack, Fortitude DC 12, incubation 
period 1d3 days, damage 1d3 Dex and 1d3 Con. The save DC is Constitution-based.

\textbf{Distraction (Ex):} Any living creature that begins its turn with a swarm 
in its square must succeed on a DC 12 Fortitude save or be nauseated for 1 round. 
The save DC is Constitution-based.

\textbf{Skills: }A rat swarm has a +4 racial bonus on Hide and Move Silently checks, 
and a +8 racial bonus on Balance, Climb, and Swim checks. A rat swarm can always 
choose to take 10 on all Climb checks, even if rushed or threatened. A rat swarm 
uses its Dexterity modifier instead of its Strength modifier for Climb and Swim 
checks. A rat swarm has a +8 racial bonus on any Swim check to perform some special 
action or avoid a hazard. It can always choose to take 10 on a Swim check, even 
if distracted or endangered. It can use the run action while swimming, provided 
it swims in a straight line.

\vspace{12pt}
SPIDER SWARM

\begin{tabular}{|>{\raggedright}p{86pt}|>{\raggedright}p{239pt}|}
\hline
  & Diminutive Vermin (Swarm)\tabularnewline
\hline
H\textbf{it Dice:} & 2d8 (9 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +3\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), climb 20 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 17 (+4 size, +3 Dex), touch 17, flat-footed 14\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +1/---\tabularnewline
\hline
A\textbf{ttack:} & Swarm (1d6 plus poison)\tabularnewline
\hline
F\textbf{ull Attack:} & Swarm (1d6 plus poison)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./0 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Distraction, poison\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., swarm traits, tremorsense 30 ft., 
vermin traits\tabularnewline
\hline
S\textbf{aves:} & Fort +3, Ref +3, Will +0\tabularnewline
\hline
A\textbf{bilities:} & Str 1, Dex 17, Con 10, Int ---, Wis 10, Cha 2\tabularnewline
\hline
S\textbf{kills:} & Climb +11, Listen +4, Spot +4\tabularnewline
\hline
E\textbf{nvironment:} & Warm forests\tabularnewline
\hline
O\textbf{rganization:} & Solitary, tangle (2-4 swarms), or colony (7-12 swarms)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 1\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & None\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

Combat

A spider swarm seeks to surround and attack any living prey it encounters. A swarm 
deals 1d6 points of damage to any creature whose space it occupies at the end of 
its move.

\textbf{Distraction (Ex):} Any living creature that begins its turn with a spider 
swarm in its space must succeed on a DC 11 Fortitude save or be nauseated for 1 
round. The save DC is Constitution-based.

\textbf{Poison (Ex): }Injury, Fortitude DC 11, initial and secondary damage 1d3 
Str. The save DC is Constitution-based.

\textbf{Skills:} A spider swarm has a +4 racial bonus on Hide and Spot checks and 
a +8 racial bonus on Climb checks. It uses its Dexterity modifier instead of its 
Strength modifier for Climb checks. It can always choose to take 10 on a Climb 
check, even if rushed or threatened.

\newpage

\end{document}
