%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{array}
\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

This material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

{\LARGE{}SPELLS (P-R)}

\vspace{12pt}
Passwall

Transmutation

\textbf{Level:} Sor/Wiz 5

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Effect:} 5 ft. by 8 ft. opening, 10 ft. deep plus 5 ft. deep per three 
additional levels

\textbf{Duration:} 1 hour/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You create a passage through wooden, plaster, or stone walls, but not through metal 
or other harder materials. The passage is 10 feet deep plus an additional 5 feet 
deep per three caster levels above 9th (15 feet at 12th, 20 feet at 15th, and a 
maximum of 25 feet deep at 18th level). If the wall's thickness is more than the 
depth of the passage created, then a single \textit{passwall }simply makes a niche 
or short tunnel. Several \textit{passwall }spells can then form a continuing passage 
to breach very thick walls. When \textit{passwall }ends, creatures within the passage 
are ejected out the nearest exit. If someone dispels the \textit{passwall }or you 
dismiss it, creatures in the passage are ejected out the far exit, if there is 
one, or out the sole exit if there is only one.

\textit{Material Component: }A pinch of sesame seeds.

\vspace{12pt}
Pass without Trace

Transmutation

\textbf{Level:} Drd 1, Rgr 1

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Targets:} One creature/level touched

\textbf{Duration:} 1 hour/level (D)

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

The subject or subjects can move through any type of terrain and leave neither 
footprints nor scent. Tracking the subjects is impossible by nonmagical means.

\vspace{12pt}
Permanency

Universal

\textbf{Level:} Sor/Wiz 5

\textbf{Components:} V, S, XP

\textbf{Casting Time:} 2 rounds

\textbf{Range:} See text

\textbf{Target, Effect, or Area: }See text

\textbf{Duration:} Permanent; see text

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell makes certain other spells permanent.

Depending on the spell, you must be of a minimum caster level and must expend a 
number of XP.

You can make the following spells permanent in regard to yourself.

\begin{tabular}{|>{\raggedright}p{98pt}|>{\raggedright}p{53pt}|>{\raggedright}p{41pt}|}
\hline
S\textbf{pell} & M\textbf{inimum Caster Level} & \subsubsection*{X\textbf{P Cost}}\tabularnewline
\hline
A\textit{rcane sight} & 11th & 1,500 XP\tabularnewline
\hline
C\textit{omprehend languages} & 9th & 500 XP\tabularnewline
\hline
D\textit{arkvision} & 10th & 1,000 XP\tabularnewline
\hline
D\textit{etect magic} & 9th & 500 XP\tabularnewline
\hline
R\textit{ead magic} & 9th & 500 XP\tabularnewline
\hline
S\textit{ee invisibility} & 10th & 1,000 XP\tabularnewline
\hline
T\textit{ongues} & 11th & 1,500 XP\tabularnewline
\hline
\end{tabular}

You cast the desired spell and then follow it with the \textit{permanency }spell. 
You cannot cast these spells on other creatures. This application of \textit{permanency 
}can be dispelled only by a caster of higher level than you were when you cast 
the spell.

In addition to personal use, \textit{permanency }can be used to make the following 
spells permanent on yourself, another creature, or an object (as appropriate).

\begin{tabular}{|>{\raggedright}p{83pt}|>{\raggedright}p{59pt}|>{\raggedright}p{56pt}|}
\hline
S\textbf{pell} & M\textbf{inimum Caster Level} & \subsubsection*{X\textbf{P Cost}}\tabularnewline
\hline
E\textit{nlarge person} & 9th & 500 XP\tabularnewline
\hline
M\textit{agic fang} & 9th & 500 XP\tabularnewline
\hline
M\textit{agic fang, greater} & 11th & 1,500 XP\tabularnewline
\hline
R\textit{educe person} & 9th & 500 XP\tabularnewline
\hline
R\textit{esistance} & 9th & 500 XP\tabularnewline
\hline
T\textit{elepathic bond}\textsuperscript{\textbf{1}} & 13th & 2,500 XP\tabularnewline
\hline
\multicolumn{3}{|p{199pt}|}{1 Only bonds two creatures per casting of \textit{permanency}.}\tabularnewline
\hline
\end{tabular}

Additionally, the following spells can be cast upon objects or areas only and rendered 
permanent.

\begin{tabular}{|>{\raggedright}p{99pt}|>{\raggedright}p{61pt}|>{\raggedright}p{41pt}|}
\hline
S\textbf{pell} & \section*{M\textbf{inimum Caster Level}} & \subsubsection*{X\textbf{P 
Cost}}\tabularnewline
\hline
A\textit{larm} & 9th & 500 XP\tabularnewline
\hline
A\textit{nimate objects} & 14th & 3,000 XP\tabularnewline
\hline
D\textit{ancing lights} & 9th & 500 XP\tabularnewline
\hline
G\textit{host sound} & 9th & 500 XP\tabularnewline
\hline
G\textit{ust of wind} & 11th & 1,500 XP\tabularnewline
\hline
I\textit{nvisibility} & 10th & 1,000 XP\tabularnewline
\hline
M\textit{age's private sanctum} & 13th & 2,500 XP\tabularnewline
\hline
M\textit{agic mouth} & 10th & 1,000 XP\tabularnewline
\hline
P\textit{hase door} & 15th & 3,500 XP\tabularnewline
\hline
P\textit{rismatic sphere} & 17th & 4,500 XP\tabularnewline
\hline
P\textit{rismatic wall} & 16th & 4,000 XP\tabularnewline
\hline
S\textit{hrink item} & 11th & 1,500 XP\tabularnewline
\hline
S\textit{olid fog} & 12th & 2,000 XP\tabularnewline
\hline
S\textit{tinking cloud} & 11th & 1,500 XP\tabularnewline
\hline
S\textit{ymbol of death} & 16th & 4,000 XP\tabularnewline
\hline
S\textit{ymbol of fear} & 14th & 3,000 XP\tabularnewline
\hline
S\textit{ymbol of insanity} & 16th & 4,000 XP\tabularnewline
\hline
S\textit{ymbol of pain} & 13th & 2,500 XP\tabularnewline
\hline
S\textit{ymbol of persuasion} & 14th & 3,000 XP\tabularnewline
\hline
S\textit{ymbol of sleep} & 16th & 4,000 XP\tabularnewline
\hline
S\textit{ymbol of stunning} & 15th & 3,500 XP\tabularnewline
\hline
S\textit{ymbol of weakness} & 15th & 3,500 XP\tabularnewline
\hline
T\textit{eleportation circle} & 17th & 4,500 XP\tabularnewline
\hline
W\textit{all of fire} & 12th & 2,000 XP\tabularnewline
\hline
W\textit{all of force} & 13th & 2,500 XP\tabularnewline
\hline
W\textit{eb} & 10th & 1,000 XP\tabularnewline
\hline
\end{tabular}

Spells cast on other creatures, objects, or locations (not on you) are vulnerable 
to \textit{dispel magic }as normal.

\textit{XP Cost: }See tables above.

\vspace{12pt}
Permanent Image

Illusion (Figment)

\textbf{Level:} Brd 6, Sor/Wiz 6

\textbf{Effect:} Figment that cannot extend beyond a 20-ft. cube + one 10-ft. cube/level 
(S)

\textbf{Duration:} Permanent (D)

This spell functions like \textit{silent image, }except that the figment includes 
visual, auditory, olfactory, and thermal elements, and the spell is permanent. 
By concentrating, you can move the image within the limits of the range, but it 
is static while you are not concentrating.

\textit{Material Component: }A bit of fleece plus powdered jade worth 100 gp.

\vspace{12pt}
Persistent Image

Illusion (Figment)

\textbf{Level:} Brd 5, Sor/Wiz 5

\textbf{Duration:} 1 min./level (D)

This spell functions like \textit{silent image, }except that the figment includes 
visual, auditory, olfactory, and thermal components, and the figment follows a 
script determined by you. The figment follows that script without your having to 
concentrate on it. The illusion can include intelligible speech if you wish. 

\textit{Material Component: }A bit of fleece and several grains of sand.

\vspace{12pt}
Phantasmal Killer

Illusion (Phantasm) [Fear, Mind-Affecting]

\textbf{Level:} Sor/Wiz 4

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Target:} One living creature

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will disbelief (if interacted with), then Fortitude partial; 
see text

\textbf{Spell Resistance:} Yes

You create a phantasmal image of the most fearsome creature imaginable to the subject 
simply by forming the fears of the subject's subconscious mind into something that 
its conscious mind can visualize: this most horrible beast. Only the spell's subject 
can see the phantasmal killer. You see only a vague shape. The target first gets 
a Will save to recognize the image as unreal. If that save fails, the phantasm 
touches the subject, and the subject must succeed on a Fortitude save or die from 
fear. Even if the Fortitude save is successful, the subject takes 3d6 points of 
damage.

If the subject of a \textit{phantasmal killer }attack succeeds in disbelieving 
and is wearing a \textit{helm of telepathy, }the beast can be turned upon you. 
You must then disbelieve it or become subject to its deadly fear attack.

\vspace{12pt}
Phantom Steed

Conjuration (Creation)

\textbf{Level:} Brd 3, Sor/Wiz 3

\textbf{Components:} V, S

\textbf{Casting Time:} 10 minutes

\textbf{Range:} 0 ft.

\textbf{Effect:} One quasi-real, horselike creature

\textbf{Duration:} 1 hour/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You conjure a Large, quasi-real, horselike creature. The steed can be ridden only 
by you or by the one person for whom you specifically created the mount. A phantom 
steed has a black head and body, gray mane and tail, and smoke-colored, insubstantial 
hooves that make no sound. It has what seems to be a saddle, bit, and bridle. It 
does not fight, but animals shun it and refuse to attack it.

The mount has an AC of 18 (-1 size, +4 natural armor, +5 Dex) and 7 hit points 
+1 hit point per caster level. If it loses all its hit points, the phantom steed 
disappears. A phantom steed has a speed of 20 feet per caster level, to a maximum 
of 240 feet. It can bear its rider's weight plus up to 10 pounds per caster level.

These mounts gain certain powers according to caster level. A mount's abilities 
include those of mounts of lower caster levels. 

\textit{8th Level: }The mount can ride over sandy, muddy, or even swampy ground 
without difficulty or decrease in speed.

\textit{10th Level: }The mount can use \textit{water walk }at will (as the spell, 
no action required to activate this ability).

\textit{12th Level: }The mount can use \textit{air walk }at will (as the spell, 
no action required to activate this ability) for up to 1 round at a time, after 
which it falls to the ground.

\textit{14th Level: }The mount can fly at its speed (average maneuverability).

\vspace{12pt}
Phantom Trap

Illusion (Glamer)

\textbf{Level:} Sor/Wiz 2

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Object touched

\textbf{Duration:} Permanent (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell makes a lock or other small mechanism seem to be trapped to anyone who 
can detect traps. You place the spell upon any small mechanism or device, such 
as a lock, hinge, hasp, cork, cap, or ratchet. Any character able to detect traps, 
or who uses any spell or device enabling trap detection, is 100\% certain a real 
trap exists. Of course, the effect is illusory and nothing happens if the trap 
is ``sprung''; its primary purpose is to frighten away thieves or make them waste 
precious time.

If another \textit{phantom trap }is active within 50 feet when the spell is cast, 
the casting fails.

\textit{Material Component: }A piece of iron pyrite touched to the object to be 
trapped while the object is sprinkled with a special dust requiring 50 gp to prepare.

\vspace{12pt}
Phase Door

Conjuration (Creation)

\textbf{Level:} Sor/Wiz 7, Travel 8

\textbf{Components:} V

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 0 ft.

\textbf{Effect:} Ethereal 5 ft. by 8 ft. opening, 10 ft. deep + 5 ft. deep per 
three levels

\textbf{Duration:} One usage per two levels

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell creates an ethereal passage through wooden, plaster, or stone walls, 
but not other materials. The \textit{phase door }is invisible and inaccessible 
to all creatures except you, and only you can use the passage. You disappear when 
you enter the \textit{phase door }and appear when you exit. If you desire, you 
can take one other creature (Medium or smaller) through the door. This counts as 
two uses of the door. The door does not allow light, sound, or spell effects through 
it, nor can you see through it without using it. Thus, the spell can provide an 
escape route, though certain creatures, such as phase spiders, can follow with 
ease. A \textit{gem of true seeing }or similar magic reveals the presence of a 
\textit{phase door }but does not allow its use.

A \textit{phase door }is subject to \textit{dispel magic. }If anyone is within 
the passage when it is dispelled, he is harmlessly ejected just as if he were inside 
a \textit{passwall }effect.

You can allow other creatures to use the \textit{phase door }by setting some triggering 
condition for the door. Such conditions can be as simple or elaborate as you desire. 
They can be based on a creature's name, identity, or alignment, but otherwise must 
be based on observable actions or qualities. Intangibles such as level, class, 
Hit Dice, and hit points don't qualify.

\textit{Phase door }can be made permanent with a \textit{permanency }spell.

\vspace{12pt}
Planar Ally

Conjuration (Calling) [see text for \textit{lesser planar ally}]

\textbf{Level:} Clr 6

\textbf{Effect:} One or two called elementals or outsiders, totaling no more than 
12 HD, which cannot be more than 30 ft. apart when they appear

This spell functions like \textit{lesser planar ally}, except you may call a single 
creature of 12 HD or less, or two creatures of the same kind whose Hit Dice total 
no more than 12. The creatures agree to help you and request your return payment 
together.

\textit{XP Cost: }250 XP.

\vspace{12pt}
Planar Ally, Greater

Conjuration (Calling) [see text for \textit{lesser planar ally}]

\textbf{Level:} Clr 8

\textbf{Effect:} Up to three called elementals or outsiders, totaling no more than 
18 HD, no two of which can be more than 30 ft. apart when they appear.

This spell functions like \textit{lesser planar ally}, except that you may call 
a single creature of 18 HD or less, or up to three creatures of the same kind whose 
Hit Dice total no more than 18. The creatures agree to help you and request your 
return payment together.

\textit{XP Cost: }500 XP.

\vspace{12pt}
Planar Ally, Lesser

Conjuration (Calling) [see text]

\textbf{Level:} Clr 4

\textbf{Components:} V, S, DF, XP

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} One called elemental or outsider of 6 HD or less

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

By casting this spell, you request your deity to send you an elemental or outsider 
(of 6 HD or less) of the deity's choice. If you serve no particular deity, the 
spell is a general plea answered by a creature sharing your philosophical alignment. 
If you know an individual creature's name, you may request that individual by speaking 
the name during the spell (though you might get a different creature anyway).

You may ask the creature to perform one task in exchange for a payment from you. 
Tasks might range from the simple to the complex. You must be able to communicate 
with the creature called in order to bargain for its services.

The creature called requires a payment for its services. This payment can take 
a variety of forms, from donating gold or magic items to an allied temple, to a 
gift given directly to the creature, to some other action on your part that matches 
the creature's alignment and goals. Regardless, this payment must be made before 
the creature agrees to perform any services. The bargaining takes at least 1 round, 
so any actions by the creature begin in the round after it arrives.

A task taking up to 1 minute per caster level requires a payment of 100 gp per 
HD of the creature called. For a task taking up to 1 hour per caster level, the 
creature requires a payment of 500 gp per HD. A long-term task, one requiring up 
to one day per caster level, requires a payment of 1,000 gp per HD.

A nonhazardous task requires only half the indicated payment, while an especially 
hazardous task might require a greater gift. Few if any creatures will accept a 
task that seems suicidal (remember, a called creature actually dies when it is 
killed, unlike a summoned creature). However, if the task is strongly aligned with 
the creature's ethos, it may halve or even waive the payment. 

At the end of its task, or when the duration bargained for expires, the creature 
returns to its home plane (after reporting back to you, if appropriate and possible).

\textit{Note: }When you use a calling spell that calls an air, chaotic, earth, 
evil, fire, good, lawful, or water creature, it is a spell of that type.

\textit{XP Cost: }100 XP.

\vspace{12pt}
Planar Binding

Conjuration (Calling) [see text for \textit{lesser planar binding}]

\textbf{Level:} Sor/Wiz 6

\textbf{Components:} V, S

\textbf{Targets:} Up to three elementals or outsiders, totaling no more than 12 
HD, no two of which can be more than 30 ft. apart when they appear

This spell functions like \textit{lesser planar binding, }except that you may call 
a single creature of 12 HD or less, or up to three creatures of the same kind whose 
Hit Dice total no more than 12. Each creature gets a save, makes an independent 
attempt to escape, and must be individually persuaded to aid you.

\vspace{12pt}
Planar Binding, Greater

Conjuration (Calling) [see text for \textit{lesser planar binding}]

\textbf{Level:} Sor/Wiz 8

\textbf{Components:} V, S

\textbf{Targets:} Up to three elementals or outsiders, totaling no more than 18 
HD, no two of which can be more than 30 ft. apart when they appear.

This spell functions like \textit{lesser planar binding, }except that you may call 
a single creature of 18 HD or less, or up to three creatures of the same kind whose 
Hit Dice total no more than 18. Each creature gets a saving throw, makes independent 
attempts to escape, and must be persuaded to aid you individually.

\vspace{12pt}
Planar Binding, Lesser

Conjuration (Calling) [see text]

\textbf{Level:} Sor/Wiz 5

\textbf{Components:} V, S

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Close (25 ft. + 5 ft./2 levels); see text

\textbf{Target:} One elemental or outsider with 6 HD or less

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} No and Yes; see text

Casting this spell attempts a dangerous act: to lure a creature from another plane 
to a specifically prepared trap, which must lie within the spell's range. The called 
creature is held in the trap until it agrees to perform one service in return for 
its freedom.

To create the trap, you must use a \textit{magic circle }spell, focused inward. 
The kind of creature to be bound must be known and stated. If you wish to call 
a specific individual, you must use that individual's proper name in casting the 
spell.

The target creature is allowed a Will saving throw. If the saving throw succeeds, 
the creature resists the spell. If the saving throw fails, the creature is immediately 
drawn to the trap (spell resistance does not keep it from being called). The creature 
can escape from the trap with by successfully pitting its spell resistance against 
your caster level check, by dimensional travel, or with a successful Charisma check 
(DC 15 + 1/2 your caster level + your Cha modifier). It can try each method once 
per day. If it breaks loose, it can flee or attack you. A \textit{dimensional anchor 
}cast on the creature prevents its escape via dimensional travel. You can also 
employ a calling diagram (see \textit{magic circle against evil}) to make the trap 
more secure.

If the creature does not break free of the trap, you can keep it bound for as long 
as you dare. You can attempt to compel the creature to perform a service by describing 
the service and perhaps offering some sort of reward. You make a Charisma check 
opposed by the creature's Charisma check. The check is assigned a bonus of +0 to 
+6 based on the nature of the service and the reward. If the creature wins the 
opposed check, it refuses service. New offers, bribes, and the like can be made 
or the old ones reoffered every 24 hours. This process can be repeated until the 
creature promises to serve, until it breaks free, or until you decide to get rid 
of it by means of some other spell. Impossible demands or unreasonable commands 
are never agreed to. If you roll a 1 on the Charisma check, the creature breaks 
free of the binding and can escape or attack you.

Once the requested service is completed, the creature need only so inform you to 
be instantly sent back whence it came. The creature might later seek revenge. If 
you assign some open-ended task that the creature cannot complete though its own 
actions the spell remains in effect for a maximum of one day per caster level, 
and the creature gains an immediate chance to break free. Note that a clever recipient 
can subvert some instructions.

When you use a calling spell to call an air, chaotic, earth, evil, fire, good, 
lawful, or water creature, it is a spell of that type. 

\vspace{12pt}
Plane Shift

Conjuration (Teleportation)

\textbf{Level:} Clr 5, Sor/Wiz 7

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched, or up to eight willing creatures joining hands

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

You move yourself or some other creature to another plane of existence or alternate 
dimension. If several willing persons link hands in a circle, as many as eight 
can be affected by the \textit{plane shift }at the same time. Precise accuracy 
as to a particular arrival location on the intended plane is nigh impossible. From 
the Material Plane, you can reach any other plane, though you appear 5 to 500 miles 
(5d\%) from your intended destination.

\textit{Note: Plane shift }transports creatures instantaneously and then ends. 
The creatures need to find other means if they are to travel back.

\textit{Focus: }A small, forked metal rod. The size and metal type dictates to 
which plane of existence or alternate dimension the spell sends the affected creatures. 

\vspace{12pt}
Plant Growth

Transmutation

\textbf{Level:} Drd 3, Plant 3, Rgr 3

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} See text

\textbf{Target or Area:} See text

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

\textit{Plant growth }has different effects depending on the version chosen.

\textit{Overgrowth: }This effect causes normal vegetation (grasses, briars, bushes, 
creepers, thistles, trees, vines) within long range (400 feet + 40 feet per caster 
level) to become thick and overgrown. The plants entwine to form a thicket or jungle 
that creatures must hack or force a way through. Speed drops to 5 feet, or 10 feet 
for Large or larger creatures. The area must have brush and trees in it for this 
spell to take effect.

At your option, the area can be a 100-foot-radius circle, a 150-foot-radius semicircle, 
or a 200-foot-radius quarter circle.

You may designate places within the area that are not affected.

\textit{Enrichment: }This effect targets plants within a range of one-half mile, 
raising their potential productivity over the course of the next year to one-third 
above normal.

\textit{Plant growth }counters \textit{diminish plants}.

This spell has no effect on plant creatures.

\vspace{12pt}
Poison

Necromancy

\textbf{Level:} Clr 4, Drd 3

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Living creature touched

\textbf{Duration:} Instantaneous; see text

\textbf{Saving Throw:} Fortitude negates; see text

\textbf{Spell Resistance:} Yes

Calling upon the venomous powers of natural predators, you infect the subject with 
a horrible poison by making a successful melee touch attack. The poison deals 1d10 
points of temporary Constitution damage immediately and another 1d10 points of 
temporary Constitution damage 1 minute later. Each instance of damage can be negated 
by a Fortitude save (DC 10 + 1/2 your caster level + your Wis modifier).

\vspace{12pt}
Polar Ray

Evocation [Cold]

\textbf{Level:} Sor/Wiz 8

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} Ray

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

A blue-white ray of freezing air and ice springs from your hand. You must succeed 
on a ranged touch attack with the ray to deal damage to a target. The ray deals 
1d6 points of cold damage per caster level (maximum 25d6).

\textit{Focus: }A small, white ceramic cone or prism.

\vspace{12pt}
Polymorph

Transmutation

\textbf{Level:} Sor/Wiz 4

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target: }Willing living creature touched

\textbf{Duration:} 1 min./level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell functions like \textit{alter self, }except that you change the willing 
subject into another form of living creature. The new form may be of the same type 
as the subject or any of the following types: aberration, animal, dragon, fey, 
giant, humanoid, magical beast, monstrous humanoid, ooze, plant, or vermin. The 
assumed form can't have more Hit Dice than your caster level (or the subject's 
HD, whichever is lower), to a maximum of 15 HD at 15th level. You can't cause a 
subject to assume a form smaller than Fine, nor can you cause a subject to assume 
an incorporeal or gaseous form. The subject's creature type and subtype (if any) 
change to match the new form.

Upon changing, the subject regains lost hit points as if it had rested for a night 
(though this healing does not restore temporary ability damage and provide other 
benefits of resting; and changing back does not heal the subject further). If slain, 
the subject reverts to its original form, though it remains dead.

The subject gains the Strength, Dexterity, and Constitution scores of the new form 
but retains its own Intelligence, Wisdom, and Charisma scores. It also gains all 
extraordinary special attacks possessed by the form but does not gain the extraordinary 
special qualities possessed by the new form or any supernatural or spell-like abilities.

Incorporeal or gaseous creatures are immune to being \textit{polymorphed}, and 
a creature with the shapechanger subtype can revert to its natural form as a standard 
action.

\textit{Material Component: }An empty cocoon.

\vspace{12pt}
Polymorph Any Object

Transmutation

\textbf{Level:} Sor/Wiz 8, Trickery 8

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One creature, or one nonmagical object of up to 100 cu. ft./level

\textbf{Duration:} See text

\textbf{Saving Throw:} Fortitude negates (object); see text

\textbf{Spell Resistance:} Yes (object)

This spell functions like \textit{polymorph, }except that it changes one object 
or creature into another. The duration of the spell depends on how radical a change 
is made from the original state to its enchanted state. The duration is determined 
by using the following guidelines.

\begin{tabular}{|>{\raggedright}p{194pt}|>{\raggedright}p{74pt}|}
\hline
C\textbf{hanged Subject Is:} & I\textbf{ncrease to Duration Factor}\textsuperscript{\textbf{1}}\tabularnewline
\hline
Same kingdom (animal, vegetable, mineral) & +5\tabularnewline
\hline
Same class (mammals, fungi, metals, etc.) & +2\tabularnewline
\hline
Same size & +2\tabularnewline
\hline
Related (twig is to tree, wolf fur is to wolf, etc.) & +2\tabularnewline
\hline
Same or lower Intelligence & +2\tabularnewline
\hline
\multicolumn{2}{|p{269pt}|}{1 Add all that apply. Look up the total on the next 
table.}\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{38pt}|>{\raggedright}p{47pt}|>{\raggedright}p{88pt}|}
\hline
\subsection*{D\textbf{uration Factor}} & \subsection*{D\textbf{uration}} & \subsection*{E\textbf{xample}}\tabularnewline
\hline
0 & 20 minutes & Pebble to human\tabularnewline
\hline
2 & 1 hour & Marionette to human\tabularnewline
\hline
4 & 3 hours & Human to marionette\tabularnewline
\hline
5 & 12 hours & Lizard to manticore\tabularnewline
\hline
6 & 2 days & Sheep to wool coat\tabularnewline
\hline
7 & 1 week & Shrew to manticore\tabularnewline
\hline
9+ & Permanent & Manticore to shrew\tabularnewline
\hline
\end{tabular}

Unlike \textit{polymorph, polymorph any object }does grant the creature the Intelligence 
score of its new form. If the original form didn't have a Wisdom or Charisma score, 
it gains those scores as appropriate for the new form.

Damage taken by the new form can result in the injury or death of the polymorphed 
creature. In general, damage occurs when the new form is changed through physical 
force.

A nonmagical object cannot be made into a magic item with this spell. Magic items 
aren't affected by this spell.

This spell cannot create material of great intrinsic value, such as copper, silver, 
gems, silk, gold, platinum, mithral, or adamantine. It also cannot reproduce the 
special properties of cold iron in order to overcome the damage reduction of certain 
creatures.

This spell can also be used to duplicate the effects of \textit{baleful polymorph, 
polymorph, flesh to stone}, \textit{stone to flesh}, \textit{transmute mud to rock}, 
\textit{transmute metal to wood, }or \textit{transmute rock to mud.}

\textit{Arcane Material Component: }Mercury, gum arabic, and smoke.

\vspace{12pt}
Power Word Blind

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Sor/Wiz 7, War 7

\textbf{Components:} V

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One creature with 200 hp or less

\textbf{Duration:} See text

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

You utter a single word of power that causes one creature of your choice to become 
blinded, whether the creature can hear the word or not. The duration of the spell 
depends on the target's current hit point total. Any creature that currently has 
201 or more hit points is unaffected by \textit{power word blind}.

\begin{tabular}{|>{\raggedright}p{41pt}|>{\raggedright}p{58pt}|}
\hline
\subsection*{H\textbf{it Points}} & \subsection*{D\textbf{uration}}\tabularnewline
\hline
50 or less & Permanent\tabularnewline
\hline
51-100 & 1d4+1 minutes\tabularnewline
\hline
101-200 & 1d4+1 rounds\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
Power Word Kill

Enchantment (Compulsion) [Death, Mind-Affecting]

\textbf{Level:} Sor/Wiz 9, War 9

\textbf{Components:} V

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One living creature with 100 hp or less

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

You utter a single word of power that instantly kills one creature of your choice, 
whether the creature can hear the word or not. Any creature that currently has 
101 or more hit points is unaffected by \textit{power word kill}.

\vspace{12pt}
Power Word Stun

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Sor/Wiz 8, War 8

\textbf{Components:} V

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One creature with 150 hp or less

\textbf{Duration:} See text

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

You utter a single word of power that instantly causes one creature of your choice 
to become stunned, whether the creature can hear the word or not. The duration 
of the spell depends on the target's current hit point total. Any creature that 
currently has 151 or more hit points is unaffected by \textit{power word stun}.

\begin{tabular}{|>{\raggedright}p{41pt}|>{\raggedright}p{43pt}|}
\hline
\subsection*{H\textbf{it Points}} & \subsection*{D\textbf{uration}}\tabularnewline
\hline
50 or less & 4d4 rounds\tabularnewline
\hline
51-100 & 2d4 rounds\tabularnewline
\hline
101-150 & 1d4 rounds\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
Prayer

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Clr 3, Pal 3

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 40 ft.

\textbf{Area:} All allies and foes within a 40-ft.-radius burst centered on you

\textbf{Duration:} 1 round/level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

You bring special favor upon yourself and your allies while bringing disfavor to 
your enemies. You and your each of your allies gain a +1 luck bonus on attack rolls, 
weapon damage rolls, saves, and skill checks, while each of your foes takes a -1 
penalty on such rolls.

\vspace{12pt}
Prestidigitation

Universal

\textbf{Level:} Brd 0, Sor/Wiz 0

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 10 ft.

\textbf{Target, Effect, or Area: }See text

\textbf{Duration:} 1 hour

\textbf{Saving Throw:} See text

\textbf{Spell Resistance:} No

Prestidigitations are minor tricks that novice spellcasters use for practice. Once 
cast, a \textit{prestidigitation }spell enables you to perform simple magical effects 
for 1 hour. The effects are minor and have severe limitations. A prestidigitation 
can slowly lift 1 pound of material. It can color, clean, or soil items in a 1-foot 
cube each round. It can chill, warm, or flavor 1 pound of nonliving material. It 
cannot deal damage or affect the concentration of spellcasters. \textit{Prestidigitation 
}can create small objects, but they look crude and artificial. The materials created 
by a \textit{prestidigitation }spell are extremely fragile, and they cannot be 
used as tools, weapons, or spell components. Finally, a \textit{prestidigitation 
}lacks the power to duplicate any other spell effects. Any actual change to an 
object (beyond just moving, cleaning, or soiling it) persists only 1 hour.

\vspace{12pt}
Prismatic Sphere

Abjuration

\textbf{Level:} Protection 9, Sor/Wiz 9, Sun 9

\textbf{Components:} V

\textbf{Range:} 10 ft.

\textbf{Effect:} 10-ft.-radius sphere centered on you

This spell functions like \textit{prismatic wall, }except you conjure up an immobile, 
opaque globe of shimmering, multicolored light that surrounds you and protects 
you from all forms of attack. The sphere flashes in all colors of the visible spectrum. 

The sphere's \textit{blindness }effect on creatures with less than 8 HD lasts 2d4x10 
minutes.

You can pass into and out of the \textit{prismatic sphere }and remain near it without 
harm. However, when you're inside it, the sphere blocks any attempt to project 
something through the sphere (including spells). Other creatures that attempt to 
attack you or pass through suffer the effects of each color, one at a time.

Typically, only the upper hemisphere of the globe will exist, since you are at 
the center of the sphere, so the lower half is usually excluded by the floor surface 
you are standing on.

The colors of the sphere have the same effects as the colors of a \textit{prismatic 
wall.}

\textit{Prismatic sphere }can be made permanent with a \textit{permanency }spell.

\vspace{12pt}
Prismatic Spray

Evocation

\textbf{Level:} Sor/Wiz 7

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 60 ft.

\textbf{Area:} Cone-shaped burst

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} See text

\textbf{Spell Resistance:} Yes

This spell causes seven shimmering, intertwined, multicolored beams of light to 
spray from your hand. Each beam has a different power. Creatures in the area of 
the spell with 8 HD or less are automatically blinded for 2d4 rounds. Every creature 
in the area is randomly struck by one or more beams, which have additional effects.

\begin{tabular}{|>{\raggedright}p{13pt}|>{\raggedright}p{53pt}|>{\raggedright}p{248pt}|}
\hline
1\textbf{d8} & C\textbf{olor of Beam} & E\textbf{ffect}\tabularnewline
\hline
1 & Red & 20 points fire damage (Reflex half)\tabularnewline
\hline
2 & Orange & 40 points acid damage (Reflex half)\tabularnewline
\hline
3 & Yellow & 80 points electricity damage (Reflex half)\tabularnewline
\hline
4 & Green & Poison (Kills; Fortitude partial, take 1d6 points of Con damage instead)\tabularnewline
\hline
5 & Blue & Turned to stone (Fortitude negates)\tabularnewline
\hline
6 & Indigo & Insane, as \textit{insanity }spell (Will negates)\tabularnewline
\hline
7 & Violet & Sent to another plane (Will negates)\tabularnewline
\hline
8 & \multicolumn{2}{p{301pt}|}{Struck by two rays; roll twice more, ignoring any 
``8'' results.}\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
Prismatic Wall

Abjuration

\textbf{Level:} Sor/Wiz 8

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:}Wall 4 ft./level wide, 2 ft./level high

\textbf{Duration:} 10 min./level (D)

\textbf{Saving Throw:} See text

\textbf{Spell Resistance:} See text

\textit{Prismatic wall }creates a vertical, opaque wall---a shimmering, multicolored 
plane of light that protects you from all forms of attack. The wall flashes with 
seven colors, each of which has a distinct power and purpose. The wall is immobile, 
and you can pass through and remain near the wall without harm. However, any other 
creature with less than 8 HD that is within 20 feet of the wall is blinded for 
2d4 rounds by the colors if it looks at the wall.

The wall's maximum proportions are 4 feet wide per caster level and 2 feet high 
per caster level. A \textit{prismatic wall }spell cast to materialize in a space 
occupied by a creature is disrupted, and the spell is wasted.

Each color in the wall has a special effect. The accompanying table shows the seven 
colors of the wall, the order in which they appear, their effects on creatures 
trying to attack you or pass through the wall, and the magic needed to negate each 
color.

The wall can be destroyed, color by color, in consecutive order, by various magical 
effects; however, the first color must be brought down before the second can be 
affected, and so on. A \textit{rod of cancellation }or a m\textit{age's disjunction 
}spell destroys a \textit{prismatic wall}, but an \textit{antimagic field }fails 
to penetrate it. \textit{Dispel magic }and \textit{greater dispel magic }cannot 
dispel the wall or anything beyond it. Spell resistance is effective against a 
\textit{prismatic wall}, but the caster level check must be repeated for each color 
present.

\textit{Prismatic wall }can be made permanent with a \textit{permanency }spell.

\begin{tabular}{|>{\raggedright}p{33pt}|>{\raggedright}p{29pt}|>{\raggedright}p{184pt}|>{\raggedright}p{55pt}|}
\hline
\subsection*{C\textbf{olor}} & \subsection*{O\textbf{rder}} & \subsection*{E\textbf{ffect 
of Color}} & \subsection*{N\textbf{egated By}}\tabularnewline
\hline
Red & 1st & Stops nonmagical ranged weapons.\linebreak{}
Deals 20 points of fire damage (Reflex half). & C\textit{one of cold}\tabularnewline
\hline
Orange & 2nd & Stops magical ranged weapons.\linebreak{}
Deals 40 points of acid damage (Reflex half). & G\textit{ust of wind}\tabularnewline
\hline
Yellow & 3rd & Stops poisons, gases, and petrification.\linebreak{}
Deals 80 points of electricity damage (Reflex half). & D\textit{isintegrate}\tabularnewline
\hline
Green & 4th & Stops breath weapons.\linebreak{}
Poison (Kills; Fortitude partial for 1d6 points of Con damage instead). & P\textit{asswall}\tabularnewline
\hline
Blue & 5th & Stops divination and mental attacks.\linebreak{}
Turned to stone (Fortitude negates). & M\textit{agic missile}\tabularnewline
\hline
Indigo & 6th & Stops all spells.\linebreak{}
Will save or become insane (as \textit{insanity }spell). & D\textit{aylight}\tabularnewline
\hline
Violet & 7th & Energy field destroys all objects and effects.\textsuperscript{\textbf{1}}\linebreak{}
Creatures sent to another plane (Will negates). & D\textit{ispel magic}\tabularnewline
\hline
\multicolumn{4}{|p{302pt}|}{1 The violet effect makes the special effects of the 
other six colors redundant, but these six effects are included here because certain 
magic items can create prismatic effects one color at a time, and spell resistance 
might render some colors ineffective (see above).}\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
Produce Flame

Evocation [Fire]

\textbf{Level:} Drd 1, Fire 2

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 0 ft.

\textbf{Effect:} Flame in your palm

\textbf{Duration:} 1 min./level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

Flames as bright as a torch appear in your open hand. The flames harm neither you 
nor your equipment.

In addition to providing illumination, the flames can be hurled or used to touch 
enemies. You can strike an opponent with a melee touch attack, dealing fire damage 
equal to 1d6 +1 point per caster level (maximum +5). Alternatively, you can hurl 
the flames up to 120 feet as a thrown weapon. When doing so, you attack with a 
ranged touch attack (with no range penalty) and deal the same damage as with the 
melee attack. No sooner do you hurl the flames than a new set appears in your hand. 
Each attack you make reduces the remaining duration by 1 minute. If an attack reduces 
the remaining duration to 0 minutes or less, the spell ends after the attack resolves.

This spell does not function underwater.

\vspace{12pt}
Programmed Image

Illusion (Figment)

\textbf{Level:} Brd 6, Sor/Wiz 6

\textbf{Effect:} Visual figment that cannot extend beyond a 20-ft. cube + one 10-ft. 
cube/level (S)

\textbf{Duration:} Permanent until triggered, then 1 round/level

This spell functions like \textit{silent image, }except that this spell's figment 
activates when a specific condition occurs. The figment includes visual, auditory, 
olfactory, and thermal elements, including intelligible speech.

You set the triggering condition (which may be a special word) when casting the 
spell. The event that triggers the illusion can be as general or as specific and 
detailed as desired but must be based on an audible, tactile, olfactory, or visual 
trigger. The trigger cannot be based on some quality not normally obvious to the 
senses, such as alignment. (See \textit{magic mouth }for more details about such 
triggers.)

\textit{Material Component: }A bit of fleece and jade dust worth 25 gp.

\vspace{12pt}
Project Image

Illusion (Shadow)

\textbf{Level:} Brd 6, Sor/Wiz 7

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Medium (100 ft. + 10 ft./level)

\textbf{Effect:} One shadow duplicate

\textbf{Duration:} 1 round/level (D)

\textbf{Saving Throw: }Will disbelief (if interacted with)

\textbf{Spell Resistance:} No

You tap energy from the Plane of Shadow to create a quasi-real, illusory version 
of yourself. The projected image looks, sounds, and smells like you but is intangible. 
The projected image mimics your actions (including speech) unless you direct it 
to act differently (which is a move action).

You can see through its eyes and hear through its ears as if you were standing 
where it is, and during your turn you can switch from using its senses to using 
your own, or back again, as a free action. While you are using its senses, your 
body is considered blinded and deafened.

If you desire, any spell you cast whose range is touch or greater can originate 
from the projected image instead of from you. The projected image can't cast any 
spells on itself except for illusion spells. The spells affect other targets normally, 
despite originating from the projected image.

Objects are affected by the projected image as if they had succeeded on their Will 
save.

You must maintain line of effect to the projected image at all times. If your line 
of effect is obstructed, the spell ends. If you use \textit{dimension door, teleport, 
plane shift, }or a similar spell that breaks your line of effect, even momentarily, 
the spell ends.

\textit{Material Component: }A small replica of you (a doll), which costs 5 gp 
to create.

\vspace{12pt}
Protection from Arrows

Abjuration

\textbf{Level:} Sor/Wiz 2

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 1 hour/level or until discharged

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

The warded creature gains resistance to ranged weapons. The subject gains damage 
reduction 10/magic against ranged weapons. (This spell doesn't grant you the ability 
to damage creatures with similar damage reduction.) Once the spell has prevented 
a total of 10 points of damage per caster level (maximum 100 points), it is discharged.

\textit{Focus: }A piece of shell from a tortoise or a turtle.

\vspace{12pt}
Protection from Chaos

Abjuration [Lawful]

\textbf{Level:} Clr 1, Law 1, Pal 1, Sor/Wiz 1

This spell functions like \textit{protection from evil}, except that the deflection 
and resistance bonuses apply to attacks from chaotic creatures, and chaotic summoned 
creatures cannot touch the subject.

\vspace{12pt}
Protection from Energy

Abjuration

\textbf{Level:} Clr 3, Drd 3, Luck 3, Protection 3, Rgr 2, Sor/Wiz 3

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 10 min./level or until discharged

\textbf{Saving Throw:} Fortitude negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

\textit{Protection from energy }grants temporary immunity to the type of energy 
you specify when you cast it (acid, cold, electricity, fire, or sonic). When the 
spell absorbs 12 points per caster level of energy damage (to a maximum of 120 
points at 10th level), it is discharged.

\textit{Note: Protection from energy }overlaps (and does not stack with) \textit{resist 
energy. }If a character is warded by \textit{protection from energy }and \textit{resist 
energy, }the \textit{protection }spell absorbs damage until its power is exhausted.

\vspace{12pt}
Protection from Evil

Abjuration [Good]

\textbf{Level:} Clr 1, Good 1, Pal 1, Sor/Wiz 1

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 1 min./level (D)

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} No; see text

This spell wards a creature from attacks by evil creatures, from mental control, 
and from summoned creatures. It creates a magical barrier around the subject at 
a distance of 1 foot. The barrier moves with the subject and has three major effects.

First, the subject gains a +2 deflection bonus to AC and a +2 resistance bonus 
on saves. Both these bonuses apply against attacks made or effects created by evil 
creatures.

Second, the barrier blocks any attempt to possess the warded creature (by a \textit{magic 
jar }attack, for example) or to exercise mental control over the creature (including 
enchantment (charm) effects and enchantment (compulsion) effects that grant the 
caster ongoing control over the subject, such as \textit{dominate person}). The 
protection does not prevent such effects from targeting the protected creature, 
but it suppresses the effect for the duration of the \textit{protection from evil 
}effect. If the \textit{protection from evil }effect ends before the effect granting 
mental control does, the would-be controller would then be able to mentally command 
the controlled creature. Likewise, the barrier keeps out a possessing life force 
but does not expel one if it is in place before the spell is cast. This second 
effect works regardless of alignment.

Third, the spell prevents bodily contact by summoned creatures. This causes the 
natural weapon attacks of such creatures to fail and the creatures to recoil if 
such attacks require touching the warded creature. Good summoned creatures are 
immune to this effect. The protection against contact by summoned creatures ends 
if the warded creature makes an attack against or tries to force the barrier against 
the blocked creature. Spell resistance can allow a creature to overcome this protection 
and touch the warded creature.

\textit{Arcane Material Component: }A little powdered silver with which you trace 
a 3-foot -diameter circle on the floor (or ground) around the creature to be warded.

\vspace{12pt}
Protection from Good

Abjuration [Evil]

\textbf{Level:} Clr 1, Evil 1, Sor/Wiz 1

This spell functions like \textit{protection from evil}, except that the deflection 
and resistance bonuses apply to attacks from good creatures, and good summoned 
creatures cannot touch the subject.

\vspace{12pt}
Protection from Law

Abjuration [Chaotic]

\textbf{Level:} Chaos 1, Clr 1, Sor/Wiz 1

This spell functions like \textit{protection from evil}, except that the deflection 
and resistance bonuses apply to attacks from lawful creatures, and lawful summoned 
creatures cannot touch the subject.

\vspace{12pt}
Protection from Spells

Abjuration

\textbf{Level: }Magic 8, Sor/Wiz 8

\textbf{Components:} V, S, M, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Targets:} Up to one creature touched per four levels

\textbf{Duration:} 10 min./level

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

The subject gains a +8 resistance bonus on saving throws against spells and spell-like 
abilities (but not against supernatural and extraordinary abilities).

\textit{Material Component: }A diamond of at least 500 gp value, which must be 
crushed and sprinkled over the targets.

\textit{Focus: }One 1,000 gp diamond per creature to be granted the protection. 
Each subject must carry one such gem for the duration of the spell. If a subject 
loses the gem, the spell ceases to affect him.

\vspace{12pt}
Prying Eyes

Divination

\textbf{Level:} Sor/Wiz 5

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 minute

\textbf{Range:} One mile

\textbf{Effect:} Ten or more levitating eyes

\textbf{Duration:} 1 hour/level; see text (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You create a number of semitangible, visible magical orbs (called ``eyes'') equal 
to 1d4 + your caster level. These eyes move out, scout around, and return as you 
direct them when casting the spell. Each eye can see 120 feet (normal vision only) 
in all directions.

While the individual eyes are quite fragile, they're small and difficult to spot. 
Each eye is a Fine construct, about the size of a small apple, that has 1 hit point, 
AC 18 (+8 bonus for its size), flies at a speed of 30 feet with perfect maneuverability, 
and has a +16 Hide modifier. It has a Spot modifier equal to your caster level 
(maximum +15) and is subject to illusions, darkness, fog, and any other factors 
that would affect your ability to receive visual information about your surroundings. 
An eye traveling through darkness must find its way by touch.

When you create the eyes, you specify instructions you want them to follow in a 
command of no more than twenty-five words. Any knowledge you possess is known by 
the eyes as well.

In order to report their findings, the eyes must return to your hand. Each replays 
in your mind all it has seen during its existence. It takes an eye 1 round to replay 
1 hour of recorded images. After relaying its findings, an eye disappears. 

If an eye ever gets more than 1 mile away from you, it instantly ceases to exist. 
However, your link with the eye is such that you won't know if the eye was destroyed 
because it wandered out of range or because of some other event.

The eyes exist for up to 1 hour per caster level or until they return to you. \textit{Dispel 
magic }can destroy eyes. Roll separately for each eye caught in an area dispel. 
Of course, if an eye is sent into darkness, it could hit a wall or similar obstacle 
and destroy itself.

\textit{Material Component: }A handful of crystal marbles.

\vspace{12pt}
Prying Eyes, Greater

Divination

\textbf{Level:} Sor/Wiz 8

This spell functions like \textit{prying eyes, }except that the eyes can see all 
things as they actually are, just as if they had \textit{true seeing }with a range 
of 120 feet. Thus, they can navigate darkened areas at full normal speed. Also, 
a \textit{greater prying eye'}s maximum Spot modifier is +25 instead of +15.

\vspace{12pt}
Purify Food and Drink

Transmutation

\textbf{Level:} Clr 0, Drd 0

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 10 ft.

\textbf{Target:} 1 cu. ft./level of contaminated food and water

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates (object)

\textbf{Spell Resistance:} Yes (object)

This spell makes spoiled, rotten, poisonous, or otherwise contaminated food and 
water pure and suitable for eating and drinking. This spell does not prevent subsequent 
natural decay or spoilage. Unholy water and similar food and drink of significance 
is spoiled by \textit{purify food and drink, }but the spell has no effect on creatures 
of any type nor upon magic potions.

\textit{Note: }Water weighs about 8 pounds per gallon. One cubic foot of water 
contains roughly 8 gallons and weighs about 60 pounds.

\vspace{12pt}
Pyrotechnics

Transmutation

\textbf{Level:} Brd 2, Sor/Wiz 2

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Long (400 ft. + 40 ft./level)

\textbf{Target:} One fire source, up to a 20-ft. cube

\textbf{Duration:} 1d4+1 rounds, or 1d4+1 rounds after creatures leave the smoke 
cloud; see text

\textbf{Saving Throw: }Will negates or Fortitude negates; see text

\textbf{Spell Resistance:} Yes or No; see text

\textit{Pyrotechnics }turns a fire into either a burst of blinding fireworks or 
a thick cloud of choking smoke, depending on the version you choose.

\textit{Fireworks: }The fireworks are a flashing, fiery, momentary burst of glowing, 
colored aerial lights. This effect causes creatures within 120 feet of the fire 
source to become blinded for 1d4+1 rounds (Will negates). These creatures must 
have line of sight to the fire to be affected. Spell resistance can prevent blindness.

\textit{Smoke Cloud: }A writhing stream of smoke billows out from the source, forming 
a choking cloud. The cloud spreads 20 feet in all directions and lasts for 1 round 
per caster level. All sight, even darkvision, is ineffective in or through the 
cloud. All within the cloud take -4 penalties to Strength and Dexterity (Fortitude 
negates). These effects last for 1d4+1 rounds after the cloud dissipates or after 
the creature leaves the area of the cloud. Spell resistance does not apply.

\textit{Material Component: }The spell uses one fire source, which is immediately 
extinguished. A fire so large that it exceeds a 20-foot cube is only partly extinguished. 
Magical fires are not extinguished, although a fire-based creature used as a source 
takes 1 point of damage per caster level.

\vspace{12pt}
Quench

Transmutation

\textbf{Level:} Drd 3

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Area or Target:} One 20-ft. cube/level (S) or one fire-based magic item

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None or Will negates (object)

\textbf{Spell Resistance:} No or Yes (object)

\textit{Quench }is often used to put out forest fires and other conflagrations. 
It extinguishes all nonmagical fires in its area. The spell also dispels any fire 
spells in its area, though you must succeed on a dispel check (1d20 +1 per caster 
level, maximum +15) against each spell to dispel it. The DC to dispel such spells 
is 11 + the caster level of the fire spell.

Each elemental (fire) creature within the area of a \textit{quench }spell takes 
1d6 points of damage per caster level (maximum 15d6, no save allowed).

Alternatively, you can target the spell on a single magic item that creates or 
controls flame\textit{. }The item loses all its fire-based magical abilities for 
1d4 hours unless it succeeds on a Will save. (Artifacts are immune to this effect.)

\vspace{12pt}
Rage

Enchantment (Compulsion) [Mind-Affecting]

\textbf{Level:} Brd 2, Sor/Wiz 3

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Targets:} One willing living creature per three levels, no two of which 
may be more than 30 ft. apart

\textbf{Duration:} Concentration + 1 round/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

Each affected creature gains a +2 morale bonus to Strength and Constitution, a 
+1 morale bonus on Will saves, and a -2 penalty to AC. The effect is otherwise 
identical with a barbarian's rage except that the subjects aren't fatigued at the 
end of the rage.

\vspace{12pt}
Rainbow Pattern

Illusion (Pattern) [Mind-Affecting]

\textbf{Level:} Brd 4, Sor/Wiz 4

\textbf{Components:} V (Brd only), S, M, F; see text

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Effect:} Colorful lights with a 20-ft.-radius spread

\textbf{Duration:} Concentration +1 round/ level (D)

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

A glowing, rainbow-hued pattern of interweaving colors fascinates those within 
it. \textit{Rainbow pattern }fascinates a maximum of 24 Hit Dice of creatures. 
Creatures with the fewest HD are affected first. Among creatures with equal HD, 
those who are closest to the spell's point of origin are affected first. An affected 
creature that fails its saves is fascinated by the pattern.

With a simple gesture (a free action), you can make the rainbow pattern move up 
to 30 feet per round (moving its effective point of origin). All fascinated creatures 
follow the moving rainbow of light, trying to get or remain within the effect. 
Fascinated creatures who are restrained and removed from the pattern still try 
to follow it. If the pattern leads its subjects into a dangerous area each fascinated 
creature gets a second save. If the view of the lights is completely blocked creatures 
who can't see them are no longer affected.

The spell does not affect sightless creatures.

\textit{Verbal Component: }A wizard or sorcerer need not utter a sound to cast 
this spell, but a bard must sing, play music, or recite a rhyme as a verbal component.

\textit{Material Component: }A piece of phosphor.

\textit{Focus: }A crystal prism.

\vspace{12pt}
Raise Dead

Conjuration (Healing)

\textbf{Level:} Clr 5

\textbf{Components:} V, S, M, DF

\textbf{Casting Time:} 1 minute

\textbf{Range:} Touch

\textbf{Target:} Dead creature touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None; see text

\textbf{Spell Resistance:} Yes (harmless)

You restore life to a deceased creature. You can raise a creature that has been 
dead for no longer than one day per caster level. In addition, the subject's soul 
must be free and willing to return. If the subject's soul is not willing to return, 
the spell does not work; therefore, a subject that wants to return receives no 
saving throw.

Coming back from the dead is an ordeal. The subject of the spell loses one level 
(or 1 Hit Die) when it is raised, just as if it had lost a level or a Hit Die to 
an energy-draining creature. If the subject is 1st level, it loses 2 points of 
Constitution instead (if this would reduce its Con to 0 or less, it can't be raised). 
This level/HD loss or Constitution loss cannot be repaired by any means. A character 
who died with spells prepared has a 50\% chance of losing any given spell upon 
being raised, in addition to losing spells for losing a level. A spellcasting creature 
that doesn't prepare spells (such as a sorcerer) has a 50\% chance of losing any 
given unused spell slot as if it had been used to cast a spell, in addition to 
losing spell slots for losing a level.

A raised creature has a number of hit points equal to its current Hit Dice. Any 
ability scores damaged to 0 are raised to 1. Normal poison and normal disease are 
cured in the process of raising the subject, but magical diseases and curses are 
not undone. While the spell closes mortal wounds and repairs lethal damage of most 
kinds, the body of the creature to be raised must be whole. Otherwise, missing 
parts are still missing when the creature is brought back to life. None of the 
dead creature's equipment or possessions are affected in any way by this spell.

A creature who has been turned into an undead creature or killed by a death effect 
can't be raised by this spell. Constructs, elementals, outsiders, and undead creatures 
can't be raised. The spell cannot bring back a creature that has died of old age.

\textit{Material Component: }Diamonds worth a total of least 5,000 gp.

\vspace{12pt}
Ray of Enfeeblement

Necromancy

\textbf{Level:} Sor/Wiz 1

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} Ray

\textbf{Duration:} 1 min./level

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

A coruscating ray springs from your hand. You must succeed on a ranged touch attack 
to strike a target. The subject takes a penalty to Strength equal to 1d6+1 per 
two caster levels (maximum 1d6+5). The subject's Strength score cannot drop below 
1.

\vspace{12pt}
Ray of Exhaustion

Necromancy

\textbf{Level:} Sor/Wiz 3

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} Ray

\textbf{Duration:} 1 min./level

\textbf{Saving Throw:} Fortitude partial; see text

\textbf{Spell Resistance:} Yes

A black ray projects from your pointing finger. You must succeed on a ranged touch 
attack with the ray to strike a target.

The subject is immediately exhausted for the spell's duration. A successful Fortitude 
save means the creature is only fatigued.

A character that is already fatigued instead becomes exhausted.

This spell has no effect on a creature that is already exhausted. Unlike normal 
exhaustion or fatigue, the effect ends as soon as the spell's duration expires.

\textit{Material Component: }A drop of sweat.

\vspace{12pt}
Ray of Frost

Evocation [Cold]

\textbf{Level:} Sor/Wiz 0

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} Ray

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} Yes

A ray of freezing air and ice projects from your pointing finger. You must succeed 
on a ranged touch attack with the ray to deal damage to a target. The ray deals 
1d3 points of cold damage.

\vspace{12pt}
Read Magic

Divination

\textbf{Level:} Brd 0, Clr 0, Drd 0, Pal 1, Rgr 1, Sor/Wiz 0

\textbf{Components:} V, S, F

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 10 min./level

By means of \textit{read magic, }you can decipher magical inscriptions on objects---books, 
scrolls, weapons, and the like---that would otherwise be unintelligible. This deciphering 
does not normally invoke the magic contained in the writing, although it may do 
so in the case of a cursed scroll. Furthermore, once the spell is cast and you 
have read the magical inscription, you are thereafter able to read that particular 
writing without recourse to the use of \textit{read magic. }You can read at the 
rate of one page (250 words) per minute. The spell allows you to identify a \textit{glyph 
of warding }with a DC 13 Spellcraft check, a \textit{greater glyph of warding }with 
a DC 16 Spellcraft check, or any \textit{symbol }spell with a Spellcraft check 
(DC 10 + spell level).

\textit{Read magic }can be made permanent with a \textit{permanency }spell.

\textit{Focus: }A clear crystal or mineral prism.

\vspace{12pt}
Reduce Animal

Transmutation

\textbf{Level:} Drd 2, Rgr 3

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} One willing animal of Small, Medium, Large, or Huge size

\textbf{Duration:} 1 hour/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

This spell functions like \textit{reduce person, }except that it affects a single 
willing animal. Reduce the damage dealt by the animal's natural attacks as appropriate 
for its new size.

\vspace{12pt}
Reduce Person

Transmutation

\textbf{Level:} Sor/Wiz 1

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 round

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Target:} One humanoid creature

\textbf{Duration:} 1 min./level (D)

\textbf{Saving Throw:} Fortitude negates

\textbf{Spell Resistance:} Yes

This spell causes instant diminution of a humanoid creature, halving its height, 
length, and width and dividing its weight by 8. This decrease changes the creature's 
size category to the next smaller one. The target gains a +2 size bonus to Dexterity, 
a -2 size penalty to Strength (to a minimum of 1), and a +1 bonus on attack rolls 
and AC due to its reduced size.

A Small humanoid creature whose size decreases to Tiny has a space of 2-1/2 feet 
and a natural reach of 0 feet (meaning that it must enter an opponent's square 
to attack). A Large humanoid creature whose size decreases to Medium has a space 
of 5 feet and a natural reach of 5 feet. This spell doesn't change the target's 
speed.

All equipment worn or carried by a creature is similarly reduced by the spell.

Melee and projectile weapons deal less damage. Other magical properties are not 
affected by this spell. Any \textit{reduced }item that leaves the \textit{reduced 
}creature's possession (including a projectile or thrown weapon) instantly returns 
to its normal size. This means that thrown weapons deal their normal damage (projectiles 
deal damage based on the size of the weapon that fired them).

Multiple magical effects that reduce size do not stack.

\textit{Reduce person }counters and dispels \textit{enlarge person}.

\textit{Reduce person }can be made permanent with a \textit{permanency }spell.

\textit{Material Component: }A pinch of powdered iron.

\vspace{12pt}
Reduce Person, Mass

Transmutation

\textbf{Level:} Sor/Wiz 4

\textbf{Target:} One humanoid creature/level, no two of which can be more than 
30 ft. apart

This spell functions like \textit{reduce person}, except that it affects multiple 
creatures.

\vspace{12pt}
Refuge

Conjuration (Teleportation)

\textbf{Level:} Clr 7, Sor/Wiz 9

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Object touched

\textbf{Duration:} Permanent until discharged

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

You create powerful magic in some specially prepared object. This object contains 
the power to instantly transport its possessor across any distance within the same 
plane to your abode. Once the item is transmuted, you must give it willingly to 
a creature and at the same time inform it of a command word to be spoken when the 
item is used. To make use of the item, the subject speaks the command word at the 
same time that it rends or breaks the item (a standard action). When this is done, 
the individual and all objects it is wearing and carrying (to a maximum of the 
character's heavy load) are instantly transported to your abode. No other creatures 
are affected (aside from a familiar that is touching the subject).

You can alter the spell when casting it so that it transports you to within 10 
feet of the possessor of the item when it is broken and the command word spoken. 
You will have a general idea of the location and situation of the item possessor 
at the time the \textit{refuge }spell is discharged, but once you decide to alter 
the spell in this fashion, you have no choice whether or not to be transported.

\textit{Material Component: }The specially prepared object, whose construction 
requires gems worth 1,500 gp.

\vspace{12pt}
Regenerate

Conjuration (Healing)

\textbf{Level:} Clr 7, Drd 9, Healing 7

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 3 full rounds

\textbf{Range:} Touch

\textbf{Target:} Living creature touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} Fortitude negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

The subject's severed body members (fingers, toes, hands, feet, arms, legs, tails, 
or even heads of multiheaded creatures), broken bones, and ruined organs grow back. 
After the spell is cast, the physical regeneration is complete in 1 round if the 
severed members are present and touching the creature. It takes 2d10 rounds otherwise.

\textit{Regenerate }also cures 4d8 points of damage +1 point per caster level (maximum 
+35), rids the subject of exhaustion and/or fatigue, and eliminates all nonlethal 
damage the subject has taken. It has no effect on nonliving creatures (including 
undead).

\vspace{12pt}
Reincarnate

Transmutation

\textbf{Level:} Drd 4

\textbf{Components:} V, S, M, DF

\textbf{Casting Time:} 10 minutes

\textbf{Range:} Touch

\textbf{Target:} Dead creature touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} None; see text

\textbf{Spell Resistance:} Yes (harmless)

With this spell, you bring back a dead creature in another body, provided that 
its death occurred no more than one week before the casting of the spell and the 
subject's soul is free and willing to return. If the subject's soul is not willing 
to return, the spell does not work; therefore, a subject that wants to return receives 
no saving throw.

Since the dead creature is returning in a new body, all physical ills and afflictions 
are repaired. The condition of the remains is not a factor. So long as some small 
portion of the creature's body still exists, it can be reincarnated, but the portion 
receiving the spell must have been part of the creature's body at the time of death. 
The magic of the spell creates an entirely new young adult body for the soul to 
inhabit from the natural elements at hand. This process takes 1 hour to complete. 
 When the body is ready, the subject is reincarnated.

A reincarnated creature recalls the majority of its former life and form. It retains 
any class abilities, feats, or skill ranks it formerly possessed. Its class, base 
attack bonus, base save bonuses, and hit points are unchanged. Strength, Dexterity, 
and Constitution scores depend partly on the new body. First eliminate the subject's 
racial adjustments (since it is no longer of his previous race) and then apply 
the adjustments found below to its remaining ability scores. The subject's level 
(or Hit Dice) is reduced by 1. If the subject was 1st level, its new Constitution 
score is reduced by 2. (If this reduction would put its Con at 0 or lower, it can't 
be reincarnated). This level/HD loss or Constitution loss cannot be repaired by 
any means.

It's possible for the change in the subject's ability scores to make it difficult 
for it to pursue its previous character class. If this is the case, the subject 
is well advised to become a multiclass character.

For a humanoid creature, the new incarnation is determined using the following 
table. For nonhumanoid creatures, a similar table of creatures of the same type 
should be created.

A creature that has been turned into an undead creature or killed by a death effect 
can't be returned to life by this spell.  Constructs, elementals, outsiders, and 
undead creatures can't be reincarnated. The spell cannot bring back a creature 
who has died of old age.

\begin{tabular}{|>{\raggedright}p{23pt}|>{\raggedright}p{49pt}|>{\raggedright}p{12pt}|>{\raggedright}p{15pt}|>{\raggedright}p{16pt}|}
\hline
d\textbf{\%} & I\textbf{ncarnation} & S\textbf{tr} & D\textbf{ex} & C\textbf{on}\tabularnewline
\hline
01 & Bugbear & +4 & +2 & +2\tabularnewline
\hline
02-13 & Dwarf & +0 & +0 & +2\tabularnewline
\hline
14-25 & Elf & +0 & +2- & 2\tabularnewline
\hline
26 & Gnoll & +4 & +0 & +2\tabularnewline
\hline
27-38 & Gnome- & 2 & +0 & +2\tabularnewline
\hline
39-42 & Goblin- & 2 & +2 & +0\tabularnewline
\hline
43-52 & Half-elf & +0 & +0 & +0\tabularnewline
\hline
53-62 & Half-orc & +2 & +0 & +0\tabularnewline
\hline
63-74 & Halfling- & 2 & +2 & +0\tabularnewline
\hline
75-89 & Human & +0 & +0 & +0\tabularnewline
\hline
90-93 & Kobold- & 4 & +2- & 2\tabularnewline
\hline
94 & Lizardfolk & +2 & +0 & +2\tabularnewline
\hline
95-98 & Orc & +4 & +0 & +0\tabularnewline
\hline
99 & Troglodyte & +0- & 2 & +4\tabularnewline
\hline
100 & Other & ? & ? & ?\tabularnewline
\hline
\end{tabular}

The reincarnated creature gains all abilities associated with its new form, including 
forms of movement and speeds, natural armor, natural attacks, extraordinary abilities, 
and the like, but it doesn't automatically speak the language of the new form. 

A \textit{wish }or a \textit{miracle }spell can restore a reincarnated character 
to his or her original form.

\textit{Material Component: }Rare oils and unguents worth a total of least 1,000 
gp, spread over the remains.

\vspace{12pt}
Remove Blindness/Deafness

Conjuration (Healing)

\textbf{Level:} Clr 3, Pal 3

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} Fortitude negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

\textit{Remove blindness/deafness }cures blindness or deafness (your choice), whether 
the effect is normal or magical in nature. The spell does not restore ears or eyes 
that have been lost, but it repairs them if they are damaged.

\textit{Remove blindness/deafness }counters and dispels \textit{blindness/deafness}.

\vspace{12pt}
Remove Curse

Abjuration

\textbf{Level:} Brd 3, Clr 3, Pal 3, Sor/Wiz 4

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature or item touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

\textit{Remove curse }instantaneously removes all curses on an object or a creature. 
\textit{Remove curse }does not remove the curse from a cursed shield, weapon, or 
suit of armor, although the spell typically enables the creature afflicted with 
any such cursed item to remove and get rid of it. Certain special curses may not 
be countered by this spell or may be countered only by a caster of a certain level 
or higher.

\textit{Remove curse }counters and dispels \textit{bestow curse}.

\vspace{12pt}
Remove Disease

Conjuration (Healing)

\textbf{Level:} Clr 3, Drd 3, Rgr 3

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw:} Fortitude negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

\textit{Remove disease }cures all diseases that the subject is suffering from. 
The spell also kills parasites, including green slime and others. Certain special 
diseases may not be countered by this spell or may be countered only by a caster 
of a certain level or higher.

\textit{Note: }Since the spell's duration is instantaneous, it does not prevent 
reinfection after a new exposure to the same disease at a later date.

\vspace{12pt}
Remove Fear

Abjuration

\textbf{Level:} Brd 1, Clr 1

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} One creature plus one additional creature per four levels, no 
two of which can be more than 30 ft. apart

\textbf{Duration:} 10 minutes; see text

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

You instill courage in the subject, granting it a +4 morale bonus against \textit{fear 
}effects for 10 minutes. If the subject is under the influence of a \textit{fear 
}effect when receiving the spell, that effect is suppressed for the duration of 
the spell.

\textit{Remove fear }counters and dispels \textit{cause fear}.

\vspace{12pt}
Remove Paralysis

Conjuration (Healing)

\textbf{Level:} Clr 2, Pal 2

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Targets:} Up to four creatures, no two of which can be more than 30 ft. 
apart

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

You can free one or more creatures from the effects of any temporary paralysis 
or related magic, including a ghoul's touch or a \textit{slow }spell. If the spell 
is cast on one creature, the paralysis is negated. If cast on two creatures, each 
receives another save with a +4 resistance bonus against the effect that afflicts 
it. If cast on three or four creatures, each receives another save with a +2 resistance 
bonus.

The spell does not restore ability scores reduced by penalties, damage, or drain.

\vspace{12pt}
Repel Metal or Stone

Abjuration [Earth]

\textbf{Level:} Drd 8

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 60 ft.

\textbf{Area:} 60-ft. line from you

\textbf{Duration:} 1 round/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

Like \textit{repel wood, }this spell creates waves of invisible and intangible 
energy that roll forth from you. All metal or stone objects in the path of the 
spell are pushed away from you to the limit of the range. Fixed metal or stone 
objects larger than 3 inches in diameter and loose objects weighing more than 500 
pounds are not affected. Anything else, including animated objects, small boulders, 
and creatures in metal armor, moves back. Fixed objects 3 inches in diameter or 
smaller bend or break, and the pieces move with the wave of energy. Objects affected 
by the spell are repelled at the rate of 40 feet per round.

Objects such as metal armor, swords, and the like are pushed back, dragging their 
bearers with them. Even magic items with metal components are repelled, although 
an \textit{antimagic field }blocks the effects.

The waves of energy continue to sweep down the set path for the spell's duration. 
After you cast the spell, the path is set, and you can then do other things or 
go elsewhere without affecting the spell's power.

\vspace{12pt}
Repel Vermin

Abjuration

\textbf{Level:} Brd 4, Clr 4, Drd 4, Rgr 3

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 10 ft.

\textbf{Area:} 10-ft.-radius emanation centered on you

\textbf{Duration:} 10 min./level (D)

\textbf{Saving Throw:} None or Will negates; see text

\textbf{Spell Resistance:} Yes

An invisible barrier holds back vermin. A vermin with Hit Dice of less than one-third 
your level cannot penetrate the barrier.

A vermin with Hit Dice of one-third your level or more can penetrate the barrier 
if it succeeds on a Will save. Even so, crossing the barrier deals the vermin 2d6 
points of damage, and pressing against the barrier causes pain, which deters most 
vermin.

\vspace{12pt}
Repel Wood

Transmutation

\textbf{Level:} Drd 6, Plant 6

\textbf{Components:} V, S

\textbf{Casting Time:} 1 standard action

\textbf{Range:} 60 ft.

\textbf{Area:} 60-ft. line-shaped emanation from you

\textbf{Duration:} 1 min./level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

Waves of energy roll forth from you, moving in the direction that you determine, 
causing all wooden objects in the path of the spell to be pushed away from you 
to the limit of the range. Wooden objects larger than 3 inches in diameter that 
are fixed firmly are not affected, but loose objects are. Objects 3 inches in diameter 
or smaller that are fixed in place splinter and break, and the pieces move with 
the wave of energy. Objects affected by the spell are repelled at the rate of 40 
feet per round.

Objects such as wooden shields, spears, wooden weapon shafts and hafts, and arrows 
and bolts are pushed back, dragging those carrying them along. (A creature being 
dragged by an item it is carrying can let go. A creature being dragged by a shield 
can loose it as a move action and drop it as a free action.) If a spear is planted 
(set) to prevent this forced movement, it splinters. Even magic items with wooden 
sections are repelled, although an \textit{antimagic field }blocks the effects.

The waves of energy continue to sweep down the set path for the spell's duration. 
After you cast the spell, the path is set, and you can then do other things or 
go elsewhere without affecting the spell's power.

\vspace{12pt}
Repulsion

Abjuration

\textbf{Level:} Clr 7, Protection 7, Sor/Wiz 6

\textbf{Components:} V, S, F/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Up to 10 ft./level

\textbf{Area:} Up to 10-ft.-radius/level emanation centered on you

\textbf{Duration:} 1 round/level (D)

\textbf{Saving Throw: }Will negates

\textbf{Spell Resistance:} Yes

An invisible, mobile field surrounds you and prevents creatures from approaching 
you. You decide how big the field is at the time of casting (to the limit your 
level allows). Any creature within or entering the field must attempt a save. If 
it fails, it becomes unable to move toward you for the duration of the spell. Repelled 
creatures' actions are not otherwise restricted.

They can fight other creatures and can cast spells and attack you with ranged weapons. 
If you move closer to an affected creature, nothing happens. (The creature is not 
forced back.) The creature is free to make melee attacks against you if you come 
within reach. If a repelled creature moves away from you and then tries to turn 
back toward you, it cannot move any closer if it is still within the spell's area.

\textit{Arcane Focus: }A pair of small iron bars attached to two small canine statuettes, 
one black and one white, the whole array worth 50 gp.

\vspace{12pt}
Resilient Sphere

Evocation [Force]

\textbf{Level:} Sor/Wiz 4

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Close (25 ft. + 5 ft./2 levels)

\textbf{Effect:} 1-ft.-diameter/level sphere, centered around a creature

\textbf{Duration:} 1 min./level (D)

\textbf{Saving Throw:} Reflex negates

\textbf{Spell Resistance:} Yes

A globe of shimmering force encloses a creature, provided the creature is small 
enough to fit within the diameter of the sphere. The sphere contains its subject 
for the spell's duration. The sphere is not subject to damage of any sort except 
from a \textit{rod of cancellation}, a rod \textit{of negation}, a \textit{disintegrate 
}spell, or a targeted \textit{dispel magic }spell. These effects destroy the sphere 
without harm to the subject. Nothing can pass through the sphere, inside or out, 
though the subject can breathe normally.

The subject may struggle, but the sphere cannot be physically moved either by people 
outside it or by the struggles of those within.

\textit{Material Component: }A hemispherical piece of clear crystal and a matching 
hemispherical piece of gum arabic.

\vspace{12pt}
Resistance

Abjuration

\textbf{Level:} Brd 0, Clr 0, Drd 0, Pal 1, Sor/Wiz 0

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 1 minute

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

You imbue the subject with magical energy that protects it from harm, granting 
it a +1 resistance bonus on saves.

\textit{Resistance }can be made permanent with a \textit{permanency }spell.

\textit{Arcane Material Component: }A miniature cloak.

\vspace{12pt}
Resist Energy

Abjuration

\textbf{Level:} Clr 2, Drd 2, Fire 3, Pal 2, Rgr 1, Sor/Wiz 2

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} 10 min./level

\textbf{Saving Throw:} Fortitude negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

This abjuration grants a creature limited protection from damage of whichever one 
of five energy types you select: acid, cold, electricity, fire, or sonic. The subject 
gains energy resistance 10 against the energy type chosen, meaning that each time 
the creature is subjected to such damage (whether from a natural or magical source), 
that damage is reduced by 10 points before being applied to the creature's hit 
points. The value of the energy resistance granted increases to 20 points at 7th 
level and to a maximum of 30 points at 11th level. The spell protects the recipient's 
equipment as well.

\textit{Resist energy }absorbs only damage. The subject could still suffer unfortunate 
side effects.

\textit{Note: Resist energy }overlaps (and does not stack with) \textit{protection 
from energy. }If a character is warded by \textit{protection from energy }and \textit{resist 
energy, }the \textit{protection }spell absorbs damage until its power is exhausted.

\vspace{12pt}
Restoration

Conjuration (Healing)

\textbf{Level:} Clr 4, Pal 4

\textbf{Components:} V, S, M

This spell functions like \textit{lesser restoration, }except that it also dispels 
negative levels and restores one experience level to a creature who has had a level 
drained. The drained level is restored only if the time since the creature lost 
the level is equal to or less than one day per caster level. A character who has 
a level restored by r\textit{estoration has }exactly the minimum number of experience 
points necessary to restore him or her to his or her previous level.

\textit{Restoration }cures all temporary ability damage, and it restores all points 
permanently drained from a single ability score (your choice if more than one is 
drained). It also eliminates any fatigue or exhaustion suffered by the target. 

\textit{Restoration }does not restore levels or Constitution points lost due to 
death.

\textit{Material Component: }Diamond dust worth 100 gp that is sprinkled over the 
target.

\vspace{12pt}
Restoration, Greater

Conjuration (Healing)

\textbf{Level:} Clr 7

\textbf{Components:} V, S, XP

\textbf{Casting Time:} 10 minutes

This spell functions like \textit{lesser restoration, }except that it dispels all 
negative levels afflicting the healed creature. This effect also reverses level 
drains by a force or creature, restoring the creature to the highest level it had 
previously attained. The drained levels are restored only if the time since the 
creature lost the level is no more than one week per caster level.

\textit{Greater restoration }also dispels all magical effects penalizing the creature's 
abilities, cures all temporary ability damage, and restores all points permanently 
drained from all ability scores. It also eliminates fatigue and exhaustion, and 
removes all forms of insanity, \textit{confusion, }and similar mental effects. 
\textit{Greater restoration }does not restore levels or Constitution points lost 
due to death.

\textit{XP Cost: }500 XP.

\vspace{12pt}
Restoration, Lesser

Conjuration (Healing)

\textbf{Level:} Clr 2, Drd 2, Pal 1

\textbf{Components:} V, S

\textbf{Casting Time:} 3 rounds

\textbf{Range:} Touch

\textbf{Target:} Creature touched

\textbf{Duration:} Instantaneous

\textbf{Saving Throw: }Will negates (harmless)

\textbf{Spell Resistance:} Yes (harmless)

\textit{Lesser restoration }dispels any magical effects reducing one of the subject's 
ability scores or cures 1d4 points of temporary ability damage to one of the subject's 
ability scores. It also eliminates any fatigue suffered by the character, and improves 
an exhausted condition to fatigued. It does not restore permanent ability drain.

\vspace{12pt}
Resurrection

Conjuration (Healing)

\textbf{Level:} Clr 7

\textbf{Casting Time:} 10 minutes

This spell functions like \textit{raise dead, }except that you are able to restore 
life and complete strength to any deceased creature.

The condition of the remains is not a factor. So long as some small portion of 
the creature's body still exists, it can be resurrected, but the portion receiving 
the spell must have been part of the creature's body at the time of death. (The 
remains of a creature hit by a \textit{disintegrate }spell count as a small portion 
of its body.) The creature can have been dead no longer than 10 years per caster 
level.

Upon completion of the spell, the creature is immediately restored to full hit 
points, vigor, and health, with no loss of prepared spells. However, the subject 
loses one level, or 2 points of Constitution if the subject was 1st level. (If 
this reduction would bring its Con to 0 or lower, it can't be resurrected). This 
level loss or Constitution loss cannot be repaired by any means.

You can resurrect someone killed by a death effect or someone who has been turned 
into an undead creature and then destroyed. You cannot resurrect someone who has 
died of old age. Constructs, elementals, outsiders, and undead creatures can't 
be resurrected.

\textit{Material Component: }A sprinkle of holy water and diamonds worth a total 
of at least 10,000 gp.

\vspace{12pt}
Reverse Gravity

Transmutation

\textbf{Level:} Drd 8, Sor/Wiz 7

\textbf{Components:} V, S, M/DF

\textbf{Casting Time:} 1 standard action

\textbf{Range: }Medium (100 ft. + 10 ft./level)

\textbf{Area:} Up to one 10-ft. cube per two levels (S)

\textbf{Duration:} 1 round/level (D)

\textbf{Saving Throw:} None; see text

\textbf{Spell Resistance:} No

This spell reverses gravity in an area, causing all unattached objects and creatures 
within that area to fall upward and reach the top of the area in 1 round. If some 
solid object (such as a ceiling) is encountered in this fall, falling objects and 
creatures strike it in the same manner as they would during a normal downward fall. 
If an object or creature reaches the top of the area without striking anything, 
it remains there, oscillating slightly, until the spell ends. At the end of the 
spell duration, affected objects and creatures fall downward.

Provided it has something to hold onto, a creature caught in the area can attempt 
a Reflex save to secure itself when the spell strikes. Creatures who can fly or 
levitate can keep themselves from falling.

\textit{Arcane Material Component: }A lodestone and iron filings.

\vspace{12pt}
Righteous Might

Transmutation

\textbf{Level:} Clr 5, Strength 5

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Personal

\textbf{Target:} You

\textbf{Duration:} 1 round/level (D)

Your height immediately doubles, and your weight increases by a factor of eight. 
This increase changes your size category to the next larger one, and you gain a 
+8 size bonus to Strength and a +4 size bonus to Constitution. You gain a +4 enhancement 
bonus to your natural armor. You gain damage reduction 5/evil (if you normally 
channel positive energy) or damage reduction 5/good (if you normally channel negative 
energy). At 12th level this damage reduction becomes 10/evil or 10/good, and at 
15th level it becomes 15/evil or 15/good (the maximum). Your size modifier for 
AC and attacks changes as appropriate to your new size category. This spell doesn't 
change your speed.  Determine space and reach as appropriate to your new size.

If insufficient room is available for the desired growth, you attain the maximum 
possible size and may make a Strength check (using your increased Strength) to 
burst any enclosures in the process. If you fail, you are constrained without harm 
by the materials enclosing you--- the spell cannot crush you by increasing your 
size.

All equipment you wear or carry is similarly enlarged by the spell. Melee and projectile 
weapons deal more damage. Other magical properties are not affected by this spell. 
Any enlarged item that leaves your possession (including a projectile or thrown 
weapon) instantly returns to its normal size. This means that thrown weapons deal 
their normal damage (projectiles deal damage based on the size of the weapon that 
fired them).

Multiple magical effects that increase size do not stack.

\vspace{12pt}
Rope Trick

Transmutation

\textbf{Level:} Sor/Wiz 2

\textbf{Components:} V, S, M

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} One touched piece of rope from 5 ft. to 30 ft. long

\textbf{Duration:} 1 hour/level (D)

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

When this spell is cast upon a piece of rope from 5 to 30 feet long, one end of 
the rope rises into the air until the whole rope hangs perpendicular to the ground, 
as if affixed at the upper end. The upper end is, in fact, fastened to an extradimensional 
space that is outside the multiverse of extradimensional spaces (``planes''). Creatures 
in the extradimensional space are hidden, beyond the reach of spells (including 
divinations), unless those spells work across planes. The space holds as many as 
eight creatures (of any size). Creatures in the space can pull the rope up into 
the space, making the rope ``disappear.'' In that case, the rope counts as one 
of the eight creatures that can fit in the space. The rope can support up to 16,000 
pounds. A weight greater than that can pull the rope free.

Spells cannot be cast across the extradimensional interface, nor can area effects 
cross it. Those in the extradimensional space can see out of it as if a 3-foot-by- 
5-foot window were centered on the rope. The window is present on the Material 
Plane, but it's invisible, and even creatures that can see the window can't see 
through it. Anything inside the extradimensional space drops out when the spell 
ends. The rope can be climbed by only one person at a time. The \textit{rope trick 
}spell enables climbers to reach a normal place if they do not climb all the way 
to the extradimensional space.

\textit{Note: }It is hazardous to create an extradimensional space within an existing 
extradimensional space or to take an extradimensional space into an existing one.

\textit{Material Component: }Powdered corn extract and a twisted loop of parchment.

\vspace{12pt}
Rusting Grasp

Transmutation

\textbf{Level:} Drd 4

\textbf{Components:} V, S, DF

\textbf{Casting Time:} 1 standard action

\textbf{Range:} Touch

\textbf{Target:} One nonmagical ferrous object (or the volume of the object within 
3 ft. of the touched point) or one ferrous creature

\textbf{Duration:} See text

\textbf{Saving Throw:} None

\textbf{Spell Resistance:} No

Any iron or iron alloy item you touch becomes instantaneously rusted, pitted, and 
worthless, effectively destroyed. If the item is so large that it cannot fit within 
a 3-foot radius a 3-foot-radius volume of the metal is rusted and destroyed. Magic 
items made of metal are immune to this spell.

You may employ \textit{rusting grasp }in combat with a successful melee touch attack. 
\textit{Rusting grasp }used in this way instantaneously destroys 1d6 points of 
Armor Class gained from metal armor (to the maximum amount of protection the armor 
offered) through corrosion. 

Weapons in use by an opponent targeted by the spell are more difficult to grasp. 
You must succeed on a melee touch attack against the weapon. A metal weapon that 
is hit is destroyed.

\textit{Note: }Striking at an opponent's weapon provokes an attack of opportunity. 
Also, you must touch the weapon and not the other way around.

Against a ferrous creature, \textit{rusting grasp }instantaneously deals 3d6 points 
of damage +1 per caster level (maximum +15) per successful attack. The spell lasts 
for 1 round per level, and you can make one melee touch attack per round.

\newpage

\end{document}
