%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{array}
\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

This material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

{\LARGE{}MAGIC ITEMS III (POTIONS, RINGS, \& RODS)}

\vspace{12pt}
POTIONS AND OILS

A potion is a magic liquid that produces its effect when imbibed. Magic oils are 
similar to potions, except that oils are applied externally rather than imbibed. 
A potion or oil can be used only once. It can duplicate the effect of a spell of 
up to 3rd level that has a casting time of less than 1 minute.

Potions are like spells cast upon the imbiber. The character taking the potion 
doesn't get to make any decisions about the effect ---the caster who brewed the 
potion has already done so. The drinker of a potion is both the effective target 
and the caster of the effect (though the potion indicates the caster level, the 
drinker still controls the effect).

The person applying an oil is the effective caster, but the object is the target.

\textbf{Physical Description:} A typical potion or oil consists of 1 ounce of liquid 
held in a ceramic or glass vial fitted with a tight stopper. The stoppered container 
is usually no more than 1 inch wide and 2 inches high. The vial has AC 13, 1 hit 
point, hardness 1, and a break DC of 12. Vials hold 1 ounce of liquid.

\textbf{Identifying Potions:} In addition to the standard methods of identification, 
PCs can sample from each container they find to attempt to determine the nature 
of the liquid inside. An experienced character learns to identify potions by memory---

for example, the last time she tasted a liquid that reminded her of almonds, it 
turned out to be a \textit{potion of cure moderate wounds}. 

\textbf{Activation:} Drinking a potion or applying an oil requires no special skill. 
The user merely removes the stopper and swallows the potion or smears on the oil. 
The following rules govern potion and oil use.

Drinking a potion or using an oil on an item of gear is a standard action. The 
potion or oil takes effect immediately. Using a potion or oil provokes attacks 
of opportunity. A successful attack (including grappling attacks) against the character 
forces a Concentration check (as for casting a spell). If the character fails this 
check, she cannot drink the potion. An enemy may direct an attack of opportunity 
against the potion or oil container rather than against the character. A successful 
attack of this sort can destroy the container. 

A creature must be able to swallow a potion or smear on an oil. Because of this, 
incorporeal creatures cannot use potions or oils. 

Any corporeal creature can imbibe a potion. The potion must be swallowed. Any corporeal 
creature can use an oil.

A character can carefully administer a potion to an unconscious creature as a full-round 
action, trickling the liquid down the creature's throat. Likewise, it takes a full-round 
action to apply an oil to an unconscious creature.

\vspace{12pt}
Potion Descriptions

The caster level for a standard potion is the minimum caster level needed to cast 
the spell (unless otherwise specified).

\begin{tabular}{|>{\raggedright}p{27pt}|>{\raggedright}p{34pt}|>{\raggedright}p{26pt}|>{\raggedright}p{149pt}|>{\raggedright}p{52pt}|}
\hline
\multicolumn{5}{|p{290pt}|}{T\textbf{able: Potions and Oils}}\tabularnewline
\hline
M\textbf{inor} & M\textbf{edium} & M\textbf{ajor} & P\textbf{otion or Oil} & \subsubsection*{M\textbf{arket 
Price}}\tabularnewline
\hline
01-10--- & --- &  & C\textit{ure light wounds }(potion) & 50 gp\tabularnewline
\hline
11-13--- & --- &  & E\textit{ndure elements }(potion) & 50 gp\tabularnewline
\hline
14-15--- & --- &  & H\textit{ide from animals }(potion) & 50 gp\tabularnewline
\hline
16-17--- & --- &  & H\textit{ide from undead }(potion) & 50 gp\tabularnewline
\hline
18-19--- & --- &  & J\textit{ump }(potion) & 50 gp\tabularnewline
\hline
20-22--- & --- &  & M\textit{age armor }(potion) & 50 gp\tabularnewline
\hline
23-25--- & --- &  & M\textit{agic fang }(potion) & 50 gp\tabularnewline
\hline
26--- & --- &  & M\textit{agic stone }(oil) & 50 gp\tabularnewline
\hline
27-29--- & --- &  & M\textit{agic weapon }(oil) & 50 gp\tabularnewline
\hline
30--- & --- &  & P\textit{ass without trace }(potion) & 50 gp\tabularnewline
\hline
31-32--- & --- &  & P\textit{rotection from (alignment) }(potion) & 50 gp\tabularnewline
\hline
33-34--- & --- &  & R\textit{emove fear }(potion) & 50 gp\tabularnewline
\hline
35--- & --- &  & S\textit{anctuary }(potion) & 50 gp\tabularnewline
\hline
36-38--- & --- &  & S\textit{hield of faith +2 }(potion) & 50 gp\tabularnewline
\hline
39--- & --- &  & S\textit{hillelagh }(oil) & 50 gp\tabularnewline
\hline
40-41 & 01-02--- &  & B\textit{less weapon }(oil) & 100 gp\tabularnewline
\hline
42-44 & 03-04--- &  & E\textit{nlarge person }(potion) & 250 gp\tabularnewline
\hline
45 & 05--- &  & R\textit{educe person }(potion) & 250 gp\tabularnewline
\hline
46-47 & 06--- &  & A\textit{id }(potion) & 300 gp\tabularnewline
\hline
48-50 & 07--- &  & B\textit{arkskin +2 }(potion) & 300 gp\tabularnewline
\hline
51-53 & 08-10--- &  & B\textit{ear's endurance }(potion) & 300 gp\tabularnewline
\hline
54-56 & 11-13 & 01-02 & B\textit{lur }(potion) & 300 gp\tabularnewline
\hline
57-59 & 14-16--- &  & B\textit{ull's strength }(potion) & 300 gp\tabularnewline
\hline
60-62 & 17-19--- &  & C\textit{at's grace }(potion) & 300 gp\tabularnewline
\hline
63-67 & 20-27 & 03-07 & C\textit{ure moderate wounds }(potion) & 300 gp\tabularnewline
\hline
68 & 28--- &  & D\textit{arkness }(oil) & 300 gp\tabularnewline
\hline
69-71 & 29-30 & 08-09 & D\textit{arkvision }(potion) & 300 gp\tabularnewline
\hline
72-74 & 31--- &  & D\textit{elay poison }(potion) & 300 gp\tabularnewline
\hline
75-76 & 32-33--- &  & E\textit{agle's splendor }(potion) & 300 gp\tabularnewline
\hline
77-78 & 34-35--- &  & F\textit{ox's cunning }(potion) & 300 gp\tabularnewline
\hline
79-81 & 36-37 & 10-11 & I\textit{nvisibility }(potion or oil) & 300 gp\tabularnewline
\hline
82-84 & 38 & 12 & L\textit{esser restoration }(potion) & 300 gp\tabularnewline
\hline
85-86 & 39--- &  & L\textit{evitate }(potion or oil) & 300 gp\tabularnewline
\hline
87 & 40--- &  & M\textit{isdirection }(potion) & 300 gp\tabularnewline
\hline
88-89 & 41-42--- &  & O\textit{wl's wisdom }(potion) & 300 gp\tabularnewline
\hline
90-91 & 43--- &  & P\textit{rotection from arrows 10/magic }(potion) & 300 gp\tabularnewline
\hline
92-93 & 44 & 13 & R\textit{emove paralysis }(potion) & 300 gp\tabularnewline
\hline
94-96 & 45-46--- &  & R\textit{esist energy (type) 10 }(potion) & 300 gp\tabularnewline
\hline
97 & 47-48 & 14 & S\textit{hield of faith +3 }(potion) & 300 gp\tabularnewline
\hline
98-99 & 49--- &  & S\textit{pider climb }(potion) & 300 gp\tabularnewline
\hline
100 & 50 & 15 & U\textit{ndetectable alignment }(potion) & 300 gp\tabularnewline
\hline
--- & 51 & 16 & B\textit{arkskin +3 }(potion) & 600 gp\tabularnewline
\hline
--- & 52 & 17-18 & S\textit{hield of faith +4 }(potion) & 600 gp\tabularnewline
\hline
--- & 53-55 & 19-20 & R\textit{esist energy (type) 20 }(potion) & 700 gp\tabularnewline
\hline
--- & 56-60 & 21-28 & C\textit{ure serious wounds }(potion) & 750 gp\tabularnewline
\hline
--- & 61 & 29 & D\textit{aylight }(oil) & 750 gp\tabularnewline
\hline
--- & 62-64 & 30-32 & D\textit{isplacement }(potion) & 750 gp\tabularnewline
\hline
--- & 65 & 33 & F\textit{lame arrow }(oil) & 750 gp\tabularnewline
\hline
--- & 66-68 & 34-38 & F\textit{ly }(potion) & 750 gp\tabularnewline
\hline
--- & 69 & 39 & G\textit{aseous form }(potion) & 750 gp\tabularnewline
\hline
--- & 70-71--- &  & G\textit{reater magic fang +1 }(potion) & 750 gp\tabularnewline
\hline
--- & 72-73--- &  & G\textit{reater magic weapon +1 }(oil) & 750 gp\tabularnewline
\hline
--- & 74-75 & 40-41 & H\textit{aste }(potion) & 750 gp\tabularnewline
\hline
--- & 76-78 & 42-44 & H\textit{eroism }(potion) & 750 gp\tabularnewline
\hline
--- & 79-80 & 45-46 & K\textit{een edge }(oil) & 750 gp\tabularnewline
\hline
--- & 81 & 47 & M\textit{agic circle against (alignment) }(potion) & 750 gp\tabularnewline
\hline
--- & 82-83--- &  & M\textit{agic vestment +1 }(oil) & 750 gp\tabularnewline
\hline
--- & 84-86 & 48-50 & N\textit{eutralize poison }(potion) & 750 gp\tabularnewline
\hline
--- & 87-88 & 51-52 & N\textit{ondetection }(potion) & 750 gp\tabularnewline
\hline
--- & 89-91 & 53-54 & P\textit{rotection from energy (type) }(potion) & 750 gp\tabularnewline
\hline
--- & 92-93 & 55 & R\textit{age }(potion) & 750 gp\tabularnewline
\hline
--- & 94 & 56 & R\textit{emove blindness/deafness }(potion) & 750 gp\tabularnewline
\hline
--- & 95 & 57 & R\textit{emove curse }(potion) & 750 gp\tabularnewline
\hline
--- & 96 & 58 & R\textit{emove disease }(potion) & 750 gp\tabularnewline
\hline
--- & 97 & 59 & T\textit{ongues }(potion) & 750 gp\tabularnewline
\hline
--- & 98-99 & 60 & W\textit{ater breathing }(potion) & 750 gp\tabularnewline
\hline
--- & 100 & 61 & W\textit{ater walk }(potion) & 750 gp\tabularnewline
\hline
------ &  & 62-63 & B\textit{arkskin +4 }(potion) & 900 gp\tabularnewline
\hline
------ &  & 64 & S\textit{hield of faith +5 }(potion) & 900 gp\tabularnewline
\hline
------ &  & 65 & G\textit{ood hope }(potion) & 1,050 gp\tabularnewline
\hline
------ &  & 66-68 & R\textit{esist energy (type) 30 }(potion) & 1,100 gp\tabularnewline
\hline
------ &  & 69 & B\textit{arkskin +5 }(potion) & 1,200 gp\tabularnewline
\hline
------ &  & 70-73 & G\textit{reater magic fang +2 }(potion) & 1,200 gp\tabularnewline
\hline
------ &  & 74-77 & G\textit{reater magic weapon +2 }(oil) & 1,200 gp\tabularnewline
\hline
------ &  & 78-81 & M\textit{agic vestment +2 }(oil) & 1,200 gp\tabularnewline
\hline
------ &  & 82 & P\textit{rotection from arrows 15/magic }(potion) & 1,500 gp\tabularnewline
\hline
------ &  & 83-85 & G\textit{reater magic fang +3 }(potion) & 1,800 gp\tabularnewline
\hline
------ &  & 86-88 & G\textit{reater magic weapon +3 }(oil) & 1,800 gp\tabularnewline
\hline
------ &  & 89-91 & M\textit{agic vestment +3 }(oil) & 1,800 gp\tabularnewline
\hline
------ &  & 92-93 & G\textit{reater magic fang +4 }(potion) & 2,400 gp\tabularnewline
\hline
------ &  & 94-95 & G\textit{reater magic weapon +4 }(oil) & 2,400 gp\tabularnewline
\hline
------ &  & 96-97 & M\textit{agic vestment +4 }(oil) & 2,400 gp\tabularnewline
\hline
------ &  & 98 & G\textit{reater magic fang +5 }(potion) & 3,000 gp\tabularnewline
\hline
------ &  & 99 & G\textit{reater magic weapon +5 }(oil) & 3,000 gp\tabularnewline
\hline
------ &  & 100 & M\textit{agic vestment +5 }(oil) & 3,000 gp\tabularnewline
\hline
\end{tabular}

\vspace{24pt}
RINGS

Rings bestow magical powers upon their wearers. Only a rare few have charges. Anyone 
can use a ring.

A character can only effectively wear two magic rings. A third magic ring doesn't 
work if the wearer is already wearing two magic rings.

\textbf{Physical Description:} Rings have no appreciable weight. Although exceptions 
exist that are crafted from glass or bone, the vast majority of rings are forged 
from metal---usually precious metals such as gold, silver, and platinum. A ring 
has AC 13, 2 hit points, hardness 10, and a break DC of 25.

\textbf{Activation:} Usually, a ring's ability is activated by a command word (a 
standard action that does not provoke attacks of opportunity) or it works continually. 
Some rings have exceptional activation methods, according to their descriptions.

\textbf{Special Qualities:} Roll d\%. A result of 01 indicates the ring is intelligent, 
02-31 indicates that something (a design, inscription, or the like) provides a 
clue to its function, and 32-100 indicates no special qualities. Intelligent items 
have extra abilities and sometimes extraordinary powers and special purposes. Rings 
with charges can never be intelligent.

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{34pt}|>{\raggedright}p{40pt}|>{\raggedright}p{31pt}|>{\raggedright}p{117pt}|>{\raggedright}p{60pt}|}
\hline
\multicolumn{5}{|p{284pt}|}{T\textbf{able: Rings}}\tabularnewline
\hline
M\textbf{inor} & M\textbf{edium} & M\textbf{ajor} & R\textbf{ing} & \subsubsection*{M\textbf{arket 
Price}}\tabularnewline
\hline
01-18--- & --- &  & P\textit{rotection +1} & 2,000 gp\tabularnewline
\hline
19-28--- & --- &  & F\textit{eather falling} & 2,200 gp\tabularnewline
\hline
29-36--- & --- &  & S\textit{ustenance} & 2,500 gp\tabularnewline
\hline
37-44--- & --- &  & C\textit{limbing} & 2,500 gp\tabularnewline
\hline
45-52--- & --- &  & J\textit{umping} & 2,500 gp\tabularnewline
\hline
53-60--- & --- &  & S\textit{wimming} & 2,500 gp\tabularnewline
\hline
61-70 & 01-05--- &  & C\textit{ounterspells} & 4,000 gp\tabularnewline
\hline
71-75 & 06-08--- &  & M\textit{ind shielding} & 8,000 gp\tabularnewline
\hline
76-80 & 09-18--- &  & P\textit{rotection +2} & 8,000 gp\tabularnewline
\hline
81-85 & 19-23--- &  & F\textit{orce shield} & 8,500 gp\tabularnewline
\hline
86-90 & 24-28--- &  & R\textit{am} & 8,600 gp\tabularnewline
\hline
--- & 29-34--- &  & C\textit{limbing, improved} & 10,000 gp\tabularnewline
\hline
--- & 35-40--- &  & J\textit{umping, improved} & 10,000 gp\tabularnewline
\hline
--- & 41-46--- &  & S\textit{wimming, improved} & 10,000 gp\tabularnewline
\hline
91-93 & 47-51--- &  & A\textit{nimal friendship} & 10,800 gp\tabularnewline
\hline
94-96 & 50-56 & 01-02 & E\textit{nergy resistance, minor} & 12,000 gp\tabularnewline
\hline
97-98 & 57-61--- &  & C\textit{hameleon power} & 12,700 gp\tabularnewline
\hline
99-100 & 62-66--- &  & W\textit{ater walking} & 15,000 gp\tabularnewline
\hline
--- & 67-71 & 03-07 & P\textit{rotection +3} & 18,000 gp\tabularnewline
\hline
--- & 72-76 & 08-10 & S\textit{pell storing, minor} & 18,000 gp\tabularnewline
\hline
--- & 77-81 & 11-15 & I\textit{nvisibility} & 20,000 gp\tabularnewline
\hline
--- & 82-85 & 16-19 & W\textit{izardry (I)} & 20,000 gp\tabularnewline
\hline
--- & 86-90 & 20-25 & E\textit{vasion} & 25,000 gp\tabularnewline
\hline
--- & 91-93 & 26-28 & X\textit{-ray vision} & 25,000 gp\tabularnewline
\hline
--- & 94-97 & 29-32 & B\textit{linking} & 27,000 gp\tabularnewline
\hline
--- & 98-100 & 33-39 & E\textit{nergy resistance, major} & 28,000 gp\tabularnewline
\hline
------ &  & 40-49 & P\textit{rotection +4} & 32,000 gp\tabularnewline
\hline
------ &  & 50-55 & W\textit{izardry (II)} & 40,000 gp\tabularnewline
\hline
------ &  & 56-60 & F\textit{reedom of movement} & 40,000 gp\tabularnewline
\hline
------ &  & 61-63 & E\textit{nergy resistance, greater} & 44,000 gp\tabularnewline
\hline
------ &  & 64-65 & F\textit{riend shield }(pair) & 50,000 gp\tabularnewline
\hline
------ &  & 66-70 & P\textit{rotection +5} & 50,000 gp\tabularnewline
\hline
------ &  & 71-74 & S\textit{hooting stars} & 50,000 gp\tabularnewline
\hline
------ &  & 75-79 & S\textit{pell storing} & 50,000 gp\tabularnewline
\hline
------ &  & 80-83 & W\textit{izardry (III)} & 70,000 gp\tabularnewline
\hline
------ &  & 84-86 & T\textit{elekinesis} & 75,000 gp\tabularnewline
\hline
------ &  & 87-88 & R\textit{egeneration} & 90,000 gp\tabularnewline
\hline
------ &  & 89 & T\textit{hree wishes} & 97,950 gp\tabularnewline
\hline
------ &  & 90-92 & S\textit{pell turning} & 98,280 gp\tabularnewline
\hline
------ &  & 93-94 & W\textit{izardry (IV)} & 100,000 gp\tabularnewline
\hline
------ &  & 95 & D\textit{jinni calling} & 125,000 gp\tabularnewline
\hline
------ &  & 96 & E\textit{lemental command (air)} & 200,000 gp\tabularnewline
\hline
------ &  & 97 & E\textit{lemental command (earth)} & 200,000 gp\tabularnewline
\hline
------ &  & 98 & E\textit{lemental command (fire)} & 200,000 gp\tabularnewline
\hline
------ &  & 99 & E\textit{lemental command (water)} & 200,000 gp\tabularnewline
\hline
------ &  & 100 & S\textit{pell storing, major} & 200,000 gp\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
Ring Descriptions

Standard rings are described below.

\textbf{Animal Friendship:} On command, this ring affects an animal as if the wearer 
had cast \textit{charm animal}.

Faint enchantment; CL 3rd; Forge Ring, \textit{charm animal; }Price 10,800 gp.

\textbf{Blinking: }On command, this ring makes the wearer blink, as with the \textit{blink 
}spell.

Moderate transmutation; CL 7th; Forge Ring, \textit{blink; }Price 27,000 gp.

\textbf{Chameleon Power: }As a free action, the wearer of this ring can gain the 
ability to magically blend in with the surroundings. This provides a +10 competence 
bonus on her Hide checks. As a standard action, she can also command the ring to 
utilize the spell \textit{disguise self }as often as she wants.

Faint illusion; CL 3rd; Forge Ring, \textit{disguise self, invisibility; }Price 
12,700 gp.

\textbf{Climbing: }This ring is actually a magic leather cord that ties around 
a finger. It continually grants the wearer a +5 competence bonus on Climb checks.

Faint transmutatation; CL 5th; Forge Ring, creator must have 5 ranks in the Climb 
skill; Price 2,500 gp.

\textbf{Climbing, Improved: }As climbing, except it grants a +10 competence bonus 
on its wearer's Climb checks.

Faint transmutation; CL 5th; Forge Ring, creator must have 10 ranks in the Climb 
skill\textit{; }Price 10,000 gp.

\textbf{Counterspells:} This ring might seem to be a \textit{ring of spell storing 
}upon first examination. However, while it allows a single spell of 1st through 
6th level to be cast into it, that spell cannot be cast out of the ring again. 
Instead, should that spell ever be cast upon the wearer, the spell is immediately 
countered, as a counterspell action, requiring no action (or even knowledge) on 
the wearer's part. Once so used, the spell cast within the ring is gone. A new 
spell (or the same one as before) may be placed in it again.

Moderate evocation; CL 11th; Forge Ring, \textit{imbue with spell ability; }Price 
4,000 gp.

\textbf{Djinni Calling: }One of the many rings of fable, this ``genie'' ring is 
most useful indeed. It serves as a special \textit{gate }by means of which a specific 
djinni can be called from the Elemental Plane of Air. When the ring is rubbed (a 
standard action), the call goes out, and the djinni appears on the next round. 
The djinni faithfully obeys and serves the wearer of the ring, but never for more 
than 1 hour per day. If the djinni of the ring is ever killed, the ring becomes 
nonmagical and worthless.

Strong conjuration; CL 17th; Forge Ring, \textit{gate; }Price 125,000 gp.

\textbf{Elemental Command: }All four kinds of \textit{elemental command }rings 
are very powerful. Each appears to be nothing more than a lesser magic ring until 
fully activated (by meeting a special condition, such as single-handedly slaying 
an elemental of the appropriate type or exposure to a sacred material of the appropriate 
element), but each has certain other powers as well as the following common properties.

Elementals of the plane to which the ring is attuned can't attack the wearer, or 
even approach within 5 feet of him. If the wearer desires, he may forego this protection 
and instead attempt to \textit{charm }the elemental (as \textit{charm monster, 
}Will DC 17 negates). If the \textit{charm }attempt fails, however, absolute protection 
is lost and no further attempt at \textit{charming }can be made.

Creatures from the plane to which the ring is attuned who attack the wearer take 
a -1 penalty on their attack rolls. The ring wearer makes applicable saving throws 
from the extraplanar creature's attacks with a +2 resistance bonus. He gains a 
+4 morale bonus on all attack rolls against such creatures. Any weapon he uses 
bypasses the damage reduction of such creatures, regardless of any qualities the 
weapon may or may not have.

The wearer of the ring is able to converse with creatures from the plane to which 
his ring is attuned. These creatures recognize that he wears the ring. They show 
a healthy respect for the wearer if alignments are similar. If alignments are opposed, 
creatures fear the wearer if he is strong. If he is weak, they hate and desire 
to slay him.

The possessor of a \textit{ring of elemental command }takes a saving throw penalty 
as follows:

\begin{tabular}{|>{\raggedright}p{34pt}|>{\raggedright}p{164pt}|}
\hline
\subsection*{E\textbf{lement}} & \subsection*{S\textbf{aving Throw Penalty}}\tabularnewline
\hline
Air- & 2 against earth-based effects\tabularnewline
\hline
Earth- & 2 against air- or electricity-based effects\tabularnewline
\hline
Fire- & 2 against water- or cold-based effects\tabularnewline
\hline
Water- & 2 against fire-based effects\tabularnewline
\hline
\end{tabular}

In addition to the powers described above, each specific ring gives its wearer 
the following abilities according to its kind.

\textit{Ring of Elemental Command (Air)• }

\textit{Feather fall }(unlimited use, wearer only)• 

\textit{Resist energy (electricity) }(unlimited use, wearer only)• 

\textit{Gust of wind }(twice per day)• 

\textit{Wind wall }(unlimited use)• 

\textit{Air walk }(once per day, wearer only)• 

\textit{Chain lightning }(once per week)

The ring appears to be a \textit{ring of feather falling }until a certain condition 
is met to activate its full potential. It must be reactivated each time a new wearer 
acquires it.

\textit{Ring of Elemental Command (Earth)• }

\textit{Meld into stone }(unlimited use, wearer only)• 

\textit{Soften earth or stone }(unlimited use)• 

\textit{Stone shape }(twice per day)• 

\textit{Stoneskin }(once per week, wearer only)• 

\textit{Passwall }(twice per week)• 

\textit{Wall of stone }(once per day)

The ring appears to be a \textit{ring of meld into stone }until the established 
condition is met.

\textit{Ring of Elemental Command (Fire)• }

\textit{Resist energy (fire) }(as a \textit{major ring of energy resistance [fire]})• 

\textit{Burning hands }(unlimited use)• 

\textit{Flaming sphere }(twice per day)• 

\textit{Pyrotechnics }(twice per day)• 

\textit{Wall of fire }(once per day)• 

\textit{Flame strike }(twice per week)

The ring appears to be a \textit{major ring of energy resistance (fire) }until 
the established condition is met.

\textit{Ring of Elemental Command (Water)• }

\textit{Water walk }(unlimited use)• 

\textit{Create water }(unlimited use)• 

\textit{Water breathing }(unlimited use)• 

\textit{Wall of ice }(once per day)• 

\textit{Ice storm }(twice per week)• 

\textit{Control }water (twice per week)

The ring appears to be a \textit{ring of water walking }until the established condition 
is met.

Strong conjuration; CL 15th; Forge Ring, \textit{summon monster VI}, all appropriate 
spells; Price 200,000 gp.

\textbf{Energy Resistance:} This reddish iron ring continually protects the wearer 
from damage from one type of energy---acid, cold, electricity, fire, or sonic (chosen 
by the creator of the item; determine randomly if found as part of a treasure hoard). 
Each time the wearer would normally take such damage, subtract the ring's resistance 
value from the damage dealt.

A \textit{minor ring of energy resistance }grants 10 points of resistance. A \textit{major 
ring of energy resistance }grants 20 points of resistance. A \textit{greater ring 
of energy resistance }grants 30 points of resistance.

Faint (minor or major) or moderate (greater) abjuration; CL 3rd  (minor), 7th (major), 
or 11th (greater); Forge Ring, \textit{resist energy; }Price 12,000 gp (minor), 
28,000 gp (major), 44,000 gp (greater).

\textbf{Evasion:} This ring continually grants the wearer the ability to avoid 
damage as if she had evasion. Whenever she makes a Reflex saving throw to determine 
whether she takes half damage, a successful save results in no damage.

Moderate transmutation; CL 7th; Forge Ring, \textit{jump; }Price 25,000 gp.

\textbf{Feather Falling:} This ring is crafted with a feather pattern all around 
its edge. It acts exactly like a \textit{feather fall }spell, activated immediately 
if the wearer falls more than 5 feet.

Faint transmutation; CL 1st; Forge Ring, \textit{feather fall; }Price 2,200 gp.

\textbf{Force Shield:} An iron band, this simple ring generates a shield-sized 
(and shield-shaped) \textit{wall of force }that stays with the ring and can be 
wielded by the wearer as if it were a heavy shield (+2 AC). This special creation 
has no armor check penalty or arcane spell failure chance since it is weightless 
and encumbrance-free. It can be activated and deactivated at will as a free action.

Moderate evocation; CL 9th; Forge Ring, \textit{wall of force; }Price 8,500 gp.

\textbf{Freedom of Movement:} This gold ring allows the wearer to act as if continually 
under the effect of a \textit{freedom of movement }spell.

Moderate abjuration; CL 7th; Forge Ring, \textit{freedom of movement; }Price 40,000 
gp.

\textbf{Friend Shield:} These curious rings always come in pairs. A \textit{friend 
shield }ring without its mate is useless. Either wearer of one of a pair of the 
rings can, at any time, command his or her ring to cast a \textit{shield other 
}spell with the wearer of the mated ring as the recipient. This effect has no range 
limitation. 

Moderate abjuration; CL 10th; Forge Ring, \textit{shield other; }Price 50,000 gp 
(for a pair).

\textbf{Invisibility:} By activating this simple silver ring, the wearer can benefit 
from \textit{invisibility}, as the spell.

Faint illusion; CL 3rd; Forge Ring, \textit{invisibility; }Price 20,000 gp.

\textbf{Jumping:} This ring continually allows the wearer to leap about, providing 
a +5 competence bonus on all his Jump checks.

Faint transmutation; CL 2nd; Forge Ring, creator must have 5 ranks in the Jump 
skill\textit{; }Price 2,500 gp.

\textbf{Jumping, Improved:} As jumping, except it grants a +10 competence bonus 
on its wearer's Jump check.

Moderate transmutation; CL 7th; Forge Ring, creator must have 10 ranks in the Jump 
skill\textit{; }Price 10,000 gp.

\textbf{Mind Shielding:} This ring is usually of fine workmanship and wrought from 
heavy gold. The wearer is continually immune to \textit{detect thoughts, discern 
lies, }and any attempt to magically discern her alignment\textit{.}

Faint aburation; CL 3rd; Forge Ring, \textit{nondetection; }Price 8,000 gp.

\textbf{Protection:} This ring offers continual magical protection in the form 
of a deflection bonus of +1 to +5 to AC.

Faint abjuration; CL 5th; Forge Ring, \textit{shield of faith, }caster must be 
of a level at least three times greater than the bonus of the ring; Price 2,000 
gp (\textit{ring +1}); 8,000 gp (\textit{ring +2}); 18,000 gp (\textit{ring +3}); 
32,000 gp (\textit{ring +4}); 50,000 gp (\textit{ring +5}).

\textbf{Ram:} The \textit{ring of the ram }is an ornate ring forged of hard metal, 
usually iron or an iron alloy. It has the head of a ram as its device.

The wearer can command the ring to give forth a ramlike force, manifested by a 
vaguely discernible shape that resembles the head of a ram or a goat. This force 
strikes a single target, dealing 1d6 points of damage if 1 charge is expended, 
2d6 points if 2 charges are used, or 3d6 points if 3 charges (the maximum) are 
used. Treat this as a ranged attack with a 50-foot maximum range and no penalties 
for distance. 

The force of the blow is considerable, and those struck by the ring are subject 
to a bull rush if within 30 feet of the ring-wearer. (The ram has Strength 25 and 
is Large.) The ram gains a +1 bonus on the bull rush attempt if 2 charges are expended, 
or +2 if 3 charges are expended.

In addition to its attack mode, the \textit{ring of the ram }also has the power 
to open doors as if it were a character with Strength 25. If 2 charges are expended, 
the effect is equivalent to a character with Strength 27. If 3 charges are expended, 
the effect is that of a character with Strength 29.

A newly created ring has 50 charges. When all the charges are expended, the ring 
becomes a nonmagical item.

Moderate transmutation; CL 9th; Forge Ring, \textit{bull's strength, telekinesis; 
}Price 8,600 gp.

\textbf{Regeneration:} This white gold ring continually allows a living wearer 
to heal 1 point of damage per level every hour rather than every day. (This ability 
cannot be aided by the Heal skill.) Nonlethal damage heals at a rate of 1 point 
of damage per level every 5 minutes. If the wearer loses a limb, an organ, or any 
other body part while wearing this ring, the ring \textit{regenerates }it as the 
spell. In either case, only damage taken while wearing the ring is regenerated.

Strong conjuration; CL 15th; Forge Ring, \textit{regenerate; }Price 90,000 gp.

\textbf{Shooting Stars:} This ring has two modes of operation, one for being in 
shadowy darkness or outdoors at night and a second one when the wearer is underground 
or indoors at night.

During the night under the open sky or in areas of shadow or darkness, the \textit{ring 
of shooting stars }can perform the following functions on command.• 

\textit{Dancing lights }(once per hour)• 

\textit{Light }(twice per night)• 

\textit{Ball lightning }(special, once per night)• 

\textit{Shooting stars }(special, three per week)

The first special function, \textit{ball lightning, }releases one to four balls 
of lightning (ring wearer's choice). These glowing globes resemble \textit{dancing 
lights}, and the ring wearer controls them in the same fashion (see the \textit{dancing 
lights }spell description). The spheres have a 120-foot range and a duration of 
4 rounds. They can be moved at 120 feet per round. Each sphere is about 3 feet 
in diameter, and any creature who comes within 5 feet of one causes its charge 
to dissipate, taking electricity damage in the process according to the number 
of balls created.

\begin{tabular}{|>{\raggedright}p{68pt}|>{\raggedright}p{106pt}|}
\hline
N\textbf{umber of Balls} & D\textbf{amage per Ball}\tabularnewline
\hline
4 lightning balls & 1d6 points of damage each\tabularnewline
\hline
3 lightning balls & 2d6 points of damage each\tabularnewline
\hline
2 lightning balls & 3d6 points of damage each\tabularnewline
\hline
1 lightning ball & 4d6 points of damage\tabularnewline
\hline
\end{tabular}

Once the \textit{ball lightning }function is activated, the balls can be released 
at any time before the sun rises. (Multiple balls can be released in the same round.)

The second special function produces three \textit{shooting stars }that can be 
released from the ring each week, simultaneously or one at a time. They impact 
for 12 points of damage and spread (as a \textit{fireball}) in a 5-foot-radius 
sphere for 24 points of fire damage.

Any creature struck by a \textit{shooting star }takes full damage from impact plus 
full fire damage from the spread unless it makes a DC 13 Reflex save. Creatures 
not struck but within the spread ignore the impact damage and take only half damage 
from the fire spread on a successful DC 13 Reflex save. Range is 70 feet, at the 
end of which the \textit{shooting star }explodes, unless it strikes a creature 
or object before that. A \textit{shooting star }always follows a straight line, 
and any creature in its path must make a save or be hit by the projectile.

Indoors at night, or underground, the \textit{ring of shooting stars }has the following 
properties.• 

\textit{Faerie fire }(twice per day)• 

\textit{Spark shower }(special, once per day)

The \textit{spark shower }is a flying cloud of sizzling purple sparks that fan 
out from the ring for a distance of 20 feet in an arc 10 feet wide. Creatures within 
this area take 2d8 points of damage each if not wearing metal armor or carrying 
a metal weapon. Those wearing metal armor and/or carrying a metal weapon take 4d8 
points of damage.

Strong evocation; CL 12th; Forge Ring, \textit{light, faerie fire, fireball, lightning 
bolt; }Price 50,000 gp.

\textbf{Spell Storing, Minor:} A \textit{minor ring of spell storing }contains 
up to three levels of spells that the wearer can cast. Each spell has a caster 
level equal to the minimum level needed to cast that spell. The user need not provide 
any material components or focus, or pay an XP cost to cast the spell, and there 
is no arcane spell failure chance for wearing armor (because the ring wearer need 
not gesture). The activation time for the ring is same as the casting time for 
the relevant spell, with a minimum of 1 standard action.

For a randomly generated ring, treat it as a scroll to determine what spells are 
stored in it. If you roll a spell that would put the ring over the three-level 
limit, ignore that roll; the ring has no more spells in it. (Not every newly discovered 
ring need be fully charged.)

A spellcaster can cast any spells into the ring, so long as the total spell levels 
do not add up to more than three. Metamagic versions of spells take up storage 
space equal to their spell level modified by the metamagic feat. A spellcaster 
can use a scroll to put a spell into the \textit{minor ring of spell storing.}

The ring magically imparts to the wearer the names of all spells currently stored 
within it.

Faint evocation; CL 5th; Forge Ring, \textit{imbue with spell ability; }Price 18,000 
gp.

\textbf{Spell Storing:} As the \textit{minor ring of spell storing, }except it 
holds up to five levels of spells.

Moderate evocation; CL 9th; Forge Ring, \textit{imbue with spell ability; }Price 
50,000 gp.

\textbf{Spell Storing, Major:} As the \textit{minor ring of spell storing, }except 
it holds up to ten levels of spells.

Strong evocation; CL 17th; Forge Ring, \textit{imbue with spell ability; }Price 
200,000 gp.

\textbf{Spell Turning:} Up to three times per day on command, this simple platinum 
band automatically reflects the next nine levels of spells cast at the wearer, 
exactly as if \textit{spell turning }had been cast upon the wearer.

Strong abjuration; CL 13th; Forge Ring, \textit{spell turning; }Price 98,280 gp.

\textbf{Sustenance: }This ring continually provides its wearer with life-sustaining 
nourishment. The ring also refreshes the body and mind, so that its wearer needs 
only sleep 2 hours per day to gain the benefit of 8 hours of sleep. The ring must 
be worn for a full week before it begins to work. If it is removed, the owner must 
wear it for another week to reattune it to himself.

Faint conjuration; CL 5th; Forge Ring, \textit{create food and water; }Price 2,500 
gp.

\textbf{Swimming:} This silver ring has a wave pattern etched into the band. It 
continually grants the wearer a +5 competence bonus on Swim checks.

Faint transmutation; CL 2nd; Forge Ring, creator must have 5 ranks in the Swim 
skill; Price 2,500 gp.

\textbf{Swimming, Improved:} As swimming, except it grants a +10 competence bonus 
on its wearer's Swim checks.

Moderate transmutation; CL 7th; Forge Ring, creator must have 10 ranks in the Swim 
skill\textit{; }Price 10,000 gp.

\textbf{Telekinesis: }This ring allows the caster to use the spell \textit{telekinesis 
}on command.

Moderate transmutation; CL 9th; Forge Ring, \textit{telekinesis; }Price 75,000 
gp.

\textbf{Three Wishes:} This ring is set with three rubies. Each ruby stores a \textit{wish 
}spell, activated by the ring. When a \textit{wish }is used, that ruby disappears. 
For a randomly generated ring, roll 1d3 to determine the remaining number of rubies. 
When all the \textit{wishes }are used, the ring becomes a nonmagical item.

Strong evocation (if \textit{miracle }is used); CL 20th; Forge Ring, \textit{wish 
}or \textit{miracle; }Price 97,950 gp; Cost 11,475 gp + 15,918 XP.

\textbf{Water Walking: }This ring, set with an opal, allows the wearer to continually 
utilize the effects of the spell \textit{water walk}.

Moderate transmutation; CL 9th; Forge Ring, \textit{water walk; }Price 15,000 gp.

\textbf{Wizardry:} This special ring comes in four kinds (\textit{ring of wizardry 
I, ring of wizardry II, ring of wizardry III, }and \textit{ring of wizardry IV}), 
all of them useful only to arcane spellcasters. The wearer's arcane spells per 
day are doubled for one specific spell level. A \textit{ring of wizardry I }doubles 
1st-level spells, a \textit{ring of wizardry II }doubles 2nd-level spells, a \textit{ring 
of wizardry III }doubles 3rd-level spells, and a \textit{ring of wizardry IV }doubles 
4th-level spells. Bonus spells from high ability scores or school specialization 
are not doubled.

Moderate (\textit{wizardry I}) or strong (\textit{wizardry II-IV}) (no school); 
CL 11th (\textit{I}), 14th (\textit{II}), 17th (\textit{III}), 20th (\textit{IV}); 
Forge Ring, \textit{limited wish}; Price 20,000 gp (\textit{I}), 40,000 gp (\textit{II}), 
70,000 gp (\textit{III}), 100,000 gp (\textit{IV}).

\textbf{X-Ray Vision:} On command, this ring gives its possessor the ability to 
see into and through solid matter. Vision range is 20 feet, with the viewer seeing 
as if he were looking at something in normal light even if there is no illumination. 
X-ray vision can penetrate 1 foot of stone, 1 inch of common metal, or up to 3 
feet of wood or dirt. Thicker substances or a thin sheet of lead blocks the vision.

Using the ring is physically exhausting, causing the wearer 1 point of Constitution 
damage per minute after the first 10 minutes of use in a single day.

Moderate divination; CL 6th; Forge Ring, \textit{true seeing; }Price 25,000 gp.

\vspace{12pt}
RODS

Rods are scepterlike devices that have unique magical powers and do not usually 
have charges. Anyone can use a rod.

\textbf{Physical Description:} Rods weigh approximately 5 pounds.

They range from 2 feet to 3 feet long and are usually made of iron or some other 
metal. (Many, as noted in their descriptions, can function as light maces or clubs 
due to their sturdy construction.)

These sturdy items have AC 9, 10 hit points, hardness 10, and a break DC of 27.

\textbf{Activation:} Details relating to rod use vary from item to item. See the 
individual descriptions for specifics.

\textbf{Special Qualities:} Roll d\%. A 01 result indicates the rod is intelligent, 
02-31 indicates that something (a design, inscription, or the like) provides a 
clue to its function, and 32-100 indicates no special qualities. Intelligent items 
have extra abilities and sometimes extraordinary powers and special purposes.

Rods with charges can never be intelligent.

\begin{tabular}{|>{\raggedright}p{40pt}|>{\raggedright}p{31pt}|>{\raggedright}p{127pt}|>{\raggedright}p{60pt}|}
\hline
\multicolumn{4}{|p{259pt}|}{T\textbf{able: Rods}}\tabularnewline
\hline
M\textbf{edium} & M\textbf{ajor} & R\textbf{od} & \subsection*{M{\small{}\textbf{arket 
Price}}}\tabularnewline
\hline
01-07--- &  & Metamagic, \textit{Enlarge, lesser} & 3,000 gp\tabularnewline
\hline
08-14--- &  & Metamagic, \textit{Extend, lesser} & 3,000 gp\tabularnewline
\hline
15-21--- &  & Metamagic, \textit{Silent, lesser} & 3,000 gp\tabularnewline
\hline
22-28--- &  & I\textit{mmovable} & 5,000 gp\tabularnewline
\hline
29-35--- &  & Metamagic, \textit{Empower, lesser} & 9,000 gp\tabularnewline
\hline
36-42--- &  & M\textit{etal and mineral detection} & 10,500 gp\tabularnewline
\hline
43-53 & 01-04 & C\textit{ancellation} & 11,000 gp\tabularnewline
\hline
54-57 & 05-06 & Metamagic, \textit{Enlarge} & 11,000 gp\tabularnewline
\hline
58-61 & 07-08 & Metamagic, \textit{Extend} & 11,000 gp\tabularnewline
\hline
62-65 & 09-10 & Metamagic, \textit{Silent} & 11,000 gp\tabularnewline
\hline
66-71 & 11-14 & W\textit{onder} & 12,000 gp\tabularnewline
\hline
72-79 & 15-18 & P\textit{ython} & 13,000 gp\tabularnewline
\hline
80-83--- &  & Metamagic, \textit{Maximize, lesser} & 14,000 gp\tabularnewline
\hline
84-89 & 19-21 & F\textit{lame extinguishing} & 15,000 gp\tabularnewline
\hline
90-97 & 22-25 & V\textit{iper} & 19,000 gp\tabularnewline
\hline
--- & 26-30 & E\textit{nemy detection} & 23,500 gp\tabularnewline
\hline
--- & 31-36 & Metamagic, \textit{Enlarge, greater} & 24,500 gp\tabularnewline
\hline
--- & 37-42 & Metamagic, \textit{Extend, greater} & 24,500 gp\tabularnewline
\hline
--- & 43-48 & Metamagic, \textit{Silent, greater} & 24,500 gp\tabularnewline
\hline
--- & 49-53 & S\textit{plendor} & 25,000 gp\tabularnewline
\hline
--- & 54-58 & W\textit{ithering} & 25,000 gp\tabularnewline
\hline
98-99 & 59-64 & Metamagic, \textit{Empower} & 32,500 gp\tabularnewline
\hline
--- & 65-69 & T\textit{hunder and lightning} & 33,000 gp\tabularnewline
\hline
100 & 70-73 & Metamagic, \textit{Quicken, lesser} & 35,000 gp\tabularnewline
\hline
--- & 74-77 & N\textit{egation} & 37,000 gp\tabularnewline
\hline
--- & 78-80 & A\textit{bsorption} & 50,000 gp\tabularnewline
\hline
--- & 81-84 & F\textit{lailing} & 50,000 gp\tabularnewline
\hline
--- & 85-86 & Metamagic, \textit{Maximize} & 54,000 gp\tabularnewline
\hline
--- & 87-88 & R\textit{ulership} & 60,000 gp\tabularnewline
\hline
--- & 89-90 & S\textit{ecurity} & 61,000 gp\tabularnewline
\hline
--- & 91-92 & L\textit{ordly might} & 70,000 gp\tabularnewline
\hline
--- & 93-94 & Metamagic, \textit{Empower, greater} & 73,000 gp\tabularnewline
\hline
--- & 95-96 & Metamagic, \textit{Quicken} & 75,500 gp\tabularnewline
\hline
--- & 97-98 & A\textit{lertness} & 85,000 gp\tabularnewline
\hline
--- & 99 & Metamagic, \textit{Maximize, greater} & 121,500 gp\tabularnewline
\hline
--- & 100 & Metamagic, \textit{Quicken, greater} & 170,000 gp\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
Rod Descriptions

Although all rods are generally scepterlike, their configurations and abilities 
run the magical gamut. Standard rods are described below.

\textbf{Absorption:} This rod acts as a magnet, drawing spells or spell-like abilities 
into itself. The magic absorbed must be a single-target spell or a ray directed 
at either the character possessing the rod or her gear. The rod then nullifies 
the spell's effect and stores its potential until the wielder releases this energy 
in the form of spells of her own. She can instantly detect a spell's level as the 
rod absorbs that spell's energy. Absorption requires no action on the part of the 
user if the rod is in hand at the time.

A running total of absorbed (and used) spell levels should be kept. The wielder 
of the rod can use captured spell energy to cast any spell she has prepared, without 
expending the preparation itself. The only restrictions are that the levels of 
spell energy stored in the rod must be equal to or greater than the level of the 
spell the wielder wants to cast, that any material components required for the 
spell be present, and that the rod be in hand when casting. For casters such as 
bards or sorcerers who do not prepare spells, the rod's energy can be used to cast 
any spell of the appropriate level or levels that they know.

A \textit{rod of absorption }absorbs a maximum of fifty spell levels and can thereafter 
only discharge any remaining potential it might have. The rod cannot be recharged. 
The wielder knows the rod's remaining absorbing potential and current amount of 
stored energy.

To determine the absorption potential remaining in a newly found rod, roll d\% 
and divide the result by 2. Then roll d\% again: On a result of 71-100, half the 
levels already absorbed by the rod are still stored within.

Strong abjuration; CL 15th;Craft Rod, \textit{spell turning; }Price 50,000 gp.

\textbf{Alertness:} This rod is indistinguishable from a \textit{+1 light mace. 
}It has eight flanges on its macelike head. The rod bestows a +1 insight bonus 
on initiative checks. If grasped firmly, the rod enables the holder to use \textit{detect 
evil, detect good, detect chaos, detect law}, \textit{detect magic, discern lies, 
light, }or \textit{see invisibility. }Each different use is a standard action.

If the head of a \textit{rod of alertness }is planted in the ground, and the possessor 
wills it to alertness (a standard action), the rod senses any creature within 120 
feet who intends to harm the possessor. At the same time, the rod creates the effect 
of a \textit{prayer }spell upon all creatures friendly to the possessor in a 20-foot 
radius. Immediately thereafter, the rod sends forth a mental alert to these friendly 
creatures, warning them of possible danger from the unfriendly creature or creatures 
within the 120-foot radius. These effects last for 10 minutes, and the rod can 
perform this function once per day. Last, the rod can be used to simulate the casting 
of an \textit{animate objects }spell, utilizing any eleven (or fewer) Small objects 
located roughly around the perimeter of a 5-foot-radius circle centered on the 
rod when planted in the ground. Objects remain animated for 11 rounds. The rod 
can perform this function once per day.

Moderate abjuration, divination, enchantment, and evocation; CL 11th; Craft Rod, 
\textit{alarm, detect chaos, detect evil, detect good, detect law, detect magic, 
discern lies, light, see invisibility, prayer, animate objects; }Price 85,000 gp.

\textbf{Cancellation:} This dreaded rod is a bane to magic items, for its touch 
drains an item of all magical properties. The item touched must make a DC 23 Will 
save to prevent the rod from draining it. If a creature is holding it at the time, 
then the item can use the holder's Will save bonus in place of its own if the holder's 
is better. In such cases, contact is made by making a melee touch attack roll. 
Upon draining an item, the rod itself becomes brittle and cannot be used again. 
Drained items are only restorable by \textit{wis}h or \textit{miracle. }(If a \textit{sphere 
of annihilation }and a \textit{rod of cancellation }negate each other, nothing 
can restore either of them.)

Strong abjuration; CL 17th; Craft Rod, \textit{mage's disjunction; }Price 11,000 
gp.

\textbf{Enemy Detection: }This device pulses in the wielder's hand and points in 
the direction of any creature or creatures hostile to the bearer of the device 
(nearest ones first). These creatures can be invisible, ethereal, hidden, disguised, 
or in plain sight. Detection range is 60 feet. If the bearer of the rod concentrates 
for a full round, the rod pinpoints the location of the nearest enemy and indicates 
how many enemies are within range. The rod can be used three times each day, each 
use lasting up to 10 minutes. Activating the rod is a standard action.

Moderate divination; CL 10th; Craft Rod, \textit{true seeing; }Price 23,500 gp.

\textbf{Flailing: }Upon the command of its possessor, the rod activates, changing 
from a normal-seeming rod to a \textit{+3 dire flail}. The dire flail is a double 
weapon, which means that each of the weapon's heads can be used to attack. The 
wielder can gain an extra attack (with the second head) at the cost of making all 
attacks at a -2 penalty (as if she had the Two-Weapon Fighting feat).

Once per day the wielder can use a free action to cause the rod to grant her a 
+4 deflection bonus to Armor Class and a +4 resistance bonus on saving throws for 
10 minutes. The rod need not be in weapon form to grant this benefit.

Transforming it into a weapon or back into a rod is a move action.

Moderate enchantment; CL 9th; Craft Rod, Craft Magic Arms and Armor, \textit{bless; 
}Price 50,000 gp.

\textbf{Flame Extinguishing: }This rod can extinguish Medium or smaller nonmagical 
fires with simply a touch (a standard action). For the rod to be effective against 
other sorts of fires, the wielder must expend 1 or more of the rod's charges.

Extinguishing a Large or larger nonmagical fire, or a magic fire of Medium or smaller 
(such as that of a flaming weapon or a \textit{burning hands }spell), expends 1 
charge. Continual magic flames, such as those of a weapon or a fire creature, are 
suppressed for 6 rounds and flare up again after that time. To extinguish an instantaneous 
fire spell, the rod must be within the area of the effect and the wielder must 
have used a ready action, effectively countering the entire spell.

When applied to Large or larger magic fires, such as those caused by \textit{fireball, 
flame strike, }or \textit{wall of fire}, extinguishing the flames expends 2 charges 
from the rod.

If the device is used upon a fire creature (a melee touch attack), it deals 6d6 
points of damage to the creature. This use requires 3 charges.

A \textit{rod of flame extinguishing }has 10 charges when found. Spent charges 
are renewed every day, so that a wielder can expend up to 10 charges in any 24-hour 
period.

Strong transmutation; CL 12th; Craft Rod, \textit{pyrotechnics; }Price 15,000 gp.

\textbf{Immovable Rod:} This rod is a flat iron bar with a small button on one 
end. When the button is pushed (a move action), the rod does not move from where 
it is, even if staying in place defies gravity. Thus, the owner can lift or place 
the rod wherever he wishes, push the button, and let go. Several \textit{immovable 
rods }can even make a ladder when used together (although only two are needed). 
An \textit{immovable rod }can support up to 8,000 pounds before falling to the 
ground. If a creature pushes against an \textit{immovable rod, }it must make a 
DC 30 Strength check to move the rod up to 10 feet in a single round.

Moderate transmutation; CL 10th; Craft Rod, \textit{levitate; }Price 5,000 gp.

\textbf{Lordly Might: }This rod has functions that are spell-like, and it can also 
be used as a magic weapon of various sorts. It also has several more mundane uses. 
The \textit{rod of lordly might }is metal, thicker than other rods, with a flanged 
ball at one end and six studlike buttons along its length. (Pushing any of the 
rod's buttons is equivalent to drawing a weapon.) It weighs 10 pounds.

The following spell-like functions of the rod can each be used once per day.• 

\textit{Hold person }upon touch, if the wielder so commands (Will DC 14 negates). 
The wielder must choose to use this power and then succeed on a melee touch attack 
to activate the power. If the attack fails, the effect is lost.• 

\textit{Fear }upon all enemies viewing it, if the wielder so desires (10-foot maximum 
range, Will DC 16 partial). Invoking this power is a standard action.• 

\parindent=3pt
Deal 2d4 hit points of damage to an opponent on a successful touch attack (Will 
DC 17 half ) and cure the wielder of a like amount of damage. The wielder must 
choose to use this power before attacking, as with \textit{hold person.}

\parindent=0pt
The following weapon functions of the rod have no limit on the number of times 
they can be employed.• 

\parindent=3pt
In its normal form, the rod can be used as a \textit{+2 light mace.• }

When button 1 is pushed, the rod becomes a \textit{+1 flaming longsword. }A blade 
springs from the ball, with the ball itself becoming the sword's hilt. The weapon 
lengthens to an overall length of 4 feet.• 

When button 2 is pushed, the rod becomes a \textit{+4 battleaxe. }A wide blade 
springs forth at the ball, and the whole lengthens to 4 feet.• 

When button 3 is pushed, the rod becomes a \textit{+3 shortspear }or \textit{+3 
longspear. }The spear blade springs forth, and the handle can be lengthened up 
to 12 feet (wielder's choice), for an overall length of from 6 feet to 15 feet. 
At its 15-foot length, the rod is suitable for use as a lance.

\parindent=0pt
The following other functions of the rod also have no limit on the number of times 
they can be employed.• 

\parindent=3pt
Climbing pole/ladder. When button 4 is pushed, a spike that can anchor in granite 
is extruded from the ball, while the other end sprouts three sharp hooks. The rod 
lengthens to anywhere between 5 and 50 feet in a single round, stopping when button 
4 is pushed again. Horizontal bars three inches long fold out from the sides, 1 
foot apart, in staggered progression. The rod is firmly held by the spike and hooks 
and can bear up to 4,000 pounds. The wielder can retract the pole by pushing button 
5.• 

The ladder function can be used to force open doors. The wielder plants the rod's 
base 30 feet or less from the portal to be forced and in line with it, then pushes 
button 4. The force exerted has a Strength modifier of +12.• 

\parindent=7pt
When button 6 is pushed, the rod indicates magnetic north and gives the wielder 
a knowledge of his approximate depth beneath the surface or height above it.

\parindent=0pt
Strong enchantment, evocation, necromancy, and transmutation; CL 19th; Craft Rod, 
Craft Magic Arms and Armor, \textit{inflict light wounds, bull's strength, flame 
blade, hold person, fear; }Price 70,000 gp.

\textbf{Metal and Mineral Detection:} This rod pulses in the wielder's hand and 
points to the largest mass of metal within 30 feet. However, the wielder can concentrate 
on a specific metal or mineral. If the specific mineral is within 30 feet, the 
rod points to any places it is located, and the rod wielder knows the approximate 
quantity as well. If more than one deposit of the specified metal or mineral is 
within range, the rod points to the largest cache first. Each operation requires 
a full-round action.

Moderate divination; CL 9th; Craft Rod, \textit{locate object; }Price 10,500 gp.

\textbf{Metamagic Rods:} Metamagic rods hold the essence of a metamagic feat but 
do not change the spell slot of the altered spell. All the rods described here 
are use-activated (but casting spells in a threatened area still draws an attack 
of opportunity). A caster may only use one metamagic rod on any given spell, but 
it is permissible to combine a rod with metamagic feats possessed by the rod's 
wielder. In this case, only the feats possessed by the wielder adjust the spell 
slot of the spell being cast.

Possession of a metamagic rod does not confer the associated feat on the owner, 
only the ability to use the given feat a specified number of times per day. A sorcerer 
still must take a full-round action when using a metamagic rod, just as if using 
a metamagic feat he possesses.

\textit{Lesser and Greater Metamagic Rods: }Normal metamagic rods can be used with 
spells of 6th level or lower. Lesser rods can be used with spells of 3rd level 
or lower, while greater rods can be used with spells of 9th level or lower.

\textbf{Metamagic, Empower: }The wielder can cast up to three spells per day that 
are empowered as though using the Empower Spell feat.

Strong (no school); CL 17th; Craft Rod, Empower Spell; Price 9,000 gp (lesser), 
32,500 gp (normal), 73,000 gp (greater).

\textbf{Metamagic, Enlarge:} The wielder can cast up to three spells per day that 
are enlarged as though using the Enlarge Spell feat.

Strong (no school); CL 17th; Craft Rod, Enlarge Spell; Price 3,000 gp (lesser), 
11,000 gp (normal), 24,500 gp (greater).

\textbf{Metamagic, Extend: }The wielder can cast up to three spells per day that 
are extended as though using the Extend Spell feat.

Strong (no school); CL 17th; Craft Rod, Extend Spell; Price 3,000 gp (lesser), 
11,000 gp (normal), 24,500 gp (greater).

\textbf{Metamagic, Maximize:} The wielder can cast up to three spells per day that 
are maximized as though using the Maximize Spell feat.

Strong (no school); CL 17th; Craft Rod, Maximize Spell feat; Price 14,000 gp (lesser), 
54,000 gp (normal), 121,500 gp (greater).

\textbf{Metamagic, Quicken:} The wielder can cast up to three spells per day that 
are quickened as though using the Quicken Spell feat.

Strong (no school); CL 17th; Craft Rod, Quicken Spell; Price 35,000 gp (lesser), 
75,500 gp (normal), 170,000 gp (greater).

\textbf{Metamagic, Silent:} The wielder can cast up to three spells per day without 
verbal components as though using the Silent Spell feat.

Strong (no school); CL 17th; Craft Rod, Silent Spell; Price 3,000 gp (lesser), 
11,000 gp (normal), 24,500 gp (greater).

\textbf{Negation: }This device negates the spell or spell-like function or functions 
of magic items. The wielder points the rod at the magic item, and a pale gray beam 
shoots forth to touch the target device, attacking as a ray (a ranged touch attack). 
The ray functions as a \textit{greater dispel magic }spell, except it only affects 
magic items. To negate instantaneous effects from an item, the rod wielder needs 
to have used a ready action. The dispel check uses the rod's caster level (15th). 
The target item gets no saving throw, although the rod can't negate artifacts (even 
minor artifacts). The rod can function three times per day.

Strong varied; CL 15th; Craft Rod, \textit{dispel magic, }and \textit{limited wish 
}or \textit{miracle; }Price 37,000 gp.

\textbf{Python: }This rod is longer than normal rods. It is about 4 feet long and 
weighs 10 pounds. It strikes as a \textit{+1/+1 quarterstaff}. If the user throws 
the rod to the ground (a standard action), it grows to become a giant constrictor 
snake by the end of the round. The python obeys all commands of the owner. (In 
animal form, it retains the +1 enhancement bonus on attacks and damage possessed 
by the rod form.) The serpent returns to rod form (a full-round action) whenever 
the wielder desires, or whenever it moves farther than 100 feet from the owner. 
If the snake form is slain, it returns to rod form and cannot be activated again 
for three days. A \textit{python rod }only functions if the possessor is good.

Moderate transmutation; CL 10th; Craft Rod, Craft Magic Arms and Armor, \textit{baleful 
polymorph, }creator must be good; Price 13,000 gp.

\textbf{Rulership:} This rod looks like a royal scepter worth at least 5,000 gp 
in materials and workmanship alone. The wielder can command the obedience and fealty 
of creatures within 120 feet when she activates the device (a standard action). 
Creatures totaling 300 Hit Dice can be ruled, but creatures with Intelligence scores 
of 12 or higher are entitled to a DC 16 Will save to negate the effect. Ruled creatures 
obey the wielder as if she were their absolute sovereign. Still, if the wielder 
gives a command that is contrary to the nature of the creatures commanded, the 
magic is broken. The rod can be used for 500 total minutes before crumbling to 
dust. This duration need not be continuous.

Strong enchantment; CL 20th; Craft Rod, \textit{mass charm monster; }Price 60,000 
gp; Cost 27,500 gp + 2,200 XP.

\textbf{Security: }This item creates a nondimensional space, a pocket paradise. 
There the rod's possessor and as many as 199 other creatures can stay in complete 
safety for a period of time, up to 200 days divided by the number of creatures 
affected. All fractions are rounded down.

In this pocket paradise, creatures don't age, and natural healing take place at 
twice the normal rate. Fresh water and food (fruits and vegetables only) are in 
abundance. The climate is comfortable for all creatures involved.

Activating the rod (a standard action) causes the wielder and all creatures touching 
the rod to be transported instantaneously to the paradise. Members of large groups 
can hold hands or otherwise maintain physical contact, allowing all connected creatures 
in a circle or a chain to be affected by the rod. Unwilling creatures get a DC 
17 Will save to negate the effect. If such a creature succeeds on its save, other 
creatures beyond that point in a chain can still be affected by the rod.

When the rod's effect expires or is dispelled, all the affected creatures instantly 
reappear in the location they occupied when the rod was activated. If something 
else occupies the space that a traveler would be returning to, then his body is 
displaced a sufficient distance to provide the space required for reentry. The 
rod's possessor can dismiss the effect whenever he wishes before the maximum time 
period expires, but the rod can only be activated once per week.

Strong conjuuration; CL 20th; Craft Rod, \textit{gate; }Price 61,000 gp.

\textbf{Splendor: }The possessor of this rod gains a +4 enhancement bonus to her 
Charisma score for as long as she holds or carries the item. Once per day, the 
rod creates and garbs her in clothing of the finest fabrics, plus adornments of 
furs and jewels. 

Apparel created by the magic of the rod remains in existence for 12 hours. However, 
if the possessor attempts to sell or give away any part of it, to use it for a 
spell component, or the like, all the apparel immediately disappears. The same 
applies if any of it is forcibly taken from her.

The value of noble garb created by the rod ranges from 7,000 to 10,000 gp (1d4+6 
x$ $1,000 gp)---1,000 gp for the fabric alone, 5,000 gp for the furs, and the rest 
for the jewel trim (maximum of twenty gems, maximum value 200 gp each).

In addition, the rod has a second special power, usable once per week. Upon command, 
it creates a palatial tent---a huge pavilion of silk 60 feet across. Inside the 
tent are temporary furnishings and food suitable to the splendor of the pavilion 
and sufficient to entertain as many as one hundred persons. The tent and its trappings 
last for one day. At the end of that time, the tent and all objects associated 
with it (including any items that were taken out of the tent) disappear.

Strong conjuration and transmutation; CL 12th; Craft Rod, \textit{eagle's splendor, 
fabricate, major creation; }Price 25,000 gp.

\textbf{Thunder and Lightning: }Constructed of iron set with silver rivets, this 
rod has the properties of a \textit{+2 light mace}. Its other magical powers are 
as follows.• 

\textit{Thunder: }Once per day, the rod can strike as a \textit{+3 light mace, 
}and the opponent struck is stunned from the noise of the rod's impact (Fortitude 
DC 16 negates). Activating this power counts as a free action, and it works if 
the wielder strikes an opponent within 1 round.• 

\textit{Lightning: }Once per day, when the wielder desires, a short spark of electricity 
can leap forth when the rod strikes an opponent to deal the normal damage for a 
\textit{+2 light mace }(1d6+2) and an extra 2d6 points of electricity damage. Even 
when the rod might not score a normal hit in combat, if the roll was good enough 
to count as a successful melee touch attack hit, then the 2d6 points of electricity 
damage still applies. The wielder activates this power as a free action, and it 
works if he strikes an opponent within 1 round.• 

\textit{Thunderclap: }Once per day as a standard action, the wielder can cause 
the rod to give out a deafening noise, just as a \textit{shout }spell (Fortitude 
DC 16 partial, 2d6 points of sonic damage, target deafened for 2d6 rounds).• 

\textit{Lightning Stroke: }Once per day as a standard action, the wielder can cause 
the rod to shoot out a 5-foot-wide lightning bolt (9d6 points of electricity damage, 
Reflex DC 16 half ) to a range of 200 feet.• 

\textit{Thunder and Lightning: }Once per week as a standard action, the wielder 
of the rod can combine the \textit{thunderclap }described above with a lightning 
bolt, as in the \textit{lightning stroke. }The thunderclap affects all within 10 
feet of the bolt. The \textit{lightning stroke }deals 9d6 points of electricity 
damage (count rolls of 1 or 2 as rolls of 3, for a range of 27 to 54 points), and 
the \textit{thunderclap }deals 2d6 points of sonic damage. A single DC 16 Reflex 
save applies for both effects.

Moderate evocation; CL 9th; Craft Rod, Craft Magic Arms and Armor, \textit{lightning 
bolt}, \textit{shout}; Price 33,000 gp.

\textbf{Viper:} This rod strikes as a \textit{+2 heavy mace}. Once per day, upon 
command, the head of the rod becomes that of an actual serpent for 10 minutes. 
During this period, any successful strike with the rod deals its usual damage and 
also poisons the creature hit. The poison deals 1d10 points of Constitution damage 
immediately (Fortitude DC 14 negates) and another 1d10 points of Constitution damage 
1 minute later (Fortitude DC 14 negates). The rod only functions if its possessor 
is evil.

Moderate necromancy; CL 10th; Craft Rod, Craft Magic Arms and Armor, \textit{poison, 
}creator must be evil; Price 19,000 gp.

\textbf{Withering:} A \textit{rod of withering }acts as a \textit{+1 light mace 
}that deals no hit point damage. Instead, the wielder deals 1d4 points of Strength 
damage and 1d4 points of Constitution damage to any creature she touches with the 
rod (by making a melee touch attack). If she scores a critical hit, the damage 
from that hit is permanent ability drain. In either case, the defender negates 
the effect with a DC 17 Fortitude save.

Strong necromancy; CL 13th; Craft Rod, Craft Magic Arms and Armor, \textit{contagion; 
}Price 25,000 gp.

\textbf{Wonder:} A \textit{rod of wonder }is a strange and unpredictable device 
that randomly generates any number of weird effects each time it is used. (Activating 
the rod is a standard action.) Typical powers of the rod include the following.

\begin{tabular}{|>{\raggedright}p{24pt}|>{\raggedright}p{302pt}|}
\hline
d\textbf{\%} & W\textbf{ondrous Effect}\tabularnewline
\hline
01-05 & S\textit{low }creature pointed at for 10 rounds (Will DC 15 negates).\tabularnewline
\hline
06-10 & F\textit{aerie fire }surrounds the target.\tabularnewline
\hline
11-15 & Deludes wielder for 1 round into believing the rod functions as indicated 
by a second die roll (no save).\tabularnewline
\hline
16-20 & G\textit{ust of wind, }but at windstorm force (Fortitude DC 14 negates).\tabularnewline
\hline
21-25 & Wielder learns target's surface thoughts (as with \textit{detect thoughts) 
}for 1d4 rounds (no save).\tabularnewline
\hline
26-30 & S\textit{tinking cloud }at 30-ft. range (Fortitude DC 15 negates).\tabularnewline
\hline
31-33 & Heavy rain falls for 1 round in 60-ft. radius centered on rod wielder.\tabularnewline
\hline
34-36 & S\textit{ummon }an animal---a rhino (01-25 on d\%), elephant (26-50), or 
mouse (51-100).\tabularnewline
\hline
37-46 & L\textit{ightning bolt }(70 ft. long, 5 ft. wide), 6d6 damage (Reflex DC 
15 half).\tabularnewline
\hline
47-49 & Stream of 600 large butterflies pours forth and flutters around for 2 rounds, 
blinding everyone (including wielder) within 25 ft. (Reflex DC 14 negates).\tabularnewline
\hline
50-53 & E\textit{nlarge person }if within 60 ft. of rod (Fortitude DC 13 negates).\tabularnewline
\hline
54-58 & D\textit{arkness}, 30-ft.-diameter hemisphere, centered 30 ft. away from 
rod.\tabularnewline
\hline
59-62 & Grass grows in 160-sq.-ft. area before the rod, or grass existing there 
grows to ten times normal size.\tabularnewline
\hline
63-65 & Turn ethereal any nonliving object of up to 1,000 lb. mass and up to 30 
cu. ft. in size.\tabularnewline
\hline
66-69 & Reduce wielder to 1/12 height (no save).\tabularnewline
\hline
70-79 & F\textit{ireball }at target or 100 ft. straight ahead, 6d6 damage (Reflex 
DC 15 half).\tabularnewline
\hline
80-84 & I\textit{nvisibility }covers rod wielder.\tabularnewline
\hline
85-87 & Leaves grow from target if within 60 ft. of rod. These last 24 hours.\tabularnewline
\hline
88-90 & 10-40 gems, value 1 gp each, shoot forth in a 30-ft.-long stream. Each 
gem deals 1 point of damage to any creature in its path: Roll 5d4 for the number 
of hits and divide them among the available targets.\tabularnewline
\hline
91-95 & Shimmering colors dance and play over a 40-ft.-by-30-ft. area in front 
of rod. Creatures therein are blinded for 1d6 rounds (Fortitude DC 15 negates).\tabularnewline
\hline
96-97 & Wielder (50\% chance) or target (50\% chance) turns permanently blue, green, 
or purple (no save).\tabularnewline
\hline
98-100 & F\textit{lesh to stone }(or \textit{stone to flesh }if target is stone 
already) if target is within 60 ft. (Fortitude DC 18 negates).\tabularnewline
\hline
\end{tabular}

Moderate enchantment; CL 10th; Craft Rod, \textit{confusion}, creator must be chaotic; 
Price 12,000 gp.

\newpage

\end{document}
