%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{array}
\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

This material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

{\LARGE{}EPIC MAGIC ITEMS 1}

\vspace{12pt}
While not truly an artifact, the epic magic item is a creation of such power that 
it surpasses other magic items\textit{. }Epic magic items are objects of great 
power and value. The following are typical characteristics of an epic magic item. 
In general, an item with even one of these characteristics is an epic magic item. 
• 

\parindent=3pt
Grants a bonus on attacks or damage greater than +5. • 

Grants an enhancement bonus to armor higher than +5. • 

\parindent=7pt
Has a special ability with a market price modifier greater than +5. • 

\parindent=3pt
Grants an armor bonus of greater than +10 (not including magic armor's enhancement 
bonus). • 

Grants a natural armor, deflection, or resistance bonus greater than +5. • 

\parindent=7pt
Grants an enhancement bonus to an ability score greater than +6. • 

\parindent=3pt
Grants an enhancement bonus on a skill check greater than +30. • 

Mimics a spell of an effective level higher than 9th. • 

\parindent=7pt
Has a caster level above 20th. • 

\parindent=3pt
Has a market price above 200,000 gp, not including material costs for armor or 
weapons, material component- or experience point-based costs, or additional value 
for intelligent items. 

\parindent=0pt
An epic magic item that grants a bonus beyond those allowed for normal magic items\textit{ 
}has a higher market price than indicated by the formulas for non-epic items. 

Epic magic items are not artifacts. They are not unique, though they are certainly 
very rare, and anyone with the proper item creation feats can build them. Even 
an epic magic item can never grant a dodge bonus, and the maximum inherent bonus 
that can be applied to an ability score is +5. An epic magic item cannot be created 
that uses or mimics an epic spell. A major artifact might be able to mimic such 
a spell, however. 

\vspace{12pt}
\subsection*{CREATING EPIC MAGIC ITEMS }

The process of creating an epic magic item is very similar to creating a nonepic 
magic item. However, certain important differences exist. 

\vspace{12pt}
\subsection*{CASTER LEVEL }

Spells with an effective level of 10th or higher are possible at epic levels. Because 
these spell slots aren't automatically gained at a particular level like 0- to 
9th-level spells are, they don't have a minimum caster level. For this reason, 
the minimum caster level for any spell of 10th level or higher is set at 11 + spell 
level. 

\vspace{12pt}
\subsection*{PREREQUISITES }

In addition to the materials and tools required for nonepic magic items\textit{, 
}any epic magic item requires at least two item creation feats: the epic and nonepic 
version.

\vspace{12pt}
\subsection*{MARKET PRICE }

Use the guidelines for nonepic magic items\textit{ }to determine the market price 
of an epic magic item, with one addition: If the item gives a bonus beyond the 
limit allowed in for normal, nonepic magic items\textit{, }multiply the portion 
of the market price derived from that characteristic by 10. Some epic characteristics, 
such as caster level, don't trigger this multiplier. 

\vspace{12pt}
\subsection*{EXPERIENCE POINT COST }

The experience point cost to create an epic magic item is determined differently 
than for a normal magic item. For all epic magic items other than scrolls, divide 
the market price by 100, then add 10,000 XP to the result. The final number is 
the experience point cost to create the item. 

For epic scrolls, divide the market price by 25 (as normal for creating a nonepic 
scroll), then add 1,000 XP to the result. The final number is the experience point 
cost to create the epic scroll. 

\vspace{12pt}
\subsection*{MAGIC ITEM DESCRIPTIONS }

In the following sections, each general type of magic item, such as armor or scrolls, 
has an overall description, followed by descriptions of specific items, if any. 
Each magic item description and table follows the same format used for nonepic 
magic items\textit{. }Specific exceptions are noted as necessary. 

\vspace{12pt}
{\LARGE{}ARMOR }

Except when otherwise stated, epic magic armor and shields follow the rules for 
nonepic magic armor and shields\textit{.}

There is no limit to the enhancement bonus of epic magic armor or shields, to the 
market price modifier of epic magic armor or shield special abilities, or to the 
total enhancement bonus and market price modifier of epic magic armor or shields. 

\vspace{12pt}
\subsection*{EPIC ARMOR AND SHIELD BASE PRICE }

To find the base price of an epic suit of magic armor or an epic magic shield, 
roll on Table: Armor and Shields. Note that the +6 to +10 rows apply only to armor 
and shields that provide an enhancement bonus of +6 to +10 or armor and shields 
with a single special ability whose market price modifier is +6 to +10. Magic armor 
and shields with a total effective bonus of +6 to +10 but that have an enhancement 
bonus of +5 or less and special abilities whose individual market price modifiers 
are +5 or less use the table for nonepic magic armor\textit{ }to determine price. 

\vspace{12pt}
\textbf{Table: Armor and Shields }

\begin{tabular}{|>{\raggedright}p{28pt}|>{\raggedright}p{144pt}|>{\raggedright}p{92pt}|}
\hline
d\textbf{\%} & E\textbf{nhancement Bonus} & \section*{M\textbf{arket Price}}\tabularnewline
\hline
01 & +1 shield \textsuperscript{\textbf{1}} & +1,000 gp \tabularnewline
\hline
02 & +1 armor \textsuperscript{\textbf{2}} & +1,000 gp \tabularnewline
\hline
03 & +2 shield \textsuperscript{\textbf{1}} & +4,000 gp \tabularnewline
\hline
04 & +2 armor \textsuperscript{\textbf{2}} & +4,000 gp \tabularnewline
\hline
05-06 & +3 shield \textsuperscript{\textbf{1}} & +9,000 gp \tabularnewline
\hline
07-08 & +3 armor \textsuperscript{\textbf{2}} & +9,000 gp \tabularnewline
\hline
09-12 & +4 shield \textsuperscript{\textbf{1}} & +16,000 gp \tabularnewline
\hline
13-16 & +4 armor \textsuperscript{\textbf{2}} & +16,000 gp \tabularnewline
\hline
17-21 & +5 shield \textsuperscript{\textbf{1}} & +25,000 gp\tabularnewline
\hline
22-26 & +5 armor \textsuperscript{\textbf{2}} & +25,000 gp \tabularnewline
\hline
27-30 & +6 shield & +360,000 gp \tabularnewline
\hline
31-34 & +6 armor & +360,000 gp \tabularnewline
\hline
35-38 & +7 shield & +490,000 gp \tabularnewline
\hline
39-42 & +7 armor & +490,000 gp \tabularnewline
\hline
43-45 & +8 shield & +640,000 gp \tabularnewline
\hline
46-48 & +8 armor & +640,000 gp \tabularnewline
\hline
49-50 & +9 shield & +810,000 gp \tabularnewline
\hline
51-52 & +9 armor & +810,000 gp \tabularnewline
\hline
53-54 & +10 shield & +1,000,000 gp \tabularnewline
\hline
55-56 & +10 armor & +1,000,000 gp \tabularnewline
\hline
57-62 & \multicolumn{2}{p{237pt}|}{Specific armor or shield \textsuperscript{\textbf{3 
}}}\tabularnewline
\hline
63-75 & \multicolumn{2}{p{237pt}|}{Armor with special ability and roll again \textsuperscript{\textbf{4}}\textbf{ 
}}\tabularnewline
\hline
76-98 & \multicolumn{2}{p{237pt}|}{Shield with special ability and roll again\textbf{ 
}\textsuperscript{\textbf{4}}\textbf{ }}\tabularnewline
\hline
99 & \multicolumn{2}{p{237pt}|}{Epic shield \textsuperscript{\textbf{5}} }\tabularnewline
\hline
100 & \multicolumn{2}{p{237pt}|}{Epic armor \textsuperscript{\textbf{5}} }\tabularnewline
\hline
\multicolumn{3}{|p{266pt}|}{1 Roll also on nonepic magic item Table: Shield Special 
Abilities. }\tabularnewline
\hline
\multicolumn{3}{|p{266pt}|}{2 Roll also on nonepic magic item Table: Armor Special 
Abilities. }\tabularnewline
\hline
\multicolumn{3}{|p{266pt}|}{3 Roll on Table: Specific Epic Magic Armor and Shields. 
}\tabularnewline
\hline
\multicolumn{3}{|p{266pt}|}{4\textsuperscript{ }Roll on Table: Armor Special Abilities 
or Table: Shield Special Abilities. }\tabularnewline
\hline
\multicolumn{3}{|p{266pt}|}{5 Roll on Table: Epic Armor and Shields. }\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\textbf{Table: Epic Armor and Shields }

\begin{tabular}{|>{\raggedright}p{29pt}|>{\raggedright}p{172pt}|>{\raggedright}p{78pt}|}
\hline
d\textbf{\%} & E\textbf{nhancement Bonus} & \section*{M\textbf{arket Price}}\tabularnewline
\hline
01-21 & +11 & +1,210,000 gp \tabularnewline
\hline
22-39 & +12 & +1,440,000 gp \tabularnewline
\hline
40-54 & +13 & +1,690,000 gp \tabularnewline
\hline
55-66 & +14 & +1,960,000 gp \tabularnewline
\hline
67-76 & +15 & +2,250,000 gp \tabularnewline
\hline
77-84 & +16 & +2,560,000 gp \tabularnewline
\hline
85-90 & +17 & +2,890,000 gp \tabularnewline
\hline
91-94 & +18 & +3,240,000 gp \tabularnewline
\hline
95-97 & +19 & +3,610,000 gp \tabularnewline
\hline
98-99 & +20 & +4,000,000 gp \tabularnewline
\hline
100 & Roll again and add +10 to bonus \textsuperscript{\textbf{1}} & 2\tabularnewline
\hline
\multicolumn{2}{|p{201pt}|}{1 This is cumulative if rolled multiple times. } & \tabularnewline
\hline
\multicolumn{3}{|p{279pt}|}{2 For enhancement bonuses higher than +20, the market 
price modifier is equal to the square of the bonus x10,000 gp. }\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\textbf{Table: Random Armor Type }

\begin{tabular}{|>{\raggedright}p{23pt}|>{\raggedright}p{193pt}|>{\raggedright}p{92pt}|}
\hline
d\textbf{\%} & A\textbf{rmor} & A\textbf{rmor Cost}\textsuperscript{\textbf{ 1}}\tabularnewline
\hline
01 & Padded & +155 gp \tabularnewline
\hline
02 & Leather & +160 gp \tabularnewline
\hline
03-12 & Hide & +165 gp \tabularnewline
\hline
13-27 & Studded leather & +175 gp \tabularnewline
\hline
28-42 & Chain shirt \textsuperscript{\textbf{2}} & +250 gp\tabularnewline
\hline
43 & Scale mail \textsuperscript{\textbf{2}} & +200 gp \tabularnewline
\hline
44 & Chainmail \textsuperscript{\textbf{2}} & +300 gp \tabularnewline
\hline
45-57 & Breastplate \textsuperscript{\textbf{2}} & +350 gp \tabularnewline
\hline
58 & Splint mail \textsuperscript{\textbf{2}} & +350 gp \tabularnewline
\hline
59 & Banded mail \textsuperscript{\textbf{2}} & +400 gp \tabularnewline
\hline
60 & Half-plate \textsuperscript{\textbf{2}} & +750 gp \tabularnewline
\hline
61-100 & Full plate \textsuperscript{\textbf{2}} & +1,650 gp \tabularnewline
\hline
\multicolumn{3}{|p{309pt}|}{All magic armor is masterwork armor (with an armor 
check penalty 1 lower than normal). }\tabularnewline
\hline
\multicolumn{3}{|p{309pt}|}{1 Add to enhancement bonus on Table: Epic Armor and 
Shields to determine total market price. }\tabularnewline
\hline
\multicolumn{3}{|p{309pt}|}{2 Roll d\% to determine material: 01-70 steel; 71-90 
mithral; 91-99 adamantine; 100 other. Adjust price accordingly. }\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\textbf{Table: Random Shield Type }

\begin{tabular}{|>{\raggedright}p{22pt}|>{\raggedright}p{186pt}|>{\raggedright}p{100pt}|}
\hline
d\textbf{\%} & S\textbf{hield} & \section*{S\textbf{hield Cost}\textsuperscript{\textbf{ 
1}}\textbf{ }}\tabularnewline
\hline
01-10 & Buckler & +165 gp \tabularnewline
\hline
11-15 & Shield, light wooden \textsuperscript{\textbf{2}} & +153 gp \tabularnewline
\hline
16-20 & Shield, light steel \textsuperscript{\textbf{3}} & +159 gp \tabularnewline
\hline
21-30 & Shield, heavy wooden \textsuperscript{\textbf{2}} & +157 gp \tabularnewline
\hline
31-95 & Shield, heavy steel \textsuperscript{\textbf{3}} & +170 gp \tabularnewline
\hline
96-100 & Shield, tower \textsuperscript{\textbf{2}} & +180 gp \tabularnewline
\hline
\multicolumn{3}{|p{309pt}|}{All magic shields are masterwork shields (with an armor 
check penalty 1 lower than normal). }\tabularnewline
\hline
\multicolumn{3}{|p{309pt}|}{1 Add to enhancement bonus on Table: Epic Armor and 
Shields to determine total market price. }\tabularnewline
\hline
\multicolumn{3}{|p{309pt}|}{2\textsuperscript{ }Roll d\% to determine material: 
01-70 wood; 71-99 dark-wood; 100 other. Adjust price accordingly. }\tabularnewline
\hline
\multicolumn{3}{|p{309pt}|}{3\textsuperscript{ }Roll d\% to determine material: 
01-70 steel; 71-90 mithral; 91-99 adamantine; 100 other. Adjust price accordingly. 
}\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\textbf{Table: Armor Special Abilities }

\begin{tabular}{|>{\raggedright}p{22pt}|>{\raggedright}p{184pt}|>{\raggedright}p{103pt}|}
\hline
d\textbf{\%} & S\textbf{pecial Ability} & \section*{M\textbf{arket Price Modifier 
}\textsuperscript{\textbf{1}}\textbf{ }}\tabularnewline
\hline
01-06 & Great invulnerability (10/magic) & +4 bonus \tabularnewline
\hline
07-11 & Great invulnerability (15/magic) & +5 bonus \tabularnewline
\hline
12-19 & Negating & +5 bonus \tabularnewline
\hline
20-26 & Acid warding & +6 bonus \tabularnewline
\hline
27-33 & Cold warding & +6 bonus \tabularnewline
\hline
34-40 & Fire warding & +6 bonus \tabularnewline
\hline
41-44 & Great invulnerability (5/epic) & +6 bonus \tabularnewline
\hline
45-50 & Great spell resistance (SR 21) & +6 bonus \tabularnewline
\hline
51-57 & Lightning warding & +6 bonus \tabularnewline
\hline
58-64 & Sonic warding & +6 bonus \tabularnewline
\hline
65-67 & Great invulnerability (10/epic) & +7 bonus \tabularnewline
\hline
68-72 & Great spell resistance (SR 23) & +7 bonus \tabularnewline
\hline
73-76 & Great spell resistance (SR 25) & +8 bonus \tabularnewline
\hline
77-79 & Great spell resistance (SR 27) & +9 bonus \tabularnewline
\hline
80-87 & \multicolumn{2}{p{287pt}|}{Roll on nonepic magic item Table: Armor Special 
Abilities, then roll again on this table. }\tabularnewline
\hline
88-95 & \multicolumn{2}{p{287pt}|}{Roll twice on nonepic magic item Table: Armor 
Special Abilities. }\tabularnewline
\hline
96-100 & \multicolumn{2}{p{287pt}|}{Roll twice again \textsuperscript{\textbf{2}}\textbf{ 
}}\tabularnewline
\hline
\multicolumn{3}{|p{309pt}|}{1\textsuperscript{ }Add to enhancement bonus on Table: 
Epic Armor and Shields to determine total market price. }\tabularnewline
\hline
\multicolumn{3}{|p{309pt}|}{2 If a special ability is rolled twice, only one counts. 
If two versions of the same special ability are rolled, use the better. }\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\textbf{Table: Shield Special Abilities }

\begin{tabular}{|>{\raggedright}p{23pt}|>{\raggedright}p{173pt}|>{\raggedright}p{112pt}|}
\hline
d\textbf{\%} & S\textbf{hield Special Ability} & \section*{M\textbf{arket Price 
Modifier }\textsuperscript{\textbf{1}}\textbf{ }}\tabularnewline
\hline
01-06 & Great invulnerability (10/magic) & +4 bonus \tabularnewline
\hline
07-12 & Great invulnerability (15/magic) & +5 bonus \tabularnewline
\hline
13-19 & Acid warding & +6 bonus \tabularnewline
\hline
20-26 & Cold warding & +6 bonus \tabularnewline
\hline
27-33 & Fire warding & +6 bonus \tabularnewline
\hline
34-37 & Great invulnerability (5/epic) & +6 bonus \tabularnewline
\hline
38-43 & Great spell resistance (SR 21) & +6 bonus \tabularnewline
\hline
44-46 & Infinite arrow deflection & +6 bonus \tabularnewline
\hline
47-53 & Lightning warding & +6 bonus \tabularnewline
\hline
54-60 & Sonic warding & +6 bonus \tabularnewline
\hline
61-63 & Great invulnerability (10/epic) & +7 bonus \tabularnewline
\hline
64-68 & Great spell resistance (SR 23) & +7 bonus \tabularnewline
\hline
69-71 & Exceptional arrow deflection & +8 bonus \tabularnewline
\hline
72-75 & Great spell resistance (SR 25) & +8 bonus \tabularnewline
\hline
76-78 & Great spell resistance (SR 27) & +9 bonus \tabularnewline
\hline
79 & Great reflection & +10 bonus \tabularnewline
\hline
80-87 & \multicolumn{2}{p{286pt}|}{Roll on nonepic magic item Table: Shield Special 
Abilities, then roll again on this table.}\tabularnewline
\hline
88-95 & \multicolumn{2}{p{286pt}|}{Roll twice on nonepic magic item Table: Shield 
Special Abilities.}\tabularnewline
\hline
96-100 & \multicolumn{2}{p{286pt}|}{Roll twice again\textbf{ }\textsuperscript{\textbf{2 
}}}\tabularnewline
\hline
\multicolumn{3}{|p{309pt}|}{1 Add to enhancement bonus on Table: Epic Armor and 
Shields to determine total market price. }\tabularnewline
\hline
\multicolumn{3}{|p{309pt}|}{2 If you roll a special ability twice, only one counts. 
If you roll two versions of the same special ability, use the better. }\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\subsection*{EPIC ARMOR AND SHIELD SPECIAL ABILITY DESCRIPTIONS }

Most magic armor and shields only have enhancement bonuses. Such items can also 
have special abilities, such as those detailed below and nonepic abilities\textit{. 
}Armor or a shield with a special ability must have at least a +1 enhancement bonus. 

\vspace{12pt}
\textbf{Acid Warding: }The armor absorbs the first 50 points of acid damage per 
round that the wearer would normally take (similar to the \textit{resist energy 
}spell). 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{protection from energy; Market Price: 
}+6 bonus. 

\textbf{Cold Warding: }The armor absorbs the first 50 points of cold damage per 
round that the wearer would normally take (similar to the \textit{resist energy 
}spell). 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{protection from energy; Market Price: 
}+6 bonus. 

\textbf{Exceptional Arrow Deflection: }This shield functions like a \textit{shield 
of arrow deflection }except that it can deflect any type of ranged attack (including 
spells that require a ranged touch attack) as if it were an arrow. Any time the 
bearer would normally be struck by a ranged attack, he or she can make a Reflex 
saving throw (DC 20). If the ranged attack has an enhancement bonus (or a spell 
level), the DC increases by that amount. If he or she succeeds, the shield deflects 
the attack. The bearer must be aware of the attack and not flat-footed. Attempting 
to deflect a ranged attack doesn't count as an action. 

\textit{Caster Level: }23rd; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{shield, wall of force; Market Price: +}8 
bonus. 

\textbf{Fire Warding: }The armor absorbs the first 50 points of fire damage per 
round that the wearer would normally take (similar to the \textit{resist energy 
}spell). 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{protection from energy; Market Price: 
+}6 bonus. 

\textbf{Great Invulnerability: }This suit of armor or shield grants the wearer 
damage reduction. The damage reduction can be 10/magic, 15/magic, 5/epic, or 10/epic, 
depending on the armor. 

\textit{Caster Level: }19th (10/magic), 20th (15/magic), 21st (5/epic), 22nd (10/epic); 
\textit{Prerequisites: }Craft Magic Arms and Armor, Craft Epic Magic Arms and Armor, 
\textit{stoneskin, wish, }or \textit{miracle; Market Price:+}4 bonus (10/magic),+5 
bonus (15/magic),+6 bonus (5/epic),+7 bonus (10/epic). 

\textbf{Great Reflection: }Any time its bearer of this shield is targeted with 
a spell, it automatically reflects the spell back at the caster (as the \textit{spell 
turning }spell). The wearer can lower or raise this effect as a free action (thus 
allowing beneficial spells in as desired). 

\textit{Caster Level: }25th; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{spell turning; Market Price: +}10 bonus. 

\textbf{Great Spell Resistance: }This suit of armor or shield grants the wearer 
spell resistance. The spell resistance can be 21, 23, 25, or 27, depending on the 
armor. 

\textit{Caster Level: }21st (SR 21), 22nd (SR 23), 23rd (SR 25), 24th (SR 27); 
\textit{Prerequisites: }Craft Magic Arms and Armor, Craft Epic Magic Arms and Armor, 
\textit{spell resistance; Market Price: +}6 bonus (SR 21), +7 bonus (SR 23), +8 
bonus (SR 25), +9 bonus (SR 27). 

\textbf{Infinite Arrow Deflection: }This shield functions like a \textit{shield 
of arrow deflection}, though it can deflect any number of projectiles or thrown 
weapons each round. Any time the bearer would normally be struck by a ranged weapon, 
he or she can make a Reflex saving throw (DC 20). If the ranged weapon has an enhancement 
bonus, the DC increases by that amount. If he or she succeeds, the shield deflects 
the weapon. The bearer must be aware of the attack and not flat-footed. Attempting 
to deflect a ranged weapon doesn't count as an action. Exceptional ranged weapons, 
such as boulders hurled by giants or ranged spells\textit{, }can't be deflected. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{protection from arrows, shield; Market 
Price: +}6 bonus. 

\textbf{Lightning Warding: }The armor absorbs the first 50 points of lightning 
damage per round that the wearer would normally take (similar to the \textit{resist 
energy }spell). 

\textit{Caster Level: }21st; \textit{Prereq-uisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{protection from energy; Market Price: 
+}6 bonus. 

\textbf{Negating: }Immediately after the wearer of this armor is hit with a magic 
weapon, the armor casts \textit{greater dispelling }on the weapon. (In the case 
of projectile weapons, the armor casts \textit{greater dispelling }on the weapon 
that fired the projectile if it is in range. If it is out of range, the armor does 
nothing.) No weapon can be affected by the armor more than once per day (regardless 
of the success of the dispel check). 

\textit{Caster Level: }20th; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{greater dispelling; Market Price: +}5 
bonus. 

\textbf{Sonic Warding: }The armor absorbs the first 50 points of sonic damage per 
round that the wearer would normally take (similar to the \textit{resist energy 
}spell). 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{protection from energy; Market Price: 
+}6 bonus. 

\vspace{12pt}
\textbf{Table: Specific Epic Magic Armor and Shields }

\begin{tabular}{|>{\raggedright}p{36pt}|>{\raggedright}p{176pt}|>{\raggedright}p{63pt}|}
\hline
d\textbf{\%} & S\textbf{pecific Armor or Shield} & \subsection*{M\textbf{arket 
Price }}\tabularnewline
\hline
01-20 & Shapeshifter's armor & 400,165 gp \tabularnewline
\hline
21-40 & Warlord's breastplate & 416,200 gp \tabularnewline
\hline
41-57 & Dragonskin armor & 564,550 gp \tabularnewline
\hline
58-71 & Armor of the celestial battalion & 616,300 gp \tabularnewline
\hline
72-82 & Armor of the abyssal horde & 768,260 gp \tabularnewline
\hline
83-93 & Antimagic armor & 871,500 gp \tabularnewline
\hline
94-100 & Bulwark of the great dragon & 1,612,980 gp\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\subsection*{SPECIFIC ARMORS AND SHIELDS }

\vspace{12pt}
\textbf{Antimagic Armor: }This +\textit{1 negating full plate armor of invulnerability 
}is crafted of adamantine (and thus has damage reduction 3/-). The armor provides 
a -5 penalty on dispel checks made against it or its wearer. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{greater dispelling, stoneskin, wish }or 
\textit{miracle; Market Price: }871,500 gp; \textit{Cost to Create: }436,500 gp 
+ 18,700 XP. 

\textbf{Armor of the Abyssal Horde: }This+\textit{6 full plate armor's} clawed 
gauntlets are effectively +4 keen weapons (1d10/19-20) that afflict the target 
as if she had been struck by an \textit{energy drain }spell (Fortitude negates 
DC 23). The armor bestows two negative levels on any nonevil creature wearing it. 
These negative levels persist as long as the armor is worn and disappear when the 
armor is removed. The negative levels never result in actual level loss, but they 
cannot be overcome in any way (including \textit{restoration }spells) while the 
armor is worn. 

\textit{Caster Level: }20th; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, creator must be evil, \textit{energy drain; Market 
Price: }768,260 gp; \textit{Cost to Create: }385,260 gp + 17,660 XP. 

\textbf{Armor of the Celestial Battalion: }This+\textit{7 chainmail }is so fine 
and light that it can be worn under normal clothing without revealing its presence. 
It has a maximum Dexterity bonus of +10, no armor check penalty, and an arcane 
spell failure chance of 10\%. It is considered light armor, and it allows the wearer 
to fly at will (as the \textit{fly }spell). Furthermore, the wearer is at all times 
surrounded by a \textit{magic circle against evil }effect (as the spell) which, 
if dispelled, can be created again as a free action. 

\textit{Caster Level: }20th; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, creator must be good, \textit{magic circle against 
evil; Market Price: }616,300 gp; \textit{Cost to Create: }308,300 gp + 16,160 XP. 

\textbf{Bulwark of the Great Dragon: }This +\textit{6 heavy shield} bears the image 
of a dragon's head. Three times per day, the bearer of the shield can command it 
to belch forth a breath weapon of the appropriate type. The range of this breath 
weapon is 80 feet (if a line) or 40 feet (if a cone). Regardless of the type, the 
breath weapon deals 20d6 points of damage. In addition, the shield provides the 
bearer with resistance 50 to the energy type that matches its breath weapon. To 
determine what type of \textit{dragon shield }is found, roll d\% and consult the 
following table:

\begin{tabular}{|>{\raggedright}p{36pt}|>{\raggedright}p{36pt}|>{\raggedright}p{111pt}|}
\hline
d\textbf{\%} & C\textbf{olor} & \subsection*{B\textbf{reath }}\tabularnewline
\hline
01-10 & Black & Line of acid\tabularnewline
\hline
11-20 & Blue & Line of lightning \tabularnewline
\hline
21-30 & Brass & Line of fire \tabularnewline
\hline
31-40 & Bronze & Line of lightning \tabularnewline
\hline
41-50 & Copper & Line of acid \tabularnewline
\hline
51-60 & Gold & Cone of fire \tabularnewline
\hline
61-70 & Green & Cone of corrosive (acid) gas \tabularnewline
\hline
71-80 & Red & Cone of fire \tabularnewline
\hline
81-90 & Silver & Cone of cold \tabularnewline
\hline
91-100 & White & Cone of cold \tabularnewline
\hline
\end{tabular}

\textit{Caster Level: }20th; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{protection from energy; Market Price: 
}1,612,970 gp; \textit{Cost to Create: }806,570 gp + 26,128 XP. 

\textbf{Dragonskin Armor: }This+\textit{5 full plate armor }is crafted from the 
hide of a great wyrm dragon. At the wearer's command, the armor sprouts enormous 
dragon wings, allowing the wearer to fly at a speed of 90 feet (clumsy) for a total 
of 4 hours each day. The armor also grants immunity to a specific type of energy, 
based on the color of dragon that supplied the armor. Roll d\% on the following 
table to determine the color and immunity.

\begin{tabular}{|>{\raggedright}p{28pt}|>{\raggedright}p{28pt}|>{\raggedright}p{41pt}|}
\hline
d\textbf{\%} & C\textbf{olor} & I\textbf{mmunity}\tabularnewline
\hline
01-10 & Black & Acid\tabularnewline
\hline
11-20 & Blue & Lightning\tabularnewline
\hline
21-30 & Brass & Fire\tabularnewline
\hline
31-40 & Bronze & Lightning\tabularnewline
\hline
41-50 & Copper & Acid\tabularnewline
\hline
51-60 & Gold & Fire\tabularnewline
\hline
61-70 & Green & Acid\tabularnewline
\hline
71-80 & Red & Fire\tabularnewline
\hline
81-90 & Silver & Cold\tabularnewline
\hline
91-100 & White & Cold\tabularnewline
\hline
\end{tabular}

The wearer of the armor takes a -4 circumstance penalty on Diplomacy checks with 
dragons, but gains a +4 circumstance bonus on Intimidate checks against dragons. 

\textit{Caster Level: }24th; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{protection from energy, shapechange; Market 
Price: }564,550 gp; \textit{Cost to Create: }283,250 gp + 15,629 XP. 

\textbf{Shapeshifter's Armor: }This suit of +\textit{6 hide armor }grants its full 
Armor Class bonus regardless of any form the wearer takes (with \textit{polymorph, 
shapechange, wild shape, }or similar abilities). 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{shapechange }or \textit{wild shape }ability; 
\textit{Market Price: }400,165 gp; \textit{Cost to Create: }200,165 gp + 14,000 
XP. 

\textbf{Warlord's Breastplate: }This +\textit{6 mithral breastplate }has an armor 
check penalty of -1, a maximum Dexterity bonus of +5, and an arcane spell failure 
chance of 15\%. It is considered light armor and weighs 15 pounds. It grants the 
wearer a +4 enhancement bonus to Charisma and allows the wearer to attract and 
lead a number of followers as if he or she had the Leadership feat (though this 
power doesn't allow the wearer to attract a cohort). If the wearer already has 
the Leadership feat, this armor has no cumulative effect on his or her followers. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{charm monster, sympathy; Market Price: 
}416,200 gp; \textit{Cost to Create: }210,200 gp + 14,120 XP. 

\vspace{12pt}
\subsection*{{\LARGE{}WEAPONS }}

Except when otherwise stated, epic magic weapons follow the rules for nonepic magic 
weapons.  There is no limit to an epic magic weapon's enhancement bonus, to the 
market price modifier of an epic magic weapon special ability, or to the total 
of an epic magic weapon's enhancement bonus and market price modifier. 

\vspace{12pt}
\subsection*{EPIC WEAPON BASE PRICE }

To find the base price of an epic magic weapon, roll on Table: Weapons. Note that 
the +6 to +10 rows apply only to weapons that provide an enhancement bonus of +6 
to +10 or weapons with a single special ability whose market price modifier is 
+6 to +10. Magic weapons with a total effective bonus of +6 to +10 but that have 
an enhancement bonus of +5 or less and special abilities whose individual market 
price modifiers are +5 or less use the table for nonepic magic weapons\textit{ 
}to determine price. 

\vspace{12pt}
\textbf{Table: Weapons }

\begin{tabular}{|>{\raggedright}p{44pt}|>{\raggedright}p{136pt}|>{\raggedright}p{133pt}|}
\hline
d\textbf{\%} & E\textbf{nhancement Bonus } & M\textbf{arket Price}\textsuperscript{\textbf{1}} 
\tabularnewline
\hline
01-03 & +1  & +2,000 gp \tabularnewline
\hline
04-07 & +2  & +8,000 gp \tabularnewline
\hline
08-13 & +3  & +18,000 gp \tabularnewline
\hline
14-20 & +4  & +32,000 gp \tabularnewline
\hline
21-28 & +5  & +50,000 gp \tabularnewline
\hline
29-36 & +6  & +720,000 gp \tabularnewline
\hline
37-43 & +7  & +980,000 gp \tabularnewline
\hline
44-49 & +8  & +1,280,000 gp \tabularnewline
\hline
50-53 & +9  & +1,620,000 gp \tabularnewline
\hline
54-56 & +10  & +2,000,000 gp \tabularnewline
\hline
57-61 & \multicolumn{2}{p{269pt}|}{Specific weapon\textsuperscript{\textbf{2}} 
}\tabularnewline
\hline
62-80 & \multicolumn{2}{p{269pt}|}{Melee weapon with special ability and roll again\textsuperscript{\textbf{3}} 
}\tabularnewline
\hline
81-99 & \multicolumn{2}{p{269pt}|}{Ranged weapon with special ability and roll 
again\textsuperscript{\textbf{3}} }\tabularnewline
\hline
100 & \multicolumn{2}{p{269pt}|}{Roll on Table: Epic Weapons }\tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{1 This price is for 50 arrows, crossbow bolts, or sling 
bullets. }\tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{2 See Table: Specific Weapons. }\tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{3 See Table: Melee Weapon Special Abilities for melee 
weapons or Table: Ranged Weapon Special Abilities for ranged weapons. }\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\textbf{Table: Epic Weapons }

\begin{tabular}{|>{\raggedright}p{44pt}|>{\raggedright}p{136pt}|>{\raggedright}p{133pt}|}
\hline
d\textbf{\%} & E\textbf{nhancement Bonus } & \section*{M\textbf{arket Price }}\tabularnewline
\hline
01-21 & +11  & +2,420,000 gp \tabularnewline
\hline
22-39 & +12  & +2,880,000 gp \tabularnewline
\hline
40-54 & +13  & +3,380,000 gp \tabularnewline
\hline
55-66 & +14  & +3,920,000 gp \tabularnewline
\hline
67-76 & +15  & +4,500,000 gp \tabularnewline
\hline
77-84 & +16  & +5,120,000 gp \tabularnewline
\hline
85-90 & +17  & +5,780,000 gp \tabularnewline
\hline
91-94 & +18  & +6,480,000 gp \tabularnewline
\hline
95-97 & +19  & +7,220,000 gp \tabularnewline
\hline
98-99 & +20  & +8,000,000 gp \tabularnewline
\hline
100 & Roll again and add +10 to bonus\textsuperscript{\textbf{1}}  & 2\tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{1This is cumulative if rolled multiple times. }\tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{2 For enhancement bonuses higher than +20, the market 
price modifier is equal to the square of the bonus x20,000 gp. }\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\textbf{Table: Weapon Type Determination }

\begin{tabular}{|>{\raggedright}p{45pt}|>{\raggedright}p{281pt}|}
\hline
d\textbf{\%} & W\textbf{eapon Type }\tabularnewline
\hline
01-70 & Roll on Table: Common Melee Weapon \tabularnewline
\hline
71-80 & Roll on Table: Uncommon Melee Weapons\tabularnewline
\hline
81-100 & Roll on Table: Ranged Weapons\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\textbf{Table: Common Melee Weapons }

\begin{tabular}{|>{\raggedright}p{44pt}|>{\raggedright}p{136pt}|>{\raggedright}p{133pt}|}
\hline
d\textbf{\%} & W\textbf{eapon}\textsuperscript{\textbf{1}} & W\textbf{eapon Cost}\textsuperscript{\textbf{2}}\tabularnewline
\hline
01-04 & Dagger  & +302 gp \tabularnewline
\hline
05-14 & Greataxe  & +320 gp \tabularnewline
\hline
15-24 & Greatsword  & +350 gp \tabularnewline
\hline
25-28 & Kama  & +302 gp \tabularnewline
\hline
29-41 & Longsword  & +315 gp \tabularnewline
\hline
42-45 & Mace, light  & +305 gp \tabularnewline
\hline
46-50 & Mace, heavy  & +312 gp \tabularnewline
\hline
51-54 & Nunchaku  & +302 gp \tabularnewline
\hline
55-57 & Quarterstaff \textsuperscript{\textbf{3}}  & +600 gp \tabularnewline
\hline
58-61 & Rapier  & +320 gp \tabularnewline
\hline
62-66 & Scimitar  & +315 gp \tabularnewline
\hline
67-70 & Shortspear  & +302 gp \tabularnewline
\hline
71-74 & Siangham  & +303 gp \tabularnewline
\hline
75-84 & Sword, bastard  & +335 gp \tabularnewline
\hline
85-89 & Sword, short  & +310 gp \tabularnewline
\hline
90-100 & Waraxe, dwarven  & +330 gp \tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{All magic weapons are masterwork weapons. }\tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{1 For weapons normally made of steel, roll d\% to determine 
material: 01-85 steel; 86-99 adamantine; 100 other. Adjust price accordingly. }\tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{2 Add to enhancement bonus on Table: Weapons to determine 
total market price. }\tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{3 Masterwork double weapons incur double the masterwork 
cost to account for each head (+300 gp masterwork cost per head for a total of 
+600 gp). Double weapons have separate bonuses for their different heads. If randomly 
determined, the second head of a double weapon has the same enhancement bonus as 
the main head (01-50 on d\%), doubling the cost of the bonus, or its enhancement 
bonus is one less (51-100 on d\%) and it has no special abilities. }\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\textbf{Table: Uncommon Melee Weapons }

\begin{tabular}{|>{\raggedright}p{44pt}|>{\raggedright}p{198pt}|>{\raggedright}p{71pt}|}
\hline
d\textbf{\%} & W\textbf{eapon}\textsuperscript{\textbf{1}} & W\textbf{eapon Cost 
}\textsuperscript{\textbf{2}} \tabularnewline
\hline
01-03 & Axe, orc double\textsuperscript{\textbf{3}} & +660 gp \tabularnewline
\hline
04-07 & Battleaxe & +310 gp \tabularnewline
\hline
08-10 & Chain, spiked & +325 gp \tabularnewline
\hline
11-12 & Club & +300 gp \tabularnewline
\hline
13-16 & Crossbow, hand & +400 gp \tabularnewline
\hline
17-19 & Crossbow, repeating & +550 gp \tabularnewline
\hline
20-21 & Dagger, punching & +302 gp \tabularnewline
\hline
22-23 & Falchion & +375 gp \tabularnewline
\hline
24-26 & Flail, dire \textsuperscript{\textbf{3}} & +690 gp \tabularnewline
\hline
27-31 & Flail, heavy & +315 gp \tabularnewline
\hline
32-35 & Flail, light & +308 gp \tabularnewline
\hline
36-37 & Gauntlet & +302 gp \tabularnewline
\hline
38-39 & Gauntlet, spiked & +305 gp \tabularnewline
\hline
40-41 & Glaive & +308 gp \tabularnewline
\hline
42-43 & Greatclub & +305 gp \tabularnewline
\hline
44-45 & Guisarme & +309 gp \tabularnewline
\hline
46-48 & Halberd & +310 gp \tabularnewline
\hline
49-51 & Halfspear & +301 gp \tabularnewline
\hline
52-54 & Hammer, gnome hooked \textsuperscript{\textbf{3}} & +620 gp \tabularnewline
\hline
55-56 & Hammer, light & +301 gp \tabularnewline
\hline
57-58 & Handaxe & +306 gp \tabularnewline
\hline
59-61 & Kukri & +308 gp \tabularnewline
\hline
62-63 & Lance, heavy & +310 gp \tabularnewline
\hline
64-65 & Lance, light & +306 gp \tabularnewline
\hline
66-67 & Longspear & +305 gp \tabularnewline
\hline
68-70 & Morningstar & +308 gp \tabularnewline
\hline
71-72 & Net & +320 gp \tabularnewline
\hline
73-74 & Pick, heavy & +308 gp \tabularnewline
\hline
75-76 & Pick, light & +304 gp \tabularnewline
\hline
77-78 & Ranseur & +310 gp \tabularnewline
\hline
79-80 & Sap & +301 gp \tabularnewline
\hline
81-82 & Scythe & +318 gp \tabularnewline
\hline
83-84 & Shuriken & +301 gp \tabularnewline
\hline
85-86 & Sickle & +306 gp \tabularnewline
\hline
87-89 & Sword, two-bladed \textsuperscript{\textbf{3}} & +700 gp \tabularnewline
\hline
90-91 & Trident & +315 gp \tabularnewline
\hline
92-94 & Urgrosh, dwarven \textsuperscript{\textbf{3}} & +650 gp \tabularnewline
\hline
95-97 & Warhammer & +312 gp \tabularnewline
\hline
98-100 & Whip & +301 gp \tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{All magic weapons are masterwork weapons. }\tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{1 For weapons normally made of steel, roll d\% to determine 
material: 01-85 steel; 86-99 adamantine; 100 other. Adjust price accordingly. }\tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{2 Add to enhancement bonus on Table: Weapons to determine 
total market price. }\tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{3 Masterwork double weapons incur double the masterwork 
cost to account for each head (+300 gp masterwork cost per head for a total of 
+600 gp). Double weapons have separate bonuses for their different heads. If randomly 
determined, the second head of a double weapon has the same enhancement bonus as 
the main head (01-50 on d\%), doubling the cost of the bonus, or its enhancement 
bonus is one less (51-100) and it has no special abilities.}\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\textbf{Table: Ranged Weapons }

\begin{tabular}{|>{\raggedright}p{58pt}|>{\raggedright}p{191pt}|>{\raggedright}p{64pt}|}
\hline
d\textbf{\%} & W\textbf{eapon } & W\textbf{eapon Cost}\textsuperscript{\textbf{1}} 
\tabularnewline
\hline
01-10 & Ammunition  & \tabularnewline
\hline
01-50 & Arrows (50) & +350 gp \tabularnewline
\hline
51-80 & Bolts, crossbow (50) & +350 gp \tabularnewline
\hline
81-100 & Bullets, sling (50) & +350 gp \tabularnewline
\hline
11-15 & Axe, throwing & +308 gp \tabularnewline
\hline
16-25 & Crossbow, heavy & +350 gp \tabularnewline
\hline
26-35 & Crossbow, light & +335 gp \tabularnewline
\hline
36-39 & Dart & +300 gp 5 sp \tabularnewline
\hline
40-41 & Javelin & +301 gp \tabularnewline
\hline
42-46 & Shortbow & +330 gp \tabularnewline
\hline
47-51 & Shortbow, composite & +375 gp \tabularnewline
\hline
52-56 & Shortbow, composite (+1 Str bonus) & +450 gp \tabularnewline
\hline
57-61 & Shortbow, composite (+2 Str bonus) & +525 gp \tabularnewline
\hline
62-65 & Sling & +300 gp \tabularnewline
\hline
66-75 & Longbow & +375 gp \tabularnewline
\hline
76-80 & Longbow, composite & +400 gp \tabularnewline
\hline
81-85 & Longbow, composite (+1 Str bonus) & +500 gp \tabularnewline
\hline
86-90 & Longbow, composite (+2 Str bonus) & +600 gp \tabularnewline
\hline
91-95 & Longbow, composite (+3 Str bonus) & +700 gp \tabularnewline
\hline
96-100 & Longbow, composite (+4 Str bonus) & +800 gp \tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{All magic weapons are masterwork weapons.}\tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{1 Add to enhancement bonus on Table: Weapons to determine 
total market price.}\tabularnewline
\hline
\end{tabular} 

\vspace{12pt}
\textbf{Table: Melee Weapon Special Abilities }

\begin{tabular}{|>{\raggedright}p{30pt}|>{\raggedright}p{179pt}|>{\raggedright}p{104pt}|}
\hline
d\textbf{\%} & S\textbf{pecial Ability } & M\textbf{arket Price Modifier}\textsuperscript{\textbf{1}} 
\tabularnewline
\hline
01-08 & Acidic blast & +6 bonus \tabularnewline
\hline
09-16 & Fiery blast & +6 bonus \tabularnewline
\hline
17-24 & Icy blast & +6 bonus \tabularnewline
\hline
25-32 & Lightning blast & +6 bonus \tabularnewline
\hline
33-40 & Mighty disruption & +6 bonus \tabularnewline
\hline
41-48 & Sonic blast & +6 bonus \tabularnewline
\hline
49-56 & Dread & +7 bonus \tabularnewline
\hline
57-61 & Anarchic power & +8 bonus \tabularnewline
\hline
62-66 & Everdancing & +8 bonus \tabularnewline
\hline
67-71 & Holy power & +8 bonus \tabularnewline
\hline
72-76 & Axiomatic power & +8 bonus \tabularnewline
\hline
77-81 & Unholy power & +8 bonus \tabularnewline
\hline
82-89 & \multicolumn{2}{p{283pt}|}{Roll on nonepic magic item Table: Melee Weapon 
Special Abilities, then roll again on this table.}\tabularnewline
\hline
90-97 & \multicolumn{2}{p{283pt}|}{Roll twice on nonepic magic item Table: Melee 
Weapon Special Abilities.}\tabularnewline
\hline
98-100 & \multicolumn{2}{p{283pt}|}{Roll twice again\textsuperscript{\textbf{2}}\textsuperscript{ 
}}\tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{1 Add to enhancement bonus on Table: Epic Weapons to 
determine total market price. }\tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{2 If a special ability is rolled twice, only one counts. 
If two versions of the same special ability are rolled, use the better. }\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\textbf{Table: Ranged Weapon Special Abilities }

\begin{tabular}{|>{\raggedright}p{30pt}|>{\raggedright}p{179pt}|>{\raggedright}p{104pt}|}
\hline
d\textbf{\%} & S\textbf{pecial Ability } & M\textbf{arket Price Modifier}\textsuperscript{\textbf{1}} 
\tabularnewline
\hline
01-07 & Acidic blast & +6 bonus \tabularnewline
\hline
08-14 & Distant shot & +6 bonus \tabularnewline
\hline
15-21 & Fiery blast & +6 bonus \tabularnewline
\hline
22-28 & Icy blast & +6 bonus \tabularnewline
\hline
29-35 & Lightning blast & +6 bonus \tabularnewline
\hline
36-41 & Sonic blast & +6 bonus \tabularnewline
\hline
42-48 & Triple-throw & +6 bonus \tabularnewline
\hline
49-53 & Unerring accuracy & +6 bonus \tabularnewline
\hline
54-60 & Dread & +7 bonus \tabularnewline
\hline
61-65 & Anarchic power & +8 bonus \tabularnewline
\hline
66-70 & Holy power & +8 bonus \tabularnewline
\hline
71-75 & Axiomatic power & +8 bonus \tabularnewline
\hline
76-80 & Unholy power & +8 bonus \tabularnewline
\hline
81-88 & \multicolumn{2}{p{283pt}|}{Roll on nonepic magic item Table: Ranged Weapon 
Special Abilities, then roll again on this table.}\tabularnewline
\hline
89-96 & \multicolumn{2}{p{283pt}|}{Roll twice on nonepic magic item Table: Ranged 
Weapon Special Abilities.}\tabularnewline
\hline
97-100 & \multicolumn{2}{p{283pt}|}{ Roll twice again\textsuperscript{\textbf{2}} 
}\tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{1 Add to enhancement bonus on Table: Weapons to determine 
total market price. }\tabularnewline
\hline
\multicolumn{3}{|p{314pt}|}{2 If a special ability is rolled twice, only one counts. 
If two versions of the same special ability are rolled, use the better.}\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\textbf{Table: Specific Weapons }

\begin{tabular}{|>{\raggedright}p{30pt}|>{\raggedright}p{198pt}|>{\raggedright}p{85pt}|}
\hline
d\textbf{\%} & S\textbf{pecific Weapon } & \section*{M\textbf{arket Price }}\tabularnewline
\hline
01-18 & S\textit{tormbrand } & 235,350 gp \tabularnewline
\hline
19-33 & Q\textit{uarterstaff of alacrity } & 462,600 gp \tabularnewline
\hline
34-48 & S\textit{ouldrinker } & 478,335 gp \tabularnewline
\hline
49-60 & B\textit{ackstabber } & 770,310 gp \tabularnewline
\hline
61-68 & M\textit{ace of ruin } & 1,000,312 gp \tabularnewline
\hline
69-72 & G\textit{ripsoul } & 1,856,500 gp \tabularnewline
\hline
73-78 & E\textit{lven greatbow } & 2,900,400 gp \tabularnewline
\hline
79-64 & F\textit{inaldeath } & 3,580,308 gp \tabularnewline
\hline
85-90 & C\textit{haosbringer } & 4,025,350 gp \tabularnewline
\hline
91-94 & H\textit{oly devastator } & 4,620,315 gp \tabularnewline
\hline
95-98 & U\textit{nholy despoiler } & 4,620,315 gp \tabularnewline
\hline
99-100 & E\textit{verwhirling chain } & 5,220,325 gp \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\subsection*{EPIC WEAPON SPECIAL ABILITY DESCRIPTIONS }

Most magic weapons only have enhancement bonuses. They can also have special abilities, 
such as those detailed below and nonepic abilities.\textit{ }A weapon with a special 
ability must have at least a +1 enhancement bonus. 

\vspace{12pt}
\textbf{Acidic Blast: }On command, an acidic blast weapon drips acid (though this 
deals no damage to the wielder). On any hit, this acid splashes the creature struck, 
dealing +3d6 points of bonus acid damage. On a successful critical hit it instead 
deals +6d6 points of acid damage (or +9d6 if the critical multiplier is x3, or 
+12d6 if the critical multiplier is x4). Bows, crossbows, and slings with this 
special ability bestow the bonus acid damage upon their ammunition. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{acid fog; Market Price: +}6 bonus. 

\textbf{Anarchic Power: }This weapon is chaos-aligned and thus bypasses the corresponding 
damage reduction. When a weapon of anarchic power strikes a lawful target, this 
power deals +3d6 points of bonus chaotic damage to the target, and the target gains 
one negative level (Fortitude DC 23 to remove 24 hours later). On a successful 
critical hit it instead deals +6d6 points of chaotic damage and bestows two negative 
levels (or +9d6 and three negative levels if the critical multiplier is x3, or 
+12d6 and four negative levels if the critical multiplier is x4). The weapon bestows 
three negative levels on any lawful creature attempting to wield it. These negative 
levels remain as long as the weapon is in hand and disappear when the weapon is 
no longer wielded. These negative levels never result in actual level loss, but 
they cannot be overcome in any way (including \textit{restoration }spells) while 
the weapon is wielded. Bows, crossbows, and slings with this special ability bestow 
the anarchic power upon their ammunition. This special ability does not stack with 
the nonepic anarchic special ability\textit{. }

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{word of chaos; Market Price: +}8 bonus. 

\textbf{Axiomatic Power: }This weapon is lawful-aligned and thus bypasses the corresponding 
damager reduction. When a weapon of axiomatic power strikes a chaotic target, this 
power erupts forth and deals +3d6 points of bonus lawful damage to the target, 
and the target gains one negative level (Fortitude DC 23 to remove 24 hours later). 
On a successful critical hit it instead deals +6d6 points of lawful damage and 
bestows two negative levels (or +9d6 and three negative levels if the critical 
multiplier is x3, or +12d6 and four negative levels if the critical multiplier 
is x4). The weapon bestows three negative levels on any chaotic creature attempting 
to wield it. These negative levels remain as long as the weapon is in hand and 
disappear when the weapon is no longer wielded. These negative levels never result 
in actual level loss, but they cannot be overcome in any way (including \textit{restoration 
}spells) while the weapon is wielded. Bows, crossbows, and slings with this special 
ability bestow the lawful power upon their ammunition. This special ability does 
not stack with the nonepic axiomatic special ability.\textit{ }

\textit{Caster Level: }23rd; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{dictum; Market Price: +}8 bonus. 

\textbf{Distant Shot: }A distant shot weapon can be used against any target within 
line of sight at no penalty for range. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{discern location; Market Price: +}6 bonus. 

\textbf{Dread: }A dread weapon excels at attacking one type of creature. Against 
its designated foe, its effective enhancement bonus is +4 better than its normal 
enhancement bonus. Further, it deals +4d6 points of bonus damage against the foe, 
and if it scores a successful critical hit against the foe, that creature must 
make a Fortitude save (DC 27) or be destroyed instantly and turned to dust. (This 
even affects creatures immune to critical hits or death magic.) To randomly determine 
a dread weapon's designated foe, roll on the following table. 

\textit{Caster Level: }22nd; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{summon monster IX; Market Price: +}7 bonus. 

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{34pt}|>{\raggedright}p{112pt}|>{\raggedright}p{40pt}|>{\raggedright}p{112pt}|}
\hline
d\textbf{\%} & D\textbf{esignated Foe} & d\textbf{\%} & D\textbf{esignated Foe}\tabularnewline
\hline
01-05 & Aberrations & 58-60 & Humanoids, orc\tabularnewline
\hline
06-09 & Animals & 61-65 & Magical beasts\tabularnewline
\hline
10-16 & Constructs & 66-70 & Monstrous humanoids\tabularnewline
\hline
17-22 & Dragons & 71-72 & Oozes\tabularnewline
\hline
23-27 & Elementals & 73 & Outsiders, air\tabularnewline
\hline
28-32 & Fey & 74-76 & Outsiders, chaotic\tabularnewline
\hline
33-39 & Giants & 77 & Outsiders, earth\tabularnewline
\hline
40 & Humanoids, aquatic & 78-80 & Outsiders, evil\tabularnewline
\hline
41-42 & Humanoids, dwarf & 81 & Outsiders, fire\tabularnewline
\hline
43-44 & Humanoids, elf & 82-84 & Outsiders, good\tabularnewline
\hline
45 & Humanoids, gnoll & 85-87 & Outsiders, lawful\tabularnewline
\hline
46 & Humanoids, gnome & 88 & Outsiders, water\tabularnewline
\hline
47-49 & Humanoids, goblinoid & 89-90 & Plants\tabularnewline
\hline
50 & Humanoids, halfling & 91-98 & Undead\tabularnewline
\hline
51-54 & Humanoids, human & 99-100 & Vermin\tabularnewline
\hline
55-57 & Humanoids, reptilian &  & \tabularnewline
\hline
\end{tabular}\newpage

\newpage
\vspace{12pt}
\textbf{Everdancing: }An everdancing weapon is much like a dancing weapon, though 
it can be loosed with a free action and will fight as long as desired. It can move 
up to 60 feet away from its owner. Its owner can instruct it to move to a different 
target as a move-equivalent action. If its owner is rendered unconscious or otherwise 
unable to direct it, it will fight the same opponent as long as that opponent is 
conscious and within range. The owner of an everdancing weapon can grasp it again 
as a free action (assuming it is within reach). 

\textit{Caster Level: }23rd; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{animate objects; Market Price: +}8 bonus. 

\textbf{Fiery Blast: }On command, a fiery blast weapon is sheathed in fire (though 
this deals no damage to the wielder). On any hit, this fire engulfs the creature 
struck, dealing +3d6 points of bonus fire damage. On a successful critical hit 
it instead deals +6d6 points of fire damage (or +9d6 if the critical multiplier 
is x3, or +12d6 if the critical multiplier is x4). Bows, crossbows, and slings 
with this special ability bestow the bonus fire damage upon their ammunition. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{fireball; Market Price: +}6 bonus. 

\textbf{Holy Power: }This weapon is good-aligned and thus bypasses the corresponding 
damager reduction. When a weapon of holy power strikes an evil target, this power 
erupts forth and deals +3d6 points of bonus holy (good) damage to the target, and 
the target gains one negative level (Fortitude DC 23 to remove 24 hours later). 
On a successful critical hit it instead deals +6d6 points of holy (good) damage 
and bestows two negative levels (or +9d6 and three negative levels if the critical 
multiplier is x3, or +12d6 and four negative levels if the critical multiplier 
is x4). The weapon bestows three negative levels on any evil creature attempting 
to wield it. These negative levels remain as long as the weapon is in hand and 
disappear when the weapon is no longer wielded. These negative levels never result 
in actual level loss, but they cannot be overcome in any way (including \textit{restoration 
}spells) while the weapon is wielded. Bows, crossbows, and slings with this special 
ability bestow the holy power upon their ammunition. This special ability does 
not stack with the nonepic holy special ability.\textit{ }

\textit{Caster Level: }23rd; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{holy word; Market Price: +}8 bonus. 

\textbf{Icy Blast: }On command, an icy blast weapon is sheathed in icy cold (though 
this deals no damage to the wielder). On any hit, this cold washes over the creature 
struck, dealing +3d6 points of bonus cold damage. On a successful critical hit 
it instead deals +6d6 points of cold damage (or +9d6 if the critical multiplier 
is x3, or +12d6 if the critical multiplier is x4). Bows, crossbows, and slings 
with this special ability bestow the bonus cold damage upon their ammunition. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{cone of cold; Market Price: +}6 bonus. 

\textbf{Lightning Blast: }On command, a lightning blast weapon crackles with electrical 
energy (though this deals no damage to the wielder). On any hit, lightning coruscates 
around the creature struck, dealing  +3d6 points of bonus electricity damage. On 
a successful critical hit it instead deals +6d6 points of electricity damage (or 
+9d6 if the critical multiplier is x3, or +12d6 if the critical multiplier is x4). 
Bows, crossbows, and slings with this special ability bestow the bonus electricity 
damage upon their ammunition. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{lightning bolt; Market Price: +}6 bonus. 

\textbf{Mighty Disruption: }Any undead creature struck in combat must succeed at 
a Fortitude save (DC 21) or be destroyed. A weapon of mighty disruption must be 
a bludgeoning weapon. (If this property is rolled for a piercing or slashing weapon, 
reroll.) 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{true resurrection; Market Price: +}6 bonus. 

\textbf{Sonic Blast: }On command, a sonic blast weapon emits a low thrumming hum 
(though this deals no damage to the wielder). On any hit, this becomes a thunderous 
roar that deals +3d6 points of bonus sonic damage to the creature struck. On a 
successful critical hit it instead deals +6d6 points of sonic damage (or +9d6 if 
the critical multiplier is x3, or +12d6 if the critical multiplier is x4). Bows, 
crossbows, and slings with this special ability bestow the bonus sonic damage upon 
their ammunition. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{shout; Market Price: +}6 bonus. 

\textbf{Triple-Throw: }This special ability can only be placed on a weapon that 
can be thrown. (If this property is rolled for a weapon that cannot be thrown, 
reroll.) A triple-throw weapon creates two duplicates of itself when thrown. Both 
the original and the duplicate weapons attack separately (at the same attack bonus). 
Regardless of the success of any of the attacks, the duplicates immediately disappear 
after the attack is completed. Any bonuses on damage due to accuracy or precision 
(including those from sneak attacks, the Precise Shot feat, or the ranger's favored 
enemy bonus) apply only to the original weapon's damage, not to the duplicates. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{shades; Market Price: +}6 bonus. 

\textbf{Unerring Accuracy: }Ranged attacks made with this weapon negate the AC 
bonus granted by any cover short of total cover.  The weapon's ranged attacks also 
ignore any miss chance from concealment (including total concealment, but the must 
still aim his or her attacks at the correct square). 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{true seeing; Market Price: +}6 bonus. 

\textbf{Unholy Power: }This weapon is evil-aligned and this bypasses the corresponding 
damage reduction. When a weapon of unholy power strikes a good target, this power 
erupts forth and deals +3d6 points of bonus unholy (evil) damage to the target, 
and the target gains one negative level (Fortitude DC 23 to remove 24 hours later). 
On a successful critical hit it instead deals +6d6 points of unholy (evil) damage 
and bestows two negative levels (or +9d6 and three negative levels if the critical 
multiplier is x3, or +12d6 and four negative levels if the critical multiplier 
is x4). The weapon bestows three negative levels on any good creature attempting 
to wield it. These negative levels remain as long as the weapon is in hand and 
disappear when the weapon is no longer wielded. These negative levels never result 
in actual level loss, but they cannot be overcome in any way (including \textit{restoration 
}spells) while the weapon is wielded. Bows, crossbows, and slings with this special 
ability bestow the unholy power upon their ammunition. This special ability does 
not stack with the nonepic unholy special ability.\textit{ }

\textit{Caster Level: }23rd; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{unholy word; Market Price: +}8 bonus. 

\vspace{12pt}
\subsection*{SPECIFIC WEAPONS}

\vspace{12pt}
\textbf{Backstabber: }This +\textit{6 short sword }adds +2d6 to the wielder's sneak 
attack damage. If the wielder does not have the sneak attack ability, this weapon 
does not grant it. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{inflict moderate} \textit{wounds; Market 
Price: }770,310 gp; \textit{Cost to Create: }385,310 gp + 17,700 XP. 

\textbf{Chaosbringer: }This +\textit{6 greataxe of anarchic power }grants its wielder 
the ability to fly into a rage (identical to a barbarian's rage) once per day (or 
one additional time per day if the wielder already has the rage class feature). 
If the wielder has the greater rage class feature, the weapon also grants the wielder 
the Incite Rage epic feat. 

\textit{Caster Level: }23rd; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{rage, mass charm monster; Market Price: 
}4,025,350 gp; \textit{Cost to Create: }2,012,850 gp + 50,250 XP. 

\textbf{Elven Greatbow: }In the hands of any nonelf, this bow performs only as 
a +\textit{2 composite longbo}w. In the hands of an elf, this weapon functions 
as a +\textit{5 composite longbow of unerring accurac}y with a Strength bonus that 
matches its elven wielder's current Strength at all times. Furthermore, any arrows 
loosed from the bow are considered keen, regardless of the enhancement bonus of 
the arrow fired. 

\textit{Caster Level: }23rd; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{bull's strength, keen edge, true seeing; 
Market Price: }2,900,400 gp; \textit{Cost to Create: }1,450,400 gp + 39,400 XP. 

\textbf{Everwhirling Chain: }This +\textit{4 defending everdancing spiked chain 
of speed }continuously twitches in its wielder's hands. The wielder of the \textit{everwhirling 
chain }can use it to make any number of attacks of opportunity per round (as if 
he or she had the Improved Combat Reflexes feat).

\textit{Caster Level: }23rd; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{animate objects, haste, shield }(or \textit{shield 
of fait}h); \textit{Market Price: }5,220,325 gp; \textit{Cost to Create: }2,610,325 
gp + 52,200 XP. 

\textbf{Finaldeath: }This +\textit{5 undead dread ghost touch morningstar }also 
grants its wielder immunity to energy drain attacks. Furthermore, if its wielder 
is capable of turning undead, he or she gains the Positive Energy Aura feat.

\textit{Caster Level: }22nd; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{death ward, plane shift, summon monster 
IX, }creator must be able to turn undead; \textit{Market Price: }3,580,308 gp; 
\textit{Cost to Create: }1,790,308 gp + 45,800 XP. 

\textbf{Gripsoul: }\textit{Gripsoul }is a +\textit{6 keen long-sword, }but instead 
of dealing additional damage on a critical hit, the weapon imprisons the victim 
in a gem set in the pommel of the sword as per a \textit{binding }spell heightened 
to 16th level (DC 30). The same is true of any blow that would otherwise kill a 
foe or knock him or her unconscious. Only one creature can be so held, but the 
wielder can release the bound soul at any time with a command word. 

\textit{Caster Level: }27th; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{binding; Market Price: }1,856,500 gp; 
\textit{Cost to Create: }934,500 gp + 28,440 XP. 

\textbf{Holy Devastator: }In the hands of any character other than a paladin, this 
sword performs as a +\textit{3 holy longswor}d. In the hands of a paladin, this 
weapon functions as a +\textit{7 longsword of holy power }and grants a +5 sacred 
bonus on the wielder's saving throws against spells with the evil descriptor or 
spells cast by evil characters. If the paladin wielder smites evil with the \textit{holy 
devastator, }he or she adds twice his or her paladin level to damage (rather than 
his or her paladin level). 

\textit{Caster Level: }23rd; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{holy aura, holy smite, holy word; Market 
Price: }4,620,315 gp; \textit{Cost to Create: }2,310,315 gp + 56,200 XP. 

\textbf{Mace of Ruin: }This +\textit{7 heavy mace }ignores the hardness or damage 
reduction of any object or creature it strikes. Furthermore, the weapon can deal 
critical hits to objects and constructs as if they were living creatures. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{disintegrate; Market Price: }1,000,312 
gp; \textit{Cost to Create: }500,312 gp + 20,000 XP. 

\textbf{Quarterstaff of Alacrity: }Both ends of this +\textit{5 quarterstaff of 
speed }have equal enhancement and special powers, meaning that it allows an additional 
attack with each end every round. While the \textit{quarterstaff of alacrity }is 
held, it grants its wielder a +5 resistance bonus on Reflex saves. It also deflects 
ranged weapons as if the wielder had the Deflect Arrows and Infinite Deflection 
feats. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{protection from arrows, shield; Market 
Price: }462,600 gp; \textit{Cost to Create: }231,600 gp + 14,620 XP. 

\textbf{Souldrinker: }This +\textit{5 bastard sword }bestows 2d4 negative levels 
on its target whenever it deals damage, just as if its target had been struck by 
the \textit{energy drain }spell. Each negative level bestowed grants the wielder 
5 temporary hit points. One day after being struck, the subject must make a Fortitude 
save (DC 25) for each negative level or lose a level. If this sword's power causes 
a character to have negative levels at least equal to his or her current level, 
the character is instantly slain and the wielder gains an additional 10 temporary 
hit points. Temporary hit points gained from this weapon last for a maximum of 
1 hour. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, Spell Focus (Necromancy), \textit{energy drain; 
Market Price: }478,335 gp; \textit{Cost to Create: }239,315 gp + 14,780 XP. 

\textbf{Stormbrand: }This +\textit{4 thundering shocking burst greatsword }allows 
its wielder to fly at will (as the \textit{fly }spell). Furthermore, the wielder 
can move normally (including flying) even in the strongest winds. When the weapon 
is drawn, the wielder gains electricity resistance 30 and sonic resistance 30. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{blindness/ deafness, call lightning }or 
\textit{lightning bolt, control winds, fly, protection from energy; Market Price: 
}235,350 gp; \textit{Cost to Create: }117,850 gp + 12,350 XP. 

\textbf{Unholy Despoiler: }In the hands of any character other than a blackguard, 
this sword performs as a +\textit{3 unholy longswor}d. In the hands of a blackguard, 
this weapon functions as a +\textit{7 longsword of unholy power }and grants a +5 
profane bonus on the wielder's saving throws against spells with the good descriptor 
or spells cast by good characters. If a blackguard wielder smites good with the 
\textit{unholy despoile}r, he or she adds twice his or her blackguard level to 
damage (rather than his or her blackguard level). 

\textit{Caster Level: }23rd; \textit{Prerequisites: }Craft Magic Arms and Armor, 
Craft Epic Magic Arms and Armor, \textit{blasphemy, unholy aura, unholy blight; 
Market Price: }4,650,315 gp; \textit{Cost to Create: }2,325,315 gp + 56,500 XP. 

\vspace{24pt}
\subsection*{{\LARGE{}RINGS }}

\vspace{12pt}
\subsection*{NONEPIC MAGIC ITEM }

\textbf{Universal Energy Resistance, Minor: }This ring functions as a \textit{ring 
of minor energy resistance }for all types of energy: fire, cold, electricity, acid, 
and sonic. When the wearer would normally take such damage, subtract 10 points 
of damage per round from the amount before applying. This is a nonepic magic item. 

\textit{Caster Level: }15th; \textit{Prerequisites: }Forge Ring, \textit{protection 
from energy; Market Price: }84,000 gp.

\vspace{24pt}
\textbf{Table: Epic Rings }

\begin{tabular}{|>{\raggedright}p{23pt}|>{\raggedright}p{212pt}|>{\raggedright}p{78pt}|}
\hline
d\textbf{\%} & R\textbf{ing } & \section*{M\textbf{arket Price }}\tabularnewline
\hline
01-08 & U\textit{niversal energy resistance, greater} & 308,000 gp\tabularnewline
\hline
09-13 & E\textit{nergy immunity (acid) } & 240,000 gp\tabularnewline
\hline
14-15 & E\textit{nergy immunity (cold) } & 240,000 gp\tabularnewline
\hline
19-23 & E\textit{nergy immunity (electricity) } & 240,000 gp \tabularnewline
\hline
24-28 & E\textit{nergy immunity (fire) } & 240,000 gp \tabularnewline
\hline
29-33 & E\textit{nergy immunity (sonic) } & 240,000 gp \tabularnewline
\hline
34-38 & A\textit{damant law } & 250,000 gp \tabularnewline
\hline
39-43 & C\textit{haotic fury } & 250,000 gp \tabularnewline
\hline
44-48 & E\textit{pic wizardry (V) } & 250,000 gp \tabularnewline
\hline
49-53 & I\textit{neffable evil } & 250,000 gp \tabularnewline
\hline
54-58 & V\textit{irtuous good } & 250,000 gp \tabularnewline
\hline
59-63 & R\textit{apid healing } & 300,000 gp \tabularnewline
\hline
64-68 & S\textit{equestering } & 300,000 gp \tabularnewline
\hline
69-72 & E\textit{pic wizardry (VI) } & 360,000 gp \tabularnewline
\hline
73-76 & I\textit{ronskin } & 400,000 gp \tabularnewline
\hline
77-80 & E\textit{pic wizardry (VII) } & 490,000 gp \tabularnewline
\hline
81-83 & W\textit{eaponbreaking } & 600,000 gp \tabularnewline
\hline
84-86 & E\textit{pic wizardry (VIII) } & 640,000 gp \tabularnewline
\hline
87-89 & E\textit{pic protection +6 } & 720,000 gp \tabularnewline
\hline
90-92 & E\textit{pic wizardry (IX)} & 810,000 gp \tabularnewline
\hline
93-95 & E\textit{pic protection +7 } & 980,000 gp \tabularnewline
\hline
96-97 & E\textit{pic protection +8 } & 1,280,000 gp \tabularnewline
\hline
98 & E\textit{pic protection +9 } & 1,620,000 gp \tabularnewline
\hline
99 & E\textit{pic protection +10 } & 2,000,000 gp \tabularnewline
\hline
100 & U\textit{niversal energy immunity } & 2,160,000 gp\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\subsection*{Epic Ring Descriptions }

Standard epic rings are described below.

\vspace{12pt}
\textbf{Adamant Law:  }The wearer of this ring is constantly sheathed in a \textit{shield 
of law }effect. It bestows one negative level on any chaotic creature that puts 
it on. The negative level remains as long as the ring is worn and disappears when 
the ring is removed. This negative level never results in actual level loss, but 
it cannot be overcome in any way (including \textit{restoration }spells) while 
the ring is worn. 

\textit{Caster Level: }15th; \textit{Prerequisites: }Forge Ring, Forge Epic Ring, 
\textit{shield of law, }creator must be lawful; \textit{Market Price: }250,000 
gp. 

\textbf{Chaotic Fury: }The wearer of this ring is constantly sheathed in a \textit{cloak 
of chaos }effect. It bestows one negative level on any lawful creature that puts 
it on. The negative level remains as long as the ring is worn and disappears when 
the ring is removed. This negative level never results in actual level loss, but 
it cannot be overcome in any way (including \textit{restoration }spells) while 
the ring is worn. 

\textit{Caster Level: }15th; \textit{Prerequisites: }Forge Ring, Forge Epic Ring, 
\textit{cloak of chaos, }creator must be chaotic; \textit{Market Price: }250,000 
gp. 

\textbf{Energy Immunity: }This band continually provides the wearer with immunity 
to a single type of energy: fire, cold, electricity, acid, or sonic. The wearer 
takes no damage from the energy of the specific type. 

\textit{Caster Level: }20th; \textit{Prerequisites: }Forge Ring, Forge Epic Ring, 
\textit{protection from energy; Market Price: }240,000 gp. Epic 

\textbf{Protection: }This ring offers continual magical protection in the form 
of a deflection bonus to Armor Class of +6 or higher. 

\textit{Caster Level: }20th: \textit{Prerequisites: }Forge Ring, Forge Epic Ring, 
\textit{shield of faith, }creator's caster level must be three times the ring's 
bonus; \textit{Market Price: }720,000 gp (+6), 980,000 gp (+7), 1,280,000 gp (+8), 
1,620,000 gp (+9), 2,000,000 gp (+10). 

\textbf{Epic Wizardry: }Like the \textit{ring of wizardry, }this ring comes in 
a variety of types useful only to arcane spellcasters. The wearer's arcane spells 
per day are doubled for one particular spell level. An \textit{epic ring of wizardry 
V }doubles 5th-level spells, an \textit{epic ring of wizardry VI }doubles 6th-level 
spells, an \textit{epic ring of wizardry VII }doubles 7th-level spells, an \textit{epic 
ring of wizardry VIII }doubles 8th-level spells, and an \textit{epic ring of wizardry 
IX }doubles 9th-level spells. Bonus spells from high ability scores, school specialization, 
or any other source are not doubled. 

\textit{Caster Level: }23rd \textit{(epic wizardry }V), 26th \textit{(epic wizardry 
V}I), 29th \textit{(epic wizardry VI}I), 32nd \textit{(epic wizardry VII}I), 35th 
\textit{(epic wizardry I}X); \textit{Prerequisites: }Forge Ring, Forge Epic Ring, 
\textit{wish; Market Price: }250,000 gp \textit{(epic wizardry }V), 360,000 gp 
\textit{(epic wizardry V}I), 490,000 gp \textit{(epic wizardry VI}I), 640,000 gp 
\textit{(epic wizardry VII}I), 810,000 gp \textit{(epic wizardry I}X). 

\textbf{Ineffable Evil: }The wearer of this ring is constantly sheathed in an \textit{unholy 
aura }effect. It bestows one negative level on any good creature that puts it on. 
The negative level remains as long as the ring is worn and disappears when the 
ring is removed. This negative level never results in actual level loss, but it 
cannot be overcome in any way (including \textit{restoration }spells) while the 
ring is worn. 

\textit{Caster Level: }15th; \textit{Prerequisites: }Forge Ring, Forge Epic Ring, 
\textit{unholy aura, }creator must be evil; \textit{Market Price: }250,000 gp. 

\textbf{Ironskin: }This ring grants its wearer damage reduction 15/adamantine. 

\textit{Caster Level: }20th; \textit{Prerequisites: }Forge Ring, Forge Epic Ring, 
\textit{iron body; Market Price: }400,000 gp. 

\textbf{Rapid Healing: }This ring grants a living wearer fast healing 3. It must 
be worn for 24 hours before its powers activate, and if removed it will not function 
again until it has been worn for 24 hours by the same individual. 

\textit{Caster Level: }20th; \textit{Prerequisites: }Forge Ring, Forge Epic Ring, 
\textit{regenerate; Market Price: }300,000 gp. 

\textbf{Sequestering: }This ring becomes invisible when worn. Upon command, the 
wearer gains the benefits of a \textit{sequester }spell (though he or she does 
not become comatose as normal for the spell). 

\textit{Caster Level: }20th; \textit{Prerequisites: }Forge Ring, Forge Epic Ring, 
\textit{sequester; Market Price: }300,000 gp. 

\textbf{Universal Energy Immunity: }This ring functions as a \textit{ring of energy 
immunity }for all types of energy--- fire, cold, electricity, acid, and sonic. 
The wearer takes no damage from energy of any of these types. 

\textit{Caster Level: }20th; \textit{Prerequisites: }Forge Ring, Forge Epic Ring, 
\textit{protection from energy; Market Price: }2,160,000 gp. 

\textbf{Universal Energy Resistance, Greater: }This ring functions as a \textit{ring 
of greater energy resistance }for all types of energy---fire, cold, electricity, 
acid, and sonic. When the wearer would normally take such damage, subtract 30 points 
of damage per round from the amount before applying. 

\textit{Caster Level: }20th; \textit{Prerequisites: }Forge Ring, Forge Epic Ring, 
\textit{protection from energy; Market Price:} 308,000 gp; \textit{Cost to Create: 
}154,000 gp + 13,080 XP. 

\textbf{Virtuous Good: }The wearer of this ring is constantly sheathed in a \textit{holy 
aura }effect. It bestows one negative level on any evil creature that puts it on. 
The negative level remains as long as the ring is worn and disappears when the 
ring is removed. This negative level never results in actual level loss, but it 
cannot be over-come in any way (including \textit{restoration }spells) while the 
ring is worn. 

\textit{Caster Level: }15th; \textit{Prerequisites: }Forge Ring, Forge Epic Ring, 
\textit{holy aura, }creator must be good; \textit{Market Price: }250,000 gp; \textit{Cost 
to Create: }125,000 gp + 12,500 XP. 

\textbf{Weaponbreaking: }A \textit{ring of weaponbreaking }is identical to a \textit{ring 
of ironskin, }and has one additional power. Any weapon that successfully strikes 
the wearer must also make a Fortitude saving throw (DC 20) or be shattered into 
pieces. 

\textit{Caster Level: }20th; \textit{Prerequisites: }Forge Ring, Forge Epic Ring, 
\textit{iron body, shatter; Market Price: }600,000 gp; \textit{Cost to Create: 
}300,000 gp + 16,000 XP. 

\textbf{Epic Psionics: }This special crystal ring comes in a variety of types useful 
only to psionic characters (characters who have power points per day). The wearer's 
total power points per day are increased, depending on the ring. The points granted 
are not bonus points---while the ring is worn, it actually increases the wearer's 
points per day, but a night's rest is required before gaining access to the increased 
power point per day total. (Power points are not actually stored in the ring, as 
would be the case for a crystal capacitor. Instead, the ring grants power points 
by magnifying the manifester's own power.) 

A \textit{ring of epic psionics V }increases the wearer's daily power points by 
43 points, a \textit{ring of epic psionics VI }grants 63 power points, a \textit{ring 
of epic psionics VII }grants 87 power points, a \textit{ring of epic psionics VIII 
}grants 115 points, and a \textit{ring of epic psionics IX }grants 147 power points. 

\textit{Manifester Level: }23rd \textit{(epic psionics }V), 26th \textit{(epic 
psionics V}I), 29th \textit{(epic psionics VI}I), 32nd \textit{(epic psionics VII}I), 
35th \textit{(epic psionics I}X); \textit{Prerequisites: }Craft Universal Item, 
Craft Epic Universal Item, \textit{great emulation; Market Price: }250,000 gp \textit{(epic 
psionics }V), 360,000 gp \textit{(epic psionics V}I), 490,000 gp \textit{(epic 
psionics VI}I), 640,000 gp \textit{(epic psionics VIII}), 810,000 gp \textit{(epic 
psionics I}X). 

\vspace{12pt}
\subsection*{{\LARGE{}RODS }}

\vspace{12pt}
\textbf{Table: Epic Rods }

\begin{tabular}{|>{\raggedright}p{30pt}|>{\raggedright}p{212pt}|>{\raggedright}p{71pt}|}
\hline
d\textbf{\%} & R\textbf{od } & \section*{M\textbf{arket Price }}\tabularnewline
\hline
01-08 & E\textit{pic spellcaster } & 245,000 gp \tabularnewline
\hline
09-16 & N\textit{ightmares } & 284,000 gp \tabularnewline
\hline
17-24 & E\textit{pic splendor } & 297,000 gp \tabularnewline
\hline
25-31 & T\textit{he path } & 306,870 gp \tabularnewline
\hline
32-38 & E\textit{pic cancellation } & 330,000 gp \tabularnewline
\hline
39-45 & E\textit{pic negation } & 446,000 gp \tabularnewline
\hline
46-51 & B\textit{esiegement } & 447,745 gp \tabularnewline
\hline
52-57 & F\textit{ortification } & 465,665 gp \tabularnewline
\hline
58-63 & E\textit{pic rulership } & 575,000 gp \tabularnewline
\hline
64-69 & I\textit{nvulnerability} & 600,000 gp \tabularnewline
\hline
70-75 & P\textit{aradise } & 610,000 gp \tabularnewline
\hline
76-80 & R\textit{estless death } & 625,000 gp \tabularnewline
\hline
81-85 & E\textit{xcellent magic } & 650,000 gp \tabularnewline
\hline
86 & W\textit{yrm }(white)  & 1,458,200 gp \tabularnewline
\hline
87 & W\textit{yrm }(brass)  & 1,458,200 gp \tabularnewline
\hline
88-90 & E\textit{pic absorption } & 1,500,000 gp \tabularnewline
\hline
91 & W\textit{yrm }(copper)  & 1,562,600 gp \tabularnewline
\hline
92 & W\textit{yrm }(black)  & 1,562,600 gp \tabularnewline
\hline
93 & W\textit{yrm }(bronze)  & 1,670,600 gp \tabularnewline
\hline
94 & W\textit{yrm }(green)  & 1,670,600 gp \tabularnewline
\hline
95 & W\textit{yrm }(blue)  & 1,782,200 gp \tabularnewline
\hline
96 & W\textit{yrm }(silver)  & 1,782,200 gp \tabularnewline
\hline
97 & W\textit{yrm }(gold)  & 1,897,400 gp \tabularnewline
\hline
98 & W\textit{yrm }(red)  & 1,897,400 gp \tabularnewline
\hline
99-100 & E\textit{pic might } & 4,293,432 gp \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\subsection*{EPIC ROD DESCRIPTIONS }

Standard epic rods are described below. 

\vspace{12pt}
\textbf{Besiegement:  }This rod functions as a +\textit{3 light mace. }In addition, 
it is useful for besieging fortifications. Whenever the wielder makes a charge 
attack, the rod improves to a +6 weapon. Twice per day, the rod can create a battering 
ram that lasts for 24 rounds. This ram can strike once per round, dealing 20 points 
of damage with each hit. It cannot be used to target individuals, only fortifications. 
It can be damaged by normal means (65 hp, AC 22), and \textit{disintegrate }or 
\textit{dispel magic }destroys it. The rod also has the following powers: \textit{Siege 
Engine: }One heavy catapult, two light catapults, or three siege towers may be 
generated with each use of this power. Each weapon created lasts for 12 hours. 
 The power can be used three times per day. Ammunition for 20 shots is included 
with each weapon created. 

\textit{Transmute Rock to Mud: }This power can be used three times per day (caster 
level 24th, save DC 17). 

\textit{Caster Level: }24th; \textit{Prerequisites: }Craft Rod, Craft Epic Rod, 
Craft Magic Arms and Armor, \textit{clenched fist, passwall,  telepathic bond, 
transmute rock to mud; Market Price: }447,745 gp; \textit{Cost to Create: }224,025 
gp + 14,474 XP. 

\textbf{Epic Absorption: }Like a \textit{rod of absorption, }this rod draws single-target 
or ray spells and spell-like abilities into itself, nullifying the effect and storing 
the potential spell levels until the wielder releases the energy in the form of 
spells of his or her own. Spells of any level (including those boosted beyond 9th 
level by metamagic) can be absorbed, although epic spells cannot. The rod absorbs 
a maximum of 150 spell levels and can thereafter only discharge any remaining potential 
it might have. The rod cannot be recharged. 

\textit{Caster Level: }23rd; \textit{Prerequisites: }Craft Rod, Craft Epic Rod, 
Empower Spell, Maximize Spell, empowered maximized \textit{spell turning; Market 
Price: }1,500,000 gp. 

\textbf{Epic Cancellation: }This rod's touch drains an item of all magical properties, 
including the magical energy in epic magic items (but not most artifacts). The 
item touched gets a Will saving throw (DC 26). If a creature is holding the magic 
item at the time, then the item can use the holder's Will save bonus in place of 
its own. In such cases, contact is established by making a melee touch attack roll. 
Upon draining three items, the rod becomes brittle and useless. Drained items can 
only be restored by \textit{wish, miracle, }or epic spells specifically designed 
to restore lost power. A \textit{rod of epic cancellation }can neutralize a normal 
\textit{sphere of annihilation }without itself being cancelled. 

\textit{Caster Level: }25th; \textit{Prerequisites: }Craft Rod, Craft Epic Rod, 
\textit{dispel magic; Market Price: }330,000 gp. 

\textbf{Epic Might: }This rod is similar to a \textit{rod of lordly might, }although 
it is far more powerful. It is larger than its normal counterpart, and it is constructed 
of adamantine rather than normal metal. It has six buttons, several spell-like 
functions, and several mundane uses, and it can also be used as a magic weapon 
of various sorts. 

The following spell-like functions of the rod can each be used once per day. • 

\textit{Dominate Person: }Touched foe is recipient of a \textit{dominate person 
}spell, if the wielder so commands (Will save DC 24). The wielder must choose to 
use this power and then succeed with a melee touch attack to activate the power. 
If the attack fails, the effect is lost. • 

\textit{Stun: }Upon command, all enemies viewing the rod are stunned, as per the 
\textit{power word, stun }spell (10-foot maximum range, Will save DC 24). Invoking 
this power is a standard action. • 

\textit{Damage: }Upon command, the rod deals 10d8 points of damage to an opponent 
on a successful touch attack and cures the wielder of a like amount of damage (Will 
save DC 26). The wielder must choose to use this power before attacking, as with 
\textit{dominate person.}

The following weapon uses of the rod have no limits on their use. • 

\parindent=3pt
In its normal form, the rod can be used as a  +\textit{6 heavy mace. • }

When button 1 is pushed, the rod becomes a +\textit{3 longsword of fiery blasting. 
• }

When button 2 is pushed, the rod becomes a +\textit{8 battleaxe. • }

When button 3 is pushed, the rod becomes a +\textit{10 shortspear }or +\textit{10 
longspear.}

\parindent=0pt
The following mundane uses of the rod also have no limits on their use. • 

\parindent=3pt
Climbing pole/ladder: When button 4 is pushed, a spike that can anchor in granite 
is extruded from the ball, while the other end sprouts three sharp hooks. The rod 
lengthens to anywhere between 5 and 150 feet in a single round, stopping when button 
4 is pushed again. Horizontal bars 3 inches long fold out from the sides, 1 foot 
apart, in staggered progression. The rod is firmly held by the spike and hooks 
and can bear up to 10,000 pounds. The wielder can retract the pole by pushing button 
5. • 

The ladder function can be used to force open doors. The wielder plants the rod's 
base 30 feet or less from the portal to be forced and in line with it, then pushes 
button 4. The force exerted has a Strength bonus of +24. • 

\parindent=7pt
When button 6 is pushed, the rod indicates magnetic north and gives the wielder 
a knowledge of his or her approximate depth beneath the surface or height above 
it. 

\parindent=0pt
\textit{Caster Level: }30th; \textit{Prerequisites: }Craft Rod, Craft Epic Rod, 
Craft Magic Arms and Armor, Craft Epic Magic Arms and Armor, \textit{cure critical 
wounds, inflict critical wounds, bull's strength, fireball, dominate person, power 
word stun; Market Price: }4,293,432 gp; \textit{Cost to Create: }2,151,372 gp + 
52,841 XP. 

\textbf{Epic Negation: }This rod negates the spell or spell-like function or functions 
of magic items, including epic magic items (but not artifacts). The wielder points 
the rod at the magic item, and a beam shoots forth to touch the target device, 
attacking as a ray (a ranged touch attack). The ray negates any currently active 
item function and has a 75\% chance to negate any other spell or spell-like functions 
of that device, regardless of the level or power of the functions, for 2d4 rounds. 
To negate instantaneous effects, the rod wielder needs to have readied an action. 
The target item gets no saving throw or means to resist this effect. The rod can 
function three times per day. 

\textit{Caster Level: }24th; \textit{Prerequisites: }Craft Rod, Craft Epic Rod, 
\textit{dispel magic, limited wish }or \textit{miracle; Market Price: }446,000 
gp. 

\textbf{Epic Rulership: }This rod appears to be a royal scepter worth at least 
25,000 gp in materials and work-manship alone. The wielder can command the obedience 
and fealty of creatures within 360 feet when he or she activates the device (a 
standard action). Creatures totaling 900 Hit Dice can be ruled, but creatures with 
Intelligence scores of 17 or higher are entitled to a Will saving throw (DC 29) 
to negate the effect. Ruled creatures obey the wielder as if he or she were their 
absolute sovereign. Still, if the wielder gives a command that is contrary to the 
nature of the creatures commanded, the magic is broken. The rod can be used for 
1,500 total minutes before crumbling to dust. This duration need not be continuous. 

\textit{Caster Level: }25th; \textit{Prerequisites: }Craft Rod, Craft Epic Rod, 
Improved Heighten Spell, improved heightened \textit{mass charm; Market Price: 
}575,000 gp; \textit{Cost to Create: }300,000 gp + 15,500 XP. 

\textbf{Epic Spellcaster: }This magnificent adamantine rod grants its wielder a 
+10 insight bonus on Spellcraft checks made to cast epic spells for as long as 
he or she holds or carries the rod. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Rod, Craft Epic Rod; 
\textit{Market Price: }245,400 gp. 

\textbf{Epic Splendor: }This rod grants its wielder a +8 enhancement bonus to Charisma 
for as long as he or she holds or carries the item. Three times per day, the rod 
creates and garbs him or her in clothing of the finest fabrics, plus adornments 
of fur and jewels. Apparel created by the rod can remain in existence up to 24 
hours. The value of the garb ranges from 70,000 to 100,000 gp (1d4+6 x10,000 gp)---10,000 
gp for the fabric, 50,000 gp for the furs, and the rest for jewel trim (maximum 
of 40 gems, maximum value 1,000 gp each). In addition, the rod can, once per week, 
create a palatial mansion in any floor plan the user desires. The mansion is palatial, 
able to accommodate up to 250 people, housing them in private chambers and serving 
them fine banquets. The mansion lasts for three days, after which time it, and 
everything originally in it (including items removed from the mansion), disappear. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Rod, Craft Epic Rod, 
\textit{fabricate, mage's magnificent mansion; Market Price: }297,400 gp; \textit{Cost 
to Create: }191,200 gp + 12,124 XP. 

\textbf{Excellent Magic: }Once per day when casting a nonepic or epic spell that 
has an experience point component, the rod supplies up to 2,000 XP, not the caster. 
If more experience points are required to cast the spell, the caster provides them. 
As a special use of the rod, the caster can substitute the power inherent in the 
rod for the experience point development cost of an epic spell. Doing so drains 
all the power from the rod, rendering it useless. 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Rod, Craft Epic Rod; 
\textit{Market Price: }650,000 gp; \textit{Cost to Create: }325,000 gp + 26,000 
XP. 

\textbf{Fortification: }This rod functions as a +\textit{3 light mace. }In addition, 
it is useful for the construction and defense of fortifications. Whenever the wielder 
is benefiting from cover, the rod improves that to nine-tenths cover (+10 cover 
AC bonus and +4 cover Reflex save bonus; failure on the save results in half damage 
and success results in no damage). Three times per day, the rod can create food 
and water, as per the cleric spell, for twenty-four people. The rod also has the 
following powers: \textit{Fortify: }Four times per day, a stone wall can be created 
that is 12 inches thick, 10 feet high, and 30 feet long. This wall has a parapet 
and battlements across the top. (Alternatively, the power can be used to mend a 
broken existing wall. The gap to be filled can be no more than 300 square feet 
in area.) In addition, once per day, a great door of iron can be created that is 
set into one of the newly created walls. This door, 4 inches thick, may be a double 
door, a drawbridge, or a portcullis, as chosen by the wielder. \textit{Siege Engine: 
}One heavy catapult, two light catapults, or three ballistae may be generated with 
each use of this power. The power can be used four times per day. Ammunition for 
20 shots is included with each weapon created. 

\textit{Caster Level: }24th; \textit{Prerequisites: }Craft Rod, Craft Epic Rod, 
Craft Magic Arms and Armor, \textit{inter-posing hand, create food and water, major 
creation, wall of iron, wall of stone; Market Price: }465,665 gp; \textit{Cost 
to Create: }231,985 gp + 14,634 XP. 

\textbf{Invulnerability: }When held, this rod grants its wielder the following 
powers. • 

\parindent=3pt
+5 enhancement bonus to natural armor. • 

+5 resistance bonus on saving throws. • 

\parindent=7pt
Damage reduction 15/adamantine. • 

\parindent=3pt
Immunity to critical hits. • 

Spell resistance 32. 

\parindent=0pt
\textit{Caster Level: }20th; \textit{Prerequisites: }Craft Rod, Craft Epic Rod, 
\textit{iron body, resistance, spell resistance; Market Price: }600,000 gp. 

\textbf{Nightmares: }Anyone who comes within 20 feet of the wielder of this rod 
feels a grave sense of unease. Each person so affected must make a Will save (DC 
17) or suffer the effects of a \textit{nightmare }spell the next time he or she 
falls asleep. The wielder is immune to this effect. Three times per day, the wielder 
can utter a command word that causes the rod to emit a horrid, inhuman scream. 
Up to twenty of the closest creatures within a 30-foot radius who hear this terrible 
wail believe that their worst nightmares have become reality and suffer the effects 
of a \textit{wail of the banshee }spell (DC 23). 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Rod, Craft Epic Rod, 
\textit{nightmare, permanency, wail of the banshee; Market Price: }284,000 gp. 

\textbf{Paradise: }This item creates a nondimensional space, similar in effect 
to a \textit{rod of securit}y. However, the rod's possessor and up to 999 other 
creatures can stay in complete safety for up to 1,000 days divided by the number 
of creatures affected. Natural healing takes place at five times the normal pace. 
The rod functions like its nonepic counterpart in all other ways.

\textit{Caster Level: }24th; \textit{Prerequisites: }Craft Rod, Craft Epic Rod, 
Improved Heighten Spell, improved heightened \textit{gate; Market Price: }610,000 
gp. 

\textbf{The Path: }This rod serves as an aid to trailblazing and travel. It grants 
the wielder a +30 enhancement bonus on Wilderness Lore checks for tracking and 
Intuit Direction checks. The handle of the rod is hollow, functioning like a telescope. 
When the wielder peers through it, the limits of vision are three times normal 
(and spotting distances for encounters are tripled). In addition, the telescoping 
handle enables the wielder to view things as though affected by a \textit{true 
seeing }spell. The rod also has the following powers. \textit{Map: }Three times 
per day a section of the rod unrolls like a scroll from a tube, revealing a map 
of the surrounding area, centered on the location of the rod. The area shown on 
the map covers an area as small as 50 feet in radius to as large as 24 miles in 
radius, zooming in or out with a set of command words. The map reveals natural 
topography and all types of structures (even hidden ones), but it will not show 
the location of creatures. 

\textit{Passage: }Three times per day, this power allows the wielder and up to 
five others in a 20-foot radius to move unhindered through natural plant growth 
or bodies of water (as per the \textit{freedom of movement }spell). 

\textit{Bridge: }Once per day, this power allows the user to create a 5-foot-wide, 
40-foot-long stone cause-way across chasms and canyons. The bridge created lasts 
for 1 hour.

\textit{Pass without Trace: }Once per day, this power can be used on the wielder 
and twenty others, for 21 minutes. It is otherwise as the spell of the same name 
(caster level 24th). 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Rod, Craft Epic Rod, 
\textit{find the path, freedom of movement, pass without trace, wall of stone; 
Market Price: }306,870 gp. 

\textbf{Restless Death: }The holder of this rod can rebuke or control undead as 
if he or she were four levels higher. (The rod doesn't grant the ability to rebuke 
or control undead if the holder doesn't already have it.) In addition, the wielder 
may speak a command word to cast \textit{animate dead}. Any skeletons or zombies 
animated by the rod's power are automatically controlled by the rod, up to the 
rod's maximum limit of 42 HD, and they follow the orders of the rod's wielder. 
These undead don't count toward the wielder's limit of controlled undead. Finally, 
the wielder can speak a second command word to cast \textit{slay living }(heightened 
to 10th level; DC 25). 

\textit{Caster Level: }21st; \textit{Prerequisites: }Craft Rod, Craft Epic Rod, 
\textit{animate dead, control undead, slay living; Market Price: }625,000 gp. 

\textbf{Wyrm: }Any of the various rods in this series functions as a +\textit{5 
quarterstaff. }Upon casting the rod to the ground (a standard action) and uttering 
a command word, the rod grows into a specific type of dragon (depending on the 
specific type of rod) by the end of the round. The dragon created is a wyrm and 
obeys the commands of the owner. The dragon returns to rod form (a full-round action) 
whenever the wielder desires, or whenever it moves farther than 500 feet from the 
owner. If the dragon form is slain, it returns to rod form and cannot be activated 
again for three days. A \textit{wyrm rod }only functions if the possessor is of 
the same alignment as the dragon type. 

\textit{Caster Level: }34th (white or brass), 35th (black or copper), 36th (green 
or bronze), 37th (blue or silver), 38th (red or gold); \textit{Prerequisites: }Craft 
Rod, Craft Epic Rod, Craft Magic Arms and Armor, Craft Epic Magic Arms and Armor, 
Improved Heighten Spell, improved heightened \textit{polymorph, }creator must be 
same alignment as dragon type; \textit{Market Price: }1,458,200 gp (brass or white), 
1,562,600 gp (black or copper), 1,670,600 gp (bronze or green), 1,782,200 gp (blue 
or silver), 1,897,400 gp (gold or red); \textit{Cost to Create: }729,400 gp + 24,576 
XP (brass or white), 781,600 gp + 25,620 XP (black or copper), 835,600 gp + 26,700 
XP (bronze or green), 891,400 gp + 27,816 XP (blue or silver), 949,000 gp + 28,968 
XP (gold or red). 

\newpage

\end{document}
