%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{array}
\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

This material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

{\LARGE{}EPIC MONSTERS (A-E)}

\vspace{12pt}
\section*{{\LARGE{}ABOMINATION }}

\vspace{12pt}
BUILDING ABOMINATIONS 

Abominations are a grouping of the outsider type in the same way that demons are 
a grouping of the outsider type. The accompanying table indicates the average values 
an abomination of a given size might have for its basic physical scores. Abominations 
also share several other characteristics.

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{49pt}|>{\raggedright}p{22pt}|>{\raggedright}p{22pt}|>{\raggedright}p{22pt}|>{\raggedright}p{47pt}|>{\raggedright}p{19pt}|>{\raggedright}p{18pt}|>{\raggedright}p{20pt}|>{\raggedright}p{19pt}|}
\hline
\section*{S\textbf{ize}} & S\textbf{tr} & D\textbf{ex} & C\textbf{on} & \#\textbf{ 
of HD} & S\textbf{lam} & B\textbf{ite} & C\textbf{law} & G\textbf{ore}\tabularnewline
\hline
Fine & 12-13 & 26-27 & 12-13 & 4d8-9d8--- &  & 1d6 & 1d4 & 1d4\tabularnewline
\hline
Diminutive & 14-15 & 24-25 & 14-15 & 7d8-13d8 & 1d4 & 1d8 & 1d6 & 1d6\tabularnewline
\hline
Tiny & 18-19 & 22-23 & 16-17 & 10d8-21d8 & 1d6 & 2d6 & 1d8 & 1d8\tabularnewline
\hline
Small & 26-27 & 20-21 & 20-21 & 19d8-33d8 & 1d8 & 2d8 & 2d6 & 2d6\tabularnewline
\hline
Medium-size & 34-35 & 18-19 & 24-25 & 27d8-38d8 & 2d6 & 4d6 & 2d8 & 2d8\tabularnewline
\hline
Large & 42-43 & 16-17 & 28-29 & 36d8-50d8+ & 2d8 & 4d8 & 4d6 & 4d6\tabularnewline
\hline
Huge & 50-51 & 14-15 & 32-33 & 47d8-58d8+ & 4d6 & 8d6 & 4d8 & 4d8\tabularnewline
\hline
Gargantuan & 58-59 & 12-13 & 36-37 & 56d8-70d8+ & 4d8 & 8d8 & 8d6 & 8d6\tabularnewline
\hline
Colossal & 66-67 & 10-11 & 40-41 & 71d8+ & 8d6 & 16d6 & 8d8 & 8d8\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\textbf{Abomination Traits}

All abominations are born directly (or indirectly) from a god and some lesser creature 
(or idea), but none are favored, wanted, or loved. Still, they all share a tiny 
spark of deific energy, which grants them the qualities described in below. (Note: 
deity rules are used\textit{, }abominations are rank 0 deities.) 

\textbf{Immunities (Ex):} Abominations are immune to polymorphing, petrification, 
or any form-altering attack. They are not subject to energy drain, ability drain, 
or ability damage. They are immune to mind-affecting effects (charms, compulsions, 
phantasms, patterns, and morale effects), and are immune to one of five energy 
types (specific to the abomination). 

\textbf{Resistances (Ex):} Abominations have fire resistance 20, cold resistance 
20, and damage reduction of at least 10/epic (certain abominations have higher 
damage reduction). Abominations all have significant spell resistance. Abominations 
resist detection, and are all treated as if affected by a \textit{nondetection 
}spell of a caster level equal to the abomination's HD. 

\textbf{Special Qualities (Ex):} All abominations have the spell-like ability to 
use \textit{true seeing }at will. Abominations are not subject to death from massive 
damage, and they have maximum hit points per Hit Die. Abominations have the blindsight 
extraordinary ability to a range of 500 feet. Abominations can choose both nonepic 
and epic feats as part of their feat selection. 

\textbf{Telepathy (Su):} Abominations can communicate telepathically with any creature 
within 1,000 feet that has a language. 

\textit{\textbf{Summon Creature }}\textbf{(Sp):} Abominations can summon creatures 
associated either with the portfolio of their godly progenitor or with their imprisonment. 
Summoned creatures serve the abomination without question. Summoned creatures automatically 
return whence they came after 1 hour, or sooner if slain. See specific abomination 
entries for additional details on summoning. 

\textbf{Fast Healing and Regeneration: }Most abominations have some measure of 
fast healing and regeneration, ranging from 5 to 55 for both abilities. Usually, 
higher HD abominations have higher values for fast healing and regeneration, though 
this not always the case. If an abomination has regeneration, it is usually subject 
to normal damage from at least two sources, one of which is the antithesis for 
their subtype, and the other of which is somehow tied to some unique feature of 
the abomination. 

\textbf{Natural Armor: }All abominations have some degree of natural armor bonus 
ranging from +10 to +100. Usually, higher HD abominations have higher natural armor 
bonuses, but this is not always the case. 

\textbf{Spell-Like Abilities: }All abominations have access to a variety of spells 
in the form of spell-like abilities, which they can use as 20th- or higher-level 
casters, depending on the abomination. No rule governs how many or how few spell-like 
abilities an abomination may claim. 

\textbf{Unique Abilities: }All abominations have at least one or more unique abilities 
tied to their godly parent's portfolio, or to the manner of their incarceration. 

\textbf{Spell Resistance: }As a general rule of thumb, abominations have spell 
resistance equal to their CR +12. 

\vspace{12pt}
Abominations speak Abyssal, Celestial, Infernal, and often a language associated 
with their imprisonment. 

{\LARGE{}ANAXIM}

\begin{tabular}{|>{\raggedright}p{65pt}|>{\raggedright}p{260pt}|}
\hline
 & Medium-Size Construct (Extraplanar, Lawful) \tabularnewline
\hline
H\textbf{it Dice:} & 38d10 (420 hp) \tabularnewline
\hline
I\textbf{nitiative:} & +7 (Dex) \tabularnewline
\hline
S\textbf{peed:} & 60 ft., fly 200 ft. (perfect) \tabularnewline
\hline
A\textbf{C:} & 37 (+7 Dex, +20 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +28/+40\tabularnewline
\hline
A\textbf{ttack:} & Spinning blade +43 (2d6+12/19-20 (plus 1d6 on critical)) or 
electricity ray +35 ranged touch\tabularnewline
\hline
F\textbf{ull Attack:} & 2 spinning blades +43 (2d6+12/19-20(plus 1d6 on critical)) 
melee, 2 slams +35 melee (2d6+6), shocking touch +35 melee touch (2d6+6); or electricity 
ray +35 ranged touch (10d6 electricity), 6 spikes +30 ranged (2d6+12) (120 ft. 
range increment)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} & Rend 4d6+18, sonic blast, spell-like abilities, summon 
iron golem \tabularnewline
\hline
S\textbf{pecial Qualities:} & Abomination traits, magic immunity, construct traits, 
fast healing 15, SR 34, damage reduction 10/chaotic and epic and adamantine\tabularnewline
\hline
S\textbf{kills:} & Climb +53, Jump +65\tabularnewline
\hline
F\textbf{eats:} & Blind-Fight, Cleave, Great Cleave, Improved Bull Rush, Improved 
Critical (blade), Improved Initiative, Improved Sunder, Power Attack, Weapon Focus 
(blade)\tabularnewline
\hline
E\textbf{pic Feats:} & Devastating Critical (blade), Epic Toughness, Epic Weapon 
Focus (blade), Overwhelming Critical (blade)\tabularnewline
\hline
S\textbf{aves:} & Fort +12, Ref +19, Will +17 \tabularnewline
\hline
A\textbf{bilities:} & Str 35, Dex 25, Con ---, Int 10, Wis 20, Cha 20 \tabularnewline
\hline
C\textbf{limate/Terrain:} & Any \tabularnewline
\hline
O\textbf{rganization:} & Solitary, binary (pair), or command line (2-5 anaxims 
and 5-12 iron golems) \tabularnewline
\hline
C\textbf{hallenge Rating:} & 22 \tabularnewline
\hline
T\textbf{reasure:} & Standard \tabularnewline
\hline
A\textbf{lignment:} & Always lawful neutral \tabularnewline
\hline
A\textbf{dvancement:} & 39-48 HD (Large); 49-55 HD (Huge); 56-70 HD (Gargantuan); 
71-140 HD (Colossal)\tabularnewline
\hline
\end{tabular}

An anaxim's natural weapons are treated as epic and lawful-aligned for the purpose 
of overcoming damage reduction.

\section*{Combat\textbf{ }}

\textbf{Sonic Blast (Ex):} As a standard action, an anaxim can emit a 60-foot cone 
of sonic energy that deals 20d6 points of sonic damage to all creatures that fail 
a Reflex save (DC 29); those that succeed take half damage. 

The save DC is Constitution-based

\textbf{Rend (Ex):} If the anaxim hits with both spinning blades, it slices particularly 
well. This attack automatically deals an additional 4d6+18 points of damage. 

\textbf{Spell-Like Abilities:} At will---\textit{greater dispel magic, displacement} 
(DC 18)\textit{, greater invisibility} (DC 19)\textit{, ethereal jaunt. }Caster 
level 22nd. The save DCs are Charisma-based.

\textit{\textbf{Summon Iron Golem }}\textbf{(Sp):} An anaxim can summon an iron 
golem up to four times per day. 

\textbf{Abomination Traits:} Immune to polymorphing, petrification, and other form-altering 
attacks; not subject to energy drain, ability drain, ability damage, or death from 
massive damage; immune to mind-affecting effects; fire resistance 20; cold resistance 
20; nondetection; true seeing at will; blindsight 500 ft.; telepathy out to 1,000 
ft. 

\textbf{Construct Traits:} Immune to mind-affecting effects (charms, compulsions, 
phantasms, patterns, and morale effects), and to poison, \textit{sleep, }paralysis, 
stunning, disease, death effects, necromantic effects, and any effect that requires 
a Fortitude save unless it also works on objects. Cannot heal damage (though regeneration 
and fast healing still apply, if present). Not subject to critical hits, subdual 
damage, ability damage, ability drain, or energy drain. Not at risk of death from 
massive damage, but destroyed when reduced to 0 hit points or less; cannot be raised 
or resurrected. Darkvision 60 ft. 

\vspace{12pt}
{\LARGE{}ATROPAL }

\begin{tabular}{|>{\raggedright}p{67pt}|>{\raggedright}p{258pt}|}
\hline
 & Large Undead (Evil, Extraplanar, Lawful) \tabularnewline
\hline
H\textbf{it Dice:}  & 66d12 (792 hp) \tabularnewline
\hline
I\textbf{nitiative:}  & +6 (+2 Dex, +4 Improved Initiative) \tabularnewline
\hline
S\textbf{peed:}  & 5 ft., fly 240 ft. (perfect) \tabularnewline
\hline
A\textbf{C:}  & 51 (-1 size, +2 Dex, +40 natural) \tabularnewline
\hline
B\textbf{ase Attack /Grapple:} & +33/+53\tabularnewline
\hline
A\textbf{ttack:}  & Touch +49 (2d6 Con drain/19-20) melee touch\tabularnewline
\hline
F\textbf{ull Attack:}  & 2 touches +49 (2d6 Con drain/19-20) melee touch, eye ray 
+30 (negative level damage/19-20) ranged touch \tabularnewline
\hline
S\textbf{pace/Reach:}  & 10 ft/10 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:}  & Constitution drain, energy drain (2d4 negative levels, 
Fort DC 59), spell-like abilities, \textit{summon nightcrawler }\tabularnewline
\hline
S\textbf{pecial Qualities:}  & Abomination traits, undead traits, rebuke/command 
undead, regeneration 20, SR 42, DR 15/good and epic and silver, negative energy 
aura \tabularnewline
\hline
S\textbf{aves:}  & Fort +22, Ref +26, Will +43 \tabularnewline
\hline
A\textbf{bilities:}  & Str 43, Dex 15, Con -, Int 28, Wis 22, Cha 42 \tabularnewline
\hline
S\textbf{kills:}  & Bluff +85, Concentration +85, Diplomacy +97, Hide -2, Jump 
+1, Intimidate +91, Knowledge (arcana, history, religion, the planes) +78, Listen 
+77, Search +78, Sense Motive +75, Spellcraft +84, Spot +77\tabularnewline
\hline
F\textbf{eats:}  & Alertness, Cleave, Dodge, Expertise, Great Cleave, Improved 
Bull Rush, Improved Critical (touch), Improved Critical (eye ray), Improved Initiative, 
Iron Will, Lightning Reflexes, Mobility, Power Attack, Quicken Spell-Like Ability, 
Run, Spring Attack, Weapon Focus (touch), Weapon Focus (eye ray), Whirlwind Attack 
\tabularnewline
\hline
E\textbf{pic Feats:}  & Devastating Critical (touch), Overwhelming Critical (touch), 
Undead Mastery, Zone of Animation \tabularnewline
\hline
C\textbf{limate/Terrain:}  & Any \tabularnewline
\hline
O\textbf{rganization:}  & Solitary \tabularnewline
\hline
C\textbf{hallenge Rating:}  & 30 \tabularnewline
\hline
T\textbf{reasure:}  & Standard \tabularnewline
\hline
A\textbf{lignment:}  & Always lawful evil \tabularnewline
\hline
A\textbf{dvancement:}  & 67-80 HD (Large); 81-100 HD (Huge) \tabularnewline
\hline
\end{tabular}

An atropal's natural weapons are treated as epic, evil, and lawful-aligned for 
the purpose of overcoming damage reduction.

\section*{Combat\textbf{ }}

\textbf{Spell-Like Abilities:} At will---\textit{animate dead, blasphemy, create 
greater undead, create undead, cone of cold, desecrate, greater dispelling, finger 
of death, improved invisibility, plane shift, slay living, speak with dead, spectral 
hand, teleport with-out error, unholy aur}a; 5/day---\textit{haste, project image, 
weird. }Caster level 30th; save DC 26 + spell level.

The save DCs are Charisma-based

\textbf{Rebuke/Command Undead (Su):} Atropals can rebuke or command undead as a 
cleric with a level equal to the atropal's HD + 6.

\textbf{Negative Energy Aura (Su):} A 30-foot-radius spread negative energy aura 
spreads from each atropal. All undead in the field (including the atropal) are 
treated as if having turn resistance +20 and a negative energy version of fast 
healing 20. Living creatures in the aura are treated as having ten negative levels 
unless they have some sort of negative energy protection or protection from evil. 
Creatures with 10 or fewer HD or levels perish (and, at the atropal's option, rise 
as spectres under the atropal's command 1 minute later). 

\textbf{Constitution Drain (Su):} When the atropal hits a living opponent with 
a touch attack, the opponent takes 5 points of permanent Constitution drain, or 
10 points on a critical hit. The atropal heals 20 points of damage, or 40 points 
on a critical hit, whenever it drains Constitution, gaining any excess as temporary 
hit points. These temporary hit points last a maximum of 1 hour. The attack allows 
a Fortitude save (DC 59). The DC is Charisma-based.

\textbf{Energy Drain (Su): }When the atropal hits with a ranged touch attack (a 
ray of darkness that it shoots from one eye to a range of 400 feet), the resultant 
energy drain bestows four negative levels, or eight negative levels on a critical 
hit. For each negative level bestowed on an opponent, the atropal heals 10 points 
of damage, or 20 on a critical hit, gaining any excess as temporary hit points. 
These temporary hit points last a maximum of 1 hour. After 24 hours have passed, 
the afflicted opponent must attempt a Fortitude save (DC 59) for each negative 
level. If successful, the negative level goes away with no harm to the creature; 
otherwise, the creature's level decreases by one. The DC is Charisma-based.

\textbf{Regeneration (Ex):} Atropals take normal damage from good weapons or sentient 
weapons (or otherwise living weapons). 

\textit{\textbf{Summon Nightcrawler }}\textbf{(Sp): }Five times per day an atropal 
can summon a nightcrawler. 

\textbf{Abomination Traits:} Immune to polymorphing, petrification, and other form-altering 
attacks; not subject to energy drain, ability drain, ability damage, or death from 
massive damage; immune to mind-affecting effects; fire resistance 20; cold resistance 
20; nondetection; true seeing at will; blindsight 500 ft.; telepathy out to 1,000 
ft. 

\textbf{Undead Traits:} Immune to poison, \textit{sleep, }paralysis, stunning, 
disease, death, effects, necromantic effects, mind-affecting effects, and any effect 
requiring a Fortitude save unless it also works on objects. Not subject to critical 
hits, subdual damage, ability damage, ability drain, or energy drain. Negative 
energy heals. Not at risk of death from massive damage, but destroyed at 0 hit 
points or less. Darkvision 60 ft. Cannot be raised; resurrection works only if 
creature is willing. 

\vspace{12pt}
{\LARGE{}CHICHIMEC}

\begin{tabular}{|>{\raggedright}p{63pt}|>{\raggedright}p{262pt}|}
\hline
  & Medium-Size Outsider (Air, Evil, Extraplanar) \tabularnewline
\hline
H\textbf{it Dice:}  & 27d8+189 (425 hp) \tabularnewline
\hline
I\textbf{nitiative:}  & +7 \tabularnewline
\hline
S\textbf{peed:}  & 5 ft., fly 200 ft. (perfect) \tabularnewline
\hline
A\textbf{C:}  & 39 (+7 Dex, +22 natural) \tabularnewline
\hline
B\textbf{ase Attack /Grapple:}  & +27/+39\tabularnewline
\hline
A\textbf{ttack:} &  Primary wing buffet +39 (2d6+12) melee\tabularnewline
\hline
F\textbf{ull Attack:} &  2 primary wing buffets +39 (2d6+12) melee, 6 secondary 
wing buffets +37 (1d6+6) melee, tail slam +37 (1d6+6 plus Cha drain) melee\tabularnewline
\hline
S\textbf{pace/Reach:} &  5 ft./5 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} &  Spell-like abilities, summon air elemental, Charisma 
drain \tabularnewline
\hline
S\textbf{pecial Qualities:} &  Abomination traits, fast healing 10, SR 33, DR 10/ 
good and epic, electricity immunity \tabularnewline
\hline
S\textbf{aves:} &  Fort +22, Ref +22, Will +17 \tabularnewline
\hline
A\textbf{bilities:} &  Str 34, Dex 25, Con 24, Int 12, Wis 14, Cha 30 \tabularnewline
\hline
S\textbf{kills:} &  Diplomacy +14, Hide +37, Intimidate +40, Jump -3, Knowledge 
(religion) +31, Knowledge (the planes) +31, Listen +34, Move Silently +37, Search 
+31, Sense Motive +32, Spot +34\tabularnewline
\hline
F\textbf{eats:} &  Alertness, Dodge, Flyby Attack, Improved Bull Rush, Mobility, 
Multiattack, Power Attack \tabularnewline
\hline
E\textbf{pic Feats:} &  Blinding Speed (x 2), Epic Toughness\tabularnewline
\hline
C\textbf{limate/Terrain:} &  Any \tabularnewline
\hline
O\textbf{rganization:} &  Solitary, pair, or flock (6-9) \tabularnewline
\hline
C\textbf{hallenge Rating:} &  21 \tabularnewline
\hline
T\textbf{reasure:} &  None \tabularnewline
\hline
A\textbf{lignment:} &  Always neutral evil \tabularnewline
\hline
A\textbf{dvancement:} &  28-38 HD (Medium-size); 39-50 HD (Large) \tabularnewline
\hline
\end{tabular}

An chichimec's natural weapons are treated as epic and evil-aligned for the purpose 
of overcoming damage reduction.

\section*{Combat\textbf{ }}

\textbf{Spell-Like Abilities:} At will---\textit{darkness, telekinesis; }3/da\textit{y---improved 
invisibility, control weather, call lightning, lightning bolt, chain lightning, 
wail of the banshe}e. Caster level 20th; save DC 20 + spell level. The DC is Charisma-based.

\textit{\textbf{Summon Air Elemental }}\textbf{(Sp):} Three times per day a chichimec 
can summon an elder air elemental. 

\textbf{Charisma Drain (Su):} This effect permanently reduces a living opponent's 
Charisma score by 2 points when the chichimec hits with a tail slam, or 4 points 
on a critical hit. The chichimec heals 10 points of damage, or 20 on a critical 
hit, whenever it drains Charisma, gaining any excess as temporary hit points. These 
temporary hit points last a maximum of 1 hour. The attack allows a Fortitude save 
(DC 33)---on a successful save, only 1 point of Charisma is drained and the chichimec 
heals 5 points of damage. The DC is Charisma-based.

\textbf{Abomination Traits:} Immune to polymorphing, petrification, and other form-altering 
attacks; not subject to energy drain, ability drain, ability damage, or death from 
massive damage; immune to mind-affecting effects; fire resistance 20; cold resistance 
20; nondetection; true seeing at will; blindsight 500 ft.; telepathy out to 1,000 
ft. 

\vspace{12pt}
{\LARGE{}DREAM LARVA }

\begin{tabular}{|>{\raggedright}p{63pt}|>{\raggedright}p{262pt}|}
\hline
  & Large Outsider (Chaotic, Evil, Extraplanar) \tabularnewline
\hline
H\textbf{it Dice:} &  40d8+360 (700 hp) \tabularnewline
\hline
I\textbf{nitiative:} &  +3 (Dex) \tabularnewline
\hline
S\textbf{peed:} &  80 ft.; fly 240 ft. (perfect) \tabularnewline
\hline
A\textbf{C:} &  52 (-1 size, +3 Dex, +40 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} &  +40/+60\tabularnewline
\hline
A\textbf{ttack:} &  Bite +56 (4d8+16)melee\tabularnewline
\hline
F\textbf{ull Attack:} &  Bite +56 (4d8+16) melee, 1 gore +51 (4d6+8) melee, 4 pincers 
+53 (4d6+8) melee, 4 claws +53 (4d6+8) melee\tabularnewline
\hline
S\textbf{pace/Reach:} &  10 ft./10 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} &  Worst nightmare, improved grab, sending, spell-like 
abilities, summon nightwalker\tabularnewline
\hline
S\textbf{pecial Qualities:} &  Abomination traits, sonic immunity, regeneration 
15, fast healing 15, SR 44, DR 15/ Good and Epic or Lawful and Epic \tabularnewline
\hline
S\textbf{aves:} &  Fort +31, Ref +25, Will +29 \tabularnewline
\hline
A\textbf{bilities:} &  Str 42, Dex 17, Con 29, Int 16, Wis 24, Cha 36 \tabularnewline
\hline
S\textbf{kills:} & Concentration +52, Craft (dreamweaving) +46, Diplomacy +60, 
Escape Artist +46, Hide +46, Jump +36, Knowledge (arcana) +46, Listen +50, Move 
Silently +46, Search +46, Sense Motive +50, Spot +50\tabularnewline
\hline
F\textbf{eats:} &  Cleave, Combat Reflexes, Flyby Attack, Great Cleave, Improved 
Sunder, Power Attack, Quicken Spell-Like Ability, Weapon Focus (pincers), Weapon 
Focus (gore), Weapon Focus (bite), Weapon Focus (claw) \tabularnewline
\hline
E\textbf{pic Feats:} &  Epic Toughness, Epic Weapon Focus (pincers), Epic Weapon 
Focus (claw)\tabularnewline
\hline
C\textbf{limate/Terrain:} &  Any \tabularnewline
\hline
O\textbf{rganization:} &  Solitary, pair, or solitary plus 1-4 nightwalkers \tabularnewline
\hline
C\textbf{hallenge Rating:} &  31 \tabularnewline
\hline
T\textbf{reasure:} &  Standard \tabularnewline
\hline
A\textbf{lignment:} &  Always chaotic evil \tabularnewline
\hline
A\textbf{dvancement:} &  41-65 HD (Large); 66-84 HD (Huge); 85-110 HD (Gargantuan) 
\tabularnewline
\hline
\end{tabular}

A dream larvas's natural weapons are treated as epic, evil, and chaotic-aligned 
for the purpose of overcoming damage reduction.

Combat\textbf{ }

\textbf{Worst Nightmare (Su):} Each time a living creature first views a specific 
dream larva from a distance of 30 feet or less (or from a scrying effect), the 
subject sees the image of the most fearsome creature imaginable. This is not an 
illusion or phantasm; the dream larva truly becomes, for just that instant, the 
subject's worst nightmare. Even if simultaneously viewed by dozens of different 
creatures, the dream larva appears differently to each one of them. Creatures immune 
to fear or mind-affecting effects (or warded by \textit{protection from evil }or 
\textit{death ward }spells) are immune to worst nightmare; all others must make 
a Will save (DC 43) or die from the supernatural horror revealed. Survivors (and 
those resurrected) are immune to the effect from that individual dream larva in 
the future. The DC is Charisma-based.

\textbf{Spell-Like Abilities:} At will---\textit{fly, haste, nightmare, prismatic 
spray; }2/day---\textit{dreamscape}(epic spell)\textit{. }Caster level 31st; save 
DC 23 + spell level. The DC is Charisma-based.

\textbf{Improved Grab (Ex): }If the dream larva hits with a claw or pincer, it 
deals normal damage and attempts to start a grapple as a free action without provoking 
an attack of opportunity. Dream larvae can use improved grab on a creature of any 
size. The dream larva has the option to conduct the grapple normally, simply use 
the claw or pincer to hold the opponent, or use sending (see below) on the opponent. 
Each successful grapple check it makes during successive rounds automatically deals 
the damage listed for the attack that established the hold, unless it used sending 
on the victim.

\textbf{Sending (Su):} Grappled victims may be physically sent into a nightmare, 
at the dream larva's option, on the dream larva's next action after establishing 
the grapple. Victims must make a Will save (DC 43) or be apparently swarmed by 
thousands of worms making up the dream larva's body. The victim is actually cast 
into a shrieking maelstrom of dream where it can take no actions but observe. The 
victim physically returns 2d4 rounds later, apparently ejected from the dream larva's 
body (but appearing even if if the dream larva has left or is otherwise absent). 
The victim of the nightmare has taken 4d6 points of temporary Wisdom damage, but 
is otherwise free to act on the round it returns. If the victim takes more Wisdom 
damage than it has points of Wisdom, the extra points are instead treated as temporary 
Constitution damage. The DC is Charisma-based.

\textit{\textbf{Summon Nightwalker }}\textbf{(Sp): }Five times per day, a dream 
larva can summon a nightwalker. 

\textbf{Abomination Traits:} Immune to polymorphing, petrification, and other form-altering 
attacks; not subject to energy drain, ability drain, ability damage, or death from 
massive damage; immune to mind-affecting effects; fire resistance 20; cold resistance 
20; nondetection; true seeing at will; blindsight 500 ft.; telepathy out to 1,000 
ft. 

\textbf{Regeneration (Ex): }Dream larvae take normal damage from good or lawful 
weapons or weapons forged by a sleepwalking weaponsmith. 

\vspace{12pt}
{\LARGE{}HECATONCHEIRES}

\begin{tabular}{|>{\raggedright}p{67pt}|>{\raggedright}p{258pt}|}
\hline
  & Huge Outsider (Chaotic, Evil, Extraplanar) \tabularnewline
\hline
H\textbf{it Dice:} &  52d8+572 (1028 hp) \tabularnewline
\hline
I\textbf{nitiative:} &  +10 (+2 Dex, +8 Superior Initiative) \tabularnewline
\hline
S\textbf{peed:} &  100 ft. \tabularnewline
\hline
A\textbf{C:} &  70 (-2 size, +30 natural, +20 insight, +12 armor [\textit{+5 half 
plat}e]) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} &  \textbf{+}52/+80\tabularnewline
\hline
A\textbf{ttack:} &  Greatsword +73 (2d6+20/17-20) melee; or boulder +55 (2d8+20/19-20) 
ranged\tabularnewline
\hline
F\textbf{ull Attack:} &  100 greatswords +73 (2d6+20/17-20) melee; or 100 boulders 
+55 (2d8+20/19-20) ranged\tabularnewline
\hline
S\textbf{pace/Reach:} &  15 ft./15 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} &  Superior multiweapon fighting, spell-like abilities, 
summon hecatoncheires \tabularnewline
\hline
S\textbf{pecial Qualities:} &  Abomination traits, electricity immunity, regeneration 
40, fast healing 50, SR 70, DR 20/Good and Epic and Cold Iron \tabularnewline
\hline
S\textbf{aves:} &  Fort +39, Ref +30, Will +27 \tabularnewline
\hline
A\textbf{bilities:} &  Str 50, Dex 15, Con 32, Int 10, Wis 8, Cha 24 \tabularnewline
\hline
S\textbf{kills:} & Climb +75, Diplomacy +13, Hide -12, Intimidate +40, Jump +89, 
Knowledge (history) +33, Listen +104, Search +105, Sense Motive +49, Spot +104, 
Survival +54\tabularnewline
\hline
F\textbf{eats:} &  Cleave, Combat Reflexes, Great Cleave, Improved Critical (greatsword), 
Improved Critical (boulder), Improved Initiative, Multidexterity, Multiweapon Fighting, 
Power Attack, Sunder, Weapon Focus (greatsword), Weapon Focus (boulder) \tabularnewline
\hline
E\textbf{pic Feats:} &  Epic Toughness (x2), Epic Weapon Focus (greatsword), Epic 
Weapon Focus (boulder), Multiweapon Rend \tabularnewline
\hline
C\textbf{limate/Terrain:} &  Any \tabularnewline
\hline
O\textbf{rganization:} &  Solitary or pair \tabularnewline
\hline
C\textbf{hallenge Rating:} &  57 \tabularnewline
\hline
T\textbf{reasure:} &  Standard \tabularnewline
\hline
A\textbf{lignment:} &  Always chaotic evil \tabularnewline
\hline
A\textbf{dvancement:} &  53-58 HD (Huge); 59-70 HD (Gargantuan); 71-140 HD (Colossal) 
\tabularnewline
\hline
\end{tabular}

A hecatoncheires' natural weapons are treated as epic, evil, and chaotic-aligned 
for the purpose of overcoming damage reduction.

Combat\textbf{ }

\textbf{Superior Multiweapon Fighting (Ex):} A hecatoncheires fights with a greatsword 
or a boulder in each hand. The hecatoncheires does not suffer an attack or damage 
penalty for attacking with one hundred weapons. However, the press of limbs prevents 
the creature from making iterative attacks with any of its arms, nor can it make 
more than ten attacks against a Small or smaller creature, fifteen attacks against 
a Medium-size creature, or twenty attacks against a Large creature in the same 
action (it can make all its attacks against a Huge or larger creature in one action). 

\textbf{Skills:} A hecatoncheires' fifty heads give it a +50 racial bonus on Listen, 
Spot, and Search checks. 

\textbf{Spell-Like Abilities:} At will--- \textit{greater magic weapon, fly, shield. 
}Caster level 50th; save DC 17 + spell level. The DC is Charisma-based.

\textit{\textbf{Summon Hecatoncheires }}\textbf{(Sp):} A hecatoncheires can summon 
one other hecatoncheires once per day, though is loath to do so because then it 
will be similarly obligated to answer its sibling's summoning. A summoned hecatoncheires 
cannot use its summoning power while ``summoned.'' 

\textbf{Abomination Traits: }Immune to polymorphing, petrification, and other form-altering 
attacks; not subject to energy drain, ability drain, ability damage, or death from 
massive damage; immune to mind-affecting effects; fire resistance 20; cold resistance 
20; nondetection; true seeing at will; blindsight 500 ft.; telepathy out to 1,000 
ft. 

\textbf{Regeneration (Ex):} Hecatoncheires take normal damage from good weapons 
or weapons tempered with the blood of a deity. 

\vspace{12pt}
{\LARGE{}INFERNAL }

\begin{tabular}{|>{\raggedright}p{65pt}|>{\raggedright}p{260pt}|}
\hline
  & Large Outsider (Evil) (Chaotic or Lawful) \tabularnewline
\hline
H\textbf{it Dice:} &  40d8+360 (680 hp) \tabularnewline
\hline
I\textbf{nitiative:} &  +11 (+7 Dex, +4 Improved Initiative) \tabularnewline
\hline
S\textbf{peed:} &  80 ft., fly 240 ft. (perfect) \tabularnewline
\hline
A\textbf{C:} &  50 (+7 Dex, -1 size, +34 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} &  +40/+60\tabularnewline
\hline
A\textbf{ttack:} &  +58 claw (4d6+16) melee\tabularnewline
\hline
F\textbf{ull Attack:} &  2 claws +58 (4d6+16) melee, 1 bite +52 (4d8+8 +spell suck) 
melee, 2 wings +50 (2d6+8) melee, 1 tail slam +50 (4d8+8) melee\tabularnewline
\hline
S\textbf{pace/Reach:} &  10ft./10 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} &  Improved grab, spell suck, learned spell immunity, 
spell-like abilities, summon fiend \tabularnewline
\hline
S\textbf{pecial Qualities:} &  Abomination traits, fast healing 15, regeneration 
15, SR 38, DR 35/+7 \tabularnewline
\hline
S\textbf{kills:} &  Balance +50, Concentration +52, Diplomacy +52, Escape Artist 
+50, Hide +50, Jump +36, Knowledge (arcana) +49, Knowledge (the planes) +49, Listen 
+53, Move Silently +50, Search +49, Sense Motive +51, Spellcraft +49, Spot +53, 
Tumble +50\tabularnewline
\hline
F\textbf{eats:} &  Alertness, Cleave, Great Cleave, Improved Initiative, Power 
Attack, Sunder, Weapon Focus (claw) \tabularnewline
\hline
E\textbf{pic Feats:} &  Blinding Speed (x2), Epic Toughness, Epic Weapon Focus 
(bite), Epic Weapon Focus (claw), Tenacious Magic \textit{(greater invisibilit}y), 
Tenacious Magic \textit{(blur}) \tabularnewline
\hline
S\textbf{aves:} &  Fort +31, Ref +29, Will +30 \tabularnewline
\hline
A\textbf{bilities:} &  Str 43, Dex 25, Con 28, Int 22, Wis 26, Cha 29 \tabularnewline
\hline
C\textbf{limate/Terrain:} &  Any \tabularnewline
\hline
O\textbf{rganization:} &  Solitary or hell brigade (1 infernal and 1d4 balors, 
or 1 infernal and 1d4 pit fiends) \tabularnewline
\hline
C\textbf{hallenge Rating:} &  26 \tabularnewline
\hline
T\textbf{reasure:} &  Standard \tabularnewline
\hline
A\textbf{lignment:} &  Lawful evil or chaotic evil \tabularnewline
\hline
A\textbf{dvancement:} &  41-50 HD (Large); 51-56 HD (Huge); 57-72 HD (Gargantuan) 
\tabularnewline
\hline
\end{tabular}

An infernal's natural weapons are treated as epic, evil, and chaotic-aligned or 
lawful-aligned for the purpose of overcoming damage reduction.

Combat\textbf{ }

\textbf{Improved Grab (Ex):} If an infernal hits with a claw, it deals normal damage 
and attempts to start a grapple as a free action without provoking an attack of 
opportunity. Infernals can use this ability on Large and smaller creatures. The 
infernal has the option to conduct the grapple normally, or simply use its claw 
to hold the opponent. Each successful grapple check it makes during successive 
rounds automatically deals standard claw damage, in addition to automatic spell 
suck damage. 

\textbf{Spell Suck (Su):} If an infernal bites a foe, the foe loses one of its 
highest-level prepared spells or one of the spell slots for this day that it has 
not used. The victim chooses the prepared spell to lose. If the victim has no prepared 
spells or unused spell slots (either because it has exhausted its spellcasting 
for the day or because the victim is not a spellcaster), the bite instead deals 
2 points of temporary Intelligence damage. 

\textbf{Learned Spell Immunity (Su):} If an infernal is affected by a spell cast 
by a particular spellcaster, the infernal thereafter becomes immune to that spell 
when cast by that spellcaster. 

\textbf{Spell-Like Abilities:} At will---\textit{animate dead, blasphemy, blur, 
charm person, create undead, darkness, deeper darkness, desecrate, detect chaos, 
detect good, detect law, detect magic, fear, fireball, greater dispel magic, hold 
person, greater invisibility, magic circle against good, major image, produce flame, 
polymorph, pyrotechnics, read magic, scrying, suggestion, symbol of death, telekinesis, 
greater teleport }(self plus 1,000 pounds), \textit{unholy aura, unholy blight, 
unhallow, wall of fire; }1/da\textit{y---fire storm, hellball }(epic spell), \textit{implosion, 
meteor swar}m. Caster level 26th; save DC 19 + spell level. The DCs are Charisma-based.

\textit{\textbf{Summon Fiend }}\textbf{(Sp):} An infernal can summon four balors 
or four pit fiends per day (lawful infernals summon pit fiends, and chaotic infernals 
summon balors). 

\textbf{Abomination Traits:} Immune to polymorphing, petrification, and other form-altering 
attacks; not subject to energy drain, ability drain, ability damage, or death from 
massive damage; immune to mind-affecting effects; fire resistance 20; cold resistance 
20; nondetection ; true seeing at will; blindsight 500 ft.; telepathy out to 1,000 
ft. 

\textbf{Regeneration (Ex):} Infernals take normal damage from good weapons. Lawful 
infernals also take normal damage from chaotic weapons, and vice versa. 

\vspace{12pt}
{\LARGE{}PHAETHON }

\begin{tabular}{|>{\raggedright}p{65pt}|>{\raggedright}p{260pt}|}
\hline
  & Gargantuan Outsider (Evil, Extraplanar, Fire)\tabularnewline
\hline
H\textbf{it Dice:} &  62d8+806 (1,362 hp) \tabularnewline
\hline
I\textbf{nitiative:} &  +15 (+7 Dex, +8 Superior Initiative) \tabularnewline
\hline
S\textbf{peed:} &  120 ft., burrow 120 ft. \tabularnewline
\hline
A\textbf{C:} &  47 (-4 size, +7 Dex, +34 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} &  +62/+98\tabularnewline
\hline
A\textbf{ttack:} &  pseudopod slam +85 (4d8+24 plus fire damage) melee\tabularnewline
\hline
F\textbf{ull Attack:} &  8 pseudopod slams +85 (4d8+24 plus fire damage) melee 
\tabularnewline
\hline
S\textbf{pace/Reach:} &  20 ft./10 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} &  Fiery touch, fiery overrun, spell-like abilities, 
improved grab, swallow whole, \textit{summon elder fire elemental }\tabularnewline
\hline
S\textbf{pecial Qualities:} &  Abomination traits, immunity to fire, vulnerability 
to cold, oozelike immunities, regeneration 25, fast healing 25, SR 46, DR 15/epic 
lawful and epic \tabularnewline
\hline
S\textbf{aves:} &  Fort +48, Ref +42, Will +43 \tabularnewline
\hline
A\textbf{bilities:} &  Str 58, Dex 25, Con 36, Int 8, Wis 18, Cha 39 \tabularnewline
\hline
S\textbf{kills:} &  Climb +89, Craft (metalworking, stoneworking) +29, Hide +12, 
Jump +69, Knowledge (geography) +64, Listen +71, Move Silently +72, Search +64, 
Spot +71\tabularnewline
\hline
F\textbf{eats:} &  Alertness, Cleave, Great Cleave, Great Fortitude, Improved Initiative, 
Iron Will, Lightning Reflexes, Power Attack, Sunder, Weapon Focus (slam) \tabularnewline
\hline
E\textbf{pic Feats:} &  Blinding Speed (x5), Epic Toughness (x3), Epic Weapon Focus 
(slam), Epic Will, Superior Initiative \tabularnewline
\hline
C\textbf{limate/Terrain:} &  Any \tabularnewline
\hline
O\textbf{rganization:} &  Solitary, or in company of 1-10 elder fire elementals 
\tabularnewline
\hline
C\textbf{hallenge Rating:} &  34 \tabularnewline
\hline
T\textbf{reasure:} &  Standard \tabularnewline
\hline
A\textbf{lignment:} &  Always chaotic evil \tabularnewline
\hline
A\textbf{dvancement:} &  63-80 HD (Gargantuan); 81-186 HD (Colossal) \tabularnewline
\hline
\end{tabular}

A phaethon's natural weapons are treated as epic and evil for the purpose of overcoming 
damage reduction.

Combat\textbf{ }

\textbf{Fiery Touch (Ex):} Touching or being touched by a phaethon deals 2d6 points 
of fire damage. 

\textbf{Fiery Overrun (Ex):} A foe who is successfully overrun by a phaethon is 
treated as if swallowed whole. 

\textbf{Improved Grab (Ex):} If the phaethon hits with a pseudopod, it deals normal 
damage and attempts to start a grapple as a free action without provoking an attack 
of opportunity. Phaethons can use improved grab on a Huge or smaller creature. 
The phaethon has the option to conduct the grapple normally, simply maintaining 
a pseudopod hold, or attempting to absorb the opponent by swallowing the foe whole. 
Each successful grapple check it makes during successive rounds automatically deals 
the damage listed for the pseudopod. 

\textbf{Swallow Whole (Ex):} The phaethon can absorb opponents it holds with a 
second successful grapple check after a grab. The opponent must be Huge or smaller. 
Absorbed creatures take 20d6 points of fire damage and 10d6 points of bludgeoning 
damage each round they remain inside a phaethon. Victims must make a successful 
grapple check to ``swim'' free of the living magma of the phaethon. 

\textbf{Spell-Like Abilities:} At will---\textit{fireball, fire storm, delayed 
blast fireball, flame strike, wall of fire; }1/day---\textit{scrying. }Caster level 
34th; save DC 24 + spell level. 

\textit{\textbf{Summon Elder Fire Elemental }}\textbf{(Sp):} Phaethons can summon 
up to ten elder fire elementals per day. 

\textbf{Abomination Traits:} Immune to polymorphing, petrification, and other form-altering 
attacks; not subject to energy drain, ability drain, ability damage, or death from 
massive damage; immune to mind-affecting effects; fire resistance 20; cold resistance 
20; nondetection; true seeing at will; blindsight 500 ft.; telepathy out to 1,000 
ft. 

\textbf{Oozelike Immunities (Ex):} Immune to poison, \textit{sleep, }paralysis, 
stunning, polymorphing, and mind-influencing effects; not subject to critical hits 
or flanking; blind. 

\textbf{Regeneration (Ex):} Phaethons take normal damage from lawful weapons and 
cold-forged weapons, and double damage from cold or ice weapons. 

\vspace{12pt}
{\LARGE{}PHANE }

\begin{tabular}{|>{\raggedright}p{66pt}|>{\raggedright}p{259pt}|}
\hline
  & Large Outsider (Chaotic, Evil, Extraplanar, Incorporeal) \tabularnewline
\hline
H\textbf{it Dice:} &  36d8+324 (612 hp) \tabularnewline
\hline
I\textbf{nitiative:} &  +11 (+7 Dex, +4 Improved Initiative) \tabularnewline
\hline
S\textbf{peed:} &  80 ft., fly 120 ft. (perfect) \tabularnewline
\hline
A\textbf{C:} &  50 (-1 size, +7 Dex, +11 deflection, +23 insight) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} &  +36/---\tabularnewline
\hline
A\textbf{ttack:} &  Incorporeal touch +43 (1d6 plus stasis touch) melee\tabularnewline
\hline
F\textbf{ull Attack:} &  Incorporeal touch +43 (1d6 plus stasis touch) melee\tabularnewline
\hline
S\textbf{pace/Reach:} &  10 ft./10 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} &  Spell-like abilities, stasis touch, chronal blast, 
time leach, \textit{summon past time duplicate }\tabularnewline
\hline
S\textbf{pecial Qualities:} &  Abomination traits, null time field, time regression, 
fast healing 15, regeneration 15, SR 37, DR 15/epic, sonic immunity \tabularnewline
\hline
S\textbf{aves:} &  Fort +31, Ref +29, Will +29 \tabularnewline
\hline
A\textbf{bilities:} &  Str ---, Dex 25, Con 28, Int 24, Wis 16, Cha 33 \tabularnewline
\hline
S\textbf{kills:} & Balance +46, Concentration +48, Diplomacy +50, Hide +42, Jump 
+20, Knowledge (arcana, religion, the planes) +46, Listen +44, Move Silently +46, 
Search +46, Sense Motive +42, Spellcraft +46, Spot +44, Swim +39, Tumble +46\tabularnewline
\hline
F\textbf{eats:} &  Alertness, Combat Reflexes, Dodge, Great Fortitude, Improved 
Initiative, Iron Will, Lightning Reflexes, Weapon Focus (incorporeal touch) \tabularnewline
\hline
E\textbf{pic Feats:} &  Blinding Speed, Epic Toughness (x2), Epic Will, Spell Stowaway 
\textit{(time sto}p)\tabularnewline
\hline
C\textbf{limate/Terrain:} &  Any land and underground \tabularnewline
\hline
O\textbf{rganization:} &  Solitary or pair \tabularnewline
\hline
C\textbf{hallenge Rating:} &  25 \tabularnewline
\hline
T\textbf{reasure:} &  None \tabularnewline
\hline
A\textbf{lignment:} &  Always chaotic evil \tabularnewline
\hline
A\textbf{dvancement:} &  37-50 HD (Large); 51-65 HD (Huge) \tabularnewline
\hline
\end{tabular}

A phanes' natural weapons are treated as epic, evil, and chaotic-aligned for the 
purpose of overcoming damage reduction.

\section*{Combat}

\textbf{Spell-Like Abilities:} At will---\textit{detect good, detect magic, improved 
invisibility }(self only); 5/da\textit{y---haste, slow, greater teleport, tongues, 
trap the soul, true strike, unholy aura; 2/day---safe time, time duplicate }(epic 
spell); \textit{1/day---time stop. }Caster level 21st; save DC 21 + spell level. 
The DC is Charisma-based.

\textbf{Stasis Touch (Su): }As \textit{temporal stasis, }except as an at-will supernatural 
ability. Caster level 25th; save DC 21 + spell level. The DC is Charisma-based.

\textbf{Chronal Blast (Su): }As a standard action, the phane can make a ranged 
touch attack against any creature within 100 feet. If it succeeds, the subject 
is targeted by a spasm of space-time flux, dealing 15d6 points of damage. 

\textbf{Time Leach (Su):} For every round of apparent time experienced by the phane, 
it automatically absorbs the ``future'' from any creature it has successfully encapsulated 
in static time via its stasis touch (not its null time field), no matter the distance 
separating victim and phane, and no matter the number of victims. Of course, to 
the victim no time passes at all, but each apparent round experienced by the phane 
ages the victim 1d4 years, at the same time healing the phane of 20 hit points 
of damage. A victim who is not somehow released from static time by a friend who 
can cast \textit{dispel magic, greater dispel magic, }or some other likely spell, 
eventually ages to death. Victims killed in this manner automatically fall out 
of static time as desiccated husks that disintegrate to a fine dust with even the 
lightest touch. Victims who are released prior to death immediately apply the physical 
effects of aging, but not the mental effects. 

\textit{\textbf{Summon Past Time Duplicate }}\textbf{(Sp):} Once per day, a phane 
can summon a duplicate of one its foes stolen from a parallel alternate past. The 
stolen \textit{time duplicate }has the same stats and possessions as the original, 
but is treated as if having two negative levels (which simulates a less experienced 
version of the original). The phane can never summon a past \textit{time duplicate 
}of a creature with more than 25 HD (add sufficient negative levels to compensate 
for high foe HD, if necessary). The past \textit{time duplicat}e, despite having 
most of the knowledge of the original, serves the phane loyally like any summoned 
creature. If the past \textit{time duplicate }is slain, the original is not harmed 
because the duplicate was pulled from a parallel past. However, the original does 
not necessarily realize this, and must make a Will save (DC 30) or be shaken for 
1d4 rounds after witnessing the death of a duplicate for the first time. 

\textbf{Abomination Traits:} Immune to polymorphing, petrification, and other form-altering 
attacks; not subject to energy drain, ability drain, ability damage, or death from 
massive damage; immune to mind-affecting effects; fire resistance 20; cold resistance 
20; nondetection; true seeing at will; blindsight 500 ft.; telepathy out to 1,000 
ft. 

\textbf{Null Time Field (Ex):} Phanes continually generate a 30-foot-radius spread 
null time field. All creatures and objects in the field, except the phane, must 
make a Will saving throw (DC 30) each round to take any actions. On a failed save, 
subjects are stuck in a static time stream until their next round of actions, at 
which time they must make another saving throw. While a subject is stuck in a static 
time stream induced by a null time field, the phane can use its static touch on 
the subject, though in all other ways, the subject is invulnerable to attacks and 
damage as if in temporal stasis. 

\textbf{Time Regression (Su):} If the phane spends an action per round for four 
rounds, at the end of the 4th round the phane regresses back in time 4 rounds, 
to the very 1st round it originally began concentrating on time regression. On 
its second pass through the time stream, it can take completely different actions, 
based on its knowledge of the future (though if it takes different actions from 
its first pass through the time stream, the events of the original time stream 
are also changed). 

\textbf{Regeneration (Ex):} Phanes take normal damage from holy and good weapons, 
and from weapons forged in the future or an alternate reality's time stream (if 
any). 

\vspace{12pt}
{\LARGE{}XIXECAL }

\begin{tabular}{|>{\raggedright}p{68pt}|>{\raggedright}p{258pt}|}
\hline
  & Colossal Outsider (Chaotic, Cold, Evil, Extraplanar) \tabularnewline
\hline
H\textbf{it Dice:} &  72d8+1,080 (1,676 hp) \tabularnewline
\hline
I\textbf{nitiative:} &  +11 (+7 Dex. +4 Improved Initiative) \tabularnewline
\hline
S\textbf{peed:} &  140 ft. \tabularnewline
\hline
A\textbf{C:} &  58 (-8 size, +1 Dex, +55 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} &  +72/+116\tabularnewline
\hline
A\textbf{ttack:} &  Claw +95 (2d8+28/19-20) melee\tabularnewline
\hline
F\textbf{ull Attack:} &  2 claws +95 (2d8+28/19-20) melee, bite +89 (4d6+14) melee, 
stamp +89 (4d6+14) melee \tabularnewline
\hline
S\textbf{pace/Reach:} &  50 ft./50 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} &  Rend 4d8+42, cold, spell-like abilities, breath weapon, 
\textit{summon white dragon, }Constitution drain \tabularnewline
\hline
S\textbf{pecial Qualities:} &  Abomination traits, cold immunity, fire vulnerability, 
dire winter, fast healing 30, regeneration 30, SR 48, DR 20/good and epic and adamantine 
\tabularnewline
\hline
S\textbf{aves:} &  Fort +55, Ref +47, Will +39 \tabularnewline
\hline
A\textbf{bilities:} &  Str 66, Dex 13, Con 40, Int 12, Wis 8, Cha 34 \tabularnewline
\hline
S\textbf{kills:} & Climb +103, Concentration +90, Hide -9, Jump +28, Knowledge 
(arcana, religion, the planes) +76, Listen +70, Search +76, Spellcraft +76, Spot 
+76\tabularnewline
\hline
F\textbf{eats:} & Alertness, Cleave, Dodge, Great Cleave, Great Fortitude, Improved 
Critical (claw), Improved Initiative, Improved Sunder, Iron Will, Lightning Reflexes, 
Power Attack, Weapon Focus (bite), Weapon Focus (claw), Weapon Focus (stomp)\tabularnewline
\hline
E\textbf{pic Feats:} & Blinding Speed (x4), Devastating Critical (claw), Epic Toughness, 
Epic Weapon Focus (bite), Epic Weapon Focus (claw), Epic Weapon Focus (stomp), 
Overwhelming Critical (claw), Spell Stowaway \textit{(greater invisibilit}y) \tabularnewline
\hline
C\textbf{limate/Terrain:} &  Any \tabularnewline
\hline
O\textbf{rganization:} &  Solitary or in the company of 1d4+1 old white dragons 
\tabularnewline
\hline
C\textbf{hallenge Rating:} &  36 \tabularnewline
\hline
T\textbf{reasure:} &  Standard \tabularnewline
\hline
A\textbf{lignment:} &  Always chaotic evil \tabularnewline
\hline
A\textbf{dvancement:} &  73-150 HD (Colossal) \tabularnewline
\hline
\end{tabular}

A xixecal's natural weapons are treated as epic, evil, and chaotic-aligned for 
the purpose of overcoming damage reduction.

Combat\textbf{ }

\textbf{Rend (Ex):} If the xixecal hits with both claw attacks, it latches onto 
the opponent's body and tears the flesh. This attack automatically deals an additional 
4d8+42 points of damage. 

\textbf{Cold (Su):} A hit from a xixecal's melee attack induces numbing cold. The 
opponent must succeed at a Fortitude save (DC 58) or be affected as though by a 
\textit{slow }spell for 10 rounds. The DC is Charisma-based.

\textbf{Spell-Like Abilities:} At will---\textit{blasphemy, cone of cold, desecrate, 
greater dispel magic, dominate monster, hold person, greater invisibility, unholy 
aura, wall of ic}e; 5/day---\textit{haste, meteor swarm. }Caster level 36th; save 
DC 22 + spell level. The DC is Charisma-based.

\textbf{Breath Weapon (Su):} Cone of cold, 300 feet, 3/day, damage 12d6, Fort half 
DC 61. The DC is Constitution-based.

\textit{\textbf{Summon White Dragon }}\textbf{(Sp):} Five times per day a xixecal 
can summon an old white dragon. 

\textbf{Constitution Drain (Su):} This effect permanently reduces a living opponent's 
Constitution score by 4 when the creature hits with a slam, bite, or rend, or twice 
that amount on a rend or critical hit. The creature heals 20 points of damage (or 
40 on a rend or critical hit) whenever it drains Constitution, gaining any excess 
as temporary hit points. These temporary hit points last a maximum of 1 hour. The 
attack allows a Fortitude save (DC 58). The DC is Charisma-based.

\textbf{Abomination Traits:} Immune to polymorphing, petrification, and other form-altering 
attacks; not subject to energy drain, ability drain, ability damage, or death from 
massive damage; immune to mind-affecting effects; nondetection; true seeing at 
will; blindsight 500 ft.; telepathy out to 1,000 ft.

The standard fire resistance 20 most abominations have is not shared by xixecales

\textbf{Dire Winter (Su):} A xixecal always stands at the center of a permanent 
\textit{dire winter} (epic spell)\textit{ }spell effect. If dispelled, the effect 
returns 1 minute later--- the only way to permanently remove the effect is to slay 
the xixecal that generates it. 

\textbf{Regeneration(Ex):} Xixecals take normal damage from Lawful weapons, and 
double damage from burning or fiery weapons. 

\vspace{24pt}
\section*{{\LARGE{}BEHEMOTH }}

\section*{\begin{tabular}{|>{\raggedright}p{66pt}|>{\raggedright}p{123pt}|>{\raggedright}p{123pt}|}
\hline
} & B\textbf{ehemoth Eagle } & B\textbf{ehemoth Gorilla }\tabularnewline
\hline
 & Colossal Outsider (Extraplanar)  & Huge Outsider (Extraplanar)\tabularnewline
\hline
H\textbf{it Dice:}  & 21d8+126 (220 hp)  & 21d8+252 (366 hp) \tabularnewline
\hline
I\textbf{nitiative:}  & +20 (+12 Dex, +8 Superior Initiative)  & +12 (+8 Dex, +4 
Improved Initiative) \tabularnewline
\hline
S\textbf{peed:}  & 20 ft., fly 160 ft. (poor)  & 60 ft., climb 30 ft. \tabularnewline
\hline
A\textbf{C:}  & 24 (+12 Dex, -8 size, +10 natural)  & 38 (+8 Dex, -2 size, +22 
natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:}  & +21/+44 & +21/+39\tabularnewline
\hline
A\textbf{ttack:} & Claw +25 (5d10+7) melee & Claw +29 (4d8+10) melee\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +25 (5d10+7) melee, bite +20 (4d12+3) melee & 2 
claws +29 (4d8+10) melee, bite +24 (3d10+5) melee\tabularnewline
\hline
S\textbf{pace/Reach:}  & 30ft./20 ft.  & 20 ft./20 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} --- &   & Rend 8d8+20 \tabularnewline
\hline
S\textbf{pecial Qualities:}  & Evasion, SR 30  & Scent, SR 30 \tabularnewline
\hline
S\textbf{aves:}  & Fort +18, Ref +24, Will +16  & Fort +24, Ref +20, Will +16 \tabularnewline
\hline
A\textbf{bilities:}  & Str 25, Dex 34, Con 23 Int 17, Wis 19, Cha 16  & Str 31, 
Dex 26, Con 35 Int 6, Wis 15, Cha 12 \tabularnewline
\hline
S\textbf{kills:} &  Diplomacy +27, Jump +3, Knowledge (geography, history, nature, 
religion, the planes) +27, Listen +30, Search +27, Sense Motive +28, Spot +41, 
Survival +28 & Balance +32, Climb +42, Intimidate +25, Jump +34, Move Silently 
+32, Spot +26\tabularnewline
\hline
F\textbf{eats:}  & Alertness, Endurance, Flyby Attack, Improved Initiative, Iron 
Will, Skill Focus (spot), Weapon Finesse & Dodge, Endurance, Improved Initiative, 
Iron Will, Mobility, Power Attack, Spring Attack\tabularnewline
\hline
E\textbf{pic Feats:}  & Superior Initiative  & Epic Toughness\tabularnewline
\hline
C\textbf{limate/Terrain:}  & Any forest, hill, mountains, and plains  & Warm forest 
and warm mountains\tabularnewline
\hline
O\textbf{rganization:}  & Solitary or pair  & Solitary or company (5-8) \tabularnewline
\hline
C\textbf{hallenge Rating:}  & 18  & 19 \tabularnewline
\hline
T\textbf{reasure:}  & None  & None \tabularnewline
\hline
A\textbf{lignment:}  & Always neutral  & Always neutral \tabularnewline
\hline
A\textbf{dvancement:}  & 21-42 HD (Colossal)  & 22-42 HD (Huge) \tabularnewline
\hline
\end{tabular}

\section*{BEHEMOTH EAGLE }

\textbf{Skills:} *A behemoth eagle receives a +8 racial bonus on Spot checks. 

\textbf{Carrying Capacity:} A light load for a behemoth eagle is up to 2,400 pounds; 
a medium load, 2,401-4,800 pounds; and a heavy load, 4,801-7,200 pounds. 

\section*{BEHEMOTH GORILLA }

\section*{Combat\textbf{ }}

\textbf{Rend (Ex):} A behemoth gorilla that hits with both claw attacks latches 
onto the opponent's body and tears the flesh. This automatically deals an additional 
8d8+20 points of damage. 

\vspace{12pt}
{\LARGE{}BRACHYURUS }

\begin{tabular}{|>{\raggedright}p{63pt}|>{\raggedright}p{263pt}|}
\hline
  & Large Magical Beast \tabularnewline
\hline
H\textbf{it Dice:} &  38d10+684 (893 hp) \tabularnewline
\hline
I\textbf{nitiative:} &  +18 (+14 Dex, +4 Improved Initiative) \tabularnewline
\hline
S\textbf{peed:} &  80 ft. \tabularnewline
\hline
A\textbf{C:} &  40 (+14 Dex, -1 size, +17 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} &  +38/+52\textbf{ }\tabularnewline
\hline
A\textbf{ttack:} &  Claw +48 (2d6+10/19-20) melee\tabularnewline
\hline
F\textbf{ull Attack:} &  2 claws +48 (2d6+10/19-20) melee, 1 bite +46 (3d6+5/19-20) 
melee\tabularnewline
\hline
S\textbf{pace/Reach:} &  10 ft./5 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} &  Frightful howl, savage 15d6+25 \tabularnewline
\hline
S\textbf{pecial Qualities:} &  Blindsight 240 ft., DR 10/epic, dark-vision 60 ft., 
fast healing 10, lowlight vision, SR 32, scent, acid, cold, electricity, fire, 
and sonic resistance 20 \tabularnewline
\hline
S\textbf{aves:} &  Fort +41, Ref +35, Will +25 \tabularnewline
\hline
A\textbf{bilities:} &  Str 30, Dex 38, Con 47, Int 18, Wis 32, Cha 19 \tabularnewline
\hline
S\textbf{kills:} & Hide +55, Jump +51, Listen +54, Move Silently +55, Spot +54, 
Survival +52\tabularnewline
\hline
F\textbf{eats:} &  Alertness, Great Fortitude, Improved Bull Rush, Improved Critical 
(claw), Improved Critical (bite), Improved Initiative, Improved Trip, Iron Will, 
Multiattack, Weapon Focus (claw), Weapon Focus (bite) \tabularnewline
\hline
E\textbf{pic Feats:} &  Blinding Speed (x2) \tabularnewline
\hline
C\textbf{limate/Terrain:} &  Any \tabularnewline
\hline
O\textbf{rganization:} &  Solitary, pair, or pack (5-10). \tabularnewline
\hline
C\textbf{hallenge Rating:} &  23 \tabularnewline
\hline
T\textbf{reasure:} &  Standard \tabularnewline
\hline
A\textbf{lignment:} &  Usually lawful neutral \tabularnewline
\hline
A\textbf{dvancement:} &  39-48 HD (Large); 49-58 HD (Huge); 59-66 HD (Gargantuan); 
67-135 HD (Colossal) \tabularnewline
\hline
\end{tabular}

COMBAT\textbf{ }

\textbf{Frightful Howl (Su):} As a standard action, a brachyurus can emit a howl 
in a 60-foot-radius spread that affects opponents with fewer Hit Dice or levels 
than the brachyurus has. The affected creature must make a successful Will save 
(DC 33) or become shaken. Success indicates that the target is immune to that brachyurus's 
frightful howl for one day. The DC is Charisma-based.

\textbf{Savage (Ex):} If a brachyurus successfully trips an opponent, it latches 
onto the opponent's body and tears the flesh. This attack automatically deals an 
additional 15d6+25 points of damage. If an opponent goes prone for any reason in 
an area a brachyurus threatens, the brachyurus can also savage the victim as a 
free action (treat as the brachyurus's attack of opportunity for the round), even 
though it had nothing to do with tripping the foe. 

\vspace{12pt}
\section*{{\LARGE{}COLOSSUS }}

\textbf{Construct Traits:} Immune to mind-affecting effects (charms, compulsions, 
phantasms, patterns, and morale effects), and to poison, \textit{sleep, }paralysis, 
stunning, disease, death effects, necromantic effects, and any effect that requires 
a Fortitude save unless it also works on objects. Cannot heal damage (though regeneration 
and fast healing still apply, if present). Not subject to critical hits, subdual 
damage, ability damage, ability drain, or energy drain. Not at risk of death from 
massive damage, but destroyed when reduced to 0 hit points or less; cannot be raised 
or resurrected. Darkvision 60 ft. 

\textbf{Magic Immunity (Ex):} A colossus completely resists most magical and supernatural 
effects, except where otherwise noted below. 

\textbf{Antimagic Field (Ex):} A colossus constantly generates an antimagic field 
in a 100-foot-radius. The field is an invisible barrier that is impervious to most 
magical effects, including spells, spell-like abilities, and supernatural abilities. 
Likewise, it prevents the functioning of any magic items or spells within its confines, 
except for the colossus's own supernatural abilities. This effect is otherwise 
as an \textit{antimagic }field cast by a 25th-level caster. 

A colossus's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

\section*{Construction }

The construction cost given for each colossus takes into account the physical body 
and all the materials and spell components that are consumed or become a permanent 
part of it. Except for the much tougher requirements, constructing a colossus is 
not unlike constructing a golem. 

\begin{tabular}{|>{\raggedright}p{70pt}|>{\raggedright}p{77pt}|>{\raggedright}p{77pt}|>{\raggedright}p{77pt}|}
\hline
   & S\textbf{tone Colossus } & F\textbf{lesh Colossus } & I\textbf{ron Colossus 
}\tabularnewline
\hline
 & Colossal Construct  & Colossal Construct  & Colossal Construct \tabularnewline
\hline
H\textbf{it Dice:}  & 64d10 (432 hp)  & 100d10 (630 hp) or as controlling spirit & 96d10 
(603 hp) \tabularnewline
\hline
I\textbf{nitiative:} - & 3 (Dex)  & +2 (-2 Dex, +4 Improved - & 2 (Dex) Initiative) 
\tabularnewline
\hline
S\textbf{peed:}  & 40 ft. (can't run)  & 40 ft. (can't run)  & 50 ft. (can't run) 
\tabularnewline
\hline
A\textbf{C:}  & 44 (-8 size, -3 Dex, +45 natural) & 45 (-8 size, -2 Dex, +25 natural, 
+20 profane)  & 60 (-8 size, -2 Dex, +60 natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +48/+94 & +75/+103 & +72/+123\tabularnewline
\hline
A\textbf{ttack:} & Slam +70 (8d6+30) melee & Colossal club +79 (6d6+18) melee or 
slam +79 (4d6+12) melee or thrown object (weighing 10 tons or less) +65 (4d6+12) 
ranged & Slam +99 (12d6+35) melee\tabularnewline
\hline
F\textbf{ull Attack:} & 2 slams +70 (8d6+30) melee & Colossal club +79 (6d6+18) 
melee or 2 slams +79 (4d6+12) melee or thrown object (weighing 10 tons or less) 
+65 (4d6+12) ranged & 2 slams +99 (12d6+35) melee\tabularnewline
\hline
S\textbf{pace/Reach:}  & 30 ft./30 ft.  & 30 ft./30 ft.  & 30 ft./30 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:}  & Shatter  & Frightful presence, horrific appearance, 
stomp, stench & Breath weapon \tabularnewline
\hline
S\textbf{pecial Qualities:}  & Construct traits, magic immunity, antimagic field, 
DR 10/Epic and adamantine & Construct traits, magic immunity, DR 15/Epic and adamantine, 
negative energy affinity & Construct traits, magic immunity, antimagic field, rustproof, 
DR 20/Epic and adamantine \tabularnewline
\hline
S\textbf{aves:}  & Fort +21, Ref +18, Will +21  & Fort +33, Ref +31, Will +33  & Fort 
+32, Ref +30, Will +33 \tabularnewline
\hline
A\textbf{bilities:}  & Str 70, Dex 5, Con -, Int 7, Wis 10, Cha 1 & Str 35, Dex 
6, Con -, Int 1 or as controlling spirit, Wis 11 or as controlling spirit, Cha 
1 or as controlling spirit & Str 80, Dex 7, Con ---,Int 9, Wis 12, Cha 1 \tabularnewline
\hline
S\textbf{kills:}  & Jump +34 & Jump +16 or as controlling spirit  & Jump +43\tabularnewline
\hline
C\textbf{limate/Terrain:}  & Any  & Any  & Any \tabularnewline
\hline
O\textbf{rganization:}  & Solitary  & Solitary  & Solitary \tabularnewline
\hline
C\textbf{hallenge Rating:}  & 24  & 27  & 33 \tabularnewline
\hline
T\textbf{reasure:}  & None  & None  & None\tabularnewline
\hline
A\textbf{lignment:}  & Usually neutral  & Neutral evil or neutral  & Usually neutral 
\tabularnewline
\hline
A\textbf{dvancement:}  & 65-192 HD (Colossal)  & None (abilities may vary by level 
of possessing spirit)  & 97-288 HD (Colossal)\tabularnewline
\hline
\end{tabular}

\section*{STONE COLOSSUS }

A stone colossus cannot speak or make any vocal noise. 

\section*{Combat\textbf{ }}

\textbf{Shatter (Su):} A stone colossus can use \textit{shatter}(DC 7)\textit{ 
}as a free action once every 2 rounds against an opponent's weapon or other item 
as a 24th-level caster. The DC is Charisma-based.

\textbf{Magic Immunity (Ex):} A stone colossus is immune to all spells, spell-like 
abilities, and supernatural effects, except as follows. A \textit{transmute rock 
to mud }spell slows it (as the \textit{slow }spell) for 1 round, with no saving 
throw, while \textit{transmute mud to rock }heals all its lost hit points. A \textit{stone 
to flesh }spell has no effect on a stone colossus. 

\section*{\textbf{Construction }}

A stone colossus's body is chiseled from a single block of hard stone, such as 
granite, weighing at least 375,000 pounds. The colossus costs 760,000 gp to create, 
which includes 100,000 gp for the body. Assembling the body requires a successful 
Craft (sculpting) or Craft (masonry) check (DC 37). The creator must be 25th level 
and able to cast epic spells. Completing the ritual drains 30,600 XP from the creator 
and requires \textit{geas/quest, limited wish, polymorph any object, }and \textit{shatte}r. 

\section*{FLESH COLOSSUS }

A flesh colossus knows all languages spoken in life by the ghost who acts as its 
control-ling spirit. When the controlling spirit is suppressed, it mutters in all 
the tongues spoken by its component corpses. 

\section*{Combat\textbf{ }}

\textbf{Frightful Presence (Ex):} All who come within a mile of the flesh colossus 
must make Will saves (DC 55). Those who make their saves are shaken. Those who 
fail by 20 points or less are frightened. Those who fail by 21 to 40 points are 
panicked. Those who fail by 41 points or more are cowering. The DC is Charisma-based.

\textbf{Horrific Appearance (Su):} Once per day, the controlling spirit within 
a flesh colossus may draw upon the negative energy latent in the construct to transform 
its countenance into a horrible leering visage: All within 300 feet who see this 
must make a Fortitude save (DC 55) or take 2d4 points of temporary Strength, Dexterity, 
and Constitution damage. The DC is Charisma-based.

\textbf{Stomp (Ex): }As a full-round action, a flesh colossus may stomp upon any 
adjacent foe (that is, any within the area it threatens). Its attack roll modifier 
for this special attack is +65 (+75 base, -2 Dex, -8 size). If successful, the 
stomp deals double damage (8d6 +24) and pins the victim until the colossus chooses 
to release him or her. 

\textbf{Stench (Ex):} All creatures with a sense of smell who come within 300 feet 
of the construct must make a Fortitude save (DC 60) or be nauseated (unable to 
attack, cast spells, concentrate on spells, or do anything else requiring attention, 
except for a single move or move-equivalent action per round). Those who succeed 
on their saves take a -4 circumstance penalty on all attack rolls, saving throws, 
and skill checks for the next 2d6+4 minutes. The DC is Constitution-based.

\textbf{Magic Immunity (Ex):} A flesh colossus is immune to all magical and supernatural 
effects except those that use negative energy (see below). 

\textbf{Negative Energy Affinity (Su): }Any negative energy spell targeted at the 
colossus repairs damage taken by the colossus at the rate of 1d10 hit points per 
level of the spell. 

\section*{\textbf{Construction }}

Only a master necromancer of surpassing power can create a flesh colossus. The 
bodies of not less than three hundred previously animated Medium-size humanoid 
creatures must be stripped of their flesh and the bones, which are then rendered 
in two great vats. The resulting semiliquid bone-paste is then cast into a new 
articulated skeleton of great size, to which the rendered flesh is applied like 
wet clay to shape the figure. Finally, the necromancer must bind a ghost of at 
least 10 HD (the more powerful the better) into the colossus to act as its controlling 
spirit. Without the controlling spirit, the colossus is just an oversized zombie; 
with it, it functions as a golem of great size and repugnant appearance. If no 
ghost is available, the necromancer may occupy the colossus himself by casting 
\textit{magic jar}. The colossus costs 760,000 gp to create, which includes 100,000 
gp for the body. Assembling the body takes no great skill, and many flesh colossi 
reveal that fact in their uncouth forms. The creator must be 29th level and able 
to cast epic spells. Completing the ritual drains 30,600 XP from the creator and 
requires \textit{magic jar, limited wish, polymorph any object, }and \textit{finger 
of death}. 

IRON COLOSSUS 

\textbf{Breath Weapon (Su):} Cone of poisonous gas, 80 feet, as a free action every 
1d4 rounds, initial damage 3d6 temporary Con, secondary damage 2d6 temporary Con, 
Fort half (DC 58). The DC is Constitution-based.

\textbf{Magic Immunity (Ex):} An iron colossus is immune to all magical and supernatural 
effects, except as follows. An electricity effect restores 1 hit point to the iron 
colossus for each 3 points of damage it would otherwise deal. 

\textbf{Rustproof (Ex):} An iron colossus is not affected by rust attacks, unlike 
smaller iron golems. 

\section*{\textbf{Construction }}

An iron colossus's body is sculpted from 150,000 pounds of pure iron. The colossus 
costs 1,000,000 gp to create, which includes 100,500 gp for the body. Assembling 
the body requires a successful Craft (armorsmithing) or Craft (weaponsmithing) 
check (DC 42). The creator must be 35th level and able to cast epic spells. Completing 
the ritual drains 50,000 XP from the creator and requires \textit{cloudkill, geas/quest, 
limited wish, }and \textit{polymorph any object. }

\vspace{12pt}
{\LARGE{}DEMILICH }

\begin{tabular}{|>{\raggedright}p{64pt}|>{\raggedright}p{262pt}|}
\hline
  & Diminutive Undead \tabularnewline
\hline
H\textbf{it Dice:} & 21d12 (130 hp) \tabularnewline
\hline
I\textbf{nitiative:} &  +7 (+3 Dex, +4 Improved Initiative) \tabularnewline
\hline
S\textbf{peed:} &  Fly 180 ft. (perfect) \tabularnewline
\hline
A\textbf{C:} &  51 (+4 size, +3 Dex, +5 natural armor, +8 \textit{bracers of armor, 
}+2 \textit{ring of protection, }+21 insight) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} &  +10/-2\tabularnewline
\hline
A\textbf{ttack:} &  Touch +35 (10d6+20 plus paralyzing touch) melee\tabularnewline
\hline
F\textbf{ull Attack:} &  Touch +35 (10d6+20 plus paralyzing touch) melee\tabularnewline
\hline
S\textbf{pace/Reach:} &  1 ft./0 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:} &  Trap the soul, fear aura, paralyzing touch, 21st-level 
wizard spellcaster, Perfect Automatic Still Spell, spell-like abilities \tabularnewline
\hline
S\textbf{pecial Qualities:} &  Magic immunity, phylactery transference, turn resistance 
+20, DR 15/Epic and bludgeoning, undead traits, acid resistance 20, fire resistance 
20, sonic resistance 20, immune to cold, electricity, polymorph, and mind-affecting 
attacks. \tabularnewline
\hline
S\textbf{aves:} &  Fort +10, Ref +12, Will +17 \tabularnewline
\hline
A\textbf{bilities:} &  Str 10, Dex 16 (with \textit{glove}s), Con -, Int 39 (with 
\textit{headband}), Wis 24, Cha 20 \tabularnewline
\hline
S\textbf{kills:} & Concentration +25, Craft (alchemy) +34, Hide +35, Knowledge 
(arcana, religion, the planes) +34, Knowledge (history) +26, Listen +27, Move Silently 
+23, Search +34, Sense Motive +27, Spellcraft +36, Spot +27\tabularnewline
\hline
F\textbf{eats:} &  Brew Potion, Combat Casting, Craft Wondrous Item, Enlarge Spell, 
Heighten Spell, Improved Initiative, Lightning Reflexes, Maximize Spell, Quicken 
Spell, Scribe Scroll, Spell Penetration, Toughness \tabularnewline
\hline
E\textbf{pic Feats:} &  Automatic Quicken Spell, Blinding Speed, Improved Spell 
Capacity (10th), Tenacious Magic (any) \tabularnewline
\hline
C\textbf{limate/Terrain:} &  Any \tabularnewline
\hline
O\textbf{rganization:} &  Solitary \tabularnewline
\hline
C\textbf{hallenge Rating:} &  29 \tabularnewline
\hline
T\textbf{reasure:} &  Standard \tabularnewline
\hline
A\textbf{lignment:} &  Neutral evil \tabularnewline
\hline
A\textbf{dvancement:} &  By character class \tabularnewline
\hline
\end{tabular}

Demiliches speak the languages they knew as liches and as living creatures. 

A demilich's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

\section*{Combat\textbf{ }}

\textbf{Trap the Soul (Su):} A demilich can trap the souls of up to eight living 
creatures per day. To use this power, it selects any target it can see within 300 
feet. The target is allowed a Fortitude saving throw (DC 36). If the target makes 
its saving throw, it gains four negative levels (this does not count as a use of 
trap the soul). If the target fails its save, the soul of the target is instantly 
drawn from its body and trapped within one of the gems incorporated into the demilich's 
form. The gem gleams for 24 hours, indicating the captive soul within. The soulless 
body collapses in a mass of corruption and molders in a single round, reduced to 
dust. If left to its own devices, the demilich slowly devours the soul over 24 
hours---at the end of that time the soul is completely absorbed, and the victim 
is forever gone. If the demilich is overcome before the soul is eaten, crushing 
the gem releases the soul, after which time it is free to seek the afterlife or 
be returned to its body by the use of either \textit{resurrectio}n, \textit{true 
resurrection, clone, }or \textit{miracle. }A potential victim protected by a \textit{death 
ward }spell is not immune to trap the soul, but receives a +5 bonus on its Fortitude 
saving throw and\textit{ }is effective against the level loss on a successful save. 
The DC is Charisma-based

\textbf{Fear Aura (Su):} Demiliches are shrouded in a dreadful aura of death and 
evil. Creatures of less than 5 HD in a 60-foot radius that look at the demilich 
must succeed at a Will save (DC 19) or be affected as though by \textit{fear }as 
cast by a 21st-level caster. The DC is Charisma-based.

\textbf{Paralyzing Touch (Su):} Any living creature a demilich touches must succeed 
at a Fortitude save (DC 36) or be permanently paralyzed. \textit{Remove paralysis 
}or any spell that can remove a curse can free the victim. The effect cannot be 
dispelled. Anyone paralyzed by a demilich seems dead, though a successful Spot 
check (DC 20) or Heal check (DC 15) reveals that the victim is still alive. 

\textbf{Spells: }The demilich can cast any spells it could cast as a lich. This 
sample demilich was a 21st-level wizard (spells per day: 4/12/8/7/7/7/7/6/6/6/3). 
The last three spell slots are 10th-level slots and available for 0-9th level metamagic 
spells. 

\textbf{Perfect Automatic Still Spell (Ex): }A demilich can cast all the spells 
it knows without gestures. 

\textbf{Spell-Like Abilities:} At will---\textit{alter self, astral projection, 
create greater undead, create undead, death knell, enervation, greater dispel magic, 
harm }(usually used to heal itself ), \textit{summon monster I-IX, telekinesis, 
weird; }2/day---\textit{greater planar ally. }Demiliches use these abilities as 
casters of a level equal to their spellcaster level, but the save DCs are all 36. 
The DC is Charisma-based.

\textbf{Magic Immunity (Ex):} Demiliches are immune to all magical and supernatural 
effects, except as follows. A \textit{shatter }spell affects a demilich as if it 
were a crystalline creature, but deals half the damage normally indicated. A \textit{dispel 
evil }spell deals 3d6 points of damage (Fort save for half damage). \textit{Holy 
smite }spells affect demiliches normally. 

\textbf{Phylactery Transference (Su): }Headbands, belts, rings, cloaks, and other 
wearable items kept in close association with the demilich's phylactery transfer 
all their benefits to the demilich no matter how far apart the demilich and the 
phylactery are located. The standard limits on types of items utilized simultaneously 
still apply. The sample demilich ``wears'' \textit{bracers of armor +8, headband 
of intellect +6, pink Ioun stone, ring of wizardry }I, \textit{ring of protection 
+2, }and \textit{gloves of Dexterity +2. }

\textbf{Undead Traits:} Immune to poison, \textit{sleep, }paralysis, stunning, 
disease, death, effects, necromantic effects, mind-affecting effects, and any effect 
requiring a Fortitude save unless it also works on objects. Not subject to critical 
hits, subdual damage, ability damage, ability drain, or energy drain. Negative 
energy heals. Not at risk of death from massive dam-age, but destroyed at 0 hit 
points or less. Darkvision 60 ft. Cannot be raised; resurrection works only if 
creature is willing. 

\textbf{Immunities (Ex):} Demiliches are immune to cold, electricity, polymorph, 
and mind-affecting attacks. 

\vspace{12pt}
\section*{CREATING A DEMILICH }

``Demilich'' is a tem-plate that can be added to any lich. It uses all the lich's 
statistics and special abilities except as noted here. A demilich's form is concentrated 
into a single portion of its original body, usually its skull. Part of the process 
of becoming a demilich includes the incorporation of costly gems into the retained 
body part; see Creating Soul Gems, below. 

Size: Medium-size and Large liches become Diminutive demiliches, Huge liches become 
Small demiliches, Gargantuan liches become Medium-size demiliches, and Colossal 
liches become Large demiliches. 

\textbf{Hit Dice:} As lich. 

\textbf{Speed:} Replace with fly 180 ft. (perfect) or the lich's supernatural (not 
natural) fly speed, whichever is better. 

\textbf{AC:} The demilich retains the lich's +5 natural armor bonus and gains an 
insight bonus equal to AC equal to its Hit Dice, as well as a probable size adjustment 
to AC. 

\textbf{Attack:} The demilich gains an insight bonus equal to its Hit Dice as a 
bonus on its touch attacks.

\textbf{Damage:} The demilich gains an enhanced touch attack over that of its previous 
lich form (it now uses its entire flying skull to make the touch attack), including 
paralyzing touch. The demilich's touch attack uses negative energy to deal 10d6+20 
points of damage to living creatures (no saving throw). Liches with other natural 
attacks lose them. 

\textbf{Special Attacks:} The demilich retains all the lich's special attacks and 
also gains those described below. 

\textit{\textbf{Trap the Soul (Su):}}\textit{ }A demilich can trap the souls of 
up to eight living creatures per day. To use this power, it selects any target 
it can see within 300 feet. The target is allowed a Fort saving throw (DC 10 + 
demilich's HD + demilich's Cha modifier). If the target makes its saving throw, 
it gains four negative levels (this does not count as a use of trap the soul). 
If the target fails its save, the soul of the target is instantly drawn from its 
body and trapped within one of the gems incorporated into the demilich's form. 
The gem gleams wickedly for 24 hours, indicating the captive soul within. The soulless 
body collapses in a mass of corruption and molders in a single round, reduced to 
dust. If left to its own devices, the demilich slowly devours the soul over 24 
hours---at the end of that time the soul is completely absorbed, and the victim 
is forever gone. If the demilich is overcome before the soul is eaten, crushing 
the gem releases the soul, after which time it is free to seek the afterlife or 
be returned to its body by the use of either \textit{resurrectio}n, \textit{true 
resurrection, clone, }or \textit{miracle. }If the demilich is overcome before the 
soul is eaten, crushing the gem releases the soul, after which time it is free 
to seek the afterlife or be returned to its body by the use of either \textit{resurrectio}n, 
\textit{true resurrection, clone, }or \textit{miracle. }A potential victim protected 
by a \textit{death ward }spell is not immune to trap the soul, but receives a +5 
bonus on its Fortitude saving throw and\textit{ }is effective against the level 
loss on a successful save. 

\textit{\textbf{Fear Aura (Su): }}Demiliches are shrouded in a dreadful aura of 
death and evil. Creatures of less than 5 HD in a 60- foot radius that look at the 
demilich must succeed at a Will save (DC 14 + demilich's Cha modifier) or be affected 
as though by \textit{fear }as cast by a 21st-level caster. 

\textit{\textbf{Paralyzing Touch (Su): }}Any living creature a demilich touches 
must succeed at a Fortitude save (DC 10 + demilich's HD + demilich's Cha modifier) 
or be permanently paralyzed. \textit{Remove paralysis }or any spell that can remove 
a curse can free the victim. The effect cannot be dispelled. Anyone paralyzed by 
a demilich seems dead, though a successful Spot check (DC 20) or Heal check (DC 
15) reveals that the victim is still alive. 

Spells: The demilich can cast any spells it could cast as a lich. 

\textit{Perfect Automatic Still Spell: }The demilich can cast all the spells it 
knows without gestures. 

\textbf{Spell-Like Abilities:} At will---\textit{alter self, astral projection, 
create greater undead, create undead, death knell, enervation, greater dispel magic, 
harm }(usually used to heal itself ), \textit{summon monster I-IX, telekinesis, 
}and \textit{weird; }2/day--- \textit{greater planar ally. }Demiliches use these 
abilities as casters of a level equal to their spellcaster level, but the save 
DCs are equal to 10 + the demilich's HD + the demilich's Charisma modifier. 

\textbf{Special Qualities:} The demilich retains all the lich's special qualities 
and also has those described below. 

\textit{\textbf{Magic Immunity (Ex):}}\textit{ }Demiliches are immune to all magical 
and supernatural effects, except as follows. A \textit{shatter }spell affects a 
demilich as if it were a crystalline creature, but deals half the damage normally 
indicated. A \textit{dispel evil }spell deals 3d6 points of damage (Fort save for 
half damage). \textit{Holy smite }spells affect demiliches normally. 

\textit{\textbf{Phylactery Transference (Su):}}\textit{ }Headbands, belts, rings, 
cloaks, and other wearable items kept in close association with the demilich's 
phylactery transfer all their benefits to the demilich no matter how far apart 
the demilich and the phylactery are located. The standard limits on types of items 
utilized simultaneously still apply. 

\textit{\textbf{Undead Traits:}}\textit{ }Immune to poison, \textit{sleep, }paralysis, 
stunning, disease, death, effects, necromantic effects, mind-affecting effects, 
and any effect requiring a Fortitude save unless it also works on objects. Not 
subject to critical hits, subdual damage, ability damage, ability drain, or energy 
drain. Negative energy heals. Not at risk of death from massive damage, but destroyed 
at 0 hit points or less. Darkvision 60 ft. Cannot be raised; resurrection works 
only if creature is willing. 

\textit{\textbf{Immunities (Ex):}}\textit{ }Demiliches are immune to cold, electricity, 
polymorph, and mind-affecting attacks. 

\textit{\textbf{Turn Resistance (Ex): }}A demilich has turn resistance +20. 

\textit{\textbf{Damage Reduction (Su):}}\textit{ }A demilich loses any previous 
damage reduction and instead has damage reduction 15/Epic and bludgeoning (15  
points of damage is subtracted from all melee attacks unless the weapon used is 
both an epic and a bludgeoning weapon). Vorpal weapons, no matter their enhancement 
bonus, ignore this damage reduction but do only half damage to a demilich (demiliches 
cannot be beheaded). 

\textit{\textbf{Resistances (Ex):}}\textit{ }Demiliches have acid resistance 20, 
fire resistance 20, and sonic resistance 20. 

\textbf{Saves:} Same as the lich. 

\textbf{Abilities:} A demilich gains +10 to Intelligence, Wisdom, and Charisma. 

\textbf{Skills:} Demiliches receive a +20 racial bonus on Hide, Listen, Move Silently, 
Search, Sense Motive, and Spot checks. Otherwise same as the lich (this overlaps 
with the previous racial bonus gained by the lich; it does not stack). 

\textbf{Feats:} Same as the lich. 

\textbf{Epic Feats:} Demiliches gain the feats Blinding Speed, Tenacious Magic, 
and Automatic Quicken Spell. 

\textbf{Climate/Terrain:} Same as the lich. 

\textbf{Organization:} Solitary or consistory (1 demilich and 3-6 liches). 

\textbf{Challenge Rating:} Same as the lich + 6. 

\textbf{Treasure:} Same as the lich. 

\textbf{Alignment:} Any evil. 

\textbf{Advancement:} By character class. 

\vspace{12pt}
\section*{DEMILICH CHARACTERS }

The process of becoming a demilich can be undertaken only by a lich acting of its 
own free will. The demilich retains all class abilities it had as a lich. 

\vspace{12pt}
\section*{CREATING SOUL GEMS }

Liches have phylacteries that allow them to reappear 1d10 days after their apparent 
death, as do demiliches. Demiliches also have eight soul gems, each of which acts 
like a phylactery in its own right. If all the soul gems, as well as the demilich's 
phylactery, are not destroyed after a demilich is downed, the demilich reappears 
1d10 days after its apparent death. The soul gems also allow the demilich to use 
its most devastating ability, trap the soul (see above). Each demilich must make 
its own soul gems, which requires the Craft Wondrous Item feat. The lich must be 
a sorcerer, wizard, or cleric of at least 21st level. Each soul gem costs 120,000 
gp and 4,800 XP to create and has a caster level equal to that of its creator at 
the time of creation. Soul gems appear as egg-shaped gems of wondrous quality. 
They are always incorporated directly into the concentrated form of the demilich. 

\vspace{12pt}
\section*{{\LARGE{}DEVASTATION VERMIN }}

\textbf{Vermin Traits: }Immune to mind-affecting effects (charms, compulsions, 
phantasms, patterns, and morale effects). 

\begin{tabular}{|>{\raggedright}p{63pt}|>{\raggedright}p{125pt}|>{\raggedright}p{125pt}|}
\hline
  & D\textbf{evastation Centipede } & D\textbf{evastation Spider }\tabularnewline
\hline
 & Colossal Vermin  & Colossal Vermin \tabularnewline
\hline
H\textbf{it Dice:}  & 128d8+1,152 (1,728 hp)  & 128d8+1,280 (1,856 hp) \tabularnewline
\hline
I\textbf{nitiative:}  & +13 (Dex)  & +14 (Dex) \tabularnewline
\hline
S\textbf{peed:}  & 80 ft.  & 60 ft., climb 40 ft. (80 ft., climb 40 ft.) \tabularnewline
\hline
A\textbf{C:}  & 55 (-8 size, +13 Dex, +40 natural)  & 58 (-8 size, +14 Dex, +42 
natural) \tabularnewline
\hline
B\textbf{ase Attack/ Grapple:} & +96/+128 & +96/+125\tabularnewline
\hline
A\textbf{ttack:} & Bite +99 (20d10+16 plus poison) melee & Bite +101 (18d10+19 
plus poison) melee\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +99 (20d10+16 plus poison) melee & Bite +101 (18d10+19 
plus poison) melee\tabularnewline
\hline
S\textbf{pace/Reach:}  & 50 ft./40 ft.  & 50 ft./40 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:}  & Poison  & Poison \tabularnewline
\hline
S\textbf{pecial Qualities:}  & Darkvision 300 ft., SR 50, DR 10/-  & Darkvision 
300 ft., SR 50, DR 10/-\tabularnewline
\hline
S\textbf{aves:}  & Fort +75, Ref +55, Will +42  & Fort +76, Ref +56, Will +42 \tabularnewline
\hline
A\textbf{bilities:}  & Str 33, Dex 37, Con 29, Int -, Wis 10, Cha 2 & Str 37, Dex 
39, Con 30, Int -, Wis 10, Cha 2\tabularnewline
\hline
S\textbf{kills:} & Climb +19, Hide +1, Jump +31, Spot +8 & Climb +21, Hide +2 (-2), 
Jump +25 (+39), Move Silently +14, Spot +8 (+12)\tabularnewline
\hline
C\textbf{limate/Terrain:}  & Any land  & Any land \tabularnewline
\hline
O\textbf{rganization:}  & Solitary or cluster (2-5)  & Solitary or cluster (2-5) 
\tabularnewline
\hline
C\textbf{hallenge Rating:}  & 39  & 41 \tabularnewline
\hline
T\textbf{reasure:}  & None  & None \tabularnewline
\hline
A\textbf{lignment:}  & Always neutral  & Always neutral \tabularnewline
\hline
A\textbf{dvancement:}  & None  & None \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{59pt}|>{\raggedright}p{127pt}|>{\raggedright}p{127pt}|}
\hline
  & D\textbf{evastation Scorpion } & D\textbf{evastation Beetle }\tabularnewline
\hline
 & Colossal Vermin  & Colossal Vermin \tabularnewline
\hline
H\textbf{it Dice:}  & 128d8+1,408 (1,984 hp)  & 128d8+2,304 (2,880 hp) \tabularnewline
\hline
I\textbf{nitiative:}  & +12 (Dex)  & +10 (Dex) \tabularnewline
\hline
S\textbf{peed:}  & 100 ft.  & 70 ft. \tabularnewline
\hline
A\textbf{C:}  & 60 (-8 size, +12 Dex, +46 natural)  & 72 (-8 size, +10 Dex, +60 
natural) \tabularnewline
\hline
A\textbf{ttacks:}  & 2 claws +102 melee, sting +97 melee  & Bite +104 melee \tabularnewline
\hline
D\textbf{amage:}  & Claw 10d10+14, sting 8d10+7 plus poison  & Bite 25d10+24 \tabularnewline
\hline
S\textbf{pace/Reach:}  & 50 ft./40 ft.  & 50 ft./40 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:}  & Improved grab, squeeze, poison  & Trample 30d10+24, 
acid cloud \tabularnewline
\hline
S\textbf{pecial Qualities:}  & Darkvision 300 ft., SR 50, DR 15/-   & Darkvision 
300 ft., SR 60, DR 20/-  \tabularnewline
\hline
S\textbf{aves:}  & Fort +77, Ref +54, Will +42  & Fort +84, Ref +52, Will +42 \tabularnewline
\hline
A\textbf{bilities:}  & Str 38, Dex 35, Con 32, Int -, Wis 10, Cha 2  & Str 42, 
Dex 31, Con 46, Int -, Wis 10, Cha 9 \tabularnewline
\hline
S\textbf{kills:}  & Climb +22, Hide +4, Jump +42, Spot +8 & Jump +31, Listen +0, 
Spot +0\tabularnewline
\hline
C\textbf{limate/Terrain:}  & Any land  & Any land \tabularnewline
\hline
O\textbf{rganization:}  & Solitary or cluster (2-5)  & Solitary or cluster (2-5) 
\tabularnewline
\hline
C\textbf{hallenge Rating:}  & 42  & 50 \tabularnewline
\hline
T\textbf{reasure:}  & None  & None \tabularnewline
\hline
A\textbf{lignment:}  & Always neutral  & Always neutral \tabularnewline
\hline
A\textbf{dvancement:}  & None  & None\tabularnewline
\hline
\end{tabular}

\section*{DEVASTATION CENTIPEDE }

\textbf{Poison (Ex): }Bite, Fort save (DC 93); initial and secondary damage 2d12 
temporary Dex. 

\textbf{Skills:} Devastation centipedes receive a +8 racial bonus on Climb, Hide, 
and Spot checks. 

\section*{DEVASTATION SPIDER }

\textbf{Poison (Ex):} Bite, Fort save (DC 94); initial and secondary damage 2d12 
temporary Con. 

\textbf{Web (Ex):} A single web strand is strong enough to support the spider and 
one creature of Gargantuan size. Web-spinning devastation spiders can cast a web 
eight times per day. This is similar to an attack with a net but has a maximum 
range of 150 feet, with a range increment of 30 feet, and is effective against 
targets of up to Gargantuan size. The web anchors the target in place, allowing 
no movement. An entangled creature can escape with a successful Escape Artist check 
(DC 38) or burst it with a Strength check (DC 44). Both are standard actions. Web-spinning 
spiders often create sheets of sticky webbing about 1,000 feet square. They usually 
position these sheets to snare flying creatures but can also try to trap prey on 
the ground. Approaching creatures must succeed at a Spot check (DC 20) to notice 
a web; otherwise, they stumble into it and become trapped as though by a successful 
web attack. Attempts to escape or burst the webbing gain a +5 bonus if the trapped 
creature has something to walk on or grab while pulling free. Each 5-foot section 
has 24 hit points, and sheet webs have damage reduction 20/fire. A devastation 
spider can move across its own sheet web at its climb speed and can determine the 
exact location of any creature caught in the web. 

\textbf{Skills:} Web-spinning devastation spiders receive a +4 racial bonus on 
Hide and Spot checks. Hunting spiders receive a +6 racial bonus on Jump checks 
and a +8 racial bonus on Spot checks (using the modifiers given in parentheses). 
*All devastation spiders gain a +8 competence bonus on Hide and Move Silently checks 
when using their webs. 

\section*{DEVASTATION SCORPION }

\textbf{Improved Grab (Ex):} To use this ability, the devastation scorpion must 
hit with its claw attack. If it gets a hold, it hangs on and squeezes. 

\textbf{Squeeze (Ex):} A devastation scorpion that gets a hold on a Gargantuan 
or smaller opponent automatically deals damage with both claws and its sting. 

\textbf{Poison (Ex):} Sting, Fort save (DC 95); initial and secondary damage 2d12 
temporary Str. 

\textbf{Skills:} Devastation scorpions receive a +4 racial bonus on Climb, Hide, 
and Spot checks. 

\section*{DEVASTATION BEETLE }

\textbf{Trample (Ex):} A devastation beetle can trample Gargantuan and smaller 
creatures for 30d10+24 points of damage. Opponents who do not make attacks of opportunity 
against the devastation beetle can attempt a Reflex save (DC 90) to halve the damage. 

\textbf{Acid Cloud (Ex):} A devastation beetle exudes a constant vapor that radiates 
outward in every direction for 60 feet. This vapor deals 6d6 points of acid damage 
each round to anyone caught in the cloud.

\vspace{12pt}
\section*{{\LARGE{}DRAGON, ADVANCED }}

The standard advancement rules allow dragons theoretically infinite progression 
even beyond the stage of great wyrm. The following rules let dragons improve more 
than their Hit Dice as they progress to epic heights of power. 

\vspace{12pt}
\textbf{Age Category:} A standard dragon gains one ``virtual age category'' for 
every 3 Hit Dice it gains beyond the great wyrm stage. Abilities that function 
once per day per age category or otherwise use the dragon's age category as part 
of a calculation use this adjusted number. Epic dragons gain one age category per 
5 Hit Dice beyond great wyrm. 

\vspace{12pt}
\textbf{Size: }One important element of dragon advancement is increasing size. 
The dragons that don't reach Colossal size by the great wyrm stage can never reach 
it according to the standard advancement rules. When advancing a dragon, consider 
its basic size group: lesser (white, black, brass, and copper dragons), ordinary 
(green, blue, and bronze dragons), greater (silver, red, and gold), or epic (force 
and prismatic). A dragon that is Tiny as a wyrmling is in the lesser group, a dragon 
that is Small as a wyrmling and never reaches Colossal size is in the ordinary 
group, and a dragon that is Small to Large as a wyrmling and reaches Colossal by 
the great wyrm stage is in the greater group. A lesser dragon becomes Colossal 
when it gains two age categories (6 Hit Dice) beyond great wyrm. It increases to 
Colossal+ when it gains an additional four age categories (12 HD). An ordinary 
dragon becomes Colossal when it gains one age category (3 Hit Dice) beyond great 
wyrm. It increases to Colossal+ when it gains an additional four age categories 
(12 HD). A greater dragon becomes Colossal+ when it gains four age categories (12 
HD) more than it needed to reach the Colossal size. 

Epic dragons, already Colossal+ by the time they reach the great wyrm stage, do 
not increase in size thereafter. Use the information presented under the epic dragon 
entry, below, to determine the statistics of a Colossal+ dragon. 

\vspace{12pt}
\textbf{Armor Class:} A dragon's natural armor bonus increases by +1 for every 
Hit Die it gains beyond the great wyrm stage. (This rule applies for lesser dragon 
advancement as well, since natural armor and Hit Dice always increase at the same 
rate.) 

\vspace{12pt}
\textbf{Breath Weapon:} If a dragon's breath weapon deals damage, the damage typically 
increases by 2 dice for every virtual age category the dragon gains. The two exceptions 
among standard dragons\textit{ }are the brass and white dragons, whose breath weapon 
damage increases by only 1 die per age category. The saving throw DC against a 
dragon's breath weapon remains 10 + 1/2 the dragon's Hit Dice + its Constitution 
modifier. 

\vspace{12pt}
\textbf{Spell Resistance:} For standard advanced dragons, spell resistance increases 
by 2 per additional age category. For epic dragons, spell resistance increases 
by 3 (for force dragons) or 6 (for prismatic dragons) per additional age category. 

\vspace{12pt}
\textbf{Speed:} When a dragon becomes Colossal, its fly speed increases by 50 feet 
and its maneuverability becomes clumsy. When it reaches Colossal+, its fly speed 
increases by another 50 feet, and its maneuverability remains clumsy. A dragon's 
land speed and other special movement types (swim, burrow, and so on) do not change. 

\vspace{12pt}
\textbf{Ability Scores: }A great wyrm's Strength and Constitution scores both increase 
by +2 for every virtual age category the dragon gains. Its Dexterity remains unchanged. 
Its Intelligence, Wisdom, and Charisma increase by +2 for every two age categories 
the dragon gains. 

\vspace{12pt}
\textbf{Special Abilities:} Dragons do not gain additional spell-like abilities, 
but their damage reduction continues to increase as they gain Hit Dice. For most 
dragons (the exceptions being the force and prismatic dragons), the great wyrm's 
damage reduction increases to 15/epic after exceeding the great wyrm age category 
and the dragon's natural weapons count as epic weapons for the purposes of bypassing 
damage reduction.

Force and prismatic dragons have the following DR progression: wyrmling 10/magic, 
young 15/magic, young adult 15/epic, mature adult 20/epic, very old 25/epic.

\textit{Caster Level: }A great wyrm's caster level increases by 2 for every virtual 
age category the dragon gains. As with the epic dragons described later, advanced 
dragons gain the Improved Spell Capacity feat as a bonus feat once for every three 
caster levels above 20th. 

\vspace{12pt}
\textbf{Feats:} Like ordinary dragons, advanced dragons receive one feat for every 
3 Hit Dice they have. Feats gained after the dragon attains great wyrm status can 
be epic feats. 

\vspace{12pt}
\textbf{Challenge Rating:} For standard dragons, CR increases by 2 per additional 
age category. For epic dragons, CR increases by 3 per additional age category. 
All other dragon statistics are as presented for dragons in general and specific 
dragon varieties. 

\vspace{12pt}
\section*{{\LARGE{}DRAGON,EPIC }}

All epic dragons speak Draconic. 

\textbf{Epic Dragon Age Categories }

\begin{tabular}{|>{\raggedright}p{8pt}|>{\raggedright}p{50pt}|>{\raggedright}p{49pt}|}
\hline
  & C\textbf{ategory } & A\textbf{ge (Years)}\tabularnewline
\hline
1  & Wyrmling  & 0-10\tabularnewline
\hline
2  & Very young  & 11-30\tabularnewline
\hline
3  & Young  & 31-50\tabularnewline
\hline
4  & Juvenile  & 51-100\tabularnewline
\hline
5  & Young adult  & 101-200\tabularnewline
\hline
6  & Adult  & 201-400\tabularnewline
\hline
7  & Mature adult  & 401-800\tabularnewline
\hline
8  & Old  & 801-1,200\tabularnewline
\hline
9  & Very old  & 1,201-1,600\tabularnewline
\hline
10  & Ancient  & 1,601-2,000\tabularnewline
\hline
11  & Wyrm  & 2,001-2,400\tabularnewline
\hline
12  & Great wyrm  & 2,401+\tabularnewline
\hline
\end{tabular} 

\vspace{12pt}
\section*{COMBAT\textbf{ }}

Epic dragons share all the combat characteristics of standard dragons\textit{. 
}Even wyrmling epic dragons are Huge, however, and are capable of making tail sweep 
and crush attacks. Although there is no size category larger than Colossal, the 
oldest epic dragons deal more damage with their attacks than other Colossal dragons, 
as shown on the Epic Dragon Face and Reach and Epic Dragon Attacks tables below. 
In addition, the breath weapon of the oldest epic dragons is a larger cone than 
most Colossal dragons possess. The size modifier for these dragons remains -8. 

\vspace{12pt}
\textbf{Epic Dragon Space and Reach }

\begin{tabular}{|>{\raggedright}p{45pt}|>{\raggedright}p{53pt}|}
\hline
S\textbf{ize } & S\textbf{pace/Reach}\tabularnewline
\hline
Huge  & 15 ft./10 ft. \tabularnewline
\hline
Gargantuan  & 20 ft./15 ft.\tabularnewline
\hline
Colossal  & 30 ft./20 ft. \tabularnewline
\hline
Colossal+  & 30 ft./20 ft. \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\textbf{Epic Dragon Attacks }

\begin{tabular}{|>{\raggedright}p{45pt}|>{\raggedright}p{23pt}|>{\raggedright}p{32pt}|>{\raggedright}p{33pt}|>{\raggedright}p{44pt}|>{\raggedright}p{33pt}|>{\raggedright}p{53pt}|}
\hline
S\textbf{ize } & 1\textbf{ Bite} & 2\textbf{ Claws} & 2\textbf{ Wings} & 1\textbf{ 
Tail Slap} & 1\textbf{ Crush} & 1\textbf{ Tail Sweep}\tabularnewline
\hline
Huge  & 2d8 & 2d6 & 1d8 & 2d6 & 2d8 & 2d4\tabularnewline
\hline
Gargantuan & 4d6 & 2d8 & 2d6 & 2d8 & 4d6 & 2d6\tabularnewline
\hline
Colossal & 4d8 & 4d6 & 2d8 & 4d6 & 4d8 & 2d8\tabularnewline
\hline
Colossal+ & 8d6 & 4d8 & 4d6 & 4d8 & 8d6 & 4d6\tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\textbf{Epic Dragon Breath Weapons }

\begin{tabular}{|>{\raggedright}p{64pt}|>{\raggedright}p{90pt}|}
\hline
S\textbf{ize } & C\textbf{one* (Length) }\tabularnewline
\hline
Huge  & 50 ft. \tabularnewline
\hline
Gargantuan  & 60 ft. \tabularnewline
\hline
Colossal  & 70 ft. \tabularnewline
\hline
Colossal+  & 80 ft. \tabularnewline
\hline
\multicolumn{2}{|p{154pt}|}{* A cone is as high and wide as its length.}\tabularnewline
\hline
\end{tabular} 

\vspace{12pt}
\parindent=3pt
Wyrmling epic dragons possess certain abilities that common dragons do not acquire 
until later in life, including frightful presence, spellcasting, and spell resistance. 
An epic dragon with a caster level above 20th gains the Improved Spell Capacity 
feat as a bonus feat once for every three caster levels above 20th. Any epic dragon 
can choose epic feats for which it otherwise meets the prerequisites. Epic dragons 
also have access to dragon feats\textit{, }as well as the additional feats described 
in this book. 

\vspace{12pt}
\parindent=0pt
\textbf{Empower Spell-Like Ability:} An epic dragon can apply the Empower Spell 
feat to one of its spell-like abilities that it can use at least twice per day. 
If the dragon has limited use of the spell-like ability, the empowered ability 
uses up two uses of the ability. If the dragon can use the ability at will, there 
is no cost for empowering the ability. The dragon does not need to specify a spell-like 
ability when it chooses this feat; it can apply the feat to any spell-like ability 
it possesses. 

\vspace{12pt}
\textbf{Maximize Spell-Like Ability: }The dragon can apply the Maximize Spell feat 
to one of its spell-like abilities that it can use at least three times per day. 
If the dragon has limited use of the spell-like ability, the maximized ability 
uses up three uses of the ability. If the dragon can use the ability at will, there 
is no cost for maximizing the ability. The dragon does not need to specify a spell-like 
ability when it chooses this feat; it can apply the feat to any spell-like ability 
it possesses. 

\vspace{12pt}
Epic dragons fly even faster than common dragons, and can cover greater distances. 

\textbf{Epic Dragon Overland Flying Speeds }

\begin{tabular}{|>{\raggedright}p{41pt}|>{\raggedright}p{44pt}|>{\raggedright}p{44pt}|>{\raggedright}p{44pt}|}
\hline
  --------------- & \multicolumn{3}{p{133pt}|}{ \textbf{Flight Speed --------------- 
}}\tabularnewline
\hline
 & 2\textbf{50 feet } & 3\textbf{00 feet } & 3\textbf{50 feet }\tabularnewline
\hline
O\textbf{ne Hour } & \multicolumn{3}{p{133pt}|}{}\tabularnewline
\hline
Normal  & 40 miles  & 50 miles  & 60 miles \tabularnewline
\hline
Hustle  & 80 miles  & 100 miles  & 120 miles \tabularnewline
\hline
\multicolumn{4}{|p{175pt}|}{}\tabularnewline
\hline
O\textbf{ne Day } & \multicolumn{3}{p{133pt}|}{}\tabularnewline
\hline
Normal  & 320 miles  & 400 miles  & 480 miles \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
Except as noted here, epic dragons conform to the general information that applies 
to all dragons given\textit{.}

\vspace{12pt}
FORCE DRAGON 

Dragon (Force) 

\textbf{Climate/Terrain:} Any land and underground 

\textbf{Organization:} Wyrmling, very young, young, juvenile and young adult: solitary 
or clutch (2-5); adult, mature adult, old, very old, ancient, wyrm, or great wyrm: 
solitary, pair, or family (1-2 and 2-5 offspring) 

\textbf{Challenge Rating:} Wyrmling 13; very young 19; young 21; juvenile 25; young 
adult 28; adult 31; mature adult 35; old 37; very old 39; ancient 43; wyrm 50; 
great wyrm 59 

\textbf{Treasure:} Double standard 

\textbf{Alignment:} Usually neutral 

Advancement Range: Wyrmling 21-24 HD (Huge); very young 26-29 HD (Huge); young 
31-34 HD (Gargantuan); juvenile 36-39 HD (Gargantuan); young adult 41-44 HD (Gargantuan); 
adult 46-49 HD (Gargantuan); mature adult 51-54 HD (Colossal); old 56-59 HD (Colossal); 
very old 61-64 HD (Colossal); ancient 66-69 HD (Colossal); wyrm 71-74 HD (Colossal+); 
great wyrm 76+ HD (Colossal+) 

\vspace{12pt}
\textbf{Force Dragons by Age }

\begin{tabular}{|>{\raggedright}p{17pt}|>{\raggedright}p{7pt}|>{\raggedright}p{19pt}|>{\raggedright}p{6pt}|>{\raggedright}p{7pt}|>{\raggedright}p{6pt}|>{\raggedright}p{4pt}|>{\raggedright}p{5pt}|>{\raggedright}p{6pt}|>{\raggedright}p{13pt}|>{\raggedright}p{10pt}|>{\raggedright}p{6pt}|>{\raggedright}p{6pt}|>{\raggedright}p{6pt}|>{\raggedright}p{18pt}|>{\raggedright}p{15pt}|}
\hline
A{\small{}\textbf{ge }} & \subsubsection*{S{\small{}\textbf{ize}}} & H{\small{}\textbf{it 
Dice (hp)}} & S{\small{}\textbf{tr}} & D{\small{}\textbf{ex}} & C{\small{}\textbf{on}} & I{\small{}\textbf{nt}} & W{\small{}\textbf{is}} & C{\small{}\textbf{ha}} & B{\small{}\textbf{ase 
Attack/}}\linebreak{}
{\small{}\textbf{Grapple}} & A{\small{}\textbf{ttack}} & F{\small{}\textbf{ort 
Save}} & R{\small{}\textbf{ef Save}} & W{\small{}\textbf{ill Save}} & B{\small{}\textbf{reath 
Weapon (DC)}} & F{\small{}\textbf{rightful Presence DC}}\tabularnewline
\hline
W{\small{}yrmling } & H & 2{\small{}0d12+100 (230)} & 3{\small{}1} & 1{\small{}0} & 2{\small{}1} & 1{\small{}8} & 1{\small{}9} & 1{\small{}8} & +{\small{}20/+38} & +{\small{}28} & +{\small{}17} & +{\small{}12} & +{\small{}16} & 5{\small{}d12 
(25)} & 2{\small{}4}\tabularnewline
\hline
V{\small{}ery young} & H & 2{\small{}5d12+150 (312)} & 3{\small{}5} & 1{\small{}0} & 2{\small{}3} & 2{\small{}2} & 2{\small{}3} & 2{\small{}2} & +{\small{}25/+45} & +{\small{}35} & +{\small{}20} & +{\small{}14} & +{\small{}20} & 1{\small{}0d12 
(28)} & 2{\small{}8}\tabularnewline
\hline
Y{\small{}oung} & G & 3{\small{}0d12+210 (405)} & 3{\small{}9} & 1{\small{}0} & 2{\small{}5} & 2{\small{}6} & 2{\small{}7} & 2{\small{}6} & +{\small{}30/+56} & +{\small{}40} & +{\small{}24} & +{\small{}17} & +{\small{}25} & 1{\small{}5d12 
(32)} & 3{\small{}3}\tabularnewline
\hline
J{\small{}uvenile} & G & 3{\small{}5d12+315 (542)} & 4{\small{}3} & 1{\small{}0} & 2{\small{}9} & 3{\small{}0} & 3{\small{}1} & 3{\small{}0} & +{\small{}35/+63} & +{\small{}47} & +{\small{}28} & +{\small{}19} & +{\small{}29} & 2{\small{}0d12 
(36)} & 3{\small{}7}\tabularnewline
\hline
Y{\small{}oung adult} & G & 4{\small{}0d12+440 (700)} & 4{\small{}7} & 1{\small{}0} & 3{\small{}3} & 3{\small{}4} & 3{\small{}5} & 3{\small{}4} & +{\small{}40/+70} & +{\small{}54} & +{\small{}33} & +{\small{}22} & +{\small{}34} & 2{\small{}5d12 
(41)} & 4{\small{}2}\tabularnewline
\hline
A{\small{}dult} & G & 4{\small{}5d12+585 (877)} & 5{\small{}1} & 1{\small{}0} & 3{\small{}7} & 3{\small{}8} & 3{\small{}9} & 3{\small{}8} & +{\small{}45/+77} & +{\small{}61} & +{\small{}37} & +{\small{}24} & +{\small{}38} & 3{\small{}0d12 
(45)} & 4{\small{}6}\tabularnewline
\hline
M{\small{}ature adult} & C & 5{\small{}0d12+750 (1,075)} & 5{\small{}5} & 1{\small{}0} & 4{\small{}1} & 4{\small{}2} & 4{\small{}3} & 4{\small{}2} & +{\small{}50/+88} & +{\small{}64} & +{\small{}42} & +{\small{}27} & +{\small{}43} & 3{\small{}5d12 
(50)} & 5{\small{}1}\tabularnewline
\hline
O{\small{}ld} & C & 5{\small{}5d12+935 (1,292)} & 5{\small{}9} & 1{\small{}0} & 4{\small{}5} & 4{\small{}6} & 4{\small{}7} & 4{\small{}6} & +{\small{}55/+95} & +{\small{}71} & +{\small{}46} & +{\small{}29} & +{\small{}47} & 4{\small{}0d12 
(54)} & 5{\small{}5}\tabularnewline
\hline
V{\small{}ery old} & C & 6{\small{}0d12+1,140 (1,530)} & 6{\small{}3} & 1{\small{}0} & 4{\small{}9} & 5{\small{}0} & 5{\small{}1} & 5{\small{}0} & +{\small{}60/+102} & +{\small{}78} & +{\small{}51} & +{\small{}32} & +{\small{}52} & 4{\small{}5d12 
(59)} & 6{\small{}0}\tabularnewline
\hline
A{\small{}ncient} & C & 6{\small{}5d12+1,365 (1,787)} & 6{\small{}7} & 1{\small{}0} & 5{\small{}3} & 5{\small{}4} & 5{\small{}5} & 5{\small{}4} & +{\small{}65/+109} & +{\small{}85} & +{\small{}55} & +{\small{}34} & +{\small{}56} & 5{\small{}0d12 
(63)} & 6{\small{}4}\tabularnewline
\hline
W{\small{}yrm} & C{\small{}+} & 7{\small{}0d12+1,610 (2,065)} & 7{\small{}1} & 1{\small{}0} & 5{\small{}7} & 5{\small{}8} & 5{\small{}9} & 5{\small{}8} & +{\small{}70/+116} & +{\small{}92} & +{\small{}60} & +{\small{}37} & +{\small{}61} & 5{\small{}5d12 
(68)} & 6{\small{}9}\tabularnewline
\hline
G{\small{}reat wyrm} & C{\small{}+} & 7{\small{}5d12+1,875 (2,362)} & 7{\small{}5} & 1{\small{}0} & 6{\small{}1} & 6{\small{}2} & 6{\small{}3} & 6{\small{}2} & +{\small{}75/+123} & +{\small{}99} & +{\small{}64} & +{\small{}39} & +{\small{}65} & 6{\small{}0d12 
(72)} & 7{\small{}3}\tabularnewline
\hline
\end{tabular} 

\vspace{24pt}
\textbf{Force Dragon Abilities by Age }

\begin{tabular}{|>{\raggedright}p{25pt}|>{\raggedright}p{57pt}|>{\raggedright}p{22pt}|>{\raggedright}p{69pt}|>{\raggedright}p{62pt}|>{\raggedright}p{19pt}|>{\raggedright}p{8pt}|}
\hline
A{\small{}\textbf{ge }} & S{\small{}\textbf{peed }} & I{\small{}\textbf{nitiative}} & A{\small{}\textbf{C 
}} & S{\small{}\textbf{pecial Abilities }} & C{\small{}\textbf{aster Level}} & S{\small{}\textbf{R}}\tabularnewline
\hline
W{\small{}yrmling} & 6{\small{}0 ft., fly 250 ft. (poor)} & +{\small{}0} & 3{\small{}1 
(-2 size, +4 deflection, +19 natural), touch 12, flat-footed 31} & D{\small{}amage 
reduction 10/magic, deflecting force, immune to force } & 3{\small{}rd} & 2{\small{}4}\tabularnewline
\hline
V{\small{}ery young} & 6{\small{}0 ft., fly 250 ft. (poor)} & +{\small{}0} & 3{\small{}8 
(-2 size, +6 deflection, +24 natural), touch 14, flat-footed 38} &  & 6{\small{}th} & 2{\small{}7}\tabularnewline
\hline
Y{\small{}oung} & 6{\small{}0 ft., fly 300 ft. (clumsy)} & +{\small{}0} & 4{\small{}3 
(-4 size, +8 deflection, +29 natural), touch 14, flat-footed 43} & D{\small{}amage 
reduction 15/magic } & 9{\small{}th} & 3{\small{}0}\tabularnewline
\hline
J{\small{}uvenile} & 6{\small{}0 ft., fly 300 ft. (clumsy)} & +{\small{}0} & 5{\small{}0 
(-4 size, +10 deflection, +34 natural), touch 16, flat-footed 50} & r{\small{}\textit{esilient 
sphere, }}{\small{}blur} & 1{\small{}2th} & 3{\small{}3}\tabularnewline
\hline
Y{\small{}oung adult} & 6{\small{}0 ft., fly 300 ft. (clumsy)} & +{\small{}0} & 5{\small{}7 
(-4 size, +12 deflection, +39 natural), touch 18, flat-footed 57} & D{\small{}amage 
reduction 15/epic} & 1{\small{}5th} & 3{\small{}6}\tabularnewline
\hline
A{\small{}dult} & 6{\small{}0 ft., fly 300 ft. (clumsy) } & +{\small{}0} & 6{\small{}4 
(-4 size, +14 deflection, +44 natural), touch 20, flat-footed 64} & W{\small{}\textit{all 
of force }} & 1{\small{}8th} & 3{\small{}9}\tabularnewline
\hline
M{\small{}ature adult} & 6{\small{}0 ft., fly 300 ft. (clumsy)} & +{\small{}0} & 6{\small{}7 
(-8 size, +16 deflection, +49 natural), touch 18, flat-footed 67} & D{\small{}amage 
reduction 20/epic} & 2{\small{}1st} & 4{\small{}2}\tabularnewline
\hline
O{\small{}ld} & 6{\small{}0 ft., fly 300 ft. (clumsy)} & +{\small{}0} & 7{\small{}4 
(-8 size, +18 deflection, +54 natural), touch 20, flat-footed 74} & F{\small{}\textit{orcecage, 
}}{\small{}displacement} & 2{\small{}4th} & 4{\small{}5}\tabularnewline
\hline
V{\small{}ery old} & 6{\small{}0 ft., fly 300 ft. (clumsy)} & +{\small{}0} & 8{\small{}1 
(-8 size, +20 deflection, +59 natural), touch 22, flat-footed 81} & D{\small{}amage 
reduction 25/epic} & 2{\small{}7th} & 4{\small{}8}\tabularnewline
\hline
A{\small{}ncient} & 6{\small{}0 ft., fly 300 ft. (clumsy)} & +{\small{}0} & 8{\small{}8 
(-8 size, +22 deflection, +64 natural), touch 24, flat-footed 88} & M{\small{}\textit{aze}} & 3{\small{}0th} & 5{\small{}1}\tabularnewline
\hline
W{\small{}yrm} & 6{\small{}0 ft., fly 350 ft. (clumsy)} & +{\small{}0} & 9{\small{}5 
(-8 size, +24 deflection, +69 natural), touch 26, flat-footed 95} &  & 3{\small{}3rd} & 5{\small{}4}\tabularnewline
\hline
G{\small{}reat wyrm} & 6{\small{}0 ft., fly 350 ft. (clumsy)} & +{\small{}0} & 1{\small{}02 
(-8 size, +26 deflection, +74 natural), touch 28, flat-footed 102} & t{\small{}\textit{elekinetic 
sphere, }}{\small{}invisibility} & 3{\small{}6th} & 5{\small{}7}\tabularnewline
\hline
\end{tabular}

A force dragon's natural weapons are treated as magic (wyrmling - juvenile) or 
epic (young adult - great wyrm) for the purpose of overcoming damage reduction.

Combat\textbf{ }

\textbf{Breath Weapon (Su):} A force dragon has one type of breath weapon, a cone 
of force that tears through creatures and objects in its area to devastating effect. 

\textbf{Immune to Force (Ex):} A force dragon cannot be harmed by any force effect, 
including \textit{magic missiles, explosive runes, mage's sword, }the \textit{sphere 
}spells, or any other spell or effect with the Force descriptor. It can move freely 
through barriers made of force, such as a \textit{wall of forc}e. 

\textbf{Deflecting Force (Su):} A force dragon is protected by a shield of force 
that provides a deflection bonus to its AC. The deflection bonus is equal to the 
dragon's Charisma bonus. 

\textbf{Blur (Su): }When a force dragon reaches juvenile age, the force surrounding 
its body distorts the light around it so that it is hard to discern the dragon's 
exact location. This functions as a continuous \textit{blur }spell, giving the 
dragon concealment (20\% miss chance). 

\textbf{Displacement (Su):} An old force dragon's deflecting force bends light 
even more ex-tremely, giving the dragon the benefit of the \textit{displacement 
}spell (50\% miss chance). This supersedes its \textit{blur }ability. 

\textbf{Invisibility (Su):} A great wyrm force dragon is a creature of pure force, 
and is completely invisible (as though protected by\textit{ greater invisibilit}y). 
A creature that can see invisible creatures must still contend with the dragon's 
\textit{displacement }ability.

\textbf{Spell-Like Abilities:} 3/day---\textit{forcecage, maze, resilient sphere, 
telekinetic sphere, wall of forc}e. The DCs are Charisma-based.

\vspace{12pt}
\section*{\textbf{Sample Force Dragon,Adult }}

Gargantuan Dragon (Force) 

\textbf{Hit Dice:} 45d12+585 (877 hp) 

\textbf{Initiative:} +0 

\textbf{Speed:} 60 ft., fly 300 ft. (clumsy) 

\textbf{AC:} 64 (-4 size, +14 deflection, +44 natural) 

\textbf{Base Attack/Grapple:} +45/+77

\textbf{Attack: } Bite +61 (4d6+20/19-20) melee

\textbf{Full Attack: }Bite +61 (4d6+20/19-20) melee, 2 claws +56 (2d6+10/19-20) 
melee, 2 wings +56 (2d6+10/19-20) melee, tail slap +56 (2d8+30/19-20) melee

\textbf{Space/Reach:} 20 ft./15 ft. 

\textbf{Special Attacks:} Crush 4d6+30 (DC 45), tail sweep 2d6+30 (DC 45), breath 
weapon, frightful presence (DC 46), spells (caster level 18th), spell-like abilities 

\textbf{Special Qualities:} Immunities, DR 15/epic, SR 39, blindsense 60ft., keen 
senses, deflecting force, blur (20\% miss chance) 

\textbf{Saves:} Fort +37, Ref +24, Will +38 

\textbf{Abilities:} Str 51, Dex 10, Con 37, Int 38, Wis 39, Cha 38 

\textbf{Skills:} Appraise +50, Bluff +62, Concentration +61, Craft (alchemy) +50, 
Diplomacy +74, Escape Artist +48, Gather Information +20, Intimidate +68, Jump 
+32, Knowledge (all) +56, Listen +64, Search +72, Sense Motive +50, Spellcraft 
+68, Spot +74

\textbf{Feats:} Alertness, Cleave, Combat Reflexes, Empower Spell-Like Ability, 
Flyby Attack, Hover, Improved Critical (bite), Improved Critical (claws), Improved 
Critical (wings), Improved Critical (tail slap), Improved Initiative, Improved 
Sunder, Maximize Spell-Like Ability, Power Attack, Spell Knowledge, Wingover 

\textbf{Climate/Terrain:} Any land and underground 

\textbf{Organization:} Solitary, pair, or family (1-2 and 2-5 offspring) 

\textbf{Challenge Rating:} 31 

\textbf{Treasure:} Double standard 

\textbf{Alignment:} Usually neutral 

\textbf{Advancement:} 46-49 HD (Gargantuan) 

A force dragon's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

\textbf{Breath Weapon (Su):} Cone of force, 60 ft., damage 30d12, Ref half (DC 
45). The DC is Constitution-based.

\textbf{Spell-Like Abilities:} 3/day---\textit{resilient sphere, wall of forc}e. 
Caster level 18th; save DC 24 + spell level. 

\textbf{Immunities (Ex):} Immune to \textit{sleep }and paralysis, immune to force 
effects. 

\textit{\textbf{Sorcerer Spells Known}}\textit{ }(6/10/10/ 9/9/9/9/8/7/5; base 
DC 24 + spell level): \textit{0---arcane mark, dancing lights, detect magic, flare, 
ghost sound, light, mending, open/close, read magic; }1st---\textit{disguise self, 
comprehend languages, detect secret doors, magic missile, true strike; }2nd--- 
\textit{bull's strength, cat's grace, darkness, locate object, mirror image; }3rd---\textit{clairaudience/clairvoyance, 
fireball, haste, lightning bolt; }4th---\textit{charm monster, greater invisibility, 
scrying, stoneskin; }5th---\textit{contact other plane, dream, hold monster, teleport; 
}6th---\textit{analyze dweomer, greater dispel magic, true seeing; }7th---\textit{forcecage, 
spell turning, summon monster VII; }8th---\textit{polymorph any object, protection 
from spells; }9th--- \textit{time stop. }

\textit{Possessions: Crystal ball with true seeing, circlet of eyes*, pearl of 
power }(7th-level spell), 8,000 gp. *Variant of \textit{robe of eyes; }benefit 
included in Search and Spot skill modifiers. 

\vspace{12pt}
PRISMATIC DRAGON 

Dragon (Light) 

\textbf{Climate/Terrain:} Any land and underground 

\textbf{Organization:} Wyrmling, very young, young, juvenile and young adult: solitary 
or clutch (2-5); adult, mature adult, old, very old, ancient, wyrm, or great wyrm: 
solitary, pair, or family (1-2 and 2-5 offspring) 

\textbf{Challenge Rating:} Wyrmling 14; very young 16; young 22; juvenile 30; young 
adult 34; adult 39; mature adult 43; old 48; very old 52; ancient 57; wyrm 61; 
great wyrm 66 

\textbf{Treasure:} Double standard 

\textbf{Alignment:} Usually neutral 

Advancement Range: Wyrmling 24-27 HD (Huge); very young 29-32 HD (Huge); young 
34-37 HD (Gargantuan); juvenile 39-42 HD (Gargantuan); young adult 44-47 HD (Gargantuan); 
adult 49-52 HD (Gargantuan); mature adult 54-57 HD (Colossal); old 59-62 HD (Colossal); 
very old 64-67 HD (Colossal); ancient 69-72 HD (Colossal+); wyrm 74-77 HD (Colossal+); 
great wyrm 79+ HD (Colossal+) 

\vspace{12pt}
\section*{\textbf{Prismatic Dragons by Age }}

\section*{\begin{tabular}{|>{\raggedright}p{17pt}|>{\raggedright}p{7pt}|>{\raggedright}p{19pt}|>{\raggedright}p{6pt}|>{\raggedright}p{6pt}|>{\raggedright}p{7pt}|>{\raggedright}p{5pt}|>{\raggedright}p{7pt}|>{\raggedright}p{6pt}|>{\raggedright}p{13pt}|>{\raggedright}p{10pt}|>{\raggedright}p{6pt}|>{\raggedright}p{6pt}|>{\raggedright}p{6pt}|>{\raggedright}p{14pt}|>{\raggedright}p{15pt}|}
\hline
A{\small{}\textbf{ge }}} & \subsection*{S{\small{}\textbf{ize}}} & H{\small{}\textbf{it 
Dice (hp)}} & S{\small{}\textbf{tr}} & D{\small{}\textbf{ex}} & C{\small{}\textbf{on}} & I{\small{}\textbf{nt}} & W{\small{}\textbf{is}} & C{\small{}\textbf{ha}} & B{\small{}\textbf{ase 
Attack/}}\linebreak{}
{\small{}\textbf{Grapple}} & A{\small{}\textbf{ttack}} & F{\small{}\textbf{ort 
Save}} & R{\small{}\textbf{ef Save}} & W{\small{}\textbf{ill Save}} & B{\small{}\textbf{reath 
Weapon DC}} & F{\small{}\textbf{rightful Presence DC}}\tabularnewline
\hline
W{\small{}yrmling } & H & 2{\small{}3d12+138 (287)} & 2{\small{}9} & 1{\small{}0} & 2{\small{}3} & 2{\small{}0} & 2{\small{}1} & 2{\small{}0} & +{\small{}23/+40} & +{\small{}30} & +{\small{}19} & +{\small{}13} & +{\small{}18} & 2{\small{}7} & 2{\small{}6}\tabularnewline
\hline
V{\small{}ery young } & H & 2{\small{}8d12+196 (378)} & 3{\small{}3} & 1{\small{}0} & 2{\small{}5} & 2{\small{}4} & 2{\small{}5} & 2{\small{}4} & +{\small{}28/+47} & +{\small{}37} & +{\small{}23} & +{\small{}16} & +{\small{}23} & 3{\small{}1} & 3{\small{}1}\tabularnewline
\hline
Y{\small{}oung } & G & 3{\small{}3d12+297 (511)} & 3{\small{}7} & 1{\small{}0} & 2{\small{}9} & 2{\small{}8} & 2{\small{}9} & 2{\small{}8} & +{\small{}33/+58} & +{\small{}42} & +{\small{}27} & +{\small{}18} & +{\small{}27} & 3{\small{}5} & 3{\small{}5}\tabularnewline
\hline
J{\small{}uvenile } & G & 3{\small{}8d12+418 (665)} & 4{\small{}1} & 1{\small{}0} & 3{\small{}3} & 3{\small{}2} & 3{\small{}3} & 3{\small{}2} & +{\small{}38/+65} & +{\small{}49} & +{\small{}32} & +{\small{}21} & +{\small{}32} & 4{\small{}0} & 4{\small{}0}\tabularnewline
\hline
Y{\small{}oung adult } & G & 4{\small{}3d12+559 (838)} & 4{\small{}5} & 1{\small{}0} & 3{\small{}7} & 3{\small{}6} & 3{\small{}7} & 3{\small{}6} & +{\small{}43/+72} & +{\small{}56} & +{\small{}36} & +{\small{}23} & +{\small{}36} & 4{\small{}4} & 4{\small{}4}\tabularnewline
\hline
A{\small{}dult } & C & 4{\small{}8d12+720 (1,032)} & 4{\small{}9} & 1{\small{}0} & 4{\small{}1} & 4{\small{}0} & 4{\small{}1} & 4{\small{}0} & +{\small{}48/+83} & +{\small{}59} & +{\small{}41} & +{\small{}26} & +{\small{}41} & 4{\small{}9} & 4{\small{}9}\tabularnewline
\hline
M{\small{}ature adult } & C & 5{\small{}3d12+901 (1,245)} & 5{\small{}3} & 1{\small{}0} & 4{\small{}5} & 4{\small{}4} & 4{\small{}5} & 4{\small{}4} & +{\small{}53/+90} & +{\small{}66} & +{\small{}45} & +{\small{}28} & +{\small{}45} & 5{\small{}3} & 5{\small{}3}\tabularnewline
\hline
O{\small{}ld } & C & 5{\small{}8d12+1,102 (1,479)} & 5{\small{}7} & 1{\small{}0} & 4{\small{}9} & 4{\small{}8} & 4{\small{}9} & 4{\small{}8} & +{\small{}58/+97} & +{\small{}73} & +{\small{}50} & +{\small{}31} & +{\small{}50} & 5{\small{}8} & 5{\small{}8}\tabularnewline
\hline
V{\small{}ery old } & C & 6{\small{}3d12+1,323 (1,732)} & 6{\small{}1} & 1{\small{}0} & 5{\small{}3} & 5{\small{}2} & 5{\small{}3} & 5{\small{}2} & +{\small{}63/+104} & +{\small{}80} & +{\small{}54} & +{\small{}33} & +{\small{}54} & 6{\small{}2} & 6{\small{}2}\tabularnewline
\hline
A{\small{}ncient } & C{\small{}+} & 6{\small{}8d12+1,564 (2,006)} & 6{\small{}5} & 1{\small{}0} & 5{\small{}7} & 5{\small{}6} & 5{\small{}7} & 5{\small{}6} & +{\small{}68/+111} & +{\small{}87} & +{\small{}59} & +{\small{}36} & +{\small{}59} & 6{\small{}7} & 6{\small{}7}\tabularnewline
\hline
W{\small{}yrm } & C{\small{}+} & 7{\small{}3d12+1,825 (2,299)} & 6{\small{}9} & 1{\small{}0} & 6{\small{}1} & 6{\small{}0} & 6{\small{}1} & 6{\small{}0} & +{\small{}73/+118} & +{\small{}94} & +{\small{}63} & +{\small{}38} & +{\small{}63} & 7{\small{}1} & 7{\small{}1}\tabularnewline
\hline
G{\small{}reat wyrm } & C{\small{}+} & 7{\small{}8d12+2,106 (2,613)} & 7{\small{}3} & 1{\small{}0} & 6{\small{}5} & 6{\small{}4} & 6{\small{}5} & 6{\small{}4} & +{\small{}78/+125} & +{\small{}101} & +{\small{}68} & +{\small{}41} & +{\small{}68} & 7{\small{}6} & 7{\small{}6}\tabularnewline
\hline
\end{tabular}

\vspace{24pt}
\textbf{Prismatic Dragon Abilities by Age }

\begin{tabular}{|>{\raggedright}p{31pt}|>{\raggedright}p{54pt}|>{\raggedright}p{23pt}|>{\raggedright}p{68pt}|>{\raggedright}p{59pt}|>{\raggedright}p{20pt}|>{\raggedright}p{8pt}|}
\hline
A{\small{}\textbf{ge }} & S{\small{}\textbf{peed }} & I{\small{}\textbf{nititative}} & A{\small{}\textbf{C}} & S{\small{}\textbf{pecial 
Abilities }} & C{\small{}\textbf{aster Level}} & S{\small{}\textbf{R}}\tabularnewline
\hline
W{\small{}yrmling } & 6{\small{}0 ft., fly 250 ft. (poor) } & +{\small{}0} & 3{\small{}5 
(-2 size, +5 deflection, +22 natural), touch 13, flat-footed 35} & D{\small{}amage 
reduction 10/magic, deflecting force, }{\small{}\textit{hypnotic pattern, }}{\small{}immune 
to light and blindness } & 5{\small{}th} & 2{\small{}4}\tabularnewline
\hline
V{\small{}ery young } & 6{\small{}0 ft., fly 250 ft. (poor) } & +{\small{}0} & 4{\small{}2 
(-2 size, +7 deflection, +27 natural), touch 15, flat-footed 42} &   & 8{\small{}th} & 3{\small{}0}\tabularnewline
\hline
Y{\small{}oung } & 6{\small{}0 ft., fly 300 ft. (clumsy) } & +{\small{}0} & 4{\small{}7 
(-4 size, +9 deflection, +32 natural), touch 15, flat-footed 47} & D{\small{}amage 
reduction 15/magic } & 1{\small{}1th} & 3{\small{}6}\tabularnewline
\hline
J{\small{}uvenile } & 6{\small{}0 ft., fly 300 ft. (clumsy) } & +{\small{}0} & 5{\small{}4 
(-4 size, +11 deflection, +37 natural), touch 17, flat-footed 54 } & R{\small{}\textit{ainbow 
pattern}} & 1{\small{}3th} & 4{\small{}2}\tabularnewline
\hline
Y{\small{}oung adult } & 6{\small{}0 ft., fly 300 ft. (clumsy) } & +{\small{}0} & 6{\small{}1 
(-4 size, +13 deflection, +42 natural), touch 19, flat-footed 61} & D{\small{}amage 
reduction 15/epic} & 1{\small{}7th} & 4{\small{}8}\tabularnewline
\hline
A{\small{}dult } & 6{\small{}0 ft., fly 300 ft. (clumsy) } & +{\small{}0} & 6{\small{}4 
(-8 size, +15 deflection, +47 natural), touch 17, flat-footed 64 } & S{\small{}\textit{unbeam 
}} & 2{\small{}0th} & 5{\small{}4}\tabularnewline
\hline
M{\small{}ature adult } & 6{\small{}0 ft., fly 300 ft. (clumsy) } & +{\small{}0} & 7{\small{}1 
(-8 size, +17 deflection, +52 natural), touch 19, flat-footed 71 } & D{\small{}amage 
reduction 20/epic} & 2{\small{}3rd} & 6{\small{}0}\tabularnewline
\hline
O{\small{}ld } & 6{\small{}0 ft., fly 300 ft. (clumsy)} & +{\small{}0} & 7{\small{}8 
(-8 size, +19 deflection, +57 natural), touch 21, flat-footed 78 } & S{\small{}\textit{unburst 
}} & 2{\small{}6th} & 6{\small{}6}\tabularnewline
\hline
V{\small{}ery old } & 6{\small{}0 ft., fly 300 ft. (clumsy) } & +{\small{}0} & 8{\small{}5 
(-8 size, +21 deflection, +62 natural), touch 23, flat-footed 85 } & D{\small{}amage 
reduction 25/epic} & 2{\small{}9th} & 6{\small{}8}\tabularnewline
\hline
A{\small{}ncient } & 6{\small{}0 ft., fly 350 ft. (clumsy) } & +{\small{}0} & 9{\small{}2 
(-8 size, +23 deflection, +67 natural), touch 25, flat-footed 92 } & P{\small{}\textit{rismatic 
wall}} & 3{\small{}2nd} & 7{\small{}4}\tabularnewline
\hline
W{\small{}yrm } & 6{\small{}0 ft., fly 350 ft. (clumsy) } & +{\small{}0} & 9{\small{}9 
(-8 size, +25 deflection, +72 natural), touch 27, flat-footed 99 } &  & 3{\small{}5th} & 8{\small{}0}\tabularnewline
\hline
G{\small{}reat wyrm } & 6{\small{}0 ft., fly 350 ft. (clumsy) } & +{\small{}0} & 1{\small{}06 
(-8 size, +27 deflection, +77 natural), touch 29, flat-footed 106 } & P{\small{}\textit{rismatic 
sphere}} & 3{\small{}8th} & 8{\small{}6}\tabularnewline
\hline
\end{tabular}

A prismatic dragon's natural weapons are treated as magic (wyrmling - juvenile) 
or epic (young adult - great wyrm) for the purpose of overcoming damage reduction.

Combat\textbf{ }

\textbf{Breath Weapon (Su):} A prismatic dragon has one type of breath weapon, 
a \textit{prismatic spray }effect. The breath weapon's cone is of the size indicated 
by the dragon's size, not the normal cone for the spell. 

\textbf{Deflecting Force (Su):} A prismatic dragon is protected by a shield of 
shimmering light similar to a \textit{cloak of chaos }that provides a deflection 
bonus to its AC. The deflection bonus is equal to the dragon's Charisma bonus. 

\textbf{Immune to Light and Blindness (Ex): }Prismatic dragons cannot be harmed 
by any light, including Evocation [Light] spells, \textit{searing light, }and the 
various \textit{prismatic (sphere, spray, wall}) spells. In addition, they are 
immune to any spell or effect that causes blindness. 

\textbf{Spell-Like Abilities:} At will---\textit{hypnotic pattern, rainbow pattern; 
}3/day---\textit{prismatic sphere, prismatic wall, sunbeam, sunburst.} The DC is 
Charisma-based.

\vspace{12pt}
\section*{\textbf{Sample Prismatic Dragon, Old }}

Colossal Dragon 

\textbf{Hit Dice:} 58d12+1,102 (1,479 hp)

\textbf{Initiative:} +4 (Improved Initiative) 

\textbf{Speed:} 60 ft., fly 300 ft. (clumsy) 

\textbf{AC:} 78 (-8 size, +19 deflection, +57 natural) 

\textbf{Base Attack/Grapple: }+58/+97

\textbf{Attack: } Bite +74 (4d8+23/19-20) melee 

\textbf{Full Attack:}  Bite +74 melee, 2 claws +69 (4d6+11) melee, 2 wings +68 
(2d8+11) melee, tail slap +69 (4d6+34)\textbf{ }melee

\textbf{Space/Reach:} 30 ft./20 ft

\textbf{Special Attacks:} Crush 4d8+34 (DC 58), tail sweep 2d8+34 (DC 58), breath 
weapon, frightful presence (DC 58), spells (caster level 26th), spell-like abilities 

\textbf{Special Qualities:} Immunities, DR 20/epic, SR 66, blindsight, keen senses, 
deflecting force 

\textbf{Saves:} Fort +50, Ref +31, Will +50 

\textbf{Abilities:} Str 57, Dex 10, Con 49, Int 48, Wis 49, Cha 48 

\textbf{Skills:} Balance +67, Bluff +80, Climb +84, Concentration +80, Diplomacy 
+92, Disable Device +80, Escape Artist +61, Gather Information +86, Heal +80, Intimidate 
+86, Jump +102, Knowledge (all) +80, Listen +80, Move Silently +61, Open Lock +61, 
Perform (any two) +80, Search +80, Sense Motive +80, Spellcraft +86, Spot +80, 
Survival +80, Tumble +67, Use Magic Device +80 

\textbf{Feats:} Blinding Speed (x2), Cleave, Flyby Attack, Great Cleave, Improved 
Critical (bite), Improved Initiative, Improved Sunder, Improved Spell Capacity 
(9th), Improved Spell Capacity (9th), Power Attack, Quicken Spell-Like Ability, 
Snatch, Spell Knowledge (x4), Weapon Focus (bite), Weapon Focus (claw), Weapon 
Focus (tail slap) 

\textbf{Climate/Terrain:} Any land and underground 

\textbf{Organization:} Solitary, pair, or family (1-2 and 2-5 off-spring) 

\textbf{Challenge Rating:} 48 

\textbf{Treasure:} Double standard 

\textbf{Alignment:} Usually neutral 

\textbf{Advancement:} 59-62 HD (Colossal) 

A prismatic dragon's natural weapons are treated as epic for the purpose of overcoming 
damage reduction.

\textbf{Breath Weapon (Su):} Cone of \textit{prismatic spray, }70 ft., effect as 
spell, Ref half (DC 58). 

\textbf{Spell-Like Abilities:} At will---\textit{hypnotic pattern, rainbow pattern; 
}3/day---\textit{sunbeam, sunburst. }Caster level 26th; save DC 29 + spell level. 
The DC is Charisma-based.

\textbf{Immunities (Ex):} Immune to \textit{sleep }and paralysis effects, immune 
to light and blindness effects. 

\textit{\textbf{Sorcerer Spells Known}}\textit{ }(6/11/11/11/10/10/10/10/9/11; 
base DC 29 + spell level): 0---\textit{dancing lights, daze, detect magic, ghost 
sound, light, prestidigitation, ray of frost, read magic, resistance; }1st---\textit{enlarge 
person, mage armor, magic missile, shield, true strike, unseen servant; }2nd---\textit{alter 
self, blur, darkness, detect thoughts, levitate; }3rd---\textit{blink, fireball, 
greater magic weapon, haste, tongues, wind wall; }4th---\textit{confusion, polymorph, 
scrying, stoneskin; }5th---\textit{ baleful polymorph, dismissal, fabricate, passwall, 
wall of force; }6th---\textit{antimagic field, greater dispel magic, true seeing; 
}7th---\textit{forcecage, sequester, spell turning, greater teleport; }8th---\textit{greater 
planar binding, mass charm monster, symbol of death, temporal stasis; }9th---\textit{dominate 
monster, gate, imprisonment, time stop, wish }

\textit{Possessions: Staff of the cosmos, ring of regeneration, }15,563 gp. 

\vspace{12pt}
\section*{{\LARGE{}ELEMENTAL,PRIMAL }}

\textbf{Elemental Traits:} Immune to poison, sleep, paralysis, and stunning. Not 
subject to critical hits.

\begin{tabular}{|>{\raggedright}p{66pt}|>{\raggedright}p{123pt}|>{\raggedright}p{123pt}|}
\hline
   & A\textbf{ir Elemental, Primal } & E\textbf{arth Elemental, Primal }\tabularnewline
\hline
 & Colossal Elemental (Air, Extraplanar)  & Colossal Elemental (Earth, Extraplanar) 
\tabularnewline
\hline
H\textbf{it Dice:}  & 64d8+576 (944 hp)  & 64d8+640 (1,048 hp) \tabularnewline
\hline
I\textbf{nitiative:}  & +24 (+16 Dex, +8 Superior Initiative) - & 1 (Dex) \tabularnewline
\hline
S\textbf{peed:} & Fly 100 ft. (perfect)  & 20 ft., burrow 40 ft. \tabularnewline
\hline
A\textbf{C:}  & 66 (-8 size, +16 Dex, +48 natural)  & 49 (-8 size, -1 Dex, +48 
natural) \tabularnewline
\hline
B\textbf{ase  Attack/Grapple:} & +48/+75 & +48/+80\tabularnewline
\hline
A\textbf{ttack:} & Slam +57 (4d8+16) melee & Slam +57 (4d10+24/19-20) melee\tabularnewline
\hline
F\textbf{ull Attack:} & 2 Slams +57 (4d8+16) melee & 2 slams +57 (4d10+24/19-20) 
melee\tabularnewline
\hline
S\textbf{pace/Reach:}  & 30 ft./30 ft.  & 30 ft./30 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:}  & Air mastery, whirlwind  & Earth mastery, push \tabularnewline
\hline
S\textbf{pecial Qualities:}  & Elemental traits, DR 10/-,SR 42  & Elemental traits, 
DR 15/-, SR 42 \tabularnewline
\hline
S\textbf{aves:}  & Fort +42, Ref +52, Will +28  & Fort +46, Ref +21, Will +28 \tabularnewline
\hline
A\textbf{bilities:}  & Str 32, Dex 43, Con 28, Int 8, Wis 13, Cha 13  & Str 43, 
Dex 8, Con 31, Int 8, Wis 13, Cha 13 \tabularnewline
\hline
S\textbf{kills:}  & Listen +48, Spot +57 & Jump +10, Listen +57, Spot +48\tabularnewline
\hline
F\textbf{eats:}  & Alertness, Combat Reflexes, Dodge, Flyby Attack, Great Fortitude, 
Improved Initiative, Iron Will, Mobility, Lightning Reflexes, Spring Attack, Weapon 
Finesse, Weapon Focus (slam)  & Alertness, Cleave, Combat Reflexes, Great Cleave, 
Improved Critical (slam), Improved Sunder, Iron Will, Lightning Reflexes, Power 
Attack, Weapon Focus (slam)\tabularnewline
\hline
E\textbf{pic Feats:}  & Blinding Speed (x2), Epic Prowess (x2), Epic Toughness 
(x4), Epic Will, Superior Initiative & Devastating Critical (slam), Epic Prowess 
(x3), Epix Toughness (x6), Epic Will, Overwhelming Critical (slam) \tabularnewline
\hline
C\textbf{limate/Terrain:}  & Any land  & Any land \tabularnewline
\hline
O\textbf{rganization:}  & Solitary  & Solitary \tabularnewline
\hline
C\textbf{hallenge Rating:}  & 35  & 35 \tabularnewline
\hline
T\textbf{reasure:}  & None  & None \tabularnewline
\hline
A\textbf{lignment:}  & Usually neutral  & Usually neutral \tabularnewline
\hline
A\textbf{dvancement:}  & 65-288 HD (Colossal)  & 65-288 HD (Colossal) \tabularnewline
\hline
\end{tabular}

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{65pt}|>{\raggedright}p{124pt}|>{\raggedright}p{124pt}|}
\hline
   & F\textbf{ire Elemental, Primal } & W\textbf{ater Elemental, Primal }\tabularnewline
\hline
 & Colossal Elemental (Fire, Extraplanar)  & Colossal Elemental (Water, Extraplanar) 
\tabularnewline
\hline
H\textbf{it Dice:}  & 64d8+576 (944 hp)  & 64d8+640 (1,048 hp) \tabularnewline
\hline
I\textbf{nitiative:}  & +22 (+14 Dex, +8 Superior Initiative)  & +14 (Dex) \tabularnewline
\hline
S\textbf{peed:}  & 50 ft.  & 20 ft., swim 90 ft. \tabularnewline
\hline
A\textbf{C:}  & 64 (-8 size, +14 Dex, +48 natural)  & 64 (-8 size, +14 Dex, +48 
natural) \tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +48/+75 & +48/+80\tabularnewline
\hline
A\textbf{ttack:} & Slam +55 (4d8+16 plus 4d8 fire) melee & Slam +57 (4d10+24/19-20)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 slams +55 (4d8+16 plus 4d8 fire) melee & 2 slam +57 (4d10+24/19-20)\tabularnewline
\hline
S\textbf{pace/Reach:}  & 30 ft./30 ft.  & 30 ft./30 ft. \tabularnewline
\hline
S\textbf{pecial Attacks:}  & Burn  & Water mastery, drench, vortex \tabularnewline
\hline
S\textbf{pecial Qualities:}  & Elemental traits, DR 10/-, SR 42, fire immunity, 
cold vulnerability & Elemental traits, DR 15/-, SR 42 \tabularnewline
\hline
S\textbf{aves:} & Fort +32, Ref +50, Will +28  & Fort +46, Ref +21, Will +28 \tabularnewline
\hline
A\textbf{bilities:}  & Str 32, Dex 39, Con 28, Int 8, Wis 13, Cha 13  & Str 42, 
Dex 38, Con 31, Int 8, Wis 13, Cha 13 \tabularnewline
\hline
S\textbf{kills:}  & Jump +19, Listen +48, Spot +57 & Jump +7, Listen +57, Spot 
+48, Swim +19\tabularnewline
\hline
F\textbf{eats:} & Alertness, Combat Reflexes, Dodge, Great Fortitude, Improved 
Initiative, Iron Will, Lightning Reflexes, Mobility, Power Attack, Spring Attack, 
Weapon Finesse, Weapon Focus (slam) & Alertness, Cleave, Combat Reflexes, Dodge, 
Great Cleave, Improved Critical (slam), Improved Sunder, Iron Will, Lightning Reflexes, 
Power Attack, Weapon Focus (slam)\tabularnewline
\hline
E\textbf{pic Feats:}  & Blinding Speed, Dire Charge, Epic Prowess, (x2), Epic Toughness 
(x4), Epic Will, Superior Initiative  & Devastating Critical (slam), Epic Prowess 
(x2), Epic Toughness (x6), Epic Will, Overwhelming Critical (slam) \tabularnewline
\hline
C\textbf{limate/Terrain:}  & Any land  & Any land \tabularnewline
\hline
O\textbf{rganization:}  & Solitary  & Solitary \tabularnewline
\hline
C\textbf{hallenge Rating:}  & 35  & 35 \tabularnewline
\hline
T\textbf{reasure:}  & None  & None \tabularnewline
\hline
A\textbf{lignment:}  & Usually neutral  & Usually neutral \tabularnewline
\hline
A\textbf{dvancement:}  & 65-288 HD (Colossal)  & 65-288 HD (Colossal) \tabularnewline
\hline
\end{tabular} 

\parindent=3pt
PRIMAL AIR ELEMENTAL 

\parindent=0pt
Primal air elementals speak Auran. 

Combat\textbf{ }

\textbf{Air Mastery (Ex):} Airborne creatures take a -1 penalty on attack and damage 
rolls against an air elemental. 

\textbf{Whirlwind (Su):} The elemental can transform itself into a whirlwind once 
every 10 minutes and remain in that form for up to 1 round for every 2 HD it has. 
In this form, the elemental can move through the air or along a surface at its 
fly speed.  A primal air whirlwind is 10 feet wide at the base, 40 feet wide at 
the top, and up to 80 feet tall. The elemental controls its exact height, but it 
must be at least 10 feet. Gargantuan or smaller creatures might take damage when 
caught in the whirlwind and may be lifted into the air. An affected creature must 
succeed at a Reflex save (DC 53) when it comes into contact with the whirlwind 
or take 4d8 points of damage. It must also succeed at a second Reflex save (DC 
53) or be picked up bodily and held suspended in the powerful winds, automatically 
taking 4d8 points of damage each round. A creature that can fly is allowed a Reflex 
save (DC 53) each round to escape the whirlwind. The creature still takes damage 
but can leave if the save is successful. The elemental can eject any carried creatures 
whenever it wishes, depositing them wherever the whirlwind happens to be. A summoned 
elemental always ejects trapped creatures before returning to its home plane. If 
the whirlwind's base touches the ground, it creates a swirling cloud of debris. 
This cloud is centered on the elemental and has a diameter equal to half the whirlwind's 
height. The cloud obscures all vision, including darkvision, beyond 5 feet. Creatures 
5 feet away have one-half concealment, while those farther away have total concealment. 
Those caught in the cloud must succeed at a Concentration check (DC 53) to cast 
a spell. The DC is Strength-based.

\section*{PRIMAL EARTH ELEMENTAL }

Earth elementals speak Terran. 

Combat\textbf{ }

\textbf{Earth Mastery (Ex): }An earth elemental gains a +1 attack and damage bonus 
if both it and its foe touch the ground. If an opponent is airborne or water-borne, 
the elemental takes a -4 penalty on attack and damage rolls. (These modifiers are 
not included in the statistics block.) 

\textbf{Push (Ex):} An earth elemental can start a bull rush maneuver without provoking 
an attack of opportunity. The combat modifiers given for earth mastery, above, 
also apply to the elemental's opposed Strength checks. 

\section*{PRIMAL FIRE ELEMENTAL }

A primal fire elemental cannot enter water or any other nonflammable liquid. Unlike 
lesser fire elementals, a primal fire elemental may attempt to evaporate a lesser 
body of water if blocked. Fire elementals speak Ignan. 

Combat\textbf{ }

\textbf{Burn (Ex):} Those hit by a primal fire elemental's slam attack must succeed 
at a Reflex save (DC 51) or catch fire. The flame burns for 1d4 rounds. A burning 
creature can take a move-equivalent action to put out the flame. Creatures hitting 
a fire elemental with natural weapons or unarmed attacks take fire damage as though 
hit by the elemental's attack, and also catch fire unless they succeed at a Reflex 
save. The DC is Constitution-based.

\section*{PRIMAL WATER ELEMENTAL }

A primal water elemental can't venture more than 360 feet from the body of water 
from which it emerged. Water elementals speak Aquan. 

Combat\textbf{ }

\textbf{Water Mastery (Ex):} A water elemental gains a +1 bonus on attack and damage 
rolls if both it and its opponent touch water. If the opponent or elemental is 
land-bound, the elemental takes a -4 penalty on attack and damage rolls. (These 
modifiers are not included in the statistics block.) A primal water elemental can 
be a serious threat to a ship that crosses its path. It can easily overturn even 
large craft (ships of up to 5 feet in length per Hit Dice of the elemental) and 
stop massive vessels (ships of less than 10 feet in length per HD). Even ocean-going 
behemoths (ships of less than 20 feet in length per HD) can be slowed to half speed. 

\textbf{Drench (Ex):} The elemental's touch puts out torches, campfires, exposed 
lanterns, and other open flames of non-magical origin if these are Gargantuan or 
smaller. The creature can dispel magical fire it touches as \textit{dispel magic 
}cast by a caster whose level equals the elemental's HD total (64). 

\textbf{Vortex (Su):} The elemental can transform itself into a whirlpool once 
every 10 minutes, provided it is underwater, and remain in that form for up to 
1 round for every 2 HD it has (48 rounds). In vortex form, the elemental can move 
through the water or along the bottom at its swim speed. The vortex is 5 feet wide 
at the base, up to 50 feet wide at the top, and up to 80 feet tall. The elemental 
controls its exact height, but it must be at least 10 feet. Gargantuan and smaller 
creatures might take damage when caught in the vortex and may be swept up by it. 
An affected creature must succeed at a Reflex save (DC 58) when it comes into contact 
with the vortex or take 4d10 points of damage. It must also succeed at a second 
Reflex save (DC 68) or be picked up bodily and held suspended in the powerful currents, 
automatically taking 4d10 points of damage each round. A creature that can swim 
is allowed a Reflex save (DC 58) each round to escape the vortex. The creature 
still takes damage, but can leave if the save is successful. The elemental can 
eject any carried creatures whenever it wishes, depositing them wherever the vortex 
hap-pens to be. A summoned elemental always ejects trapped creatures before returning 
to its home plane. If the vortex's base touches the bottom, it creates a swirling 
cloud of debris. This cloud is centered on the elemental and has a diameter equal 
to half the vortex's height. The cloud obscures all vision, including dark-vision, 
beyond 5 feet. Creatures 5 feet away have one-half concealment, while those farther 
away have total concealment. Those caught in the cloud must succeed at a Concentration 
check (DC 58) to cast a spell.  The DC is Strength-based.

\newpage

\end{document}
