%&pdfLaTeX
% !TEX encoding = UTF-8 Unicode
\documentclass{article}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\setmainfont[Mapping=tex-text]{STIXGeneral}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi
\usepackage{textcomp}

\usepackage{array}
\usepackage{amssymb}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}

This material is Open Game Content, and is licensed for public use under the terms 
of the Open Game License v1.0a.

{\LARGE{}MONSTERS (B-C)}

\vspace{12pt}
{\LARGE{}BARGHEST}

\begin{tabular}{|>{\raggedright}p{49pt}|>{\raggedright}p{124pt}|>{\raggedright}p{139pt}|}
\hline
  & B\textbf{arghest} & G\textbf{reater Barghest}\tabularnewline
\hline
  & Medium Outsider (Evil, Extraplanar, Lawful, Shapechanger) & Large Outsider 
(Evil, Extraplanar, Lawful, Shapechanger) \tabularnewline
\hline
H\textbf{it Dice:} & 6d8+6 (33 hp) & 9d8+27 (67 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +6 & +6\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares) & 40 ft. (8 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 18 (+2 Dex, +6 natural), touch 12, flat-footed 16 & 20 
(-1 size, +2 Dex, +9 natural), touch 11, flat-footed 18\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +6/+9  & +9/+18\tabularnewline
\hline
A\textbf{ttack:} & Bite +9 melee (1d6+3)) & Bite +13 melee (1d8+5\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +9 melee (1d6+3) and 2 claws +4 melee (1d4+1) & Bite 
+13 melee (1d8+5) and 2 claws +8 melee (1d6+2)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Spell-like abilities, feed & Spell-like abilities, 
feed\tabularnewline
\hline
S\textbf{pecial Qualities:} & Change shape, damage reduction 5/magic, darkvision 
60 ft., scent & Change shape, damage reduction 10/magic, darkvision 60 ft., scent\tabularnewline
\hline
S\textbf{aves:} & Fort +6, Ref +7, Will +7 & Fort +9, Ref +8, Will +10\tabularnewline
\hline
A\textbf{bilities:} & Str 17, Dex 15, Con 13, Int 14, Wis 14, Cha 14 & Str 20, 
Dex 15, Con 16, Int 18, Wis 18, Cha 18\tabularnewline
\hline
S\textbf{kills:} & Bluff +11, Diplomacy +6, Disguise +2 (+4 acting), Hide +11*, 
Intimidate +13, Jump +12, Listen +11, Move Silently +10, Search +11, Sense Motive 
+11, Spot +11, Survival +11 (+13 following tracks) & Bluff +16, Climb +17, Concentration 
+15, Diplomacy +8, Disguise +4 (+6 acting), Hide +10*, Intimidate +18, Jump +21, 
Listen +16, Move Silently +14, Sense Motive +16, Spot +16 Survival +16 (+18 following 
tracks), Tumble +16\tabularnewline
\hline
F\textbf{eats:} & Combat Reflexes, Improved Initiative, Track & Combat Casting, 
Combat Reflexes, Improved Initiative, Track\tabularnewline
\hline
E\textbf{nvironment} & An evil-aligned plane & An evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary or pack (3-6) & Solitary or pack (3-6)\tabularnewline
\hline
C{\small{}\textbf{hallenge Rating:}} & 4 & 5\tabularnewline
\hline
T\textbf{reasure:} & Double standard & Double standard\tabularnewline
\hline
A\textbf{lignment:} & Always lawful evil & Always lawful evil\tabularnewline
\hline
A\textbf{dvancement:} & Special (see below) & Special (see below)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & --- & \tabularnewline
\hline
\end{tabular}

A barghest is a lupine fiend that can take the shape of a wolf or a goblin. In 
its natural form, it resembles a goblin-wolf hybrid with terrible jaws and sharp 
claws. As whelps, barghests are nearly indistinguishable from wolves, except for 
their size and claws. As they grow larger and stronger, their skin darkens to bluishred 
and eventually becomes blue altogether.

A full-grown barghest is about 6 feet long and weighs 180 pounds. A barghest's 
eyes glow orange when the creature becomes excited.

Barghests speak Goblin, Worg, and Infernal.

COMBAT

Barghests can claw and bite, no matter what their form, and usually disdain weapons. 
Though they love killing, they have little stomach for direct combat and attack 
from ambush whenever possible. Barghests start a combat by using \textit{crushing 
despair }and \textit{charm monster }to keep opponents off balance. They try to 
stay away from the enemy's main strength.

A barghest's natural weapons, as well as any weapons it wields, are treated as 
evil-aligned and lawful-aligned for the purpose of overcoming damage reduction. 
Its natural weapons are treated as magic weapons for the purpose of overcoming 
damage reduction.

\textbf{Spell-Like Abilities:} At will---\textit{blink, levitate, misdirection 
}(DC 14), \textit{rage }(DC 15); 1/day---\textit{charm monster }(DC 16), \textit{crushing 
despair }(DC 16), \textit{dimension door. }Caster level equals the barghest's HD.

The save DCs are Charisma-based.

\textbf{Feed (Su):} When a barghest slays a humanoid opponent, it can feed on the 
corpse, devouring both flesh and life force, as a full-round action. Feeding destroys 
the victim's body and prevents any form of raising or resurrection that requires 
part of the corpse. There is a 50\% chance that a \textit{wish, miracle, }or \textit{true 
resurrection }spell can restore a devoured victim to life. Check once for each 
destroyed creature. If the check fails, the creature cannot be brought back to 
life by mortal magic.

A barghest advances in Hit Dice by consuming corpses in this fashion. For every 
three suitable corpses a barghest devours, it gains 1 Hit Die, and its Strength, 
Constitution, and natural armor increase by +1. Its attack bonus and saves improve 
as normal for an outsider of its Hit Dice, and it gains skill points, feats, and 
ability score improvements normally.

The barghest only advances by consuming the corpses of creatures whose Hit Dice 
or levels are equal to or greater than its own current total. A barghest that reaches 
9 Hit Dice through feeding immediately becomes a greater barghest upon completion 
of the act.

\textbf{Change Shape (Su):} A barghest can assume the shape of a goblin or a wolf 
as a standard action. In goblin form, a barghest cannot use its natural weapons 
but can wield weapons and wear armor. In wolf form, a barghest loses its claw attacks 
but retains its bite attack.

\textbf{Pass Without Trace (Ex):} A barghest in wolf form can use pass without 
trace (as the spell) as a free action.

\textbf{Skills:} *A barghest in wolf form gains a +4 circumstance bonus on Hide 
checks.

\vspace{12pt}
GREATER BARGHEST

A barghest that reaches 9 Hit Dice through feeding becomes a greater barghest. 
These creatures can change shape into a goblinlike creature of Large size (about 
8 feet tall and 400 pounds) or a dire wolf. In goblin form, a greater barghest 
cannot use its natural weapons but can wield weapons and wear armor. In dire wolf 
form, a greater barghest loses its claw attacks but retains its bite attack.

A greater barghest can reach a maximum of 18 Hit Dice through feeding.

\textbf{Spell-Like Abilities:} In addition to the spelllike abilities all barghests 
possess, a greater barghest gains the following abilities. At will---\textit{invisibility 
sphere; }1/day---\textit{mass bull's strength}, \textit{mass enlarge. }Caster level 
equals the greater barghest's HD.

COMBAT

Occasionally, a greater barghest uses a magic two-handed weapon in combat instead 
of its claws, giving it multiple attacks (attack bonus +13/+8). It can also make 
one bite attack (attack bonus +8) each round. The save DC against a greater barghest's 
spell-like abilities is 14 + spell level.

\vspace{12pt}
{\LARGE{}BASILISK}

\begin{tabular}{|>{\raggedright}p{50pt}|>{\raggedright}p{134pt}|>{\raggedright}p{129pt}|}
\hline
  & B\textbf{asilisk} & A\textbf{byssal Greater Basilisk}\tabularnewline
\hline
  & Medium Magical Beast & Large Outsider (Augmented Magical Beast, Extraplanar)\tabularnewline
\hline
H\textbf{it Dice:} & 6d10+12 (45 hp) & 18d10+90 (189 hp)\tabularnewline
\hline
I\textbf{nitiative:}- & 1- & 1\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares) & 20 ft. (4 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 16 (-1 Dex, +7 natural), touch 9, flat-footed 16 & 17 (-1 
Dex, -1 size, +9 natural) touch 8, flat-footed 17\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +6/+8 & +18/+29\tabularnewline
\hline
A\textbf{ttack:} & Bite +8 melee (1d8+3) & Bite +25 melee (2d8+10)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +8 melee (1d8+3) & Bite +25 melee (2d8+10)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft. & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Petrifying gaze & Petrifying gaze, smite good \tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., low-light vision & Resistance 
to cold 10 and fire 10, damage reduction 10/magic, darkvision 60 ft., low-light 
vision, spell resistance 23\tabularnewline
\hline
S\textbf{aves:} & Fort +9, Ref +4, Will +3 & Fort +18, Ref +12, Will +8\tabularnewline
\hline
A\textbf{bilities:} & Str 15, Dex 8, Con 15, Int 2, Wis 12, Cha 11 & Str 24, Dex 
8, Con 21, Int 3, Wis 10, Cha 15\tabularnewline
\hline
S\textbf{kills:} & Hide +0*, Listen +7, Spot +7 & Hide +0*, Listen +10, Spot +10\tabularnewline
\hline
F\textbf{eats:} & Alertness, Blind-Fight, Great Fortitude & Alertness, Blind-Fight, 
Great Fortitude, Iron Will, Improved Natural Attack (bite), Lightning Reflexes, 
Weapon Focus (bite)\tabularnewline
\hline
E\textbf{nvironment:} & Warm deserts & A chaotic evil plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary or colony (3-6) & Solitary or colony (3-6)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 5 & 12\tabularnewline
\hline
T\textbf{reasure:} & None & Standard\tabularnewline
\hline
A\textbf{lignment:} & Always neutral & Always chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 7-10 HD (Medium); 11-18 HD (Large)--- & \tabularnewline
\hline
L\textbf{evel Adjustment:}--- & --- & \tabularnewline
\hline
\end{tabular}

A basilisk is a reptilian monster that petrifies living creatures with a mere gaze. 
A basilisk usually has a dull brown body with a yellowish underbelly. Some specimens 
sport a short, curved horn atop the nose. An adult basilisk's body grows to about 
6 feet long, not including its tail, which can reach an additional length of 5 
to 7 feet. The creature weighs about 300 pounds.

COMBAT

A basilisk relies on its gaze attack, biting only when opponents come within reach. 
Though it has eight legs, its slow metabolism renders it relatively sluggish, so 
it does not expend energy unnecessarily. Intruders who flee a basilisk rather than 
fight can expect, at best, a halfhearted pursuit. These creatures tend to spend 
most of their time lying in wait for prey, which includes small mammals, birds, 
reptiles, and similar creatures.

\textbf{Petrifying Gaze (Su):} Turn to stone permanently, range 30 feet; Fortitude 
DC 13 negates. The save DC is Charisma-based.

\textbf{Skills:} *The basilisk's dull coloration and its ability to remain motionless 
for long periods of time grant it a +4 racial bonus on Hide checks in natural settings.

\vspace{12pt}
ABYSSAL GREATER BASILISK

COMBAT

The save DC for the abyssal greater basilisk's petrifying gaze (DC 21) is adjusted 
for its greater Hit Dice and higher Charisma score.

An abyssal greater basilisk's natural weapons are treated as magic weapons for 
the purpose of overcoming damage reduction.

\textbf{Smite Good (Su): }Once per day an abyssal greater basilisk can make a normal 
melee attack to deal an extra 18 points of damage against a good foe.

\vspace{12pt}
{\LARGE{}BEHIR}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{231pt}|}
\hline
  & Huge Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 9d10+45 (94 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares), climb 15 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 20 (-2 size, +1 Dex, +11 natural), touch 9, flat-footed 
19\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +9/+25\tabularnewline
\hline
A\textbf{ttack:} & Bite +15 melee (2d4+12)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +15 melee (2d4+12)\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Breath weapon, constrict 2d8+8, improved grab, rake 
1d4+4, swallow whole\tabularnewline
\hline
S\textbf{pecial Qualities:} & Can't be tripped, darkvision 60 ft., immunity to 
electricity, low-light vision, scent\tabularnewline
\hline
S\textbf{aves:} & Fort +11, Ref +7, Will +5\tabularnewline
\hline
A\textbf{bilities:} & Str 26, Dex 13, Con 21, Int 7, Wis 14, Cha 12\tabularnewline
\hline
S\textbf{kills:} & Climb +16, Hide +5, Listen +4, Spot +4, Survival +2\tabularnewline
\hline
F\textbf{eats:} & Alertness, Cleave, Power Attack, Track\tabularnewline
\hline
E\textbf{nvironment:} & Warm hills\tabularnewline
\hline
O\textbf{rganization:} & Solitary or pair\tabularnewline
\hline
C\textbf{hallenge Rating:} & 8\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Often neutral\tabularnewline
\hline
A\textbf{dvancement:} & 10-13 HD (Huge); 14-27 HD (Gargantuan)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

The behir is a serpentine monster that can slither like a snake or use its dozen 
legs to move with considerable speed. A behir is about 40 feet long and weighs 
about 4,000 pounds. It can fold its limbs close to its long, narrow body and slither 
in snake fashion if it desires. The coloration of behirs ranges from ultramarine 
to deep blue with bands of gray-brown. The belly is pale blue. The two large horns 
curving back over the head look dangerous but are actually used for preening the 
creature's scales, not for fighting.

Behirs speak Common.

COMBAT

A behir usually bites and grabs its prey first, then either swallows or constricts 
the opponent. It can employ its claws only against foes caught in its coils. If 
beset by a large number of foes, it uses its breath weapon.

\textbf{Breath Weapon (Su):} 20-foot line, once every 10 rounds, damage 7d6 electricity, 
Reflex DC 19 half. The save DC is Constitution-based.

\textbf{Constrict (Ex):} A behir deals 2d8+8 points of damage with a successful 
grapple check. It can make six rake attacks against a grappled foe as well.

\textbf{Improved Grab (Ex):} To use this ability, a behir must hit a creature of 
any size with its bite attack. It can then attempt to start a grapple as a free 
action without provoking an attack of opportunity.

If it wins the grapple check, it it establishes a hold and can attempt to constrict 
the opponent or swallow the opponent in the following round.

\textbf{Rake (Ex):} Six claws, attack bonus +15 melee, damage 1d4+4.

\textbf{Swallow Whole (Ex): }A behir can try to swallow a grabbed Medium or smaller 
opponent by making a successful grapple check.

A behir that swallows an opponent can use its Cleave feat to bite and grab another 
opponent.

A swallowed creature takes 2d8+8 points of bludgeoning damage and 8 points of acid 
damage per round from the behir's gizzard. A swallowed creature can cut its way 
out by using a light slashing or piercing weapon to deal 25 points of damage to 
the gizzard (AC 15). Once the creature exits, muscular action closes the hole; 
another swallowed opponent must cut its own way out.

A behir's gizzard can hold 2 Medium, 8 Small, 32 Tiny, or 128 Diminutive or smaller 
opponents.

\textbf{Skills: }Behirs have a +8 racial bonus on Climb checks and can always choose 
to take 10 on Climb checks, even if rushed or threatened.

\vspace{12pt}
\subsection*{{\LARGE{}BELKER}}

\subsection*{\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{231pt}|}
\hline
 } & Large Elemental (Air, Extraplanar)\tabularnewline
\hline
H\textbf{it Dice:} & 7d8+7 (38 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +5\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares), fly 50 ft. (perfect)\tabularnewline
\hline
A\textbf{rmor Class:} & 22 (-1 size, +5 Dex, +8 natural), touch 14, flat-footed 
17\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +5/+11\tabularnewline
\hline
A\textbf{ttack:} & Wing +9 melee (1d6+2)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 wings +9 melee (1d6+2) and bite +4 melee (1d4+1) and 
2 claws +4 melee (1d3+1)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Smoke claws\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., elemental traits, smoke form\tabularnewline
\hline
S\textbf{aves:} & Fort +3, Ref +10, Will +2\tabularnewline
\hline
A\textbf{bilities:} & Str 14, Dex 21, Con 13, Int 6, Wis 11, Cha 11\tabularnewline
\hline
S\textbf{kills:} & Listen +7, Move Silently +9, Spot +7\tabularnewline
\hline
F\textbf{eats:} & Alertness, Multiattack, Weapon Finesse\tabularnewline
\hline
E\textbf{nvironment:} & Elemental Plane of Air\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or clutch (3-4)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 6\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Usually neutral evil\tabularnewline
\hline
A\textbf{dvancement:} & 8-10 HD (Large); 11-21 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

Belkers are creatures from the Plane of Air. They are composed primarily of smoke. 
Although undeniably evil, they are very reclusive and usually have no interest 
in the affairs of others. A belker's winged shape makes it look distinctly demonic. 
It is about 7 feet long and weighs about 8 pounds.

Belkers speak Auran.

COMBAT

In most cases, a belker fights with its nasty claws and painful bite.

\textbf{Smoke Claws (Ex):} A belker in smoke form (see below) can engulf opponents 
by moving on top of them. It fills the air around one Medium or smaller opponent 
without provoking an attack of opportunity. The target must succeed on a DC 14 
Fortitude save or inhale part of the creature. The save DC is Constitution-based. 
Smoke inside the victim solidifies into a claw and begins to rip at the surrounding 
organs, dealing 3d4 points of damage per round. An affected creature can attempt 
another Fortitude save each subsequent round to cough out the semivaporous menace.

\textbf{Smoke Form (Su):} Most of the time a belker is more or less solid, but 
at will it can assume smoke form. It can switch forms once per round as a free 
action and can spend up to 20 rounds per day in smoke form. A belker in smoke form 
can fly at a speed of 50 feet (perfect). The ability is otherwise similar to a 
\textit{gaseous form }spell (caster level 7th).

\textbf{Skills:} Belkers have a +4 racial bonus on Move Silently checks.

\vspace{12pt}
{\LARGE{}BLINK DOG}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{204pt}|}
\hline
  & Medium Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 4d10 (22 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +3\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 16 (+3 Dex, +3 natural), touch 13, flat-footed 13\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +4/+4\tabularnewline
\hline
A\textbf{ttack:} & Bite +4 melee (1d6)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +4 melee (1d6)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:}--- & \tabularnewline
\hline
S\textbf{pecial Qualities:} & Blink, darkvision 60 ft., dimension door, low-light 
vision, scent\tabularnewline
\hline
S\textbf{aves:} & Fort +4, Ref +7, Will +4\tabularnewline
\hline
A\textbf{bilities:} & Str 10, Dex 17, Con 10, Int 10, Wis 13, Cha 11\tabularnewline
\hline
S\textbf{kills:} & Hide +3, Listen +5, Sense Motive +3, Spot +5, Survival +4\tabularnewline
\hline
F\textbf{eats:} & Iron Will, Run, Track\textsuperscript{B}\tabularnewline
\hline
E\textbf{nvironment:} & Temperate plains\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or pack (7-16)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Usually lawful good\tabularnewline
\hline
A\textbf{dvancement:} & 5-7 HD (Medium); 8-12 HD (Large)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +2 (cohort)\tabularnewline
\hline
\end{tabular}

The blink dog is an intelligent canine that has a limited teleportation ability.

Blink dogs have their own language, a mixture of barks, yaps, whines, and growls 
that can transmit complex information.

COMBAT

Blink dogs hunt in packs, teleporting in a seemingly random fashion until they 
surround their prey, allowing some of them to take advantage of flanking.

\textbf{Blink (Su): }A blink dog can use \textit{blink }as the spell (caster level 
8th), and can evoke or end the effect as a free action.

\textbf{Dimension Door (Su):} A blink dog can teleport, as \textit{dimension door 
}(caster level 8th), once per round as a free action. The ability affects only 
the blink dog, which never appears within a solid object and can act immediately 
after teleporting.

\vspace{12pt}
\subsection*{{\LARGE{}BODAK}}

\subsection*{\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{204pt}|}
\hline
 } & Medium Undead (Extraplanar)\tabularnewline
\hline
H\textbf{it Dice:} & 9d12 (58 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +6\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 20 (+2 Dex, +8 natural), touch 12, flat-footed 18\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +4/+5\tabularnewline
\hline
A\textbf{ttack:} & Slam +6 melee (1d8+1)\tabularnewline
\hline
F\textbf{ull Attack:} & Slam +6 melee (1d8+1)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Death gaze\tabularnewline
\hline
S\textbf{pecial Qualities:} & Damage reduction 10/cold iron, darkvision 60 ft., 
immunity to electricity, resistance to acid 10 and fire 10, undead traits, vulnerability 
to sunlight\tabularnewline
\hline
S\textbf{aves:} & Fort +3, Ref +5, Will +7\tabularnewline
\hline
A\textbf{bilities:} & Str 13, Dex 15, Con ---, Int 6, Wis 12, Cha 12\tabularnewline
\hline
S\textbf{kills:} & Listen +11, Move Silently +10, Spot +11\tabularnewline
\hline
F\textbf{eats:} & Alertness, Dodge, Improved Initiative, Weapon Focus (slam)\tabularnewline
\hline
E\textbf{nvironment:} & A chaotic evil-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary or gang (2-4)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 8\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 10-13 HD (Medium); 14-27 HD (Large)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

Bodaks are the undead remnants of humanoids who have been destroyed by the touch 
of absolute evil.

A bodak retains fleeting memories of its past life and can speak Common (or some 
other humanoid language).

COMBAT

\textbf{Death Gaze (Su):} Death, range 30 feet, Fortitude DC 15 negates. Humanoids 
who die from this attack are transformed into bodaks 24 hours later. The save DC 
is Charisma-based.

\textbf{Vulnerability to Sunlight (Ex):} Bodaks loathe sunlight, for its merest 
touch burns their impure flesh. Each round of exposure to the direct rays of the 
sun deals 1 point of damage to the creature.

\vspace{12pt}
{\LARGE{}BRALANI}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{226pt}|}
\hline
  & Medium Outsider (Chaotic, Extraplanar, Good)\tabularnewline
\hline
H\textbf{it Dice:} & 6d8+18 (45 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +8\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares), fly 100 ft. (perfect)\tabularnewline
\hline
A\textbf{rmor Class:} & 20 (+4 Dex, +6 natural), touch 14, flat-footed 16\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +6/+10\tabularnewline
\hline
A\textbf{ttack:} & +\textit{1 holy scimitar }+11 melee (1d6+4/18-20) or \textit{+1 
holy composite longbow }(+4 Str bonus) +11 ranged (1d8+5/x3) or slam +10 melee 
(1d6+4)\tabularnewline
\hline
F\textbf{ull Attack:} & +\textit{1 holy scimitar }+11/+6 melee (1d6+4/18-20) or 
\textit{+1 holy composite longbow }(+4 Str bonus) +11/+6 ranged (1d8+5/x3) or slam 
+10 melee (1d6+4)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Spell-like abilities, whirlwind blast\tabularnewline
\hline
S\textbf{pecial Qualities:} & Alternate form, damage reduction 10/cold iron or 
evil, darkvision 60 ft., immunity to electricity and petrification, low-light vision, 
resistance to cold 10 and fire 10, spell resistance 17, tongues\tabularnewline
\hline
S\textbf{aves:} & Fort +8, Ref +9, Will +7\tabularnewline
\hline
A\textbf{bilities:} & Str 18, Dex 18, Con 17, Int 13, Wis 14, Cha 14\tabularnewline
\hline
S\textbf{kills:} & Concentration +12, Diplomacy +4, Escape Artist +13, Handle Animal 
+11, Hide +13, Jump +10, Listen +13, Move Silently +13, Ride +6, Sense Motive +11, 
Spot +13, Tumble +13, Use Rope +4 (+6 with bindings)\tabularnewline
\hline
F\textbf{eats:} & Alertness, Blind-Fight, Improved Initiative\tabularnewline
\hline
E\textbf{nvironment:} & A chaotic good-aligned plane\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or squad (3-5)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 6\tabularnewline
\hline
T\textbf{reasure:} & No coins; double goods; standard items\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic good\tabularnewline
\hline
A\textbf{dvancement:} & 7-12 HD (Medium); 13-18 HD (Large)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +5\tabularnewline
\hline
\end{tabular}

In addition to their natural form, bralanis can assume the shape of a whirlwind 
or zephyr of dust, snow, or sand.

Bralanis speak Celestial, Infernal, and Draconic, but can communicate with almost 
any creature, thanks to their \textit{tongues }ability.

Combat

Bralanis prefer the scimitar and bow, the weapons of the desert nomads they most 
closely resemble.

A bralani's natural weapons, as well as any weapons it wields, are treated as chaotic-aligned 
and good-aligned for the purpose of overcoming damage reduction.

\textbf{Spell-Like Abilities:} At will--- \textit{blur, charm person }(DC 13), 
\textit{gust of wind }(DC 14), \textit{mirror image, wind wall; }2/day---\textit{lightning 
bolt }(DC 15), \textit{cure serious wounds }(DC 15). Caster level 6th. The save 
DCs are Charisma-based.

\textbf{Whirlwind Blast (Su): }When in whirlwind form, a bralani can attack with 
a scouring blast of wind, dealing 3d6 points of damage in a 20-foot line (Reflex 
DC 16 half ). The save DC is Constitution-based.

\textbf{Alternate Form (Su): }A bralani can shift between its humanoid and whirlwind 
forms as a standard action. In humanoid form, it cannot fly or use its whirlwind 
blast, but it can use its spell-like abilities and its weapons. In whirlwind form, 
it can fly, make slam attacks and whirlwind blast attacks, and use spell-like abilities. 

A bralani remains in one form until it chooses to assume a new one. A change in 
form cannot be dispelled, nor does the bralani revert to any particular form when 
killed. A \textit{true seeing }spell, however, reveals both forms simultaneously.

\textbf{Tongues (Su):} Bralanis can speak with any creature that has a language, 
as though using a \textit{tongues }spell (caster level 14th). This ability is always 
active.

\vspace{12pt}
{\LARGE{}BUGBEAR}

\begin{tabular}{|>{\raggedright}p{85pt}|>{\raggedright}p{240pt}|}
\hline
  & Medium Humanoid (Goblinoid)\tabularnewline
\hline
H\textbf{it Dice:} & 3d8+3 (16 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 17 (+1 Dex, +3 natural, +2 leather armor, +1 light wooden 
shield), touch 11, flat-footed 16\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +2/+4\tabularnewline
\hline
A\textbf{ttack:} & Morningstar +5 melee (1d8+2) or javelin +3 ranged (1d6+2)\tabularnewline
\hline
F\textbf{ull Attack:} & Morningstar +5 melee (1d8+2) or javelin +3 ranged (1d6+2)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:}--- & \tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., scent\tabularnewline
\hline
S\textbf{aves:} & Fort +2, Ref +4, Will +1\tabularnewline
\hline
A\textbf{bilities:} & Str 15, Dex 12, Con 13, Int 10, Wis 10, Cha 9\tabularnewline
\hline
S\textbf{kills:} & Climb +3, Hide +4, Listen +4, Move Silently +6, Spot +4\tabularnewline
\hline
F\textbf{eats:} & Alertness, Weapon Focus (morningstar)\tabularnewline
\hline
E\textbf{nvironment:} & Temperate mountains\tabularnewline
\hline
O\textbf{rganization:} & Solitary, gang (2-4), or band (11-20 plus 150\% noncombatants 
plus 2 2nd-level sergeants and 1 leader of 2nd-5th level)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & By character class\tabularnewline
\hline
L\textbf{evel Adjustment:} & +1\tabularnewline
\hline
\end{tabular}

Bugbears speak Goblin and Common.

COMBAT

Bugbears prefer to ambush opponents whenever possible. When hunting, they normally 
send scouts ahead of the main group that, if they spy prey, return to report and 
bring up reinforcements.

Bugbear attacks are coordinated, and their tactics are sound if not brilliant.

\textbf{Skills: }Bugbears have a +4 racial bonus on Move Silently checks.

BUGBEARS AS CHARACTERS

Bugbear characters possess the following racial traits.--- 

\parindent=3pt
+4 Strength, +2 Dexterity, +2 Constitution, -2 Charisma.---

\parindent=0pt
Medium size.---

A bugbear's base land speed is 30 feet.---

Darkvision out to 60 feet.---

Racial Hit Dice: A bugbear begins with three levels of humanoid, which provide 
3d8 Hit Dice, a base attack bonus of +2, and base saving throw bonuses of Fort 
+1, Ref +3, and Will +1.---

Racial Skills: A bugbear's humanoid levels give it skill points equal to 6 x $ $(2 
+ Int modifier). Its class skills are Climb, Hide, Listen, Move Silently, Search, 
and Spot.---

Racial Feats: A bugbear's humanoid levels give it two feats.--- 

\parindent=3pt
+3 natural armor bonus.--- 

+4 racial bonus on Move Silently checks.---

Automatic Languages: Common, Goblin. Bonus Languages: Draconic, Elven, Giant, Gnoll, 
Orc.---

Favored Class: Rogue.---

Level adjustment +1.

\vspace{12pt}
\parindent=0pt
{\LARGE{}BULETTE}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{234pt}|}
\hline
  & Huge Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 9d10+45 (94 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +2\tabularnewline
\hline
S\textbf{peed:} & 40 ft. (8 squares), burrow 10 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 22 (-2 size, +2 Dex, +12 natural), touch 10, flat-footed 
20\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +9/+25\tabularnewline
\hline
A\textbf{ttack:} & Bite +16 melee (2d8+8)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +16 melee (2d8+8) and 2 claws +10 melee (2d6+4)\tabularnewline
\hline
S\textbf{pace/Reach:} & 15 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Leap\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., lowlight vision, scent, tremorsense 
60 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +11, Ref +8, Will +6\tabularnewline
\hline
A\textbf{bilities:} & Str 27, Dex 15, Con 20, Int 2, Wis 13, Cha 6\tabularnewline
\hline
S\textbf{kills:} & Jump +18, Listen +9, Spot +3\tabularnewline
\hline
F\textbf{eats:} & Alertness, Iron Will, Track, Weapon Focus (bite)\tabularnewline
\hline
E\textbf{nvironment:} & Temperate hills\tabularnewline
\hline
O\textbf{rganization:} & Solitary or pair\tabularnewline
\hline
C\textbf{hallenge Rating:} & 7\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & 10-16 HD (Huge); 17-27 HD (Gargantuan)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

Also known as the landshark, the bulette is a terrifying predator that lives only 
to eat. 

COMBAT

A bulette attacks anything it regards as edible, choosing the easiest or closest 
prey first. The only creatures it refuses to eat are elves (and it dislikes the 
taste of dwarves). When burrowing underground, a landshark relies on its tremorsense 
ability to detect prey.

When it senses something edible (that is, senses movement), it breaks to the surface, 
crest first, and begins its attack.

\textbf{Leap (Ex):} A bulette can jump into the air during combat. This allows 
it to make four claw attacks instead of two, each with a +15 attack bonus, but 
it cannot bite.

\vspace{12pt}
\subsection*{{\LARGE{}CELESTIAL CREATURE}}

Celestial creatures dwell on the upper planes, the realms of good, although they 
resemble beings found on the Material Plane. They are more regal and more beautiful 
than their earthly counterparts.

Celestial creatures often come in metallic colors (usually silver, gold, or platinum). 
They can be mistaken for half-celestials, more powerful creatures that are created 
when a celestial mates with a non-celestial creature.

CREATING A CELESTIAL CREATURE

``Celestial'' is an inherited template that can be added to any corporeal animal, 
aberration, animal, dragon, fey, giant, humanoid, magical beast, monstrous humanoid, 
plant, or vermin of good or neutral alignment (referred to hereafter as the base 
creature).

A celestial creature uses all the base creature's statistics and abilities except 
as noted here. Do not recalculate the creature's Hit Dice, base attack bonus, saves, 
or skill points if its type changes.

\textbf{Size and Type:} Animals or vermin with this template become magical beasts, 
but otherwise the creature type is unchanged. Size is unchanged. Celestial creatures 
encountered on the Material Plane have the extraplanar subtype.

\textbf{Special Attacks:} A celestial creature retains all the special attacks 
of the base creature and also gains the following attack.

\textit{Smite Evil (Su): }Once per day a celestial creature can make a normal melee 
attack to deal extra damage equal to its HD (maximum of +20) against an evil foe.

\textbf{Special Qualities:} A celestial creature retains all the special qualities 
of the base creature and also gains the following qualities.---

Darkvision out to 60 feet.---

Damage reduction (see the table below).---

Resistance to acid, cold, and electricity (see the table below).---

Spell resistance equal to HD + 5 (maximum 25).

\vspace{12pt}
\begin{tabular}{|>{\raggedright}p{42pt}|>{\raggedright}p{152pt}|>{\raggedright}p{80pt}|}
\hline
H\textbf{it Dice} & R\textbf{esistance to Acid, Cold, Electricity} & D\textbf{amage 
Reduction}\tabularnewline
\hline
1-3 & 5--- & \tabularnewline
\hline
4-7 & 5 & 5/magic\tabularnewline
\hline
8-11 & 10 & 5/magic\tabularnewline
\hline
12 or more & 10 & 10/magic\tabularnewline
\hline
\end{tabular}

If the base creature already has one or more of these special qualities, use the 
better value.

If a celestial creature gains damage reduction, its natural weapons are treated 
as magic weapons for the purpose of overcoming damage reduction.

\textbf{Abilities:} Same as the base creature, but Intelligence is at least 3.

\textbf{Environment:} Any good-aligned plane.

\textbf{Challenge Rating:} HD 3 or less, as base creature; HD 4 to 7, as base creature 
+1; HD 8 or more, as base creature +2.

\textbf{Alignment:} Always good (any).

\textbf{Level Adjustment:} Same as the base creature +2.

\vspace{12pt}
{\LARGE{}CENTAUR}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{231pt}|}
\hline
  & Large Monstrous Humanoid\tabularnewline
\hline
H\textbf{it Dice:} & 4d8+8 (26 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +2\tabularnewline
\hline
S\textbf{peed:} & 50 ft. (10 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 14 (-1 size, +2 Dex, +3 natural), touch 11, flat-footed 
12\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +4/+12\tabularnewline
\hline
A\textbf{ttack:} & Longsword +7 melee (2d6+6/19-20) or composite longbow (+4 Str 
bonus) +5 ranged (2d6+4/x3)\tabularnewline
\hline
F\textbf{ull Attack:} & Longsword +7 melee (2d6+6/19-20) and 2 hooves +3 melee 
(1d6+2); or composite longbow (+4 Str bonus) +5 ranged (2d6+4/x3)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:}--- & \tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +3, Ref +6, Will +5\tabularnewline
\hline
A\textbf{bilities:} & Str 18, Dex 14, Con 15, Int 8, Wis 13, Cha 11\tabularnewline
\hline
S\textbf{kills:} & Listen +3, Move Silently +4, Spot +3, Survival +2 \tabularnewline
\hline
F\textbf{eats:} & Dodge, Weapon Focus (hoof )\tabularnewline
\hline
E\textbf{nvironment:} & Temperate forests\tabularnewline
\hline
O\textbf{rganization:} & Solitary, company (5-8), troop (8-18 plus 1 leader of 
2nd-5th level), or tribe (20-150 plus 30\% noncombatants plus 10 3rd-level sergeants, 
5 5th-level lieutenants, and 1 leader of 5th-9th level)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 3\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Usually neutral good\tabularnewline
\hline
A\textbf{dvancement:} & By character class\tabularnewline
\hline
L\textbf{evel Adjustment:} & +2\tabularnewline
\hline
\end{tabular}

A centaur is as big as a heavy horse, but much taller and slightly heavier. A centaur 
is about 7 feet tall and weighs about 2,100 pounds.

Centaurs speak Sylvan and Elven.

COMBAT

A centaur employing a lance deals double damage when it charges, just as a rider 
on a mount does.

CENTAURS AS CHARACTERS

A centaur druid is usually a tribe's designated leader and speaker. Centaur characters 
possess the following racial traits.--- 

\parindent=3pt
+8 Strength, +4 Dexterity, +4 Constitution, -2 Intelligence, +2 Wisdom.---

\parindent=0pt
Large size. -1 penalty to Armor Class, -1 penalty on attack rolls, -4 penalty on 
Hide checks, +4 bonus on grapple checks, lifting and carrying limits double those 
of Medium characters.---

Space/Reach: 10 feet/5 feet.---

A centaur's base land speed is 50 feet.---

Darkvision out to 60 feet.---

Racial Hit Dice: A centaur begins with four levels of monstrous humanoid, which 
provide 4d8 Hit Dice, a base attack bonus of +4, and base saving throw bonuses 
of Fort +1, Ref +4, and Will +4.---

Racial Skills: A centaur's monstrous humanoid levels give it skill points equal 
to 7 x$ $ (2 + Int modifier). Its class skills are Listen, Move Silently, Spot, 
and Survival.---

Racial Feats: A centaur's monstrous humanoid levels give it two feats.--- 

\parindent=3pt
+3 natural armor bonus.--- 

Automatic Languages: Sylvan, Elven. Bonus Languages: Common, Gnome, Halfling.---

Favored Class: Ranger.---

Level adjustment +2.

\vspace{12pt}
\parindent=0pt
{\LARGE{}CHAOS BEAST}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{202pt}|}
\hline
  & Medium Outsider (Chaotic, Extraplanar)\tabularnewline
\hline
H\textbf{it Dice:} & 8d8+8 (44 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +5\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares)\tabularnewline
\hline
A\textbf{rmor Class:} & 16 (+1 Dex, +5 natural), touch 11, flat-footed 15\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +8/+10\tabularnewline
\hline
A\textbf{ttack:} & Claw +10 melee (1d3+2 plus corporeal instability)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +10 melee (1d3+2 plus corporeal instability)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Corporeal instability\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., immunity to critical hits and 
transformation, spell resistance 15\tabularnewline
\hline
S\textbf{aves:} & Fort +7, Ref +7, Will +6\tabularnewline
\hline
A\textbf{bilities:} & Str 14, Dex 13, Con 13, Int 10, Wis 10, Cha 10\tabularnewline
\hline
S\textbf{kills:} & Climb +13, Escape Artist +12, Hide +12, Jump +9, Listen +11, 
Search +11, Spot +11, Survival +0 (+2 following tracks), Tumble +14, Use Rope +1 
(+3 with bindings)\tabularnewline
\hline
F\textbf{eats:} & Dodge, Improved Initiative, Mobility\tabularnewline
\hline
E\textbf{nvironment:} & Ever-Changing Chaos of Limbo\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 7\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always chaotic neutral\tabularnewline
\hline
A\textbf{dvancement:} & 9-12 HD (Medium); 13-24 HD (Large)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

The horrific creatures known as chaos beasts have mutable, ever-changing forms. 
A chaos beast's dimensions vary, but it always weighs about 200 pounds.

Chaos beasts do not speak.

COMBAT

For all its fearsome appearances, whether it has claws, fangs, pincers, tentacles, 
or spines, a chaos beast does little physical harm. Regardless of form, the creature 
seems unable to manage more than two attacks per round. Its continual transmutations 
prevent the coordination needed to do more.

A chaos beast's claw attacks, as well as any weapons it wields, are treated as 
chaotic-aligned for the purpose of overcoming damage reduction.

\textbf{Corporeal Instability (Su):} A blow from a chaos beast against a living 
creature can cause a terrible transformation. The creature must succeed on a DC 
15 Fortitude save or become a spongy, amorphous mass. Unless the victim manages 
to control the effect (see below), its shape melts, flows, writhes, and boils. 
The save DC is Constitution-based.

An affected creature is unable to hold or use any item. Clothing, armor, rings, 
and helmets become useless. Large items worn or carried---armor, backpacks, even 
shirts---hamper more than help, reducing the victim's Dexterity score by 4. Soft 
or misshapen feet and legs reduce speed to 10 feet or one-quarter normal, whichever 
is less. Searing pain courses along the nerves, so strong that the victim cannot 
act coherently. The victim cannot cast spells or use magic items, and it attacks 
blindly, unable to distinguish friend from foe (-4 penalty on attack rolls and 
a 50\% miss chance, regardless of the attack roll).

Each round the victim spends in an amorphous state causes 1 point of Wisdom drain 
from mental shock. If the victim's Wisdom score falls to 0, it becomes a chaos 
beast.

A victim can regain its own shape by taking a standard action to attempt a DC 15 
Charisma check (this check DC does not vary for a chaos beast with different Hit 
Dice or ability scores). A success reestablishes the creature's normal form for 
1 minute. On a failure, the victim can still repeat this check each round until 
successful.

Corporeal instability is not a disease or a curse and so is hard to remove. A \textit{shapechange 
}or \textit{stoneskin }spell does not cure an afflicted creature but fixes its 
form for the duration of the spell. A \textit{restoration, heal, }or \textit{greater 
restoration }spell removes the affliction (a separate \textit{restoration }is necessary 
to restore any drained points of Wisdom).

\textbf{Immunity to Transformation (Ex): }No mortal magic can permanently affect 
or fix a chaos beast's form. Effects such as polymorphing or petrification force 
the creature into a new shape, but at the start of its next turn it immediately 
returns to its mutable form as a free action.

\vspace{12pt}
{\LARGE{}CHIMERA}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{231pt}|}
\hline
  & Large Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 9d10+27 (76 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +1\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares), fly 50 ft. (poor)\tabularnewline
\hline
A\textbf{rmor Class:} & 19 (-1 size, +1 Dex, +9 natural), touch 10, flat-footed 
18\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +9/+17\tabularnewline
\hline
A\textbf{ttack:} & Bite +12 melee (2d6+4)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +12 melee (2d6+4) and bite +12 melee (1d8+4) and gore 
+12 melee (1d8+4) and 2 claws +10 melee (1d6+2)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Breath weapon\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., low-light vision, scent\tabularnewline
\hline
S\textbf{aves:} & Fort +9, Ref +7, Will +6\tabularnewline
\hline
A\textbf{bilities:} & Str 19, Dex 13, Con 17, Int 4, Wis 13, Cha 10\tabularnewline
\hline
S\textbf{kills:} & Hide +1*, Listen +9, Spot +9\tabularnewline
\hline
F\textbf{eats:} & Alertness, Hover, Iron Will, Multiattack\tabularnewline
\hline
E\textbf{nvironment:} & Temperate hills\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pride (3-5), or flight (6-13)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 7\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 10-13 HD (Large); 14-27 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +2 (cohort)\tabularnewline
\hline
\end{tabular}

A chimera is about 5 feet tall at the shoulder, nearly 10 feet long, and weighs 
about 4,000 pounds. A chimera's dragon head might be black, blue, green, red, or 
white.

Chimeras can speak Draconic but seldom bother to do so, except when toadying to 
more powerful creatures.

COMBAT

The chimera prefers to surprise prey. It often swoops down from the sky or lies 
concealed until it charges. The dragon head can loose a breath weapon instead of 
biting. Several chimeras attack in concert.

\textbf{Breath Weapon (Su):} A chimera's breath weapon depends on the color of 
its dragon head, as summarized on the table below. Regardless of its type, a chimera's 
breath weapon is usable once every 1d4 rounds, deals 3d8 points of damage, and 
allows a DC 17 Reflex save for half damage. The save DC is Constitution-based.

To determine a chimera's head color and breath weapon randomly, roll 1d10 and consult 
the table below.

\begin{tabular}{|>{\raggedright}p{19pt}|>{\raggedright}p{48pt}|>{\raggedright}p{102pt}|}
\hline
1\textbf{d10} & H\textbf{ead Color} & B\textbf{reath Weapon}\tabularnewline
\hline
1-2 & Black & 40-foot line of acid\tabularnewline
\hline
3-4 & Blue & 40-foot line of lightning\tabularnewline
\hline
5-6 & Green & 20-foot cone of gas (acid)\tabularnewline
\hline
7-8 & Red & 20-foot cone of fire\tabularnewline
\hline
9-10 & White & 20-foot cone of cold\tabularnewline
\hline
\end{tabular}

\textbf{Skills:} A chimera's three heads give it a +2 racial bonus on Spot and 
Listen checks.

*In areas of scrubland or brush, a chimera gains a +4 racial bonus on Hide checks.

\textbf{Carrying Capacity:} A light load for a chimera is up to 348 pounds; a medium 
load, 349-699 pounds, and a heavy load, 700-1,050 pounds.

\vspace{12pt}
{\LARGE{}CHOKER}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{226pt}|}
\hline
  & Small Aberration\tabularnewline
\hline
H\textbf{it Dice:} & 3d8+3 (16 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +6\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), climb 10 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 17 (+1 size, +2 Dex, +4 natural), touch 13, flat-footed 
15\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +2/+5\tabularnewline
\hline
A\textbf{ttack:} & Tentacle +6 melee (1d3+3)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 tentacles +6 melee (1d3+3)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./10 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Improved grab, constrict 1d3+3\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., quickness\tabularnewline
\hline
S\textbf{aves:} & Fort +2, Ref +5, Will +4\tabularnewline
\hline
A\textbf{bilities:} & Str 16, Dex 14, Con 13, Int 4, Wis 13, Cha 7\tabularnewline
\hline
S\textbf{kills:} & Climb +13, Hide +10, Move Silently +6\tabularnewline
\hline
F\textbf{eats:} & Improved Initiative\textsuperscript{B}, Lightning Reflexes, Stealthy\tabularnewline
\hline
E\textbf{nvironment:} & Underground\tabularnewline
\hline
O\textbf{rganization:} & Solitary\tabularnewline
\hline
C\textbf{hallenge Rating:} & 2\tabularnewline
\hline
T\textbf{reasure:} & 1/10 coins; 50\% goods; 50\% items\tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 4-6 HD (Small); 7-12 HD (Medium)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

These vicious little predators lurk underground, grabbing whatever prey happens 
by. Its hands and feet have spiny pads that help the choker grip almost any surface. 
The creature weighs about 35 pounds.

Chokers speak Undercommon.

COMBAT

A choker likes to perch near the ceiling, often at intersections, archways, wells, 
or staircases, and reach down to attack its prey. A choker attacks creatures of 
almost any size, but prefers lone prey of its size or larger. If one is very hungry, 
it may attack a group, but it waits to grab the last creature in line.

\textbf{Constrict (Ex): }A choker deals 1d3+3 points of damage with a successful 
grapple check against a Large or smaller creature. Because it seizes its victim 
by the neck, a creature in the choker's grasp cannot speak or cast spells with 
verbal components.

\textbf{Improved Grab (Ex): }To use this ability, a choker must hit a Large or 
smaller opponent with a tentacle attack. It can then attempt to start a grapple 
as a free action without provoking an attack of opportunity. If it wins the grapple 
check, it establishes a hold and can constrict. Chokers receive a +4 racial bonus 
on grapple checks, which is already included in the statistics block.

\textbf{Quickness (Su): }Although not particularly dexterous, a choker is supernaturally 
quick. It can take an extra standard action or move action during its turn each 
round.

\textbf{Skills: }A chokers has a +8 racial bonus on Climb checks and can always 
choose to take 10 on Climb checks, even if rushed or threatened.

\vspace{12pt}
{\LARGE{}CHUUL}

\begin{tabular}{|>{\raggedright}p{89pt}|>{\raggedright}p{236pt}|}
\hline
  & Large Aberration (Aquatic)\tabularnewline
\hline
H\textbf{it Dice:} & 11d8+44 (93 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +7\tabularnewline
\hline
S\textbf{peed:} & 30 ft. (6 squares), swim 20 ft.\tabularnewline
\hline
A\textbf{rmor Class:} & 22 (-1 size, +3 Dex, +10 natural), touch 12, flat-footed 
19\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +8/+17\tabularnewline
\hline
A\textbf{ttack:} & Claw +12 melee (2d6+5)\tabularnewline
\hline
F\textbf{ull Attack:} & 2 claws +12 melee (2d6+5)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Constrict 3d6+5, improved grab, paralytic tentacles\tabularnewline
\hline
S\textbf{pecial Qualities:} & Amphibious, darkvision 60 ft., immunity to poison\tabularnewline
\hline
S\textbf{aves:} & Fort +7, Ref +6, Will +9\tabularnewline
\hline
A\textbf{bilities:} & Str 20, Dex 16, Con 18, Int 10, Wis 14, Cha 5\tabularnewline
\hline
S\textbf{kills:} & Hide +13, Listen +11, Spot +11, Swim +13\tabularnewline
\hline
F\textbf{eats:} & Alertness, Blind-Fight,Combat Reflexes, Improved Initiative\tabularnewline
\hline
E\textbf{nvironment:} & Temperate marshes\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or pack (3-5)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 7\tabularnewline
\hline
T\textbf{reasure:} & 1/10th coins; 50\% goods; standard items\tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic evil\tabularnewline
\hline
A\textbf{dvancement:} & 12-16 HD (Large); 17-33 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A horrible mix of crustacean, insect, and serpent, the chuul is an abomination 
that lurks submerged or partially submerged, awaiting intelligent prey to devour. 
Although amphibious, chuuls are not good swimmers and actually prefer to be on 
land or in very shallow water when they attack. A chuul is about 8 feet long and 
weighs 650 pounds.

Chuuls speak Common (or Undercommon, for the underground variety).

COMBAT

A chuul prefers to wait by the shore, submerged in murky water, until it hears 
nearby prey (in or out of the water) that it can attack with surprise. A chuul 
grabs with its claws and constricts its foe, then passes the opponent to its paralytic 
tentacles. It tries to always have one claw free, so if it faces a large number 
of opponents, it drops a paralyzed or dead victim and continues attempting to grab, 
constrict, and paralyze the rest.

\textbf{Constrict (Ex): }On a successful grapple check, a chuul deals 3d6+5 points 
of damage.

\textbf{Improved Grab (Ex):} To use this ability, a chuul must hit with a claw 
attack. It can then attempt to start a grapple as a free action without provoking 
an attack of opportunity. If it wins the grapple check, it establishes a hold and 
can constrict or on its next turn transfer a grabbed opponent to its tentacles.

\textbf{Paralytic Tentacles (Ex):} A chuul can transfer grabbed victims from a 
claw to its ten- tacles as a move action. The tentacles grapple with the same strength 
as the claw but deal no damage. However, they exude a paralytic secretion. Anyone 
held in the tentacles must succeed on a DC 19 Fortitude save each round on the 
chuul's turn or be paralyzed for 6 rounds. The save DC is Constitution-based. While 
held in the tentacles, paralyzed or not, a victim automatically takes 1d8+2 points 
of damage each round from the creature's mandibles.

\textbf{Amphibious (Ex):} Although chuuls are aquatic, they can survive indefinitely 
on land.

\textbf{Skills:} A chuul has a +8 racial bonus on any Swim check to perform some 
special action or avoid a hazard. It can always choose to take 10 on a Swim check, 
even if distracted or endangered. It can use the run action while swimming, provided 
it swims in a straight line.

\vspace{12pt}
{\LARGE{}CLOAKER}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{226pt}|}
\hline
  & Large Aberration\tabularnewline
\hline
H\textbf{it Dice:} & 6d8+18 (45 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +7\tabularnewline
\hline
S\textbf{peed:} & 10 ft. (2 squares), fly 40 ft. (average)\tabularnewline
\hline
A\textbf{rmor Class:} & 19 (-1 size, +3 Dex, +7 natural), touch 12, flat-footed 
16\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +4/+13 \tabularnewline
\hline
A\textbf{ttack:} & Tail slap +8 melee (1d6+5)\tabularnewline
\hline
F\textbf{ull Attack:} & Tail slap +8 melee (1d6+5) and bite +3 melee (1d4+2)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./10 ft. (5 ft. with bite)\tabularnewline
\hline
S\textbf{pecial Attacks:} & Moan, engulf\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., shadow shift\tabularnewline
\hline
S\textbf{aves:} & Fort +5, Ref +5, Will +7\tabularnewline
\hline
A\textbf{bilities:} & Str 21, Dex 16, Con 17, Int 14, Wis 15, Cha 15\tabularnewline
\hline
S\textbf{kills:} & Hide +8, Listen +13, Move Silently +12, Spot +13\tabularnewline
\hline
F\textbf{eats:} & Alertness, Combat Reflexes, Improved Initiative\tabularnewline
\hline
E\textbf{nvironment:} & Underground\tabularnewline
\hline
O\textbf{rganization:} & Solitary, mob (3-6), or flock (7-12)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 5\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Usually chaotic neutral\tabularnewline
\hline
A\textbf{dvancement:} & 7-9 HD (Large); 10-18 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

When resting or lying in wait, these creatures are almost impossible to distinguish 
from common black cloaks (the cloaker's ivory claws look very much like bone clasps). 
Only when it unfurls does the horrific nature of the creature become apparent.

A cloaker has a wingspan of about 8 feet. It weighs about 100 pounds.

Cloakers speak Undercommon.

COMBAT

Cloakers usually lie still, watching and listening for prey. If facing a single 
opponent, a cloaker uses its engulf attack. Against multiple foes, it lashes with 
its tail in concert with its moan and shadow shift abilities to reduce the opposition's 
numbers, then engulfs a survivor. Multiple cloakers usually split up, leaving one 
or two behind to use special abilities while the rest make melee attacks.

\textbf{Moan (Ex): }A cloaker can emit a dangerous subsonic moan as a standard 
action. By changing the frequency, the cloaker can cause one of four effects. Cloakers 
are immune to these sonic, mindaffecting attacks. Unless otherwise specified, a 
creature that successfully saves against one of these effects cannot be affected 
by the same moan effect from the same cloaker for 24 hours. All save DCs for moan 
effects are Charisma-based.

\textit{Unnerve: }Anyone within a 60-foot spread automatically takes a -2 penalty 
on attack and damage rolls. Those forced to hear the moan for more than 6 consecutive 
rounds must succeed on a DC 15 Will save or enter a trance, unable to attack or 
defend themselves until the moaning stops.

\textit{Fear: }Anyone within a 30-foot spread must succeed on a DC 15 Will save 
or become panicked for 2 rounds.

\textit{Nausea: }Anyone in a 30-foot cone must succeed on a DC 15 Fortitude save 
or be overcome by nausea and weakness.

Affected characters fall prone and become nauseated for 1d4+1 rounds.

\textit{Stupor: }A single creature within 30 feet of the cloaker must succeed on 
a DC 15 Fortitude save or be affected as though by a \textit{hold monster }spell 
for 5 rounds. Even after a successful save, the creature must repeat the save if 
the cloaker uses this effect again.

\textbf{Engulf (Ex): }A cloaker can try to wrap a Medium or smaller creature in 
its body as a standard action. The cloaker attempts a grapple that does not provoke 
an attack of opportunity. If it wins the grapple check, it establishes a hold and 
bites the engulfed victim with a +4 bonus on its attack roll. It can still use 
its whiplike tail to strike at other targets.

Attacks that hit an engulfing cloaker deal half their damage to the monster and 
half to the trapped victim.

\textbf{Shadow Shift (Su): }A cloaker can manipulate shadows. This ability is effective 
only in shadowy areas and has three possible effects.

\textit{Obscure Vision: }The cloaker gains concealment (20\% miss chance) for 1d4 
rounds.

\textit{Dancing Images: }This effect duplicates a \textit{mirror image }spell (caster 
level 6th).

\textit{Silent Image: }This effect duplicates a \textit{silent image }spell (DC 
15, caster level 6th). The save DC is Charisma-based.

\vspace{12pt}
{\LARGE{}COCKATRICE}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{180pt}|}
\hline
  & Small Magical Beast\tabularnewline
\hline
H\textbf{it Dice:} & 5d10 (27 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +3\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), fly 60 ft. (poor)\tabularnewline
\hline
A\textbf{rmor Class:} & 14 (+1 size, +3 Dex), touch 14, flat-footed 11\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +5/-1\tabularnewline
\hline
A\textbf{ttack:} & Bite +9 melee (1d4-2 plus petrification)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +9 melee (1d4-2 plus petrification)\tabularnewline
\hline
S\textbf{pace/Reach:} & 5 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Petrification\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., low-light vision\tabularnewline
\hline
S\textbf{aves:} & Fort +4, Ref +7, Will +2\tabularnewline
\hline
A\textbf{bilities:} & Str 6, Dex 17, Con 11, Int 2, Wis 13, Cha 9\tabularnewline
\hline
S\textbf{kills:} & Listen +7, Spot +7\tabularnewline
\hline
F\textbf{eats:} & Alertness, Dodge, Weapon Finesse\textsuperscript{B}\tabularnewline
\hline
E\textbf{nvironment:} & Temperate plains\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, flight (3-5), or flock (6-13)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 3\tabularnewline
\hline
T\textbf{reasure:} & None\tabularnewline
\hline
A\textbf{lignment:} & Always neutral\tabularnewline
\hline
A\textbf{dvancement:} & 6-8 HD (Small); 9-15 HD (Medium)\tabularnewline
\hline
L\textbf{evel Adjustment:}--- & \tabularnewline
\hline
\end{tabular}

A male cockatrice has wattles and a comb, just like a rooster. Females, much rarer 
than males, differ only in that they have no wattles or comb. A cockatrice weighs 
about 25 pounds.

COMBAT

A cockatrice fiercely attacks anything that it deems a threat to itself or its 
lair. Flocks of cockatrices do their utmost to overwhelm and confuse their foes, 
and sometimes fly directly into their opponents' faces.

\textbf{Petrification (Su): }Creatures hit by a cockatrice's bite attack must succeed 
on a DC 12 Fortitude save or instantly turn to stone. The save DC is Constitution-based. 
 Cockatrices have immunity to the petrification ability of other cockatrices, but 
other petrification attacks affect them normally.

\vspace{12pt}
{\LARGE{}COUATL}

\begin{tabular}{|>{\raggedright}p{91pt}|>{\raggedright}p{231pt}|}
\hline
  & Large Outsider (Native)\tabularnewline
\hline
H\textbf{it Dice:} & 9d8+18 (58 hp)\tabularnewline
\hline
I\textbf{nitiative:} & +7\tabularnewline
\hline
S\textbf{peed:} & 20 ft. (4 squares), fly 60 ft. (good)\tabularnewline
\hline
A\textbf{rmor Class:} & 21 (-1 size, +3 Dex, +9 natural), touch 12, flat-footed 
18\tabularnewline
\hline
B\textbf{ase Attack/Grapple:} & +9/+17\tabularnewline
\hline
A\textbf{ttack:} & Bite +12 melee (1d3+6 plus poison)\tabularnewline
\hline
F\textbf{ull Attack:} & Bite +12 melee (1d3+6 plus poison)\tabularnewline
\hline
S\textbf{pace/Reach:} & 10 ft./5 ft.\tabularnewline
\hline
S\textbf{pecial Attacks:} & Constrict 2d8+6, improved grab, poison, psionics, spells\tabularnewline
\hline
S\textbf{pecial Qualities:} & Darkvision 60 ft., ethereal jaunt, telepathy 90 ft.\tabularnewline
\hline
S\textbf{aves:} & Fort +8, Ref +9, Will +10\tabularnewline
\hline
A\textbf{bilities:} & Str 18, Dex 16, Con 14, Int 17, Wis 19, Cha 17\tabularnewline
\hline
S\textbf{kills:} & Concentration +14, Diplomacy +17, Jump +0, Knowledge (any two) 
+15, Listen +16, Search +15, Sense Motive +16, Spellcraft +15 (+17 scrolls), Spot 
+16, Survival +4 (+6 following tracks), Tumble +15, Use Magic Device +15 (+17 scrolls)\tabularnewline
\hline
F\textbf{eats:} & Dodge, Empower Spell, Eschew Materials\textsuperscript{B}, Hover, 
Improved Initiative\tabularnewline
\hline
E\textbf{nvironment:} & Warm forests\tabularnewline
\hline
O\textbf{rganization:} & Solitary, pair, or flight (3-6)\tabularnewline
\hline
C\textbf{hallenge Rating:} & 10\tabularnewline
\hline
T\textbf{reasure:} & Standard\tabularnewline
\hline
A\textbf{lignment:} & Always lawful good\tabularnewline
\hline
A\textbf{dvancement:} & 10-13 HD (Large); 14-27 HD (Huge)\tabularnewline
\hline
L\textbf{evel Adjustment:} & +7\tabularnewline
\hline
\end{tabular}

A couatl is about 12 feet long, with a wingspan of about 15 feet. It weighs about 
1,800 pounds.

Couatls speak Celestial, Common, and Draconic, and also have the power of telepathy 
(see below).

COMBAT

A couatl uses its \textit{detect thoughts }ability on any creature that arouses 
its suspicions. Since it is highly intelligent, a couatl usually casts spells from 
a distance before closing. If more than one couatl is involved, they discuss their 
strategy before a battle.

\textbf{Constrict (Ex):} A couatl deals 2d8+6 points of damage with a successful 
grapple check.

\textbf{Improved Grab (Ex): }To use this ability, a couatl must hit a creature 
of up to two size categories larger with its bite attack. It can then attempt to 
start a grapple as a free action without provoking an attack of opportunity. If 
it wins the grapple check, it establishes a hold and can constrict.

\textbf{Poison (Ex)}: Injury, Fortitude DC 16, initial damage 2d4 Str, secondary 
damage 4d4 Str.

The save DC is Constitution-based.

\textbf{Psionics (Sp):} At will---\textit{detect chaos, detect evil, detect good, 
detect law, detect thoughts }(DC 15), \textit{invisibility, plane shift }(DC 20), 
\textit{polymorph }(self only). Effective caster level 9th. The save DCs are Charisma-based.

\textbf{Spells:} A couatl casts spells as a 9th-level sorcerer.

It can choose its spells known from the sorcerer list, the cleric list, and from 
the lists for the Air, Good, and Law domains. The cleric spells and domain spells 
are considered arcane spells for a couatl, meaning that the creature does not need 
a divine focus to cast them.

\textit{Typical Spells Known }(6/7/7/7/4; save DC 13 + spell level): 0---\textit{cure 
minor wounds, daze, disrupt undead, light, obscuring mist, ray of frost, read magic, 
resistance; }1st---\textit{endure elements, mage armor, protection from chaos, 
true strike, wind wall; }2nd---\textit{cure moderate wounds, eagle's splendor, 
scorching ray, silence; }3rd---\textit{gaseous form, magic circle against evil, 
summon monster III; }4th--- \textit{charm monster, freedom of movement.}

\textbf{Ethereal Jaunt (Su):} This ability works like the \textit{ethereal jaunt 
}spell (caster level 16th).

\textbf{Telepathy (Su):} A couatl can communicate telepathically with any creature 
within 90 feet that has an Intelligence score. The creature can respond to the 
couatl if it wishes---no common language is needed.

\newpage

\end{document}
